### 2017-09-21 Manuscript.Rmd
First version of the report

---
title: "Efectividad de reservas marinas a nivel nacional"
subtitle: "Casos de estudio del Pacífico, Golfo de California, y Sistema Arrecifal Mesoamericano"
header-includes:
- \usepackage{setspace}
- \doublespacing
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    toc: yes
fontsize: 12pt
bibliography:
- exported-references.bib
- packages.bib
csl: huct.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
options(knitr.table.format = "latex")

windowsFonts(Times=windowsFont("TT Times New Roman"))

```

Juan Carlos Villaseñor-Derbez^1\*^, Stuart Fulton^2^, Jorge Torre^2^

^1^ Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, California, United States

^2^ Comunidad y Biodiversidad A.C., Calle Isla del Peruano, Guaymas, Sonora, México

# Resumen ejecutivo

# Introducción

La sobrepesca y prácticas pesqueras no sostenibles son unas de las mayores amenazas para la conservación de los ecosistemas marinos del mundo [@halpern_2008-dK; @halpern_2017-Zi]. Esto es particularmente cierto para el caso de las pesquerías artesanales. La implementación de reservas marinas ( i.e. áreas donde la captura de una o más especies está prohibida) es una medida de manejo frecuentemente propuesta para recuperar stocks pesqueros e impulsar la productividad pesquera en aguas cercanas [@afflerbach_2014-HP; @krueck_2017-J1; @sala_2017-69]. Recientes trabajos han demostrado que también pueden mitigar y proveer amortiguamiento ante el cambio climático [@roberts_2017-J9], solve bycatch problems [@hastings_2017-sm] y en general, incrementar la biomasa, riqueza y densidades de organismos dentro de sus fronteras [@lester_2009-Ks; @giakoumi_2017-V2; @sala_2017-69].

En México, las reservas marinas han sido comúnmente establecidas como zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por la Comisión Nacional de Áreas Naturales Protegidas (CONANP). Al día de hoy, **41** RBs protegen una porción del ambiente marino en México. Sin embargo, solamente **27** de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas. Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, éste proceso aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores. La escasez de recursos monetarios y humanos de la comisión limitan también el monitoreo y vigilancia de las reservas, limitando el desempeño de la reserva.

Buscando promover una alternativa con procesos ascendentes para implementar reservas marinas, las OSCs comenzaron a trabajar con comunidades pesqueras para establecer reservas comunitarias [@uribe_2010-u2] . Estas son comúnmente establecidas dentro de zonas de concesión, una forma de derechos de uso territoriales para pesquerías (TURF, en inglés). Al permitir a los pescadores diseñar sus propias reservas, una mayor proporción de la comunidad está de acuerdo con los perímetros establecidos, y por lo tanto los respetan [@gelcich_2015-Gw; @espinosaromero_2014-PY; @beger_2004-Y8] . Adicionalmente, los pescadores pueden implementar sus reservas por un periodo acordado (usualmente cinco años), después del cual la reserva puede ser abierta a la pesca. Esto provee a los pescadores con un sentido de confianza de que, en caso de ser necesario, aún tienen acceso a pescar esa zona[^1]. Las reservas son directamente vigiladas y monitoreadas por la comunidad, quienes comúnmente utilizan pequeñas embarcaciones (*e.g.* pangas) para patrullar la zona, o realizan avistamientos desde la costa en búsqueda de pescadores ilegales Aún así, las reservas comunitarias carecen de reconocimiento legal; por lo tanto, no hay forma de penalizar a los infractores.

[^1]: Hasta ahora, solamente una comunidad ha decidido abrir sur reservas a la pesca.

Sin embargo, en el 2014 una nueva norma [@nom049sagpesc_2014-V6] permite a los pescadores solicitar el establecimiento de reservas marinas bajo el nombre de “Zonas de refugio Pesquero” (ZRP). El manejo de las ZRP combina procesos ascendentes y descendentes al reconocer legalmente las reservas propuestas por las comunidades. Posterior a la revisión por parte de la Comisión Nacional de Acuacultura y Pesca (CONAPESCA) y la opinión técnica del Instituto Nacional de Acuacultura y Pesca (INAPESCA) las ZRP son establecidas por el periodo solicitado por los pescadores [^2] . El monitoreo y la vigilancia de las ZRP es típicamente llevado a cabo por la comunidad , con ayuda de OSCs locales. Hasta este cambio regulatorio, las reservas comunitarias no contaban con el soporte legal, y eran solamente reconocidas por la comunidad. Al día de hoy, existen **35** ZRP establecidas en el Pacífico, Golfo de California y Caribe Mexicano.

[^2]: Existen excepciones a esto, como la “Zona de Refugio Pesquero Golfo de Ulloa” y la “Zona de Refugio Pesquero Akumal”, creadas por CONAPESCA para cerrar la pesca y prevenir la captura incidental de tortugas marinas.

Aunque existen tres aproximaciones generales para implementar reservas marinas en México (*i.e.* Zonas núcleo dentro de AMP, reservas comunitarias y Zonas de Refugio Pesuqero), aún no comprendemos a fondo las características sociales que permiten su efectividad. La ciencia de reservas marinas se ha enfocado ampliamente en los efectos biológicos que estas tienen [@lester_2009-Ks; @giakoumi_2017-V2; @sala_2017-69; @afflerbach_2014-HP; @krueck_2017-J1]. Aunque el aspecto ecológico de las reservas es importante para su éxito, su efectividad también depende del estado socioeconómico y los sistemas de gobernanza de las comunidades pesqueras. La literatura indica que diferentes características influyen en el éxito de una reserva. En Palau, por ejemplo, la edad (*i.e.* tiempo de implementación), tamaño y hábitat contenido son características claves que determinan la efectividad [@friedlander_2017-oI]. Por otro lado, en el Mar Meditarráneo, @difranco_2016-Xw identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks, el ingreso a los pescadores, y se presenta una mayor aceptación social de las prácticas de manejo. En una aproximación global, @edgar_2014-UO encuentran que la procuración, edad, tamaño y aislamiento son determinantes de la efectividad de las reservas. Por lo tanto, observamos que las características que habilitan el éxito varían a través de regiones, y poco esfuerzo se ha hecho por comprender estas interacciones en México.

# Objetivos

El objetivo de este trabajo este trabajo es realizar una evaluación de la efectividad de reservas marinas en México, presentando resultados de cinco comunidades costeras como caso de estudio. Las comunidades utilizadas en este reporte se distribuyen a lo largo de la costa Pacífica de Baja Claifornia, el Golfo de California, y el Sistema Arrecifal Mesoamericano. Con el fin de obtener una visión holística del sistema, la evaluación se realizará tomando en cuenta indicadores biológicos, socioeconómicos y de gobernanza (**Referencia aqui**).

La evaluación de éstos cinco casos de estudios nos permitirá identificar patrones en los que características socioeconómicas y de gobernanza que se encuentran ligadas a la efectividad (biológica) de las reservas marinas evaluadas. Los patrones identificados podrán utilizarse para informar la toma de desiciones para la implementación de la red de reservas marinas en la Región de las Grandes Islas del Golfo de California.

# Materiales y Métodos

## Zonas de estudio

## Caraterísticas de las comunidades y reservas seleccionadas

## Análisis de datos

# Resultados

# Conclusiones

# Referencias

### 2017-09-21 Manuscript.tex
First version of the report

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Efectividad de reservas marinas a nivel nacional},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Efectividad de reservas marinas a nivel nacional}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Casos de estudio del Pacífico, Golfo de California, y Sistema Arrecifal
Mesoamericano}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{setspace}
\doublespacing

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
Juan Carlos Villaseñor-Derbez\textsuperscript{1*}, Stuart
Fulton\textsuperscript{2}, Jorge Torre\textsuperscript{2}

\textsuperscript{1} Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara, California,
United States

\textsuperscript{2} Comunidad y Biodiversidad A.C., Calle Isla del
Peruano, Guaymas, Sonora, México

\section{Resumen ejecutivo}\label{resumen-ejecutivo}

\section{Introducción}\label{introduccion}

La sobrepesca y prácticas pesqueras no sostenibles son unas de las
mayores amenazas para la conservación de los ecosistemas marinos del
mundo (Halpern et al., 2008, 2017). Esto es particularmente cierto para
el caso de las pesquerías artesanales. La implementación de reservas
marinas ( i.e.~áreas donde la captura de una o más especies está
prohibida) es una medida de manejo frecuentemente propuesta para
recuperar stocks pesqueros e impulsar la productividad pesquera en aguas
cercanas (Afflerbach et al., 2014; Krueck et al., 2017; Sala \&
Giakoumi, 2017). Recientes trabajos han demostrado que también pueden
mitigar y proveer amortiguamiento ante el cambio climático (Roberts et
al., 2017), solve bycatch problems (Hastings, Gaines \& Costello, 2017)
y en general, incrementar la biomasa, riqueza y densidades de organismos
dentro de sus fronteras (Lester et al., 2009; Giakoumi et al., 2017;
Sala \& Giakoumi, 2017).

En México, las reservas marinas han sido comúnmente establecidas como
zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por
la Comisión Nacional de Áreas Naturales Protegidas (CONANP). Al día de
hoy, \textbf{41} RBs protegen una porción del ambiente marino en México.
Sin embargo, solamente \textbf{27} de estas incluyen (pequeñas) zonas
núcleo donde las actividades pesqueras están prohibidas. Aunque la
CONANP ha hecho esfuerzos importantes por involucrar a los actores
durante la implementación de las reservas, éste proceso aún se
caracteriza por un proceso descendente, el cual conlleva a la falta de
cumplimiento por parte de los actores. La escasez de recursos monetarios
y humanos de la comisión limitan también el monitoreo y vigilancia de
las reservas, limitando el desempeño de la reserva.

Buscando promover una alternativa con procesos ascendentes para
implementar reservas marinas, las OSCs comenzaron a trabajar con
comunidades pesqueras para establecer reservas comunitarias (Uribe et
al., 2010) . Estas son comúnmente establecidas dentro de zonas de
concesión, una forma de derechos de uso territoriales para pesquerías
(TURF, en inglés). Al permitir a los pescadores diseñar sus propias
reservas, una mayor proporción de la comunidad está de acuerdo con los
perímetros establecidos, y por lo tanto los respetan (Beger et al.,
2004; Espinosa-Romero et al., 2014; Gelcich \& Donlan, 2015) .
Adicionalmente, los pescadores pueden implementar sus reservas por un
periodo acordado (usualmente cinco años), después del cual la reserva
puede ser abierta a la pesca. Esto provee a los pescadores con un
sentido de confianza de que, en caso de ser necesario, aún tienen acceso
a pescar esa zona\footnote{Hasta ahora, solamente una comunidad ha
  decidido abrir sur reservas a la pesca.}. Las reservas son
directamente vigiladas y monitoreadas por la comunidad, quienes
comúnmente utilizan pequeñas embarcaciones (\emph{e.g.} pangas) para
patrullar la zona, o realizan avistamientos desde la costa en búsqueda
de pescadores ilegales Aún así, las reservas comunitarias carecen de
reconocimiento legal; por lo tanto, no hay forma de penalizar a los
infractores.

Sin embargo, en el 2014 una nueva norma (NOM-049-SAG/PESC, 2014) permite
a los pescadores solicitar el establecimiento de reservas marinas bajo
el nombre de ``Zonas de refugio Pesquero'' (ZRP). El manejo de las ZRP
combina procesos ascendentes y descendentes al reconocer legalmente las
reservas propuestas por las comunidades. Posterior a la revisión por
parte de la Comisión Nacional de Acuacultura y Pesca (CONAPESCA) y la
opinión técnica del Instituto Nacional de Acuacultura y Pesca (INAPESCA)
las ZRP son establecidas por el periodo solicitado por los
pescadores\footnote{Existen excepciones a esto, como la ``Zona de
  Refugio Pesquero Golfo de Ulloa'' y la ``Zona de Refugio Pesquero
  Akumal'', creadas por CONAPESCA para cerrar la pesca y prevenir la
  captura incidental de tortugas marinas.} . El monitoreo y la
vigilancia de las ZRP es típicamente llevado a cabo por la comunidad ,
con ayuda de OSCs locales. Hasta este cambio regulatorio, las reservas
comunitarias no contaban con el soporte legal, y eran solamente
reconocidas por la comunidad. Al día de hoy, existen \textbf{35} ZRP
establecidas en el Pacífico, Golfo de California y Caribe Mexicano.

Aunque existen tres aproximaciones generales para implementar reservas
marinas en México (\emph{i.e.} Zonas núcleo dentro de AMP, reservas
comunitarias y Zonas de Refugio Pesuqero), aún no comprendemos a fondo
las características sociales que permiten su efectividad. La ciencia de
reservas marinas se ha enfocado ampliamente en los efectos biológicos
que estas tienen (Lester et al., 2009; Afflerbach et al., 2014; Giakoumi
et al., 2017; Krueck et al., 2017; Sala \& Giakoumi, 2017). Aunque el
aspecto ecológico de las reservas es importante para su éxito, su
efectividad también depende del estado socioeconómico y los sistemas de
gobernanza de las comunidades pesqueras. La literatura indica que
diferentes características influyen en el éxito de una reserva. En
Palau, por ejemplo, la edad (\emph{i.e.} tiempo de implementación),
tamaño y hábitat contenido son características claves que determinan la
efectividad (Friedlander et al., 2017). Por otro lado, en el Mar
Meditarráneo, Di Franco et al. (2016) identifican que la procuración y
vigilancia, presencia de un plan de manejo, participación de pescadores
en el manejo, representación de pescadores en la toma de decisiones y
promoción de la pesca sustentable son los cinco factores que incrementan
la salud de los stocks, el ingreso a los pescadores, y se presenta una
mayor aceptación social de las prácticas de manejo. En una aproximación
global, Edgar et al. (2014) encuentran que la procuración, edad, tamaño
y aislamiento son determinantes de la efectividad de las reservas. Por
lo tanto, observamos que las características que habilitan el éxito
varían a través de regiones, y poco esfuerzo se ha hecho por comprender
estas interacciones en México.

\section{Objetivos}\label{objetivos}

El objetivo de este trabajo este trabajo es realizar una evaluación de
la efectividad de reservas marinas en México, presentando resultados de
cinco comunidades costeras como caso de estudio. Las comunidades
utilizadas en este reporte se distribuyen a lo largo de la costa
Pacífica de Baja Claifornia, el Golfo de California, y el Sistema
Arrecifal Mesoamericano. Con el fin de obtener una visión holística del
sistema, la evaluación se realizará tomando en cuenta indicadores
biológicos, socioeconómicos y de gobernanza (\textbf{Referencia aqui}).

La evaluación de éstos cinco casos de estudios nos permitirá identificar
patrones en los que características socioeconómicas y de gobernanza que
se encuentran ligadas a la efectividad (biológica) de las reservas
marinas evaluadas. Los patrones identificados podrán utilizarse para
informar la toma de desiciones para la implementación de la red de
reservas marinas en la Región de las Grandes Islas del Golfo de
California.

\section{Materiales y Métodos}\label{materiales-y-metodos}

\subsection{Zonas de estudio}\label{zonas-de-estudio}

\subsection{Caraterísticas de las comunidades y reservas
seleccionadas}\label{carateristicas-de-las-comunidades-y-reservas-seleccionadas}

\subsection{Análisis de datos}\label{analisis-de-datos}

\section{Resultados}\label{resultados}

\section{Conclusiones}\label{conclusiones}

\section*{Referencias}\label{referencias}
\addcontentsline{toc}{section}{Referencias}

\hypertarget{refs}{}
\hypertarget{ref-afflerbach_2014-HP}{}
Afflerbach, J.C., Lester, S.E., Dougherty, D.T. \& Poon, S.E. 2014. A
global survey of -reserves, territorial use rights for fisheries coupled
with marine reserves. \emph{Global Ecology and Conservation}. 2:97--106.
DOI:
\href{https://doi.org/10.1016/j.gecco.2014.08.001}{10.1016/j.gecco.2014.08.001}.

\hypertarget{ref-beger_2004-Y8}{}
Beger, M., Harborne, A.R., Dacles, T.P., Solandt, J.-L. \& Ledesma, G.L.
2004. A framework of lessons learned from community-based marine
reserves and its effectiveness in guiding a new coastal management
initiative in the philippines. \emph{Environ Manage}. 34(6):786--801.
DOI:
\href{https://doi.org/10.1007/s00267-004-0149-z}{10.1007/s00267-004-0149-z}.

\hypertarget{ref-difranco_2016-Xw}{}
Di Franco, A., Thiriet, P., Di Carlo, G., Dimitriadis, C., Francour, P.,
Gutiérrez, N.L., Jeudy de Grissac, A., Koutsoubas, D., et al. 2016. Five
key attributes can increase marine protected areas performance for
small-scale fisheries management. \emph{Sci Rep}. 6(1):38135. DOI:
\href{https://doi.org/10.1038/srep38135}{10.1038/srep38135}.

\hypertarget{ref-edgar_2014-UO}{}
Edgar, G.J., Stuart-Smith, R.D., Willis, T.J., Kininmonth, S., Baker,
S.C., Banks, S., Barrett, N.S., Becerro, M.A., et al. 2014. Global
conservation outcomes depend on marine protected areas with five key
features. \emph{Nature}. 506(7487):216--220. DOI:
\href{https://doi.org/10.1038/nature13022}{10.1038/nature13022}.

\hypertarget{ref-espinosaromero_2014-PY}{}
Espinosa-Romero, M.J., Rodriguez, L.F., Weaver, A.H., Villanueva-Aznar,
C. \& Torre, J. 2014. The changing role of ngos in mexican small-scale
fisheries: From environmental conservation to multi-scale governance.
\emph{Marine Policy}. 50:290--299. DOI:
\href{https://doi.org/10.1016/j.marpol.2014.07.005}{10.1016/j.marpol.2014.07.005}.

\hypertarget{ref-friedlander_2017-oI}{}
Friedlander, A.M., Golbuu, Y., Ballesteros, E., Caselle, J.E., Gouezo,
M., Olsudong, D. \& Sala, E. 2017. Size, age, and habitat determine
effectiveness of palau's marine protected areas. \emph{PLoS ONE}.
12(3):e0174787. DOI:
\href{https://doi.org/10.1371/journal.pone.0174787}{10.1371/journal.pone.0174787}.

\hypertarget{ref-gelcich_2015-Gw}{}
Gelcich, S. \& Donlan, C.J. 2015. Incentivizing biodiversity
conservation in artisanal fishing communities through territorial user
rights and business model innovation. \emph{Conserv Biol}.
29(4):1076--1085. DOI:
\href{https://doi.org/10.1111/cobi.12477}{10.1111/cobi.12477}.

\hypertarget{ref-giakoumi_2017-V2}{}
Giakoumi, S., Scianna, C., Plass-Johnson, J., Micheli, F.,
Grorud-Colvert, K., Thiriet, P., Claudet, J., Di Carlo, G., et al. 2017.
Ecological effects of full and partial protection in the crowded
mediterranean sea: A regional meta-analysis. \emph{Sci Rep}. 7(1):8940.
DOI:
\href{https://doi.org/10.1038/s41598-017-08850-w}{10.1038/s41598-017-08850-w}.

\hypertarget{ref-halpern_2017-Zi}{}
Halpern, B.S., Frazier, M., Afflerbach, J., O'Hara, C., Katona, S.,
Stewart Lowndes, J.S., Jiang, N., Pacheco, E., et al. 2017. Drivers and
implications of change in global ocean health over the past five years.
\emph{PLoS ONE}. 12(7):e0178267. DOI:
\href{https://doi.org/10.1371/journal.pone.0178267}{10.1371/journal.pone.0178267}.

\hypertarget{ref-halpern_2008-dK}{}
Halpern, B.S., Walbridge, S., Selkoe, K.A., Kappel, C.V., Micheli, F.,
D'Agrosa, C., Bruno, J.F., Casey, K.S., et al. 2008. A global map of
human impact on marine ecosystems. \emph{Science}. 319(5865):948--952.
DOI:
\href{https://doi.org/10.1126/science.1149345}{10.1126/science.1149345}.

\hypertarget{ref-hastings_2017-sm}{}
Hastings, A., Gaines, S.D. \& Costello, C. 2017. Marine reserves solve
an important bycatch problem in fisheries. \emph{Proc Natl Acad Sci U S
A}. (August, 9). DOI:
\href{https://doi.org/10.1073/pnas.1705169114}{10.1073/pnas.1705169114}.

\hypertarget{ref-krueck_2017-J1}{}
Krueck, N.C., Ahmadia, G.N., Possingham, H.P., Riginos, C., Treml, E.A.
\& Mumby, P.J. 2017. Marine reserve targets to sustain and rebuild
unregulated fisheries. \emph{PLoS Biol}. 15(1):e2000537. DOI:
\href{https://doi.org/10.1371/journal.pbio.2000537}{10.1371/journal.pbio.2000537}.

\hypertarget{ref-lester_2009-Ks}{}
Lester, S., Halpern, B., Grorud-Colvert, K., Lubchenco, J., Ruttenberg,
B., Gaines, S., Airamé, S. \& Warner, R. 2009. Biological effects within
no-take marine reserves: A global synthesis. \emph{Mar. Ecol. Prog.
Ser.} 384:33--46. DOI:
\href{https://doi.org/10.3354/meps08029}{10.3354/meps08029}.

\hypertarget{ref-nom049sagpesc_2014-V6}{}
NOM-049-SAG/PESC. 2014. NORMA oficial mexicana nom-049-sag/pesc-2014,
que determina el procedimiento para establecer zonas de refugio para los
recursos pesqueros en aguas de jurisdicción federal de los estados
unidos mexicanos. \emph{DOF}.

\hypertarget{ref-roberts_2017-J9}{}
Roberts, C.M., O'Leary, B.C., McCauley, D.J., Cury, P.M., Duarte, C.M.,
Lubchenco, J., Pauly, D., Sáenz-Arroyo, A., et al. 2017. Marine reserves
can mitigate and promote adaptation to climate change. \emph{Proc Natl
Acad Sci U S A}. 114(24):6167--6175. DOI:
\href{https://doi.org/10.1073/pnas.1701262114}{10.1073/pnas.1701262114}.

\hypertarget{ref-sala_2017-69}{}
Sala, E. \& Giakoumi, S. 2017. No-take marine reserves are the most
effective protected areas in the ocean. \emph{ICES Journal of Marine
Science}. DOI:
\href{https://doi.org/10.1093/icesjms/fsx059}{10.1093/icesjms/fsx059}.

\hypertarget{ref-uribe_2010-u2}{}
Uribe, P., Moguel, S., Torre, J., Bourillon, L. \& Saenz, A. 2010.
\emph{Implementación de reservas marinas en méxico}. 1st ed. (nos.).
Mexico.

\end{document}

### 2017-09-25 analisis.R
Agregue funciones como wrappers

@@ -1,0 +1,49 @@
+analisis <- function(peces, invert, comunidad, ano.imp){
+  
+  # "Índice de diversidad de shannon", "Riqueza", "Organismos > LT50", "Densidad", "Densidad de especies objetivo", "Nivel trófico", "Biomasa", "Biomasa de especies objetivo"
+  
+  results <- tibble(Indicador = c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI"),
+                    Modelo = list(NA))
+  
+  results$Modelo[[1]] <- shannon(data = peces, location = comunidad) %>% 
+    mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
+           Ano = as.factor(Ano)) %>% 
+    modelo()
+  
+  results$Modelo[[2]] <- richness(data = peces, location = comunidad) %>% 
+    mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
+           Ano = as.factor(Ano)) %>% 
+    modelo()
+  
+  results$Modelo[[3]] <- density(data = peces, location = comunidad) %>% 
+    mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
+           Ano = as.factor(Ano)) %>% 
+    modelo()
+  
+  results$Modelo[[4]] <- trophic(data = peces, location = comunidad) %>% 
+    mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
+           Ano = as.factor(Ano)) %>% 
+    modelo()
+  
+  results$Modelo[[5]] <- fish_biomass(data = peces, location = comunidad) %>% 
+    mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
+           Ano = as.factor(Ano)) %>% 
+    modelo()
+  
+  results$Modelo[[6]] <- shannon(data = peces, location = comunidad) %>% 
+    mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
+           Ano = as.factor(Ano)) %>% 
+    modelo()
+  
+  results$Modelo[[7]] <- richness(data = peces, location = comunidad) %>% 
+    mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
+           Ano = as.factor(Ano)) %>% 
+    modelo()
+  
+  results$Modelo[[8]] <- density(data = peces, location = comunidad) %>% 
[…]

### 2017-09-25 extract_effect.R

@@ -1,0 +1,7 @@
+extract_effect <- function(x){
+  
+  x$Modelo %>%
+    map_df(robust_se) %>% 
+    filter(term == "ZonaReserva:Post") %>% 
+    mutate(Indicador = c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI"))
+}

### 2017-09-25 modelo.R

@@ -1,0 +1,3 @@
+modelo <- function(data){
+  lm(Indicador ~ Ano + Zona + Post*Zona, data)
+}

### 2017-09-25 my_ggplot.R

@@ -1,0 +1,16 @@
+my_ggplot <- function(x, y.label){
+  
+  x$model %>% 
+    mutate(Ano = as.numeric(as.character(Ano))) %>% 
+    ggplot(aes(x = Ano, y = Indicador, color = Zona, fill = Zona)) +
+    geom_jitter(height = 0, width = 0.1) +
+    stat_summary(geom = "ribbon", fun.data = "mean_se", alpha = 0.5)+
+    stat_summary(geom = "point", fun.data = "mean_se", color = "black")+
+    stat_summary(geom = "line", fun.data = "mean_se", color = "black") +
+    theme_bw() +
+    theme(legend.position = "None") +
+    scale_color_brewer(palette = "Set1") +
+    scale_fill_brewer(palette = "Set1") +
+    labs(x = "Año", y = y.label)
+  
+}

### 2017-09-25 my_plot.R

@@ -1,0 +1,14 @@
+my_plot <- function(x) {
+  
+  p1 <- my_ggplot(x$Modelo[[1]], y.label = "Shannon")
+  p2 <- my_ggplot(x$Modelo[[2]], y.label = "Riqueza")
+  p3 <- my_ggplot(x$Modelo[[3]], y.label = "Densidad")
+  p4 <- my_ggplot(x$Modelo[[4]], y.label = "Nivel trófico")
+  p5 <- my_ggplot(x$Modelo[[5]], y.label = "Biomasa")
+  p6 <- my_ggplot(x$Modelo[[6]], y.label = "Shannon")
+  p7 <- my_ggplot(x$Modelo[[7]], y.label = "Riqueza")
+  p8 <- my_ggplot(x$Modelo[[8]], y.label = "Densidad")
+  
+  gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8)
+  
+}

### 2017-09-25 Manuscript.Rmd
Actualice manuscrito

@@
 subtitle: "Casos de estudio del Pacífico, Golfo de California, y Sistema Arrecifal Mesoamericano"
+author: "Juan Carlos Villaseñor-Derbez, Stuart Fulton, Jorge Torre"
 header-includes:
@@
 - exported-references.bib
-- packages.bib
 csl: huct.csl
@@
 windowsFonts(Times=windowsFont("TT Times New Roman"))
-
 ```
-
-Juan Carlos Villaseñor-Derbez^1\*^, Stuart Fulton^2^, Jorge Torre^2^
-
-^1^ Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, California, United States
-
-^2^ Comunidad y Biodiversidad A.C., Calle Isla del Peruano, Guaymas, Sonora, México
 
@@
 La sobrepesca y prácticas pesqueras no sostenibles son unas de las mayores amenazas para la conservación de los ecosistemas marinos del mundo [@halpern_2008-dK; @halpern_2017-Zi].
-Esto es particularmente cierto para el caso de las pesquerías artesanales.
-La implementación de reservas marinas ( i.e.
-áreas donde la captura de una o más especies está prohibida) es una medida de manejo frecuentemente propuesta para recuperar stocks pesqueros e impulsar la productividad pesquera en aguas cercanas [@afflerbach_2014-HP; @krueck_2017-J1; @sala_2017-69].
-Recientes trabajos han demostrado que también pueden mitigar y proveer amortiguamiento ante el cambio climático [@roberts_2017-J9], solve bycatch problems [@hastings_2017-sm] y en general, incrementar la biomasa, riqueza y densidades de organismos dentro de sus fronteras [@lester_2009-Ks; @giakoumi_2017-V2; @sala_2017-69].
+La implementación de reservas marinas (*i.e.* áreas donde la captura de una o más especies está prohibida) es una medida de manejo frecuentemente propuesta para recuperar stocks pesqueros e impulsar la productividad pesquera en aguas cercanas [@afflerbach_2014-HP; @krueck_2017-J1; @sala_2017-69].
+Recientes trabajos han demostrado que también pueden mitigar y proveer amortiguamiento ante el cambio climático [@roberts_2017-J9], variabilidad ambiental [@micheli_2012-EU], resolver problemas de pesca incidental [@hastings_2017-sm] y, en general, incrementar la biomasa, riqueza y densidades de organismos dentro de sus fronteras [@lester_2009-Ks; @giakoumi_2017-V2; @sala_2017-69].
 
@@
 Sin embargo, solamente **27** de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas.
-Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, éste proceso aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores.
-La escasez de recursos monetarios y humanos de la comisión limitan también el monitoreo y vigilancia de las reservas, limitando el desempeño de la reserva.
+Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, esto aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores.
+La escasez de recursos monetarios y humanos de la limitan también el monitoreo y vigilancia de las reservas, y a su vez, el desempeño de la reserva.
 
@@
 Estas son comúnmente establecidas dentro de zonas de concesión, una forma de derechos de uso territoriales para pesquerías (TURF, en inglés).
-Al permitir a los pescadores diseñar sus propias reservas, una mayor proporción de la comunidad está de acuerdo con los perímetros establecidos, y por lo tanto los respetan [@gelcich_2015-Gw; @espinosaromero_2014-PY; @beger_2004-Y8] .
+Al permitir a los pescadores diseñar sus propias reservas, una mayor proporción de la comunidad está de acuerdo con los perímetros y reglas establecidas, y por lo tanto los respetan [@gelcich_2015-Gw; @espinosaromero_2014-PY; @beger_2004-Y8] .
 Adicionalmente, los pescadores pueden implementar sus reservas por un periodo acordado (usualmente cinco años), después del cual la reserva puede ser abierta a la pesca.
@@
 Aunque el aspecto ecológico de las reservas es importante para su éxito, su efectividad también depende del estado socioeconómico y los sistemas de gobernanza de las comunidades pesqueras.
+
 La literatura indica que diferentes características influyen en el éxito de una reserva.
-En Palau, por ejemplo, la edad (*i.e.* tiempo de implementación), tamaño y hábitat contenido son características claves que determinan la efectividad [@friedlander_2017-oI].
-Por otro lado, en el Mar Meditarráneo, @difranco_2016-Xw identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks, el ingreso a los pescadores, y se presenta una mayor aceptación social de las prácticas de manejo.
+En Palau, por ejemplo, la edad (*i.e.* tiempo transcurrido desde implementación), tamaño y hábitat contenido son características claves que determinan la efectividad [@friedlander_2017-oI].
+Por otro lado, en el Mar Meditarráneo, @difranco_2016-Xw identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks y el ingreso económicos a los pescadores, a la vez que se presenta una mayor aceptación social de las prácticas de manejo.
 En una aproximación global, @edgar_2014-UO encuentran que la procuración, edad, tamaño y aislamiento son determinantes de la efectividad de las reservas.
@@
 
-La evaluación de éstos cinco casos de estudios nos permitirá identificar patrones en los que características socioeconómicas y de gobernanza que se encuentran ligadas a la efectividad (biológica) de las reservas marinas evaluadas.
+La evaluación de éstos cinco casos de estudios nos permitirá identificar l amanera en que las características socioeconómicas y de gobernanza se relacionan con la efectividad (biológica) de las reservas marinas evaluadas.
 Los patrones identificados podrán utilizarse para informar la toma de desiciones para la implementación de la red de reservas marinas en la Región de las Grandes Islas del Golfo de California.
 
-# Materiales y Métodos
+# Métodos
 
@@
 
-## Caraterísticas de las comunidades y reservas seleccionadas
+- Maria Elena
+- Punta Herrero
+- Isla Natividad
+- Guaymas
+- Puerto Libertad (Cerro bola)
 
-## Análisis de datos
+```{r}
+suppressPackageStartupMessages({
+  library(sp)
+  library(tmap)
+  library(tmaptools)
+  library(MPAtools)
+  library(kableExtra)
+  library(tidyverse)
+})
+```
+
+```{r}
+proj <- CRS("+proj=longlat +datum=NAD27")
+
+xy1 <- data.frame(X = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087), Y = c(27.85, 19.3722361, 19.3183386, 27.969079, 29.8904957))
+coordinates(xy1) <- c("X", "Y")
+proj4string(xy1) <- proj
+res1 <- spTransform(xy1, proj)
+
+load(file = "Data/coastline_mx.Rda")
+
+tm_shape(baja2) +
+  tm_polygons() +
+  tm_shape(res1) +
+  tm_markers(shape = 21, col = "red", border.col = "black", size = 0.5)
+
+# tm_shape(baja2, bbox = bb(xlim = c(-90.269600, -86.726716),
+#                           ylim = c(18.165540, 21.536636))) +
+#   tm_polygons() +
+#   tm_shape(res1) +
+#   tm_markers(shape = 21, col = "red", border.col = "black", size = 0.5)
+```
+
+## Datos y análisis de datos
+
+Para evaluar las reservas, utilizamos tres fuentes de información.
+La información ecológica proviene de los monitoreos ecológicos realizados anualmente en las zonas reserva y control.
+Cada año, se realizan censos visuales para evaluar las comunidades de peces e invertebrados, registrando riquezas, abundancias y tallas (en peces).
+Esta información nos permite calcular los indicadores biológicos.
+La información socioeconómica proviene de los avisos de arribo de CONAPESCA.
+En este caso, se tienen registros mensuales de cada una de las especies aprovechadas por las diferentes comunidades, en los que se reportan los arribos (Kg) y el valor de los arribos ($).
+Los ingresos generados por arribos son ajustados por medio del índice de precio al consumidor.
+La información de governanza fue obtenida a nivel de comunidad, pidiendo a personas familiares con las comunidades que proveyeran la información necesaria.
+
+Dada la similitud de objetivos entre las reservas, la evaluación se realiza con los mismos indicadores. En este caso, se utilizan nueve indicadores biológicos, 4 socioeconómicos y 14 de gobernanza.
+
+```{r}
+data(IndListG)
+
+data.frame(Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Ingresos", "Arribos", "Ingresos por especies objetivo", "Arribos de especies objetivo", IndListG$Indicators), stringsAsFactors = F) %>%
+  knitr::kable(caption = "Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.", booktabs = T) %>%
+  kable_styling() %>%
+  group_rows("Biológicos", 1, 5) %>%
+  group_rows("Socioeconómicos", 6, 9) %>% 
+  group_rows("Gobernanza", 9, 23)
+```
+
+Dado que el interés principal de este trabajo es identificar las relaciones entre la efectividad de una reserva y las características socioeconómicas y de gobernanza de su comunidad, los análisis se realizan a nivel de comunidad, agrupando los datos para todas las reservas y sitios control.
+Sin embargo, para controlar por respuestas diferenciales de cada sitio (para comunidades con más de una reserva), debemos tomar en cuenta factores específicos a cada sitio.
+Para esto, modificamos el modelo de diferencia en diferencias utilizado para indicadores biológicos agregándo el término para sitio de la siguiente manera:
+
+$$I = \beta_0 + \sum \gamma Ano + \beta_1 Zona + \beta_2 Post\times Zona + \beta_3Sitio + \epsilon$$
+
+```{r}
+fish <- read.csv("Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
+
+invert <- read.csv("Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
+
+gob <- read.csv("Data/governance.csv")
+```
+
+```{r}
+source(file = "analisis.R")
+source(file = "modelo.R")
+source(file = "my_ggplot.R")
+source(file = "my_plot.R")
+source(file = "robust_se.R")
+source(file = "extract_effect.R")
+```
+
+```{r}
+
+MariaElena <- analisis(peces = fish, invert = invert, comunidad = "Maria Elena", ano.imp = 2012)
+PuntaHerrero <- analisis(peces = fish, invert = invert, comunidad = "Punta Herrero", ano.imp = 2013)
+IslaNatividad <- analisis(peces = fish, invert = invert, comunidad = "Isla Natividad", ano.imp = 2006)
+Guaymas <- analisis(peces = fish, invert = invert, comunidad = "Guaymas", ano.imp = 2012)
+PuertoLibertad <- analisis(peces = fish, invert = invert, comunidad = "Puerto Libertad", ano.imp = 2012)
+
+```
 
 # Resultados
+
+## Isla Natividad
+
+```{r}
+my_plot(IslaNatividad)
+```
+
+## Puerto Libertad
+
+```{r}
+my_plot(PuertoLibertad)
+```
+
+## Guaymas
+
+```{r}
+my_plot(Guaymas)
+```
+
+## Maria Elena
+
+```{r}
+my_plot(MariaElena)
+```
+
+## Punta Herrero
+
+```{r}
+my_plot(PuntaHerrero)
+```
+
+## Efectos generales
+
+```{r}
+rbind(extract_effect(IslaNatividad),
+      extract_effect(PuertoLibertad),
+      extract_effect(Guaymas),
+      extract_effect(MariaElena),
+      extract_effect(PuntaHerrero)) %>% 
+  mutate(Comunidad = c(rep("IN", 8),
+                       rep("PL", 8),
+                       rep("G", 8),
+                       rep("ME", 8),
+                       rep("PH", 8))) %>% 
+  ggplot(aes(x = Comunidad, y = estimate)) +
+  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0) +
+  geom_point(size = 3, color = "steelblue") +
+  coord_flip() +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  theme_bw() +
+  facet_wrap(~Indicador, scales = "free") +
+  theme(strip.background = element_rect(fill = "transparent"))
+```
 

### 2017-09-25 Manuscript.tex
Actualice manuscrito

@@
             pdftitle={Efectividad de reservas marinas a nivel nacional},
+            pdfauthor={Juan Carlos Villaseñor-Derbez, Stuart Fulton, Jorge Torre},
             pdfborder={0 0 0},
@@
 Mesoamericano}
-  \author{}
-  \preauthor{}\postauthor{}
+  \author{Juan Carlos Villaseñor-Derbez, Stuart Fulton, Jorge Torre}
+  \preauthor{\centering\large\emph}
+  \postauthor{\par}
   \date{}
@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage[table]{xcolor}
+\usepackage{wrapfig}
 \usepackage{setspace}
@@
 }
-Juan Carlos Villaseñor-Derbez\textsuperscript{1*}, Stuart
-Fulton\textsuperscript{2}, Jorge Torre\textsuperscript{2}
-
-\textsuperscript{1} Bren School of Environmental Science and Management,
-University of California Santa Barbara, Santa Barbara, California,
-United States
-
-\textsuperscript{2} Comunidad y Biodiversidad A.C., Calle Isla del
-Peruano, Guaymas, Sonora, México
-
 \section{Resumen ejecutivo}\label{resumen-ejecutivo}
@@
 mayores amenazas para la conservación de los ecosistemas marinos del
-mundo (Halpern et al., 2008, 2017). Esto es particularmente cierto para
-el caso de las pesquerías artesanales. La implementación de reservas
-marinas ( i.e.~áreas donde la captura de una o más especies está
+mundo (Halpern et al., 2008, 2017). La implementación de reservas
+marinas (\emph{i.e.} áreas donde la captura de una o más especies está
 prohibida) es una medida de manejo frecuentemente propuesta para
@@
 mitigar y proveer amortiguamiento ante el cambio climático (Roberts et
-al., 2017), solve bycatch problems (Hastings, Gaines \& Costello, 2017)
-y en general, incrementar la biomasa, riqueza y densidades de organismos
+al., 2017), variabilidad ambiental (Micheli et al., 2012), resolver
+problemas de pesca incidental (Hastings, Gaines \& Costello, 2017) y, en
+general, incrementar la biomasa, riqueza y densidades de organismos
 dentro de sus fronteras (Lester et al., 2009; Giakoumi et al., 2017;
@@
 CONANP ha hecho esfuerzos importantes por involucrar a los actores
-durante la implementación de las reservas, éste proceso aún se
-caracteriza por un proceso descendente, el cual conlleva a la falta de
-cumplimiento por parte de los actores. La escasez de recursos monetarios
-y humanos de la comisión limitan también el monitoreo y vigilancia de
-las reservas, limitando el desempeño de la reserva.
+durante la implementación de las reservas, esto aún se caracteriza por
+un proceso descendente, el cual conlleva a la falta de cumplimiento por
+parte de los actores. La escasez de recursos monetarios y humanos de la
+limitan también el monitoreo y vigilancia de las reservas, y a su vez,
+el desempeño de la reserva.
 
@@
 reservas, una mayor proporción de la comunidad está de acuerdo con los
-perímetros establecidos, y por lo tanto los respetan (Beger et al.,
-2004; Espinosa-Romero et al., 2014; Gelcich \& Donlan, 2015) .
+perímetros y reglas establecidas, y por lo tanto los respetan (Beger et
+al., 2004; Espinosa-Romero et al., 2014; Gelcich \& Donlan, 2015) .
 Adicionalmente, los pescadores pueden implementar sus reservas por un
@@
 efectividad también depende del estado socioeconómico y los sistemas de
-gobernanza de las comunidades pesqueras. La literatura indica que
-diferentes características influyen en el éxito de una reserva. En
-Palau, por ejemplo, la edad (\emph{i.e.} tiempo de implementación),
-tamaño y hábitat contenido son características claves que determinan la
-efectividad (Friedlander et al., 2017). Por otro lado, en el Mar
-Meditarráneo, Di Franco et al. (2016) identifican que la procuración y
-vigilancia, presencia de un plan de manejo, participación de pescadores
-en el manejo, representación de pescadores en la toma de decisiones y
-promoción de la pesca sustentable son los cinco factores que incrementan
-la salud de los stocks, el ingreso a los pescadores, y se presenta una
-mayor aceptación social de las prácticas de manejo. En una aproximación
+gobernanza de las comunidades pesqueras.
+
+La literatura indica que diferentes características influyen en el éxito
+de una reserva. En Palau, por ejemplo, la edad (\emph{i.e.} tiempo
+transcurrido desde implementación), tamaño y hábitat contenido son
+características claves que determinan la efectividad (Friedlander et
+al., 2017). Por otro lado, en el Mar Meditarráneo, Di Franco et al.
+(2016) identifican que la procuración y vigilancia, presencia de un plan
+de manejo, participación de pescadores en el manejo, representación de
+pescadores en la toma de decisiones y promoción de la pesca sustentable
+son los cinco factores que incrementan la salud de los stocks y el
+ingreso económicos a los pescadores, a la vez que se presenta una mayor
+aceptación social de las prácticas de manejo. En una aproximación
 global, Edgar et al. (2014) encuentran que la procuración, edad, tamaño
@@
 La evaluación de éstos cinco casos de estudios nos permitirá identificar
-patrones en los que características socioeconómicas y de gobernanza que
-se encuentran ligadas a la efectividad (biológica) de las reservas
-marinas evaluadas. Los patrones identificados podrán utilizarse para
-informar la toma de desiciones para la implementación de la red de
-reservas marinas en la Región de las Grandes Islas del Golfo de
-California.
+l amanera en que las características socioeconómicas y de gobernanza se
+relacionan con la efectividad (biológica) de las reservas marinas
+evaluadas. Los patrones identificados podrán utilizarse para informar la
+toma de desiciones para la implementación de la red de reservas marinas
+en la Región de las Grandes Islas del Golfo de California.
 
-\section{Materiales y Métodos}\label{materiales-y-metodos}
+\section{Métodos}\label{metodos}
 
@@
 
-\subsection{Caraterísticas de las comunidades y reservas
-seleccionadas}\label{carateristicas-de-las-comunidades-y-reservas-seleccionadas}
+\begin{itemize}
+\tightlist
+\item
+  Maria Elena
+\item
+  Punta Herrero
+\item
+  Isla Natividad
+\item
+  Guaymas
+\item
+  Puerto Libertad (Cerro bola)
+\end{itemize}
 
-\subsection{Análisis de datos}\label{analisis-de-datos}
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
+
+\subsection{Datos y análisis de datos}\label{datos-y-analisis-de-datos}
+
+Para evaluar las reservas, utilizamos tres fuentes de información. La
+información ecológica proviene de los monitoreos ecológicos realizados
+anualmente en las zonas reserva y control. Cada año, se realizan censos
+visuales para evaluar las comunidades de peces e invertebrados,
+registrando riquezas, abundancias y tallas (en peces). Esta información
+nos permite calcular los indicadores biológicos. La información
+socioeconómica proviene de los avisos de arribo de CONAPESCA. En este
+caso, se tienen registros mensuales de cada una de las especies
+aprovechadas por las diferentes comunidades, en los que se reportan los
+arribos (Kg) y el valor de los arribos (\$). Los ingresos generados por
+arribos son ajustados por medio del índice de precio al consumidor. La
+información de governanza fue obtenida a nivel de comunidad, pidiendo a
+personas familiares con las comunidades que proveyeran la información
+necesaria.
+
+Dada la similitud de objetivos entre las reservas, la evaluación se
+realiza con los mismos indicadores. En este caso, se utilizan nueve
+indicadores biológicos, 4 socioeconómicos y 14 de gobernanza.
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-3}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.}
+\centering
+\begin{tabular}[t]{l}
+\toprule
+Indicador\\
+\midrule
+\addlinespace[0.5em]
+\multicolumn{1}{l}{\textbf{Biológicos}}\\
+\hspace{1em}Índice de diversidad de shannon\\
+\hspace{1em}Riqueza\\
+\hspace{1em}Densidad\\
+\hspace{1em}Nivel trófico\\
+\hspace{1em}Biomasa\\
+\addlinespace
+\addlinespace[0.5em]
+\multicolumn{1}{l}{\textbf{Socioeconómicos}}\\
+\hspace{1em}Ingresos\\
+\hspace{1em}Arribos\\
+\hspace{1em}Ingresos por especies objetivo\\
+\addlinespace[0.5em]
+\multicolumn{1}{l}{\textbf{Gobernanza}}\\
+\hspace{1em}\hspace{1em}Arribos de especies objetivo\\
+\hspace{1em}Acceso a la pesqueria\\
+\addlinespace
+\hspace{1em}Numero de pescadores\\
+\hspace{1em}Reconocimiento legal de la reserva\\
+\hspace{1em}Tipo de reserva\\
+\hspace{1em}Grado de pesca ilegal\\
+\hspace{1em}Plan de manejo\\
+\addlinespace
+\hspace{1em}Procuracion de la reserva\\
+\hspace{1em}Tamano de la reserva\\
+\hspace{1em}Razonamiento para el diseno de la reserva\\
+\hspace{1em}Pertenencia a oragnizaciones pesqueras\\
+\hspace{1em}Tipo de organizacion pesquera\\
+\addlinespace
+\hspace{1em}Representacion\\
+\hspace{1em}Reglamentacion interna\\
+\hspace{1em}Efectividad percibida\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Dado que el interés principal de este trabajo es identificar las
+relaciones entre la efectividad de una reserva y las características
+socioeconómicas y de gobernanza de su comunidad, los análisis se
+realizan a nivel de comunidad, agrupando los datos para todas las
+reservas y sitios control. Sin embargo, para controlar por respuestas
+diferenciales de cada sitio (para comunidades con más de una reserva),
+debemos tomar en cuenta factores específicos a cada sitio. Para esto,
+modificamos el modelo de diferencia en diferencias utilizado para
+indicadores biológicos agregándo el término para sitio de la siguiente
+manera:
+
+\[I = \beta_0 + \sum \gamma Ano + \beta_1 Zona + \beta_2 Post\times Zona + \beta_3Sitio + \epsilon\]
 
 \section{Resultados}\label{resultados}
+
+\subsection{Isla Natividad}\label{isla-natividad}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
+
+\subsection{Puerto Libertad}\label{puerto-libertad}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
+
+\subsection{Guaymas}\label{guaymas}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
+
+\subsection{Maria Elena}\label{maria-elena}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-10-1.pdf}
+
+\subsection{Punta Herrero}\label{punta-herrero}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-11-1.pdf}
+
+\subsection{Efectos generales}\label{efectos-generales}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
 
@@
 
+\hypertarget{ref-micheli_2012-EU}{}
+Micheli, F., Saenz-Arroyo, A., Greenley, A., Vazquez, L., Espinoza
+Montes, J.A., Rossetto, M. \& De Leo, G.A. 2012. Evidence that marine
+reserves enhance resilience to climatic impacts. \emph{PLoS ONE}.
+7(7):e40832. DOI:
+\href{https://doi.org/10.1371/journal.pone.0040832}{10.1371/journal.pone.0040832}.
+
 \hypertarget{ref-nom049sagpesc_2014-V6}{}

### 2017-09-25 analisis.R
Agregue especies objetivo

@@ -1,8 +1,6 @@
-analisis <- function(peces, invert, comunidad, ano.imp){
+analisis <- function(peces, invert, comunidad, ano.imp, obj.spp = ""){
   
-  # "Índice de diversidad de shannon", "Riqueza", "Organismos > LT50", "Densidad", "Densidad de especies objetivo", "Nivel trófico", "Biomasa", "Biomasa de especies objetivo"
-  
-  results <- tibble(Indicador = c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI"),
+  results <- tibble(Indicador = c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI", "DensidadObj1", "DensidadObj2", "DensidadObj3", "DensidadObj4", "DensidadObj5", "DensidadObj6", "DensidadObj7"),
                     Modelo = list(NA))
   
   results$Modelo[[1]] <- shannon(data = peces, location = comunidad) %>% 
@@ -30,20 +28,101 @@ analisis <- function(peces, invert, comunidad, ano.imp){
            Ano = as.factor(Ano)) %>% 
     modelo()
   
-  results$Modelo[[6]] <- shannon(data = peces, location = comunidad) %>% 
+  results$Modelo[[6]] <- shannon(data = invert, location = comunidad) %>% 
     mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
            Ano = as.factor(Ano)) %>% 
     modelo()
   
-  results$Modelo[[7]] <- richness(data = peces, location = comunidad) %>% 
+  results$Modelo[[7]] <- richness(data = invert, location = comunidad) %>% 
     mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
            Ano = as.factor(Ano)) %>% 
     modelo()
   
-  results$Modelo[[8]] <- density(data = peces, location = comunidad) %>% 
+  results$Modelo[[8]] <- density(data = invert, location = comunidad) %>% 
     mutate(Post = ifelse(Ano <= ano.imp, 0, 1),
            Ano = as.factor(Ano)) %>% 
     modelo()
+  
+  ## Especies objetivo
+  
+  ### Densidades
+  
+  if(length(obj.spp) > 1){
+    n <-  1
[…]

### 2017-09-25 extract_effect.R
Ajuste para incluir especies objetivo

@@ -1,7 +1,14 @@
 extract_effect <- function(x){
   
-  x$Modelo %>%
+  extraction <- filter(x, !is.na(Modelo)) %$%
+    Modelo %>% 
     map_df(robust_se) %>% 
-    filter(term == "ZonaReserva:Post") %>% 
-    mutate(Indicador = c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI"))
+    filter(term == "ZonaReserva:Post")
+  
+  Indicador = c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI", "DensidadObj1", "DensidadObj2", "DensidadObj3", "DensidadObj4", "DensidadObj5", "DensidadObj6", "DensidadObj7")
+  
+  extraction$Indicador <- Indicador[1:dim(extraction)[1]]
+  
+  return(extraction)
+    
 }

### 2017-09-25 my_ggplot.R
Cambios a tamano de puntos

@@ -3,7 +3,7 @@ my_ggplot <- function(x, y.label){
   x$model %>% 
     mutate(Ano = as.numeric(as.character(Ano))) %>% 
     ggplot(aes(x = Ano, y = Indicador, color = Zona, fill = Zona)) +
-    geom_jitter(height = 0, width = 0.1) +
+    geom_jitter(height = 0, width = 0.1, size = 1, alpha = 0.5) +
     stat_summary(geom = "ribbon", fun.data = "mean_se", alpha = 0.5)+
     stat_summary(geom = "point", fun.data = "mean_se", color = "black")+
     stat_summary(geom = "line", fun.data = "mean_se", color = "black") +

### 2017-09-25 Manuscript.Rmd
Actualice reporte

@@
   library(MPAtools)
+  library(magrittr)
   library(kableExtra)
@@
 Dado que el interés principal de este trabajo es identificar las relaciones entre la efectividad de una reserva y las características socioeconómicas y de gobernanza de su comunidad, los análisis se realizan a nivel de comunidad, agrupando los datos para todas las reservas y sitios control.
-Sin embargo, para controlar por respuestas diferenciales de cada sitio (para comunidades con más de una reserva), debemos tomar en cuenta factores específicos a cada sitio.
-Para esto, modificamos el modelo de diferencia en diferencias utilizado para indicadores biológicos agregándo el término para sitio de la siguiente manera:
+Utilizando un análisis de diferencia en diferencias podemos estimar el efecto que la reserva tienen en los indicadores biológicos [@moland_2013-VP].
+En este caso, modelamos los años como factores, tomando como referencia el primer año en la serie de datos de cada comunidad.
+Modelar los años como factores reduce la estructura del modelo, y relaja el ajuste al no asumir una tendencia lineal entre años; es decir, el cambio observado entre 2006 - 2007 no deberá de ser igual al observado entre el 2009 - 2010.
+Incluimos también un término para la zona, en la que la variable toma un valor de 0 si el sitio es una zona control y de 1 si es una zona de reserva.
+Finalmente, incluimos un término de interacción entre la variable de Zona y Post.
+La variable Post toma un valor de 0 para todos los años existentes antes de la implementación de la reserva, y un valor de 1 para años posteriores a la implementación.
+En este modelo, el coeficiente $\beta_2$ representa el efecto que la reservsa tuvo sobre un indicador a través del tiempo y con respecto a los sitios control.
 
-$$I = \beta_0 + \sum \gamma Ano + \beta_1 Zona + \beta_2 Post\times Zona + \beta_3Sitio + \epsilon$$
+$$I = \beta_0 + \sum \gamma Ano + \beta_1 Zona + \beta_2 Post\times Zona + \epsilon$$
 
@@
 
-gob <- read.csv("Data/governance.csv")
+# load("Data/conapesca_2000-2015.RData")
+# 
+# conapesca %<>%
+#   filter(UnidadEconomica %in% c("Scpp jose maria azcorra scl", "Scpp cozumel scl", "Scpp pesc del banco chinchorro scl", "Scpp langosteros del caribe scl", "Scpp andres quintana roo sc de rl", "Scpp buzos y pescadores de la baja california scl"))
+
+gob <- read.csv("Data/governance.csv", strip.white = T, stringsAsFactors = T)
 ```
@@
 source(file = "modelo.R")
-source(file = "my_ggplot.R")
 source(file = "my_plot.R")
@@
 source(file = "extract_effect.R")
+
+my_ggplot <- function(x, y.label){
+
+  x$model %>% 
+    mutate(Ano = as.numeric(as.character(Ano))) %>% 
+    ggplot(aes(x = Ano, y = Indicador, color = Zona, fill = Zona)) +
+    geom_jitter(height = 0, width = 0.1, size = 1, alpha = 0.5) +
+    stat_summary(geom = "ribbon", fun.data = "mean_se", alpha = 0.5)+
+    stat_summary(geom = "point", fun.data = "mean_se", color = "black")+
+    stat_summary(geom = "line", fun.data = "mean_se", color = "black") +
+    theme_bw() +
+    theme(legend.position = "None") +
+    scale_color_brewer(palette = "Set1") +
+    scale_fill_brewer(palette = "Set1") +
+    labs(x = "Año", y = y.label)
+}
+
 ```
@@
 
-MariaElena <- analisis(peces = fish, invert = invert, comunidad = "Maria Elena", ano.imp = 2012)
-PuntaHerrero <- analisis(peces = fish, invert = invert, comunidad = "Punta Herrero", ano.imp = 2013)
-IslaNatividad <- analisis(peces = fish, invert = invert, comunidad = "Isla Natividad", ano.imp = 2006)
-Guaymas <- analisis(peces = fish, invert = invert, comunidad = "Guaymas", ano.imp = 2012)
-PuertoLibertad <- analisis(peces = fish, invert = invert, comunidad = "Puerto Libertad", ano.imp = 2012)
+MariaElena <- analisis(peces = fish,
+                       invert = invert,
+                       comunidad = "Maria Elena",
+                       ano.imp = 2012)
+
+PuntaHerrero <- analisis(peces = fish,
+                         invert = invert,
+                         comunidad = "Punta Herrero",
+                         ano.imp = 2013)
+
+IslaNatividad <- analisis(peces = fish,
+                          invert = invert,
+                          comunidad = "Isla Natividad",
+                          ano.imp = 2006,
+                          obj.spp = c("Haliotis fulgens", "Haliotis corrugata"))
+
+Guaymas <- analisis(peces = fish,
+                    invert = invert,
+                    comunidad = "Guaymas",
+                    ano.imp = 2012,
+                    obj.spp = c("Mycteroperca rosacea", "Lutjanus argentiventris", "Balistes polylepis", "Octopus spp", "Panulirus inflatus"))
+
+PuertoLibertad <- analisis(peces = fish,
+                           invert = invert,
+                           comunidad = "Puerto Libertad",
+                           ano.imp = 2012,
+                           obj.spp = c("Balistes polylepis","Lutjanus argentiventris", "Mycteroperca rosacea", "Spondylus limbatus", "Hexaplex nigritus", "Octopus bimaculatus"))
 
@@
 
-```{r}
+```{r, message = F, warning = F}
 rbind(extract_effect(IslaNatividad),
@@
       extract_effect(PuntaHerrero)) %>% 
-  mutate(Comunidad = c(rep("IN", 8),
-                       rep("PL", 8),
-                       rep("G", 8),
+  mutate(Comunidad = c(rep("IN", 10),
+                       rep("PL", 14),
+                       rep("G", 11),
                        rep("ME", 8),
-                       rep("PH", 8))) %>% 
+                       rep("PH", 8)),
+         P = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
+  filter(Indicador %in% c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI")) %>% 
   ggplot(aes(x = Comunidad, y = estimate)) +
   geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0) +
-  geom_point(size = 3, color = "steelblue") +
+  geom_point(size = 3, aes(color = P)) +
   coord_flip() +
@@
   facet_wrap(~Indicador, scales = "free") +
-  theme(strip.background = element_rect(fill = "transparent"))
+  theme(strip.background = element_rect(fill = "transparent")) +
+  labs(x = "Efecto") +
+  scale_color_brewer(palette = "Set1")
+```
+
+```{r, warning = F, message = F, fig.height = 6, fig.width = 6}
+gob_summary <- group_by(gob, Comunidad) %>% 
+  summarize(Ano = mean(Ano),
+            PescadoresAntes = mean(Num.pescadores.antes),
+            PescadoresAhora = mean(Num.pescadores.actual),
+            PescaIlegal = mean(Grado.pesca.ilegal)) %>% 
+  mutate(CambioPescadores = PescadoresAntes-PescadoresAhora,
+         Edad = 2017 - Ano) %>% 
+  select(Comunidad, PescaIlegal, Edad)
+
+rbind(extract_effect(IslaNatividad),
+      extract_effect(PuertoLibertad),
+      extract_effect(Guaymas),
+      extract_effect(MariaElena),
+      extract_effect(PuntaHerrero)) %>% 
+  mutate(Comunidad = c(rep("IN", 10),
+                       rep("PL", 14),
+                       rep("G", 11),
+                       rep("ME", 8),
+                       rep("PH", 8)),
+         P = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
+  filter(Indicador %in% c("ShannonP", "DensidadP", "TroficoP", "ShannonI", "DensidadI")) %>%
+  select(Comunidad, Indicador, estimate, P) %>% 
+  left_join(gob_summary, by = "Comunidad") %>% 
+  gather(Variable, Valor, -c(Comunidad, Indicador, estimate, P)) %>% 
+  ggplot(aes(x = Valor, y = estimate)) +
+  geom_smooth(method = "lm", color = "black") +
+  geom_point(aes(color = Comunidad)) +
+  facet_grid(Indicador~Variable, scales = "free") +
+  theme_bw() +
+  labs(x = "Valor", y = "Efecto")
+
 ```

### 2017-09-25 Manuscript.tex
Actualice reporte

@@
 realizan a nivel de comunidad, agrupando los datos para todas las
-reservas y sitios control. Sin embargo, para controlar por respuestas
-diferenciales de cada sitio (para comunidades con más de una reserva),
-debemos tomar en cuenta factores específicos a cada sitio. Para esto,
-modificamos el modelo de diferencia en diferencias utilizado para
-indicadores biológicos agregándo el término para sitio de la siguiente
-manera:
+reservas y sitios control. Utilizando un análisis de diferencia en
+diferencias podemos estimar el efecto que la reserva tienen en los
+indicadores biológicos (Moland et al., 2013). En este caso, modelamos
+los años como factores, tomando como referencia el primer año en la
+serie de datos de cada comunidad. Modelar los años como factores reduce
+la estructura del modelo, y relaja el ajuste al no asumir una tendencia
+lineal entre años; es decir, el cambio observado entre 2006 - 2007 no
+deberá de ser igual al observado entre el 2009 - 2010. Incluimos también
+un término para la zona, en la que la variable toma un valor de 0 si el
+sitio es una zona control y de 1 si es una zona de reserva. Finalmente,
+incluimos un término de interacción entre la variable de Zona y Post. La
+variable Post toma un valor de 0 para todos los años existentes antes de
+la implementación de la reserva, y un valor de 1 para años posteriores a
+la implementación. En este modelo, el coeficiente \(\beta_2\) representa
+el efecto que la reservsa tuvo sobre un indicador a través del tiempo y
+con respecto a los sitios control.
 
-\[I = \beta_0 + \sum \gamma Ano + \beta_1 Zona + \beta_2 Post\times Zona + \beta_3Sitio + \epsilon\]
+\[I = \beta_0 + \sum \gamma Ano + \beta_1 Zona + \beta_2 Post\times Zona + \epsilon\]
 
@@
 \includegraphics{Manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-13-1.pdf}
 
@@
 
+\hypertarget{ref-moland_2013-VP}{}
+Moland, E., Olsen, E.M., Knutsen, H., Garrigou, P., Espeland, S.H.,
+Kleiven, A.R., André, C. \& Knutsen, J.A. 2013. Lobster and cod benefit
+from small-scale northern marine protected areas: Inference from an
+empirical before-after control-impact study. \emph{Proc Biol Sci}.
+280(1754):20122679. DOI:
+\href{https://doi.org/10.1098/rspb.2012.2679}{10.1098/rspb.2012.2679}.
+
 \hypertarget{ref-nom049sagpesc_2014-V6}{}

### 2017-09-26 Manuscript.Rmd
Actualice reporte

@@
 ---
-title: "Efectividad de reservas marinas a nivel nacional"
-subtitle: "Casos de estudio del Pacífico, Golfo de California, y Sistema Arrecifal Mesoamericano"
+title: "Efectividad de reservas marinas en México"
 author: "Juan Carlos Villaseñor-Derbez, Stuart Fulton, Jorge Torre"
@@
 - \doublespacing
+- \renewcommand{\contentsname}{Contenidos}
 output:
@@
     fig_caption: yes
-    keep_tex: yes
     toc: yes
@@
 ---
+
+\renewcommand{\contentsname}{Whatever}
 
@@
 
-# Resumen ejecutivo
+\clearpage
 
@@
 El objetivo de este trabajo este trabajo es realizar una evaluación de la efectividad de reservas marinas en México, presentando resultados de cinco comunidades costeras como caso de estudio.
-Las comunidades utilizadas en este reporte se distribuyen a lo largo de la costa Pacífica de Baja Claifornia, el Golfo de California, y el Sistema Arrecifal Mesoamericano.
 Con el fin de obtener una visión holística del sistema, la evaluación se realizará tomando en cuenta indicadores biológicos, socioeconómicos y de gobernanza (**Referencia aqui**).
-
-La evaluación de éstos cinco casos de estudios nos permitirá identificar l amanera en que las características socioeconómicas y de gobernanza se relacionan con la efectividad (biológica) de las reservas marinas evaluadas.
+La evaluación de éstos cinco casos de estudios nos permitirá identificar la manera en que las características socioeconómicas y de gobernanza se relacionan con la efectividad (biológica) de las reservas marinas evaluadas.
 Los patrones identificados podrán utilizarse para informar la toma de desiciones para la implementación de la red de reservas marinas en la Región de las Grandes Islas del Golfo de California.
@@
 
-- Maria Elena
-- Punta Herrero
-- Isla Natividad
-- Guaymas
-- Puerto Libertad (Cerro bola)
+Las comunidades utilizadas en este reporte se distribuyen a lo largo de la costa Pacífica de Baja Claifornia (n = 1), el Golfo de California (n = 2), y el Sistema Arrecifal Mesoamericano (n = 2; Fig.
+1).
+A continuación, se describen algunas características relevantes de cada una de las comunidades, y se presenta una tabla resumiendo algunas de las características de gobernanza de las comunidades mencionadas (Tabla 1).
 
@@
   library(kableExtra)
+  library(broom)
   library(tidyverse)
@@
 
-```{r}
+```{r, message = F, warning = F}
 proj <- CRS("+proj=longlat +datum=NAD27")
 
-xy1 <- data.frame(X = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087), Y = c(27.85, 19.3722361, 19.3183386, 27.969079, 29.8904957))
+xy1 <- data.frame(X = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087), Y = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957))
 coordinates(xy1) <- c("X", "Y")
@@
 
+data(World)
+W <- spTransform(World, proj)
+W <- W[!W$name == "Mexico",]
+
+mex <- tm_shape(W, bbox = bb(baja2)) +
+  tm_polygons() +
   tm_shape(baja2) +
@@
   tm_shape(res1) +
-  tm_markers(shape = 21, col = "red", border.col = "black", size = 0.5)
+  tm_markers(shape = 21, col = "red", border.col = "black", size = 0.5) +
+  tm_grid() +
+  tm_compass(position = c("left", "bottom")) +
+  tm_scale_bar(position = c("left", "bottom"))
 
-# tm_shape(baja2, bbox = bb(xlim = c(-90.269600, -86.726716),
-#                           ylim = c(18.165540, 21.536636))) +
-#   tm_polygons() +
-#   tm_shape(res1) +
-#   tm_markers(shape = 21, col = "red", border.col = "black", size = 0.5)
+sam <- tm_shape(baja2, bbox = bb(xlim = c(-88, -86.726716),
+                          ylim = c(19.25, 19.5))) +
+  tm_polygons() +
+  tm_shape(res1) +
+  tm_markers(shape = 21, col = "red", border.col = "black", size = 0.5) +
+  tm_compass() +
+  tm_scale_bar()
+
+tmap_arrange(mex, sam, ncol = 2)
+
+```
+
+\clearpage
+
+### Isla Natividad
+
+La Isla Natividad se encuentra en la costa oeste de la Península de Baja California, donde el hábitat predominante es el bosque de kelp o sargazo gigante (*Macrocystis pyrifera*) y los arrecifes rocosos.
+En la isla, la Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California SCL realiza actividades de extracción de los recursos marinos.
+Aunque la langosta roja (*Panulirus interruptus*) es la especie más importante en términos económicos, otras especie simportantes incluyen la escama (con un enfoque en Jurel; *Seriola lalandi*), el pepino de mar (*Parastichopus parvimensis*), el erizo rojo (*Mesocentrotus franciscanus*), el caracol (*Megastraea turbanica* y *M.
+undosa*) y, hasta el 2010, el abulón (*Haliotis sp.*).
+En 2006, por medio de un proceso participativo, la cooperativa decidió establecer dos reservas marinas de manera voluntaria.
+Agentes externos a la cooperativa, como personal de Comunidad y Biodiversidad A.C., académicos de la Universidad de Stanford, y personal de la CONANP (de la oficina de Reserva de la Biósfera El Vizcaíno), también participaron en el diseño e implementación de las reservas.
+Las reservas fueron establecidas como instrumento de manejo pesquero, buscando recurperar las poblaciones de abulón y otros invertebrados.
+Al día de hoy, las reservas marinas de Isla Natividad no han recibido reconocimiento legal, pero la cooperativa ha mostrado interés por reconocerlas como Zonas de Refugio Pesquero (ZRP).
+Los pescadores tienen un sistema de turnos para vigilar la reserva día y noche desde embarcaciones patrulla.
+
+### Puerto Libertad
+
+La comunidad de Puerto Libertad se encuentra en el Golfo de California, en el estado de Sonora, donde el hábitat predominante son los arrecifes rocosos.
+A diferencia de Isla Natividad, Puerto libertad no cuenta con una única cooperativa en la comunidad.
+En su lugar, existe un Comité Comunitario de Pesca y Acuacultura (CCPA), en el que hay representación de cooperativas y permisionarios de la zona.
+Con representación del CCPA, COBI, INAPESCA Guaymas, EDF y CONAPESCA, en el 2012 se implementó una reserva marina comunitaria.
+Eel 12 de Julio del 2017 la reserva comunitaria fue establecida como ZRP tras la publicación del acuerdo en el Diario Oficial de la Federación, con una vigencia de cinco años.
+La reserva fue establecida con los objetivos de conservar especies bajo régimen de protección especial, mejorar la productividad en aguas pesqueras adyacentes, y preservar la diversidad biológica y los ecosistemas.
+En específico, busca proteger especies como *Balistes polylepis, Lutjanus argentiventris, Mycteroperca rosacea, Spondylus limbatus, Hexaplex nigritus* y *Octopus bimaculatus*.
+Las reservas no cuentan con vigilancia por parte de la comunidad.
+
+### Guaymas
+
+Guaymas se localiza aproximadamente 300 km al sur de Puerto Libertad.
+Al igual que ésta última comunidad, no existe una única cooperativa pesquera en Guaymas.
+En su lugar, existen una serie de cooperativas y permisionarios productores, Comités comunitarios, Union regional pesquera Guaymas-Empalme y el sindicato de pesca deportiva.
+Siendo uno de los puertos más productivos de México, las pesquerías locales aprovechan una diversa gama de especies.
+Asociadas a Puerto Libertad, las tres zonas de reserva de Isla San Pedro Nolasco fueron establecidas en el 2012.
+Posteriormente, al igual que Puerto Libertad, las tres reservas de Isla San Pedro Nolasco fueron reconocidas legalmente como Zonas de Refugio Pesquero el 12 de Julio del 2017, con una vigencia de cinco años.
+A lo largo del proceso de establecimiento de reservas comunitarias o ZRP, existió la participación de miembros del Comité Nolasco (C.Técnica, Pesca Artesanal, Pesca Industrial, Pesca Deportiva, Turismo, Buceo Deportivo, Academia(CIAD), OSC (COBI)), y el grupo núcelo, conformado por CONANP, SEMARNAT, SEMAR, COFETUR, INAPESCA Guaymas, CONAPESCA, SAGARHPA, CEDES, H.
+ayuntamiento de Guaymas y PROFEPA.
+Las reservas fueron implementadas buscando proteger a especies como *Mycteroperca rosacea, Lutjanus argentiventris, Balistes polylepis, Octopus sp, Panulirus inflatus* y *Panulirus interruptus*.
+Las reservas no cuentan con vigilancia por parte de la comunidad.
+
+### Maria Elena
+
+Maria Elena es una ccomunidad pesquera en la costa de Quintana Roo.
+Los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona.
+El campo pesquero es utilizado por pescadores de la SCPP Cozumel scl (de la Isla de Cozumel).
+La principal especie aprovechada por ésta organización es la langosta espinoza del caribe (*Panulirus argus*).
+La cooperativa cuenta con permiso de pesca de escama y concesion de langosta.
+En el 2012, la Cooperativa, en conjunto con la Alianza Kanan Kay, COBI, CONANP, CONAPESCA, Oceanus, Fundación Claudia y Roberto Hernández, Fundación Haciendas del Mundo Maya, establecieron ocho ZRP con una vigencia de cinco años.
+La vigilancia de las reservas se realiza por medio del equipo de vigilancia comunitaria, con apoyo de la CONANP y una embarcación -donada por COBI- utilizada para realizar recorridos frecuentes.
+
+### Punta Herrero
+
+La comunidad de Punta Herrero se encuentra aproximadamente a 15 km al sur del campo pesquero de Maria Elena.
+De igual manera, los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona, y la principal especie explotada es la langosta.
+Sin embargo, la SCPP José María Azcorra también cuenta con permisos para pesca de escama y tiburon y una concesion de langosta.
+En una réplica del ejercicio realizado en Maria Elena -con presencia de los mismos actores-, cuatro ZRP fueron establecidas en el 2013, con una vigencia de cinco años.
+El equipo de vigilancia comunitaria, con apoyo de la CONANP, se encarga de la vigilancia de las reservas.
+
+```{r}
+read.csv("Data/governance_summary.csv") %>% 
+  knitr::kable(booktabs = T, col.names = c("Comunidad", "Reservas", "Año Implementación", "Año Reconocimiento", "Tipo", "Inspección y Vigilancia", "Manejo de recursos"), caption = "Características de gobernanza generales de las cinco comunidades estudiadas.") %>%
+  kable_styling(latex_options = c("scale_down")) %>%
+  column_spec(5, width = "3cm") %>%
+  column_spec(6, width = "6cm") %>%
+  column_spec(7, width = "6cm")
 ```
@@
 
+```{r}
+fish <- read.csv("Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
+
+invert <- read.csv("Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
+
+load("Data/conapesca_2000-2015.RData")
+
+conapesca %<>%
+  filter(UnidadEconomica %in% c("Scpp cozumel scl",
+                                "Scpp jose maria azcorra scl",
+                                "Scpp buzos y pescadores de la baja california scl")) %>% 
+  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
+  group_by(UnidadEconomica, Ano, NombreComun) %>%
+  summarize(PesoDesembarcado = sum(PesoDesembarcado), PesoVivo = sum(PesoVivo)) %>%
+  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" &NombreComun == "Langosta ent. fca."))
+
+gob <- read.csv("Data/governance.csv", strip.white = T, stringsAsFactors = T)
+```
+
 Para evaluar las reservas, utilizamos tres fuentes de información.
@@
 Cada año, se realizan censos visuales para evaluar las comunidades de peces e invertebrados, registrando riquezas, abundancias y tallas (en peces).
-Esta información nos permite calcular los indicadores biológicos.
-La información socioeconómica proviene de los avisos de arribo de CONAPESCA.
-En este caso, se tienen registros mensuales de cada una de las especies aprovechadas por las diferentes comunidades, en los que se reportan los arribos (Kg) y el valor de los arribos ($).
-Los ingresos generados por arribos son ajustados por medio del índice de precio al consumidor.
-La información de governanza fue obtenida a nivel de comunidad, pidiendo a personas familiares con las comunidades que proveyeran la información necesaria.
+Esta información nos permite calcular los indicadores biológicos de manera anual.
+Al tener valores de diferentes indicadores biológicos antes y después de la implementación de las reservas, para las zonas de reserva y sitios control, tenemos un diseño muestreal de Antes-Después-Control-Impacto.
+Utilizando un análisis de diferencia en diferencias podemos estimar el efecto que la reserva tienen en los indicadores biológicos [@moland_2013-VP] con el uso de un modelo de regresión lineal múltiple:
 
-Dada la similitud de objetivos entre las reservas, la evaluación se realiza con los mismos indicadores. En este caso, se utilizan nueve indicadores biológicos, 4 socioeconómicos y 14 de gobernanza.
+$$I = \beta_0 + \sum \gamma Año + \beta_1 Zona + \beta_2 Post\times Zona + \epsilon$$
 
-```{r}
-data(IndListG)
-
-data.frame(Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Ingresos", "Arribos", "Ingresos por especies objetivo", "Arribos de especies objetivo", IndListG$Indicators), stringsAsFactors = F) %>%
-  knitr::kable(caption = "Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.", booktabs = T) %>%
-  kable_styling() %>%
-  group_rows("Biológicos", 1, 5) %>%
-  group_rows("Socioeconómicos", 6, 9) %>% 
-  group_rows("Gobernanza", 9, 23)
-```
-
-Dado que el interés principal de este trabajo es identificar las relaciones entre la efectividad de una reserva y las características socioeconómicas y de gobernanza de su comunidad, los análisis se realizan a nivel de comunidad, agrupando los datos para todas las reservas y sitios control.
-Utilizando un análisis de diferencia en diferencias podemos estimar el efecto que la reserva tienen en los indicadores biológicos [@moland_2013-VP].
 En este caso, modelamos los años como factores, tomando como referencia el primer año en la serie de datos de cada comunidad.
@@
 En este modelo, el coeficiente $\beta_2$ representa el efecto que la reservsa tuvo sobre un indicador a través del tiempo y con respecto a los sitios control.
+La siguiente figura presenta un resumen del esfuerzo muestreal anual para cada comunidad.
 
-$$I = \beta_0 + \sum \gamma Ano + \beta_1 Zona + \beta_2 Post\times Zona + \epsilon$$
+```{r, fig.cap = "Esfuerzo muestreal anual para cada comunidad.
+El color indica la zona de reserva (rojo) o sitio control (azul).
+El tamaño del círculo indica el número de transectos totales realizados por año."}
+esfuerzo_peces <- filter(fish, Comunidad %in% c("Isla Natividad", "Puerto Libertad", "Guaymas", "Maria Elena", "Punta Herrero")) %>% 
+  group_by(Comunidad, Ano, Zona, Transecto) %>% 
+  count() %>% 
+  group_by(Comunidad, Ano, Zona) %>% 
+  summarize(Transectos = n()) %>% 
+  mutate(Class = "Peces")
+
+filter(invert, Comunidad %in% c("Isla Natividad", "Puerto Libertad", "Guaymas", "Maria Elena", "Punta Herrero")) %>% 
+  group_by(Comunidad, Ano, Zona, Transecto) %>% 
+  count() %>% 
+  group_by(Comunidad, Ano, Zona) %>% 
+  summarize(Transectos = n()) %>% 
+  mutate(Class = "Invertebrados") %>% 
+  rbind(esfuerzo_peces) %>% 
+  ggplot(aes(y = Ano, x = Comunidad, color = Zona)) +
+  geom_line(position=position_dodge(1)) +
+  geom_point(aes(size = Transectos), position=position_dodge(1)) +
+  theme_bw() +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  labs(x = "Año", y = "Comunidad - tratamiento") +
+  facet_wrap(~Class, ncol = 1) +
+  theme(strip.background = element_rect(fill = "transparent")) +
+  coord_flip()
+```
+
+También incluimos información socioeconómica relevanteproveniente de los avisos de arribo de CONAPESCA.
+En este caso, se tienen registros mensuales de los recursos aprovechados por las diferentes comunidades, en los que se reportan los arribos (Toneladas) y el valor de los arribos ($).
+La información se encuentra disponible para el periodo 2001 - 2014.
+Ya que las categorías registradas por CONAPESCA son amplias y existe un nivel de error, utilizamos únicamente los arribos reportados para langosta entera fresca a nivel de cooperativa.
+Los ingresos generados por arribos son ajustados por medio del índice de precio al consumidor.
+El análisis de datos socioeconómicos se aplicó únicamente a Isla Natividad, Maria Elena y Punta Herrero, siguiendo un modelo con la forma:
+
+$$I = \beta_0 + \beta_1Post$$
+
+Que nos permite comparar el cambio en el promedio de los indicadores antes (Post = 0) y después (Post = 1) de la implementación de la reserva.
+Tanto para los indicadores biológicos como los socioeconómicos, los coeficientes fueron ajustador con el estimador de muestras heterocedásticas.
+
+La información de governanza fue obtenida a nivel de comunidad, pidiendo a personas familiares con las comunidades que proveyeran la información necesaria.
+
+Dada la similitud de objetivos entre las reservas, la evaluación se realiza con los mismos indicadores.
+En este caso, se utilizan 8 indicadores biológicos, 2 socioeconómicos y 14 de gobernanza (Tabla 2).
+Según la disponibilidad de datos, se calculó la densidad de las especies objetivo presentadas en la sección de descripción de las comunidades.
+El criterio de selección fue que cada especie debía tener, por lo menos, dos observaciones anuales para las zonas de reserva y control.
 
 ```{r}
-fish <- read.csv("Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
+data(IndListG)
 
-invert <- read.csv("Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
-
-# load("Data/conapesca_2000-2015.RData")
-# 
-# conapesca %<>%
-#   filter(UnidadEconomica %in% c("Scpp jose maria azcorra scl", "Scpp cozumel scl", "Scpp pesc del banco chinchorro scl", "Scpp langosteros del caribe scl", "Scpp andres quintana roo sc de rl", "Scpp buzos y pescadores de la baja california scl"))
-
-gob <- read.csv("Data/governance.csv", strip.white = T, stringsAsFactors = T)
+data.frame(Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Densidad de especies objetivo", "Ingresos por especies objetivo", "Arribos de especies objetivo", IndListG$Indicators), stringsAsFactors = F) %>%
+  knitr::kable(caption = "Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.", booktabs = T) %>%
+  kable_styling() %>%
+  group_rows("Biológicos", 1, 6) %>%
+  group_rows("Socioeconómicos", 7, 8) %>% 
+  group_rows("Gobernanza", 9, 22)
 ```
@@
 source(file = "modelo.R")
-source(file = "my_plot.R")
 source(file = "robust_se.R")
@@
 
-my_ggplot <- function(x, y.label){
+my_ggplot <- function(x, y.label, breaks){
 
@@
     theme(legend.position = "None") +
-    scale_color_brewer(palette = "Set1") +
-    scale_fill_brewer(palette = "Set1") +
-    labs(x = "Año", y = y.label)
+    scale_color_brewer(palette = "Set1", direction = -1) +
+    scale_fill_brewer(palette = "Set1", direction = -1) +
+    labs(x = "Año", y = y.label) +
+    scale_x_continuous(breaks = breaks) +
+    theme(text = element_text(size = 8))
+}
+
+my_plot <- function(x, breaks) {
+
+  p1 <- my_ggplot(x$Modelo[[1]], y.label = "Peces\nShannon", breaks)
+  p2 <- my_ggplot(x$Modelo[[2]], y.label = "Peces\nRiqueza (sp / transecto)", breaks)
+  p3 <- my_ggplot(x$Modelo[[3]], y.label = "Peces\nDensidad (org / transecto)", breaks)
+  p4 <- my_ggplot(x$Modelo[[4]], y.label = "Peces\nNivel trófico", breaks)
+  p5 <- my_ggplot(x$Modelo[[5]], y.label = "Peces\nBiomasa (Kg / transecto)", breaks)
+  p6 <- my_ggplot(x$Modelo[[6]], y.label = "Invert.\nShannon", breaks)
+  p7 <- my_ggplot(x$Modelo[[7]], y.label = "Invert.\nRiqueza (sp / transecto)", breaks)
+  p8 <- my_ggplot(x$Modelo[[8]], y.label = "Invert.\nDensidad (org / transecto)", breaks)
+
+  gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8)
+
 }
@@
                           ano.imp = 2006,
-                          obj.spp = c("Haliotis fulgens", "Haliotis corrugata"))
-
-Guaymas <- analisis(peces = fish,
-                    invert = invert,
-                    comunidad = "Guaymas",
-                    ano.imp = 2012,
-                    obj.spp = c("Mycteroperca rosacea", "Lutjanus argentiventris", "Balistes polylepis", "Octopus spp", "Panulirus inflatus"))
+                          obj.spp = c("Haliotis fulgens", "Haliotis corrugata", "Panulirus interruptus"))
 
@@
 
+Guaymas <- analisis(peces = fish,
+                    invert = invert,
+                    comunidad = "Guaymas",
+                    ano.imp = 2012,
+                    obj.spp = c("Mycteroperca rosacea", "Lutjanus argentiventris", "Balistes polylepis"))
+
 ```
+
+\clearpage
 
@@
 
+A continuación se presentan los resultados de cada una de las comunidades.
+Los resultados biológicos se presentarán para cada comunidad, discutiendo primero los indicadores en común con otras comunidades (Shannon, Riqueza, Densidad, Nivel Trófico, Biomasa para peces e invertebrados) y, según su caso, las densidades de las especies objetivo.
+
 ## Isla Natividad
 
-```{r}
-my_plot(IslaNatividad)
+Para las dos reservas de Isla Natividad, todos los indicadores tienen un comportamiento similar a lo largo del tiempo.
+Sin embargo, hay indicios del efecto de la reserva.
+Por ejemplo, el Índice de diversidad de Shannon para peces muestra una mejora importante.
+Al inicio de la serie de tiempo, el indicador tenía un valor menor para la reserva que para el control.
+Sin embargo, la diferencia entre reserva y control se ha reducido con el paso del tiempo, indicando un incremento en el indicador para la zona de reserva.
+Un patrón similar, pero menos evidente, se observa para la riqueza de peces.
+Sin embargo, no parece existir un efecto de la reserva en la densidad de peces; los promedios anuales de las reservas y sitios control son similares, y no var[ian con el tiempo.
+Para el caso del nivel trófico de peces, el efecto es menor y mucho más lento que para el Índice de Shannon o la riqueza.
+En este caso, vemos un comportamiento similar para ambos tratamientos hasta el 2012, donde la reserva comienza a tener una mayor dispersión (rango más amplio) en los niveles tróficos.
+Finalmente, para el 2016, se observan diferencias evidentes en el nivel trófico medio.
+Para los indicadores de invertebrados, todos muestran (en diferente magnitud) un decremento, posiblemente explicado por los eventos de hypoxia reportados en la zona [@micheli_2012-EU].
+Sin embargo, es evidente que la reservas producen un efecto de amortiguamiento ante estos cambios ambientales, pues la reducción en las reservas es menor que en los sitios control.
+
+```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Isla Natividad.
+El rojo indica zonas de reserva, el azul zonas control.
+Cada punto representa un transecto.
+Las líneas y puntos negros indican el promedio anual para cada zona.
+Las barras de error indican 1 Error Estándar con respecto del promedio."}
+my_plot(IslaNatividad, breaks = c(2006, 2009, 2012, 2015))
 ```
+
+```{r, fig.width = 4, fig.height = 6, fig.cap = "Series de tiempo densidad de especies objetivo para Isla Natividad.
+El rojo indica zonas de reserva, el azul zonas control.
+Cada punto representa un transecto.
+Las líneas y puntos negros indican el promedio anual para cada zona.
+Las barras de error indican 1 Error Estándar con respecto del promedio."}
+gridExtra::grid.arrange(my_ggplot(IslaNatividad$Modelo[[9]], y.label = "H. fulgens / transecto", breaks = c(2006, 2009, 2012, 2015)),
+                        my_ggplot(IslaNatividad$Modelo[[10]], y.label = "H. corrugata / transecto", breaks = c(2006, 2009, 2012, 2015)),
+                        my_ggplot(IslaNatividad$Modelo[[10]], y.label = "P. interruptus / transecto", breaks = c(2006, 2009, 2012, 2015)))
+```
+
+\clearpage
 
@@
 
-```{r}
-my_plot(PuertoLibertad)
+```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Puerto Libertad.
+El rojo indica zonas de reserva, el azul zonas control.
+Cada punto representa un transecto.
+Las líneas y puntos negros indican el promedio anual para cada zona.
+Las barras de error indican 1 Error Estándar con respecto del promedio."}
+my_plot(PuertoLibertad, breaks = c(2012, 2013, 2014, 2015, 2016))
 ```
+
+```{r, fig.width = 4, fig.height = 4, fig.cap = "Series de tiempo densidad de especies objetivo para Puerto Libertad.
+El rojo indica zonas de reserva, el azul zonas control.
+Cada punto representa un transecto.
+Las líneas y puntos negros indican el promedio anual para cada zona.
+Las barras de error indican 1 Error Estándar con respecto del promedio."}
+
+gridExtra::grid.arrange(my_ggplot(PuertoLibertad$Modelo[[9]], y.label = "B. polylepis / transecto", breaks = c(2012, 2013, 2014, 2015, 2016)),
+                        my_ggplot(PuertoLibertad$Modelo[[12]], y.label = "S. limbatus / transecto", breaks = c(2012, 2013, 2014, 2015, 2016)))
+```
+
+\clearpage
 
@@
 
-```{r}
-my_plot(Guaymas)
+```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Guaymas.
+El rojo indica zonas de reserva, el azul zonas control.
+Cada punto representa un transecto.
+Las líneas y puntos negros indican el promedio anual para cada zona.
+Las barras de error indican 1 Error Estándar con respecto del promedio."}
+my_plot(Guaymas, breaks = c(2011, 2012, 2013, 2014))
 ```
+
+```{r, fig.width = 4, fig.height = 6, fig.cap = "Series de tiempo densidad de especies objetivo para Guaymas.
+El rojo indica zonas de reserva, el azul zonas control.
+Cada punto representa un transecto.
+Las líneas y puntos negros indican el promedio anual para cada zona.
+Las barras de error indican 1 Error Estándar con respecto del promedio."}
+gridExtra::grid.arrange(my_ggplot(Guaymas$Modelo[[9]], y.label = "M. rosacea / transecto", breaks = c(2011, 2012, 2013, 2014)),
+                        my_ggplot(Guaymas$Modelo[[10]], y.label = "L. argentiventris / transecto", breaks = c(2011, 2012, 2013, 2014)),
+                        my_ggplot(Guaymas$Modelo[[11]], y.label = "B. polylepis / transecto", breaks = c(2011, 2012, 2013, 2014)))
+```
+
+\clearpage
 
@@
 
-```{r}
-my_plot(MariaElena)
+```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Maria Elena.
+El rojo indica zonas de reserva, el azul zonas control.
+Cada punto representa un transecto.
+Las líneas y puntos negros indican el promedio anual para cada zona.
+Las barras de error indican 1 Error Estándar con respecto del promedio."}
+
+my_plot(MariaElena, breaks = c(2012, 2013, 2014, 2015, 2016))
 ```
+
+\clearpage
 
@@
 
-```{r}
-my_plot(PuntaHerrero)
+```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Punta Herrero.
+El rojo indica zonas de reserva, el azul zonas control.
+Cada punto representa un transecto.
+Las líneas y puntos negros indican el promedio anual para cada zona.
+Las barras de error indican 1 Error Estándar con respecto del promedio."}
+
+my_plot(PuntaHerrero, breaks = c(2012, 2013, 2014, 2015, 2016))
 ```
+
+\clearpage
 
@@
       extract_effect(PuntaHerrero)) %>% 
-  mutate(Comunidad = c(rep("IN", 10),
+  mutate(Comunidad = c(rep("IN", 11),
                        rep("PL", 14),
@@
 
+## Socioeconómicos
+
+```{r}
+ggplot(conapesca, aes(x = Ano, y = PesoVivo/1000, color = UnidadEconomica)) +
+  geom_smooth(method = "lm") +
+  geom_point() +
+  theme_bw() +
+  scale_color_brewer(palette = "Set1") +
+  geom_vline(xintercept = 2006, color = "red", linetype = "dashed", size = 1) +
+  geom_vline(xintercept = 2012, color = "blue", linetype = "dashed", size = 1) +
+  geom_vline(xintercept = 2013, color = "darkgreen", linetype = "dashed", size = 1) +
+  labs(x = "Año", y = "Peso (Toneladas)")
+```
+
+```{r}
+anos <- data.frame(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl", "Scpp cozumel scl", "Scpp jose maria azcorra scl"), AnoImp = c(2006, 2012, 2013), stringsAsFactors = F)
+
+conapesca %>% 
+  ungroup() %>% 
+  mutate(UnidadEconomica = as.character(UnidadEconomica)) %>% 
+  left_join(anos, by = "UnidadEconomica") %>% 
+  mutate(Post = ifelse(Ano <= AnoImp, 0, 1)) %>% 
+  split(.$UnidadEconomica) %>% 
+  map(~ lm(PesoVivo/1000 ~ Post, data = .x)) %>% 
+  map_df(tidy) %>% 
+  mutate(model = c("Isla Natividad", "Isla Natividad", "Maria Elena", "Maria Elena", "Punta Herrero", "Punta Herrero"),
+         P1 = ifelse(p.value < 0.1, "*", ""),
+         P2 = ifelse(p.value < 0.05, "*", ""),
+         P = paste0(P1, P2),
+         estimate = formatC(estimate, digits = 2, format = "f"),
+         std.error = formatC(std.error, digits = 2, format = "f"),
+         texto = paste0(estimate, "(", std.error, ")", P)) %>% 
+  select(model, term, texto) %>% 
+  spread(model, texto) %>% 
+  knitr::kable(booktabs = T)
+
+```
+
 ```{r, warning = F, message = F, fig.height = 6, fig.width = 6}

### 2017-09-26 Manuscript.tex
Actualice reporte

@@
 {
-\setcounter{tocdepth}{2}
+\setcounter{tocdepth}{4}
 \tableofcontents
@@
 \section{Resumen ejecutivo}\label{resumen-ejecutivo}
+
+\clearpage
 
@@
 la efectividad de reservas marinas en México, presentando resultados de
-cinco comunidades costeras como caso de estudio. Las comunidades
-utilizadas en este reporte se distribuyen a lo largo de la costa
-Pacífica de Baja Claifornia, el Golfo de California, y el Sistema
-Arrecifal Mesoamericano. Con el fin de obtener una visión holística del
-sistema, la evaluación se realizará tomando en cuenta indicadores
-biológicos, socioeconómicos y de gobernanza (\textbf{Referencia aqui}).
+cinco comunidades costeras como caso de estudio. Con el fin de obtener
+una visión holística del sistema, la evaluación se realizará tomando en
+cuenta indicadores biológicos, socioeconómicos y de gobernanza
+(\textbf{Referencia aqui}). La evaluación de éstos cinco casos de
+estudios nos permitirá identificar la manera en que las características
+socioeconómicas y de gobernanza se relacionan con la efectividad
+(biológica) de las reservas marinas evaluadas. Los patrones
+identificados podrán utilizarse para informar la toma de desiciones para
+la implementación de la red de reservas marinas en la Región de las
+Grandes Islas del Golfo de California.
 
-La evaluación de éstos cinco casos de estudios nos permitirá identificar
-l amanera en que las características socioeconómicas y de gobernanza se
-relacionan con la efectividad (biológica) de las reservas marinas
-evaluadas. Los patrones identificados podrán utilizarse para informar la
-toma de desiciones para la implementación de la red de reservas marinas
-en la Región de las Grandes Islas del Golfo de California.
+\clearpage
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Maria Elena
-\item
-  Punta Herrero
-\item
-  Isla Natividad
-\item
-  Guaymas
-\item
-  Puerto Libertad (Cerro bola)
-\end{itemize}
+Las comunidades utilizadas en este reporte se distribuyen a lo largo de
+la costa Pacífica de Baja Claifornia (n = 1), el Golfo de California (n
+= 2), y el Sistema Arrecifal Mesoamericano (n = 2; Fig. 1). A
+continuación, se describen algunas características relevantes de cada
+una de las comunidades.
 
-\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
+!{[}Mapa de la
+ubicaci\href{Manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}{on
+general de las 5 comunidades estudiadas. El panel de la derecha muestra
+un acercamiento a las comunidades de el Caribe.}
+
+\subsubsection{Isla Natividad}\label{isla-natividad}
+
+La Isla Natividad se encuentra en la costa oeste de la Península de Baja
+California, donde el hábitat predominante es el bosque de kelp o sargazo
+gigante (\emph{Macrocystis pyrifera}) y los arrecifes rocosos. En la
+isla, la Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y
+Pescadores de la Baja California SCL realiza actividades de extracción
+de los recursos marinos. Aunque la langosta roja (\emph{Panulirus
+interruptus}) es la especie más importante en términos económicos, otras
+especie simportantes incluyen la escama (con un enfoque en Jurel;
+\emph{Seriola lalandi}), el pepino de mar (\emph{Parastichopus
+parvimensis}), el erizo rojo (\emph{Mesocentrotus franciscanus}), el
+caracol (\emph{Megastraea turbanica} y \emph{M. undosa}) y, hasta el
+2010, el abulón (\emph{Haliotis sp.}).
+
+En 2006, por medio de un proceso participativo, la cooperativa decidió
+establecer dos reservas marinas de manera voluntaria. Agentes externos a
+la cooperativa, como personal de Comunidad y Biodiversidad A.C.,
+académicos de la Universidad de Stanford, y personal de la CONANP (de la
+oficina de Reserva de la Biósfera El Vizcaíno), también participaron en
+el diseño e implementación de las reservas. Las reservas fueron
+establecidas como instrumento de manejo pesquero, buscando recurperar
+las poblaciones de abulón y otros invertebrados. Al día de hoy, las
+reservas marinas de Isla Natividad no han recibido reconocimiento legal,
+pero la cooperativa ha mostrado interés por reconocerlas como Zonas de
+Refugio Pesquero (ZRP).
+
+\subsubsection{Puerto Libertad}\label{puerto-libertad}
+
+La comunidad de Puerto Libertad se encuentra en el Golfo de California,
+en el estado de Sonora. A diferencia de Isla Natividad, Puerto libertad
+no cuenta con una única cooperativa en la comunidad. En su lugar, existe
+un Comité Comunitario de Pesca y Acuacultura (CCPA), en el que hay
+representación de cooperativas y permisionarios de la zona. Con
+representación del CCPA, COBI, INAPESCA Guaymas, EDF y CONAPESCA, en el
+2012 se implementó una reserva marina comunitaria. Eel 12 de Julio del
+2017 la reserva comunitaria fue establecida como ZRP tras la publicación
+del acuerdo en el Diario Oficial de la Federación, con una vigencia de
+cinco años. La reserva fue establecida con los objetivos de conservar
+especies bajo régimen de protección especial, mejorar la productividad
+en aguas pesqueras adyacentes, y preservar la diversidad biológica y los
+ecosistemas. En específico, busca proteger especies como \emph{Balistes
+polylepis, Lutjanus argentiventris, Mycteroperca rosacea, Spondylus
+limbatus, Hexaplex nigritus} y \emph{Octopus bimaculatus}.
+
+\subsubsection{Guaymas}\label{guaymas}
+
+Guaymas se localiza aproximadamente 300 km al sur de Puerto Libertad. Al
+igual que en ésta última comunidad, no existe una única cooperativa
+pesquera. En su lugar, existen una serie de cooperativas y
+permisionarios productores.
+
+Al igual que Puerto Libertad, las tres reservas de Isla San Pedro
+Nolasco fueron reconocidas legalmente como Zonas de Refugio Pesquero el
+12 de Julio del 2017, con una vigencia de cinco años.
+
+\subsubsection{Maria Elena}\label{maria-elena}
+
+\subsubsection{Punta Herrero}\label{punta-herrero}
 
@@
 \addlinespace
+\hspace{1em}Densidad de especies objetivo\\
 \addlinespace[0.5em]
 \multicolumn{1}{l}{\textbf{Socioeconómicos}}\\
-\hspace{1em}Ingresos\\
-\hspace{1em}Arribos\\
 \hspace{1em}Ingresos por especies objetivo\\
+\hspace{1em}Arribos de especies objetivo\\
 \addlinespace[0.5em]
 \multicolumn{1}{l}{\textbf{Gobernanza}}\\
-\hspace{1em}\hspace{1em}Arribos de especies objetivo\\
 \hspace{1em}Acceso a la pesqueria\\
+\hspace{1em}Numero de pescadores\\
 \addlinespace
-\hspace{1em}Numero de pescadores\\
 \hspace{1em}Reconocimiento legal de la reserva\\
@@
 \hspace{1em}Plan de manejo\\
+\hspace{1em}Procuracion de la reserva\\
 \addlinespace
-\hspace{1em}Procuracion de la reserva\\
 \hspace{1em}Tamano de la reserva\\
@@
 \hspace{1em}Tipo de organizacion pesquera\\
+\hspace{1em}Representacion\\
 \addlinespace
-\hspace{1em}Representacion\\
 \hspace{1em}Reglamentacion interna\\
-\hspace{1em}Efectividad percibida\\
+Efectividad percibida\\
 \bottomrule
@@
 
-\subsection{Isla Natividad}\label{isla-natividad}
+\subsection{Isla Natividad}\label{isla-natividad-1}
 
@@
 
-\subsection{Puerto Libertad}\label{puerto-libertad}
-
 \includegraphics{Manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
 
-\subsection{Guaymas}\label{guaymas}
+\subsection{Puerto Libertad}\label{puerto-libertad-1}
 
@@
 
-\subsection{Maria Elena}\label{maria-elena}
-
 \includegraphics{Manuscript_files/figure-latex/unnamed-chunk-10-1.pdf}
 
-\subsection{Punta Herrero}\label{punta-herrero}
+\subsection{Guaymas}\label{guaymas-1}
 
@@
 
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
+
+\subsection{Maria Elena}\label{maria-elena-1}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-13-1.pdf}
+
+\subsection{Punta Herrero}\label{punta-herrero-1}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-14-1.pdf}
+
 \subsection{Efectos generales}\label{efectos-generales}
 
-\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-15-1.pdf}
 
-\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-13-1.pdf}
+\subsection{Socioeconómicos}\label{socioeconomicos}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-16-1.pdf}
+
+\begin{tabular}{lll}
+\toprule
+term & Isla Natividad & Maria Elena\\
+\midrule
+(Intercept) & 122682.57(10666.51)** & 11410.03(3307.67)**\\
+Post & 7365.28(14605.73) & 5833.97(8751.26)\\
+\bottomrule
+\end{tabular}
+
+\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-18-1.pdf}
 

### 2017-09-28 Manuscript.Rmd
reporte terminado

@@
 El objetivo de este trabajo este trabajo es realizar una evaluación de la efectividad de reservas marinas en México, presentando resultados de cinco comunidades costeras como caso de estudio.
-Con el fin de obtener una visión holística del sistema, la evaluación se realizará tomando en cuenta indicadores biológicos, socioeconómicos y de gobernanza (**Referencia aqui**).
+Con el fin de obtener una visión holística del sistema, la evaluación se realizará tomando en cuenta indicadores biológicos, socioeconómicos y de gobernanza.
 La evaluación de éstos cinco casos de estudios nos permitirá identificar la manera en que las características socioeconómicas y de gobernanza se relacionan con la efectividad (biológica) de las reservas marinas evaluadas.
@@
 
-```{r, message = F, warning = F}
+```{r, message = F, warning = F, fig.cap = "Mapa de la localización general de las comunidades de estudio. El panel de la derecha es un acercamiento a las comunidades de Maria Elena y Punta Herrero."}
+
 proj <- CRS("+proj=longlat +datum=NAD27")
@@
 read.csv("Data/governance_summary.csv") %>% 
-  knitr::kable(booktabs = T, col.names = c("Comunidad", "Reservas", "Año Implementación", "Año Reconocimiento", "Tipo", "Inspección y Vigilancia", "Manejo de recursos"), caption = "Características de gobernanza generales de las cinco comunidades estudiadas.") %>%
+  knitr::kable(booktabs = T, col.names = c("Comunidad", "Reservas", "Año Implementación", "Año Reconocimiento", "Tipo", "Manejo de recursos", "Grado de pesca ilegal", "Inspección y Vigilancia", "Organizaciones pesqueras", "Edad"), caption = "Características de gobernanza generales de las cinco comunidades estudiadas.") %>%
   kable_styling(latex_options = c("scale_down")) %>%
   column_spec(5, width = "3cm") %>%
-  column_spec(6, width = "6cm") %>%
-  column_spec(7, width = "6cm")
+  column_spec(6, width = "5cm") %>%
+  column_spec(8, width = "5cm") %>%
+  column_spec(9, width = "5cm") %>%
+  landscape()
 ```
@@
 
+cpi <- read.csv("Data/CPI.csv", stringsAsFactors = F, strip.white = T)
+
+cpi_2014 <- cpi$CPI[cpi$Ano == 2014]
+
 conapesca %<>%
@@
   group_by(UnidadEconomica, Ano, NombreComun) %>%
-  summarize(PesoDesembarcado = sum(PesoDesembarcado), PesoVivo = sum(PesoVivo)) %>%
-  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" &NombreComun == "Langosta ent. fca."))
+  summarize(PesoDesembarcado = sum(PesoDesembarcado), 
+            PesoVivo = sum(PesoVivo),
+            Valor = sum(Valor)) %>%
+  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" &NombreComun == "Langosta ent. fca.")) %>% 
+  left_join(cpi, by = "Ano") %>% 
+  mutate(Valor = Valor * (CPI / cpi_2014))
 
@@
 
-La información de governanza fue obtenida a nivel de comunidad, pidiendo a personas familiares con las comunidades que proveyeran la información necesaria.
+La información de gobernanza fue obtenida a nivel de comunidad, pidiendo a personas familiares con las comunidades que proveyeran la información necesaria.
+La información de gobernanza no es evaluada de manera cuantitativa.
+En su lugar, esta se interpreta de manera tal que podamos comprender qué decisiones, reglas y estructuras tienen un impacto en la reserva.
 
 Dada la similitud de objetivos entre las reservas, la evaluación se realiza con los mismos indicadores.
-En este caso, se utilizan 8 indicadores biológicos, 2 socioeconómicos y 14 de gobernanza (Tabla 2).
+En este caso, se utilizan 6 indicadores biológicos, 2 socioeconómicos y 5 de gobernanza (Tabla 2).
 Según la disponibilidad de datos, se calculó la densidad de las especies objetivo presentadas en la sección de descripción de las comunidades.
@@
 ```{r}
-data(IndListG)
-
-data.frame(Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Densidad de especies objetivo", "Ingresos por especies objetivo", "Arribos de especies objetivo", IndListG$Indicators), stringsAsFactors = F) %>%
+data.frame(Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Densidad de especies objetivo", "Ingresos por especies objetivo", "Arribos de especies objetivo", "Tipo de acceso a la pesquería", "Grado de pesca ilegal", "Procuración de la reserva", "Tipo de organización pesquera", "Edad de la reserva"), stringsAsFactors = F) %>%
   knitr::kable(caption = "Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.", booktabs = T) %>%
@@
   group_rows("Socioeconómicos", 7, 8) %>% 
-  group_rows("Gobernanza", 9, 22)
+  group_rows("Gobernanza", 9, 13)
 ```
@@
 
-MariaElena <- analisis(peces = fish,
-                       invert = invert,
-                       comunidad = "Maria Elena",
-                       ano.imp = 2012)
-
-PuntaHerrero <- analisis(peces = fish,
-                         invert = invert,
-                         comunidad = "Punta Herrero",
-                         ano.imp = 2013)
-
 IslaNatividad <- analisis(peces = fish,
@@
 
+MariaElena <- analisis(peces = fish,
+                       invert = invert,
+                       comunidad = "Maria Elena",
+                       ano.imp = 2012)
+
+PuntaHerrero <- analisis(peces = fish,
+                         invert = invert,
+                         comunidad = "Punta Herrero",
+                         ano.imp = 2013,
+                         obj.spp = "Panulirus argus")
+
 ```
@@
 Los resultados biológicos se presentarán para cada comunidad, discutiendo primero los indicadores en común con otras comunidades (Shannon, Riqueza, Densidad, Nivel Trófico, Biomasa para peces e invertebrados) y, según su caso, las densidades de las especies objetivo.
+Habiendo presentado los resultados biológicos, nos enfocaremos después en los socioeconómicos y de gobernanza.
+Usaremos esta información para identificar las causas (sociales) del éxito (biológico) de las reservas.
 
-## Isla Natividad
+## Biológicos
 
-Para las dos reservas de Isla Natividad, todos los indicadores tienen un comportamiento similar a lo largo del tiempo.
+### Isla Natividad
+
+Para las dos reservas de Isla Natividad, todos los indicadores tienen un comportamiento similar a lo largo del tiempo (Fig.
+3).
 Sin embargo, hay indicios del efecto de la reserva.
@@
 El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
+Cada punto representa un transecto, y los puntos se encuentran ajustados horizontalmente para evitar traslapes.
 Las líneas y puntos negros indican el promedio anual para cada zona.
@@
 
-```{r, fig.width = 4, fig.height = 6, fig.cap = "Series de tiempo densidad de especies objetivo para Isla Natividad.
+Para el caso de las especies objetivo de Ilsa Natividad, se analizaron las densidades de tres de los invertebrados de mayor valor comercial: los abulones (*H.
+fulgens* y *H.
+corrugata*), y la langosta (*P.
+interruptus*; Fig.
+4).
+Cabe mencionar que desde 2010 la Cooperativa de Isla Natividad decidió, de manera voluntaria, cerrar la pesca de abulón como una medida para recuperar sus poblaciones.
+Por lo tanto, los efectos de la reserva (diferencias entre reserva y control) únicamente son relevantes para el periodo 2006 - 2010.
+Para *H.
+fulgens*, las densidades tuvieron un decremento en el primer par de años en la reserva, pero para 2009 y 2010 ya habían regresado a los niveles anteriores.
+Al igual que el caso de invertebrados generales, ésto pudo deberse a los eventos de bajos niveles de oxígeno.
+Por el contrario, *H.
+corruugata* mostró incrementos inmediatos en la reserva; aunque la reserve, inicialmente, presentaba densidades menores (~ org.
+/ transecto), las densidades rápidamente aumentaron en 2007 y 2008, sobrepasando a la zona control.
+Después de esto, las densidades dentro y guera de la reserva se redujeron a casi 1 org.
+/ transecto, pero la disminución fue menor para la reserva que para el sitio control, evidenciando el efecto de amortiguamiento.
+
+```{r, fig.width = 7, fig.height = 2, fig.cap = "Series de tiempo densidad de especies objetivo para Isla Natividad.
 El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
+Cada punto representa un transecto, y los puntos se encuentran ajustados horizontalmente para evitar traslapes.
 Las líneas y puntos negros indican el promedio anual para cada zona.
@@
                         my_ggplot(IslaNatividad$Modelo[[10]], y.label = "H. corrugata / transecto", breaks = c(2006, 2009, 2012, 2015)),
-                        my_ggplot(IslaNatividad$Modelo[[10]], y.label = "P. interruptus / transecto", breaks = c(2006, 2009, 2012, 2015)))
+                        my_ggplot(IslaNatividad$Modelo[[10]], y.label = "P. interruptus / transecto", breaks = c(2006, 2009, 2012, 2015)),
+             nrow = 1)
 ```
@@
 
-## Puerto Libertad
+### Puerto Libertad
+
+Para Puerto Libertad, el 2014 no se realizaron monitoreos.
+Sin embargo, aún es posible ver que la reserva parece afectar positivamente todos los indicadores (Fig.
+5).
+Aunque para el Índice de Diversidad de Shannon hubo una disminución entre 2012 y 2013, la disminución se observa para ambos sitios.
+Sin embargo, para el 2015 el valor es mayor para la reserva que para el control.
+La riqueza de peces también mostró un mejor comportamiento, en el que la riqueza del sitio control disminuyó pero aumentó para la reserva.
+La densidad de peces mostró aumentos importantes dentro de la reserva, casi duplicando de 35 org.
+/ transecto a 64 org.
+/ transecto.
+El sitio control permaneció relativamente constante, oscilando entre 24 y 27 org.
+/ transecto.
+El nivel trófico de peces mostró un comportamiento similar al del Índice de Shannon, disminuyendo de 2012 a 2013 pero recuperándose para el 2015.
+Para los indicadores de invertebrados, observamos un incremento en diversidad para el 2013, posiblemente causado por falta de depredación por peces (es el mismo año en el que el nivel trófico y la diversidad de peces disminuye).
+Independientemente de la tendencia temporal de los indicadores en la zona, la reserva siempre tiene valores más altos de diversidad y riqueza de especies.
+De igual manera, la densidad de organismos de invertebrados mostró un leve incremento para la reserva (de 35 a 42 org.
+/ transecto), mientras que el sitio control tuvo disminuciones de 60 org / transecto a 42 org / transecto.
 
@@
 
-```{r, fig.width = 4, fig.height = 4, fig.cap = "Series de tiempo densidad de especies objetivo para Puerto Libertad.
+Para Puerto Libertad, las únicas dos especies objetivo con imformación suficiente para realizar el análisis fueron *B.
+polylepis* y *S.
+limbatus* (Fig.
+6).
+Aún cuando el sitio contol muestra incrementos en la densidad de *B.
+polylepis* (2 a 11 org.
+/ transecto), el incremento es mayor para la reserva (5 a 28 org.
+/ transecto).
+En cambio, para *S.
+limbatus* no parece haber un efecto importante de la reserva, mostrando densidades similares a través del tiempo para ambos sitios.
+
+```{r, fig.width = 5, fig.height = 2, fig.cap = "Series de tiempo densidad de especies objetivo para Puerto Libertad.
 El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
+Cada punto representa un transecto, y los puntos se encuentran ajustados horizontalmente para evitar traslapes.
 Las líneas y puntos negros indican el promedio anual para cada zona.
@@
 gridExtra::grid.arrange(my_ggplot(PuertoLibertad$Modelo[[9]], y.label = "B. polylepis / transecto", breaks = c(2012, 2013, 2014, 2015, 2016)),
-                        my_ggplot(PuertoLibertad$Modelo[[12]], y.label = "S. limbatus / transecto", breaks = c(2012, 2013, 2014, 2015, 2016)))
+                        my_ggplot(PuertoLibertad$Modelo[[12]], y.label = "S. limbatus / transecto", breaks = c(2012, 2013, 2014, 2015, 2016)),
+                        nrow = 1)
 ```
@@
 
-## Guaymas
+### Guaymas
+
+Para el caso de Guaymas (Isla San Pedro Nolasco), los monitoreos se llevaron a cabo del 2011 al 2014.
+Los datos disponibles no muestran que las reservas hayan tenido un comportamiento distinto al de los sitios control a lo largo del tiempo (Fig.
+7).
+La diversidad de peces e invertebrados y riqueza de invertebrados muestran decrementos de constantes.
+La riqueza de peces, densidad de peces, nivel trófico, biomasa de peces y densidad de invertebrados presentan, a lo largo del tiempo, un comportamiento constante para cada sitio y entre sitios.
 
@@
 
-```{r, fig.width = 4, fig.height = 6, fig.cap = "Series de tiempo densidad de especies objetivo para Guaymas.
+\clearpage
+
+Las especies objetivo con suficiente información para realizar el análisis en Guaymas fueron *M.
+rosacea*, *L.
+argentiventris* y *B.
+polylepis* (Fig.
+8).
+Al igual que los indicadores biológicos generales, las densidades de especies objetivo no mostraron cambios importantes a lo largo del tiempo ni diferencias entre tratamientos, indicando así una falta de efectividad.
+
+```{r, fig.width = 7, fig.height = 2, fig.cap = "Series de tiempo densidad de especies objetivo para Guaymas.
 El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
+Cada punto representa un transecto, y los puntos se encuentran ajustados horizontalmente para evitar traslapes.
 Las líneas y puntos negros indican el promedio anual para cada zona.
@@
                         my_ggplot(Guaymas$Modelo[[10]], y.label = "L. argentiventris / transecto", breaks = c(2011, 2012, 2013, 2014)),
-                        my_ggplot(Guaymas$Modelo[[11]], y.label = "B. polylepis / transecto", breaks = c(2011, 2012, 2013, 2014)))
+                        my_ggplot(Guaymas$Modelo[[11]], y.label = "B. polylepis / transecto", breaks = c(2011, 2012, 2013, 2014)),
+                        nrow = 1)
 ```
@@
 
-## Maria Elena
+### Maria Elena
+
+Para Maria Elena, las zonas de reserva muestran cambios en indicadores como diversidad de peces, nivel trófico, diversidad de invertebrados y densidad de invertebrados (Fig.
+9).
+El Índice de Diversidad de Shannon para peces incrementó de 2.3 a 2.9 para la reserva, mientras que disminuyó de 2.7 a 2.2 para el sitio control, entre 2012 y 2016.
+El nivel trófico de peces mostró poca variación, pero los valores siempre fueron mayores para las reservas que para los sitios control.
 
@@
 
-## Punta Herrero
+### Punta Herrero
+
+Los indicadores biológicos de Punta Herrero (Fig.
+10) presentan cambios positivos para la riqueza de peces ganando, en promedio, una especie.
+La densidad de peces también mostró un incremento importante de 81 org.
+/ transecto a 132 org.
+/ transecto, mientras que el cambio en el control solo fue de 36.6 a 79 org.
+/ transecto.
+Aunque el cambio fue pequeño, la diversidad, riqueza y densidad de invertebrados también mostraron un comportamiento positivo con respecto al sitio control.
 
@@
 
+Para Punta Herrero se tuilizó a la langosta (*P.
+argus*) como especie objetivo (Fig.
+11), que mostró un incremento importante en sus densidade.
+El incremento promedio para la reserva fue de 2 org.
+/ transecto, mientras que el control mostró incrementos de 0.25 org.
+/ transecto.
+
+```{r, fig.width = 4, fig.height = 2.5, fig.cap = "Series de tiempo densidad de especies objetivo para Guaymas.
+El rojo indica zonas de reserva, el azul zonas control.
+Cada punto representa un transecto, y los puntos se encuentran ajustados horizontalmente para evitar traslapes.
+Las líneas y puntos negros indican el promedio anual para cada zona.
+Las barras de error indican 1 Error Estándar con respecto del promedio."}
+my_ggplot(PuntaHerrero$Modelo[[9]], y.label = "P. argus / transecto", breaks = c(2011, 2012, 2013, 2014, 2015, 2016))
+```
+
 \clearpage
 
-## Efectos generales
+### Efectos generales
 
-```{r, message = F, warning = F}
+A continuación se presentan los efectos estimados para cada indicador y cada comunidad (Fig.
+12).
+La siguiente figura presenta el valor del coeficiente de diferencia en diferencia ($\beta_2$) junto con los errores estándar reportados.
+El valor del coeficiente es una medida de las diferencias observadas entre la tendencia temporal de la reserva y el control.
+Por lo tanto, valores de $\beta_2 < 0$ indican que, a través del tiempo, la reserva tuvo una disminución -- o un incremento más lento-- con respecto al sitio control.
+De igual manera, $\beta_2 > 0$ indican que la reserva mostró un incremento -- o una disminución más abrupta-- con respecto al sitio control.
+De manera general, asumiermos que un valor positivo de $\beta_2$ indica que la reserva fue efectiva para ese indicador.
+
+Por lo tanto, la evaluación de efectos (*i.e.* $\beta_2$) es una recapitulación y resumen de las observaciones realizadas en las series de tiempo.
+En este caso, Puerto Libertad es la comunidad con el mejor desempeño, mostrando dos indicadores positivos y significativos y cuatro indicadores positivos pero no significativos (Fig.
+13).
+Después, Isla Natividad y Maria Elena presentaron tres indicadores positivos, pero solamente uno fue significativo.
+En el lado negativo, Maria Elena mostró dos indicadores significativamente negativos e Isla Natividad únicamente uno.
+Aunque Punta Herrero tuvo cinco indicadores positivos y solamente tres negativos, ninguno de ellos fue significativo.
+Finalmente, Guaymas únicamente presentó dos indicadores positivos (pero no significativos) y tuvo seis indicadores negativos, de los cuales uno fue significativo.
+
+```{r, message = F, warning = F, fig.cap = "Estimación del efecto de la reserva en los indicadores biológicos para cinco comunidades costeras.
+Los círculos indican el tamaño del efecto según el análisis de diferencia en diferencias, y las barras muestran el error estándar del coeficiente estimado.
+Los colores indican la significancia del coeficiente, con el color rojo indicando p < 0.05.
+PL = Puerto Libertad, PH = Punta Herrero, ME = Maria Elena, IN = Isla Natividad, G = Guaymas."}
 rbind(extract_effect(IslaNatividad),
@@
                        rep("ME", 8),
-                       rep("PH", 8)),
+                       rep("PH", 9)),
          P = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
@@
 
+```{r, fig.cap = "Resumen del número de indicadores significativos (rojos) y no significativos (azul) para ocho indicadores utilizados en la evaluación de reservas de cinco comunidades costeras.
+PL = Puerto Libertad, PH = Punta Herrero, ME = Maria Elena, IN = Isla Natividad, G = Guaymas."}
+rbind(extract_effect(IslaNatividad),
+      extract_effect(PuertoLibertad),
+      extract_effect(Guaymas),
+      extract_effect(MariaElena),
+      extract_effect(PuntaHerrero)) %>% 
+  mutate(Comunidad = c(rep("IN", 11),
+                       rep("PL", 14),
+                       rep("G", 11),
+                       rep("ME", 8),
+                       rep("PH", 9)),
+         P = ifelse(p.value <= 0.05, "p < 0.05", "p > 0.05"),
+         Pos = ifelse(estimate > 0, "positive", "negative")) %>% 
+  filter(Indicador %in% c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI")) %>%
+  mutate(Count = 1) %>% 
+  spread(Pos, Count, fill = 0) %>% 
+  group_by(Comunidad, P) %>% 
+  summarize(negative = -sum(negative),
+            positive = sum(positive)) %>%
+  gather(Direccion, Indicadores, -c(Comunidad, P)) %>% 
+  ggplot(aes(x = Comunidad, y = Indicadores, fill = P)) +
+  geom_col(color = "black", width = 0.75) +
+  coord_flip() +
+  theme_bw() +
+  scale_fill_brewer(palette = "Set1") +
+  scale_y_continuous(breaks = c(-6, -4, -2, 0, 2, 4, 6))
+```
+
+\clearpage
+
 ## Socioeconómicos
 
-```{r}
-ggplot(conapesca, aes(x = Ano, y = PesoVivo/1000, color = UnidadEconomica)) +
+Para los indicadores socioeconómicos, el análisis se realizó unicamente para las comunidades de Isla Natividad, Maria Elena y Punta Herrero (Fig.
+13).
+En este caso, se utilizó la langosta roja (*P.
+interruptus*; para Isla Natividad) y la langosta espinoza (*P.
+argus*) para Maria Elena y Punta Herrero.
+Como se observa en la siguiente figura y en la Tabla 3, los arribos de Isla Natividad incrementaron, en promedio, 7.3 $\pm$ 14.6 (M $\pm$ ES) toneladas después de la implementación de la reserva.
+Para Maria Elena, el cambio fue de 5.8 $\pm$ 8.7 toneladas, y para Punta Herrero hubo un decremento en 1.8 $\pm$ 5.13 toneladas.
+Sin embargo, estas diferencias no son significativas.
+Dado el diseño experimental, no es posible asumir que lo cambios observados son por causa de la reserva.
+
+```{r, fig.cap = "Arribos (toneladas) anuales de langosta para tres comunidades costeras de México.
+Las líneas punteadas verticales indican el año de implementación de las reservas de cada comunidad.
+Las líneas sólidas indican el ajuste de un modelo lineal, y las sombras indican intervalos de confianza (95%) al rededor de la línea de regresión.", message = F, warning = F}
+
+com_coop <- data.frame(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl", "Scpp cozumel scl", "Scpp jose maria azcorra scl"),
+                       Comunidad = c("Isla Natividad", "Maria Elena", "Punta Herrero"))
+
+left_join(conapesca, com_coop, by = "UnidadEconomica") %>% 
+  ggplot(aes(x = Ano, y = PesoVivo/1000, color = Comunidad)) +
   geom_smooth(method = "lm") +
@@
          std.error = formatC(std.error, digits = 2, format = "f"),
-         texto = paste0(estimate, "(", std.error, ")", P)) %>% 
-  select(model, term, texto) %>% 
+         texto = paste0(estimate, " (", std.error, ")", P)) %>% 
+  select(model, Variable = term, texto) %>% 
   spread(model, texto) %>% 
-  knitr::kable(booktabs = T)
+  knitr::kable(booktabs = T, caption = "Resultados del análisis de regresión de arribos de langosta.
+Los números entre paréntesis indican el Error Estándar de los coeficientes estimados.
+* Indica p < 0.1; ** Indica p < 0.05.")
 
@@
 
-```{r, warning = F, message = F, fig.height = 6, fig.width = 6}
+En el caso de los ingresos generados por los arribos previamente mencionados (Fig.
+14), Isla Natividad y Maria Elena reportan, en promedio, 14.3 millones y 1.2 millones de pesos más, respectivamente.
+Para el caso de Punta Herrero, se reportan pérdidas por 60,000 pesos.
+Las cantidasdes observadas y reportadas corresponden al valor del peso en el 2014 (el año más reciente en la serie de tiempo de los datos).
+Al igual que con los arribos, no es posible asumir que los cambios observados en los ingresos se deben exclusivamente a la reserva.
+
+```{r, message = F, warning = F}
+
+com_coop <- data.frame(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl", "Scpp cozumel scl", "Scpp jose maria azcorra scl"),
+                       Comunidad = c("Isla Natividad", "Maria Elena", "Punta Herrero"))
+
+left_join(conapesca, com_coop, by = "UnidadEconomica") %>% 
+  ggplot(aes(x = Ano, y = Valor/1000, color = Comunidad)) +
+  geom_smooth(method = "lm") +
+  geom_point() +
+  theme_bw() +
+  scale_color_brewer(palette = "Set1") +
+  geom_vline(xintercept = 2006, color = "red", linetype = "dashed", size = 1) +
+  geom_vline(xintercept = 2012, color = "blue", linetype = "dashed", size = 1) +
+  geom_vline(xintercept = 2013, color = "darkgreen", linetype = "dashed", size = 1) +
+  labs(x = "Año", y = "Valor (Miles de pesos)")
+```
+
+```{r}
+anos <- data.frame(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl", "Scpp cozumel scl", "Scpp jose maria azcorra scl"), AnoImp = c(2006, 2012, 2013), stringsAsFactors = F)
+
+conapesca %>% 
+  ungroup() %>% 
+  mutate(UnidadEconomica = as.character(UnidadEconomica)) %>% 
+  left_join(anos, by = "UnidadEconomica") %>% 
+  mutate(Post = ifelse(Ano <= AnoImp, 0, 1)) %>% 
+  split(.$UnidadEconomica) %>% 
+  map(~ lm(Valor/1000 ~ Post, data = .x)) %>% 
+  map_df(tidy) %>% 
+  mutate(model = c("Isla Natividad", "Isla Natividad", "Maria Elena", "Maria Elena", "Punta Herrero", "Punta Herrero"),
+         P1 = ifelse(p.value < 0.1, "*", ""),
+         P2 = ifelse(p.value < 0.05, "*", ""),
+         P = paste0(P1, P2),
+         estimate = formatC(estimate, digits = 2, format = "f"),
+         std.error = formatC(std.error, digits = 2, format = "f"),
+         texto = paste0(estimate, "(", std.error, ")", P)) %>% 
+  select(model, Variable = term, texto) %>% 
+  spread(model, texto) %>% 
+  knitr::kable(booktabs = T, caption = "Resultados del análisis de regresión de ingresos de langosta.
+Los números entre paréntesis indican el Error Estándar de los coeficientes estimados.
+* Indica p < 0.1; ** Indica p < 0.05.")
+
+```
+
+\clearpage
+
+## Gobernanza
+
+A continuación se presentan algunos de los indicadores numéricos de gobernanza y su relación con el tamaño de efecto (*i.e.* $\beta_2$) de los indicadores biológicos previamente presentados (Fig.
+15).
+Se observa que la edad de la reserva (*i.e.* tiempo transcurrido desde su implementación) es un buen predictor para el incrmento en densidad de invertebrados y el nivel trófico de peces.
+Éste mismo indicicador también sugiere que existe una disminución en la densidad de peces.
+Sin embargo, al comparar con otros indicadores biológicos, la interpretación es que hay menos peces, pero de mayor nivel trófico, lo cual ha sido reportado para reservas marinas en las que los depredadores regresan y muchas de las presas se encuentran ahora bajo depredación o competencia [@anderson_2014-yi].
+Por el otro lado, la percepción del nivel (de 0 a 5) de la intensidad de la pesca ilegal no parece tener un efecto importante en ninguno de los indicadores biológicos.
+
+Sin embargo, éstos no son todos los indicadores de gobernanza utilizados, y es necesario incoroporal los demás para obtener la historia completa y comprender por qué algunas comunidades fueron más exitosas que otras.
+Como se observó previamente, Puerto Libertad fue la comunidad con el mayor número de indicadores positivos.
+Aunque la reserva fue implementada desde el 2012, esta no fue reconocida hasta el verano del 2017.
+A pesar de esto y de la carencia de un programa de vigilancia establecido, la presencia de los comités comunitarios parece brindar fortaleza a la reserva, donde todos los miembros de la comunidad están de acuerdo en protegerla.
+Además, a diferencia de Guaymas --con ambientes similares, y relativamente bajo los mismos forzamientos ambientales--, es una comunidad pequeña y que se encuentra más aislada.
+De igual manera, Isla Natividad y Maria Elena, las siguientes dos comunidades con indicadore spositivos y significativos, son comunidades relativamente pequeñas y aisladas, con un buen sistema de organización por medio de cooperativas pesqueras.
+
+\clearpage
+
+```{r, warning = F, message = F, fig.height = 6, fig.width = 6, fig.cap = "Relación entre tamaño de efecto (eje y) y el valor de dos indicadores de gobernanza (eje x).
+PL = Puerto Libertad, PH = Punta Herrero, ME = Maria Elena, IN = Isla Natividad, G = Guaymas."}
 gob_summary <- group_by(gob, Comunidad) %>% 
@@
 
+Con excepción de las zonas de concesión, cantidad y tipo de instrumentos de manejo no parecen tener un efecto importante en el efecto de la reserva.
+Las reservas dentro de una zona de concesión (Isla Natividad, Maria Elena, Punta Herrero) tuvieron un mejor desempeño.
+Aunque Puerto Libertad no cuenta con una concesión, la presencia de comités comunitarios y el aislamiento relativo permiten que haya cierto uso exclusivo de los recursos.
+
+```{r, include = F, eval = F}
+rbind(extract_effect(IslaNatividad),
+      extract_effect(PuertoLibertad),
+      extract_effect(Guaymas),
+      extract_effect(MariaElena),
+      extract_effect(PuntaHerrero)) %>% 
+  mutate(Comunidad = c(rep("IN", 11),
+                       rep("PL", 14),
+                       rep("G", 11),
+                       rep("ME", 8),
+                       rep("PH", 9)),
+         P = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05"),
+         Pos = ifelse(estimate > 0, 1, 0)) %>% 
+  filter(Indicador %in% c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI")) %>% 
+  group_by(Comunidad, P) %>% 
+  summarize(Indicadores = sum(Pos)) %>% 
+  left_join(gob_summary, by = "Comunidad") %>% 
+  gather(Variable, Valor, -c(Comunidad, Indicadores, P)) %>% 
+  ggplot(aes(x = Valor, y = Indicadores)) +
+  geom_smooth(method = "lm", color = "black") +
+  geom_point(aes(color = Comunidad)) +
+  facet_grid(P~Variable, scales = "free") +
+  theme_bw() +
+  labs(x = "Valor", y = "Efecto")
+```
+
+\clearpage
+
 # Conclusiones
 
+Al evaluar las cinco comunidades costeras nos fue posible identificar características relacionadas a la efectividad.
+El primer punto es el aislamiento de la comunidad.
+Ésto ha sido reportado antes [@edgar_2014-UO] como una de las características relacionadas con la efectividad de la reserva, y ocurre cuando es difícil o costoso --incluso para los pescadores furtivos-- acceder a las aguas protegidas.
+En zonas no aisladas, la vigilancia puede ayudar a reducir el incentivo de pesca ilegal al incrementar la probabilidad de ser visto.
+Ya sea por aislamiento o por una buena vigilancia de la reserva, el resultado final es un bajo grado de pesca ilegal, que se traduce en incrementos en los indicadores biológicos.
+
+También es relevante mencionar el papel que la edad de la reserva juega.
+Si bien no es posible modificar este atributo de las reservas (a diferencia de incrementar vigilancia, por ejemplo),  en ocasiones será necesario esperar más tiempo para observar cambios a corto y mediano plazo.
+En la medida de lo posible, realizar investigación que permita estimar mejor los "tiempos de recuperación" de las reservas, puede ayudar a poner metas alcanzables dentro del periodo establecido.
+
+Más importante aún es la presencia de organizaciones pesqueras.
+Aunque el esquema de cooperativas es el más reconocido por promover el comanejo de los recursos [@mccay_2014-CN; @mccay_2017-1m], la realidad es que cualquier estructura o foro que permita el diálogo entre usuarios resulta igualmente efectiva [@mccay_1990-81].
+En este caso, la presencia de Comités Comunitarios en Puerto Libertad cumple la labor de permitir la interacción, intercambio de ideas y generación de acuerdos entre usuarios.
+
+Además, la presencia de zonas de concesión también es una fuente de éxito.
+Las reservas dentro de concesiones suelen ser más efectivas [@afflerbach_2014-HP], ya que las concesiones promueven un sentido de pertenencia sobre los recursos naturales [@mccay_2014-CN; @afflerbach_2014-HP].
+Aunque el caso más afectivo (Puerto Libertad) no cuenta con una zona de concesión, la presencia del Comité Comunitario y los tipos de acceso a la pesquería presentes, junto con el aislamiento, son suficientes para establecer que existe una cierta restricción de acceso al uso de los recursos de la zona.
+
+Para los cinco casos de estudio evaluados en este trabajo, el aislamiento y vigilancia (como reductores de pesca ilegal), la presencia de organizaciones pesqueras, poblaciones relativamente pequeñas y aisladas y con acceso exclusivo al aprovechamiento de recursos podrían explicar la efectividad observada en los indicadores biológicos.
+Aunque no es posible asumir causalidad, es razonable asumir que una comunidad organizada, aislada y con buena vigilancia podrá tener mejores resultados en sus reservas.
+Por lo tanto, en la medida de lo posible, éstas características deberan de buscarse ---o generarse-- en las comunidades candidatas al establecimiento de reservas marinas.
+
+\clearpage
+
 # Referencias

### 2017-09-28 analisis.R

@@ -1,4 +1,4 @@
-analisis <- function(peces, invert, comunidad, ano.imp, obj.spp = ""){
+analisis <- function(peces, invert, comunidad, ano.imp, obj.spp = NULL){
   
   results <- tibble(Indicador = c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI", "DensidadObj1", "DensidadObj2", "DensidadObj3", "DensidadObj4", "DensidadObj5", "DensidadObj6", "DensidadObj7"),
                     Modelo = list(NA))
@@ -47,7 +47,7 @@ analisis <- function(peces, invert, comunidad, ano.imp, obj.spp = ""){
   
   ### Densidades
   
-  if(length(obj.spp) > 1){
+  if(length(obj.spp) >= 1){
     n <-  1
     if(obj.spp[n] %in% unique(peces$GeneroEspecie)){data <- peces}
     if(obj.spp[n] %in% unique(invert$GeneroEspecie)){data <- invert}

### 2017-09-29 Manuscript.Rmd
Reporte finalizado

@@
 
-\renewcommand{\contentsname}{Whatever}
-
 ```{r setup, include=FALSE}
@@
 En México, las reservas marinas han sido comúnmente establecidas como zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por la Comisión Nacional de Áreas Naturales Protegidas (CONANP).
-Al día de hoy, **41** RBs protegen una porción del ambiente marino en México.
-Sin embargo, solamente **27** de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas.
+Al día de hoy, 36 RBs protegen una porción del ambiente marino en México.
+Sin embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas.
 Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, esto aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores.
@@
 
-Buscando promover una alternativa con procesos ascendentes para implementar reservas marinas, las OSCs comenzaron a trabajar con comunidades pesqueras para establecer reservas comunitarias [@uribe_2010-u2] .
+Buscando promover una alternativa con procesos ascendentes para implementar reservas marinas, las Organizaciones de la Sociedad Civil (OSCs) comenzaron a trabajar con comunidades pesqueras para establecer reservas comunitarias [@uribe_2010-u2] .
 Estas son comúnmente establecidas dentro de zonas de concesión, una forma de derechos de uso territoriales para pesquerías (TURF, en inglés).
@@
 Hasta este cambio regulatorio, las reservas comunitarias no contaban con el soporte legal, y eran solamente reconocidas por la comunidad.
-Al día de hoy, existen **35** ZRP establecidas en el Pacífico, Golfo de California y Caribe Mexicano.
+Al día de hoy, existen 39 ZRP establecidas en el Pacífico, Golfo de California y Caribe Mexicano.
 
@@
 
-Aunque existen tres aproximaciones generales para implementar reservas marinas en México (*i.e.* Zonas núcleo dentro de AMP, reservas comunitarias y Zonas de Refugio Pesuqero), aún no comprendemos a fondo las características sociales que permiten su efectividad.
+Aunque existen tres aproximaciones generales para implementar reservas marinas en México (*i.e.* Zonas núcleo dentro de AMP, reservas comunitarias y Zonas de Refugio Pesquero), aún no comprendemos a fondo las características sociales que permiten su efectividad.
 La ciencia de reservas marinas se ha enfocado ampliamente en los efectos biológicos que estas tienen [@lester_2009-Ks; @giakoumi_2017-V2; @sala_2017-69; @afflerbach_2014-HP; @krueck_2017-J1].
@@
 En Palau, por ejemplo, la edad (*i.e.* tiempo transcurrido desde implementación), tamaño y hábitat contenido son características claves que determinan la efectividad [@friedlander_2017-oI].
-Por otro lado, en el Mar Meditarráneo, @difranco_2016-Xw identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks y el ingreso económicos a los pescadores, a la vez que se presenta una mayor aceptación social de las prácticas de manejo.
+Por otro lado, en el Mar Mediterráneo, @difranco_2016-Xw identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks y el ingreso económicos a los pescadores, a la vez que se presenta una mayor aceptación social de las prácticas de manejo.
 En una aproximación global, @edgar_2014-UO encuentran que la procuración, edad, tamaño y aislamiento son determinantes de la efectividad de las reservas.
@@
 La evaluación de éstos cinco casos de estudios nos permitirá identificar la manera en que las características socioeconómicas y de gobernanza se relacionan con la efectividad (biológica) de las reservas marinas evaluadas.
-Los patrones identificados podrán utilizarse para informar la toma de desiciones para la implementación de la red de reservas marinas en la Región de las Grandes Islas del Golfo de California.
+Los patrones identificados podrán utilizarse para informar la toma de decisiones para la implementación de la red de reservas marinas en la Región de las Grandes Islas del Golfo de California.
 
@@
 
-Las comunidades utilizadas en este reporte se distribuyen a lo largo de la costa Pacífica de Baja Claifornia (n = 1), el Golfo de California (n = 2), y el Sistema Arrecifal Mesoamericano (n = 2; Fig.
+Las comunidades utilizadas en este reporte se distribuyen a lo largo de la costa Pacífica de Baja California (n = 1), el Golfo de California (n = 2), y el Sistema Arrecifal Mesoamericano (n = 2; Fig.
 1).
@@
 
-```{r, message = F, warning = F, fig.cap = "Mapa de la localización general de las comunidades de estudio. El panel de la derecha es un acercamiento a las comunidades de Maria Elena y Punta Herrero."}
+```{r, message = F, warning = F, fig.cap = "Mapa de la localización general de las comunidades de estudio. El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero."}
 
@@
 En la isla, la Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California SCL realiza actividades de extracción de los recursos marinos.
-Aunque la langosta roja (*Panulirus interruptus*) es la especie más importante en términos económicos, otras especie simportantes incluyen la escama (con un enfoque en Jurel; *Seriola lalandi*), el pepino de mar (*Parastichopus parvimensis*), el erizo rojo (*Mesocentrotus franciscanus*), el caracol (*Megastraea turbanica* y *M.
+Aunque la langosta roja (*Panulirus interruptus*) es la especie más importante en términos económicos, otras especies importantes incluyen la escama (con un enfoque en Jurel; *Seriola lalandi*), el pepino de mar (*Parastichopus parvimensis*), el erizo rojo (*Mesocentrotus franciscanus*), el caracol (*Megastraea turbanica* y *M.
 undosa*) y, hasta el 2010, el abulón (*Haliotis sp.*).
@@
 Agentes externos a la cooperativa, como personal de Comunidad y Biodiversidad A.C., académicos de la Universidad de Stanford, y personal de la CONANP (de la oficina de Reserva de la Biósfera El Vizcaíno), también participaron en el diseño e implementación de las reservas.
-Las reservas fueron establecidas como instrumento de manejo pesquero, buscando recurperar las poblaciones de abulón y otros invertebrados.
+Las reservas fueron establecidas como instrumento de manejo pesquero, buscando recuperar las poblaciones de abulón y otros invertebrados.
 Al día de hoy, las reservas marinas de Isla Natividad no han recibido reconocimiento legal, pero la cooperativa ha mostrado interés por reconocerlas como Zonas de Refugio Pesquero (ZRP).
@@
 Posteriormente, al igual que Puerto Libertad, las tres reservas de Isla San Pedro Nolasco fueron reconocidas legalmente como Zonas de Refugio Pesquero el 12 de Julio del 2017, con una vigencia de cinco años.
-A lo largo del proceso de establecimiento de reservas comunitarias o ZRP, existió la participación de miembros del Comité Nolasco (C.Técnica, Pesca Artesanal, Pesca Industrial, Pesca Deportiva, Turismo, Buceo Deportivo, Academia(CIAD), OSC (COBI)), y el grupo núcelo, conformado por CONANP, SEMARNAT, SEMAR, COFETUR, INAPESCA Guaymas, CONAPESCA, SAGARHPA, CEDES, H.
+A lo largo del proceso de establecimiento de reservas comunitarias o ZRP, existió la participación de miembros del Comité Nolasco (C.Técnica, Pesca Artesanal, Pesca Industrial, Pesca Deportiva, Turismo, Buceo Deportivo, Academia(CIAD), OSC (COBI)), y el grupo núcleo, conformado por CONANP, SEMARNAT, SEMAR, COFETUR, INAPESCA Guaymas, CONAPESCA, SAGARHPA, CEDES, H.
 ayuntamiento de Guaymas y PROFEPA.
@@
 
-### Maria Elena
+### María Elena
 
-Maria Elena es una ccomunidad pesquera en la costa de Quintana Roo.
+María Elena es una comunidad pesquera en la costa de Quintana Roo.
 Los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona.
 El campo pesquero es utilizado por pescadores de la SCPP Cozumel scl (de la Isla de Cozumel).
-La principal especie aprovechada por ésta organización es la langosta espinoza del caribe (*Panulirus argus*).
-La cooperativa cuenta con permiso de pesca de escama y concesion de langosta.
+La principal especie aprovechada por ésta organización es la langosta espinosa del caribe (*Panulirus argus*).
+La cooperativa cuenta con permiso de pesca de escama y concesión de langosta.
 En el 2012, la Cooperativa, en conjunto con la Alianza Kanan Kay, COBI, CONANP, CONAPESCA, Oceanus, Fundación Claudia y Roberto Hernández, Fundación Haciendas del Mundo Maya, establecieron ocho ZRP con una vigencia de cinco años.
@@
 
-La comunidad de Punta Herrero se encuentra aproximadamente a 15 km al sur del campo pesquero de Maria Elena.
+La comunidad de Punta Herrero se encuentra aproximadamente a 15 km al sur del campo pesquero de María Elena.
 De igual manera, los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona, y la principal especie explotada es la langosta.
-Sin embargo, la SCPP José María Azcorra también cuenta con permisos para pesca de escama y tiburon y una concesion de langosta.
-En una réplica del ejercicio realizado en Maria Elena -con presencia de los mismos actores-, cuatro ZRP fueron establecidas en el 2013, con una vigencia de cinco años.
+Sin embargo, la SCPP José María Azcorra también cuenta con permisos para pesca de escama y tiburón y una concesión de langosta.
+En una réplica del ejercicio realizado en María Elena -con presencia de los mismos actores-, cuatro ZRP fueron establecidas en el 2013, con una vigencia de cinco años.
 El equipo de vigilancia comunitaria, con apoyo de la CONANP, se encarga de la vigilancia de las reservas.
@@
             Valor = sum(Valor)) %>%
-  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" &NombreComun == "Langosta ent. fca.")) %>% 
+  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
   left_join(cpi, by = "Ano") %>% 
@@
 La variable Post toma un valor de 0 para todos los años existentes antes de la implementación de la reserva, y un valor de 1 para años posteriores a la implementación.
-En este modelo, el coeficiente $\beta_2$ representa el efecto que la reservsa tuvo sobre un indicador a través del tiempo y con respecto a los sitios control.
+En este modelo, el coeficiente $\beta_2$ representa el efecto que la reservas tuvo sobre un indicador a través del tiempo y con respecto a los sitios control.
 La siguiente figura presenta un resumen del esfuerzo muestreal anual para cada comunidad.
@@
 
-También incluimos información socioeconómica relevanteproveniente de los avisos de arribo de CONAPESCA.
+También incluimos información socioeconómica relevante proveniente de los avisos de arribo de CONAPESCA.
 En este caso, se tienen registros mensuales de los recursos aprovechados por las diferentes comunidades, en los que se reportan los arribos (Toneladas) y el valor de los arribos ($).
@@
 Los ingresos generados por arribos son ajustados por medio del índice de precio al consumidor.
-El análisis de datos socioeconómicos se aplicó únicamente a Isla Natividad, Maria Elena y Punta Herrero, siguiendo un modelo con la forma:
+El análisis de datos socioeconómicos se aplicó únicamente a Isla Natividad, María Elena y Punta Herrero, siguiendo un modelo con la forma:
 
@@
 Un patrón similar, pero menos evidente, se observa para la riqueza de peces.
-Sin embargo, no parece existir un efecto de la reserva en la densidad de peces; los promedios anuales de las reservas y sitios control son similares, y no var[ian con el tiempo.
+Sin embargo, no parece existir un efecto de la reserva en la densidad de peces; los promedios anuales de las reservas y sitios control son similares, y no varían con el tiempo.
 Para el caso del nivel trófico de peces, el efecto es menor y mucho más lento que para el Índice de Shannon o la riqueza.
@@
 Finalmente, para el 2016, se observan diferencias evidentes en el nivel trófico medio.
-Para los indicadores de invertebrados, todos muestran (en diferente magnitud) un decremento, posiblemente explicado por los eventos de hypoxia reportados en la zona [@micheli_2012-EU].
+Para los indicadores de invertebrados, todos muestran (en diferente magnitud) un decremento, posiblemente explicado por los eventos de hipoxia reportados en la zona [@micheli_2012-EU].
 Sin embargo, es evidente que la reservas producen un efecto de amortiguamiento ante estos cambios ambientales, pues la reducción en las reservas es menor que en los sitios control.
@@
 
-Para el caso de las especies objetivo de Ilsa Natividad, se analizaron las densidades de tres de los invertebrados de mayor valor comercial: los abulones (*H.
+Para el caso de las especies objetivo de Isla Natividad, se analizaron las densidades de tres de los invertebrados de mayor valor comercial: los abulones (*H.
 fulgens* y *H.
@@
 / transecto), las densidades rápidamente aumentaron en 2007 y 2008, sobrepasando a la zona control.
-Después de esto, las densidades dentro y guera de la reserva se redujeron a casi 1 org.
+Después de esto, las densidades dentro y fuera de la reserva se redujeron a casi 1 org.
 / transecto, pero la disminución fue menor para la reserva que para el sitio control, evidenciando el efecto de amortiguamiento.
@@
 
-Para Puerto Libertad, las únicas dos especies objetivo con imformación suficiente para realizar el análisis fueron *B.
+Para Puerto Libertad, las únicas dos especies objetivo con información suficiente para realizar el análisis fueron *B.
 polylepis* y *S.
@@
 6).
-Aún cuando el sitio contol muestra incrementos en la densidad de *B.
+Aún cuando el sitio control muestra incrementos en la densidad de *B.
 polylepis* (2 a 11 org.
@@
 
-### Maria Elena
+### María Elena
 
-Para Maria Elena, las zonas de reserva muestran cambios en indicadores como diversidad de peces, nivel trófico, diversidad de invertebrados y densidad de invertebrados (Fig.
+Para María Elena, las zonas de reserva muestran cambios en indicadores como diversidad de peces, nivel trófico, diversidad de invertebrados y densidad de invertebrados (Fig.
 9).
@@
 
-```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Maria Elena.
+```{r, fig.cap = "Series de tiempo de los indicadores biológicos para María Elena.
 El rojo indica zonas de reserva, el azul zonas control.
@@
 
-Para Punta Herrero se tuilizó a la langosta (*P.
+Para Punta Herrero se utilizó a la langosta (*P.
 argus*) como especie objetivo (Fig.
-11), que mostró un incremento importante en sus densidade.
+11), que mostró un incremento importante en sus densidades.
 El incremento promedio para la reserva fue de 2 org.
@@
 De igual manera, $\beta_2 > 0$ indican que la reserva mostró un incremento -- o una disminución más abrupta-- con respecto al sitio control.
-De manera general, asumiermos que un valor positivo de $\beta_2$ indica que la reserva fue efectiva para ese indicador.
+De manera general, asumiremos que un valor positivo de $\beta_2$ indica que la reserva fue efectiva para ese indicador.
 
@@
 13).
-Después, Isla Natividad y Maria Elena presentaron tres indicadores positivos, pero solamente uno fue significativo.
-En el lado negativo, Maria Elena mostró dos indicadores significativamente negativos e Isla Natividad únicamente uno.
+Después, Isla Natividad y María Elena presentaron tres indicadores positivos, pero solamente uno fue significativo.
+En el lado negativo, María Elena mostró dos indicadores significativamente negativos e Isla Natividad únicamente uno.
 Aunque Punta Herrero tuvo cinco indicadores positivos y solamente tres negativos, ninguno de ellos fue significativo.
@@
 Los colores indican la significancia del coeficiente, con el color rojo indicando p < 0.05.
-PL = Puerto Libertad, PH = Punta Herrero, ME = Maria Elena, IN = Isla Natividad, G = Guaymas."}
+PL = Puerto Libertad, PH = Punta Herrero, ME = María Elena, IN = Isla Natividad, G = Guaymas."}
+
+suppressWarnings(
 rbind(extract_effect(IslaNatividad),
@@
   labs(x = "Efecto") +
-  scale_color_brewer(palette = "Set1")
+  scale_color_brewer(palette = "Set1"))
 ```
 
-```{r, fig.cap = "Resumen del número de indicadores significativos (rojos) y no significativos (azul) para ocho indicadores utilizados en la evaluación de reservas de cinco comunidades costeras.
-PL = Puerto Libertad, PH = Punta Herrero, ME = Maria Elena, IN = Isla Natividad, G = Guaymas."}
+```{r, message = F, warning = F, fig.cap = "Resumen del número de indicadores significativos (rojos) y no significativos (azul) para ocho indicadores utilizados en la evaluación de reservas de cinco comunidades costeras.
+PL = Puerto Libertad, PH = Punta Herrero, ME = María Elena, IN = Isla Natividad, G = Guaymas."}
 rbind(extract_effect(IslaNatividad),
@@
 
-Para los indicadores socioeconómicos, el análisis se realizó unicamente para las comunidades de Isla Natividad, Maria Elena y Punta Herrero (Fig.
+Para los indicadores socioeconómicos, el análisis se realizó únicamente para las comunidades de Isla Natividad, María Elena y Punta Herrero (Fig.
 13).
 En este caso, se utilizó la langosta roja (*P.
-interruptus*; para Isla Natividad) y la langosta espinoza (*P.
-argus*) para Maria Elena y Punta Herrero.
+interruptus*; para Isla Natividad) y la langosta espinosa (*P.
+argus*) para María Elena y Punta Herrero.
 Como se observa en la siguiente figura y en la Tabla 3, los arribos de Isla Natividad incrementaron, en promedio, 7.3 $\pm$ 14.6 (M $\pm$ ES) toneladas después de la implementación de la reserva.
-Para Maria Elena, el cambio fue de 5.8 $\pm$ 8.7 toneladas, y para Punta Herrero hubo un decremento en 1.8 $\pm$ 5.13 toneladas.
+Para María Elena, el cambio fue de 5.8 $\pm$ 8.7 toneladas, y para Punta Herrero hubo un decremento en 1.8 $\pm$ 5.13 toneladas.
 Sin embargo, estas diferencias no son significativas.
@@
 com_coop <- data.frame(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl", "Scpp cozumel scl", "Scpp jose maria azcorra scl"),
-                       Comunidad = c("Isla Natividad", "Maria Elena", "Punta Herrero"))
+                       Comunidad = c("Isla Natividad", "María Elena", "Punta Herrero"))
 
@@
   map_df(tidy) %>% 
-  mutate(model = c("Isla Natividad", "Isla Natividad", "Maria Elena", "Maria Elena", "Punta Herrero", "Punta Herrero"),
+  mutate(model = c("Isla Natividad", "Isla Natividad", "María Elena", "María Elena", "Punta Herrero", "Punta Herrero"),
          P1 = ifelse(p.value < 0.1, "*", ""),
@@
 En el caso de los ingresos generados por los arribos previamente mencionados (Fig.
-14), Isla Natividad y Maria Elena reportan, en promedio, 14.3 millones y 1.2 millones de pesos más, respectivamente.
+14), Isla Natividad y María Elena reportan, en promedio, 14.3 millones y 1.2 millones de pesos más, respectivamente.
 Para el caso de Punta Herrero, se reportan pérdidas por 60,000 pesos.
-Las cantidasdes observadas y reportadas corresponden al valor del peso en el 2014 (el año más reciente en la serie de tiempo de los datos).
+Las cantidades observadas y reportadas corresponden al valor del peso en el 2014 (el año más reciente en la serie de tiempo de los datos).
 Al igual que con los arribos, no es posible asumir que los cambios observados en los ingresos se deben exclusivamente a la reserva.
@@
 15).
-Se observa que la edad de la reserva (*i.e.* tiempo transcurrido desde su implementación) es un buen predictor para el incrmento en densidad de invertebrados y el nivel trófico de peces.
-Éste mismo indicicador también sugiere que existe una disminución en la densidad de peces.
+Se observa que la edad de la reserva (*i.e.* tiempo transcurrido desde su implementación) es un buen predictor para el incremento en densidad de invertebrados y el nivel trófico de peces.
+Éste mismo indicador también sugiere que existe una disminución en la densidad de peces.
 Sin embargo, al comparar con otros indicadores biológicos, la interpretación es que hay menos peces, pero de mayor nivel trófico, lo cual ha sido reportado para reservas marinas en las que los depredadores regresan y muchas de las presas se encuentran ahora bajo depredación o competencia [@anderson_2014-yi].
@@
 
-Sin embargo, éstos no son todos los indicadores de gobernanza utilizados, y es necesario incoroporal los demás para obtener la historia completa y comprender por qué algunas comunidades fueron más exitosas que otras.
+Sin embargo, éstos no son todos los indicadores de gobernanza utilizados, y es necesario incorporar los demás para obtener la historia completa y comprender por qué algunas comunidades fueron más exitosas que otras.
 Como se observó previamente, Puerto Libertad fue la comunidad con el mayor número de indicadores positivos.
@@
 Además, a diferencia de Guaymas --con ambientes similares, y relativamente bajo los mismos forzamientos ambientales--, es una comunidad pequeña y que se encuentra más aislada.
-De igual manera, Isla Natividad y Maria Elena, las siguientes dos comunidades con indicadore spositivos y significativos, son comunidades relativamente pequeñas y aisladas, con un buen sistema de organización por medio de cooperativas pesqueras.
+De igual manera, Isla Natividad y María Elena, las siguientes dos comunidades con indicadores positivos y significativos, son comunidades relativamente pequeñas y aisladas, con un buen sistema de organización por medio de cooperativas pesqueras.
 
@@
 ```{r, warning = F, message = F, fig.height = 6, fig.width = 6, fig.cap = "Relación entre tamaño de efecto (eje y) y el valor de dos indicadores de gobernanza (eje x).
-PL = Puerto Libertad, PH = Punta Herrero, ME = Maria Elena, IN = Isla Natividad, G = Guaymas."}
+PL = Puerto Libertad, PH = Punta Herrero, ME = María Elena, IN = Isla Natividad, G = Guaymas."}
 gob_summary <- group_by(gob, Comunidad) %>% 
@@
 Con excepción de las zonas de concesión, cantidad y tipo de instrumentos de manejo no parecen tener un efecto importante en el efecto de la reserva.
-Las reservas dentro de una zona de concesión (Isla Natividad, Maria Elena, Punta Herrero) tuvieron un mejor desempeño.
+Las reservas dentro de una zona de concesión (Isla Natividad, María Elena, Punta Herrero) tuvieron un mejor desempeño.
 Aunque Puerto Libertad no cuenta con una concesión, la presencia de comités comunitarios y el aislamiento relativo permiten que haya cierto uso exclusivo de los recursos.
@@
 Más importante aún es la presencia de organizaciones pesqueras.
-Aunque el esquema de cooperativas es el más reconocido por promover el comanejo de los recursos [@mccay_2014-CN; @mccay_2017-1m], la realidad es que cualquier estructura o foro que permita el diálogo entre usuarios resulta igualmente efectiva [@mccay_1990-81].
+Aunque el esquema de cooperativas es el más reconocido por promover el co--manejo de los recursos [@mccay_2014-CN; @mccay_2017-1m], la realidad es que cualquier estructura o foro que permita el diálogo entre usuarios resulta igualmente efectiva [@mccay_1990-81].
 En este caso, la presencia de Comités Comunitarios en Puerto Libertad cumple la labor de permitir la interacción, intercambio de ideas y generación de acuerdos entre usuarios.
@@
 Aunque no es posible asumir causalidad, es razonable asumir que una comunidad organizada, aislada y con buena vigilancia podrá tener mejores resultados en sus reservas.
-Por lo tanto, en la medida de lo posible, éstas características deberan de buscarse ---o generarse-- en las comunidades candidatas al establecimiento de reservas marinas.
+Por lo tanto, en la medida de lo posible, éstas características deberán de buscarse ---o generarse-- en las comunidades candidatas al establecimiento de reservas marinas.
 

### 2018-01-11 Frontiers_LaTex_Templates/Villasenor-Derbez.Rmd
Added latex template for manuscript skeleton

--- 2017-09-26 Manuscript.Rmd
@@
 ---
-title: "Efectividad de reservas marinas en México"
-author: "Juan Carlos Villaseñor-Derbez, Stuart Fulton, Jorge Torre"
-header-includes:
-- \usepackage{setspace}
-- \doublespacing
-- \renewcommand{\contentsname}{Contenidos}
+title: "Three archetypes of no-take marine reserves yield positive (but idiosyncratic) social and ecological benefits in Mexico"
+rtitle: "Mexican marine reserves"
+author: "Juan Carlos Villaseñor-Derbez,$^{1,*}$, Stuart Fulton,$^{2}$ Jorge Torre,$^{2}$"
+institution: "
+$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
+$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico
+"
+corrauth: "Me, at home"
+email: "jvillasenor@bren.ucsb.edu"
+abstract: ""
+keywords: "Marine Reserves, Marine Conservation, Fisheries, Citizen Science, Mexico"
+bibliography: "../exported-references.bib"
 output:
   pdf_document:
-    fig_caption: yes
-    toc: yes
-fontsize: 12pt
-bibliography:
-- exported-references.bib
-csl: huct.csl
+    keep_tex: yes
+    template: template.tex
+csl: FMS.csl
 ---
 
-\renewcommand{\contentsname}{Whatever}
-
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F)
-options(knitr.table.format = "latex")
-
-windowsFonts(Times=windowsFont("TT Times New Roman"))
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-\clearpage
-
-# Introducción
+# Introduction
 
@@
 
-[^1]: Hasta ahora, solamente una comunidad ha decidido abrir sur reservas a la pesca.
+# Materials and Methods
 
-Sin embargo, en el 2014 una nueva norma [@nom049sagpesc_2014-V6] permite a los pescadores solicitar el establecimiento de reservas marinas bajo el nombre de “Zonas de refugio Pesquero” (ZRP).
-El manejo de las ZRP combina procesos ascendentes y descendentes al reconocer legalmente las reservas propuestas por las comunidades.
-Posterior a la revisión por parte de la Comisión Nacional de Acuacultura y Pesca (CONAPESCA) y la opinión técnica del Instituto Nacional de Acuacultura y Pesca (INAPESCA) las ZRP son establecidas por el periodo solicitado por los pescadores [^2] .
-El monitoreo y la vigilancia de las ZRP es típicamente llevado a cabo por la comunidad , con ayuda de OSCs locales.
-Hasta este cambio regulatorio, las reservas comunitarias no contaban con el soporte legal, y eran solamente reconocidas por la comunidad.
-Al día de hoy, existen **35** ZRP establecidas en el Pacífico, Golfo de California y Caribe Mexicano.
+# Results
 
-[^2]: Existen excepciones a esto, como la “Zona de Refugio Pesquero Golfo de Ulloa” y la “Zona de Refugio Pesquero Akumal”, creadas por CONAPESCA para cerrar la pesca y prevenir la captura incidental de tortugas marinas.
+# Discussion
 
-Aunque existen tres aproximaciones generales para implementar reservas marinas en México (*i.e.* Zonas núcleo dentro de AMP, reservas comunitarias y Zonas de Refugio Pesuqero), aún no comprendemos a fondo las características sociales que permiten su efectividad.
-La ciencia de reservas marinas se ha enfocado ampliamente en los efectos biológicos que estas tienen [@lester_2009-Ks; @giakoumi_2017-V2; @sala_2017-69; @afflerbach_2014-HP; @krueck_2017-J1].
-Aunque el aspecto ecológico de las reservas es importante para su éxito, su efectividad también depende del estado socioeconómico y los sistemas de gobernanza de las comunidades pesqueras.
-
-La literatura indica que diferentes características influyen en el éxito de una reserva.
-En Palau, por ejemplo, la edad (*i.e.* tiempo transcurrido desde implementación), tamaño y hábitat contenido son características claves que determinan la efectividad [@friedlander_2017-oI].
-Por otro lado, en el Mar Meditarráneo, @difranco_2016-Xw identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks y el ingreso económicos a los pescadores, a la vez que se presenta una mayor aceptación social de las prácticas de manejo.
-En una aproximación global, @edgar_2014-UO encuentran que la procuración, edad, tamaño y aislamiento son determinantes de la efectividad de las reservas.
-Por lo tanto, observamos que las características que habilitan el éxito varían a través de regiones, y poco esfuerzo se ha hecho por comprender estas interacciones en México.
-
-# Objetivos
-
-El objetivo de este trabajo este trabajo es realizar una evaluación de la efectividad de reservas marinas en México, presentando resultados de cinco comunidades costeras como caso de estudio.
-Con el fin de obtener una visión holística del sistema, la evaluación se realizará tomando en cuenta indicadores biológicos, socioeconómicos y de gobernanza (**Referencia aqui**).
-La evaluación de éstos cinco casos de estudios nos permitirá identificar la manera en que las características socioeconómicas y de gobernanza se relacionan con la efectividad (biológica) de las reservas marinas evaluadas.
-Los patrones identificados podrán utilizarse para informar la toma de desiciones para la implementación de la red de reservas marinas en la Región de las Grandes Islas del Golfo de California.
-
-# Métodos
-
-## Zonas de estudio
-
-Las comunidades utilizadas en este reporte se distribuyen a lo largo de la costa Pacífica de Baja Claifornia (n = 1), el Golfo de California (n = 2), y el Sistema Arrecifal Mesoamericano (n = 2; Fig.
-1).
-A continuación, se describen algunas características relevantes de cada una de las comunidades, y se presenta una tabla resumiendo algunas de las características de gobernanza de las comunidades mencionadas (Tabla 1).
-
-```{r}
-suppressPackageStartupMessages({
-  library(sp)
-  library(tmap)
-  library(tmaptools)
-  library(MPAtools)
-  library(magrittr)
-  library(kableExtra)
-  library(broom)
-  library(tidyverse)
-})
-```
-
-```{r, message = F, warning = F}
-proj <- CRS("+proj=longlat +datum=NAD27")
-
-xy1 <- data.frame(X = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087), Y = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957))
-coordinates(xy1) <- c("X", "Y")
-proj4string(xy1) <- proj
-res1 <- spTransform(xy1, proj)
-
-load(file = "Data/coastline_mx.Rda")
-
-data(World)
-W <- spTransform(World, proj)
-W <- W[!W$name == "Mexico",]
-
-mex <- tm_shape(W, bbox = bb(baja2)) +
-  tm_polygons() +
-  tm_shape(baja2) +
-  tm_polygons() +
-  tm_shape(res1) +
-  tm_markers(shape = 21, col = "red", border.col = "black", size = 0.5) +
-  tm_grid() +
-  tm_compass(position = c("left", "bottom")) +
-  tm_scale_bar(position = c("left", "bottom"))
-
-sam <- tm_shape(baja2, bbox = bb(xlim = c(-88, -86.726716),
-                          ylim = c(19.25, 19.5))) +
-  tm_polygons() +
-  tm_shape(res1) +
-  tm_markers(shape = 21, col = "red", border.col = "black", size = 0.5) +
-  tm_compass() +
-  tm_scale_bar()
-
-tmap_arrange(mex, sam, ncol = 2)
-
-```
-
-\clearpage
-
-### Isla Natividad
-
-La Isla Natividad se encuentra en la costa oeste de la Península de Baja California, donde el hábitat predominante es el bosque de kelp o sargazo gigante (*Macrocystis pyrifera*) y los arrecifes rocosos.
-En la isla, la Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California SCL realiza actividades de extracción de los recursos marinos.
-Aunque la langosta roja (*Panulirus interruptus*) es la especie más importante en términos económicos, otras especie simportantes incluyen la escama (con un enfoque en Jurel; *Seriola lalandi*), el pepino de mar (*Parastichopus parvimensis*), el erizo rojo (*Mesocentrotus franciscanus*), el caracol (*Megastraea turbanica* y *M.
-undosa*) y, hasta el 2010, el abulón (*Haliotis sp.*).
-En 2006, por medio de un proceso participativo, la cooperativa decidió establecer dos reservas marinas de manera voluntaria.
-Agentes externos a la cooperativa, como personal de Comunidad y Biodiversidad A.C., académicos de la Universidad de Stanford, y personal de la CONANP (de la oficina de Reserva de la Biósfera El Vizcaíno), también participaron en el diseño e implementación de las reservas.
-Las reservas fueron establecidas como instrumento de manejo pesquero, buscando recurperar las poblaciones de abulón y otros invertebrados.
-Al día de hoy, las reservas marinas de Isla Natividad no han recibido reconocimiento legal, pero la cooperativa ha mostrado interés por reconocerlas como Zonas de Refugio Pesquero (ZRP).
-Los pescadores tienen un sistema de turnos para vigilar la reserva día y noche desde embarcaciones patrulla.
-
-### Puerto Libertad
-
-La comunidad de Puerto Libertad se encuentra en el Golfo de California, en el estado de Sonora, donde el hábitat predominante son los arrecifes rocosos.
-A diferencia de Isla Natividad, Puerto libertad no cuenta con una única cooperativa en la comunidad.
-En su lugar, existe un Comité Comunitario de Pesca y Acuacultura (CCPA), en el que hay representación de cooperativas y permisionarios de la zona.
-Con representación del CCPA, COBI, INAPESCA Guaymas, EDF y CONAPESCA, en el 2012 se implementó una reserva marina comunitaria.
-Eel 12 de Julio del 2017 la reserva comunitaria fue establecida como ZRP tras la publicación del acuerdo en el Diario Oficial de la Federación, con una vigencia de cinco años.
-La reserva fue establecida con los objetivos de conservar especies bajo régimen de protección especial, mejorar la productividad en aguas pesqueras adyacentes, y preservar la diversidad biológica y los ecosistemas.
-En específico, busca proteger especies como *Balistes polylepis, Lutjanus argentiventris, Mycteroperca rosacea, Spondylus limbatus, Hexaplex nigritus* y *Octopus bimaculatus*.
-Las reservas no cuentan con vigilancia por parte de la comunidad.
-
-### Guaymas
-
-Guaymas se localiza aproximadamente 300 km al sur de Puerto Libertad.
-Al igual que ésta última comunidad, no existe una única cooperativa pesquera en Guaymas.
-En su lugar, existen una serie de cooperativas y permisionarios productores, Comités comunitarios, Union regional pesquera Guaymas-Empalme y el sindicato de pesca deportiva.
-Siendo uno de los puertos más productivos de México, las pesquerías locales aprovechan una diversa gama de especies.
-Asociadas a Puerto Libertad, las tres zonas de reserva de Isla San Pedro Nolasco fueron establecidas en el 2012.
-Posteriormente, al igual que Puerto Libertad, las tres reservas de Isla San Pedro Nolasco fueron reconocidas legalmente como Zonas de Refugio Pesquero el 12 de Julio del 2017, con una vigencia de cinco años.
-A lo largo del proceso de establecimiento de reservas comunitarias o ZRP, existió la participación de miembros del Comité Nolasco (C.Técnica, Pesca Artesanal, Pesca Industrial, Pesca Deportiva, Turismo, Buceo Deportivo, Academia(CIAD), OSC (COBI)), y el grupo núcelo, conformado por CONANP, SEMARNAT, SEMAR, COFETUR, INAPESCA Guaymas, CONAPESCA, SAGARHPA, CEDES, H.
-ayuntamiento de Guaymas y PROFEPA.
-Las reservas fueron implementadas buscando proteger a especies como *Mycteroperca rosacea, Lutjanus argentiventris, Balistes polylepis, Octopus sp, Panulirus inflatus* y *Panulirus interruptus*.
-Las reservas no cuentan con vigilancia por parte de la comunidad.
-
-### Maria Elena
-
-Maria Elena es una ccomunidad pesquera en la costa de Quintana Roo.
-Los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona.
-El campo pesquero es utilizado por pescadores de la SCPP Cozumel scl (de la Isla de Cozumel).
-La principal especie aprovechada por ésta organización es la langosta espinoza del caribe (*Panulirus argus*).
-La cooperativa cuenta con permiso de pesca de escama y concesion de langosta.
-En el 2012, la Cooperativa, en conjunto con la Alianza Kanan Kay, COBI, CONANP, CONAPESCA, Oceanus, Fundación Claudia y Roberto Hernández, Fundación Haciendas del Mundo Maya, establecieron ocho ZRP con una vigencia de cinco años.
-La vigilancia de las reservas se realiza por medio del equipo de vigilancia comunitaria, con apoyo de la CONANP y una embarcación -donada por COBI- utilizada para realizar recorridos frecuentes.
-
-### Punta Herrero
-
-La comunidad de Punta Herrero se encuentra aproximadamente a 15 km al sur del campo pesquero de Maria Elena.
-De igual manera, los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona, y la principal especie explotada es la langosta.
-Sin embargo, la SCPP José María Azcorra también cuenta con permisos para pesca de escama y tiburon y una concesion de langosta.
-En una réplica del ejercicio realizado en Maria Elena -con presencia de los mismos actores-, cuatro ZRP fueron establecidas en el 2013, con una vigencia de cinco años.
-El equipo de vigilancia comunitaria, con apoyo de la CONANP, se encarga de la vigilancia de las reservas.
-
-```{r}
-read.csv("Data/governance_summary.csv") %>% 
-  knitr::kable(booktabs = T, col.names = c("Comunidad", "Reservas", "Año Implementación", "Año Reconocimiento", "Tipo", "Inspección y Vigilancia", "Manejo de recursos"), caption = "Características de gobernanza generales de las cinco comunidades estudiadas.") %>%
-  kable_styling(latex_options = c("scale_down")) %>%
-  column_spec(5, width = "3cm") %>%
-  column_spec(6, width = "6cm") %>%
-  column_spec(7, width = "6cm")
-```
-
-## Datos y análisis de datos
-
-```{r}
-fish <- read.csv("Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
-
-invert <- read.csv("Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
-
-load("Data/conapesca_2000-2015.RData")
-
-conapesca %<>%
-  filter(UnidadEconomica %in% c("Scpp cozumel scl",
-                                "Scpp jose maria azcorra scl",
-                                "Scpp buzos y pescadores de la baja california scl")) %>% 
-  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
-  group_by(UnidadEconomica, Ano, NombreComun) %>%
-  summarize(PesoDesembarcado = sum(PesoDesembarcado), PesoVivo = sum(PesoVivo)) %>%
-  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" &NombreComun == "Langosta ent. fca."))
-
-gob <- read.csv("Data/governance.csv", strip.white = T, stringsAsFactors = T)
-```
-
-Para evaluar las reservas, utilizamos tres fuentes de información.
-La información ecológica proviene de los monitoreos ecológicos realizados anualmente en las zonas reserva y control.
-Cada año, se realizan censos visuales para evaluar las comunidades de peces e invertebrados, registrando riquezas, abundancias y tallas (en peces).
-Esta información nos permite calcular los indicadores biológicos de manera anual.
-Al tener valores de diferentes indicadores biológicos antes y después de la implementación de las reservas, para las zonas de reserva y sitios control, tenemos un diseño muestreal de Antes-Después-Control-Impacto.
-Utilizando un análisis de diferencia en diferencias podemos estimar el efecto que la reserva tienen en los indicadores biológicos [@moland_2013-VP] con el uso de un modelo de regresión lineal múltiple:
-
-$$I = \beta_0 + \sum \gamma Año + \beta_1 Zona + \beta_2 Post\times Zona + \epsilon$$
-
-En este caso, modelamos los años como factores, tomando como referencia el primer año en la serie de datos de cada comunidad.
-Modelar los años como factores reduce la estructura del modelo, y relaja el ajuste al no asumir una tendencia lineal entre años; es decir, el cambio observado entre 2006 - 2007 no deberá de ser igual al observado entre el 2009 - 2010.
-Incluimos también un término para la zona, en la que la variable toma un valor de 0 si el sitio es una zona control y de 1 si es una zona de reserva.
-Finalmente, incluimos un término de interacción entre la variable de Zona y Post.
-La variable Post toma un valor de 0 para todos los años existentes antes de la implementación de la reserva, y un valor de 1 para años posteriores a la implementación.
-En este modelo, el coeficiente $\beta_2$ representa el efecto que la reservsa tuvo sobre un indicador a través del tiempo y con respecto a los sitios control.
-La siguiente figura presenta un resumen del esfuerzo muestreal anual para cada comunidad.
-
-```{r, fig.cap = "Esfuerzo muestreal anual para cada comunidad.
-El color indica la zona de reserva (rojo) o sitio control (azul).
-El tamaño del círculo indica el número de transectos totales realizados por año."}
-esfuerzo_peces <- filter(fish, Comunidad %in% c("Isla Natividad", "Puerto Libertad", "Guaymas", "Maria Elena", "Punta Herrero")) %>% 
-  group_by(Comunidad, Ano, Zona, Transecto) %>% 
-  count() %>% 
-  group_by(Comunidad, Ano, Zona) %>% 
-  summarize(Transectos = n()) %>% 
-  mutate(Class = "Peces")
-
-filter(invert, Comunidad %in% c("Isla Natividad", "Puerto Libertad", "Guaymas", "Maria Elena", "Punta Herrero")) %>% 
-  group_by(Comunidad, Ano, Zona, Transecto) %>% 
-  count() %>% 
-  group_by(Comunidad, Ano, Zona) %>% 
-  summarize(Transectos = n()) %>% 
-  mutate(Class = "Invertebrados") %>% 
-  rbind(esfuerzo_peces) %>% 
-  ggplot(aes(y = Ano, x = Comunidad, color = Zona)) +
-  geom_line(position=position_dodge(1)) +
-  geom_point(aes(size = Transectos), position=position_dodge(1)) +
-  theme_bw() +
-  scale_color_brewer(palette = "Set1", direction = -1) +
-  labs(x = "Año", y = "Comunidad - tratamiento") +
-  facet_wrap(~Class, ncol = 1) +
-  theme(strip.background = element_rect(fill = "transparent")) +
-  coord_flip()
-```
-
-También incluimos información socioeconómica relevanteproveniente de los avisos de arribo de CONAPESCA.
-En este caso, se tienen registros mensuales de los recursos aprovechados por las diferentes comunidades, en los que se reportan los arribos (Toneladas) y el valor de los arribos ($).
-La información se encuentra disponible para el periodo 2001 - 2014.
-Ya que las categorías registradas por CONAPESCA son amplias y existe un nivel de error, utilizamos únicamente los arribos reportados para langosta entera fresca a nivel de cooperativa.
-Los ingresos generados por arribos son ajustados por medio del índice de precio al consumidor.
-El análisis de datos socioeconómicos se aplicó únicamente a Isla Natividad, Maria Elena y Punta Herrero, siguiendo un modelo con la forma:
-
-$$I = \beta_0 + \beta_1Post$$
-
-Que nos permite comparar el cambio en el promedio de los indicadores antes (Post = 0) y después (Post = 1) de la implementación de la reserva.
-Tanto para los indicadores biológicos como los socioeconómicos, los coeficientes fueron ajustador con el estimador de muestras heterocedásticas.
-
-La información de governanza fue obtenida a nivel de comunidad, pidiendo a personas familiares con las comunidades que proveyeran la información necesaria.
-
-Dada la similitud de objetivos entre las reservas, la evaluación se realiza con los mismos indicadores.
-En este caso, se utilizan 8 indicadores biológicos, 2 socioeconómicos y 14 de gobernanza (Tabla 2).
-Según la disponibilidad de datos, se calculó la densidad de las especies objetivo presentadas en la sección de descripción de las comunidades.
-El criterio de selección fue que cada especie debía tener, por lo menos, dos observaciones anuales para las zonas de reserva y control.
-
-```{r}
-data(IndListG)
-
-data.frame(Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Densidad de especies objetivo", "Ingresos por especies objetivo", "Arribos de especies objetivo", IndListG$Indicators), stringsAsFactors = F) %>%
-  knitr::kable(caption = "Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.", booktabs = T) %>%
-  kable_styling() %>%
-  group_rows("Biológicos", 1, 6) %>%
-  group_rows("Socioeconómicos", 7, 8) %>% 
-  group_rows("Gobernanza", 9, 22)
-```
-
-```{r}
-source(file = "analisis.R")
-source(file = "modelo.R")
-source(file = "robust_se.R")
-source(file = "extract_effect.R")
-
-my_ggplot <- function(x, y.label, breaks){
-
-  x$model %>% 
-    mutate(Ano = as.numeric(as.character(Ano))) %>% 
-    ggplot(aes(x = Ano, y = Indicador, color = Zona, fill = Zona)) +
-    geom_jitter(height = 0, width = 0.1, size = 1, alpha = 0.5) +
-    stat_summary(geom = "ribbon", fun.data = "mean_se", alpha = 0.5)+
-    stat_summary(geom = "point", fun.data = "mean_se", color = "black")+
-    stat_summary(geom = "line", fun.data = "mean_se", color = "black") +
-    theme_bw() +
-    theme(legend.position = "None") +
-    scale_color_brewer(palette = "Set1", direction = -1) +
-    scale_fill_brewer(palette = "Set1", direction = -1) +
-    labs(x = "Año", y = y.label) +
-    scale_x_continuous(breaks = breaks) +
-    theme(text = element_text(size = 8))
-}
-
-my_plot <- function(x, breaks) {
-
-  p1 <- my_ggplot(x$Modelo[[1]], y.label = "Peces\nShannon", breaks)
-  p2 <- my_ggplot(x$Modelo[[2]], y.label = "Peces\nRiqueza (sp / transecto)", breaks)
-  p3 <- my_ggplot(x$Modelo[[3]], y.label = "Peces\nDensidad (org / transecto)", breaks)
-  p4 <- my_ggplot(x$Modelo[[4]], y.label = "Peces\nNivel trófico", breaks)
-  p5 <- my_ggplot(x$Modelo[[5]], y.label = "Peces\nBiomasa (Kg / transecto)", breaks)
-  p6 <- my_ggplot(x$Modelo[[6]], y.label = "Invert.\nShannon", breaks)
-  p7 <- my_ggplot(x$Modelo[[7]], y.label = "Invert.\nRiqueza (sp / transecto)", breaks)
-  p8 <- my_ggplot(x$Modelo[[8]], y.label = "Invert.\nDensidad (org / transecto)", breaks)
-
-  gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8)
-
-}
-
-```
-
-```{r}
-
-MariaElena <- analisis(peces = fish,
-                       invert = invert,
-                       comunidad = "Maria Elena",
-                       ano.imp = 2012)
-
-PuntaHerrero <- analisis(peces = fish,
-                         invert = invert,
-                         comunidad = "Punta Herrero",
-                         ano.imp = 2013)
-
-IslaNatividad <- analisis(peces = fish,
-                          invert = invert,
-                          comunidad = "Isla Natividad",
-                          ano.imp = 2006,
-                          obj.spp = c("Haliotis fulgens", "Haliotis corrugata", "Panulirus interruptus"))
-
-PuertoLibertad <- analisis(peces = fish,
-                           invert = invert,
-                           comunidad = "Puerto Libertad",
-                           ano.imp = 2012,
-                           obj.spp = c("Balistes polylepis","Lutjanus argentiventris", "Mycteroperca rosacea", "Spondylus limbatus", "Hexaplex nigritus", "Octopus bimaculatus"))
-
-Guaymas <- analisis(peces = fish,
-                    invert = invert,
-                    comunidad = "Guaymas",
-                    ano.imp = 2012,
-                    obj.spp = c("Mycteroperca rosacea", "Lutjanus argentiventris", "Balistes polylepis"))
-
-```
-
-\clearpage
-
-# Resultados
-
-A continuación se presentan los resultados de cada una de las comunidades.
-Los resultados biológicos se presentarán para cada comunidad, discutiendo primero los indicadores en común con otras comunidades (Shannon, Riqueza, Densidad, Nivel Trófico, Biomasa para peces e invertebrados) y, según su caso, las densidades de las especies objetivo.
-
-## Isla Natividad
-
-Para las dos reservas de Isla Natividad, todos los indicadores tienen un comportamiento similar a lo largo del tiempo.
-Sin embargo, hay indicios del efecto de la reserva.
-Por ejemplo, el Índice de diversidad de Shannon para peces muestra una mejora importante.
-Al inicio de la serie de tiempo, el indicador tenía un valor menor para la reserva que para el control.
-Sin embargo, la diferencia entre reserva y control se ha reducido con el paso del tiempo, indicando un incremento en el indicador para la zona de reserva.
-Un patrón similar, pero menos evidente, se observa para la riqueza de peces.
-Sin embargo, no parece existir un efecto de la reserva en la densidad de peces; los promedios anuales de las reservas y sitios control son similares, y no var[ian con el tiempo.
-Para el caso del nivel trófico de peces, el efecto es menor y mucho más lento que para el Índice de Shannon o la riqueza.
-En este caso, vemos un comportamiento similar para ambos tratamientos hasta el 2012, donde la reserva comienza a tener una mayor dispersión (rango más amplio) en los niveles tróficos.
-Finalmente, para el 2016, se observan diferencias evidentes en el nivel trófico medio.
-Para los indicadores de invertebrados, todos muestran (en diferente magnitud) un decremento, posiblemente explicado por los eventos de hypoxia reportados en la zona [@micheli_2012-EU].
-Sin embargo, es evidente que la reservas producen un efecto de amortiguamiento ante estos cambios ambientales, pues la reducción en las reservas es menor que en los sitios control.
-
-```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Isla Natividad.
-El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
-Las líneas y puntos negros indican el promedio anual para cada zona.
-Las barras de error indican 1 Error Estándar con respecto del promedio."}
-my_plot(IslaNatividad, breaks = c(2006, 2009, 2012, 2015))
-```
-
-```{r, fig.width = 4, fig.height = 6, fig.cap = "Series de tiempo densidad de especies objetivo para Isla Natividad.
-El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
-Las líneas y puntos negros indican el promedio anual para cada zona.
-Las barras de error indican 1 Error Estándar con respecto del promedio."}
-gridExtra::grid.arrange(my_ggplot(IslaNatividad$Modelo[[9]], y.label = "H. fulgens / transecto", breaks = c(2006, 2009, 2012, 2015)),
-                        my_ggplot(IslaNatividad$Modelo[[10]], y.label = "H. corrugata / transecto", breaks = c(2006, 2009, 2012, 2015)),
-                        my_ggplot(IslaNatividad$Modelo[[10]], y.label = "P. interruptus / transecto", breaks = c(2006, 2009, 2012, 2015)))
-```
-
-\clearpage
-
-## Puerto Libertad
-
-```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Puerto Libertad.
-El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
-Las líneas y puntos negros indican el promedio anual para cada zona.
-Las barras de error indican 1 Error Estándar con respecto del promedio."}
-my_plot(PuertoLibertad, breaks = c(2012, 2013, 2014, 2015, 2016))
-```
-
-```{r, fig.width = 4, fig.height = 4, fig.cap = "Series de tiempo densidad de especies objetivo para Puerto Libertad.
-El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
-Las líneas y puntos negros indican el promedio anual para cada zona.
-Las barras de error indican 1 Error Estándar con respecto del promedio."}
-
-gridExtra::grid.arrange(my_ggplot(PuertoLibertad$Modelo[[9]], y.label = "B. polylepis / transecto", breaks = c(2012, 2013, 2014, 2015, 2016)),
-                        my_ggplot(PuertoLibertad$Modelo[[12]], y.label = "S. limbatus / transecto", breaks = c(2012, 2013, 2014, 2015, 2016)))
-```
-
-\clearpage
-
-## Guaymas
-
-```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Guaymas.
-El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
-Las líneas y puntos negros indican el promedio anual para cada zona.
-Las barras de error indican 1 Error Estándar con respecto del promedio."}
-my_plot(Guaymas, breaks = c(2011, 2012, 2013, 2014))
-```
-
-```{r, fig.width = 4, fig.height = 6, fig.cap = "Series de tiempo densidad de especies objetivo para Guaymas.
-El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
-Las líneas y puntos negros indican el promedio anual para cada zona.
-Las barras de error indican 1 Error Estándar con respecto del promedio."}
-gridExtra::grid.arrange(my_ggplot(Guaymas$Modelo[[9]], y.label = "M. rosacea / transecto", breaks = c(2011, 2012, 2013, 2014)),
-                        my_ggplot(Guaymas$Modelo[[10]], y.label = "L. argentiventris / transecto", breaks = c(2011, 2012, 2013, 2014)),
-                        my_ggplot(Guaymas$Modelo[[11]], y.label = "B. polylepis / transecto", breaks = c(2011, 2012, 2013, 2014)))
-```
-
-\clearpage
-
-## Maria Elena
-
-```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Maria Elena.
-El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
-Las líneas y puntos negros indican el promedio anual para cada zona.
-Las barras de error indican 1 Error Estándar con respecto del promedio."}
-
-my_plot(MariaElena, breaks = c(2012, 2013, 2014, 2015, 2016))
-```
-
-\clearpage
-
-## Punta Herrero
-
-```{r, fig.cap = "Series de tiempo de los indicadores biológicos para Punta Herrero.
-El rojo indica zonas de reserva, el azul zonas control.
-Cada punto representa un transecto.
-Las líneas y puntos negros indican el promedio anual para cada zona.
-Las barras de error indican 1 Error Estándar con respecto del promedio."}
-
-my_plot(PuntaHerrero, breaks = c(2012, 2013, 2014, 2015, 2016))
-```
-
-\clearpage
-
-## Efectos generales
-
-```{r, message = F, warning = F}
-rbind(extract_effect(IslaNatividad),
-      extract_effect(PuertoLibertad),
-      extract_effect(Guaymas),
-      extract_effect(MariaElena),
-      extract_effect(PuntaHerrero)) %>% 
-  mutate(Comunidad = c(rep("IN", 11),
-                       rep("PL", 14),
-                       rep("G", 11),
-                       rep("ME", 8),
-                       rep("PH", 8)),
-         P = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
-  filter(Indicador %in% c("ShannonP", "RiquezaP", "DensidadP", "TroficoP", "BiomasaP", "ShannonI", "RiquezaI", "DensidadI")) %>% 
-  ggplot(aes(x = Comunidad, y = estimate)) +
-  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0) +
-  geom_point(size = 3, aes(color = P)) +
-  coord_flip() +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  theme_bw() +
-  facet_wrap(~Indicador, scales = "free") +
-  theme(strip.background = element_rect(fill = "transparent")) +
-  labs(x = "Efecto") +
-  scale_color_brewer(palette = "Set1")
-```
-
-## Socioeconómicos
-
-```{r}
-ggplot(conapesca, aes(x = Ano, y = PesoVivo/1000, color = UnidadEconomica)) +
-  geom_smooth(method = "lm") +
-  geom_point() +
-  theme_bw() +
-  scale_color_brewer(palette = "Set1") +
-  geom_vline(xintercept = 2006, color = "red", linetype = "dashed", size = 1) +
-  geom_vline(xintercept = 2012, color = "blue", linetype = "dashed", size = 1) +
-  geom_vline(xintercept = 2013, color = "darkgreen", linetype = "dashed", size = 1) +
-  labs(x = "Año", y = "Peso (Toneladas)")
-```
-
-```{r}
-anos <- data.frame(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl", "Scpp cozumel scl", "Scpp jose maria azcorra scl"), AnoImp = c(2006, 2012, 2013), stringsAsFactors = F)
-
-conapesca %>% 
-  ungroup() %>% 
-  mutate(UnidadEconomica = as.character(UnidadEconomica)) %>% 
-  left_join(anos, by = "UnidadEconomica") %>% 
-  mutate(Post = ifelse(Ano <= AnoImp, 0, 1)) %>% 
-  split(.$UnidadEconomica) %>% 
-  map(~ lm(PesoVivo/1000 ~ Post, data = .x)) %>% 
-  map_df(tidy) %>% 
-  mutate(model = c("Isla Natividad", "Isla Natividad", "Maria Elena", "Maria Elena", "Punta Herrero", "Punta Herrero"),
-         P1 = ifelse(p.value < 0.1, "*", ""),
-         P2 = ifelse(p.value < 0.05, "*", ""),
-         P = paste0(P1, P2),
-         estimate = formatC(estimate, digits = 2, format = "f"),
-         std.error = formatC(std.error, digits = 2, format = "f"),
-         texto = paste0(estimate, "(", std.error, ")", P)) %>% 
-  select(model, term, texto) %>% 
-  spread(model, texto) %>% 
-  knitr::kable(booktabs = T)
-
-```
-
-```{r, warning = F, message = F, fig.height = 6, fig.width = 6}
-gob_summary <- group_by(gob, Comunidad) %>% 
-  summarize(Ano = mean(Ano),
-            PescadoresAntes = mean(Num.pescadores.antes),
-            PescadoresAhora = mean(Num.pescadores.actual),
-            PescaIlegal = mean(Grado.pesca.ilegal)) %>% 
-  mutate(CambioPescadores = PescadoresAntes-PescadoresAhora,
-         Edad = 2017 - Ano) %>% 
-  select(Comunidad, PescaIlegal, Edad)
-
-rbind(extract_effect(IslaNatividad),
-      extract_effect(PuertoLibertad),
-      extract_effect(Guaymas),
-      extract_effect(MariaElena),
-      extract_effect(PuntaHerrero)) %>% 
-  mutate(Comunidad = c(rep("IN", 11),
-                       rep("PL", 14),
-                       rep("G", 11),
-                       rep("ME", 8),
-                       rep("PH", 8)),
-         P = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
-  filter(Indicador %in% c("ShannonP", "DensidadP", "TroficoP", "ShannonI", "DensidadI")) %>%
-  select(Comunidad, Indicador, estimate, P) %>% 
-  left_join(gob_summary, by = "Comunidad") %>% 
-  gather(Variable, Valor, -c(Comunidad, Indicador, estimate, P)) %>% 
-  ggplot(aes(x = Valor, y = estimate)) +
-  geom_smooth(method = "lm", color = "black") +
-  geom_point(aes(color = Comunidad)) +
-  facet_grid(Indicador~Variable, scales = "free") +
-  theme_bw() +
-  labs(x = "Valor", y = "Efecto")
-
-```
-
-# Conclusiones
-
-# Referencias
+# References

### 2018-01-11 Frontiers_LaTex_Templates/Villasenor-Derbez.tex
Added latex template for manuscript skeleton

\documentclass{frontiersSCNS}
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

\usepackage[utf8]{inputenc}

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villasenor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez,\(^{1,*}\), Stuart Fulton,\(^{2}\) Jorge
Torre,\(^{2}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Mexico}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Me, at home}

\def\corrEmail{\href{mailto:jvillasenor@bren.ucsb.edu}{\nolinkurl{jvillasenor@bren.ucsb.edu}}}

\begin{document}
\onecolumn
\firstpage{1}

\title[Mexican marine reserves]{Three archetypes of no-take marine reserves yield positive (but
idiosyncratic) social and ecological benefits in Mexico}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}

\maketitle

\section{Introduction}\label{introduction}

La sobrepesca y prácticas pesqueras no sostenibles son unas de las
mayores amenazas para la conservación de los ecosistemas marinos del
mundo \citep{halpern_2008-dK} (Halpern et al., 2017). La implementación
de reservas marinas (\emph{i.e.} áreas donde la captura de una o más
especies está prohibida) es una medida de manejo frecuentemente
propuesta para recuperar stocks pesqueros e impulsar la productividad
pesquera en aguas cercanas (Afflerbach et al., 2014; Krueck et al.,
2017; Sala and Giakoumi, 2017). Recientes trabajos han demostrado que
también pueden mitigar y proveer amortiguamiento ante el cambio
climático (Roberts et al., 2017), variabilidad ambiental (Micheli et
al., 2012), resolver problemas de pesca incidental (Hastings et al.,
2017) y, en general, incrementar la biomasa, riqueza y densidades de
organismos dentro de sus fronteras (Giakoumi et al., 2017; Lester et
al., 2009; Sala and Giakoumi, 2017).

En México, las reservas marinas han sido comúnmente establecidas como
zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por
la Comisión Nacional de Áreas Naturales Protegidas (CONANP). Al día de
hoy, 36 RBs protegen una porción del ambiente marino en México. Sin
embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde
las actividades pesqueras están prohibidas. Aunque la CONANP ha hecho
esfuerzos importantes por involucrar a los actores durante la
implementación de las reservas, esto aún se caracteriza por un proceso
descendente, el cual conlleva a la falta de cumplimiento por parte de
los actores. La escasez de recursos monetarios y humanos de la limitan
también el monitoreo y vigilancia de las reservas, y a su vez, el
desempeño de la reserva.

Buscando promover una alternativa con procesos ascendentes para
implementar reservas marinas, las Organizaciones de la Sociedad Civil
(OSCs) comenzaron a trabajar con comunidades pesqueras para establecer
reservas comunitarias (Uribe et al., 2010) . Estas son comúnmente
establecidas dentro de zonas de concesión, una forma de derechos de uso
territoriales para pesquerías (TURF, en inglés). Al permitir a los
pescadores diseñar sus propias reservas, una mayor proporción de la
comunidad está de acuerdo con los perímetros y reglas establecidas, y
por lo tanto los respetan (Beger et al., 2004; Espinosa-Romero et al.,
2014; Gelcich and Donlan, 2015) . Adicionalmente, los pescadores pueden
implementar sus reservas por un periodo acordado (usualmente cinco
años), después del cual la reserva puede ser abierta a la pesca. Esto
provee a los pescadores con un sentido de confianza de que, en caso de
ser necesario, aún tienen acceso a pescar esa zona{[}\^{}1{]}. Las
reservas son directamente vigiladas y monitoreadas por la comunidad,
quienes comúnmente utilizan pequeñas embarcaciones (\emph{e.g.} pangas)
para patrullar la zona, o realizan avistamientos desde la costa en
búsqueda de pescadores ilegales Aún así, las reservas comunitarias
carecen de reconocimiento legal; por lo tanto, no hay forma de penalizar
a los infractores.

\section{Materials and Methods}\label{materials-and-methods}

\section{Results}\label{results}

\section{Discussion}\label{discussion}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-afflerbach_2014-HP}{}
Afflerbach, J. C., Lester, S. E., Dougherty, D. T., and Poon, S. E.
(2014). A global survey of -reserves, territorial use rights for
fisheries coupled with marine reserves. \emph{Global Ecology and
Conservation} 2, 97--106.
doi:\href{https://doi.org/10.1016/j.gecco.2014.08.001}{10.1016/j.gecco.2014.08.001}.

\hypertarget{ref-beger_2004-Y8}{}
Beger, M., Harborne, A. R., Dacles, T. P., Solandt, J.-L., and Ledesma,
G. L. (2004). A framework of lessons learned from community-based marine
reserves and its effectiveness in guiding a new coastal management
initiative in the philippines. \emph{Environ Manage} 34, 786--801.
doi:\href{https://doi.org/10.1007/s00267-004-0149-z}{10.1007/s00267-004-0149-z}.

\hypertarget{ref-espinosaromero_2014-PY}{}
Espinosa-Romero, M. J., Rodriguez, L. F., Weaver, A. H.,
Villanueva-Aznar, C., and Torre, J. (2014). The changing role of ngos in
mexican small-scale fisheries: From environmental conservation to
multi-scale governance. \emph{Marine Policy} 50, 290--299.
doi:\href{https://doi.org/10.1016/j.marpol.2014.07.005}{10.1016/j.marpol.2014.07.005}.

\hypertarget{ref-gelcich_2015-Gw}{}
Gelcich, S., and Donlan, C. J. (2015). Incentivizing biodiversity
conservation in artisanal fishing communities through territorial user
rights and business model innovation. \emph{Conserv Biol} 29,
1076--1085.
doi:\href{https://doi.org/10.1111/cobi.12477}{10.1111/cobi.12477}.

\hypertarget{ref-giakoumi_2017-V2}{}
Giakoumi, S., Scianna, C., Plass-Johnson, J., Micheli, F.,
Grorud-Colvert, K., Thiriet, P., et al. (2017). Ecological effects of
full and partial protection in the crowded mediterranean sea: A regional
meta-analysis. \emph{Sci Rep} 7, 8940.
doi:\href{https://doi.org/10.1038/s41598-017-08850-w}{10.1038/s41598-017-08850-w}.

\hypertarget{ref-halpern_2017-Zi}{}
Halpern, B. S., Frazier, M., Afflerbach, J., O'Hara, C., Katona, S.,
Stewart Lowndes, J. S., et al. (2017). Drivers and implications of
change in global ocean health over the past five years. \emph{PLoS ONE}
12, e0178267.
doi:\href{https://doi.org/10.1371/journal.pone.0178267}{10.1371/journal.pone.0178267}.

\hypertarget{ref-hastings_2017-sm}{}
Hastings, A., Gaines, S. D., and Costello, C. (2017). Marine reserves
solve an important bycatch problem in fisheries. \emph{Proc Natl Acad
Sci U S A}.
doi:\href{https://doi.org/10.1073/pnas.1705169114}{10.1073/pnas.1705169114}.

\hypertarget{ref-krueck_2017-J1}{}
Krueck, N. C., Ahmadia, G. N., Possingham, H. P., Riginos, C., Treml, E.
A., and Mumby, P. J. (2017). Marine reserve targets to sustain and
rebuild unregulated fisheries. \emph{PLoS Biol} 15, e2000537.
doi:\href{https://doi.org/10.1371/journal.pbio.2000537}{10.1371/journal.pbio.2000537}.

\hypertarget{ref-lester_2009-Ks}{}
Lester, S., Halpern, B., Grorud-Colvert, K., Lubchenco, J., Ruttenberg,
B., Gaines, S., et al. (2009). Biological effects within no-take marine
reserves: A global synthesis. \emph{Mar. Ecol. Prog. Ser.} 384, 33--46.
doi:\href{https://doi.org/10.3354/meps08029}{10.3354/meps08029}.

\hypertarget{ref-micheli_2012-EU}{}
Micheli, F., Saenz-Arroyo, A., Greenley, A., Vazquez, L., Espinoza
Montes, J. A., Rossetto, M., et al. (2012). Evidence that marine
reserves enhance resilience to climatic impacts. \emph{PLoS ONE} 7,
e40832.
doi:\href{https://doi.org/10.1371/journal.pone.0040832}{10.1371/journal.pone.0040832}.

\hypertarget{ref-roberts_2017-J9}{}
Roberts, C. M., O'Leary, B. C., McCauley, D. J., Cury, P. M., Duarte, C.
M., Lubchenco, J., et al. (2017). Marine reserves can mitigate and
promote adaptation to climate change. \emph{Proc Natl Acad Sci U S A}
114, 6167--6175.
doi:\href{https://doi.org/10.1073/pnas.1701262114}{10.1073/pnas.1701262114}.

\hypertarget{ref-sala_2017-69}{}
Sala, E., and Giakoumi, S. (2017). No-take marine reserves are the most
effective protected areas in the ocean. \emph{ICES Journal of Marine
Science}.
doi:\href{https://doi.org/10.1093/icesjms/fsx059}{10.1093/icesjms/fsx059}.

\hypertarget{ref-uribe_2010-u2}{}
Uribe, P., Moguel, S., Torre, J., Bourillon, L., and Saenz, A. (2010).
\emph{Implementación de reservas marinas en méxico}. 1st ed. Mexico.

\end{document}

### 2018-01-11 Manuscript.Rmd
Added latex template for manuscript skeleton

@@
 title: "Efectividad de reservas marinas en México"
-author: "Juan Carlos Villaseñor-Derbez, Stuart Fulton, Jorge Torre"
+author: "Juan Carlos Villaseñor-Derbez, Stuart Fulton, Alvin Suárez-Castillo"
 header-includes:

### 2018-01-11 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Finalized LaTeX - Rmd integration, all works now. Let´s start running analyses

---
title: "Three archetypes of no-take marine reserves in Mexico yield positive (but idiosyncratic) social and ecological benefits"
rtitle: "Mexican marine reserves"
author: "Juan Carlos Villaseñor-Derbez,$^{1,*}$, Stuart Fulton,$^{2}$ Jorge Torre,$^{2}$"
institution: "
$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico
"
corrauth: "Me, at home"
email: "jvillasenor@bren.ucsb.edu"
abstract: "This is my abstract"
keywords: "Marine Reserves, Marine Conservation, Fisheries, Citizen Science, Mexico"
# biblio: test.bib
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
# csl: FMS.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

Succinct, with no subheadings.

Rationale
Objectives
Research question

La sobrepesca y prácticas pesqueras no sostenibles son unas de las mayores amenazas para la conservación de los ecosistemas marinos del mundo \citep{halpern_2008-dK,halpern_2017-Zi}. La implementación de reservas marinas (*i.e.* áreas donde la captura de una o más especies está prohibida) es una medida de manejo frecuentemente propuesta para recuperar stocks pesqueros e impulsar la productividad pesquera en aguas cercanas \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}. Recientes trabajos han demostrado que también pueden mitigar y proveer amortiguamiento ante el cambio climático \citep{roberts_2017-J9}, variabilidad ambiental \citep{micheli_2012-EU}, resolver problemas de pesca incidental \citep{hastings_2017-sm} y, en general, incrementar la biomasa, riqueza y densidades de organismos dentro de sus fronteras \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}.

En México, las reservas marinas han sido comúnmente establecidas como zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por la Comisión Nacional de Áreas Naturales Protegidas (CONANP). Al día de hoy, 36 RBs protegen una porción del ambiente marino en México. Sin embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas. Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, esto aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores. La escasez de recursos monetarios y humanos de la limitan también el monitoreo y vigilancia de las reservas, y a su vez, el desempeño de la reserva.

Buscando promover una alternativa con procesos ascendentes para implementar reservas marinas, las Organizaciones de la Sociedad Civil (OSCs) comenzaron a trabajar con comunidades pesqueras para establecer reservas comunitarias \citep{uribe_2010-u2} . Estas son comúnmente establecidas dentro de zonas de concesión, una forma de derechos de uso territoriales para pesquerías (TURF, en inglés). Al permitir a los pescadores diseñar sus propias reservas, una mayor proporción de la comunidad está de acuerdo con los perímetros y reglas establecidas, y por lo tanto los respetan \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8} . Adicionalmente, los pescadores pueden implementar sus reservas por un periodo acordado (usualmente cinco años), después del cual la reserva puede ser abierta a la pesca. Esto provee a los pescadores con un sentido de confianza de que, en caso de ser necesario, aún tienen acceso a pescar esa zona. Las reservas son directamente vigiladas y monitoreadas por la comunidad, quienes comúnmente utilizan pequeñas embarcaciones (*e.g.* pangas) para patrullar la zona, o realizan avistamientos desde la costa en búsqueda de pescadores ilegales Aún así, las reservas comunitarias carecen de reconocimiento legal; por lo tanto, no hay forma de penalizar a los infractores.

Sin embargo, en el 2014 una nueva norma \citep{nom049sagpesc_2014-V6} permite a los pescadores solicitar el establecimiento de reservas marinas bajo el nombre de “Zonas de refugio Pesquero” (ZRP). El manejo de las ZRP combina procesos ascendentes y descendentes al reconocer legalmente las reservas propuestas por las comunidades. Posterior a la revisión por parte de la Comisión Nacional de Acuacultura y Pesca (CONAPESCA) y la opinión técnica del Instituto Nacional de Acuacultura y Pesca (INAPESCA) las ZRP son establecidas por el periodo solicitado por los pescadores. El monitoreo y la vigilancia de las ZRP es típicamente llevado a cabo por la comunidad , con ayuda de OSCs locales. Hasta este cambio regulatorio, las reservas comunitarias no contaban con el soporte legal, y eran solamente reconocidas por la comunidad. Al día de hoy, existen 39 ZRP establecidas en el Pacífico, Golfo de California y Caribe Mexicano.

Aunque existen tres aproximaciones generales para implementar reservas marinas en México (*i.e.* Zonas núcleo dentro de AMP, reservas comunitarias y Zonas de Refugio Pesquero), aún no comprendemos a fondo las características sociales que permiten su efectividad. La ciencia de reservas marinas se ha enfocado ampliamente en los efectos biológicos que estas tienen \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,afflerbach_2014-HP,krueck_2017-J1}. Aunque el aspecto ecológico de las reservas es importante para su éxito, su efectividad también depende del estado socioeconómico y los sistemas de gobernanza de las comunidades pesqueras.

La literatura indica que diferentes características influyen en el éxito de una reserva. En Palau, por ejemplo, la edad (*i.e.* tiempo transcurrido desde implementación), tamaño y hábitat contenido son características claves que determinan la efectividad \citep{friedlander_2017-oI}. Por otro lado, en el Mar Mediterráneo, @difranco_2016-Xw identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks y el ingreso económicos a los pescadores, a la vez que se presenta una mayor aceptación social de las prácticas de manejo. En una aproximación global, @edgar_2014-UO encuentran que la procuración, edad, tamaño y aislamiento son determinantes de la efectividad de las reservas. Por lo tanto, observamos que las características que habilitan el éxito varían a través de regiones, y poco esfuerzo se ha hecho por comprender estas interacciones en México.

# Materials and Methods

This section may be divided by subheadings. This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated. For experiments reporting results on animal or human subject research, an ethics approval statement should be included in this section (for further information, see section Materials and Data Policies)

## Study area

```{r}
suppressPackageStartupMessages({
  library(sp)
  library(tmap)
  library(tmaptools)
  library(MPAtools)
  library(magrittr)
  library(kableExtra)
  library(broom)
  library(tidyverse)
})
```

```{r, message = F, warning = F, fig.cap = "Mapa de la localización general de las comunidades de estudio. El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero.", fig.pos = 'h'}

####

proj <- CRS("+proj=longlat +datum=NAD27")

data(World)
W <- spTransform(World, proj)
W <- W[!W$name == "Mexico",]

load(file = "../Data/coastline_mx.Rda")

locations <- cbind(x_coord = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087), y_coord = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957)) %>% 
  SpatialPoints(proj4string = proj)

location_labels <- data.frame(X = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087),
                              Y = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957),
                              Location = c("Isla Natividad", "Maria Elena", "Punta Herrero", "Puerto Libertad", "Guaymas"),
                              Region = c("Pacific", "Caribbean","Caribbean", "Gulf of California", "Gulf of California"))
coordinates(location_labels) <- c("X", "Y")
proj4string(location_labels) <- proj
location_labels <- spTransform(location_labels, proj)

SAM_polygon <- cbind(x_coord = c(-88, -87, -87, -88, -88), y_coord = c(19, 19, 20, 20, 20)) %>% 
  Polygon() %>% 
  list() %>% 
  Polygons(srl = ., ID = 1) %>% 
  list() %>% 
  SpatialPolygons(proj4string = proj)

mex <- tm_shape(W, bbox = bb(baja2)) +
  tm_polygons() +
  tm_shape(baja2) +
  tm_polygons() +
  # tm_shape(SAM_polygon) +
  # tm_borders(col = "black", lwd = 2) +
  tm_shape(location_labels) +
  tm_symbols(shape = "Location", col = "Region", palette = c("#33A02C", "#1F78B4", "#E31A1C"), border.col = "black", size = 0.5) +
  tm_compass(position = c("left", "bottom")) +
  tm_scale_bar(position = c("left", "bottom")) +
  tm_grid(n.x = 4, n.y = 5)

sam <- tm_shape(baja2, bbox = bb(xlim = c(-88, -86.726716),
                          ylim = c(19.25, 19.5))) +
  tm_polygons() +
  tm_shape(location_labels) +
  tm_symbols(shape = "Location", col = "Region", palette = c("#33A02C", "#1F78B4", "#E31A1C"), border.col = "black", size = 0.5) +
  tm_compass() +
  tm_scale_bar() +
  tm_legend(show=FALSE) +
  tm_grid(n.x = 4, n.y = 5)

tmap_arrange(mex, sam, ncol = 2)
```

## Data collection

## Data analysis

### Biological

### Socioeconomic

### Governance

# Results

# Discussion

This section may be divided by subheadings. Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.

Summary of main findings
Limitations
Conclusions

\section*{Conflict of Interest Statement}

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}

JC analyzed and interpreted data, discussed the results and wrote the manuscrip. SF and JT edited the manuscript and discussed the results.

\section*{Funding}
Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.

\section*{Acknowledgments}
This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.

\section*{Supplemental Data}
 \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder

\bibliographystyle{frontiersinSCNS_ENG_HUMS}
\bibliography{references}

\section*{Figure captions}

### 2018-01-11 Frontiers_Manuscript/Villasenor-Derbez.tex
Finalized LaTeX - Rmd integration, all works now. Let´s start running analyses

--- 2018-01-11 Frontiers_Manuscript/Villasenor-Derbez.Rmd
@@
----
-title: "Three archetypes of no-take marine reserves in Mexico yield positive (but idiosyncratic) social and ecological benefits"
-rtitle: "Mexican marine reserves"
-author: "Juan Carlos Villaseñor-Derbez,$^{1,*}$, Stuart Fulton,$^{2}$ Jorge Torre,$^{2}$"
-institution: "
-$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
-$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico
-"
-corrauth: "Me, at home"
-email: "jvillasenor@bren.ucsb.edu"
-abstract: "This is my abstract"
-keywords: "Marine Reserves, Marine Conservation, Fisheries, Citizen Science, Mexico"
-# biblio: test.bib
-output:
-  bookdown::pdf_document2:
-    keep_tex: yes
-    template: template.tex
-# csl: FMS.csl
----
+\documentclass{frontiersSCNS}
+\usepackage{url,hyperref,lineno,microtype,subcaption}
+\usepackage[onehalfspacing]{setspace}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-```
+\linenumbers
 
-# Introduction
+\usepackage[utf8]{inputenc}
+
+\def\keyFont{\fontsize{8}{11}\helveticabold }
+\def\firstAuthorLast{Villasenor-Derbez {et~al.}}
+\def\Authors{Juan Carlos Villaseñor-Derbez,\(^{1,*}\), Stuart Fulton,\(^{2}\) Jorge
+Torre,\(^{2}\)}
+% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
+% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
+\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
+of California, Santa Barbara, Santa Barbara, CA,
+USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Mexico}
+% The Corresponding Author should be marked with an asterisk
+% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
+\def\corrAuthor{Me, at home}
+
+\def\corrEmail{\href{mailto:jvillasenor@bren.ucsb.edu}{\nolinkurl{jvillasenor@bren.ucsb.edu}}}
+
+\begin{document}
+\onecolumn
+\firstpage{1}
+
+\title[Mexican marine reserves]{Three archetypes of no-take marine reserves in Mexico yield positive
+(but idiosyncratic) social and ecological benefits}
+
+\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
+\address{} %This field will be automatically populated
+\correspondance{} %This field will be automatically populated
+
+\extraAuth{}
+
+\maketitle
+
+\begin{abstract}
+
+This is my abstract
+
+\medskip
+\tiny
+ \keyFont{ \section{Keywords:} Marine Reserves, Marine Conservation, Fisheries, Citizen Science, Mexico}
+
+\end{abstract}
+
+\section{Introduction}\label{introduction}
 
@@
 
-Rationale
-Objectives
-Research question
+Rationale Objectives Research question
 
-La sobrepesca y prácticas pesqueras no sostenibles son unas de las mayores amenazas para la conservación de los ecosistemas marinos del mundo \citep{halpern_2008-dK,halpern_2017-Zi}.
-La implementación de reservas marinas (*i.e.* áreas donde la captura de una o más especies está prohibida) es una medida de manejo frecuentemente propuesta para recuperar stocks pesqueros e impulsar la productividad pesquera en aguas cercanas \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}.
-Recientes trabajos han demostrado que también pueden mitigar y proveer amortiguamiento ante el cambio climático \citep{roberts_2017-J9}, variabilidad ambiental \citep{micheli_2012-EU}, resolver problemas de pesca incidental \citep{hastings_2017-sm} y, en general, incrementar la biomasa, riqueza y densidades de organismos dentro de sus fronteras \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}.
+La sobrepesca y prácticas pesqueras no sostenibles son unas de las
+mayores amenazas para la conservación de los ecosistemas marinos del
+mundo \citep{halpern_2008-dK,halpern_2017-Zi}. La implementación de
+reservas marinas (\emph{i.e.} áreas donde la captura de una o más
+especies está prohibida) es una medida de manejo frecuentemente
+propuesta para recuperar stocks pesqueros e impulsar la productividad
+pesquera en aguas cercanas
+\citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}. Recientes
+trabajos han demostrado que también pueden mitigar y proveer
+amortiguamiento ante el cambio climático \citep{roberts_2017-J9},
+variabilidad ambiental \citep{micheli_2012-EU}, resolver problemas de
+pesca incidental \citep{hastings_2017-sm} y, en general, incrementar la
+biomasa, riqueza y densidades de organismos dentro de sus fronteras
+\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}.
 
-En México, las reservas marinas han sido comúnmente establecidas como zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por la Comisión Nacional de Áreas Naturales Protegidas (CONANP).
-Al día de hoy, 36 RBs protegen una porción del ambiente marino en México.
-Sin embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas.
-Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, esto aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores.
-La escasez de recursos monetarios y humanos de la limitan también el monitoreo y vigilancia de las reservas, y a su vez, el desempeño de la reserva.
+En México, las reservas marinas han sido comúnmente establecidas como
+zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por
+la Comisión Nacional de Áreas Naturales Protegidas (CONANP). Al día de
+hoy, 36 RBs protegen una porción del ambiente marino en México. Sin
+embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde
+las actividades pesqueras están prohibidas. Aunque la CONANP ha hecho
+esfuerzos importantes por involucrar a los actores durante la
+implementación de las reservas, esto aún se caracteriza por un proceso
+descendente, el cual conlleva a la falta de cumplimiento por parte de
+los actores. La escasez de recursos monetarios y humanos de la limitan
+también el monitoreo y vigilancia de las reservas, y a su vez, el
+desempeño de la reserva.
 
-Buscando promover una alternativa con procesos ascendentes para implementar reservas marinas, las Organizaciones de la Sociedad Civil (OSCs) comenzaron a trabajar con comunidades pesqueras para establecer reservas comunitarias \citep{uribe_2010-u2} .
-Estas son comúnmente establecidas dentro de zonas de concesión, una forma de derechos de uso territoriales para pesquerías (TURF, en inglés).
-Al permitir a los pescadores diseñar sus propias reservas, una mayor proporción de la comunidad está de acuerdo con los perímetros y reglas establecidas, y por lo tanto los respetan \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8} .
-Adicionalmente, los pescadores pueden implementar sus reservas por un periodo acordado (usualmente cinco años), después del cual la reserva puede ser abierta a la pesca.
-Esto provee a los pescadores con un sentido de confianza de que, en caso de ser necesario, aún tienen acceso a pescar esa zona.
-Las reservas son directamente vigiladas y monitoreadas por la comunidad, quienes comúnmente utilizan pequeñas embarcaciones (*e.g.* pangas) para patrullar la zona, o realizan avistamientos desde la costa en búsqueda de pescadores ilegales Aún así, las reservas comunitarias carecen de reconocimiento legal; por lo tanto, no hay forma de penalizar a los infractores.
+Buscando promover una alternativa con procesos ascendentes para
+implementar reservas marinas, las Organizaciones de la Sociedad Civil
+(OSCs) comenzaron a trabajar con comunidades pesqueras para establecer
+reservas comunitarias \citep{uribe_2010-u2} . Estas son comúnmente
+establecidas dentro de zonas de concesión, una forma de derechos de uso
+territoriales para pesquerías (TURF, en inglés). Al permitir a los
+pescadores diseñar sus propias reservas, una mayor proporción de la
+comunidad está de acuerdo con los perímetros y reglas establecidas, y
+por lo tanto los respetan
+\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8} .
+Adicionalmente, los pescadores pueden implementar sus reservas por un
+periodo acordado (usualmente cinco años), después del cual la reserva
+puede ser abierta a la pesca. Esto provee a los pescadores con un
+sentido de confianza de que, en caso de ser necesario, aún tienen acceso
+a pescar esa zona. Las reservas son directamente vigiladas y
+monitoreadas por la comunidad, quienes comúnmente utilizan pequeñas
+embarcaciones (\emph{e.g.} pangas) para patrullar la zona, o realizan
+avistamientos desde la costa en búsqueda de pescadores ilegales Aún así,
+las reservas comunitarias carecen de reconocimiento legal; por lo tanto,
+no hay forma de penalizar a los infractores.
 
-Sin embargo, en el 2014 una nueva norma \citep{nom049sagpesc_2014-V6} permite a los pescadores solicitar el establecimiento de reservas marinas bajo el nombre de “Zonas de refugio Pesquero” (ZRP).
-El manejo de las ZRP combina procesos ascendentes y descendentes al reconocer legalmente las reservas propuestas por las comunidades.
-Posterior a la revisión por parte de la Comisión Nacional de Acuacultura y Pesca (CONAPESCA) y la opinión técnica del Instituto Nacional de Acuacultura y Pesca (INAPESCA) las ZRP son establecidas por el periodo solicitado por los pescadores.
-El monitoreo y la vigilancia de las ZRP es típicamente llevado a cabo por la comunidad , con ayuda de OSCs locales.
-Hasta este cambio regulatorio, las reservas comunitarias no contaban con el soporte legal, y eran solamente reconocidas por la comunidad.
-Al día de hoy, existen 39 ZRP establecidas en el Pacífico, Golfo de California y Caribe Mexicano.
+Sin embargo, en el 2014 una nueva norma \citep{nom049sagpesc_2014-V6}
+permite a los pescadores solicitar el establecimiento de reservas
+marinas bajo el nombre de ``Zonas de refugio Pesquero'' (ZRP). El manejo
+de las ZRP combina procesos ascendentes y descendentes al reconocer
+legalmente las reservas propuestas por las comunidades. Posterior a la
+revisión por parte de la Comisión Nacional de Acuacultura y Pesca
+(CONAPESCA) y la opinión técnica del Instituto Nacional de Acuacultura y
+Pesca (INAPESCA) las ZRP son establecidas por el periodo solicitado por
+los pescadores. El monitoreo y la vigilancia de las ZRP es típicamente
+llevado a cabo por la comunidad , con ayuda de OSCs locales. Hasta este
+cambio regulatorio, las reservas comunitarias no contaban con el soporte
+legal, y eran solamente reconocidas por la comunidad. Al día de hoy,
+existen 39 ZRP establecidas en el Pacífico, Golfo de California y Caribe
+Mexicano.
 
-Aunque existen tres aproximaciones generales para implementar reservas marinas en México (*i.e.* Zonas núcleo dentro de AMP, reservas comunitarias y Zonas de Refugio Pesquero), aún no comprendemos a fondo las características sociales que permiten su efectividad.
-La ciencia de reservas marinas se ha enfocado ampliamente en los efectos biológicos que estas tienen \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,afflerbach_2014-HP,krueck_2017-J1}.
-Aunque el aspecto ecológico de las reservas es importante para su éxito, su efectividad también depende del estado socioeconómico y los sistemas de gobernanza de las comunidades pesqueras.
+Aunque existen tres aproximaciones generales para implementar reservas
+marinas en México (\emph{i.e.} Zonas núcleo dentro de AMP, reservas
+comunitarias y Zonas de Refugio Pesquero), aún no comprendemos a fondo
+las características sociales que permiten su efectividad. La ciencia de
+reservas marinas se ha enfocado ampliamente en los efectos biológicos
+que estas tienen
+\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,afflerbach_2014-HP,krueck_2017-J1}.
+Aunque el aspecto ecológico de las reservas es importante para su éxito,
+su efectividad también depende del estado socioeconómico y los sistemas
+de gobernanza de las comunidades pesqueras.
 
-La literatura indica que diferentes características influyen en el éxito de una reserva.
-En Palau, por ejemplo, la edad (*i.e.* tiempo transcurrido desde implementación), tamaño y hábitat contenido son características claves que determinan la efectividad \citep{friedlander_2017-oI}.
-Por otro lado, en el Mar Mediterráneo, @difranco_2016-Xw identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks y el ingreso económicos a los pescadores, a la vez que se presenta una mayor aceptación social de las prácticas de manejo.
-En una aproximación global, @edgar_2014-UO encuentran que la procuración, edad, tamaño y aislamiento son determinantes de la efectividad de las reservas.
-Por lo tanto, observamos que las características que habilitan el éxito varían a través de regiones, y poco esfuerzo se ha hecho por comprender estas interacciones en México.
+La literatura indica que diferentes características influyen en el éxito
+de una reserva. En Palau, por ejemplo, la edad (\emph{i.e.} tiempo
+transcurrido desde implementación), tamaño y hábitat contenido son
+características claves que determinan la efectividad
+\citep{friedlander_2017-oI}. Por otro lado, en el Mar Mediterráneo,
+@difranco\_2016-Xw identifican que la procuración y vigilancia,
+presencia de un plan de manejo, participación de pescadores en el
+manejo, representación de pescadores en la toma de decisiones y
+promoción de la pesca sustentable son los cinco factores que incrementan
+la salud de los stocks y el ingreso económicos a los pescadores, a la
+vez que se presenta una mayor aceptación social de las prácticas de
+manejo. En una aproximación global, @edgar\_2014-UO encuentran que la
+procuración, edad, tamaño y aislamiento son determinantes de la
+efectividad de las reservas. Por lo tanto, observamos que las
+características que habilitan el éxito varían a través de regiones, y
+poco esfuerzo se ha hecho por comprender estas interacciones en México.
 
-# Materials and Methods
+\section{Materials and Methods}\label{materials-and-methods}
 
-This section may be divided by subheadings.
-This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated.
-For experiments reporting results on animal or human subject research, an ethics approval statement should be included in this section (for further information, see section Materials and Data Policies)
+This section may be divided by subheadings. This section should contain
+sufficient detail so that when read in conjunction with cited
+references, all procedures can be repeated. For experiments reporting
+results on animal or human subject research, an ethics approval
+statement should be included in this section (for further information,
+see section Materials and Data Policies)
 
-## Study area
+\subsection{Study area}\label{study-area}
 
-```{r}
-suppressPackageStartupMessages({
-  library(sp)
-  library(tmap)
-  library(tmaptools)
-  library(MPAtools)
-  library(magrittr)
-  library(kableExtra)
-  library(broom)
-  library(tidyverse)
-})
-```
+\begin{figure}
+\centering
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{\label{fig:unnamed-chunk-2}Mapa de la localización general de las
+comunidades de estudio. El panel de la derecha es un acercamiento a las
+comunidades de María Elena y Punta Herrero.}
+\end{figure}
 
-```{r, message = F, warning = F, fig.cap = "Mapa de la localización general de las comunidades de estudio.
-El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero.", fig.pos = 'h'}
+\subsection{Data collection}\label{data-collection}
 
-####
+\subsection{Data analysis}\label{data-analysis}
 
-proj <- CRS("+proj=longlat +datum=NAD27")
+\subsubsection{Biological}\label{biological}
 
-data(World)
-W <- spTransform(World, proj)
-W <- W[!W$name == "Mexico",]
+\subsubsection{Socioeconomic}\label{socioeconomic}
 
-load(file = "../Data/coastline_mx.Rda")
+\subsubsection{Governance}\label{governance}
 
-locations <- cbind(x_coord = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087), y_coord = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957)) %>% 
-  SpatialPoints(proj4string = proj)
+\section{Results}\label{results}
 
-location_labels <- data.frame(X = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087),
-                              Y = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957),
-                              Location = c("Isla Natividad", "Maria Elena", "Punta Herrero", "Puerto Libertad", "Guaymas"),
-                              Region = c("Pacific", "Caribbean","Caribbean", "Gulf of California", "Gulf of California"))
-coordinates(location_labels) <- c("X", "Y")
-proj4string(location_labels) <- proj
-location_labels <- spTransform(location_labels, proj)
+\section{Discussion}\label{discussion}
 
-SAM_polygon <- cbind(x_coord = c(-88, -87, -87, -88, -88), y_coord = c(19, 19, 20, 20, 20)) %>% 
-  Polygon() %>% 
-  list() %>% 
-  Polygons(srl = ., ID = 1) %>% 
-  list() %>% 
-  SpatialPolygons(proj4string = proj)
+This section may be divided by subheadings. Discussions should cover the
+key findings of the study: discuss any prior art related to the subject
+so to place the novelty of the discovery in the appropriate context;
+discuss the potential short-comings and limitations on their
+interpretations; discuss their integration into the current
+understanding of the problem and how this advances the current views;
+speculate on the future direction of the research and freely postulate
+theories that could be tested in the future.
 
-mex <- tm_shape(W, bbox = bb(baja2)) +
-  tm_polygons() +
-  tm_shape(baja2) +
-  tm_polygons() +
-  # tm_shape(SAM_polygon) +
-  # tm_borders(col = "black", lwd = 2) +
-  tm_shape(location_labels) +
-  tm_symbols(shape = "Location", col = "Region", palette = c("#33A02C", "#1F78B4", "#E31A1C"), border.col = "black", size = 0.5) +
-  tm_compass(position = c("left", "bottom")) +
-  tm_scale_bar(position = c("left", "bottom")) +
-  tm_grid(n.x = 4, n.y = 5)
-
-sam <- tm_shape(baja2, bbox = bb(xlim = c(-88, -86.726716),
-                          ylim = c(19.25, 19.5))) +
-  tm_polygons() +
-  tm_shape(location_labels) +
-  tm_symbols(shape = "Location", col = "Region", palette = c("#33A02C", "#1F78B4", "#E31A1C"), border.col = "black", size = 0.5) +
-  tm_compass() +
-  tm_scale_bar() +
-  tm_legend(show=FALSE) +
-  tm_grid(n.x = 4, n.y = 5)
-
-tmap_arrange(mex, sam, ncol = 2)
-```
-
-## Data collection
-
-## Data analysis
-
-### Biological
-
-### Socioeconomic
-
-### Governance
-
-# Results
-
-# Discussion
-
-This section may be divided by subheadings.
-Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.
-
-Summary of main findings
-Limitations
-Conclusions
+Summary of main findings Limitations Conclusions
 
@@
 
-The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+The authors declare that the research was conducted in the absence of
+any commercial or financial relationships that could be construed as a
+potential conflict of interest.
 
@@
 
-JC analyzed and interpreted data, discussed the results and wrote the manuscrip. SF and JT edited the manuscript and discussed the results.
+JC analyzed and interpreted data, discussed the results and wrote the
+manuscrip. SF and JT edited the manuscript and discussed the results.
 
 \section*{Funding}
-Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
+
+Details of all funding sources should be provided, including grant
+numbers if applicable. Please ensure to add all necessary funding
+information, as after publication this is no longer possible.
 
 \section*{Acknowledgments}
-This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.
+
+This is a short text to acknowledge the contributions of specific
+colleagues, institutions, or agencies that aided the efforts of the
+authors.
 
 \section*{Supplemental Data}
- \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
-LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder
 
-\bibliographystyle{frontiersinSCNS_ENG_HUMS}
-\bibliography{references}
+\href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material}
+should be uploaded separately on submission, if there are Supplementary
+Figures, please include the caption in the same file as the figure.
+LaTeX Supplementary Material templates can be found in the Frontiers
+LaTeX folder
+
+\bibliographystyle{frontiersinSCNS_ENG_HUMS}\bibliography{references}
 
 \section*{Figure captions}
+
+\end{document}

### 2018-01-11 TechnicalReport/Manuscript.Rmd
Finalized LaTeX - Rmd integration, all works now. Let´s start running analyses

--- 2018-01-11 Manuscript.Rmd
@@
 
-load(file = "Data/coastline_mx.Rda")
+load(file = "../Data/coastline_mx.Rda")
 
@@
 ```{r}
-read.csv("Data/governance_summary.csv") %>% 
+read.csv("../Data/governance_summary.csv") %>% 
   knitr::kable(booktabs = T, col.names = c("Comunidad", "Reservas", "Año Implementación", "Año Reconocimiento", "Tipo", "Manejo de recursos", "Grado de pesca ilegal", "Inspección y Vigilancia", "Organizaciones pesqueras", "Edad"), caption = "Características de gobernanza generales de las cinco comunidades estudiadas.") %>%
@@
 ```{r}
-fish <- read.csv("Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
+fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
 
-invert <- read.csv("Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
+invert <- read.csv("../Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
 
-load("Data/conapesca_2000-2015.RData")
+load("../Data/conapesca_2000-2015.RData")
 
-cpi <- read.csv("Data/CPI.csv", stringsAsFactors = F, strip.white = T)
+cpi <- read.csv("../Data/CPI.csv", stringsAsFactors = F, strip.white = T)
 
@@
 
-gob <- read.csv("Data/governance.csv", strip.white = T, stringsAsFactors = T)
+gob <- read.csv("../Data/governance.csv", strip.white = T, stringsAsFactors = T)
 ```

### 2018-01-19 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Finalized map (for now) and started data curation

@@
 rtitle: "Mexican marine reserves"
-author: "Juan Carlos Villaseñor-Derbez,$^{1,*}$, Stuart Fulton,$^{2}$ Jorge Torre,$^{2}$"
+author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
 institution: "
@@
 email: "jvillasenor@bren.ucsb.edu"
-abstract: "This is my abstract"
-keywords: "Marine Reserves, Marine Conservation, Fisheries, Citizen Science, Mexico"
+abstract: "En México, las reservas marinas han sido comúnmente establecidas como zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por la Comisión Nacional de Áreas Naturales Protegidas (CONANP).
+Al día de hoy, 36 RBs protegen una porción del ambiente marino en México.
+Sin embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas.
+Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, esto aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores.
+La escasez de recursos monetarios y humanos de la limitan también el monitoreo y vigilancia de las reservas, y a su vez, el desempeño de la reserva.
+En México, las reservas marinas han sido comúnmente establecidas como zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por la Comisión Nacional de Áreas Naturales Protegidas (CONANP).
+Al día de hoy, 36 RBs protegen una porción del ambiente marino en México.
+Sin embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas.
+Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, esto aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores.
+La escasez de recursos monetarios y humanos de la limitan también el monitoreo y vigilancia de las reservas, y a su vez, el desempeño de la reserva."
+keywords: "Marine Reserves, Marine Conservation, Fisheries, Citizen Science, Mexico. "
 # biblio: test.bib
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+options(knitr.table.format = "latex")
 ```
@@
 
-Sin embargo, en el 2014 una nueva norma \citep{nom049sagpesc_2014-V6} permite a los pescadores solicitar el establecimiento de reservas marinas bajo el nombre de “Zonas de refugio Pesquero” (ZRP).
+Sin embargo, en el 2014 una nueva norma \citep{nom} permite a los pescadores solicitar el establecimiento de reservas marinas bajo el nombre de “Zonas de refugio Pesquero” (ZRP).
 El manejo de las ZRP combina procesos ascendentes y descendentes al reconocer legalmente las reservas propuestas por las comunidades.
@@
 En Palau, por ejemplo, la edad (*i.e.* tiempo transcurrido desde implementación), tamaño y hábitat contenido son características claves que determinan la efectividad \citep{friedlander_2017-oI}.
-Por otro lado, en el Mar Mediterráneo, @difranco_2016-Xw identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks y el ingreso económicos a los pescadores, a la vez que se presenta una mayor aceptación social de las prácticas de manejo.
-En una aproximación global, @edgar_2014-UO encuentran que la procuración, edad, tamaño y aislamiento son determinantes de la efectividad de las reservas.
+Por otro lado, en el Mar Mediterráneo, \citet{difranco_2016-Xw} identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks y el ingreso económicos a los pescadores, a la vez que se presenta una mayor aceptación social de las prácticas de manejo.
+En una aproximación global, \citet{edgar_2014-UO} encuentran que la procuración, edad, tamaño y aislamiento son determinantes de la efectividad de las reservas.
 Por lo tanto, observamos que las características que habilitan el éxito varían a través de regiones, y poco esfuerzo se ha hecho por comprender estas interacciones en México.
@@
 
-```{r}
+... a map of the general location of the study sites is shown in figure \ref{fig:map}.
+
+```{r loadp_kgs}
 suppressPackageStartupMessages({
   library(sp)
+  library(sf)
+  library(ggsn)
+  library(patchwork)
   library(tmap)
-  library(tmaptools)
   library(MPAtools)
@@
   library(broom)
+  library(cowplot)
   library(tidyverse)
@@
 
-```{r, message = F, warning = F, fig.cap = "Mapa de la localización general de las comunidades de estudio.
-El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero.", fig.pos = 'h'}
-
-####
+```{r create_map, include = T, fig.cap = "\\label{fig:map}Mapa de la localización general de las comunidades de estudio.
+El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero."}
 
@@
 W <- spTransform(World, proj)
-W <- W[!W$name == "Mexico",]
+W <- W[!W$name == "Mexico",] %>% 
+  st_as_sf()
 
-load(file = "../Data/coastline_mx.Rda")
-
-locations <- cbind(x_coord = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087), y_coord = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957)) %>% 
-  SpatialPoints(proj4string = proj)
+coastline_mx <- readRDS("../Data/Spatial/coastline_mx.rds")
 
@@
                               Location = c("Isla Natividad", "Maria Elena", "Punta Herrero", "Puerto Libertad", "Guaymas"),
-                              Region = c("Pacific", "Caribbean","Caribbean", "Gulf of California", "Gulf of California"))
+                              Region = c("Pacific", "Caribbean","Caribbean", "Gulf of California", "Gulf of California"),
+                              stringsAsFactors = F)
 coordinates(location_labels) <- c("X", "Y")
 proj4string(location_labels) <- proj
-location_labels <- spTransform(location_labels, proj)
+location_labels <- spTransform(location_labels, proj) %>% 
+  st_as_sf()
 
-SAM_polygon <- cbind(x_coord = c(-88, -87, -87, -88, -88), y_coord = c(19, 19, 20, 20, 20)) %>% 
-  Polygon() %>% 
-  list() %>% 
-  Polygons(srl = ., ID = 1) %>% 
-  list() %>% 
-  SpatialPolygons(proj4string = proj)
+main_map <- ggplot() +
+  geom_sf(data = W) +
+  geom_sf(data = coastline_mx) +
+  geom_sf(data = location_labels, aes(color = Location, fill = Location), size = 3) +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "none")
 
-mex <- tm_shape(W, bbox = bb(baja2)) +
-  tm_polygons() +
-  tm_shape(baja2) +
-  tm_polygons() +
-  # tm_shape(SAM_polygon) +
-  # tm_borders(col = "black", lwd = 2) +
-  tm_shape(location_labels) +
-  tm_symbols(shape = "Location", col = "Region", palette = c("#33A02C", "#1F78B4", "#E31A1C"), border.col = "black", size = 0.5) +
-  tm_compass(position = c("left", "bottom")) +
-  tm_scale_bar(position = c("left", "bottom")) +
-  tm_grid(n.x = 4, n.y = 5)
+mex <- main_map +
+  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863))
 
-sam <- tm_shape(baja2, bbox = bb(xlim = c(-88, -86.726716),
-                          ylim = c(19.25, 19.5))) +
-  tm_polygons() +
-  tm_shape(location_labels) +
-  tm_symbols(shape = "Location", col = "Region", palette = c("#33A02C", "#1F78B4", "#E31A1C"), border.col = "black", size = 0.5) +
-  tm_compass() +
-  tm_scale_bar() +
-  tm_legend(show=FALSE) +
-  tm_grid(n.x = 4, n.y = 5)
+legends <- main_map +
+  lims(x = c(-87, -86.5), y = c(19.4, 19.6)) +
+  theme_map() +
+  theme(legend.justification = c(0.5, 0.5),
+        legend.position = c(0.5, 0.5),
+        legend.background = element_rect(fill = "transparent")) +
+  ggExtra::removeGrid() +
+  blank()
 
-tmap_arrange(mex, sam, ncol = 2)
+sam <- main_map +
+  lims(x = c(-88, -87), y = c(19.1, 19.7))
+
+goc <- main_map +
+  lims(x = c(-115.5, -110.5), y = c(27, 30))
+
+plot_grid(mex, sam, goc, legends, labels = c("A", "C", "B"), ncol = 2)
+
 ```
@@
 ## Data analysis
+
+```{r load_data}
+fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
+
+invert <- read.csv("../Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
+
+load("../Data/conapesca_2000-2015.RData")
+
+cpi <- read.csv("../Data/CPI.csv", stringsAsFactors = F, strip.white = T)
+
+cpi_2014 <- cpi$CPI[cpi$Ano == 2014]
+
+conapesca %<>%
+  filter(UnidadEconomica %in% c("Scpp cozumel scl",
+                                "Scpp jose maria azcorra scl",
+                                "Scpp buzos y pescadores de la baja california scl")) %>% 
+  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
+  group_by(UnidadEconomica, Ano, NombreComun) %>%
+  summarize(PesoDesembarcado = sum(PesoDesembarcado), 
+            PesoVivo = sum(PesoVivo),
+            Valor = sum(Valor)) %>%
+  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
+  left_join(cpi, by = "Ano") %>% 
+  mutate(Valor = Valor * (CPI / cpi_2014))
+
+gob <- read.csv("../Data/governance.csv", strip.white = T, stringsAsFactors = T)
+```
+
+```{r complete_fish, eval = F}
+fish2 <- fish %>% 
+  select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
+  group_by(Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  ungroup() %>% 
+  complete(Ano, Transecto, nesting(Sitio, Zona, GeneroEspecie), fill = list(Abundancia = 0L)) %>% 
+  select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
+  arrange()
+```
+
+Table \ref{table:indicators} is the way to cite a table.
+
+```{r}
+data.frame(Category = c("Biological", "Biological","Biological","Biological","Biological","Biological", "Socioeconomic", "Socioeconomic", "Governance", "Governance", "Governance", "Governance", "Governance"),
+           Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Densidad de especies objetivo", "Ingresos por especies objetivo", "Arribos de especies objetivo", "Tipo de acceso a la pesquería", "Grado de pesca ilegal", "Procuración de la reserva", "Tipo de organización pesquera", "Edad de la reserva"), stringsAsFactors = F) %>%
+  knitr::kable(caption = "\\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.")
+```
 
@@
 
+\paragraph*{S1 Figure}
+\label{S1_Figure}
+Maps of the marine reserves and corresponding control sites at each community.
+
+\paragraph*{S2 Table}
+\label{S2_Table}
+Table with a general overview of on the governance characteristics of each community.
+
 \bibliographystyle{frontiersinSCNS_ENG_HUMS}

### 2018-01-19 Frontiers_Manuscript/Villasenor-Derbez.tex
Finalized map (for now) and started data curation

@@
 \def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{Villasenor-Derbez {et~al.}}
-\def\Authors{Juan Carlos Villaseñor-Derbez,\(^{1,*}\), Stuart Fulton,\(^{2}\) Jorge
-Torre,\(^{2}\)}
+\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
+\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Stuart Fulton\(^{2}\), Jorge
+Torre\(^{2}\)}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 
-This is my abstract
+En México, las reservas marinas han sido comúnmente establecidas como
+zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por
+la Comisión Nacional de Áreas Naturales Protegidas (CONANP). Al día de
+hoy, 36 RBs protegen una porción del ambiente marino en México. Sin
+embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde
+las actividades pesqueras están prohibidas. Aunque la CONANP ha hecho
+esfuerzos importantes por involucrar a los actores durante la
+implementación de las reservas, esto aún se caracteriza por un proceso
+descendente, el cual conlleva a la falta de cumplimiento por parte de
+los actores. La escasez de recursos monetarios y humanos de la limitan
+también el monitoreo y vigilancia de las reservas, y a su vez, el
+desempeño de la reserva. En México, las reservas marinas han sido
+comúnmente establecidas como zonas núcleo dentro de Reservas de la
+Biósfera (RBs), administradas por la Comisión Nacional de Áreas
+Naturales Protegidas (CONANP). Al día de hoy, 36 RBs protegen una
+porción del ambiente marino en México. Sin embargo, solamente 26 de
+estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras
+están prohibidas. Aunque la CONANP ha hecho esfuerzos importantes por
+involucrar a los actores durante la implementación de las reservas, esto
+aún se caracteriza por un proceso descendente, el cual conlleva a la
+falta de cumplimiento por parte de los actores. La escasez de recursos
+monetarios y humanos de la limitan también el monitoreo y vigilancia de
+las reservas, y a su vez, el desempeño de la reserva.
 
@@
 \tiny
- \keyFont{ \section{Keywords:} Marine Reserves, Marine Conservation, Fisheries, Citizen Science, Mexico}
+ \keyFont{ \section{Keywords:} Marine Reserves, Marine Conservation, Fisheries, Citizen Science,
+Mexico.}
 
@@
 
-Sin embargo, en el 2014 una nueva norma \citep{nom049sagpesc_2014-V6}
-permite a los pescadores solicitar el establecimiento de reservas
-marinas bajo el nombre de ``Zonas de refugio Pesquero'' (ZRP). El manejo
-de las ZRP combina procesos ascendentes y descendentes al reconocer
-legalmente las reservas propuestas por las comunidades. Posterior a la
-revisión por parte de la Comisión Nacional de Acuacultura y Pesca
-(CONAPESCA) y la opinión técnica del Instituto Nacional de Acuacultura y
-Pesca (INAPESCA) las ZRP son establecidas por el periodo solicitado por
-los pescadores. El monitoreo y la vigilancia de las ZRP es típicamente
-llevado a cabo por la comunidad , con ayuda de OSCs locales. Hasta este
-cambio regulatorio, las reservas comunitarias no contaban con el soporte
-legal, y eran solamente reconocidas por la comunidad. Al día de hoy,
-existen 39 ZRP establecidas en el Pacífico, Golfo de California y Caribe
-Mexicano.
+Sin embargo, en el 2014 una nueva norma \citep{nom} permite a los
+pescadores solicitar el establecimiento de reservas marinas bajo el
+nombre de ``Zonas de refugio Pesquero'' (ZRP). El manejo de las ZRP
+combina procesos ascendentes y descendentes al reconocer legalmente las
+reservas propuestas por las comunidades. Posterior a la revisión por
+parte de la Comisión Nacional de Acuacultura y Pesca (CONAPESCA) y la
+opinión técnica del Instituto Nacional de Acuacultura y Pesca (INAPESCA)
+las ZRP son establecidas por el periodo solicitado por los pescadores.
+El monitoreo y la vigilancia de las ZRP es típicamente llevado a cabo
+por la comunidad , con ayuda de OSCs locales. Hasta este cambio
+regulatorio, las reservas comunitarias no contaban con el soporte legal,
+y eran solamente reconocidas por la comunidad. Al día de hoy, existen 39
+ZRP establecidas en el Pacífico, Golfo de California y Caribe Mexicano.
 
@@
 \citep{friedlander_2017-oI}. Por otro lado, en el Mar Mediterráneo,
-@difranco\_2016-Xw identifican que la procuración y vigilancia,
+\citet{difranco_2016-Xw} identifican que la procuración y vigilancia,
 presencia de un plan de manejo, participación de pescadores en el
@@
 vez que se presenta una mayor aceptación social de las prácticas de
-manejo. En una aproximación global, @edgar\_2014-UO encuentran que la
-procuración, edad, tamaño y aislamiento son determinantes de la
+manejo. En una aproximación global, \citet{edgar_2014-UO} encuentran que
+la procuración, edad, tamaño y aislamiento son determinantes de la
 efectividad de las reservas. Por lo tanto, observamos que las
@@
 
+\ldots{} a map of the general location of the study sites is shown in
+figure \ref{fig:map}.
+
 \begin{figure}
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{\label{fig:unnamed-chunk-2}Mapa de la localización general de las
-comunidades de estudio. El panel de la derecha es un acercamiento a las
-comunidades de María Elena y Punta Herrero.}
+\includegraphics{Villasenor-Derbez_files/figure-latex/create_map-1.pdf}
+\caption{(\#fig:create\_map)\label{fig:map}Mapa de la localización
+general de las comunidades de estudio. El panel de la derecha es un
+acercamiento a las comunidades de María Elena y Punta Herrero.}
 \end{figure}
@@
 \subsection{Data analysis}\label{data-analysis}
+
+Table \ref{table:indicators} is the way to cite a table.
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-1}\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.}
+\centering
+\begin{tabular}[t]{l|l}
+\hline
+Category & Indicador\\
+\hline
+Biological & Índice de diversidad de shannon\\
+\hline
+Biological & Riqueza\\
+\hline
+Biological & Densidad\\
+\hline
+Biological & Nivel trófico\\
+\hline
+Biological & Biomasa\\
+\hline
+Biological & Densidad de especies objetivo\\
+\hline
+Socioeconomic & Ingresos por especies objetivo\\
+\hline
+Socioeconomic & Arribos de especies objetivo\\
+\hline
+Governance & Tipo de acceso a la pesquería\\
+\hline
+Governance & Grado de pesca ilegal\\
+\hline
+Governance & Procuración de la reserva\\
+\hline
+Governance & Tipo de organización pesquera\\
+\hline
+Governance & Edad de la reserva\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
+\paragraph*{S1 Figure}
+\label{S1_Figure}
+
+Maps of the marine reserves and corresponding control sites at each
+community.
+
+\paragraph*{S2 Table}
+\label{S2_Table}
+
+Table with a general overview of on the governance characteristics of
+each community.
+
 \bibliographystyle{frontiersinSCNS_ENG_HUMS}\bibliography{references}

### 2018-01-19 Test.Rmd
Finalized map (for now) and started data curation

---
title: "Testing different model configurations and specifications"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    toc_collapse: no
---

##  Cargar paquetes

```{r}
library(tidyverse)
```

# Probar especificacion del modelo

## Leer datos

```{r}
fish <- read_csv(file = "Test.csv", col_types = cols())
```

## Inspeccionar los datos

```{r}
skimr::skim(fish)
```

## Visualizar datos crudos por Comunidad y par RC

```{r}
ggplot(fish, aes(x = Year, y = Abundance, color = Zone, fill = Zone, shape = RC)) +
  geom_smooth(method = "lm", alpha = 0.1) +
  geom_point() +
  scale_color_brewer(palette = "Set1", direction = -1) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  facet_wrap(Community ~ RC) +
  theme_bw()
```

## Calcular promedios por Ano, comunidad, RC y zona

```{r}
group_by(fish, Community, RC, Year, Zone) %>% 
  summarize(mean(Abundance))
```

## Calcular la diferencia entre antes y despues

```{r}
group_by(fish, Community, RC, Year, Zone) %>% 
  summarize(m = mean(Abundance)) %>% 
  spread(Year, m) %>% 
  magrittr::set_colnames(value = c("Community", "RC", "Zone", "Before", "After")) %>% 
  mutate(Diff = After - Before)
```

## Ajustar el modelo

```{r, results = 'asis'}
model1 <- mutate(fish, Year = as.factor(Year)) %>% 
  lm(Abundance ~ Year * Zone * RC + 0, data = .)

stargazer::stargazer(model1, type = "html", single.row = T)
```

## Probar el modelo

```{r}
data.frame(Year = as.factor(c(1, 1, 2, 2, 1, 1, 2, 2, 1, 1, 2, 2, 1, 1, 2, 2)),
           Community = c("A","A","A","A","A","A","A","A","B","B","B","B","B","B","B","B"),
           RC = c("AB", "AB", "AB", "AB", "CD", "CD", "CD", "CD", "EF", "EF", "EF", "EF", "GH", "GH", "GH", "GH"),
           Zone = c("Control", "Reserve")) %>% 
  mutate(mean = predict(object = model1, .))
```

Tiene sentido, el promedio predeciro corresponde con los observados de arriba. Carguemos datos nuevos que incluyen un tercer ano

# Probando el modelo con 3 anos

## Cargar datos

```{r}
fish <- read_csv(file = "Test2.csv", col_types = cols())
```

## Inspeccionar datos

```{r}
skimr::skim(fish)
```

## Visualizar datos crudos por Comunidad y par RC

```{r}
ggplot(fish, aes(x = Year, y = Abundance, color = Zone, fill = Zone, shape = RC)) +
  geom_smooth(method = "lm", alpha = 0.1) +
  geom_point() +
  scale_color_brewer(palette = "Set1", direction = -1) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  facet_wrap(Community ~ RC) +
  theme_bw()
```

## Ajustar el nuevo modelo.

Tiene la misma estructura, pero los datos tienen 3 anos

```{r, results = 'asis'}
model1 <- mutate(fish, Year = as.factor(Year)) %>% 
  lm(Abundance ~ Year * Zone * RC + 0, data = .)

stargazer::stargazer(model1, type = "html", single.row = T)
```

## Probar el modelo

```{r}
data.frame(Year = as.factor(c(3)),
           Community = c("A"),
           RC = c("AB"),
           Zone = c("Control", "Reserve")) %>% 
  mutate(mean = predict(object = model1, .))
```

Las predicciones son correctas, pues coinciden con los valores promedios del tercer ano en los datos crudos:

```{r}
filter(fish, Year == 3) %>% 
  group_by(RC, Zone) %>% 
  summarize(mean = mean(Abundance))
```

## Visualicemos los efectos del modelo.

```{r}
broom::tidy(model1) %>% 
  ggplot(aes(x = term, y = estimate)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
  geom_point(fill = "steelblue", size = 4, shape = 21, color = "black") +
  coord_flip() +
  theme_classic()
```

# Probar efectos fijos por ano vs. pre-post

Usemos los mismos datos, pero solamente la comunidad A para comparar Pre-Post vs. Efectos fijos por ano

## Primero efectos fijos por ano:

```{r}
(test_results <- filter(fish, Community == "A") %>% 
  mutate(Year = as.factor(Year)) %>% 
  lm(Abundance ~ Year * Zone * RC + 0, data = .) %>% 
  broom::tidy())

ggplot(test_results, aes(x = term, y = estimate)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
  geom_point(fill = "steelblue", size = 4, shape = 21, color = "black") +
  coord_flip() +
  theme_classic()
```

## Ahora modificando Ano para tener pre y post:

```{r}
(test_results <- filter(fish, Community == "A") %>% 
  mutate(Year = as.factor(ifelse(Year == 1, "A", "D"))) %>% 
  lm(Abundance ~ Year * Zone * RC + 0, data = .) %>% 
  broom::tidy())

ggplot(test_results, aes(x = term, y = estimate)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
  geom_point(fill = "steelblue", size = 4, shape = 21, color = "black") +
  coord_flip() +
  theme_classic()
```

Perdemos los terminos del Ano2 = 10 y Ano3 = 15, Ano2:Reserva = 2 y Ano3:Reserva = 5, y terminamos con AnoD = 12.5 y AnoD:Reserva, que son los promedios entre los perdidos, respectivamente. Esta forma de modelarlo nos dice el efecto PROMEDIO de la reserva, que podria subestimar el efecto real. Es mejor usar los efectos fijos por ano.

Que pasa si tenemos reservas que comienzan en diferentes anos?

# Reservas con anos diferentes

```{r}
fish <- read_csv(file = "Test3.csv", col_types = cols())
```

```{r, fig.width = 4, fig.height = 8}
ggplot(fish, aes(x = Year, y = Abundance, color = Zone, fill = Zone, shape = RC)) +
  geom_smooth(method = "lm", alpha = 0.1) +
  geom_point() +
  scale_color_brewer(palette = "Set1", direction = -1) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  facet_wrap(Community ~ RC, ncol = 1) +
  theme_bw()
```

```{r, results = 'asis'}
model1 <- mutate(fish, Year = as.factor(Year)) %>% 
  lm(Abundance ~ Year * Zone * RC + 0, data = .)

stargazer::stargazer(model1, type = "html", single.row = T)
```

```{r}
data.frame(Year = as.factor(c(3)),
           Community = c("A"),
           RC = c("AB"),
           Zone = c("Control", "Reserve")) %>% 
  mutate(mean = predict(object = model1, .))
```

```{r}
broom::tidy(model1) %>% 
  ggplot(aes(x = term, y = estimate)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
  geom_point(fill = "steelblue", size = 4, shape = 21, color = "black") +
  coord_flip() +
  theme_classic()
```

Nada! Algunos coeficientes no pueden ser estimados, pero es normal. Los coeficientes estimados corresponden a valores que no existen. Por ejemplo, la reserva RC = AB existe en anos 1, 2 y 3. El valor de Ano5 para esta reserva es NA, porque la reserva no existe en ese punto en el tiempo.

# Como completar los datos?

```{r}
df <- read_csv(file = "Test4.csv", col_types = cols())
```

```{r}
skimr::skim(df)
```

```{r}

complete(df, Year, Community, Transect, nesting(Site, Zone, GenusSpecies), fill = list(Abundance = 0)) %>% 
  select(Year, Community, Site, Zone, Transect, GenusSpecies, Abundance) %>% 
  arrange(Year, Community, Site, Zone, Transect, GenusSpecies)

```

### 2018-01-20 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Added initial model fit

@@
 ```{r load_data}
-fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
+fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F) %>% 
+  filter(Comunidad %in% c("Isla Natividad", "Puerto Libertad", "Guaymas", "Maria Elena", "Punta Herrero")) %>% 
+  filter(!is.na(RC))
 
-invert <- read.csv("../Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
+invert <- read.csv("../Data/invert_data.csv", strip.white = T, stringsAsFactors = F) %>% 
+  filter(Comunidad %in% c("Isla Natividad", "Puerto Libertad", "Guaymas", "Maria Elena", "Punta Herrero")) %>% 
+  filter(!is.na(RC))
 
@@
 
-```{r complete_fish, eval = F}
-fish2 <- fish %>% 
+### Biological
+
+```{r, prepare_covars}
+fish_covars <- fish %>% 
+  select(RC, Ano, Sitio, Zona, Transecto, ProfundidadInicial, Visibilidad, Temperatura) %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize_all("mean", na.rm = T)
+
+invert_covars <- invert %>% 
+  select(RC, Ano, Sitio, Zona, Transecto, ProfundidadInicial, Visibilidad, Temperatura) %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize_all("mean", na.rm = T)
+```
+
+```{r complete_fish}
+fish_complete <- fish %>% 
   select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
@@
   complete(Ano, Transecto, nesting(Sitio, Zona, GeneroEspecie), fill = list(Abundancia = 0L)) %>% 
+  left_join(fish_covars, by = c("Ano", "Sitio", "Zona", "Transecto")) %>% 
+  select(Ano, RC, Sitio, Zona, Transecto, ProfundidadInicial, Visibilidad, Temperatura, GeneroEspecie, Abundancia) %>% 
+  arrange()
+
+invert_complete <- invert %>%
   select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
+  group_by(Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  ungroup() %>% 
+  complete(Ano, Transecto, nesting(Sitio, Zona, GeneroEspecie), fill = list(Abundancia = 0L)) %>% 
+  left_join(fish_covars, by = c("Ano", "Sitio", "Zona", "Transecto")) %>% 
+  select(Ano, RC, Sitio, Zona, Transecto, ProfundidadInicial, Visibilidad, Temperatura, GeneroEspecie, Abundancia) %>% 
   arrange()
+```
+
+```{r model_fit}
+fish_abundance_model <- fish_complete %>% 
+  mutate(Ano = as.factor(Ano)) %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  ungroup() %>% 
+  glm(Abundancia ~ Ano * Zona * RC + 0, data = ., family = "poisson")
+```
+
+```{r, eval = F}
+fish_complete %>% 
+  mutate(Ano = as.factor(Ano)) %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  ungroup() %>% 
+  mutate(predictions = predict(fish_abundance_model, ., type = "response")) %>% 
+  ggplot(aes(x = predictions, y = Abundancia, color = RC)) +
+  geom_point() +
+  theme_classic() +
+  geom_abline() +
+  theme(legend.position = "none")
+
+fish_complete %>% 
+  mutate(Ano = as.factor(Ano)) %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  ungroup() %>% 
+  mutate(predictions = predict(fish_abundance_model, ., type = "response"),
+         RMSE = predictions - Abundancia) %>% 
+  summarize_at(.vars = "RMSE", .funs = function(x){sqrt(mean(x^2, na.rm = T))})
+
+glance(fish_abundance_model)
+```
+
+```{r}
+broom::tidy(fish_abundance_model) %>% 
+  filter(str_detect(term ,":ZonaReserva"), !str_detect(term ,"RC")) %>%
+  # mutate(rows = 1:nrow(.)) %>% 
+  # filter(rows <= 10) %>% 
+  ggplot(aes(x = term, y = estimate)) +
+  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
+  geom_point(fill = "steelblue", size = 4, shape = 21, color = "black") +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  theme_classic() +
+  coord_flip()
 ```
@@
 ```
-
-### Biological
 

### 2018-01-20 Frontiers_Manuscript/Villasenor-Derbez.tex
Added initial model fit

@@
 
+\subsubsection{Biological}\label{biological}
+
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-2-1.pdf}
+
 Table \ref{table:indicators} is the way to cite a table.
@@
 
-\caption{\label{tab:unnamed-chunk-1}\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.}
+\caption{\label{tab:unnamed-chunk-3}\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.}
 \centering
@@
 \end{table}
-
-\subsubsection{Biological}\label{biological}
 

### 2018-01-20 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Fixed map and focused analysis on 3 locations

--- 2018-01-19 Frontiers_Manuscript/Villasenor-Derbez.Rmd
@@
 ---
-title: "Three archetypes of no-take marine reserves in Mexico yield positive (but idiosyncratic) social and ecological benefits"
+title: "Community-based marine reserves produce biological and economic benefits"
 rtitle: "Mexican marine reserves"
@@
 "
-corrauth: "Me, at home"
+corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
 email: "jvillasenor@bren.ucsb.edu"
-abstract: "En México, las reservas marinas han sido comúnmente establecidas como zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por la Comisión Nacional de Áreas Naturales Protegidas (CONANP).
-Al día de hoy, 36 RBs protegen una porción del ambiente marino en México.
-Sin embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas.
-Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, esto aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores.
-La escasez de recursos monetarios y humanos de la limitan también el monitoreo y vigilancia de las reservas, y a su vez, el desempeño de la reserva.
-En México, las reservas marinas han sido comúnmente establecidas como zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por la Comisión Nacional de Áreas Naturales Protegidas (CONANP).
-Al día de hoy, 36 RBs protegen una porción del ambiente marino en México.
-Sin embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas.
-Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, esto aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores.
-La escasez de recursos monetarios y humanos de la limitan también el monitoreo y vigilancia de las reservas, y a su vez, el desempeño de la reserva."
-keywords: "Marine Reserves, Marine Conservation, Fisheries, Citizen Science, Mexico. "
+abstract: ""
+keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico."
 # biblio: test.bib
@@
 })
+
+source(file = "../Scripts/robust_se.R")
+source(file = "../Scripts/abundance_model.R")
 ```
 
-```{r create_map, include = T, fig.cap = "\\label{fig:map}Mapa de la localización general de las comunidades de estudio.
-El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero."}
+```{r load_spatial}
 
+# Determine a projection for the maps
 proj <- CRS("+proj=longlat +datum=NAD27")
 
+# Load coastline for Mexico and World
 data(World)
-W <- spTransform(World, proj)
-W <- W[!W$name == "Mexico",] %>% 
+W <- spTransform(World, proj) %>% 
   st_as_sf()
 
+# Load high-res mexican coastline
 coastline_mx <- readRDS("../Data/Spatial/coastline_mx.rds")
 
-location_labels <- data.frame(X = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087),
-                              Y = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957),
-                              Location = c("Isla Natividad", "Maria Elena", "Punta Herrero", "Puerto Libertad", "Guaymas"),
-                              Region = c("Pacific", "Caribbean","Caribbean", "Gulf of California", "Gulf of California"),
+# Declare points for reference map
+location_labels <- data.frame(X = c(-115.24, -87.4611756),
+                              Y = c(27.85, 19.3183386),
+                              Location = c("Isla Natividad", "Punta Herrero"),
+                              Region = c("Pacific", "Caribbean"),
                               stringsAsFactors = F)
@@
 
-main_map <- ggplot() +
+# Load ZRP shapefile
+zrp <- st_read(dsn = "../Data/Spatial", layer = "ZRP", quiet = T) %>% 
+  filter(Name %in% c("El Cabezo", "El Faro", "Gallineros")) %>% 
+  st_transform(crs = "+proj=longlat +datum=NAD27")
+
+# Load Community reserve shapefile
+rescom <- st_read(dsn = "../Data/Spatial", layer = "Res_Com", quiet = T) %>% 
+  filter(Nombre == "La_Plana") %>% 
+  st_transform(crs = "+proj=longlat +datum=NAD27")
+
+```
+
+```{r create_map, include = T, fig.cap = "\\label{fig:map}Mapa de la localización general de las comunidades de estudio.
+El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero."}
+# Define reference map
+mex <- ggplot() +
   geom_sf(data = W) +
-  geom_sf(data = coastline_mx) +
-  geom_sf(data = location_labels, aes(color = Location, fill = Location), size = 3) +
+  geom_sf(data = location_labels, color = "steelblue", size = 3) +
   theme_classic() +
@@
   scale_fill_brewer(palette = "Set1") +
-  theme(legend.position = "none")
-
-mex <- main_map +
+  theme(legend.position = "none") +
   lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863))
 
-legends <- main_map +
-  lims(x = c(-87, -86.5), y = c(19.4, 19.6)) +
-  theme_map() +
-  theme(legend.justification = c(0.5, 0.5),
-        legend.position = c(0.5, 0.5),
-        legend.background = element_rect(fill = "transparent")) +
-  ggExtra::removeGrid() +
-  blank()
+# Map for Punta Herrero
+ph <- ggplot() +
+  geom_sf(data = coastline_mx) +
+  lims(x = c(-87.5, -87.4), y = c(19.280990, 19.35)) +
+  geom_sf(data = zrp, fill = "red") +
+  theme_classic()
 
-sam <- main_map +
-  lims(x = c(-88, -87), y = c(19.1, 19.7))
+# Map for Maria Elena
+ma <- ggplot() +
+  geom_sf(data = coastline_mx) +
+  lims(x = c(-87.54, -87.44), y = c(19.38, 19.44)) +
+  geom_sf(data = zrp, fill = "red") +
+  theme_classic()
 
-goc <- main_map +
-  lims(x = c(-115.5, -110.5), y = c(27, 30))
+# Map for Natividad
+pbc <- ggplot() +
+  geom_sf(data = coastline_mx) +
+  lims(x = c(-115.24, -115.14), y = c(27.83, 27.90)) +
+  geom_sf(data = rescom, fill = "red") +
+  theme_classic()
 
-plot_grid(mex, sam, goc, legends, labels = c("A", "C", "B"), ncol = 2)
-
+plot_grid(mex, ph, pbc, ma, labels = c("A", "B", "C", "D"), ncol = 2)
 ```
@@
 ```{r load_data}
-fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
+fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F) %>% 
+  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
+  filter(!is.na(RC),
+         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)"))
 
-invert <- read.csv("../Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
+invert <- read.csv("../Data/invert_data.csv", strip.white = T, stringsAsFactors = F) %>% 
+  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
+  filter(!is.na(RC),
+         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)"))
 
@@
 
-```{r complete_fish, eval = F}
-fish2 <- fish %>% 
+### Biological
+
+```{r, prepare_covars}
+fish_covars <- fish %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto, Comunidad) %>% 
+  count() %>% 
+  select(-n)
+
+invert_covars <- invert %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto, Comunidad) %>% 
+  count() %>% 
+  select(-n)
+```
+
+```{r complete_fish}
+fish_complete <- fish %>% 
   select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
@@
   complete(Ano, Transecto, nesting(Sitio, Zona, GeneroEspecie), fill = list(Abundancia = 0L)) %>% 
+  left_join(fish_covars, by = c("Ano", "Sitio", "Zona", "Transecto")) %>% 
+  select(Comunidad, Ano, RC, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
+  arrange() %>% 
+  filter(!is.na(RC))
+
+invert_complete <- invert %>%
   select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
+  group_by(Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  ungroup() %>% 
+  complete(Ano, Transecto, nesting(Sitio, Zona, GeneroEspecie), fill = list(Abundancia = 0L)) %>% 
+  left_join(fish_covars, by = c("Ano", "Sitio", "Zona", "Transecto")) %>% 
+  select(Comunidad, Ano, RC, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
   arrange()
+```
+
+```{r model_fit}
+fish_abundance_model <- abundance_model(fish_complete)
+
+invert_abundance_model <- abundance_model(invert_complete)
+```
+
+```{r, eval = F}
+fish_complete %>% 
+  mutate(Ano = as.factor(Ano)) %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  ungroup() %>% 
+  mutate(predictions = predict(fish_abundance_model, ., type = "response")) %>% 
+  ggplot(aes(x = predictions, y = Abundancia, color = RC)) +
+  geom_point() +
+  theme_classic() +
+  geom_abline() +
+  theme(legend.position = "none")
+
+fish_complete %>% 
+  mutate(Ano = as.factor(Ano)) %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  ungroup() %>% 
+  mutate(predictions = predict(fish_abundance_model, ., type = "response"),
+         Err = predictions - Abundancia) %>% 
+  ggplot(aes(x = Err)) +
+  geom_histogram(binwidth = 10) +
+  theme_classic()
+
+fish_complete %>% 
+  mutate(Ano = as.factor(Ano)) %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  ungroup() %>% 
+  mutate(predictions = predict(fish_abundance_model, ., type = "response"),
+         Err = predictions - Abundancia) %>% 
+  ggplot(aes(x = predictions, y = Err)) +
+  geom_point() +
+  geom_abline(slope = 0) +
+  theme_classic()
+
+glance(fish_abundance_model)
 ```
@@
 
-### Biological
-
 ### Socioeconomic
@@
 # Results
+
+```{r}
+comunidades <- fish %>% 
+  group_by(Comunidad, RC) %>% 
+  count() %>% 
+  ungroup() %>% 
+  select(-n)
+
+tidy_terms <- robust_se(fish_abundance_model) %>% 
+  filter(grepl("Ano20.*:ZonaReserva:RC", term) |
+           grepl("Ano20..:ZonaReserva", term)
+         ) %$%
+  str_split(term, pattern = ":", simplify = T) %>% 
+  as.tibble() %>% 
+  set_colnames(c("Ano", "Zona", "RC")) %>% 
+  select(-Zona) %>% 
+  mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = ""))) %>% 
+  filter(!(Ano > 2011 & RC == "")) %>% 
+  mutate(RC = gsub(RC, pattern = "RC", replacement = ""),
+         RC = ifelse(RC == "", "La Plana / Las Cuevas - La Dulce / Babencho", RC))
+
+robust_se(fish_abundance_model) %>% 
+  filter(grepl("Ano20.*:ZonaReserva:RC", term) |
+           grepl("Ano20..:ZonaReserva", term),
+         !term %in% c("Ano2012:ZonaReserva", "Ano2013:ZonaReserva", "Ano2014:ZonaReserva", "Ano2015:ZonaReserva", "Ano2016:ZonaReserva")) %>%
+  cbind(tidy_terms) %>%
+  mutate(p = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
+  left_join(comunidades, by = "RC") %>%
+  mutate(Comunidad = ifelse(is.na(Comunidad), "Isla Natividad", Comunidad)) %>%
+  ggplot(aes(x = Ano, y = estimate)) +
+  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
+  geom_line(aes(group = RC, color = RC)) +
+  geom_point(fill = "steelblue", size = 4, shape = 21, color = "black") +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  theme_classic() +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  facet_wrap(~Comunidad, ncol = 2) +
+  theme(legend.justification = c(0, 0), 
+        legend.position = c(0.55, 0.01))
+```
+
+```{r}
+comunidades <- invert %>% 
+  group_by(Comunidad, RC) %>% 
+  count() %>% 
+  ungroup() %>% 
+  select(-n)
+
+tidy_terms <- robust_se(invert_abundance_model) %>% 
+  filter(grepl("Ano20.*:ZonaReserva:RC", term) |
+           grepl("Ano20..:ZonaReserva", term)
+         ) %$%
+  str_split(term, pattern = ":", simplify = T) %>% 
+  as.tibble() %>% 
+  set_colnames(c("Ano", "Zona", "RC")) %>% 
+  select(-Zona) %>% 
+  mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = ""))) %>% 
+  filter(!(Ano > 2011 & RC == "")) %>% 
+  mutate(RC = gsub(RC, pattern = "RC", replacement = ""),
+         RC = ifelse(RC == "", "La Plana / Las Cuevas - La Dulce / Babencho", RC))
+
+robust_se(invert_abundance_model) %>% 
+  filter(grepl("Ano20.*:ZonaReserva:RC", term) |
+           grepl("Ano20..:ZonaReserva", term),
+         !term %in% c("Ano2012:ZonaReserva", "Ano2013:ZonaReserva", "Ano2014:ZonaReserva", "Ano2015:ZonaReserva", "Ano2016:ZonaReserva")) %>%
+  cbind(tidy_terms) %>%
+  mutate(p = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
+  left_join(comunidades, by = "RC") %>%
+  mutate(Comunidad = ifelse(is.na(Comunidad), "Isla Natividad", Comunidad)) %>%
+  ggplot(aes(x = Ano, y = estimate)) +
+  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
+  geom_line(aes(group = RC, color = RC)) +
+  geom_point(fill = "steelblue", size = 4, shape = 21, color = "black") +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  theme_classic() +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  facet_wrap(~Comunidad, ncol = 2) +
+  theme(legend.justification = c(0, 0), 
+        legend.position = c(0.55, 0.01))
+```
 

### 2018-01-20 Frontiers_Manuscript/Villasenor-Derbez.tex
Fixed map and focused analysis on 3 locations

--- 2018-01-19 Frontiers_Manuscript/Villasenor-Derbez.tex
@@
 % Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Me, at home}
+\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
+Santa Barbara, Santa Barbara, CA, 93106}
 
@@
 
-\title[Mexican marine reserves]{Three archetypes of no-take marine reserves in Mexico yield positive
-(but idiosyncratic) social and ecological benefits}
+\title[Mexican marine reserves]{Community-based marine reserves produce biological and economic benefits}
 
@@
 \maketitle
-
-\begin{abstract}
-
-En México, las reservas marinas han sido comúnmente establecidas como
-zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por
-la Comisión Nacional de Áreas Naturales Protegidas (CONANP). Al día de
-hoy, 36 RBs protegen una porción del ambiente marino en México. Sin
-embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde
-las actividades pesqueras están prohibidas. Aunque la CONANP ha hecho
-esfuerzos importantes por involucrar a los actores durante la
-implementación de las reservas, esto aún se caracteriza por un proceso
-descendente, el cual conlleva a la falta de cumplimiento por parte de
-los actores. La escasez de recursos monetarios y humanos de la limitan
-también el monitoreo y vigilancia de las reservas, y a su vez, el
-desempeño de la reserva. En México, las reservas marinas han sido
-comúnmente establecidas como zonas núcleo dentro de Reservas de la
-Biósfera (RBs), administradas por la Comisión Nacional de Áreas
-Naturales Protegidas (CONANP). Al día de hoy, 36 RBs protegen una
-porción del ambiente marino en México. Sin embargo, solamente 26 de
-estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras
-están prohibidas. Aunque la CONANP ha hecho esfuerzos importantes por
-involucrar a los actores durante la implementación de las reservas, esto
-aún se caracteriza por un proceso descendente, el cual conlleva a la
-falta de cumplimiento por parte de los actores. La escasez de recursos
-monetarios y humanos de la limitan también el monitoreo y vigilancia de
-las reservas, y a su vez, el desempeño de la reserva.
-
-\medskip
-\tiny
- \keyFont{ \section{Keywords:} Marine Reserves, Marine Conservation, Fisheries, Citizen Science,
-Mexico.}
-
-\end{abstract}
 
@@
 
+\subsubsection{Biological}\label{biological}
+
 Table \ref{table:indicators} is the way to cite a table.
@@
 
-\caption{\label{tab:unnamed-chunk-1}\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.}
+\caption{\label{tab:unnamed-chunk-2}\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.}
 \centering
@@
 
-\subsubsection{Biological}\label{biological}
-
 \subsubsection{Socioeconomic}\label{socioeconomic}
@@
 \section{Results}\label{results}
+
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
 

### 2018-01-20 Scripts/abundance_model.R

@@ -1,0 +1,8 @@
+abundance_model <- function(x){
+  x %>% 
+    mutate(Ano = as.factor(Ano)) %>% 
+    group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+    summarize(Abundancia = sum(Abundancia)) %>% 
+    ungroup() %>% 
+    glm(Abundancia ~ Ano * Zona * RC, data = ., family = "quasipoisson")
+}

### 2018-01-20 Scripts/robust_se.R

@@ -1,0 +1,8 @@
+robust_se <- function(model){
+  
+  RSE <- model %>%
+    lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC1")) %>%
+    broom::tidy()
+  
+  return(RSE)
+}

### 2018-01-24 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Added skeleton to the manuscript

@@
 ```
+
+___
+
+**Esqueleto:**
+
+- Introduction
+
+  - Las reservas marinas son una herramienta para X y Y, con beneficios A y B
+
+  - Existen diferencias entre reservas bottom up y top down
+
+  - En Mexico, las reservas se implementan de 3 maneras
+
+  - Los ZRP tienen X objetivos
+
+  - A X anos de la implementacion de la NOM-049, no hay estudio formal de ZRP o reservas comunitarias
+
+  - Proveer la primer evaluacion de reservas comunitarias triple bottom line
+
+- Materials and methods
+
+  - Study areas
+
+    - Main type of ecosystem
+
+    - Number of fishers (aprox)
+
+  - Data collection
+
+    - Sampling design (BACI)
+
+    - Sampling protocol (transects, richness, abundance, sizes)
+
+    - Sample sizes
+
+  - Data analysis
+
+    - Objectives and indicators
+
+    - Analysis of numeric indicators
+
+    - Analysis of descriptive indicators
+
+- Results
+
+- Discusion
+
+____
 
@@
 ```{r}
-comunidades <- fish %>% 
-  group_by(Comunidad, RC) %>% 
-  count() %>% 
+
+fish_model2 <- invert_complete %>% 
+  filter(Comunidad == "Isla Natividad") %>% 
+  mutate(Ano = as.factor(Ano)) %>% 
+  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
   ungroup() %>% 
-  select(-n)
+  glm(Abundancia ~ Ano * Zona, data = ., family = "quasipoisson")
 
-tidy_terms <- robust_se(fish_abundance_model) %>% 
-  filter(grepl("Ano20.*:ZonaReserva:RC", term) |
-           grepl("Ano20..:ZonaReserva", term)
-         ) %$%
+tidy_terms <-  fish_model2 %>% 
+  robust_se() %>%
+  filter(grepl("Ano20.*:ZonaReserva", term)) %$%
   str_split(term, pattern = ":", simplify = T) %>% 
   as.tibble() %>% 
-  set_colnames(c("Ano", "Zona", "RC")) %>% 
+  set_colnames(c("Ano", "Zona")) %>% 
   select(-Zona) %>% 
-  mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = ""))) %>% 
-  filter(!(Ano > 2011 & RC == "")) %>% 
-  mutate(RC = gsub(RC, pattern = "RC", replacement = ""),
-         RC = ifelse(RC == "", "La Plana / Las Cuevas - La Dulce / Babencho", RC))
+  mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = "")))
 
-robust_se(fish_abundance_model) %>% 
-  filter(grepl("Ano20.*:ZonaReserva:RC", term) |
-           grepl("Ano20..:ZonaReserva", term),
-         !term %in% c("Ano2012:ZonaReserva", "Ano2013:ZonaReserva", "Ano2014:ZonaReserva", "Ano2015:ZonaReserva", "Ano2016:ZonaReserva")) %>%
+robust_se(fish_model2) %>%
+  filter(grepl("Ano20.*:ZonaReserva", term)) %>% 
   cbind(tidy_terms) %>%
   mutate(p = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
-  left_join(comunidades, by = "RC") %>%
-  mutate(Comunidad = ifelse(is.na(Comunidad), "Isla Natividad", Comunidad)) %>%
   ggplot(aes(x = Ano, y = estimate)) +
   geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
-  geom_line(aes(group = RC, color = RC)) +
-  geom_point(fill = "steelblue", size = 4, shape = 21, color = "black") +
+  geom_point(aes(fill = p), size = 4, shape = 21, color = "black") +
   geom_hline(yintercept = 0, linetype = "dashed") +
@@
   scale_fill_brewer(palette = "Set1", direction = -1) +
-  facet_wrap(~Comunidad, ncol = 2) +
   theme(legend.justification = c(0, 0), 
         legend.position = c(0.55, 0.01))
+
 ```
@@
 ```{r}
-comunidades <- invert %>% 
-  group_by(Comunidad, RC) %>% 
-  count() %>% 
-  ungroup() %>% 
-  select(-n)
 
-tidy_terms <- robust_se(invert_abundance_model) %>% 
-  filter(grepl("Ano20.*:ZonaReserva:RC", term) |
-           grepl("Ano20..:ZonaReserva", term)
-         ) %$%
-  str_split(term, pattern = ":", simplify = T) %>% 
-  as.tibble() %>% 
-  set_colnames(c("Ano", "Zona", "RC")) %>% 
-  select(-Zona) %>% 
-  mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = ""))) %>% 
-  filter(!(Ano > 2011 & RC == "")) %>% 
-  mutate(RC = gsub(RC, pattern = "RC", replacement = ""),
-         RC = ifelse(RC == "", "La Plana / Las Cuevas - La Dulce / Babencho", RC))
-
-robust_se(invert_abundance_model) %>% 
-  filter(grepl("Ano20.*:ZonaReserva:RC", term) |
-           grepl("Ano20..:ZonaReserva", term),
-         !term %in% c("Ano2012:ZonaReserva", "Ano2013:ZonaReserva", "Ano2014:ZonaReserva", "Ano2015:ZonaReserva", "Ano2016:ZonaReserva")) %>%
-  cbind(tidy_terms) %>%
-  mutate(p = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
-  left_join(comunidades, by = "RC") %>%
-  mutate(Comunidad = ifelse(is.na(Comunidad), "Isla Natividad", Comunidad)) %>%
-  ggplot(aes(x = Ano, y = estimate)) +
-  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
-  geom_line(aes(group = RC, color = RC)) +
-  geom_point(fill = "steelblue", size = 4, shape = 21, color = "black") +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  theme_classic() +
-  scale_fill_brewer(palette = "Set1", direction = -1) +
-  facet_wrap(~Comunidad, ncol = 2) +
-  theme(legend.justification = c(0, 0), 
-        legend.position = c(0.55, 0.01))
 ```

### 2018-01-24 Frontiers_Manuscript/Villasenor-Derbez.tex
Added skeleton to the manuscript

@@
 \maketitle
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\textbf{Esqueleto:}
+
+\begin{itemize}
+\item
+  Introduction
+\item
+  Las reservas marinas son una herramienta para X y Y, con beneficios A
+  y B
+\item
+  Existen diferencias entre reservas bottom up y top down
+\item
+  En Mexico, las reservas se implementan de 3 maneras
+\item
+  Los ZRP tienen X objetivos
+\item
+  A X anos de la implementacion de la NOM-049, no hay estudio formal de
+  ZRP o reservas comunitarias
+\item
+  Proveer la primer evaluacion de reservas comunitarias triple bottom
+  line
+\item
+  Materials and methods
+\item
+  Study areas
+
+  \begin{itemize}
+  \item
+    Main type of ecosystem
+  \item
+    Number of fishers (aprox)
+  \end{itemize}
+\item
+  Data collection
+
+  \begin{itemize}
+  \item
+    Sampling design (BACI)
+  \item
+    Sampling protocol (transects, richness, abundance, sizes)
+  \item
+    Sample sizes
+  \end{itemize}
+\item
+  Data analysis
+
+  \begin{itemize}
+  \item
+    Objectives and indicators
+  \item
+    Analysis of numeric indicators
+  \item
+    Analysis of descriptive indicators
+  \end{itemize}
+\item
+  Results
+\item
+  Discusion
+\end{itemize}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
@@
 
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
-
 \section{Discussion}\label{discussion}

### 2018-01-24 Scripts/abundance_model.R

@@ -4,5 +4,5 @@ abundance_model <- function(x){
     group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
     summarize(Abundancia = sum(Abundancia)) %>% 
     ungroup() %>% 
-    glm(Abundancia ~ Ano * Zona * RC, data = ., family = "quasipoisson")
+    glm(Abundancia ~ Ano * Zona, data = ., family = "quasipoisson")
 }

### 2018-01-26 Scripts/model.R
Started working towards #2

@@ -1,0 +1,12 @@
+model <- function(x, com, family = "gaussian", sfe = T){
+  data <- x %>%  
+    dplyr::filter(Comunidad == com)
+  
+  if(sfe){
+    glm(Abundancia ~ Ano * Zona + Sitio -1, data = data, family = family)
+  }
+  
+  if(!sfe){
+    glm(Abundancia ~ Ano * Zona -1, data = data, family = family)
+  }
+}

### 2018-01-26 Scripts/abundance_model.R
Deleted old abundance_model fit

@@ -1,8 +1,0 @@
-abundance_model <- function(x){
-  x %>% 
-    mutate(Ano = as.factor(Ano)) %>% 
-    group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
-    summarize(Abundancia = sum(Abundancia)) %>% 
-    ungroup() %>% 
-    glm(Abundancia ~ Ano * Zona, data = ., family = "quasipoisson")
-}

### 2018-01-26 Scripts/plot_model.R
Started working in plotting function towards #1

@@ -1,0 +1,75 @@
+plot_model <- function(model, various = F, model_names = NULL){
+  
+  if(various){
+    
+    if(is.null(model_names)){model_names <- letters[1:length(model)]}
+    
+    p_all <- data.frame(term = NULL,
+                        estimate = NULL,
+                        std.error = NULL,
+                        statistic = NULL,
+                        p.value = NULL,
+                        model = NULL,
+                        stringsAsFactors = F)
+    
+    for(i in 1:length(model)){
+      tidy_terms <- model[[i]] %>% 
+        robust_se() %>%
+        filter(grepl(":ZonaReserva", term)) %$%
+        str_split(term, pattern = ":", simplify = T) %>% 
+        as.tibble() %>% 
+        set_colnames(c("Ano", "Zona")) %>% 
+        select(-Zona) %>% 
+        mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = "")),
+               model = model_names[i])
+      
+      p_i <- robust_se(model[[i]]) %>%
+        filter(grepl(":ZonaReserva", term)) %>% 
+        cbind(tidy_terms) %>%
+        mutate(p = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05"))
+      
+      p_all <- rbind(p_all,  p_i)
+    }
+    
+    p <- p_all %>% 
+      ggplot(aes(x = Ano, y = estimate)) +
+      geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
+      geom_point(aes(shape = model), size = 4) +
+      geom_hline(yintercept = 0, linetype = "dashed") +
+      scale_fill_brewer(palette = "Set1") +
+      theme(legend.justification = c(0, 1),
+            legend.position = c(0, 1)) +
+      labs(x = "Year", y = "Effect size") +
+      geom_line(aes(group = model))
+    
+  }
+  
+  if(!various){
+    tidy_terms <- model %>% 
+      robust_se() %>%
+      filter(grepl(":ZonaReserva", term)) %$%
+      str_split(term, pattern = ":", simplify = T) %>% 
+      as.tibble() %>% 
+      set_colnames(c("Ano", "Zona")) %>% 
+      select(-Zona) %>% 
+      mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = "")))
+    
[…]

### 2018-01-26 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Started streamlining analyses in the general manuscript

@@
 source(file = "../Scripts/robust_se.R")
-source(file = "../Scripts/abundance_model.R")
+source(file = "../Scripts/model.R")
+source(file = "../Scripts/plot_model.R")
 ```
@@
   filter(!is.na(RC),
-         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)"))
+         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)", "El Faro - El Faro (Control)", "Anegado de Chal - Anegado de Chal (Control)"))
 
@@
 ```{r model_fit}
-fish_abundance_model <- abundance_model(fish_complete)
+fish_abundance_model <- fish_complete %>% 
+  mutate(Ano = as.factor(Ano)) %>% 
+  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  model(com = "Isla Natividad", sfe = F, family = "quasipoisson")
 
-invert_abundance_model <- abundance_model(invert_complete)
+# invert_abundance_model <- abundance_model(invert_complete)
 ```
@@
 ```{r}
+fish_complete %>% 
+  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>%
+  filter(Comunidad == "Punta Herrero") %>% 
+  group_by(Zona, Ano) %>% 
+  summarize(Abundancia = mean(Abundancia)) %>% 
+  ggplot(aes(x = Ano, y = Abundancia, color = Zona)) +
+  geom_point() +
+  geom_line()
+```
 
-fish_model2 <- invert_complete %>% 
-  filter(Comunidad == "Isla Natividad") %>% 
+```{r}
+fish_complete %>% 
+  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>%
+  filter(Comunidad == "Punta Herrero") %>% 
+  group_by(Zona, Ano) %>% 
+  summarize(Abundancia = mean(Abundancia)) %>% 
+  spread(Zona, Abundancia) %>%
+  mutate(Abundancia = c(183.94, Reserva[-1] - Control[-1] - 183.94)) %>%
+  ggplot(aes(x = Ano, y = Abundancia)) +
+  geom_point() +
+  geom_line()
+```
+
+```{r}
+com <- "Maria Elena"
+
+jc3 <- fish_complete %>% 
   mutate(Ano = as.factor(Ano)) %>% 
-  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
   summarize(Abundancia = sum(Abundancia)) %>% 
-  ungroup() %>% 
-  glm(Abundancia ~ Ano * Zona, data = ., family = "quasipoisson")
+  filter(Comunidad == com) %>%
+  lm(Abundancia ~ Ano * Zona + GeneroEspecie, data = .)
 
-tidy_terms <-  fish_model2 %>% 
-  robust_se() %>%
-  filter(grepl("Ano20.*:ZonaReserva", term)) %$%
-  str_split(term, pattern = ":", simplify = T) %>% 
-  as.tibble() %>% 
-  set_colnames(c("Ano", "Zona")) %>% 
-  select(-Zona) %>% 
-  mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = "")))
+fish_complete %>% 
+  mutate(Ano = as.factor(Ano)) %>% 
+  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto) %>% 
+  summarize(Abundancia = sum(Abundancia)) %>% 
+  filter(Comunidad == com) %>%
+  lm(Abundancia ~ Ano * Zona, data = .) %>% #summary()
+  plot_model()
 
-robust_se(fish_model2) %>%
-  filter(grepl("Ano20.*:ZonaReserva", term)) %>% 
-  cbind(tidy_terms) %>%
-  mutate(p = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05")) %>% 
-  ggplot(aes(x = Ano, y = estimate)) +
-  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
-  geom_point(aes(fill = p), size = 4, shape = 21, color = "black") +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  theme_classic() +
-  scale_fill_brewer(palette = "Set1", direction = -1) +
-  theme(legend.justification = c(0, 0), 
-        legend.position = c(0.55, 0.01))
+```
+
+```{r}
 

### 2018-01-27 Scripts/plot_model.R
Plotting function done, close #1

@@ -1,4 +1,4 @@
-plot_model <- function(model, various = F, model_names = NULL){
+plot_model <- function(model, various = F, model_names = NULL, legend = F){
   
   if(various){
     
@@ -21,7 +21,7 @@ plot_model <- function(model, various = F, model_names = NULL){
         set_colnames(c("Ano", "Zona")) %>% 
         select(-Zona) %>% 
         mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = "")),
-               model = model_names[i])
+               Community = model_names[i])
       
       p_i <- robust_se(model[[i]]) %>%
         filter(grepl(":ZonaReserva", term)) %>% 
@@ -31,16 +31,19 @@ plot_model <- function(model, various = F, model_names = NULL){
       p_all <- rbind(p_all,  p_i)
     }
     
+    pd <- position_dodge(width = 0.4)
+    
     p <- p_all %>% 
-      ggplot(aes(x = Ano, y = estimate)) +
-      geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, color = "red", size = 1) +
-      geom_point(aes(shape = model), size = 4) +
+      ggplot(aes(x = Ano, y = estimate, color = Community)) +
+      geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, size = 1, position = pd) +
+      geom_point(aes(shape = Community), size = 4, position = pd, alpha = 0.5) +
       geom_hline(yintercept = 0, linetype = "dashed") +
       scale_fill_brewer(palette = "Set1") +
       theme(legend.justification = c(0, 1),
-            legend.position = c(0, 1)) +
+            legend.position = c(0, 1),
+            text = element_text(size = 12)) +
       labs(x = "Year", y = "Effect size") +
-      geom_line(aes(group = model))
+      scale_color_brewer(palette = "Set1")
     
   }
   
@@ -66,6 +69,11 @@ plot_model <- function(model, various = F, model_names = NULL){
       theme(legend.justification = c(1, 1), 
             legend.position = c(1, 1)) +
       labs(x = "Year", y = "Effect size")
+  }
+  
+  if(!legend){
+    p <- p +
+      theme(legend.position = "None")
   }
   
   return(p)

### 2018-01-27 Scripts/model.R
Finalized model fitting function, closes #2

@@ -1,12 +1,16 @@
-model <- function(x, com, family = "gaussian", sfe = T){
-  data <- x %>%  
-    dplyr::filter(Comunidad == com)
+model <- function(x, family = "gaussian", sfe = T, spfe = T){
+  
+  # x %<>% filter(Comunidad == com)
   
   if(sfe){
-    glm(Abundancia ~ Ano * Zona + Sitio -1, data = data, family = family)
+    if(spfe){model <- glm(Indicador ~ Ano * Zona + Sitio + GeneroEspecie, data = x, family = family)} #Site FE and Spp FE
+    if(!spfe){model <- glm(Indicador ~ Ano * Zona + Sitio, data = x, family = family)} #Site FE, no Spp FE
   }
   
   if(!sfe){
-    glm(Abundancia ~ Ano * Zona -1, data = data, family = family)
+    if(spfe){model <- glm(Indicador ~ Ano * Zona + GeneroEspecie, data = x, family = family)} #No site FE and Spp FE
+    if(!spfe){model <- glm(Indicador ~ Ano * Zona, data = x, family = family)} #No site and no Spp FE
   }
+  
+  return(model)
 }

### 2018-01-27 Scripts/abundance.R
Created first three indicator functions for #3

@@ -1,0 +1,14 @@
+abundance <- function(x, com, spp = NULL){
+  
+  if(!is.null(spp)){
+    x %<>% filter(GeneroEspecie == spp)
+  }
+  
+  results <- x %>% 
+    filter(Comunidad == com) %>%
+    mutate(Ano = as.factor(Ano)) %>% 
+    group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
+    summarize(Indicador = sum(Abundancia))
+    
+    return(results)
+}

### 2018-01-27 Scripts/richness.R

@@ -1,0 +1,10 @@
+richness <- function(x, com){
+  results <- x %>% 
+    filter(Comunidad == com) %>%
+    filter(Abundancia > 0) %>%
+    mutate(Ano = as.factor(Ano)) %>% 
+    group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+    summarize(Indicador = n())
+  
+  return(results)
+}

### 2018-01-27 Scripts/shannon.R

@@ -1,0 +1,38 @@
+shannon <- function(x, com){
+  N <- x %>%
+    filter(Comunidad == com) %>%
+    filter(Abundancia > 0) %>%
+    mutate(Ano = as.factor(Ano)) %>%
+    group_by(Ano,
+             Zona,
+             Sitio,
+             Transecto) %>%
+    summarize(N = sum(Abundancia)) %>%
+    ungroup() %>%
+    mutate(ID = paste(Ano, Zona, Sitio, Transecto)) %>%
+    select(ID, N)
+
+  H <- x %>%
+    filter(Comunidad == com) %>%
+    filter(Abundancia > 0) %>%
+    mutate(Ano = as.factor(Ano)) %>%
+    group_by(Ano,
+             Zona,
+             Sitio,
+             Transecto,
+             GeneroEspecie) %>%
+    summarize(ni = sum(Abundancia)) %>%
+    ungroup() %>%
+    mutate(ID = paste(Ano, Zona, Sitio, Transecto)) %>%
+    left_join(N, by = "ID") %>%
+    mutate(pi = ni/N) %>%
+    group_by(Ano,
+             Zona,
+             Sitio,
+             Transecto) %>%
+    summarize(H = -1*sum(pi*log2(pi))) %>%
+    ungroup() %>%
+    select(Ano, Zona, Sitio, Transecto, Indicador = H)
+
+  return(as.data.frame(H))
+}

### 2018-01-27 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Updated manuscript

@@
 
-___
+<!-- **Esqueleto:** -->
 
-**Esqueleto:**
+<!-- - Introduction -->
 
-- Introduction
+<!--   - Las reservas marinas son una herramienta para X y Y, con beneficios A y B -->
 
-  - Las reservas marinas son una herramienta para X y Y, con beneficios A y B
+<!--   - Existen diferencias entre reservas bottom up y top down -->
 
-  - Existen diferencias entre reservas bottom up y top down
+<!--   - En Mexico, las reservas se implementan de 3 maneras -->
 
-  - En Mexico, las reservas se implementan de 3 maneras
+<!--   - Los ZRP tienen X objetivos -->
 
-  - Los ZRP tienen X objetivos
+<!--   - A X anos de la implementacion de la NOM-049, no hay estudio formal de ZRP o reservas comunitarias -->
 
-  - A X anos de la implementacion de la NOM-049, no hay estudio formal de ZRP o reservas comunitarias
+<!--   - Proveer la primer evaluacion de reservas comunitarias triple bottom line -->
 
-  - Proveer la primer evaluacion de reservas comunitarias triple bottom line
+<!-- - Materials and methods -->
 
-- Materials and methods
+<!--   - Study areas -->
 
-  - Study areas
+<!--     - Main type of ecosystem -->
 
-    - Main type of ecosystem
+<!--     - Number of fishers (aprox) -->
 
-    - Number of fishers (aprox)
+<!--   - Data collection -->
 
-  - Data collection
+<!--     - Sampling design (BACI) -->
 
-    - Sampling design (BACI)
+<!--     - Sampling protocol (transects, richness, abundance, sizes) -->
 
-    - Sampling protocol (transects, richness, abundance, sizes)
+<!--     - Sample sizes -->
 
-    - Sample sizes
+<!--   - Data analysis -->
 
-  - Data analysis
+<!--     - Objectives and indicators -->
 
-    - Objectives and indicators
+<!--     - Analysis of numeric indicators -->
 
-    - Analysis of numeric indicators
+<!--     - Analysis of descriptive indicators -->
 
-    - Analysis of descriptive indicators
+<!-- - Results -->
 
-- Results
-
-- Discusion
-
-____
+<!-- - Discusion -->
 
@@
 
-Succinct, with no subheadings.
+<!-- Succinct, with no subheadings. -->
 
-Rationale
-Objectives
-Research question
+<!-- Rationale -->
+<!-- Objectives -->
+<!-- Research question -->
 
@@
 
+El objetivo de este trabajo este trabajo es realizar una evaluación de la efectividad de reservas marinas en México, presentando resultados de cinco comunidades costeras como caso de estudio.
+Con el fin de obtener una visión holística del sistema, la evaluación se realizará tomando en cuenta indicadores biológicos, socioeconómicos y de gobernanza.
+La evaluación de éstos cinco casos de estudios nos permitirá identificar la manera en que las características socioeconómicas y de gobernanza se relacionan con la efectividad (biológica) de las reservas marinas evaluadas.
+Los patrones identificados podrán utilizarse para informar la toma de decisiones para la implementación de la red de reservas marinas en la Región de las Grandes Islas del Golfo de California.
+
 # Materials and Methods
 
-This section may be divided by subheadings.
+<!-- This section may be divided by subheadings.
 This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated.
-For experiments reporting results on animal or human subject research, an ethics approval statement should be included in this section (for further information, see section Materials and Data Policies)
+For experiments reporting results on animal or human subject research, an ethics approval statement should be included in this section (for further information, see section Materials and Data Policies) -->
 
@@
 
-... a map of the general location of the study sites is shown in figure \ref{fig:map}.
+Las comunidades utilizadas en este reporte se distribuyen a lo largo de la costa Pacífica de Baja California (n = 1) y el Sistema Arrecifal Mesoamericano (n = 2; Fig \ref{fig:map}).
+
+### Isla Natividad
+
+La Isla Natividad se encuentra en la costa oeste de la Península de Baja California, donde el hábitat predominante es el bosque de kelp o sargazo gigante (*Macrocystis pyrifera*) y los arrecifes rocosos.
+En la isla, la Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California SCL realiza actividades de extracción de los recursos marinos.
+Aunque la langosta roja (*Panulirus interruptus*) es la especie más importante en términos económicos, otras especies importantes incluyen la escama (con un enfoque en Jurel; *Seriola lalandi*), el pepino de mar (*Parastichopus parvimensis*), el erizo rojo (*Mesocentrotus franciscanus*), el caracol (*Megastraea turbanica* y *M.
+undosa*) y, hasta el 2010, el abulón (*Haliotis sp.*).
+En 2006, por medio de un proceso participativo, la cooperativa decidió establecer dos reservas marinas de manera voluntaria.
+Agentes externos a la cooperativa, como personal de Comunidad y Biodiversidad A.C., académicos de la Universidad de Stanford, y personal de la CONANP (de la oficina de Reserva de la Biósfera El Vizcaíno), también participaron en el diseño e implementación de las reservas.
+Las reservas fueron establecidas como instrumento de manejo pesquero, buscando recuperar las poblaciones de abulón y otros invertebrados.
+Al día de hoy, las reservas marinas de Isla Natividad no han recibido reconocimiento legal, pero la cooperativa ha mostrado interés por reconocerlas como Zonas de Refugio Pesquero (ZRP).
+Los pescadores tienen un sistema de turnos para vigilar la reserva día y noche desde embarcaciones patrulla.
+
+### María Elena
+
+María Elena es una comunidad pesquera en la costa de Quintana Roo.
+Los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona.
+El campo pesquero es utilizado por pescadores de la SCPP Cozumel scl (de la Isla de Cozumel).
+La principal especie aprovechada por ésta organización es la langosta espinosa del caribe (*Panulirus argus*).
+La cooperativa cuenta con permiso de pesca de escama y concesión de langosta.
+En el 2012, la Cooperativa, en conjunto con la Alianza Kanan Kay, COBI, CONANP, CONAPESCA, Oceanus, Fundación Claudia y Roberto Hernández, Fundación Haciendas del Mundo Maya, establecieron ocho ZRP con una vigencia de cinco años.
+La vigilancia de las reservas se realiza por medio del equipo de vigilancia comunitaria, con apoyo de la CONANP y una embarcación -donada por COBI- utilizada para realizar recorridos frecuentes.
+
+### Punta Herrero
+
+La comunidad de Punta Herrero se encuentra aproximadamente a 15 km al sur del campo pesquero de María Elena.
+De igual manera, los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona, y la principal especie explotada es la langosta.
+Sin embargo, la SCPP José María Azcorra también cuenta con permisos para pesca de escama y tiburón y una concesión de langosta.
+En una réplica del ejercicio realizado en María Elena -con presencia de los mismos actores-, cuatro ZRP fueron establecidas en el 2013, con una vigencia de cinco años.
+El equipo de vigilancia comunitaria, con apoyo de la CONANP, se encarga de la vigilancia de las reservas.
 
@@
 
-source(file = "../Scripts/robust_se.R")
-source(file = "../Scripts/model.R")
-source(file = "../Scripts/plot_model.R")
+# Load indicator functions
+source(file = "../Scripts/abundance.R")
+source(file = "../Scripts/richness.R")
+source(file = "../Scripts/shannon.R")
+
+# Load analysis functions
+source(file = "../Scripts/model.R") #fits the model
+source(file = "../Scripts/robust_se.R") #calculates heteroskedastic-robust Standard Errors
+source(file = "../Scripts/plot_model.R") #plots effect sizes
 ```
@@
 
-```{r create_map, include = T, fig.cap = "\\label{fig:map}Mapa de la localización general de las comunidades de estudio.
-El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero."}
+```{r, fig.cap = "\\label{fig:map}Mapa de la localización general de las comunidades de estudio. El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero."}
 # Define reference map
@@
 
+Para evaluar las reservas, utilizamos tres fuentes de información.
+La información ecológica proviene de los monitoreos ecológicos realizados anualmente en las zonas reserva y control.
+Cada año, se realizan censos visuales para evaluar las comunidades de peces e invertebrados, registrando riquezas, abundancias y tallas (en peces).
+Esta información nos permite calcular los indicadores biológicos de manera anual.
+Al tener valores de diferentes indicadores biológicos antes y después de la implementación de las reservas, para las zonas de reserva y sitios control, tenemos un diseño muestreal de Antes-Después-Control-Impacto.
+
+También incluimos información socioeconómica relevante proveniente de los avisos de arribo de CONAPESCA.
+En este caso, se tienen registros mensuales de los recursos aprovechados por las diferentes comunidades, en los que se reportan los arribos (Toneladas) y el valor de los arribos ($).
+La información se encuentra disponible para el periodo 2001 - 2014.
+Ya que las categorías registradas por CONAPESCA son amplias y existe un nivel de error, utilizamos únicamente los arribos reportados para langosta entera fresca a nivel de cooperativa.
+Los ingresos generados por arribos son ajustados por medio del índice de precio al consumidor.
+
+La información de gobernanza fue obtenida a nivel de comunidad, pidiendo a personas familiares con las comunidades que proveyeran la información necesaria.
+La información de gobernanza no es evaluada de manera cuantitativa.
+En su lugar, esta se interpreta de manera tal que podamos comprender qué decisiones, reglas y estructuras tienen un impacto en la reserva.
+
 ## Data analysis
+
+Dada la similitud de objetivos entre las reservas, la evaluación se realiza con los mismos indicadores.
+En este caso, se utilizan 6 indicadores biológicos, 2 socioeconómicos y 5 de gobernanza (Tabla \ref{table:indicators}).
+Según la disponibilidad de datos, se calculó la densidad de las especies objetivo presentadas en la sección de descripción de las comunidades.
+El criterio de selección fue que cada especie debía tener, por lo menos, dos observaciones anuales para las zonas de reserva y control.
+
+```{r}
+data.frame(Category = c("Biological", "","","","","", "Socioeconomic", "", "Governance", "", "", "", ""),
+           Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Densidad de especies objetivo", "Ingresos por especies objetivo", "Arribos de especies objetivo", "Tipo de acceso a la pesquería", "Grado de pesca ilegal", "Procuración de la reserva", "Tipo de organización pesquera", "Edad de la reserva"), stringsAsFactors = F) %>%
+  knitr::kable(caption = "\\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.", format = "latex")
+```
 
@@
 ### Biological
+
+Utilizando un análisis de diferencia en diferencias podemos estimar el efecto que la reserva tienen en los indicadores biológicos \citep{moland_2013-VP} con el uso de un modelo de regresión lineal múltiple:
+
+$$I = \beta_0 + \sum \gamma Year + \beta_1 Zona + \sum \lambda Year\times Zona + \Omega + \epsilon$$
+
+En este caso, modelamos los años como factores, tomando como referencia el primer año en la serie de datos de cada comunidad.
+Modelar los años como factores reduce la estructura del modelo, y relaja el ajuste al no asumir una tendencia lineal entre años; es decir, el cambio observado entre 2006 - 2007 no deberá de ser igual al observado entre el 2009 - 2010.
+Incluimos también un término para la zona, en la que la variable toma un valor de 0 si el sitio es una zona control y de 1 si es una zona de reserva.
+Finalmente, incluimos un término de interacción entre la variable de Zona y el Año.
+En este modelo, $\lambda_i$ representa el efecto que la reservas tuvo sobre un indicador en cada año y con respecto a los sitios control.
+El termino $\Omega$ captura efectos fijos por especies y por sitio.
 
@@
 ```{r model_fit}
-fish_abundance_model <- fish_complete %>% 
-  mutate(Ano = as.factor(Ano)) %>% 
-  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>% 
-  model(com = "Isla Natividad", sfe = F, family = "quasipoisson")
 
-# invert_abundance_model <- abundance_model(invert_complete)
-```
+# Fish
 
-```{r, eval = F}
-fish_complete %>% 
-  mutate(Ano = as.factor(Ano)) %>% 
-  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>% 
-  ungroup() %>% 
-  mutate(predictions = predict(fish_abundance_model, ., type = "response")) %>% 
-  ggplot(aes(x = predictions, y = Abundancia, color = RC)) +
-  geom_point() +
-  theme_classic() +
-  geom_abline() +
-  theme(legend.position = "none")
+## Abundance
+N_IN <- fish_complete %>% 
+  abundance(com = "Isla Natividad") %>% 
+  model(sfe = F)
 
-fish_complete %>% 
-  mutate(Ano = as.factor(Ano)) %>% 
-  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>% 
-  ungroup() %>% 
-  mutate(predictions = predict(fish_abundance_model, ., type = "response"),
-         Err = predictions - Abundancia) %>% 
-  ggplot(aes(x = Err)) +
-  geom_histogram(binwidth = 10) +
-  theme_classic()
+N_ME <- fish_complete %>% 
+  abundance(com = "Maria Elena") %>% 
+  model(sfe = T)
 
-fish_complete %>% 
-  mutate(Ano = as.factor(Ano)) %>% 
-  group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>% 
-  ungroup() %>% 
-  mutate(predictions = predict(fish_abundance_model, ., type = "response"),
-         Err = predictions - Abundancia) %>% 
-  ggplot(aes(x = predictions, y = Err)) +
-  geom_point() +
-  geom_abline(slope = 0) +
-  theme_classic()
+N_PH <- fish_complete %>% 
+  abundance(com = "Punta Herrero") %>% 
+  model(sfe = F)
 
-glance(fish_abundance_model)
-```
+## Richness
+S_IN <- fish_complete %>% 
+  richness(com = "Isla Natividad") %>% 
+  model(sfe = F, spfe = F)
 
-Table \ref{table:indicators} is the way to cite a table.
+S_ME <- fish_complete %>% 
+  richness(com = "Maria Elena") %>% 
+  model(sfe = T, spfe = F)
 
-```{r}
-data.frame(Category = c("Biological", "Biological","Biological","Biological","Biological","Biological", "Socioeconomic", "Socioeconomic", "Governance", "Governance", "Governance", "Governance", "Governance"),
-           Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Densidad de especies objetivo", "Ingresos por especies objetivo", "Arribos de especies objetivo", "Tipo de acceso a la pesquería", "Grado de pesca ilegal", "Procuración de la reserva", "Tipo de organización pesquera", "Edad de la reserva"), stringsAsFactors = F) %>%
-  knitr::kable(caption = "\\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.")
+S_PH <- fish_complete %>% 
+  richness(com = "Punta Herrero") %>% 
+  model(sfe = F, spfe = F)
+
+## Shannon
+H_IN <- fish_complete %>% 
+  shannon(com = "Isla Natividad") %>% 
+  model(sfe = F, spfe = F)
+
+H_ME <- fish_complete %>% 
+  shannon(com = "Maria Elena") %>% 
+  model(sfe = T, spfe = F)
+
+H_PH <- fish_complete %>% 
+  shannon(com = "Punta Herrero") %>% 
+  model(sfe = F, spfe = F)
+
+# Invertebrates
+
+## Abundance
+Ni_IN <- invert_complete %>% 
+  abundance(com = "Isla Natividad") %>% 
+  model(sfe = F)
+
+Ni_ME <- invert_complete %>% 
+  abundance(com = "Maria Elena") %>% 
+  model(sfe = T)
+
+Ni_PH <- invert_complete %>% 
+  abundance(com = "Punta Herrero") %>% 
+  model(sfe = F)
+
+## Richness
+Si_IN <- invert_complete %>% 
+  richness(com = "Isla Natividad") %>% 
+  model(sfe = F, spfe = F)
+
+Si_ME <- invert_complete %>% 
+  richness(com = "Maria Elena") %>% 
+  model(sfe = T, spfe = F)
+
+Si_PH <- invert_complete %>% 
+  richness(com = "Punta Herrero") %>% 
+  model(sfe = F, spfe = F)
+
+## Shannon
+Hi_IN <- invert_complete %>% 
+  shannon(com = "Isla Natividad") %>% 
+  model(sfe = F, spfe = F)
+
+Hi_ME <- invert_complete %>% 
+  shannon(com = "Maria Elena") %>% 
+  model(sfe = T, spfe = F)
+
+Hi_PH <- invert_complete %>% 
+  shannon(com = "Punta Herrero") %>% 
+  model(sfe = F, spfe = F)
 ```
@@
 ### Socioeconomic
+
+El análisis de datos socioeconómicos se aplicó únicamente a Isla Natividad, María Elena y Punta Herrero, siguiendo un modelo con la forma:
+
+$$I = \beta_0 + \beta_1Post$$
+
+Que nos permite comparar el cambio en el promedio de los indicadores antes (Post = 0) y después (Post = 1) de la implementación de la reserva.
+Tanto para los indicadores biológicos como los socioeconómicos, los coeficientes fueron ajustador con el estimador de muestras heterocedásticas.
 
@@
 
-```{r}
-fish_complete %>% 
-  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>%
-  filter(Comunidad == "Punta Herrero") %>% 
-  group_by(Zona, Ano) %>% 
-  summarize(Abundancia = mean(Abundancia)) %>% 
-  ggplot(aes(x = Ano, y = Abundancia, color = Zona)) +
-  geom_point() +
-  geom_line()
-```
+A continuación se presentan los resultados de cada una de las comunidades.
+Los resultados biológicos se presentarán para cada comunidad, discutiendo primero los indicadores en común con otras comunidades (Shannon, Riqueza, Densidad, Nivel Trófico, Biomasa para peces e invertebrados) y, según su caso, las densidades de las especies objetivo.
+Habiendo presentado los resultados biológicos, nos enfocaremos después en los socioeconómicos y de gobernanza.
+Usaremos esta información para identificar las causas (sociales) del éxito (biológico) de las reservas.
 
-```{r}
-fish_complete %>% 
-  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>%
-  filter(Comunidad == "Punta Herrero") %>% 
-  group_by(Zona, Ano) %>% 
-  summarize(Abundancia = mean(Abundancia)) %>% 
-  spread(Zona, Abundancia) %>%
-  mutate(Abundancia = c(183.94, Reserva[-1] - Control[-1] - 183.94)) %>%
-  ggplot(aes(x = Ano, y = Abundancia)) +
-  geom_point() +
-  geom_line()
-```
+```{r, fig.width = 6, fig.height = 8, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares).
+Plots are ordered by survey type (left: fish; right: invertebrates) and indicators: Abundance (A, B), Richness (C, D), and Shannon's diversity index (E, F)"}
 
-```{r}
-com <- "Maria Elena"
+# Fish plots
+F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
+                  various = T,
+                  model_names = c("IN", "ME", "PH"),
+                  legend = T)
 
-jc3 <- fish_complete %>% 
-  mutate(Ano = as.factor(Ano)) %>% 
-  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>% 
-  filter(Comunidad == com) %>%
-  lm(Abundancia ~ Ano * Zona + GeneroEspecie, data = .)
+F_S <- plot_model(model = list(S_IN, S_ME, S_PH),
+                  various = T,
+                  model_names = c("IN", "ME", "PH"))
 
-fish_complete %>% 
-  mutate(Ano = as.factor(Ano)) %>% 
-  group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>% 
-  filter(Comunidad == com) %>%
-  lm(Abundancia ~ Ano * Zona, data = .) %>% #summary()
-  plot_model()
+F_H <- plot_model(model = list(H_IN, H_ME, H_PH),
+                  various = T,
+                  model_names = c("IN", "ME", "PH"))
+# Invert plots
+I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
+                  various = T,
+                  model_names = c("IN", "ME", "PH"))
 
-```
+I_S <- plot_model(model = list(Si_IN, Si_ME, Si_PH),
+                  various = T,
+                  model_names = c("IN", "ME", "PH"))
 
-```{r}
+I_H <- plot_model(model = list(Hi_IN, Hi_ME, Hi_PH),
+                  various = T,
+                  model_names = c("IN", "ME", "PH"))
 
-```
-
-```{r}
-
+cowplot::plot_grid(plotlist = list(F_N, I_N, F_S, I_S, F_H, I_S),
+                   labels = "AUTO",
+                   ncol = 2)
 ```
@@
 
-This section may be divided by subheadings.
+<!-- This section may be divided by subheadings.
 Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.
+-->
 
-Summary of main findings
-Limitations
-Conclusions
+<!-- Summary of main findings -->
+<!-- Limitations -->
+<!-- Conclusions -->
 

### 2018-01-27 Frontiers_Manuscript/Villasenor-Derbez.tex
Updated manuscript

--- 2018-01-20 Frontiers_Manuscript/Villasenor-Derbez.tex
@@
 \section{Introduction}\label{introduction}
-
-Succinct, with no subheadings.
-
-Rationale Objectives Research question
 
@@
 
+El objetivo de este trabajo este trabajo es realizar una evaluación de
+la efectividad de reservas marinas en México, presentando resultados de
+cinco comunidades costeras como caso de estudio. Con el fin de obtener
+una visión holística del sistema, la evaluación se realizará tomando en
+cuenta indicadores biológicos, socioeconómicos y de gobernanza. La
+evaluación de éstos cinco casos de estudios nos permitirá identificar la
+manera en que las características socioeconómicas y de gobernanza se
+relacionan con la efectividad (biológica) de las reservas marinas
+evaluadas. Los patrones identificados podrán utilizarse para informar la
+toma de decisiones para la implementación de la red de reservas marinas
+en la Región de las Grandes Islas del Golfo de California.
+
 \section{Materials and Methods}\label{materials-and-methods}
-
-This section may be divided by subheadings. This section should contain
-sufficient detail so that when read in conjunction with cited
-references, all procedures can be repeated. For experiments reporting
-results on animal or human subject research, an ethics approval
-statement should be included in this section (for further information,
-see section Materials and Data Policies)
 
@@
 
-\ldots{} a map of the general location of the study sites is shown in
-figure \ref{fig:map}.
+Las comunidades utilizadas en este reporte se distribuyen a lo largo de
+la costa Pacífica de Baja California (n = 1) y el Sistema Arrecifal
+Mesoamericano (n = 2; Fig \ref{fig:map}).
+
+\subsubsection{Isla Natividad}\label{isla-natividad}
+
+La Isla Natividad se encuentra en la costa oeste de la Península de Baja
+California, donde el hábitat predominante es el bosque de kelp o sargazo
+gigante (\emph{Macrocystis pyrifera}) y los arrecifes rocosos. En la
+isla, la Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y
+Pescadores de la Baja California SCL realiza actividades de extracción
+de los recursos marinos. Aunque la langosta roja (\emph{Panulirus
+interruptus}) es la especie más importante en términos económicos, otras
+especies importantes incluyen la escama (con un enfoque en Jurel;
+\emph{Seriola lalandi}), el pepino de mar (\emph{Parastichopus
+parvimensis}), el erizo rojo (\emph{Mesocentrotus franciscanus}), el
+caracol (\emph{Megastraea turbanica} y \emph{M. undosa}) y, hasta el
+2010, el abulón (\emph{Haliotis sp.}). En 2006, por medio de un proceso
+participativo, la cooperativa decidió establecer dos reservas marinas de
+manera voluntaria. Agentes externos a la cooperativa, como personal de
+Comunidad y Biodiversidad A.C., académicos de la Universidad de
+Stanford, y personal de la CONANP (de la oficina de Reserva de la
+Biósfera El Vizcaíno), también participaron en el diseño e
+implementación de las reservas. Las reservas fueron establecidas como
+instrumento de manejo pesquero, buscando recuperar las poblaciones de
+abulón y otros invertebrados. Al día de hoy, las reservas marinas de
+Isla Natividad no han recibido reconocimiento legal, pero la cooperativa
+ha mostrado interés por reconocerlas como Zonas de Refugio Pesquero
+(ZRP). Los pescadores tienen un sistema de turnos para vigilar la
+reserva día y noche desde embarcaciones patrulla.
+
+\subsubsection{María Elena}\label{maria-elena}
+
+María Elena es una comunidad pesquera en la costa de Quintana Roo. Los
+arrecifes coralinos y manglares son los principales ecosistemas
+representados en la zona. El campo pesquero es utilizado por pescadores
+de la SCPP Cozumel scl (de la Isla de Cozumel). La principal especie
+aprovechada por ésta organización es la langosta espinosa del caribe
+(\emph{Panulirus argus}). La cooperativa cuenta con permiso de pesca de
+escama y concesión de langosta. En el 2012, la Cooperativa, en conjunto
+con la Alianza Kanan Kay, COBI, CONANP, CONAPESCA, Oceanus, Fundación
+Claudia y Roberto Hernández, Fundación Haciendas del Mundo Maya,
+establecieron ocho ZRP con una vigencia de cinco años. La vigilancia de
+las reservas se realiza por medio del equipo de vigilancia comunitaria,
+con apoyo de la CONANP y una embarcación -donada por COBI- utilizada
+para realizar recorridos frecuentes.
+
+\subsubsection{Punta Herrero}\label{punta-herrero}
+
+La comunidad de Punta Herrero se encuentra aproximadamente a 15 km al
+sur del campo pesquero de María Elena. De igual manera, los arrecifes
+coralinos y manglares son los principales ecosistemas representados en
+la zona, y la principal especie explotada es la langosta. Sin embargo,
+la SCPP José María Azcorra también cuenta con permisos para pesca de
+escama y tiburón y una concesión de langosta. En una réplica del
+ejercicio realizado en María Elena -con presencia de los mismos
+actores-, cuatro ZRP fueron establecidas en el 2013, con una vigencia de
+cinco años. El equipo de vigilancia comunitaria, con apoyo de la CONANP,
+se encarga de la vigilancia de las reservas.
 
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/create_map-1.pdf}
-\caption{(\#fig:create\_map)\label{fig:map}Mapa de la localización
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-1-1.pdf}
+\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Mapa de la localización
 general de las comunidades de estudio. El panel de la derecha es un
@@
 
+Para evaluar las reservas, utilizamos tres fuentes de información. La
+información ecológica proviene de los monitoreos ecológicos realizados
+anualmente en las zonas reserva y control. Cada año, se realizan censos
+visuales para evaluar las comunidades de peces e invertebrados,
+registrando riquezas, abundancias y tallas (en peces). Esta información
+nos permite calcular los indicadores biológicos de manera anual. Al
+tener valores de diferentes indicadores biológicos antes y después de la
+implementación de las reservas, para las zonas de reserva y sitios
+control, tenemos un diseño muestreal de Antes-Después-Control-Impacto.
+
+También incluimos información socioeconómica relevante proveniente de
+los avisos de arribo de CONAPESCA. En este caso, se tienen registros
+mensuales de los recursos aprovechados por las diferentes comunidades,
+en los que se reportan los arribos (Toneladas) y el valor de los arribos
+(\$). La información se encuentra disponible para el periodo 2001 -
+2014. Ya que las categorías registradas por CONAPESCA son amplias y
+existe un nivel de error, utilizamos únicamente los arribos reportados
+para langosta entera fresca a nivel de cooperativa. Los ingresos
+generados por arribos son ajustados por medio del índice de precio al
+consumidor.
+
+La información de gobernanza fue obtenida a nivel de comunidad, pidiendo
+a personas familiares con las comunidades que proveyeran la información
+necesaria. La información de gobernanza no es evaluada de manera
+cuantitativa. En su lugar, esta se interpreta de manera tal que podamos
+comprender qué decisiones, reglas y estructuras tienen un impacto en la
+reserva.
+
 \subsection{Data analysis}\label{data-analysis}
 
-\subsubsection{Biological}\label{biological}
-
-Table \ref{table:indicators} is the way to cite a table.
+Dada la similitud de objetivos entre las reservas, la evaluación se
+realiza con los mismos indicadores. En este caso, se utilizan 6
+indicadores biológicos, 2 socioeconómicos y 5 de gobernanza (Tabla
+\ref{table:indicators}). Según la disponibilidad de datos, se calculó la
+densidad de las especies objetivo presentadas en la sección de
+descripción de las comunidades. El criterio de selección fue que cada
+especie debía tener, por lo menos, dos observaciones anuales para las
+zonas de reserva y control.
 
@@
 \hline
-Biological & Riqueza\\
+ & Riqueza\\
 \hline
-Biological & Densidad\\
+ & Densidad\\
 \hline
-Biological & Nivel trófico\\
+ & Nivel trófico\\
 \hline
-Biological & Biomasa\\
+ & Biomasa\\
 \hline
-Biological & Densidad de especies objetivo\\
+ & Densidad de especies objetivo\\
 \hline
@@
 \hline
-Socioeconomic & Arribos de especies objetivo\\
+ & Arribos de especies objetivo\\
 \hline
@@
 \hline
-Governance & Grado de pesca ilegal\\
+ & Grado de pesca ilegal\\
 \hline
-Governance & Procuración de la reserva\\
+ & Procuración de la reserva\\
 \hline
-Governance & Tipo de organización pesquera\\
+ & Tipo de organización pesquera\\
 \hline
-Governance & Edad de la reserva\\
+ & Edad de la reserva\\
 \hline
@@
 
+\subsubsection{Biological}\label{biological}
+
+Utilizando un análisis de diferencia en diferencias podemos estimar el
+efecto que la reserva tienen en los indicadores biológicos
+\citep{moland_2013-VP} con el uso de un modelo de regresión lineal
+múltiple:
+
+\[I = \beta_0 + \sum \gamma Year + \beta_1 Zona + \sum \lambda Year\times Zona + \Omega + \epsilon\]
+
+En este caso, modelamos los años como factores, tomando como referencia
+el primer año en la serie de datos de cada comunidad. Modelar los años
+como factores reduce la estructura del modelo, y relaja el ajuste al no
+asumir una tendencia lineal entre años; es decir, el cambio observado
+entre 2006 - 2007 no deberá de ser igual al observado entre el 2009 -
+2010. Incluimos también un término para la zona, en la que la variable
+toma un valor de 0 si el sitio es una zona control y de 1 si es una zona
+de reserva. Finalmente, incluimos un término de interacción entre la
+variable de Zona y el Año. En este modelo, \(\lambda_i\) representa el
+efecto que la reservas tuvo sobre un indicador en cada año y con
+respecto a los sitios control. El termino \(\Omega\) captura efectos
+fijos por especies y por sitio.
+
 \subsubsection{Socioeconomic}\label{socioeconomic}
+
+El análisis de datos socioeconómicos se aplicó únicamente a Isla
+Natividad, María Elena y Punta Herrero, siguiendo un modelo con la
+forma:
+
+\[I = \beta_0 + \beta_1Post\]
+
+Que nos permite comparar el cambio en el promedio de los indicadores
+antes (Post = 0) y después (Post = 1) de la implementación de la
+reserva. Tanto para los indicadores biológicos como los socioeconómicos,
+los coeficientes fueron ajustador con el estimador de muestras
+heterocedásticas.
 
@@
 
+A continuación se presentan los resultados de cada una de las
+comunidades. Los resultados biológicos se presentarán para cada
+comunidad, discutiendo primero los indicadores en común con otras
+comunidades (Shannon, Riqueza, Densidad, Nivel Trófico, Biomasa para
+peces e invertebrados) y, según su caso, las densidades de las especies
+objetivo. Habiendo presentado los resultados biológicos, nos enfocaremos
+después en los socioeconómicos y de gobernanza. Usaremos esta
+información para identificar las causas (sociales) del éxito (biológico)
+de las reservas.
+
+\begin{figure}
+\centering
 \includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-3-1.pdf}
-
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{\label{fig:unnamed-chunk-3}\label{fig:indicators}Effect sizes for
+marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
+blue triangles), and Punta Herrero (PH; green squares). Plots are
+ordered by survey type (left: fish; right: invertebrates) and
+indicators: Abundance (A, B), Richness (C, D), and Shannon's diversity
+index (E, F)}
+\end{figure}
 
 \section{Discussion}\label{discussion}
-
-This section may be divided by subheadings. Discussions should cover the
-key findings of the study: discuss any prior art related to the subject
-so to place the novelty of the discovery in the appropriate context;
-discuss the potential short-comings and limitations on their
-interpretations; discuss their integration into the current
-understanding of the problem and how this advances the current views;
-speculate on the future direction of the research and freely postulate
-theories that could be tested in the future.
-
-Summary of main findings Limitations Conclusions
 

### 2018-01-27 TechnicalReport/Manuscript.Rmd
Updated manuscript

@@
 Los ingresos generados por arribos son ajustados por medio del índice de precio al consumidor.
-El análisis de datos socioeconómicos se aplicó únicamente a Isla Natividad, María Elena y Punta Herrero, siguiendo un modelo con la forma:
-
-$$I = \beta_0 + \beta_1Post$$
-
-Que nos permite comparar el cambio en el promedio de los indicadores antes (Post = 0) y después (Post = 1) de la implementación de la reserva.
-Tanto para los indicadores biológicos como los socioeconómicos, los coeficientes fueron ajustador con el estimador de muestras heterocedásticas.
 

### 2018-02-08 Scripts/biomass.R
Finalized indicator functions. Close #3

@@ -1,0 +1,12 @@
+biomass <- function(x, com){
+  results <- x %>% 
+    filter(Comunidad == com) %>% 
+    filter(Abundancia > 0) %>% 
+    left_join(species_bio, by = c("GeneroEspecie")) %>% 
+    mutate(Ano = as.factor(Ano),
+           biomass = Abundancia * a * (Talla^b)/1000) %>% 
+    group_by(RC, Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
+    summarize(Indicador = sum(biomass, na.rm = T))
+  
+  return(results)
+}

### 2018-02-08 Scripts/trophic.R

@@ -1,0 +1,12 @@
+trophic <- function(x, com){
+  resutls <- x %>% 
+    filter(Comunidad == com) %>% 
+    filter(Abundancia > 0) %>% 
+    left_join(species_bio, by = c("GeneroEspecie")) %>% 
+    reshape::untable(df = ., num = .$Abundancia) %>% 
+    mutate(Ano = as.factor(Ano)) %>% 
+    group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
+    summarize(Indicador = mean(NT, na.rm = T))
+  
+  return(resutls)
+}

### 2018-02-08 Frontiers_Manuscript/Villasenor-Derbez.Rmd
updated manuscript

@@
 rtitle: "Mexican marine reserves"
-author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
+author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Stuart Fulton$^{2}$"
 institution: "
@@
 keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico."
-# biblio: test.bib
 output:
@@
     template: template.tex
-# csl: FMS.csl
 ---
@@
 <!-- - Introduction -->
-
-<!--   - Las reservas marinas son una herramienta para X y Y, con beneficios A y B -->
-
-<!--   - Existen diferencias entre reservas bottom up y top down -->
-
-<!--   - En Mexico, las reservas se implementan de 3 maneras -->
-
-<!--   - Los ZRP tienen X objetivos -->
-
-<!--   - A X anos de la implementacion de la NOM-049, no hay estudio formal de ZRP o reservas comunitarias -->
-
-<!--   - Proveer la primer evaluacion de reservas comunitarias triple bottom line -->
 
@@
 
+<!--   - Las reservas marinas son una herramienta para X y Y, con beneficios A y B -->
+
+<!--   - Existen diferencias entre reservas bottom up y top down -->
+
+<!--   - En Mexico, las reservas se implementan de 3 maneras -->
+
+<!--   - Los ZRP tienen X objetivos -->
+
+<!--   - A X anos de la implementacion de la NOM-049, no hay estudio formal de ZRP o reservas comunitarias -->
+
+<!--   - Proveer la primer evaluacion de reservas comunitarias triple bottom line -->
+
 <!-- Rationale -->
@@
 
-La sobrepesca y prácticas pesqueras no sostenibles son unas de las mayores amenazas para la conservación de los ecosistemas marinos del mundo \citep{halpern_2008-dK,halpern_2017-Zi}.
-La implementación de reservas marinas (*i.e.* áreas donde la captura de una o más especies está prohibida) es una medida de manejo frecuentemente propuesta para recuperar stocks pesqueros e impulsar la productividad pesquera en aguas cercanas \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}.
-Recientes trabajos han demostrado que también pueden mitigar y proveer amortiguamiento ante el cambio climático \citep{roberts_2017-J9}, variabilidad ambiental \citep{micheli_2012-EU}, resolver problemas de pesca incidental \citep{hastings_2017-sm} y, en general, incrementar la biomasa, riqueza y densidades de organismos dentro de sus fronteras \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}.
+Overfishing and unsustainable fishing practices threaten marine ecosystems around the world \citep{halpern_2008-dK,halpern_2017-Zi}.
+The implementation of marine reserves (*i.e.* areas where all fishing activities are off--limits; MRs) is a common management approach used to manage the spatial distribution of fishing effort and recover stocks \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}.
+Recent research poses that marine reserves can also help buffer climate change \citep{roberts_2017-J9}, provide a refuge to environmental variability \citep{micheli_2012-EU}, help reduce by-catch \citep{hastings_2017-sm} and, in general, increase biomass, richness and densities of organisms within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}.
+However, the extent to which MRs yield benefits to fisheries remains relatively unexplored \citep{krueck_2017-J1} and the focus is often only in the social and economic dimensions \citep{klein_2008,charles_2008}, rarely providing a hollistic evaluation of the social-ecological system \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
+Here, we combine causal inference techniques \citep{depalma_2018} and the social-ecological systems framework \citep{ostrom_2009-hg,mascia_2017-m_} to provide a comprehensive evaluation of four community-based marine reserves in three coastal communities in Mexico.
 
@@
 source(file = "../Scripts/shannon.R")
+source(file = "../Scripts/biomass.R")
+source(file = "../Scripts/trophic.R")
 
@@
   filter(!is.na(RC),
-         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)", "El Faro - El Faro (Control)", "Anegado de Chal - Anegado de Chal (Control)"))
+         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)", "El Faro - El Faro (Control)", "Anegado de Chal - Anegado de Chal (Control)")) %>% 
+  mutate(GeneroEspecie = case_when(
+    GeneroEspecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
+    GeneroEspecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
+    GeneroEspecie == "Xcochin spp" ~ "Balistes capriscus",
+    GeneroEspecie == "Pterois volitans/miles" ~ "Pterois volitans",
+    GeneroEspecie == "Mycteroperca spp" ~ "Mycteroperca",
+    GeneroEspecie == "Chromis spp" ~ "Chromis",
+    GeneroEspecie == "Damisela spp" ~ "Stegastes",
+    GeneroEspecie == "Haemulon spp" ~ "Haemulon",
+    GeneroEspecie == "Lancero spp" ~ "Acanthurus",
+    GeneroEspecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
+    GeneroEspecie == "Sebastes spp" ~ "Sebastes",
+    GeneroEspecie == "Loro spp" ~ "Scarus",
+    GeneroEspecie == "Lutjanus spp" ~ "Lutjanus",
+    GeneroEspecie == "Mariposa spp" ~ "Chaetodon",
+    GeneroEspecie == "Pluma spp" ~ "Calamus pennatula",
+    GeneroEspecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
+    GeneroEspecie == "Vieja spp" ~ "Bodianus rufus",
+    # GeneroEspecie == "" ~ "",
+    TRUE ~ GeneroEspecie
+  ))
 
@@
 
+data("species_bio")
+
+spp <- group_by(fish_complete, GeneroEspecie) %>% 
+  count() %>% 
+  left_join(species_bio, by = "GeneroEspecie") %>% 
+  filter(is.na(a) | is.na(b))
+
+FB_update <- rfishbase::length_weight(c("Kyphosus incisor", "Kyphosus sectatrix", "Pterois volitans", "Scarus guacamaia", "Chaetodon capistratus", "Chaetodon capistratus", "Chaetodon striatus", "Calamus pennatula"), fields = c("a", "b")) %>%
+  select(GeneroEspecie = sciname, a, b) %>% 
+  mutate(GeneroEspecie = case_when(
+    GeneroEspecie %in% c("Kyphosus sectatrix", "Kyphosus incisor") ~ "Kyphosus",
+    GeneroEspecie %in% c("Chaetodon capistratus", "Chaetodon striatus") ~ "Chaetodon",
+    TRUE ~ GeneroEspecie
+  )) %>% 
+  group_by(GeneroEspecie) %>% 
+  summarize_all(mean, na.rm = T)
+
+GenusLevelBio <- species_bio %>% 
+  filter(Genus %in% c("Mycteroperca", "Chromis", "Stegastes", "Haemulon", "Acanthurus", "Sebastes", "Lutjanus", "Scarus")) %>%
+  select(GeneroEspecie = Genus, a, b) %>% 
+  group_by(GeneroEspecie) %>% 
+  summarize_all(mean, na.rm = T)
+
+species_bio %<>%
+  select(GeneroEspecie, a, b) %>% 
+  rbind(FB_update) %>% 
+  rbind(GenusLevelBio)
+```
+
+```{r complete_inverts}
 invert_complete <- invert %>%
@@
   model(sfe = F, spfe = F)
+
+## Biomass
+B_IN <- fish %>% 
+  biomass(com = "Isla Natividad") %>% 
+  model(sfe = F)
+
+B_ME <- fish%>% 
+  biomass(com = "Maria Elena") %>% 
+  model(sfe = T)
+
+B_PH <- fish %>% 
+  biomass(com = "Punta Herrero") %>% 
+  model(sfe = F)
+
+## Trophic level
+# T_IN <- fish %>% 
+#   trophic("Isla Natividad") %>% 
+#   model(sfe = F, spfe = F)
+# 
+# T_ME <- fish %>% 
+#   trophic("Maria Elena") %>% 
+#   model(spfe = F)
+# 
+# T_PH <- fish %>% 
+#   trophic("Punta Herrero") %>% 
+#   model(sfe = F, spfe = F)
 
@@
                   model_names = c("IN", "ME", "PH"))
+
 # Invert plots
@@
                    ncol = 2)
+```
+
+```{r, fig.width = 4, fig.height = 3}
+F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
+                 various = T,
+                 model_names = c("IS", "ME", "PH"))
+
+F_B
 ```

### 2018-02-08 Frontiers_Manuscript/Villasenor-Derbez.tex
updated manuscript

@@
 \usepackage[onehalfspacing]{setspace}
+\usepackage{float}
 
@@
 \usepackage[utf8]{inputenc}
+\floatplacement{figure}{H}
 
@@
 \def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
-\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Stuart Fulton\(^{2}\), Jorge
-Torre\(^{2}\)}
+\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
+Aceves-Bueno\(^{1,*}\), Stuart Fulton\(^{2}\)}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 
-La sobrepesca y prácticas pesqueras no sostenibles son unas de las
-mayores amenazas para la conservación de los ecosistemas marinos del
-mundo \citep{halpern_2008-dK,halpern_2017-Zi}. La implementación de
-reservas marinas (\emph{i.e.} áreas donde la captura de una o más
-especies está prohibida) es una medida de manejo frecuentemente
-propuesta para recuperar stocks pesqueros e impulsar la productividad
-pesquera en aguas cercanas
-\citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}. Recientes
-trabajos han demostrado que también pueden mitigar y proveer
-amortiguamiento ante el cambio climático \citep{roberts_2017-J9},
-variabilidad ambiental \citep{micheli_2012-EU}, resolver problemas de
-pesca incidental \citep{hastings_2017-sm} y, en general, incrementar la
-biomasa, riqueza y densidades de organismos dentro de sus fronteras
-\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}.
+Overfishing and unsustainable fishing practices threaten marine
+ecosystems around the world \citep{halpern_2008-dK,halpern_2017-Zi}. The
+implementation of marine reserves (\emph{i.e.} areas where all fishing
+activities are off--limits; MRs) is a common management approach used to
+manage the spatial distribution of fishing effort and recover stocks
+\citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}. Recent research
+poses that marine reserves can also help buffer climate change
+\citep{roberts_2017-J9}, provide a refuge to environmental variability
+\citep{micheli_2012-EU}, help reduce by-catch \citep{hastings_2017-sm}
+and, in general, increase biomass, richness and densities of organisms
+within the protected regions
+\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}. However, the
+extent to which MRs yield benefits to fisheries remains relatively
+unexplored \citep{krueck_2017-J1} and the focus is often only in the
+social and economic dimensions \citep{klein_2008,charles_2008}, rarely
+providing a hollistic evaluation of the social-ecological system
+\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. Here, we combine
+causal inference techniques \citep{depalma_2018} and the
+social-ecological systems framework
+\citep{ostrom_2009-hg,mascia_2017-m_} to provide a comprehensive
+evaluation of four community-based marine reserves in three coastal
+communities in Mexico.
 
@@
 
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
+
 \section{Discussion}\label{discussion}

### 2018-02-09 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Updated plotting function and manuscript

@@
 ---
-title: "Community-based marine reserves produce biological and economic benefits"
+title: "Cool title here"
 rtitle: "Mexican marine reserves"
@@
 email: "jvillasenor@bren.ucsb.edu"
-abstract: ""
+abstract: "Nice abstract here"
 keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico."
@@
   library(ggsn)
-  library(patchwork)
   library(tmap)
@@
 
-$$I = \beta_0 + \sum \gamma Year + \beta_1 Zona + \sum \lambda Year\times Zona + \Omega + \epsilon$$
+$$I = \alpha_{i} + \gamma_{it} Year + \beta_{i} Zona + \lambda_{it} Year\times Zona + \sigma_{j} + \Pi_{k} + \epsilon$$
 
@@
   abundance(com = "Isla Natividad") %>% 
-  model(sfe = F)
+  model(sfe = F, family = "quasipoisson")
 
@@
   abundance(com = "Maria Elena") %>% 
-  model(sfe = T)
+  model(sfe = T, family = "quasipoisson")
 
@@
   abundance(com = "Punta Herrero") %>% 
-  model(sfe = F)
+  model(sfe = F, family = "quasipoisson")
 
@@
   model(sfe = F, spfe = F)
+
+## Lobsters
+L_IN <- invert_complete %>%
+  filter(GeneroEspecie == "Panulirus interruptus") %>% 
+  abundance(com = "Isla Natividad") %>% 
+  model(sfe = F, spfe = F, family = "quasipoisson")
+
+L_ME <- invert_complete %>% 
+  filter(GeneroEspecie == "Panulirus argus") %>% 
+  abundance(com = "Maria Elena") %>% 
+  model(sfe = T, spfe = F, family = "quasipoisson")
+
+L_PH <- invert_complete %>% 
+  filter(GeneroEspecie == "Panulirus argus") %>% 
+  abundance(com = "Punta Herrero") %>% 
+  model(sfe = F, spfe = F, family = "quasipoisson")
+
 ```
@@
 
-```{r, fig.width = 6, fig.height = 8, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares).
-Plots are ordered by survey type (left: fish; right: invertebrates) and indicators: Abundance (A, B), Richness (C, D), and Shannon's diversity index (E, F)"}
+```{r, fig.width = 6.5, fig.height = 8, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for community-levell indicators.
+Plots are ordered by survey type (left: fish; right: invertebrates) and indicators: Abundance (A, B), Richness (C, D), Shannon's diversity index (E, F), and fish biomass (G)"}
 
@@
                   various = T,
-                  model_names = c("IN", "ME", "PH"),
-                  legend = T)
+                  model_names = c("IN", "ME", "PH"), legend = T)  +
+  theme(legend.justification = c(0, 0),
+        legend.position = c(0, 0),
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        legend.direction = "horizontal") +
+  scale_colour_brewer(guide = guide_legend(title.position = "top", nrow = 1), palette = "Set1")
 
@@
                   model_names = c("IN", "ME", "PH"))
+
+F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
+                 various = T,
+                 model_names = c("IS", "ME", "PH"))
 
@@
 
-cowplot::plot_grid(plotlist = list(F_N, I_N, F_S, I_S, F_H, I_S),
+cowplot::plot_grid(plotlist = list(F_N, I_N, F_S, I_S, F_H, I_S, F_B),
                    labels = "AUTO",
@@
 
-```{r, fig.width = 4, fig.height = 3}
-F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
+```{r, fig.width = 4, fig.height = 6, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster abundances in marine reserves from Isla Natividad (IN; red romboids), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares)."}
+
+L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
                      various = T,
-                 model_names = c("IS", "ME", "PH"))
+                     model_names = c("IS", "ME", "PH")) +
+  theme(legend.position = c(0, 1))
 
-F_B
+L_conapesca <- tibble(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl",
+                                          "Scpp cozumel scl",
+                                          "Scpp jose maria azcorra scl"),
+                      Comunidad = c("Isla Natividad",
+                                    "María Elena",
+                                    "Punta Herrero")) %>% 
+  left_join(conapesca, ., by = "UnidadEconomica") %>% 
+  mutate(Post = case_when(Comunidad == "Isla Natividad" && Ano <= 2006 ~ "Pre",
+                          Comunidad == "Maria Elena" && Ano <= 2012 ~ "Pre",
+                          Comunidad == "Punta Herrer" && Ano <= 2012 ~ "Pre",
+                          TRUE ~ "Post"))
+
+L_catches <- ggplot(L_conapesca, aes(x = Ano, y = PesoVivo/1000, color = Comunidad, group = Comunidad, fill = Comunidad)) +
+  geom_smooth(method = "lm") +
+  geom_point(size = 2, alpha = 0.5) +
+  scale_color_brewer(palette = "Set1") +
+  geom_vline(xintercept = c(2006, 2012, 2013), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +
+  labs(x = "Year", y = "Catches (Tones)") +
+  theme(legend.position = "none")
+
+L_value <- ggplot(L_conapesca, aes(x = Ano, y = Valor/1000, color = Comunidad, group = Comunidad, fill = Comunidad)) +
+  geom_smooth(method = "lm") +
+  geom_point(size = 2, alpha = 0.5) +
+  scale_color_brewer(palette = "Set1") +
+  geom_vline(xintercept = c(2006, 2012, 2013), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +
+  labs(x = "Year", y = "Value (MXP)") +
+  theme(legend.position = "none")
+
+cowplot::plot_grid(plotlist = list(L_plot, L_catches, L_value),
+                   labels = "AUTO",
+                   ncol = 1)
+
 ```

### 2018-02-09 Frontiers_Manuscript/Villasenor-Derbez.tex
Updated plotting function and manuscript

@@
 
-\title[Mexican marine reserves]{Community-based marine reserves produce biological and economic benefits}
+\title[Mexican marine reserves]{Cool title here}
 
@@
 \maketitle
+
+\begin{abstract}
+
+Nice abstract here
+
+\medskip
+\tiny
+ \keyFont{ \section{Keywords:} Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen
+Science, Mexico.}
+
+\end{abstract}
 
@@
 marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
-blue triangles), and Punta Herrero (PH; green squares). Plots are
-ordered by survey type (left: fish; right: invertebrates) and
-indicators: Abundance (A, B), Richness (C, D), and Shannon's diversity
-index (E, F)}
+blue triangles), and Punta Herrero (PH; green squares) for
+community-levell indicators. Plots are ordered by survey type (left:
+fish; right: invertebrates) and indicators: Abundance (A, B), Richness
+(C, D), Shannon's diversity index (E, F), and fish biomass (G)}
 \end{figure}
 
+\begin{figure}
+\centering
 \includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{\label{fig:unnamed-chunk-4}\label{fig:lobsters}Effect sizes for
+lobster abundances in marine reserves from Isla Natividad (IN; red
+romboids), Maria Elena (ME; blue triangles), and Punta Herrero (PH;
+green squares).}
+\end{figure}
 

### 2018-02-09 Scripts/plot_model.R

@@ -24,7 +24,7 @@ plot_model <- function(model, various = F, model_names = NULL, legend = F){
                Community = model_names[i])
       
       p_i <- robust_se(model[[i]]) %>%
-        filter(grepl(":ZonaReserva", term)) %>% 
+        filter(grepl(":ZonaReserva", term)) %>%
         cbind(tidy_terms) %>%
         mutate(p = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05"))
       
@@ -39,9 +39,7 @@ plot_model <- function(model, various = F, model_names = NULL, legend = F){
       geom_point(aes(shape = Community), size = 4, position = pd, alpha = 0.5) +
       geom_hline(yintercept = 0, linetype = "dashed") +
       scale_fill_brewer(palette = "Set1") +
-      theme(legend.justification = c(0, 1),
-            legend.position = c(0, 1),
-            text = element_text(size = 12)) +
+      theme(text = element_text(size = 12)) +
       labs(x = "Year", y = "Effect size") +
       scale_color_brewer(palette = "Set1")
     

### 2018-02-11 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Added english text for intro and methods

@@
 email: "jvillasenor@bren.ucsb.edu"
-abstract: "Nice abstract here"
-keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico."
+abstract: "Abstract here"
+keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico, Social--Ecological Systems"
 output:
@@
 
-<!-- **Esqueleto:** -->
+# Introduction
 
-<!-- - Introduction -->
+Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}.
+A common approach to manage the spatial distribution of fishing effort and recover stocks is through the implementation of marine reserves (*i.e.* areas where all fishing activities are off--limits; MRs) \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}.
 
-<!-- - Materials and methods -->
+Marine reserve science has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, climate change mitigation \citep{roberts_2017-J9}, and protection from environmental variability \citep{micheli_2012-EU}.
+However, there is considerably less literature focusing on the relationship between socioeconomic and governance structures and their relationship to ecological effectriveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or benefits to fisheries \citep{krueck_2017-J1}; evaluations of marine reserves rarely provide a hollistic view of the social-ecological system \citep{lpezangarita_2014}.
+Here, we combine causal inference techniques \citep{depalma_2018} and the social-ecological systems framework \citep{ostrom_2009-hg} to provide a comprehensive ecological and socioeconomic evaluation of four community-based marine reserves in three coastal communities in Mexico.
 
-<!--   - Study areas -->
+Marine Reserves in Mexico have been commonly implemented as "core zones" within Biosphere Reserves (BRs) that are administered by the National Commision of Protected Areas (*Comisión Nacional de Áreas Marinas Protegidas*, CONANP).
+While CONANP has made efforts to have a participatory process, the implementation of these zones is still characterized by top--down approaches.
+This motivated Civil Society Organizations (CSOs) to work with coastal communities to implement community--based marine reserves \citep{uribe_2010-u2}, which are usually established within a Territorial Use Rights for Fisheries (TURFs); thus making them TURF--reserves \citep{afflerbach_2014-HP}.
+This bottom--up approach allows fishers to disign their own reserves, which  increases compliance and self--enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
+However, these reserves still lack legal recognition, making them vulnerable to poaching.
+In 2014, a new norm \citep{nom} allowed fishers to request the legal recognition of a community--based reserve under the name of "Fishing Refugia" (*Zona de Refugio Pesquero*, FR).
+This new norm thus combines bottom--up approaches to design marine reserves, along with a legal recognition of the management intervention.
+Since then, *39* FR have been implemented along the Pacific, Gulfof Caloifornia, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.
 
-<!--     - Main type of ecosystem -->
+While there are ecological factors defining the success of a MR (*i.e.* habitat representation, initial state of protection, connectivity to other protected areas), their effectivness also depends on the socioeconomic and governance settings under which they are implemented.
+Literature shows that many non-ecological characteristics can play an equally important role in the effectivenes of MRs.
+For example, age of a reserve (*i.e.* time since its implementation), size, and habitat contained were key to the effectiveness of MRs in Palau \citep{friedlander_2017-oI}.
+In the Mediterranean, \citet{difranco_2016-Xw} identify that surveilance and enfocrement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key fators that increased stock health and income to fishers.
+At a global level, \citet{edgar_2014-UO} indicate that enforcement, age, size, and isolation were important factors determining effectiveness of the reserves.
 
-<!--     - Number of fishers (aprox) -->
+The objective of this work is twofold: i) Provide the first evaluation of community--based marine reserves in Mexico, and ii) provide a comprehensive evaluation of the social--ecological system to identify how sociocoenomic and governance characteristics relate to ecological effectiveness.
+With the purpose of providing a hollistic evaluation, we combine ecological, socioeconomic, and governance indicators.
+We use causal inference techniques to provide a measurement of the effect of the management intervention, and combine it with the social--ecological systems framework \citep{ostrom_2009-hg}.
+
+# Materials and Methods
 
@@
 
-<!-- - Results -->
-
-<!-- - Discusion -->
-
-# Introduction
-
-<!-- Succinct, with no subheadings. -->
-
-<!--   - Las reservas marinas son una herramienta para X y Y, con beneficios A y B -->
-
-<!--   - Existen diferencias entre reservas bottom up y top down -->
-
-<!--   - En Mexico, las reservas se implementan de 3 maneras -->
-
-<!--   - Los ZRP tienen X objetivos -->
-
-<!--   - A X anos de la implementacion de la NOM-049, no hay estudio formal de ZRP o reservas comunitarias -->
-
-<!--   - Proveer la primer evaluacion de reservas comunitarias triple bottom line -->
-
-<!-- Rationale -->
-<!-- Objectives -->
-<!-- Research question -->
-
-Overfishing and unsustainable fishing practices threaten marine ecosystems around the world \citep{halpern_2008-dK,halpern_2017-Zi}.
-The implementation of marine reserves (*i.e.* areas where all fishing activities are off--limits; MRs) is a common management approach used to manage the spatial distribution of fishing effort and recover stocks \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}.
-Recent research poses that marine reserves can also help buffer climate change \citep{roberts_2017-J9}, provide a refuge to environmental variability \citep{micheli_2012-EU}, help reduce by-catch \citep{hastings_2017-sm} and, in general, increase biomass, richness and densities of organisms within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}.
-However, the extent to which MRs yield benefits to fisheries remains relatively unexplored \citep{krueck_2017-J1} and the focus is often only in the social and economic dimensions \citep{klein_2008,charles_2008}, rarely providing a hollistic evaluation of the social-ecological system \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
-Here, we combine causal inference techniques \citep{depalma_2018} and the social-ecological systems framework \citep{ostrom_2009-hg,mascia_2017-m_} to provide a comprehensive evaluation of four community-based marine reserves in three coastal communities in Mexico.
-
-En México, las reservas marinas han sido comúnmente establecidas como zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por la Comisión Nacional de Áreas Naturales Protegidas (CONANP).
-Al día de hoy, 36 RBs protegen una porción del ambiente marino en México.
-Sin embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde las actividades pesqueras están prohibidas.
-Aunque la CONANP ha hecho esfuerzos importantes por involucrar a los actores durante la implementación de las reservas, esto aún se caracteriza por un proceso descendente, el cual conlleva a la falta de cumplimiento por parte de los actores.
-La escasez de recursos monetarios y humanos de la limitan también el monitoreo y vigilancia de las reservas, y a su vez, el desempeño de la reserva.
-
-Buscando promover una alternativa con procesos ascendentes para implementar reservas marinas, las Organizaciones de la Sociedad Civil (OSCs) comenzaron a trabajar con comunidades pesqueras para establecer reservas comunitarias \citep{uribe_2010-u2} .
-Estas son comúnmente establecidas dentro de zonas de concesión, una forma de derechos de uso territoriales para pesquerías (TURF, en inglés).
-Al permitir a los pescadores diseñar sus propias reservas, una mayor proporción de la comunidad está de acuerdo con los perímetros y reglas establecidas, y por lo tanto los respetan \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8} .
-Adicionalmente, los pescadores pueden implementar sus reservas por un periodo acordado (usualmente cinco años), después del cual la reserva puede ser abierta a la pesca.
-Esto provee a los pescadores con un sentido de confianza de que, en caso de ser necesario, aún tienen acceso a pescar esa zona.
-Las reservas son directamente vigiladas y monitoreadas por la comunidad, quienes comúnmente utilizan pequeñas embarcaciones (*e.g.* pangas) para patrullar la zona, o realizan avistamientos desde la costa en búsqueda de pescadores ilegales Aún así, las reservas comunitarias carecen de reconocimiento legal; por lo tanto, no hay forma de penalizar a los infractores.
-
-Sin embargo, en el 2014 una nueva norma \citep{nom} permite a los pescadores solicitar el establecimiento de reservas marinas bajo el nombre de “Zonas de refugio Pesquero” (ZRP).
-El manejo de las ZRP combina procesos ascendentes y descendentes al reconocer legalmente las reservas propuestas por las comunidades.
-Posterior a la revisión por parte de la Comisión Nacional de Acuacultura y Pesca (CONAPESCA) y la opinión técnica del Instituto Nacional de Acuacultura y Pesca (INAPESCA) las ZRP son establecidas por el periodo solicitado por los pescadores.
-El monitoreo y la vigilancia de las ZRP es típicamente llevado a cabo por la comunidad , con ayuda de OSCs locales.
-Hasta este cambio regulatorio, las reservas comunitarias no contaban con el soporte legal, y eran solamente reconocidas por la comunidad.
-Al día de hoy, existen 39 ZRP establecidas en el Pacífico, Golfo de California y Caribe Mexicano.
-
-Aunque existen tres aproximaciones generales para implementar reservas marinas en México (*i.e.* Zonas núcleo dentro de AMP, reservas comunitarias y Zonas de Refugio Pesquero), aún no comprendemos a fondo las características sociales que permiten su efectividad.
-La ciencia de reservas marinas se ha enfocado ampliamente en los efectos biológicos que estas tienen \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,afflerbach_2014-HP,krueck_2017-J1}.
-Aunque el aspecto ecológico de las reservas es importante para su éxito, su efectividad también depende del estado socioeconómico y los sistemas de gobernanza de las comunidades pesqueras.
-
-La literatura indica que diferentes características influyen en el éxito de una reserva.
-En Palau, por ejemplo, la edad (*i.e.* tiempo transcurrido desde implementación), tamaño y hábitat contenido son características claves que determinan la efectividad \citep{friedlander_2017-oI}.
-Por otro lado, en el Mar Mediterráneo, \citet{difranco_2016-Xw} identifican que la procuración y vigilancia, presencia de un plan de manejo, participación de pescadores en el manejo, representación de pescadores en la toma de decisiones y promoción de la pesca sustentable son los cinco factores que incrementan la salud de los stocks y el ingreso económicos a los pescadores, a la vez que se presenta una mayor aceptación social de las prácticas de manejo.
-En una aproximación global, \citet{edgar_2014-UO} encuentran que la procuración, edad, tamaño y aislamiento son determinantes de la efectividad de las reservas.
-Por lo tanto, observamos que las características que habilitan el éxito varían a través de regiones, y poco esfuerzo se ha hecho por comprender estas interacciones en México.
-
-El objetivo de este trabajo este trabajo es realizar una evaluación de la efectividad de reservas marinas en México, presentando resultados de cinco comunidades costeras como caso de estudio.
-Con el fin de obtener una visión holística del sistema, la evaluación se realizará tomando en cuenta indicadores biológicos, socioeconómicos y de gobernanza.
-La evaluación de éstos cinco casos de estudios nos permitirá identificar la manera en que las características socioeconómicas y de gobernanza se relacionan con la efectividad (biológica) de las reservas marinas evaluadas.
-Los patrones identificados podrán utilizarse para informar la toma de decisiones para la implementación de la red de reservas marinas en la Región de las Grandes Islas del Golfo de California.
-
-# Materials and Methods
-
-<!-- This section may be divided by subheadings.
-This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated.
-For experiments reporting results on animal or human subject research, an ethics approval statement should be included in this section (for further information, see section Materials and Data Policies) -->
-
 ## Study area
 
-Las comunidades utilizadas en este reporte se distribuyen a lo largo de la costa Pacífica de Baja California (n = 1) y el Sistema Arrecifal Mesoamericano (n = 2; Fig \ref{fig:map}).
+We focus our evaluation in three coastal communities from the Pacific coast of Baja California (n = 1) and the Mesoamerican Reef System (n = 2; Fig \ref{fig:map}).
+Isla Natividad (IN) lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests (*Macrocystis pyrifera*) and rocky reefs are the predominant habitats.
+The island is home to a fishing cooperative (*Sociedad Cooperative de Producción Pesquera Buzos y Pescadores de la Baja California SCL*), that holds a TURF for spiny lobster (*Panulirus interruptus*).
+However, other resources like finfish (yellowtail jack, *Seriola lalandi*), sea cucumber (*Parastichopus parvimensis*), red sea urchin (*Mesocentrotus franciscanus*), snail (*Megastraea turbanica* y *M.
+undosa*), and abalone (*Haliotis spp*, until 2010) are also important sources of income.
+In 2006, the community decided to implement two community--based marine reserves within their fishing grounds, seeking to recover depleted stocks of invertebrate species (mainly lobster and abalone).
+Until today, these reserves are yet to be legally recognized as Fishing Refugia.
 
-### Isla Natividad
-
-La Isla Natividad se encuentra en la costa oeste de la Península de Baja California, donde el hábitat predominante es el bosque de kelp o sargazo gigante (*Macrocystis pyrifera*) y los arrecifes rocosos.
-En la isla, la Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California SCL realiza actividades de extracción de los recursos marinos.
-Aunque la langosta roja (*Panulirus interruptus*) es la especie más importante en términos económicos, otras especies importantes incluyen la escama (con un enfoque en Jurel; *Seriola lalandi*), el pepino de mar (*Parastichopus parvimensis*), el erizo rojo (*Mesocentrotus franciscanus*), el caracol (*Megastraea turbanica* y *M.
-undosa*) y, hasta el 2010, el abulón (*Haliotis sp.*).
-En 2006, por medio de un proceso participativo, la cooperativa decidió establecer dos reservas marinas de manera voluntaria.
-Agentes externos a la cooperativa, como personal de Comunidad y Biodiversidad A.C., académicos de la Universidad de Stanford, y personal de la CONANP (de la oficina de Reserva de la Biósfera El Vizcaíno), también participaron en el diseño e implementación de las reservas.
-Las reservas fueron establecidas como instrumento de manejo pesquero, buscando recuperar las poblaciones de abulón y otros invertebrados.
-Al día de hoy, las reservas marinas de Isla Natividad no han recibido reconocimiento legal, pero la cooperativa ha mostrado interés por reconocerlas como Zonas de Refugio Pesquero (ZRP).
-Los pescadores tienen un sistema de turnos para vigilar la reserva día y noche desde embarcaciones patrulla.
-
-### María Elena
-
-María Elena es una comunidad pesquera en la costa de Quintana Roo.
-Los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona.
-El campo pesquero es utilizado por pescadores de la SCPP Cozumel scl (de la Isla de Cozumel).
-La principal especie aprovechada por ésta organización es la langosta espinosa del caribe (*Panulirus argus*).
-La cooperativa cuenta con permiso de pesca de escama y concesión de langosta.
-En el 2012, la Cooperativa, en conjunto con la Alianza Kanan Kay, COBI, CONANP, CONAPESCA, Oceanus, Fundación Claudia y Roberto Hernández, Fundación Haciendas del Mundo Maya, establecieron ocho ZRP con una vigencia de cinco años.
-La vigilancia de las reservas se realiza por medio del equipo de vigilancia comunitaria, con apoyo de la CONANP y una embarcación -donada por COBI- utilizada para realizar recorridos frecuentes.
-
-### Punta Herrero
-
-La comunidad de Punta Herrero se encuentra aproximadamente a 15 km al sur del campo pesquero de María Elena.
-De igual manera, los arrecifes coralinos y manglares son los principales ecosistemas representados en la zona, y la principal especie explotada es la langosta.
-Sin embargo, la SCPP José María Azcorra también cuenta con permisos para pesca de escama y tiburón y una concesión de langosta.
-En una réplica del ejercicio realizado en María Elena -con presencia de los mismos actores-, cuatro ZRP fueron establecidas en el 2013, con una vigencia de cinco años.
-El equipo de vigilancia comunitaria, con apoyo de la CONANP, se encarga de la vigilancia de las reservas.
+The other two communities are Maria Elena (ME; Fig \ref{fig:map}C) and Punta Herrero (PH; Fig \ref{fig:map}D) in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems.
+ME is a fishing camp --visited intermitently during the fishing season-- belonging to the Cozumel fishing cooperative.
+PH is home to the "José María Azcorra" fishing cooperative.
+The main source of income to both communities is the caribbean spinly lobster fishery (*Panulirus argus*), which is carried out within their respective TURFs.
+These communities also target finfish in the off season, mainly snappers (Lutjanidae) and groupers (Serranidae).
+ME established eight marine reserves in 2012, and PH established four marine reserves in 2013.
+All these reserves are legally recognized as Fishing Refugia.
 
@@
   library(tmap)
-  library(MPAtools)
   library(magrittr)
@@
 
-```{r, fig.cap = "\\label{fig:map}Mapa de la localización general de las comunidades de estudio. El panel de la derecha es un acercamiento a las comunidades de María Elena y Punta Herrero."}
+```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A).
+Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are located in the yucatan Peninsula."}
 # Define reference map
@@
 
-plot_grid(mex, ph, pbc, ma, labels = c("A", "B", "C", "D"), ncol = 2)
+plot_grid(mex, ph, pbc, ma, labels = c("A", "C", "B", "D"), ncol = 2)
 ```
@@
 ## Data collection
-
-Para evaluar las reservas, utilizamos tres fuentes de información.
-La información ecológica proviene de los monitoreos ecológicos realizados anualmente en las zonas reserva y control.
-Cada año, se realizan censos visuales para evaluar las comunidades de peces e invertebrados, registrando riquezas, abundancias y tallas (en peces).
-Esta información nos permite calcular los indicadores biológicos de manera anual.
-Al tener valores de diferentes indicadores biológicos antes y después de la implementación de las reservas, para las zonas de reserva y sitios control, tenemos un diseño muestreal de Antes-Después-Control-Impacto.
-
-También incluimos información socioeconómica relevante proveniente de los avisos de arribo de CONAPESCA.
-En este caso, se tienen registros mensuales de los recursos aprovechados por las diferentes comunidades, en los que se reportan los arribos (Toneladas) y el valor de los arribos ($).
-La información se encuentra disponible para el periodo 2001 - 2014.
-Ya que las categorías registradas por CONAPESCA son amplias y existe un nivel de error, utilizamos únicamente los arribos reportados para langosta entera fresca a nivel de cooperativa.
-Los ingresos generados por arribos son ajustados por medio del índice de precio al consumidor.
-
-La información de gobernanza fue obtenida a nivel de comunidad, pidiendo a personas familiares con las comunidades que proveyeran la información necesaria.
-La información de gobernanza no es evaluada de manera cuantitativa.
-En su lugar, esta se interpreta de manera tal que podamos comprender qué decisiones, reglas y estructuras tienen un impacto en la reserva.
-
-## Data analysis
-
-Dada la similitud de objetivos entre las reservas, la evaluación se realiza con los mismos indicadores.
-En este caso, se utilizan 6 indicadores biológicos, 2 socioeconómicos y 5 de gobernanza (Tabla \ref{table:indicators}).
-Según la disponibilidad de datos, se calculó la densidad de las especies objetivo presentadas en la sección de descripción de las comunidades.
-El criterio de selección fue que cada especie debía tener, por lo menos, dos observaciones anuales para las zonas de reserva y control.
-
-```{r}
-data.frame(Category = c("Biological", "","","","","", "Socioeconomic", "", "Governance", "", "", "", ""),
-           Indicador = c("Índice de diversidad de shannon", "Riqueza", "Densidad", "Nivel trófico", "Biomasa", "Densidad de especies objetivo", "Ingresos por especies objetivo", "Arribos de especies objetivo", "Tipo de acceso a la pesquería", "Grado de pesca ilegal", "Procuración de la reserva", "Tipo de organización pesquera", "Edad de la reserva"), stringsAsFactors = F) %>%
-  knitr::kable(caption = "\\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.", format = "latex")
-```
 
@@
 
-### Biological
+To perform the evaluation of these reserves we use thre sources of information.
+Ecological data come from the annual ecological monitorings of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO "Comunidad y Biodiversidad" ([COBI](www.cobi.org.mx)).
+These monitorings record richness and abundances of fish and invertebrate species in the reservs and control sites.
+For fish census, size structures are also collected to derive biomass.
+We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had the same probability of being selected as reserves during the design phase.
+From all the reserves in these three communities, we use the ones that have data for reserve and control sites before and after the implementation of the reserve (n = 4) , thus providing us with a Before--After--Control--Impact (*i.e.* BACI) design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
+BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causaly attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
 
-Utilizando un análisis de diferencia en diferencias podemos estimar el efecto que la reserva tienen en los indicadores biológicos \citep{moland_2013-VP} con el uso de un modelo de regresión lineal múltiple:
+Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA).
+Data contain monthly lobster landings (Kg) and value (MXP) from 2000 to 2014.
+This information was aggregated by year, and economic values were adjusted by the Consumer Price Index \citep{oecd_2017-VV} via Eq \ref{eqn:cpi}.
 
-$$I = \alpha_{i} + \gamma_{it} Year + \beta_{i} Zona + \lambda_{it} Year\times Zona + \sigma_{j} + \Pi_{k} + \epsilon$$
+\begin{equation}
+I_t = RI_t\times\frac{CPI_t}{CPI_T}
+\label{eqn:cpi}
+\end{equation}
 
-En este caso, modelamos los años como factores, tomando como referencia el primer año en la serie de datos de cada comunidad.
-Modelar los años como factores reduce la estructura del modelo, y relaja el ajuste al no asumir una tendencia lineal entre años; es decir, el cambio observado entre 2006 - 2007 no deberá de ser igual al observado entre el 2009 - 2010.
-Incluimos también un término para la zona, en la que la variable toma un valor de 0 si el sitio es una zona control y de 1 si es una zona de reserva.
-Finalmente, incluimos un término de interacción entre la variable de Zona y el Año.
-En este modelo, $\lambda_i$ representa el efecto que la reservas tuvo sobre un indicador en cada año y con respecto a los sitios control.
-El termino $\Omega$ captura efectos fijos por especies y por sitio.
+Where $I_t$ represents the adjusted income for year $t$ as the product between the reported income for that year and the ratio between the consumer price index in that year ($CPI_t$) to the most recent year’s consumer price index ($CPI_T$).
+
+**Governance data were collected at the community--level, using key informants to collect the necesary information on X, Y, and Z. Algo mas de Ere.**
+
+## Data analysis
+
+Following a framework that relates reserve objectives to performance indicators \citep{Villasenor-Derbez_2018}, we use five biological, two socioeconomic, and five governance indicators to evaluate these marine reserves Table \ref{table:indicators}.
+
+```{r}
+tibble(Category = c("Biological", "","","","", "Socioeconomic", "", "Governance", "", "", "", ""),
+           Indicador = c("Abundance", "Richness", "Shannon's diversity index", "Biomass", "Abundance of target species (lobsters)", "Income from target species", "Landings from target species", "Type of access to the fishery", "Perceived degree of illegal fishing", "Reserve surveilance anf enforcement", "Type of fishing organization", "Age of the reserve")) %>%
+  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex") %>% 
+  kable_styling(latex_options = "HOLD_position")
+```
+
+Biological indicators are analyzed with a difference--in--differences analysis (Eq \ref{eqn:reg_bio}), which allows us to estimate the effect that the reserve has on the biological indicators by comparing trends across time and space (*i.e.* reserve and control sites) \citep{moland_2013-VP,Villasenor-Derbez_2018}.
+The analysis is performed with generalized linear models of the form:
+
+\begin{equation}
+I = \alpha_{i} + \gamma_{it} Year_t + \beta Zone_i + \lambda_{it} Year_t\times Zone_i + \sigma_jSpp_j + \Pi_{k}Site_k + \epsilon
+\label{eqn:reg_bio}
+\end{equation}
+
+Where year--fixed effects are represented by $\gamma_{it} Year_t$, $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites, and the interaction term $\lambda_{it} Year_t\times Zone_i$ represents the average treatment effect for each year.
+When evaluating biomass and abundances, we include species--fixed effects ($\sigma_j$), and when more than one reserve is present we also include site--fixed effects ($\Pi_k$).
+For abundances (*i.e.* count data) the model is estimated with a quasipoisson error distribution.
+
+Socioeconomic indicators are evaluated with a similar approach (Eq \ref{eqn:soc_reg}), where landings and income before and after the implementation of the reserve are compared:
+
+\begin{equation}
+I = \beta_0 + \beta_1Post
+\label{eqn:soc_reg}
+\end{equation}
+
+This approach does not allow for a causal attribution of the observed changes to the reserve, but still allows us to draw important information that can inform our conclusions.
+For both approaches, model coefficients are estimated via ordinary least--squares and heteroskedastic--robust standard errors \citep{zeileis_2004-7n}.
+
+### Governance
+
+**Texto de ere y los SES**
 
@@
 
-data("species_bio")
+species_bio <- MPAtools::species_bio
 
@@
 ```{r model_fit}
-
 # Fish
@@
 ```
-
-### Socioeconomic
-
-El análisis de datos socioeconómicos se aplicó únicamente a Isla Natividad, María Elena y Punta Herrero, siguiendo un modelo con la forma:
-
-$$I = \beta_0 + \beta_1Post$$
-
-Que nos permite comparar el cambio en el promedio de los indicadores antes (Post = 0) y después (Post = 1) de la implementación de la reserva.
-Tanto para los indicadores biológicos como los socioeconómicos, los coeficientes fueron ajustador con el estimador de muestras heterocedásticas.
-
-### Governance
 

### 2018-02-11 Frontiers_Manuscript/Villasenor-Derbez.tex
Added english text for intro and methods

@@
 
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[section]
+\theoremstyle{definition}
+\newtheorem{exercise}{Exercise}[section]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+\newtheorem*{solution}{Solution}
 \begin{document}
@@
 
-Nice abstract here
+Abstract here
 
@@
  \keyFont{ \section{Keywords:} Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen
-Science, Mexico.}
+Science, Mexico, Social--Ecological Systems}
 
@@
 
-Overfishing and unsustainable fishing practices threaten marine
-ecosystems around the world \citep{halpern_2008-dK,halpern_2017-Zi}. The
-implementation of marine reserves (\emph{i.e.} areas where all fishing
-activities are off--limits; MRs) is a common management approach used to
-manage the spatial distribution of fishing effort and recover stocks
-\citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}. Recent research
-poses that marine reserves can also help buffer climate change
-\citep{roberts_2017-J9}, provide a refuge to environmental variability
-\citep{micheli_2012-EU}, help reduce by-catch \citep{hastings_2017-sm}
-and, in general, increase biomass, richness and densities of organisms
-within the protected regions
-\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}. However, the
-extent to which MRs yield benefits to fisheries remains relatively
-unexplored \citep{krueck_2017-J1} and the focus is often only in the
-social and economic dimensions \citep{klein_2008,charles_2008}, rarely
-providing a hollistic evaluation of the social-ecological system
-\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. Here, we combine
-causal inference techniques \citep{depalma_2018} and the
-social-ecological systems framework
-\citep{ostrom_2009-hg,mascia_2017-m_} to provide a comprehensive
-evaluation of four community-based marine reserves in three coastal
-communities in Mexico.
+Marine ecosystems around the world sustain significant impacts due to
+overfishing and unsustainable fishing practices
+\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. A common approach to
+manage the spatial distribution of fishing effort and recover stocks is
+through the implementation of marine reserves (\emph{i.e.} areas where
+all fishing activities are off--limits; MRs)
+\citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}.
 
-En México, las reservas marinas han sido comúnmente establecidas como
-zonas núcleo dentro de Reservas de la Biósfera (RBs), administradas por
-la Comisión Nacional de Áreas Naturales Protegidas (CONANP). Al día de
-hoy, 36 RBs protegen una porción del ambiente marino en México. Sin
-embargo, solamente 26 de estas incluyen (pequeñas) zonas núcleo donde
-las actividades pesqueras están prohibidas. Aunque la CONANP ha hecho
-esfuerzos importantes por involucrar a los actores durante la
-implementación de las reservas, esto aún se caracteriza por un proceso
-descendente, el cual conlleva a la falta de cumplimiento por parte de
-los actores. La escasez de recursos monetarios y humanos de la limitan
-también el monitoreo y vigilancia de las reservas, y a su vez, el
-desempeño de la reserva.
+Marine reserve science has largely focused on understanding the
+ecological effects of these areas, which include increased biomass,
+richness, and densities of organisms within the protected regions
+\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, climate change
+mitigation \citep{roberts_2017-J9}, and protection from environmental
+variability \citep{micheli_2012-EU}. However, there is considerably less
+literature focusing on the relationship between socioeconomic and
+governance structures and their relationship to ecological
+effectriveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or
+benefits to fisheries \citep{krueck_2017-J1}; evaluations of marine
+reserves rarely provide a hollistic view of the social-ecological system
+\citep{lpezangarita_2014}. Here, we combine causal inference techniques
+\citep{depalma_2018} and the social-ecological systems framework
+\citep{ostrom_2009-hg} to provide a comprehensive ecological and
+socioeconomic evaluation of four community-based marine reserves in
+three coastal communities in Mexico.
 
-Buscando promover una alternativa con procesos ascendentes para
-implementar reservas marinas, las Organizaciones de la Sociedad Civil
-(OSCs) comenzaron a trabajar con comunidades pesqueras para establecer
-reservas comunitarias \citep{uribe_2010-u2} . Estas son comúnmente
-establecidas dentro de zonas de concesión, una forma de derechos de uso
-territoriales para pesquerías (TURF, en inglés). Al permitir a los
-pescadores diseñar sus propias reservas, una mayor proporción de la
-comunidad está de acuerdo con los perímetros y reglas establecidas, y
-por lo tanto los respetan
-\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8} .
-Adicionalmente, los pescadores pueden implementar sus reservas por un
-periodo acordado (usualmente cinco años), después del cual la reserva
-puede ser abierta a la pesca. Esto provee a los pescadores con un
-sentido de confianza de que, en caso de ser necesario, aún tienen acceso
-a pescar esa zona. Las reservas son directamente vigiladas y
-monitoreadas por la comunidad, quienes comúnmente utilizan pequeñas
-embarcaciones (\emph{e.g.} pangas) para patrullar la zona, o realizan
-avistamientos desde la costa en búsqueda de pescadores ilegales Aún así,
-las reservas comunitarias carecen de reconocimiento legal; por lo tanto,
-no hay forma de penalizar a los infractores.
+Marine Reserves in Mexico have been commonly implemented as ``core
+zones'' within Biosphere Reserves (BRs) that are administered by the
+National Commision of Protected Areas (\emph{Comisión Nacional de Áreas
+Marinas Protegidas}, CONANP). While CONANP has made efforts to have a
+participatory process, the implementation of these zones is still
+characterized by top--down approaches. This motivated Civil Society
+Organizations (CSOs) to work with coastal communities to implement
+community--based marine reserves \citep{uribe_2010-u2}, which are
+usually established within a Territorial Use Rights for Fisheries
+(TURFs); thus making them TURF--reserves \citep{afflerbach_2014-HP}.
+This bottom--up approach allows fishers to disign their own reserves,
+which increases compliance and self--enforcement
+\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
+these reserves still lack legal recognition, making them vulnerable to
+poaching. In 2014, a new norm \citep{nom} allowed fishers to request the
+legal recognition of a community--based reserve under the name of
+``Fishing Refugia'' (\emph{Zona de Refugio Pesquero}, FR). This new norm
+thus combines bottom--up approaches to design marine reserves, along
+with a legal recognition of the management intervention. Since then,
+\emph{39} FR have been implemented along the Pacific, Gulfof
+Caloifornia, and Mexican Caribbean coastlines, but no formal evaluation
+of their effectiveness has taken place.
 
-Sin embargo, en el 2014 una nueva norma \citep{nom} permite a los
-pescadores solicitar el establecimiento de reservas marinas bajo el
-nombre de ``Zonas de refugio Pesquero'' (ZRP). El manejo de las ZRP
-combina procesos ascendentes y descendentes al reconocer legalmente las
-reservas propuestas por las comunidades. Posterior a la revisión por
-parte de la Comisión Nacional de Acuacultura y Pesca (CONAPESCA) y la
-opinión técnica del Instituto Nacional de Acuacultura y Pesca (INAPESCA)
-las ZRP son establecidas por el periodo solicitado por los pescadores.
-El monitoreo y la vigilancia de las ZRP es típicamente llevado a cabo
-por la comunidad , con ayuda de OSCs locales. Hasta este cambio
-regulatorio, las reservas comunitarias no contaban con el soporte legal,
-y eran solamente reconocidas por la comunidad. Al día de hoy, existen 39
-ZRP establecidas en el Pacífico, Golfo de California y Caribe Mexicano.
+While there are ecological factors defining the success of a MR
+(\emph{i.e.} habitat representation, initial state of protection,
+connectivity to other protected areas), their effectivness also depends
+on the socioeconomic and governance settings under which they are
+implemented. Literature shows that many non-ecological characteristics
+can play an equally important role in the effectivenes of MRs. For
+example, age of a reserve (\emph{i.e.} time since its implementation),
+size, and habitat contained were key to the effectiveness of MRs in
+Palau \citep{friedlander_2017-oI}. In the Mediterranean,
+\citet{difranco_2016-Xw} identify that surveilance and enfocrement,
+presence of a management plan, and involvement of fishers in management
+and decision--making along with promotion of sustainable fishing
+practices were the key fators that increased stock health and income to
+fishers. At a global level, \citet{edgar_2014-UO} indicate that
+enforcement, age, size, and isolation were important factors determining
+effectiveness of the reserves.
 
-Aunque existen tres aproximaciones generales para implementar reservas
-marinas en México (\emph{i.e.} Zonas núcleo dentro de AMP, reservas
-comunitarias y Zonas de Refugio Pesquero), aún no comprendemos a fondo
-las características sociales que permiten su efectividad. La ciencia de
-reservas marinas se ha enfocado ampliamente en los efectos biológicos
-que estas tienen
-\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,afflerbach_2014-HP,krueck_2017-J1}.
-Aunque el aspecto ecológico de las reservas es importante para su éxito,
-su efectividad también depende del estado socioeconómico y los sistemas
-de gobernanza de las comunidades pesqueras.
-
-La literatura indica que diferentes características influyen en el éxito
-de una reserva. En Palau, por ejemplo, la edad (\emph{i.e.} tiempo
-transcurrido desde implementación), tamaño y hábitat contenido son
-características claves que determinan la efectividad
-\citep{friedlander_2017-oI}. Por otro lado, en el Mar Mediterráneo,
-\citet{difranco_2016-Xw} identifican que la procuración y vigilancia,
-presencia de un plan de manejo, participación de pescadores en el
-manejo, representación de pescadores en la toma de decisiones y
-promoción de la pesca sustentable son los cinco factores que incrementan
-la salud de los stocks y el ingreso económicos a los pescadores, a la
-vez que se presenta una mayor aceptación social de las prácticas de
-manejo. En una aproximación global, \citet{edgar_2014-UO} encuentran que
-la procuración, edad, tamaño y aislamiento son determinantes de la
-efectividad de las reservas. Por lo tanto, observamos que las
-características que habilitan el éxito varían a través de regiones, y
-poco esfuerzo se ha hecho por comprender estas interacciones en México.
-
-El objetivo de este trabajo este trabajo es realizar una evaluación de
-la efectividad de reservas marinas en México, presentando resultados de
-cinco comunidades costeras como caso de estudio. Con el fin de obtener
-una visión holística del sistema, la evaluación se realizará tomando en
-cuenta indicadores biológicos, socioeconómicos y de gobernanza. La
-evaluación de éstos cinco casos de estudios nos permitirá identificar la
-manera en que las características socioeconómicas y de gobernanza se
-relacionan con la efectividad (biológica) de las reservas marinas
-evaluadas. Los patrones identificados podrán utilizarse para informar la
-toma de decisiones para la implementación de la red de reservas marinas
-en la Región de las Grandes Islas del Golfo de California.
+The objective of this work is twofold: i) Provide the first evaluation
+of community--based marine reserves in Mexico, and ii) provide a
+comprehensive evaluation of the social--ecological system to identify
+how sociocoenomic and governance characteristics relate to ecological
+effectiveness. With the purpose of providing a hollistic evaluation, we
+combine ecological, socioeconomic, and governance indicators. We use
+causal inference techniques to provide a measurement of the effect of
+the management intervention, and combine it with the social--ecological
+systems framework \citep{ostrom_2009-hg}.
 
@@
 
-Las comunidades utilizadas en este reporte se distribuyen a lo largo de
-la costa Pacífica de Baja California (n = 1) y el Sistema Arrecifal
-Mesoamericano (n = 2; Fig \ref{fig:map}).
+We focus our evaluation in three coastal communities from the Pacific
+coast of Baja California (n = 1) and the Mesoamerican Reef System (n =
+2; Fig \ref{fig:map}). Isla Natividad (IN) lies west of the Baja
+California Peninsula (Fig \ref{fig:map}B), where kelp forests
+(\emph{Macrocystis pyrifera}) and rocky reefs are the predominant
+habitats. The island is home to a fishing cooperative (\emph{Sociedad
+Cooperative de Producción Pesquera Buzos y Pescadores de la Baja
+California SCL}), that holds a TURF for spiny lobster (\emph{Panulirus
+interruptus}). However, other resources like finfish (yellowtail jack,
+\emph{Seriola lalandi}), sea cucumber (\emph{Parastichopus
+parvimensis}), red sea urchin (\emph{Mesocentrotus franciscanus}), snail
+(\emph{Megastraea turbanica} y \emph{M. undosa}), and abalone
+(\emph{Haliotis spp}, until 2010) are also important sources of income.
+In 2006, the community decided to implement two community--based marine
+reserves within their fishing grounds, seeking to recover depleted
+stocks of invertebrate species (mainly lobster and abalone). Until
+today, these reserves are yet to be legally recognized as Fishing
+Refugia.
 
-\subsubsection{Isla Natividad}\label{isla-natividad}
-
-La Isla Natividad se encuentra en la costa oeste de la Península de Baja
-California, donde el hábitat predominante es el bosque de kelp o sargazo
-gigante (\emph{Macrocystis pyrifera}) y los arrecifes rocosos. En la
-isla, la Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y
-Pescadores de la Baja California SCL realiza actividades de extracción
-de los recursos marinos. Aunque la langosta roja (\emph{Panulirus
-interruptus}) es la especie más importante en términos económicos, otras
-especies importantes incluyen la escama (con un enfoque en Jurel;
-\emph{Seriola lalandi}), el pepino de mar (\emph{Parastichopus
-parvimensis}), el erizo rojo (\emph{Mesocentrotus franciscanus}), el
-caracol (\emph{Megastraea turbanica} y \emph{M. undosa}) y, hasta el
-2010, el abulón (\emph{Haliotis sp.}). En 2006, por medio de un proceso
-participativo, la cooperativa decidió establecer dos reservas marinas de
-manera voluntaria. Agentes externos a la cooperativa, como personal de
-Comunidad y Biodiversidad A.C., académicos de la Universidad de
-Stanford, y personal de la CONANP (de la oficina de Reserva de la
-Biósfera El Vizcaíno), también participaron en el diseño e
-implementación de las reservas. Las reservas fueron establecidas como
-instrumento de manejo pesquero, buscando recuperar las poblaciones de
-abulón y otros invertebrados. Al día de hoy, las reservas marinas de
-Isla Natividad no han recibido reconocimiento legal, pero la cooperativa
-ha mostrado interés por reconocerlas como Zonas de Refugio Pesquero
-(ZRP). Los pescadores tienen un sistema de turnos para vigilar la
-reserva día y noche desde embarcaciones patrulla.
-
-\subsubsection{María Elena}\label{maria-elena}
-
-María Elena es una comunidad pesquera en la costa de Quintana Roo. Los
-arrecifes coralinos y manglares son los principales ecosistemas
-representados en la zona. El campo pesquero es utilizado por pescadores
-de la SCPP Cozumel scl (de la Isla de Cozumel). La principal especie
-aprovechada por ésta organización es la langosta espinosa del caribe
-(\emph{Panulirus argus}). La cooperativa cuenta con permiso de pesca de
-escama y concesión de langosta. En el 2012, la Cooperativa, en conjunto
-con la Alianza Kanan Kay, COBI, CONANP, CONAPESCA, Oceanus, Fundación
-Claudia y Roberto Hernández, Fundación Haciendas del Mundo Maya,
-establecieron ocho ZRP con una vigencia de cinco años. La vigilancia de
-las reservas se realiza por medio del equipo de vigilancia comunitaria,
-con apoyo de la CONANP y una embarcación -donada por COBI- utilizada
-para realizar recorridos frecuentes.
-
-\subsubsection{Punta Herrero}\label{punta-herrero}
-
-La comunidad de Punta Herrero se encuentra aproximadamente a 15 km al
-sur del campo pesquero de María Elena. De igual manera, los arrecifes
-coralinos y manglares son los principales ecosistemas representados en
-la zona, y la principal especie explotada es la langosta. Sin embargo,
-la SCPP José María Azcorra también cuenta con permisos para pesca de
-escama y tiburón y una concesión de langosta. En una réplica del
-ejercicio realizado en María Elena -con presencia de los mismos
-actores-, cuatro ZRP fueron establecidas en el 2013, con una vigencia de
-cinco años. El equipo de vigilancia comunitaria, con apoyo de la CONANP,
-se encarga de la vigilancia de las reservas.
+The other two communities are Maria Elena (ME; Fig \ref{fig:map}C) and
+Punta Herrero (PH; Fig \ref{fig:map}D) in the Yucatan Peninsula, where
+coral reefs and mangroves are the representative coastal ecosystems. ME
+is a fishing camp --visited intermitently during the fishing season--
+belonging to the Cozumel fishing cooperative. PH is home to the ``José
+María Azcorra'' fishing cooperative. The main source of income to both
+communities is the caribbean spinly lobster fishery (\emph{Panulirus
+argus}), which is carried out within their respective TURFs. These
+communities also target finfish in the off season, mainly snappers
+(Lutjanidae) and groupers (Serranidae). ME established eight marine
+reserves in 2012, and PH established four marine reserves in 2013. All
+these reserves are legally recognized as Fishing Refugia.
 
@@
 \includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Mapa de la localización
-general de las comunidades de estudio. El panel de la derecha es un
-acercamiento a las comunidades de María Elena y Punta Herrero.}
+\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Location of the three
+coastal communities studied (A). Isla Natividad (B) is located off the
+Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are
+located in the yucatan Peninsula.}
 \end{figure}
@@
 
-Para evaluar las reservas, utilizamos tres fuentes de información. La
-información ecológica proviene de los monitoreos ecológicos realizados
-anualmente en las zonas reserva y control. Cada año, se realizan censos
-visuales para evaluar las comunidades de peces e invertebrados,
-registrando riquezas, abundancias y tallas (en peces). Esta información
-nos permite calcular los indicadores biológicos de manera anual. Al
-tener valores de diferentes indicadores biológicos antes y después de la
-implementación de las reservas, para las zonas de reserva y sitios
-control, tenemos un diseño muestreal de Antes-Después-Control-Impacto.
+To perform the evaluation of these reserves we use thre sources of
+information. Ecological data come from the annual ecological monitorings
+of reserve and control areas, carried out by members from each community
+and personnel from the Mexican CSO ``Comunidad y Biodiversidad''
+(\href{www.cobi.org.mx}{COBI}). These monitorings record richness and
+abundances of fish and invertebrate species in the reservs and control
+sites. For fish census, size structures are also collected to derive
+biomass. We define control sites as regions with habitat characteristics
+similar to the corresponding reserves, and that presumably had the same
+probability of being selected as reserves during the design phase. From
+all the reserves in these three communities, we use the ones that have
+data for reserve and control sites before and after the implementation
+of the reserve (n = 4) , thus providing us with a
+Before--After--Control--Impact (\emph{i.e.} BACI) design that allows us
+to capture and control for temporal and spatial dynamics
+\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
+techniques have proven effective to evaluate marine reserves, as they
+allow us to causaly attribute observed changes to the intervention
+\citep{moland_2013-VP,Villasenor-Derbez_2018}.
 
-También incluimos información socioeconómica relevante proveniente de
-los avisos de arribo de CONAPESCA. En este caso, se tienen registros
-mensuales de los recursos aprovechados por las diferentes comunidades,
-en los que se reportan los arribos (Toneladas) y el valor de los arribos
-(\$). La información se encuentra disponible para el periodo 2001 -
-2014. Ya que las categorías registradas por CONAPESCA son amplias y
-existe un nivel de error, utilizamos únicamente los arribos reportados
-para langosta entera fresca a nivel de cooperativa. Los ingresos
-generados por arribos son ajustados por medio del índice de precio al
-consumidor.
+Socioeconomic data come from landing receipts reported to the National
+Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
+Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
+(Kg) and value (MXP) from 2000 to 2014. This information was aggregated
+by year, and economic values were adjusted by the Consumer Price Index
+\citep{oecd_2017-VV} via Eq \ref{eqn:cpi}.
 
-La información de gobernanza fue obtenida a nivel de comunidad, pidiendo
-a personas familiares con las comunidades que proveyeran la información
-necesaria. La información de gobernanza no es evaluada de manera
-cuantitativa. En su lugar, esta se interpreta de manera tal que podamos
-comprender qué decisiones, reglas y estructuras tienen un impacto en la
-reserva.
+\begin{equation}
+I_t = RI_t\times\frac{CPI_t}{CPI_T}
+\label{eqn:cpi}
+\end{equation}
+
+Where \(I_t\) represents the adjusted income for year \(t\) as the
+product between the reported income for that year and the ratio between
+the consumer price index in that year (\(CPI_t\)) to the most recent
+year's consumer price index (\(CPI_T\)).
+
+\textbf{Governance data were collected at the community--level, using
+key informants to collect the necesary information on X, Y, and Z. Algo
+mas de Ere.}
 
@@
 
-Dada la similitud de objetivos entre las reservas, la evaluación se
-realiza con los mismos indicadores. En este caso, se utilizan 6
-indicadores biológicos, 2 socioeconómicos y 5 de gobernanza (Tabla
-\ref{table:indicators}). Según la disponibilidad de datos, se calculó la
-densidad de las especies objetivo presentadas en la sección de
-descripción de las comunidades. El criterio de selección fue que cada
-especie debía tener, por lo menos, dos observaciones anuales para las
-zonas de reserva y control.
+Following a framework that relates reserve objectives to performance
+indicators \citep{Villasenor-Derbez_2018}, we use five biological, two
+socioeconomic, and five governance indicators to evaluate these marine
+reserves Table \ref{table:indicators}.
 
-\begin{table}
+\begin{table}[H]
 
-\caption{\label{tab:unnamed-chunk-2}\label{table:indicators}Lista de indicadores utilizados para evaluar resvas marinas, agrupados por tipo.}
+\caption{\label{tab:unnamed-chunk-2}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
 \centering
@@
 \hline
-Biological & Índice de diversidad de shannon\\
+Biological & Abundance\\
 \hline
- & Riqueza\\
+ & Richness\\
 \hline
- & Densidad\\
+ & Shannon's diversity index\\
 \hline
- & Nivel trófico\\
+ & Biomass\\
 \hline
- & Biomasa\\
+ & Abundance of target species (lobsters)\\
 \hline
- & Densidad de especies objetivo\\
+Socioeconomic & Income from target species\\
 \hline
-Socioeconomic & Ingresos por especies objetivo\\
+ & Landings from target species\\
 \hline
- & Arribos de especies objetivo\\
+Governance & Type of access to the fishery\\
 \hline
-Governance & Tipo de acceso a la pesquería\\
+ & Perceived degree of illegal fishing\\
 \hline
- & Grado de pesca ilegal\\
+ & Reserve surveilance anf enforcement\\
 \hline
- & Procuración de la reserva\\
+ & Type of fishing organization\\
 \hline
- & Tipo de organización pesquera\\
-\hline
- & Edad de la reserva\\
+ & Age of the reserve\\
 \hline
@@
 
-\subsubsection{Biological}\label{biological}
+Biological indicators are analyzed with a difference--in--differences
+analysis (Eq \ref{eqn:reg_bio}), which allows us to estimate the effect
+that the reserve has on the biological indicators by comparing trends
+across time and space (\emph{i.e.} reserve and control sites)
+\citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis is performed
+with generalized linear models of the form:
 
-Utilizando un análisis de diferencia en diferencias podemos estimar el
-efecto que la reserva tienen en los indicadores biológicos
-\citep{moland_2013-VP} con el uso de un modelo de regresión lineal
-múltiple:
+\begin{equation}
+I = \alpha_{i} + \gamma_{it} Year_t + \beta Zone_i + \lambda_{it} Year_t\times Zone_i + \sigma_jSpp_j + \Pi_{k}Site_k + \epsilon
+\label{eqn:reg_bio}
+\end{equation}
 
-\[I = \beta_0 + \sum \gamma Year + \beta_1 Zona + \sum \lambda Year\times Zona + \Omega + \epsilon\]
+Where year--fixed effects are represented by \(\gamma_{it} Year_t\),
+\(\beta Zone_i\) captures the difference between reserve (\(Zone = 1\))
+and control (\(Zone = 0\)) sites, and the interaction term
+\(\lambda_{it} Year_t\times Zone_i\) represents the average treatment
+effect for each year. When evaluating biomass and abundances, we include
+species--fixed effects (\(\sigma_j\)), and when more than one reserve is
+present we also include site--fixed effects (\(\Pi_k\)). For abundances
+(\emph{i.e.} count data) the model is estimated with a quasipoisson
+error distribution.
 
-En este caso, modelamos los años como factores, tomando como referencia
-el primer año en la serie de datos de cada comunidad. Modelar los años
-como factores reduce la estructura del modelo, y relaja el ajuste al no
-asumir una tendencia lineal entre años; es decir, el cambio observado
-entre 2006 - 2007 no deberá de ser igual al observado entre el 2009 -
-2010. Incluimos también un término para la zona, en la que la variable
-toma un valor de 0 si el sitio es una zona control y de 1 si es una zona
-de reserva. Finalmente, incluimos un término de interacción entre la
-variable de Zona y el Año. En este modelo, \(\lambda_i\) representa el
-efecto que la reservas tuvo sobre un indicador en cada año y con
-respecto a los sitios control. El termino \(\Omega\) captura efectos
-fijos por especies y por sitio.
+Socioeconomic indicators are evaluated with a similar approach (Eq
+\ref{eqn:soc_reg}), where landings and income before and after the
+implementation of the reserve are compared:
 
-\subsubsection{Socioeconomic}\label{socioeconomic}
+\begin{equation}
+I = \beta_0 + \beta_1Post
+\label{eqn:soc_reg}
+\end{equation}
 
-El análisis de datos socioeconómicos se aplicó únicamente a Isla
-Natividad, María Elena y Punta Herrero, siguiendo un modelo con la
-forma:
-
-\[I = \beta_0 + \beta_1Post\]
-
-Que nos permite comparar el cambio en el promedio de los indicadores
-antes (Post = 0) y después (Post = 1) de la implementación de la
-reserva. Tanto para los indicadores biológicos como los socioeconómicos,
-los coeficientes fueron ajustador con el estimador de muestras
-heterocedásticas.
+This approach does not allow for a causal attribution of the observed
+changes to the reserve, but still allows us to draw important
+information that can inform our conclusions. For both approaches, model
+coefficients are estimated via ordinary least--squares and
+heteroskedastic--robust standard errors \citep{zeileis_2004-7n}.
 
 \subsubsection{Governance}\label{governance}
+
+\textbf{Texto de ere y los SES}
 

### 2018-02-13 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Updated manuscript with results section

@@
 zrp <- st_read(dsn = "../Data/Spatial", layer = "ZRP", quiet = T) %>% 
-  filter(Name %in% c("El Cabezo", "El Faro", "Gallineros")) %>% 
+  filter(Name %in% c("El Cabezo", "El Faro")) %>% 
   st_transform(crs = "+proj=longlat +datum=NAD27")
@@
   geom_sf(data = coastline_mx) +
-  lims(x = c(-87.5, -87.4), y = c(19.280990, 19.35)) +
+  lims(x = c(-87.5, -87.4), y = c(19.29, 19.36)) +
   geom_sf(data = zrp, fill = "red") +
@@
 
-plot_grid(mex, ph, pbc, ma, labels = c("A", "C", "B", "D"), ncol = 2)
+plot_grid(mex, ma, pbc, ph, labels = c("A", "C", "B", "D"), ncol = 2)
 ```
@@
   filter(!is.na(RC),
-         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)", "El Faro - El Faro (Control)", "Anegado de Chal - Anegado de Chal (Control)")) %>% 
+         RC %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
+                   "Cabezo - Cabezo (Control)",
+                   "Manchon - Manchon (Control)",
+                   "El Faro - El Faro (Control)")) %>% 
+  filter(!(RC %in% c("El Faro - El Faro (Control)") & Ano == 2012)) %>% 
   mutate(GeneroEspecie = case_when(
@@
     GeneroEspecie == "Vieja spp" ~ "Bodianus rufus",
-    # GeneroEspecie == "" ~ "",
     TRUE ~ GeneroEspecie
@@
   filter(!is.na(RC),
-         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)", "El Faro - El Faro (Control)", "Anegado de Chal - Anegado de Chal (Control)"))
+         RC %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
+                   "Cabezo - Cabezo (Control)",
+                   "Manchon - Manchon (Control)",
+                   "El Faro - El Faro (Control)")) %>% 
+  filter(!(RC %in% c("El Faro - El Faro (Control)") & Ano == 2012))
 
@@
 We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had the same probability of being selected as reserves during the design phase.
-From all the reserves in these three communities, we use the ones that have data for reserve and control sites before and after the implementation of the reserve (n = 4) , thus providing us with a Before--After--Control--Impact (*i.e.* BACI) design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
+From all the reserves in these three communities, we use the ones that have data for reserve and control sites before and after the implementation of the reserve.
+This provides us with a Before--After--Control--Impact (*i.e.* BACI) design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
 BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causaly attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
+All reserves were surveyed annualy from at least one year before implementation until 2016.
+Table \ref{tab:com_sum} shows a summary of the number of reserves, year of implementation, and number of transects for each reserve.
+
+```{r}
+inv_summary <- group_by(invert, Comunidad, RC, Zona, Sitio, Transecto, Ano) %>%
+  count() %>% 
+  ungroup() %>% 
+  group_by(Comunidad, RC, Zona) %>% 
+  summarize(n = n()) %>% 
+  ungroup() %>% 
+  spread(Zona, n) %>% 
+  select(Comunidad, RC, Control, Reserve = Reserva)
+
+group_by(fish, Comunidad, RC, Zona, Sitio, Transecto, Ano) %>%
+  count() %>% 
+  ungroup() %>% 
+  group_by(Comunidad, RC, Zona) %>% 
+  summarize(Start = min(Ano), n = n()) %>% 
+  ungroup() %>% 
+  spread(Zona, n) %>% 
+  mutate(Implementation = c(2006, 2012, 2013, 2013)) %>% 
+  select(Comunidad, RC, Implementation, Start, Control, Reserve = Reserva) %>% 
+  left_join(inv_summary, by = c("Comunidad", "RC")) %>% 
+  knitr::kable(format = "latex",
+               caption = "\\label{tab:com_sum} Summary of commuity--based marine reserves by community.
+Imp = Year of implementation, Start = Year of first sampling, number of fish transects in control (Cf) and reserve (Rf) sites, and number of invertebrate transects in Control (Ci) and Reserve (Ri) sites.",
+               col.names = c("Community", "Reserve - Control", "Imp", "Start", "Cf", "Rf", "Ci", "Ri")) #%>% 
+  # kableExtra::add_header_above(header = c(" " = 4, "Fish" = 6, "Invert" = 8))
+```
 
@@
 
-Biological indicators are analyzed with a difference--in--differences analysis (Eq \ref{eqn:reg_bio}), which allows us to estimate the effect that the reserve has on the biological indicators by comparing trends across time and space (*i.e.* reserve and control sites) \citep{moland_2013-VP,Villasenor-Derbez_2018}.
+Biological indicators are analyzed with a difference--in--differences analysis (Eq \ref{eqn:reg_bio}), which allows us to estimate the effect that the reserve has on the biological indicators by comparing trends across time and treatments \citep{moland_2013-VP,Villasenor-Derbez_2018}.
 The analysis is performed with generalized linear models of the form:
@@
 \begin{equation}
-I = \alpha_{i} + \gamma_{it} Year_t + \beta Zone_i + \lambda_{it} Year_t\times Zone_i + \sigma_jSpp_j + \Pi_{k}Site_k + \epsilon
+I_i = \alpha_{i} + \gamma_{it} Year_t + \beta Zone_i + \lambda_{it} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon
 \label{eqn:reg_bio}
@@
 
-Where year--fixed effects are represented by $\gamma_{it} Year_t$, $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites, and the interaction term $\lambda_{it} Year_t\times Zone_i$ represents the average treatment effect for each year.
-When evaluating biomass and abundances, we include species--fixed effects ($\sigma_j$), and when more than one reserve is present we also include site--fixed effects ($\Pi_k$).
-For abundances (*i.e.* count data) the model is estimated with a quasipoisson error distribution.
+Where year--fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
+The interaction term $\lambda_{it} Year_t\times Zone_i$ represents represent the mean change in the indicator inside the reserve, for year $t$, with respect to the first year of evaluation in the control site (See Table \ref{tab:com_sum}).
+When evaluating biomass and abundances, we include species--fixed effects ($\sigma_j$).
+For abundances and richness (*i.e.* count data) the model is estimated with a quasipoisson error distribution.
 
@@
 \begin{equation}
-I = \beta_0 + \beta_1Post
+I_i = \beta_0 + \beta_1Post
 \label{eqn:soc_reg}
@@
 **Texto de ere y los SES**
+
+\clearpage
 
@@
   abundance(com = "Maria Elena") %>% 
-  model(sfe = T, family = "quasipoisson")
+  model(sfe = F, family = "quasipoisson")
 
@@
   richness(com = "Isla Natividad") %>% 
-  model(sfe = F, spfe = F)
+  model(sfe = F, spfe = F, family = "quasipoisson")
 
@@
   richness(com = "Maria Elena") %>% 
-  model(sfe = T, spfe = F)
+  model(sfe = F, spfe = F, family = "quasipoisson")
 
@@
   richness(com = "Punta Herrero") %>% 
-  model(sfe = F, spfe = F)
+  model(sfe = F, spfe = F, family = "quasipoisson")
 
@@
   shannon(com = "Maria Elena") %>% 
-  model(sfe = T, spfe = F)
+  model(sfe = F, spfe = F)
 
@@
   biomass(com = "Maria Elena") %>% 
-  model(sfe = T)
+  model(sfe = F)
 
@@
   model(sfe = F)
-
-## Trophic level
-# T_IN <- fish %>% 
-#   trophic("Isla Natividad") %>% 
-#   model(sfe = F, spfe = F)
-# 
-# T_ME <- fish %>% 
-#   trophic("Maria Elena") %>% 
-#   model(spfe = F)
-# 
-# T_PH <- fish %>% 
-#   trophic("Punta Herrero") %>% 
-#   model(sfe = F, spfe = F)
 
@@
   abundance(com = "Maria Elena") %>% 
-  model(sfe = T, spfe = F, family = "quasipoisson")
+  model(sfe = F, spfe = F, family = "quasipoisson")
 
@@
   model(sfe = F, spfe = F, family = "quasipoisson")
-
 ```
@@
 
-A continuación se presentan los resultados de cada una de las comunidades.
-Los resultados biológicos se presentarán para cada comunidad, discutiendo primero los indicadores en común con otras comunidades (Shannon, Riqueza, Densidad, Nivel Trófico, Biomasa para peces e invertebrados) y, según su caso, las densidades de las especies objetivo.
-Habiendo presentado los resultados biológicos, nos enfocaremos después en los socioeconómicos y de gobernanza.
-Usaremos esta información para identificar las causas (sociales) del éxito (biológico) de las reservas.
+Our methodological approach with biological indicators allows us to make a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}.
+The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the reserve site at time $t$ after controling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
+Here we present the effect that marine reserves had on each of the biological indicators for each coastal community, along with the trends in socioeconomic indicators of lobster catches and revenues.
+We also provide an overview of the state of the socioeconomic and governance settings of each community, and discuss how these dimensions might be intertwined with each other.
+
+## Biological
+
+Effect sizes for biological indicators are shown in Figure \ref{fig:indicators}.
+Isla Natividad, the community with the oldest reserve, shows inconsistent effects across indicators and sources of data (*i.e.* fish vs.
+invertebrates).
+For example, the reserve had a small effect on fish abundances (Fig \ref{fig:indicators}A), where only year 2010 showed significat effect sizes in fish abundances ($p<0.05$) and all other years oscillated above and under zero ($p > 0.05$).
+However, in the case of invertebrate abundances (Fig \ref{fig:indicators}B), there was a clear positive effect of the reserve, as all but 1 year (2008) presented positive effect sizes relative to the control site ($p < 0.05$).
+Maria Elena and Punta Herrero showed no significant increase in fish and invertebrate abundances ($p< 0.05$), except for invertebrates in Punta Herrero for 2014 --right after the implementation of the reserves-- which showed a significant increase (*i.e.* $\lambda_{2014} = 2.5$, $p < 0.05$).
+Full tables with model coefficients are presented in the supplementary materials (**S1 Table**, **S2 Table**, **S3 Table**).
+
+While the number of fish species oscillated above and bellow zero through time for all reserves, none of these changes were statistically significant ($p > 0.05$) indicating that the reserves had no effect on fish species richness (Fig \ref{fig:indicators}C).
+For invertebrate species in Isla Natividad, all effect sizes were negative, but only significant for 2008, 2009, 2011, and 2014 ($p < 0.05$; Fig \ref{fig:indicators}D).
+For Maria Elena and Punta Herrero, the data do not show significant changes in invertebrate species richness ($p > 0.05$).
+
+Effect sizes for Shannon's diversity index for fish (Fig \ref{fig:indicators}E) in Isla Natividad oscillated between $\lambda_{2011} = -0.45$ and $\lambda_{2010} = -0.005$, but were not significantly different from null hypotheses of no change (*i.e.* $\lambda_t = 0$; $p > 0.05$).
+For invertebrates in that same community (Fig \ref{fig:indicators}F), Shannon's diversity index showed a significant decrease between 2008 and 2014, with largest decrease observed for 2011 ($\lambda_{2011} = -0.91$; $p < 0.05$).
+In the case of Maria Elena and Punta Herrero, Shannon's diversity index for fish showed increases in the order of $\lambda_t = 1$.
+For Maria Elena and Punta Herrero, these effects were only statistically significant for 2014, and 2014 and 2015 ($p < 0.05$).
+
+Biomass was only evaluated for fish data (Fig \ref{fig:indicators}G).
+In Isla Natividad, fish biomass presented a steady but small increase ($p>0.05$), and exhibited an increased variability in biomass between 2013 and 2016.
+Maria Elena and Punta Herrero also showed small, non-statistically significant increases in fish biomass ($p>0.05$).
+The last biological indicator is abundance of target species, *Panulirus interruptus* and *P.
+argus*, for the Pacific and Caribbean, respectively (Fig \ref{fig:indicators}H).
+Isla Natividad presented small constantly-positive effects but were not significantly different from the reference point of control site before the implementation of the reserve ($p > 0.05$.
+Maria Elena showed significant increases in lobster densities in the order of $\lambda_t = 10$ ($p < 0.05$).
+Finally, punta herrero presented alternating negative and positive effects, but these were not different from the baseline case ($p > 0.05$).
 
 ```{r, fig.width = 6.5, fig.height = 8, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for community-levell indicators.
-Plots are ordered by survey type (left: fish; right: invertebrates) and indicators: Abundance (A, B), Richness (C, D), Shannon's diversity index (E, F), and fish biomass (G)"}
+Plots are ordered by survey type (left: fish; right: invertebrates) and indicators: Abundance (A, B), Richness (C, D), Shannon's diversity index (E, F),fish biomass (G), and lobster (*Panulirus spp*) abundances (H).
+Points are jittered hotizontally to avoid overplotting.
+Points indicate the effect size, and errorbars are heteroskedastic-robust standard errors."}
 
@@
 
-cowplot::plot_grid(plotlist = list(F_N, I_N, F_S, I_S, F_H, I_S, F_B),
+## Lobsters
+L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
+                     various = T,
+                     model_names = c("IS", "ME", "PH"))
+
+cowplot::plot_grid(plotlist = list(F_N, I_N, F_S, I_S, F_H, I_H, F_B, L_plot),
                    labels = "AUTO",
@@
 
-```{r, fig.width = 4, fig.height = 6, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster abundances in marine reserves from Isla Natividad (IN; red romboids), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares)."}
+## Socioeconomic
 
-L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
-                     various = T,
-                     model_names = c("IS", "ME", "PH")) +
-  theme(legend.position = c(0, 1))
+```{r, fig.width = 3, fig.height = 4, fig.cap = "\\label{fig:lobsters}Time series of lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares)."}
 
@@
   labs(x = "Year", y = "Catches (Tones)") +
-  theme(legend.position = "none")
+  theme(legend.position = "none",
+        text = element_text(size = 12))
 
-L_value <- ggplot(L_conapesca, aes(x = Ano, y = Valor/1000, color = Comunidad, group = Comunidad, fill = Comunidad)) +
+L_value <- ggplot(L_conapesca, aes(x = Ano, y = Valor/1e6, color = Comunidad, group = Comunidad, fill = Comunidad)) +
   geom_smooth(method = "lm") +
@@
   geom_vline(xintercept = c(2006, 2012, 2013), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +
-  labs(x = "Year", y = "Value (MXP)") +
-  theme(legend.position = "none")
+  labs(x = "Year", y = "Value (M MXP)") +
+  theme(legend.position = "none",
+        text = element_text(size = 12))
 
-cowplot::plot_grid(plotlist = list(L_plot, L_catches, L_value),
+cowplot::plot_grid(plotlist = list(L_catches, L_value),
                    labels = "AUTO",
@@
 
+## Governance
+
 # Discussion
+
+The biological effects of the reserve contrast the existing literature. Why no effect on bio? IN hypoxia. ME and PH, age? Perhaps is the use of causal inference methods?
+
+Isla natividad no funciona por hipoxya
+
+Maria Elena funciona "bien"
+
+Punta Herrero no funciona por poaching.
+
+Aun con las mejores caracteristicas sociales, noe s posible vencer al ambiente (IN).
+Para el Caribe, parece ser que buena vigilancia y poca pesca ilegal son determinantes para obtener buenos resultados.
+Preguntarle a Stuart si ME solamente se pesca durante lanosta, y todo el ano es reserva.
+
+IN captures more dynamics. Reflects hypoxia events. Citar paper Giron
+
+Differences between fish and inverts sheds a lights on direct and indirect effects (Paper que recomendo Fio).
 
@@
 
-JC analyzed and interpreted data, discussed the results and wrote the manuscrip. SF and JT edited the manuscript and discussed the results.
+JC and EA analyzed and interpreted data, discussed the results and wrote the manuscrip. SF and JT edited the manuscript and discussed the results.
 
@@
 \label{S1_Figure}
-Maps of the marine reserves and corresponding control sites at each community.
+Timeseries of indicators for IN
+
+\paragraph*{S2 Figure}
+\label{S2_Figure}
+Timeseries of indicators for ME
+
+\paragraph*{S3 Figure}
+\label{S3_Figure}
+Timeseries of indicators for PH
+
+\paragraph*{S1 Table}
+\label{S1_Table}
+Coefficient estimates for Isla Natividad
 
@@
 \label{S2_Table}
-Table with a general overview of on the governance characteristics of each community.
+Coefficient estimates for Maria Elena
+
+\paragraph*{S3 Table}
+\label{S3_Table}
+Coefficient estimates for Punta Herrero
+
+```{r, results = "asis", eval = F}
+stargazer::stargazer(list(N_IN, Ni_IN, S_IN, Si_IN, H_IN, Hi_IN, B_IN, L_IN),
+                     type = "latex",
+                     single.row = T,
+                     omit = "GeneroEspecie",
+                     model.names = F,
+                     column.labels = c("Abundance (f)", "Abundance(i)", "Richness (f)", "Richness(i)", "Shannon (f)", "Shannon (i)", "Biomass", "Lobsters"),
+                     dep.var.labels.include = FALSE,
+                     model.numbers = FALSE,
+                     digits = 2,
+                     intercept.bottom = F, title = "\\paragraph*{S1 Table}\\label{S1_Table}")
+```
 

### 2018-02-13 Frontiers_Manuscript/Villasenor-Derbez.tex
Updated manuscript with results section

@@
 data for reserve and control sites before and after the implementation
-of the reserve (n = 4) , thus providing us with a
-Before--After--Control--Impact (\emph{i.e.} BACI) design that allows us
-to capture and control for temporal and spatial dynamics
-\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
-techniques have proven effective to evaluate marine reserves, as they
-allow us to causaly attribute observed changes to the intervention
-\citep{moland_2013-VP,Villasenor-Derbez_2018}.
+of the reserve. This provides us with a Before--After--Control--Impact
+(\emph{i.e.} BACI) design that allows us to capture and control for
+temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI
+designs and causal inference techniques have proven effective to
+evaluate marine reserves, as they allow us to causaly attribute observed
+changes to the intervention
+\citep{moland_2013-VP,Villasenor-Derbez_2018}. All reserves were
+surveyed annualy from at least one year before implementation until
+2016. Table \ref{tab:com_sum} shows a summary of the number of reserves,
+year of implementation, and number of transects for each reserve.
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-2}\label{tab:com_sum} Summary of commuity--based marine reserves by community.
+Imp = Year of implementation, Start = Year of first sampling, number of fish transects in control (Cf) and reserve (Rf) sites, and number of invertebrate transects in Control (Ci) and Reserve (Ri) sites.}
+\centering
+\begin{tabular}[t]{l|l|r|r|r|r|r|r}
+\hline
+Community & Reserve - Control & Imp & Start & Cf & Rf & Ci & Ri\\
+\hline
+Isla Natividad & La Plana / Las Cuevas - La Dulce / Babencho & 2006 & 2006 & 405 & 242 & 415 & 245\\
+\hline
+Maria Elena & Cabezo - Cabezo (Control) & 2012 & 2012 & 44 & 45 & 27 & 21\\
+\hline
+Punta Herrero & El Faro - El Faro (Control) & 2013 & 2013 & 39 & 40 & 24 & 32\\
+\hline
+Punta Herrero & Manchon - Manchon (Control) & 2013 & 2012 & 43 & 45 & 27 & 42\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
-\caption{\label{tab:unnamed-chunk-2}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
+\caption{\label{tab:unnamed-chunk-3}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
 \centering
@@
 that the reserve has on the biological indicators by comparing trends
-across time and space (\emph{i.e.} reserve and control sites)
+across time and treatments
 \citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis is performed
@@
 
-Where year--fixed effects are represented by \(\gamma_{it} Year_t\),
+Where year--fixed effects are represented by \(\gamma_{it} Year_t\), and
 \(\beta Zone_i\) captures the difference between reserve (\(Zone = 1\))
-and control (\(Zone = 0\)) sites, and the interaction term
-\(\lambda_{it} Year_t\times Zone_i\) represents the average treatment
-effect for each year. When evaluating biomass and abundances, we include
-species--fixed effects (\(\sigma_j\)), and when more than one reserve is
-present we also include site--fixed effects (\(\Pi_k\)). For abundances
+and control (\(Zone = 0\)) sites. The interaction term
+\(\lambda_{it} Year_t\times Zone_i\) represents represent the mean
+change in the indicator inside the reserve, for year \(t\), with respect
+to the first year of evaluation in the control site (See Table
+\ref{tab:com_sum}). When evaluating biomass and abundances, we include
+species--fixed effects (\(\sigma_j\)). For abundances and richness
 (\emph{i.e.} count data) the model is estimated with a quasipoisson
@@
 
+\clearpage
+
 \section{Results}\label{results}
 
-A continuación se presentan los resultados de cada una de las
-comunidades. Los resultados biológicos se presentarán para cada
-comunidad, discutiendo primero los indicadores en común con otras
-comunidades (Shannon, Riqueza, Densidad, Nivel Trófico, Biomasa para
-peces e invertebrados) y, según su caso, las densidades de las especies
-objetivo. Habiendo presentado los resultados biológicos, nos enfocaremos
-después en los socioeconómicos y de gobernanza. Usaremos esta
-información para identificar las causas (sociales) del éxito (biológico)
-de las reservas.
+Our methodological approach with biological indicators allows us to make
+a causal link between the implementation of marine reserves and the
+observed trends by accounting for temporal and spatial dynamics
+\citep{depalma_2018}. The effect of the reserve is captured by the
+\(\lambda_t\) coefficient, and represents the difference observed
+between the control site before the implementation of the reserve and
+the reserve site at time \(t\) after controling for other time and space
+variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Here
+we present the effect that marine reserves had on each of the biological
+indicators for each coastal community, along with the trends in
+socioeconomic indicators of lobster catches and revenues. We also
+provide an overview of the state of the socioeconomic and governance
+settings of each community, and discuss how these dimensions might be
+intertwined with each other.
+
+\subsection{Biological}\label{biological}
+
+Effect sizes for biological indicators are shown in Figure
+\ref{fig:indicators}. Isla Natividad, the community with the oldest
+reserve, shows inconsistent effects across indicators and sources of
+data (\emph{i.e.} fish vs.~invertebrates). For example, the reserve had
+a small effect on fish abundances (Fig \ref{fig:indicators}A), where
+only year 2010 showed significat effect sizes in fish abundances
+(\(p<0.05\)) and all other years oscillated above and under zero
+(\(p > 0.05\)). However, in the case of invertebrate abundances (Fig
+\ref{fig:indicators}B), there was a clear positive effect of the
+reserve, as all but 1 year (2008) presented positive effect sizes
+relative to the control site (\(p < 0.05\)). Maria Elena and Punta
+Herrero showed no significant increase in fish and invertebrate
+abundances (\(p< 0.05\)), except for invertebrates in Punta Herrero for
+2014 --right after the implementation of the reserves-- which showed a
+significant increase (\emph{i.e.} \(\lambda_{2014} = 2.5\),
+\(p < 0.05\)). Full tables with model coefficients are presented in the
+supplementary materials (\textbf{S1 Table}, \textbf{S2 Table},
+\textbf{S3 Table}).
+
+While the number of fish species oscillated above and bellow zero
+through time for all reserves, none of these changes were statistically
+significant (\(p > 0.05\)) indicating that the reserves had no effect on
+fish species richness (Fig \ref{fig:indicators}C). For invertebrate
+species in Isla Natividad, all effect sizes were negative, but only
+significant for 2008, 2009, 2011, and 2014 (\(p < 0.05\); Fig
+\ref{fig:indicators}D). For Maria Elena and Punta Herrero, the data do
+not show significant changes in invertebrate species richness
+(\(p > 0.05\)).
+
+Effect sizes for Shannon's diversity index for fish (Fig
+\ref{fig:indicators}E) in Isla Natividad oscillated between
+\(\lambda_{2011} = -0.45\) and \(\lambda_{2010} = -0.005\), but were not
+significantly different from null hypotheses of no change (\emph{i.e.}
+\(\lambda_t = 0\); \(p > 0.05\)). For invertebrates in that same
+community (Fig \ref{fig:indicators}F), Shannon's diversity index showed
+a significant decrease between 2008 and 2014, with largest decrease
+observed for 2011 (\(\lambda_{2011} = -0.91\); \(p < 0.05\)). In the
+case of Maria Elena and Punta Herrero, Shannon's diversity index for
+fish showed increases in the order of \(\lambda_t = 1\). For Maria Elena
+and Punta Herrero, these effects were only statistically significant for
+2014, and 2014 and 2015 (\(p < 0.05\)).
+
+Biomass was only evaluated for fish data (Fig \ref{fig:indicators}G). In
+Isla Natividad, fish biomass presented a steady but small increase
+(\(p>0.05\)), and exhibited an increased variability in biomass between
+2013 and 2016. Maria Elena and Punta Herrero also showed small,
+non-statistically significant increases in fish biomass (\(p>0.05\)).
+The last biological indicator is abundance of target species,
+\emph{Panulirus interruptus} and \emph{P. argus}, for the Pacific and
+Caribbean, respectively (Fig \ref{fig:indicators}H). Isla Natividad
+presented small constantly-positive effects but were not significantly
+different from the reference point of control site before the
+implementation of the reserve (\(p > 0.05\). Maria Elena showed
+significant increases in lobster densities in the order of
+\(\lambda_t = 10\) (\(p < 0.05\)). Finally, punta herrero presented
+alternating negative and positive effects, but these were not different
+from the baseline case (\(p > 0.05\)).
 
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{\label{fig:unnamed-chunk-3}\label{fig:indicators}Effect sizes for
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{\label{fig:unnamed-chunk-4}\label{fig:indicators}Effect sizes for
 marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
@@
 fish; right: invertebrates) and indicators: Abundance (A, B), Richness
-(C, D), Shannon's diversity index (E, F), and fish biomass (G)}
+(C, D), Shannon's diversity index (E, F),fish biomass (G), and lobster
+(\emph{Panulirus spp}) abundances (H). Points are jittered hotizontally
+to avoid overplotting. Points indicate the effect size, and errorbars
+are heteroskedastic-robust standard errors.}
 \end{figure}
+
+\subsection{Socioeconomic}\label{socioeconomic}
 
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{\label{fig:unnamed-chunk-4}\label{fig:lobsters}Effect sizes for
-lobster abundances in marine reserves from Isla Natividad (IN; red
-romboids), Maria Elena (ME; blue triangles), and Punta Herrero (PH;
-green squares).}
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{\label{fig:unnamed-chunk-5}\label{fig:lobsters}Time series of
+lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
+circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green
+squares).}
 \end{figure}
 
+\subsection{Governance}\label{governance-1}
+
 \section{Discussion}\label{discussion}
+
+The biological effects of the reserve contrast the existing literature.
+Why no effect on bio? IN hypoxia. ME and PH, age? Perhaps is the use of
+causal inference methods?
+
+Isla natividad no funciona por hipoxya
+
+Maria Elena funciona ``bien''
+
+Punta Herrero no funciona por poaching.
+
+Aun con las mejores caracteristicas sociales, noe s posible vencer al
+ambiente (IN). Para el Caribe, parece ser que buena vigilancia y poca
+pesca ilegal son determinantes para obtener buenos resultados.
+Preguntarle a Stuart si ME solamente se pesca durante lanosta, y todo el
+ano es reserva.
+
+IN captures more dynamics. Reflects hypoxia events. Citar paper Giron
+
+Differences between fish and inverts sheds a lights on direct and
+indirect effects (Paper que recomendo Fio).
 
@@
 
-JC analyzed and interpreted data, discussed the results and wrote the
-manuscrip. SF and JT edited the manuscript and discussed the results.
+JC and EA analyzed and interpreted data, discussed the results and wrote
+the manuscrip. SF and JT edited the manuscript and discussed the
+results.
 
@@
 
-Maps of the marine reserves and corresponding control sites at each
-community.
+Timeseries of indicators for IN
+
+\paragraph*{S2 Figure}
+\label{S2_Figure}
+
+Timeseries of indicators for ME
+
+\paragraph*{S3 Figure}
+\label{S3_Figure}
+
+Timeseries of indicators for PH
+
+\paragraph*{S1 Table}
+\label{S1_Table}
+
+Coefficient estimates for Isla Natividad
 
@@
 
-Table with a general overview of on the governance characteristics of
-each community.
+Coefficient estimates for Maria Elena
+
+\paragraph*{S3 Table}
+\label{S3_Table}
+
+Coefficient estimates for Punta Herrero
 

### 2018-02-13 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Added new figure summarizing biological results

@@
 
-Effect sizes for biological indicators are shown in Figure \ref{fig:indicators}.
-Isla Natividad, the community with the oldest reserve, shows inconsistent effects across indicators and sources of data (*i.e.* fish vs.
+Effect sizes for biological indicators are shown in Figure \ref{fig:indicators}, and Figure \ref{fig:res_com} shows the summarized biological effects by community.
+Isla Natividad shows inconsistent effects across data sources (*i.e.* fish vs.
 invertebrates).
 For example, the reserve had a small effect on fish abundances (Fig \ref{fig:indicators}A), where only year 2010 showed significat effect sizes in fish abundances ($p<0.05$) and all other years oscillated above and under zero ($p > 0.05$).
-However, in the case of invertebrate abundances (Fig \ref{fig:indicators}B), there was a clear positive effect of the reserve, as all but 1 year (2008) presented positive effect sizes relative to the control site ($p < 0.05$).
+However, invertebrate abundances (Fig \ref{fig:indicators}B) presented a positive trend relative to the control site before implementation ($p < 0.05$) for all but 1 year (2008).
 Maria Elena and Punta Herrero showed no significant increase in fish and invertebrate abundances ($p< 0.05$), except for invertebrates in Punta Herrero for 2014 --right after the implementation of the reserves-- which showed a significant increase (*i.e.* $\lambda_{2014} = 2.5$, $p < 0.05$).
@@
 
+```{r, fig.height = 3, fig.width = 6, fig.cap = "\\label{fig:res_com}Summarized effects of the marne reserves by direction (positive - negative) and significance."}
+observe <- function(model){      tidy_terms <- model %>% 
+        robust_se() %>%
+        filter(grepl(":ZonaReserva", term)) %$%
+        str_split(term, pattern = ":", simplify = T) %>% 
+        as.tibble() %>% 
+        set_colnames(c("Ano", "Zona")) %>% 
+        select(-Zona) %>% 
+        mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = "")))
+
+      p_i <- robust_se(model) %>%
+        filter(grepl(":ZonaReserva", term)) %>%
+        cbind(tidy_terms) %>%
+        mutate(ptest = ((p.value < 0.05)*-1)+0.05)
+       return(p_i)}
+
+IN <- list(N_IN, Ni_IN, S_IN, Si_IN, H_IN, Hi_IN, B_IN, L_IN) %>% map_df(.f = observe) %>%
+  as.tibble() %>% 
+  select(estimate, p.value) %>% 
+  mutate(Com = "Isla Natividad",
+         n_ind = nrow(.))
+
+ME <- list(N_ME, Ni_ME, S_ME, Si_ME, H_ME, Hi_ME, B_ME, L_ME) %>% map_df(.f = observe) %>%
+  as.tibble() %>% 
+  select(estimate, p.value) %>% 
+  mutate(Com = "Maria Elena",
+         n_ind = nrow(.))
+
+PH <- list(N_PH, Ni_PH, S_PH, Si_PH, H_PH, Hi_PH, B_PH, L_PH) %>% map_df(.f = observe) %>%
+  as.tibble() %>% 
+  select(estimate, p.value) %>% 
+  mutate(Com = "Punta Herrero",
+         n_ind = nrow(.))
+
+rbind(IN, ME, PH) %>% 
+  mutate(direction = case_when(estimate < 0 ~ "Negative",
+                               estimate > 0 ~ "Positive"),
+         Significance = case_when(p.value < 0.05 ~ "p < 0.05",
+                            p.value >= 0.05 ~ "p > 0.05")) %>%
+  group_by(Com, direction, Significance, n_ind) %>% 
+  summarize(A = n()) %>% 
+  ungroup() %>% 
+  mutate(A = A/n_ind,
+         A = ifelse(direction == "Negative", -A, A),
+         Significance = fct_relevel(Significance, "p > 0.05")) %>% 
+  ggplot(aes(x = Com, y = A, fill = Significance)) +
+  geom_col(color = "black", alpha = 0.8, width = 0.8) +
+  geom_hline(yintercept = 0, linetype = "dashed", size = 1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  coord_flip() +
+  labs(x = "Community", y = "Percentage of total coefficients") +
+  scale_y_continuous(breaks = c(-0.5, -0.25, 0, 0.25, 0.5, 0.75), limits = c(-0.5, 0.75),
+                     labels = scales::percent) +
+  theme(legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8))
+```
+
 ## Socioeconomic
 
-```{r, fig.width = 3, fig.height = 4, fig.cap = "\\label{fig:lobsters}Time series of lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares)."}
+```{r, fig.width = 3.5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Time series of lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares)."}
 
@@
                    ncol = 1)
-
 ```

### 2018-02-13 Frontiers_Manuscript/Villasenor-Derbez.tex
Added new figure summarizing biological results

@@
 Effect sizes for biological indicators are shown in Figure
-\ref{fig:indicators}. Isla Natividad, the community with the oldest
-reserve, shows inconsistent effects across indicators and sources of
-data (\emph{i.e.} fish vs.~invertebrates). For example, the reserve had
-a small effect on fish abundances (Fig \ref{fig:indicators}A), where
-only year 2010 showed significat effect sizes in fish abundances
-(\(p<0.05\)) and all other years oscillated above and under zero
-(\(p > 0.05\)). However, in the case of invertebrate abundances (Fig
-\ref{fig:indicators}B), there was a clear positive effect of the
-reserve, as all but 1 year (2008) presented positive effect sizes
-relative to the control site (\(p < 0.05\)). Maria Elena and Punta
-Herrero showed no significant increase in fish and invertebrate
-abundances (\(p< 0.05\)), except for invertebrates in Punta Herrero for
-2014 --right after the implementation of the reserves-- which showed a
-significant increase (\emph{i.e.} \(\lambda_{2014} = 2.5\),
-\(p < 0.05\)). Full tables with model coefficients are presented in the
-supplementary materials (\textbf{S1 Table}, \textbf{S2 Table},
-\textbf{S3 Table}).
+\ref{fig:indicators}, and Figure \ref{fig:res_com} shows the summarized
+biological effects by community. Isla Natividad shows inconsistent
+effects across data sources (\emph{i.e.} fish vs.~invertebrates). For
+example, the reserve had a small effect on fish abundances (Fig
+\ref{fig:indicators}A), where only year 2010 showed significat effect
+sizes in fish abundances (\(p<0.05\)) and all other years oscillated
+above and under zero (\(p > 0.05\)). However, invertebrate abundances
+(Fig \ref{fig:indicators}B) presented a positive trend relative to the
+control site before implementation (\(p < 0.05\)) for all but 1 year
+(2008). Maria Elena and Punta Herrero showed no significant increase in
+fish and invertebrate abundances (\(p< 0.05\)), except for invertebrates
+in Punta Herrero for 2014 --right after the implementation of the
+reserves-- which showed a significant increase (\emph{i.e.}
+\(\lambda_{2014} = 2.5\), \(p < 0.05\)). Full tables with model
+coefficients are presented in the supplementary materials (\textbf{S1
+Table}, \textbf{S2 Table}, \textbf{S3 Table}).
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{\label{fig:unnamed-chunk-5}\label{fig:res_com}Summarized effects of
+the marne reserves by direction (positive - negative) and significance.}
+\end{figure}
+
 \subsection{Socioeconomic}\label{socioeconomic}
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{\label{fig:unnamed-chunk-5}\label{fig:lobsters}Time series of
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{\label{fig:unnamed-chunk-6}\label{fig:lobsters}Time series of
 lobster catches (A) and revenues (B) in at Isla Natividad (IN; red

### 2018-02-14 Frontiers_Manuscript/Villasenor-Derbez.Rmd
ÑAdded first final version of manuscript draft

@@
 ---
-title: "Cool title here"
-rtitle: "Mexican marine reserves"
-author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Stuart Fulton$^{2}$"
+title: "Enabling conditions for effective community based marine reserves in small scale fisheries"
+rtitle: "Marine reserve effectiveness"
+author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Álvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
 institution: "
@@
 email: "jvillasenor@bren.ucsb.edu"
-abstract: "Abstract here"
+abstract: "Abstract after comments from co-authors"
 keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico, Social--Ecological Systems"
@@
 Marine reserve science has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, climate change mitigation \citep{roberts_2017-J9}, and protection from environmental variability \citep{micheli_2012-EU}.
-However, there is considerably less literature focusing on the relationship between socioeconomic and governance structures and their relationship to ecological effectriveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or benefits to fisheries \citep{krueck_2017-J1}; evaluations of marine reserves rarely provide a hollistic view of the social-ecological system \citep{lpezangarita_2014}.
+However, there is considerably less literature focusing on the relationship between socioeconomic and governance structures and their relationship to ecological effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or benefits to fisheries \citep{krueck_2017-J1}; evaluations of marine reserves rarely provide a holistic view of the social-ecological system \citep{lpezangarita_2014}.
 Here, we combine causal inference techniques \citep{depalma_2018} and the social-ecological systems framework \citep{ostrom_2009-hg} to provide a comprehensive ecological and socioeconomic evaluation of four community-based marine reserves in three coastal communities in Mexico.
 
-Marine Reserves in Mexico have been commonly implemented as "core zones" within Biosphere Reserves (BRs) that are administered by the National Commision of Protected Areas (*Comisión Nacional de Áreas Marinas Protegidas*, CONANP).
+Marine Reserves in Mexico have been commonly implemented as "core zones" within Biosphere Reserves that are administered by the National Commission of Protected Areas (*Comisión Nacional de Áreas Marinas Protegidas*, CONANP).
 While CONANP has made efforts to have a participatory process, the implementation of these zones is still characterized by top--down approaches.
 This motivated Civil Society Organizations (CSOs) to work with coastal communities to implement community--based marine reserves \citep{uribe_2010-u2}, which are usually established within a Territorial Use Rights for Fisheries (TURFs); thus making them TURF--reserves \citep{afflerbach_2014-HP}.
-This bottom--up approach allows fishers to disign their own reserves, which  increases compliance and self--enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
+This bottom--up approach allows fishers to design their own reserves, which  increases compliance and self--enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
 However, these reserves still lack legal recognition, making them vulnerable to poaching.
@@
 This new norm thus combines bottom--up approaches to design marine reserves, along with a legal recognition of the management intervention.
-Since then, *39* FR have been implemented along the Pacific, Gulfof Caloifornia, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.
+Since then, *39* FR have been implemented along the Pacific, Gulf of California, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.
 
-While there are ecological factors defining the success of a MR (*i.e.* habitat representation, initial state of protection, connectivity to other protected areas), their effectivness also depends on the socioeconomic and governance settings under which they are implemented.
-Literature shows that many non-ecological characteristics can play an equally important role in the effectivenes of MRs.
+While there are ecological factors defining the success of a MR (*i.e.* habitat representation, initial state of protection, connectivity to other protected areas), their effectiveness also depends on the socioeconomic and governance settings under which they are implemented.
+Literature shows that many non-ecological characteristics can play an equally important role in the effectiveness of MRs.
 For example, age of a reserve (*i.e.* time since its implementation), size, and habitat contained were key to the effectiveness of MRs in Palau \citep{friedlander_2017-oI}.
-In the Mediterranean, \citet{difranco_2016-Xw} identify that surveilance and enfocrement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key fators that increased stock health and income to fishers.
+In the Mediterranean, \citet{difranco_2016-Xw} identify that surveillance and enforcement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key factors that increased stock health and income to fishers.
 At a global level, \citet{edgar_2014-UO} indicate that enforcement, age, size, and isolation were important factors determining effectiveness of the reserves.
 
-The objective of this work is twofold: i) Provide the first evaluation of community--based marine reserves in Mexico, and ii) provide a comprehensive evaluation of the social--ecological system to identify how sociocoenomic and governance characteristics relate to ecological effectiveness.
-With the purpose of providing a hollistic evaluation, we combine ecological, socioeconomic, and governance indicators.
+The objective of this work is twofold: i) Provide the first evaluation of community--based marine reserves in Mexico, and ii) provide a comprehensive evaluation of the social--ecological system to identify how socioeconomic and governance characteristics relate to ecological effectiveness.
+With the purpose of providing a holistic evaluation, we combine ecological, socioeconomic, and governance indicators.
 We use causal inference techniques to provide a measurement of the effect of the management intervention, and combine it with the social--ecological systems framework \citep{ostrom_2009-hg}.
@@
 # Materials and Methods
-
-<!--   - Data collection -->
-
-<!--     - Sampling design (BACI) -->
-
-<!--     - Sampling protocol (transects, richness, abundance, sizes) -->
-
-<!--     - Sample sizes -->
-
-<!--   - Data analysis -->
-
-<!--     - Objectives and indicators -->
-
-<!--     - Analysis of numeric indicators -->
-
-<!--     - Analysis of descriptive indicators -->
 
@@
 The island is home to a fishing cooperative (*Sociedad Cooperative de Producción Pesquera Buzos y Pescadores de la Baja California SCL*), that holds a TURF for spiny lobster (*Panulirus interruptus*).
-However, other resources like finfish (yellowtail jack, *Seriola lalandi*), sea cucumber (*Parastichopus parvimensis*), red sea urchin (*Mesocentrotus franciscanus*), snail (*Megastraea turbanica* y *M.
+However, other resources like finfish (yellow-tail jack, *Seriola lalandi*), sea cucumber (*Parastichopus parvimensis*), red sea urchin (*Mesocentrotus franciscanus*), snail (*Megastraea turbanica* y *M.
 undosa*), and abalone (*Haliotis spp*, until 2010) are also important sources of income.
@@
 The other two communities are Maria Elena (ME; Fig \ref{fig:map}C) and Punta Herrero (PH; Fig \ref{fig:map}D) in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems.
-ME is a fishing camp --visited intermitently during the fishing season-- belonging to the Cozumel fishing cooperative.
+ME is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative.
 PH is home to the "José María Azcorra" fishing cooperative.
-The main source of income to both communities is the caribbean spinly lobster fishery (*Panulirus argus*), which is carried out within their respective TURFs.
+The main source of income to both communities is the Caribbean spiny lobster fishery (*Panulirus argus*), which is carried out within their respective TURFs.
 These communities also target finfish in the off season, mainly snappers (Lutjanidae) and groupers (Serranidae).
@@
 
-To perform the evaluation of these reserves we use thre sources of information.
-Ecological data come from the annual ecological monitorings of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO "Comunidad y Biodiversidad" ([COBI](www.cobi.org.mx)).
-These monitorings record richness and abundances of fish and invertebrate species in the reservs and control sites.
+To perform the evaluation of these reserves we use thee sources of information.
+Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO "Comunidad y Biodiversidad" ([COBI](www.cobi.org.mx)).
+These monitorings record richness and abundances of fish and invertebrate species in the reserves and control sites.
 For fish census, size structures are also collected to derive biomass.
@@
 From all the reserves in these three communities, we use the ones that have data for reserve and control sites before and after the implementation of the reserve.
-This provides us with a Before--After--Control--Impact (*i.e.* BACI) design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
-BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causaly attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
-All reserves were surveyed annualy from at least one year before implementation until 2016.
-Table \ref{tab:com_sum} shows a summary of the number of reserves, year of implementation, and number of transects for each reserve.
+This provides us with a Before-After-Control-Impact (*i.e.* BACI) design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
+BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
+All reserves were surveyed annually from at least one year before implementation until 2016.
+Table \ref{table:com_sum} shows a summary of the number of reserves, year of implementation, and number of transects for each reserve.
 
@@
   knitr::kable(format = "latex",
-               caption = "\\label{tab:com_sum} Summary of commuity--based marine reserves by community.
+               caption = "\\label{table:com_sum} Summary of commuity--based marine reserves by community.
 Imp = Year of implementation, Start = Year of first sampling, number of fish transects in control (Cf) and reserve (Rf) sites, and number of invertebrate transects in Control (Ci) and Reserve (Ri) sites.",
-               col.names = c("Community", "Reserve - Control", "Imp", "Start", "Cf", "Rf", "Ci", "Ri")) #%>% 
-  # kableExtra::add_header_above(header = c(" " = 4, "Fish" = 6, "Invert" = 8))
+               col.names = c("Community", "Reserve - Control", "Imp", "Start", "Cf", "Rf", "Ci", "Ri"))
 ```
@@
 
-**Governance data were collected at the community--level, using key informants to collect the necesary information on X, Y, and Z. Algo mas de Ere.**
+Governance data were collected at the community-level.
+The information was compiled by combining key informants and the authors; experience and knowledge of the communities to collect the necessary information.
+These data contain information on the ecological system where the fishing activities develop, as well as the governance structures present in the cooperative.
+We also gathered information on the resource unit (*i.e.* lobsters) and the relevant actors present in each community \citep{leslie_2015-na}.
 
@@
 
-Following a framework that relates reserve objectives to performance indicators \citep{Villasenor-Derbez_2018}, we use five biological, two socioeconomic, and five governance indicators to evaluate these marine reserves Table \ref{table:indicators}.
+Following a framework that relates reserve objectives to performance indicators \citep{Villasenor-Derbez_2018}, we use five biological and two socioeconomic indicators to evaluate these marine reserves Table \ref{table:indicators}.
+We also use a set of governance indicators to analyze the governance structures of each cooperative \citep{leslie_2015-na}.
+The indicators (Table \ref{table:gov_ind}) focus on the resource system (four indicators), governance system (seven indicators), resource units (three indicators) and actors (three indicators).
 
 ```{r}
-tibble(Category = c("Biological", "","","","", "Socioeconomic", "", "Governance", "", "", "", ""),
-           Indicador = c("Abundance", "Richness", "Shannon's diversity index", "Biomass", "Abundance of target species (lobsters)", "Income from target species", "Landings from target species", "Type of access to the fishery", "Perceived degree of illegal fishing", "Reserve surveilance anf enforcement", "Type of fishing organization", "Age of the reserve")) %>%
+tibble(Category = c("Biological", "","","","", "Socioeconomic", ""),
+           Indicador = c("Abundance", "Richness", "Shannon's diversity index", "Biomass", "Abundance of target species (lobsters)", "Income from target species", "Landings from target species")) %>% 
   knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex") %>% 
   kable_styling(latex_options = "HOLD_position")
+```
+
+```{r}
+read.csv("../Data/gov_ind.csv") %>% 
+  knitr::kable(caption = "\\label{table:gov_ind}Indicators used for the operationalization of the SES framework \\citep{leslie_2015-na}", col.names = c("Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero"), format = "latex") %>%
+  kableExtra::column_spec(1, width = "10em") %>% 
+  kableExtra::column_spec(2, width = "8em") %>% 
+  kableExtra::column_spec(3, width = "8em") %>% 
+  kableExtra::column_spec(4, width = "8em") %>% 
+  kableExtra::row_spec(1 , bold = T) %>% 
+  kableExtra::row_spec(6 , bold = T) %>% 
+  kableExtra::row_spec(14 , bold = T) %>% 
+  kableExtra::row_spec(18 , bold = T)
 ```
@@
 Where year--fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
-The interaction term $\lambda_{it} Year_t\times Zone_i$ represents represent the mean change in the indicator inside the reserve, for year $t$, with respect to the first year of evaluation in the control site (See Table \ref{tab:com_sum}).
+The interaction term $\lambda_{it} Year_t\times Zone_i$ represents represent the mean change in the indicator inside the reserve, for year $t$, with respect to the first year of evaluation in the control site (See Table \ref{table:com_sum}).
 When evaluating biomass and abundances, we include species--fixed effects ($\sigma_j$).
@@
 For both approaches, model coefficients are estimated via ordinary least--squares and heteroskedastic--robust standard errors \citep{zeileis_2004-7n}.
-
-### Governance
-
-**Texto de ere y los SES**
 
@@
 Our methodological approach with biological indicators allows us to make a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}.
-The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the reserve site at time $t$ after controling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
+The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the reserve site at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
 Here we present the effect that marine reserves had on each of the biological indicators for each coastal community, along with the trends in socioeconomic indicators of lobster catches and revenues.
@@
 invertebrates).
-For example, the reserve had a small effect on fish abundances (Fig \ref{fig:indicators}A), where only year 2010 showed significat effect sizes in fish abundances ($p<0.05$) and all other years oscillated above and under zero ($p > 0.05$).
+For example, the reserve had a small effect on fish abundances (Fig \ref{fig:indicators}A), where only year 2010 showed significant effect sizes in fish abundances ($p<0.05$) and all other years oscillated above and under zero ($p > 0.05$).
 However, invertebrate abundances (Fig \ref{fig:indicators}B) presented a positive trend relative to the control site before implementation ($p < 0.05$) for all but 1 year (2008).
@@
 Maria Elena showed significant increases in lobster densities in the order of $\lambda_t = 10$ ($p < 0.05$).
-Finally, punta herrero presented alternating negative and positive effects, but these were not different from the baseline case ($p > 0.05$).
+Finally, Punta Herrero presented alternating negative and positive effects, but these were not different from the baseline case ($p > 0.05$).
 
@@
 
-```{r, fig.width = 3.5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Time series of lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares)."}
+Lobster catches and income showed a increase after the implementation of the reserves for Isla Natividad and Maria Elena, but not for Punta Herrero (Fig \ref{fig:lobsters}).
+However, the differences in catches and and revenue were not different in the periods before and after the implementation ($p > 0.05$) except for revenues in Isla Natividad, which presented a significant increase of 14.37 (M MXP; $p<0.05$).
+Table \ref{table:soc_res} presents all the regression coefficients.
 
+```{r, results = "asis"}
 L_conapesca <- tibble(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl",
@@
                       Comunidad = c("Isla Natividad",
-                                    "María Elena",
+                                    "Maria Elena",
                                     "Punta Herrero")) %>% 
   left_join(conapesca, ., by = "UnidadEconomica") %>% 
-  mutate(Post = case_when(Comunidad == "Isla Natividad" && Ano <= 2006 ~ "Pre",
-                          Comunidad == "Maria Elena" && Ano <= 2012 ~ "Pre",
-                          Comunidad == "Punta Herrer" && Ano <= 2012 ~ "Pre",
-                          TRUE ~ "Post"))
+  mutate(Post = case_when(Comunidad == "Isla Natividad" && Ano <= 2006 ~ 0,
+                          Comunidad == "Maria Elena" && Ano <= 2012 ~ 0,
+                          Comunidad == "Punta Herrero" && Ano <= 2013 ~ 0,
+                          TRUE ~ 1))
+
+C_IN <- filter(L_conapesca, Comunidad == "Isla Natividad") %>% 
+  lm(PesoVivo/1000~Post, data = .)
+
+C_ME <- filter(L_conapesca, Comunidad == "Maria Elena") %>% 
+  lm(PesoVivo/1000~Post, data = .)
+
+C_PH <- filter(L_conapesca, Comunidad == "Punta Herrero") %>% 
+  lm(PesoVivo/1000~Post, data = .)
+
+V_IN <- filter(L_conapesca, Comunidad == "Isla Natividad") %>% 
+  lm(Valor/1000000~Post, data = .)
+
+V_ME <- filter(L_conapesca, Comunidad == "Maria Elena") %>% 
+  lm(Valor/1000000~Post, data = .)
+
+V_PH <- filter(L_conapesca, Comunidad == "Punta Herrero") %>% 
+  lm(Valor/1000000~Post, data = .)
+
+stargazer::stargazer(C_IN, C_ME, C_PH, V_IN, V_ME, V_PH, digits = 2, type = "latex", header = F, omit.stat = c("adj.rsq", "ser", "f"), title = "\\label{table:soc_res} Regression coefficients for lobster catches and revenues for Isla Natividad (1, 4), Maria Elena (2, 5), and Punta Herrero (3, 6).", single.row = T, report = c("vc*"), dep.var.labels = c("Catches (tones)", "Revenue(M MXP)"))
+
+```
+
+```{r, fig.width = 3.5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Time series of lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares)."}
 
@@
   geom_smooth(method = "lm") +
-  geom_point(size = 2, alpha = 0.5) +
+  geom_point(aes(shape = Comunidad), size = 2, alpha = 0.5) +
   scale_color_brewer(palette = "Set1") +
@@
 
+Although we have little information on the social dimension of these fisheries, using the SES framework indicators (Table \ref{table:gov_ind}), we can analyze the performance of each governance system with respect to MR enforcement (Table \ref{table:gov_res}).
+In general, the presence and success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources they depend upon \citep{jupiter_2017}.
+The enabling conditions for conservation seem to be strongly present in Isla Natividad.
+Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperative.
+These conditions have favored the development of an efficient community based enforcement system.
+In contrast, the communities of Maria Elena and Punta Herrero are located near other fishing communities and cities.
+In Maria Elena, the fishing pressure caused by outsiders can be reduced by implementing a strong enforcement system (in water and land) supported by CSOs and the local government (CONANP).
+Lastly, the community of Punta Herrero shows the highest levels of illegal activities which can be attributed to its connectedness to other communities and the lack of appropriate technologies for enforcement.
+
+```{r}
+read.csv("../Data/gov_res.csv") %>% 
+  knitr::kable(caption = "\\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \\citep{mcginnis_2014}.",
+               col.names = c("", "Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
+  kableExtra::column_spec(1, width = "10em") %>%
+  kableExtra::column_spec(2, width = "10em") %>%
+  kableExtra::row_spec(1 , bold = T) %>% 
+  kableExtra::row_spec(7 , bold = T) %>% 
+  kableExtra::row_spec(15 , bold = T) %>% 
+  kableExtra::row_spec(19 , bold = T)
+```
+
+\clearpage
+
 # Discussion
 
-The biological effects of the reserve contrast the existing literature. Why no effect on bio? IN hypoxia. ME and PH, age? Perhaps is the use of causal inference methods?
+Our results indicated idiosyncratic biological effects of the reserves across communities and indicators, with a combination of positive and negative effects.
+However, many of these effects were not statistically significant, indicating no effect of the reserve \ref{fig:res_com}.
+The socioeconomic indicators pertaining to landings and revenues associated to those landings showed little or no temporal change before and after reserve implementation.
+These contrasting effects, however, might be clarified when understanding the social-ecological context in which these communities and their reserves sit.
+In this section, we discuss potential shortcomings in our analysis, and provide plausible explanations to the observed biological and socioeconomic basing on previous literature and our analysis of the social-ecological system.
 
-Isla natividad no funciona por hipoxya
+The contrasting biological effectiveness observed is perhaps explained by our approach to evaluate the temporal and spatial changes of each indicator.
+Some works have solely focused on an inside-outside comparison of indicators \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD}, which do not address temporal variability \citep{depalma_2018}.
+Other works have compared the trend observed within a reserve through time \citep{betti_2017-lq}, which cannot distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}.
+By accounting for trends between sites and through times, we can control for time and space dynamics, and provide a better identification of the effect.
+However, it is worth looking deeper into each case, and identifying other plausible explanations.
 
-Maria Elena funciona "bien"
+Age, isolation, and enforcement are important factors influencing effectiveness of a marine reserve \citep{edgar_2014-UO}.
+Isla Natividad has the oldest reserve, is fairly isolated, and has a well-established community-based enforcement system.
+While other communities are certainly within reach, these are known to be well organized fishing communities with successful resource management \citep{mccay_2017-1m,mccay_2014-CN}.
+The reserve at Isla Natividad presented the largest percentage of significantly positive changes in biological indicators (19%), but an important portion of was also negative (15%).
+With the age, relative isolation, and enforcement level of this reserve, it would be expected for it to be considerably effective.
+The potential gap in performance can be attributed to perturbations that do not distinguish reserve boundaries, such as environmental variability (**no recuerdo esta cita**).
+The region is known to be under the influence of recurrent hypoxia and high-temperature events known to cause massive adult mortalities \citep{micheli_2012-EU}.
 
-Punta Herrero no funciona por poaching.
+Maria Elena and Punta Herrero are relatively young reserves (See Table \ref{table:com_sum}).
+From these, the Maria Elena exhibited the highest performance in terms of biological indicators (15% significantly positive).
+In contrast, Punta Herrero had a similar proportion of positive and negative effects.
 
-Aun con las mejores caracteristicas sociales, noe s posible vencer al ambiente (IN).
-Para el Caribe, parece ser que buena vigilancia y poca pesca ilegal son determinantes para obtener buenos resultados.
-Preguntarle a Stuart si ME solamente se pesca durante lanosta, y todo el ano es reserva.
+The way in which we measure changes in catches and revenues can not identify whether the observed differences are simply caused by pre-existing temporal trends or by the implementation of the reserve.
+Yet, there were no detectable changes in these indicators, except for landings in Isla Natividad.
+Other research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
+Since the trend was not detected in catches --directly related to abundance and fishing effort-- it is plausible that these differences are purely explained by an increase in market-level prices.
 
-IN captures more dynamics. Reflects hypoxia events. Citar paper Giron
+The fact that there was no detectable change in catches for Maria Elena and Punta Herrero can be explained by a combination of factors related to the design, management, age, or ecological factors.
+Reserves in these communities are relatively small and young, and may need more time for lobster abundances to increase enough to export larvae or adult organisms.
+Other community-based marine reserves in tropical ecosystems have taken up to six years to show a spillover effect \citep{dasilva_2015-zX}.
+A complimentary explanation lies in the results observed for the governance system.
+The lack of enforcement in Punta Herrero, for example, could explain the lack of effectiveness observed in their reserves.
 
-Differences between fish and inverts sheds a lights on direct and indirect effects (Paper que recomendo Fio).
+Our results show that community-based marine reserves can be effective if the environmental and social settings allow it.
+By studying the social-ecological system as a whole, we can provide a wider range of explanations to the patterns observed.
+It is interesting that even under the best enabling social conditions, climate variability can hinder the effect of a reserve --Although it is interesting to imagine what the state of that fishery had been if the reserve and organized cooperative were not present--.
+On the contrary, we show how under low climate variability, absence of proper governance structures can limit the effectiveness and benefits of a reserve.
+Whether the combination of a stable environment and governance structures are additive or multiplicative represents an interesting area for future research, especially under a changing climate.
 
@@
 
-JC and EA analyzed and interpreted data, discussed the results and wrote the manuscrip. SF and JT edited the manuscript and discussed the results.
+JC and EA analyzed and interpreted data, discussed the results, and wrote the manuscript. AS, SF and JT edited the manuscript and discussed the results.
 
 \section*{Funding}
-Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
+JCVD CONACyT + LAFF
+ASC
+SF
+JT
 
 \section*{Acknowledgments}
-This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.
+The authors wish to acknowledge Arturo Hernández and Imelda Amador for contributions on the governance data, as well as pre-processing biological data.
+This study would have not been possible without the effort by members of the communities here mentioned, who collected the biological data.
 

### 2018-02-14 Frontiers_Manuscript/Villasenor-Derbez.tex
ÑAdded first final version of manuscript draft

@@
 \def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
-Aceves-Bueno\(^{1,*}\), Stuart Fulton\(^{2}\)}
+Aceves-Bueno\(^{1,*}\), Álvin Suarez-Castillo\(^{2}\), Stuart
+Fulton\(^{2}\), Jorge Torre\(^{2}\)}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 
-\title[Mexican marine reserves]{Cool title here}
+\title[Marine reserve effectiveness]{Enabling conditions for effective community based marine reserves in
+small scale fisheries}
 
@@
 
-Abstract here
+Abstract after comments from co-authors
 
@@
 literature focusing on the relationship between socioeconomic and
-governance structures and their relationship to ecological
-effectriveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or
-benefits to fisheries \citep{krueck_2017-J1}; evaluations of marine
-reserves rarely provide a hollistic view of the social-ecological system
+governance structures and their relationship to ecological effectiveness
+\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or benefits to
+fisheries \citep{krueck_2017-J1}; evaluations of marine reserves rarely
+provide a holistic view of the social-ecological system
 \citep{lpezangarita_2014}. Here, we combine causal inference techniques
@@
 Marine Reserves in Mexico have been commonly implemented as ``core
-zones'' within Biosphere Reserves (BRs) that are administered by the
-National Commision of Protected Areas (\emph{Comisión Nacional de Áreas
-Marinas Protegidas}, CONANP). While CONANP has made efforts to have a
+zones'' within Biosphere Reserves that are administered by the National
+Commission of Protected Areas (\emph{Comisión Nacional de Áreas Marinas
+Protegidas}, CONANP). While CONANP has made efforts to have a
 participatory process, the implementation of these zones is still
@@
 (TURFs); thus making them TURF--reserves \citep{afflerbach_2014-HP}.
-This bottom--up approach allows fishers to disign their own reserves,
+This bottom--up approach allows fishers to design their own reserves,
 which increases compliance and self--enforcement
@@
 with a legal recognition of the management intervention. Since then,
-\emph{39} FR have been implemented along the Pacific, Gulfof
-Caloifornia, and Mexican Caribbean coastlines, but no formal evaluation
+\emph{39} FR have been implemented along the Pacific, Gulf of
+California, and Mexican Caribbean coastlines, but no formal evaluation
 of their effectiveness has taken place.
@@
 (\emph{i.e.} habitat representation, initial state of protection,
-connectivity to other protected areas), their effectivness also depends
+connectivity to other protected areas), their effectiveness also depends
 on the socioeconomic and governance settings under which they are
 implemented. Literature shows that many non-ecological characteristics
-can play an equally important role in the effectivenes of MRs. For
+can play an equally important role in the effectiveness of MRs. For
 example, age of a reserve (\emph{i.e.} time since its implementation),
@@
 Palau \citep{friedlander_2017-oI}. In the Mediterranean,
-\citet{difranco_2016-Xw} identify that surveilance and enfocrement,
+\citet{difranco_2016-Xw} identify that surveillance and enforcement,
 presence of a management plan, and involvement of fishers in management
 and decision--making along with promotion of sustainable fishing
-practices were the key fators that increased stock health and income to
+practices were the key factors that increased stock health and income to
 fishers. At a global level, \citet{edgar_2014-UO} indicate that
@@
 comprehensive evaluation of the social--ecological system to identify
-how sociocoenomic and governance characteristics relate to ecological
-effectiveness. With the purpose of providing a hollistic evaluation, we
+how socioeconomic and governance characteristics relate to ecological
+effectiveness. With the purpose of providing a holistic evaluation, we
 combine ecological, socioeconomic, and governance indicators. We use
@@
 California SCL}), that holds a TURF for spiny lobster (\emph{Panulirus
-interruptus}). However, other resources like finfish (yellowtail jack,
+interruptus}). However, other resources like finfish (yellow-tail jack,
 \emph{Seriola lalandi}), sea cucumber (\emph{Parastichopus
@@
 coral reefs and mangroves are the representative coastal ecosystems. ME
-is a fishing camp --visited intermitently during the fishing season--
+is a fishing camp --visited intermittently during the fishing season--
 belonging to the Cozumel fishing cooperative. PH is home to the ``José
 María Azcorra'' fishing cooperative. The main source of income to both
-communities is the caribbean spinly lobster fishery (\emph{Panulirus
+communities is the Caribbean spiny lobster fishery (\emph{Panulirus
 argus}), which is carried out within their respective TURFs. These
@@
 
-To perform the evaluation of these reserves we use thre sources of
-information. Ecological data come from the annual ecological monitorings
+To perform the evaluation of these reserves we use thee sources of
+information. Ecological data come from the annual ecological monitoring
 of reserve and control areas, carried out by members from each community
@@
 (\href{www.cobi.org.mx}{COBI}). These monitorings record richness and
-abundances of fish and invertebrate species in the reservs and control
+abundances of fish and invertebrate species in the reserves and control
 sites. For fish census, size structures are also collected to derive
@@
 data for reserve and control sites before and after the implementation
-of the reserve. This provides us with a Before--After--Control--Impact
+of the reserve. This provides us with a Before-After-Control-Impact
 (\emph{i.e.} BACI) design that allows us to capture and control for
@@
 designs and causal inference techniques have proven effective to
-evaluate marine reserves, as they allow us to causaly attribute observed
-changes to the intervention
+evaluate marine reserves, as they allow us to causally attribute
+observed changes to the intervention
 \citep{moland_2013-VP,Villasenor-Derbez_2018}. All reserves were
-surveyed annualy from at least one year before implementation until
-2016. Table \ref{tab:com_sum} shows a summary of the number of reserves,
-year of implementation, and number of transects for each reserve.
+surveyed annually from at least one year before implementation until
+2016. Table \ref{table:com_sum} shows a summary of the number of
+reserves, year of implementation, and number of transects for each
+reserve.
 
@@
 
-\caption{\label{tab:unnamed-chunk-2}\label{tab:com_sum} Summary of commuity--based marine reserves by community.
+\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of commuity--based marine reserves by community.
 Imp = Year of implementation, Start = Year of first sampling, number of fish transects in control (Cf) and reserve (Rf) sites, and number of invertebrate transects in Control (Ci) and Reserve (Ri) sites.}
@@
 
-\textbf{Governance data were collected at the community--level, using
-key informants to collect the necesary information on X, Y, and Z. Algo
-mas de Ere.}
+Governance data were collected at the community-level. The information
+was compiled by combining key informants and the authors; experience and
+knowledge of the communities to collect the necessary information. These
+data contain information on the ecological system where the fishing
+activities develop, as well as the governance structures present in the
+cooperative. We also gathered information on the resource unit
+(\emph{i.e.} lobsters) and the relevant actors present in each community
+\citep{leslie_2015-na}.
 
@@
 Following a framework that relates reserve objectives to performance
-indicators \citep{Villasenor-Derbez_2018}, we use five biological, two
-socioeconomic, and five governance indicators to evaluate these marine
-reserves Table \ref{table:indicators}.
+indicators \citep{Villasenor-Derbez_2018}, we use five biological and
+two socioeconomic indicators to evaluate these marine reserves Table
+\ref{table:indicators}. We also use a set of governance indicators to
+analyze the governance structures of each cooperative
+\citep{leslie_2015-na}. The indicators (Table \ref{table:gov_ind}) focus
+on the resource system (four indicators), governance system (seven
+indicators), resource units (three indicators) and actors (three
+indicators).
 
@@
 \hline
-Governance & Type of access to the fishery\\
+\end{tabular}
+\end{table}
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-4}\label{table:gov_ind}Indicators used for the operationalization of the SES framework \citep{leslie_2015-na}}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{8em}|>{\raggedright\arraybackslash}p{8em}|>{\raggedright\arraybackslash}p{8em}}
 \hline
- & Perceived degree of illegal fishing\\
+Indicator & Isla Natividad & Maria Elena & Punta Herrero\\
 \hline
- & Reserve surveilance anf enforcement\\
+\textbf{Resource systems (RS)} & \textbf{} & \textbf{}\\
 \hline
- & Type of fishing organization\\
+TURF presence & Yes & Yes & Yes\\
 \hline
- & Age of the reserve\\
+Type of ecosystem & Kelp Forest / Rocky Reefs & Coral Reef & Coral Reef\\
+\hline
+Intensity of environmental Disturbance & El nino event & Hurricanes & Hurricanes\\
+\hline
+Location & Island & Coastal & Coastal\\
+\hline
+\textbf{Governance systems (GS)} & \textbf{} & \textbf{}\\
+\hline
+Fishing cooperative & Yes & Yes & Yes\\
+\hline
+Involved actors & COBI, Stanford, REBIVI & Alianza Kanan Kay, COBI, CONANP, Coop, CONAPESCA, Oceanus, FCyRH, FHMM, & Alianza Kanan Kay, COBI, CONANP, Coop, CONAPESCA, Oceanus, FCyRH, FHMM,\\
+\hline
+Presence of an inter-cooperative structure & Fedecoop & Non & Non\\
+\hline
+Fishing Regulations & Size limits, seasonal closures, quotas & Size limits, seasonal closures & Size limits, seasonal closures\\
+\hline
+Enforcement technology & Boats & Boats & Land enforcement\\
+\hline
+MR enforcement &  &  & \\
+\hline
+Cooperative regulations &  &  & \\
+\hline
+\textbf{Resource Units (RU)} & \textbf{} & \textbf{}\\
+\hline
+Adult targeted species mobility & 1km & 30km & 30km\\
+\hline
+Targeted species longevity (years) &  &  & \\
+\hline
+Price of targeted species &  &  & \\
+\hline
+\textbf{Actors (A)} & \textbf{} & \textbf{}\\
+\hline
+Leadership &  &  & \\
+\hline
+Level of illegal fishing & 1 & 1 & 3\\
+\hline
+Presence of alternative livelihoods &  &  & \\
 \hline
@@
 to the first year of evaluation in the control site (See Table
-\ref{tab:com_sum}). When evaluating biomass and abundances, we include
+\ref{table:com_sum}). When evaluating biomass and abundances, we include
 species--fixed effects (\(\sigma_j\)). For abundances and richness
@@
 
-\subsubsection{Governance}\label{governance}
-
-\textbf{Texto de ere y los SES}
-
 \clearpage
@@
 between the control site before the implementation of the reserve and
-the reserve site at time \(t\) after controling for other time and space
-variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Here
-we present the effect that marine reserves had on each of the biological
-indicators for each coastal community, along with the trends in
-socioeconomic indicators of lobster catches and revenues. We also
+the reserve site at time \(t\) after controlling for other time and
+space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
+Here we present the effect that marine reserves had on each of the
+biological indicators for each coastal community, along with the trends
+in socioeconomic indicators of lobster catches and revenues. We also
 provide an overview of the state of the socioeconomic and governance
@@
 example, the reserve had a small effect on fish abundances (Fig
-\ref{fig:indicators}A), where only year 2010 showed significat effect
+\ref{fig:indicators}A), where only year 2010 showed significant effect
 sizes in fish abundances (\(p<0.05\)) and all other years oscillated
@@
 significant increases in lobster densities in the order of
-\(\lambda_t = 10\) (\(p < 0.05\)). Finally, punta herrero presented
+\(\lambda_t = 10\) (\(p < 0.05\)). Finally, Punta Herrero presented
 alternating negative and positive effects, but these were not different
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{\label{fig:unnamed-chunk-4}\label{fig:indicators}Effect sizes for
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{\label{fig:unnamed-chunk-5}\label{fig:indicators}Effect sizes for
 marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{\label{fig:unnamed-chunk-5}\label{fig:res_com}Summarized effects of
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{\label{fig:unnamed-chunk-6}\label{fig:res_com}Summarized effects of
 the marne reserves by direction (positive - negative) and significance.}
@@
 
+Lobster catches and income showed a increase after the implementation of
+the reserves for Isla Natividad and Maria Elena, but not for Punta
+Herrero (Fig \ref{fig:lobsters}). However, the differences in catches
+and and revenue were not different in the periods before and after the
+implementation (\(p > 0.05\)) except for revenues in Isla Natividad,
+which presented a significant increase of 14.37 (M MXP; \(p<0.05\)).
+Table \ref{table:soc_res} presents all the regression coefficients.
+
+\begin{table}[!htbp] \centering 
+  \caption{\label{table:soc_res} Regression coefficients for lobster catches and revenues for Isla Natividad (1, 4), Maria Elena (2, 5), and Punta Herrero (3, 6).} 
+  \label{} 
+\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-7} 
+\\[-1.8ex] & \multicolumn{3}{c}{Catches (tones)} & \multicolumn{3}{c}{Revenue(M MXP)} \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
+\hline \\[-1.8ex] 
+ Post & 7.37 & 5.83 & $-$1.26 & 14.37$^{***}$ & 1.24 & $-$0.06 \\ 
+  Constant & 122.68$^{***}$ & 11.41$^{***}$ & 6.06$^{***}$ & 10.89$^{***}$ & 2.04$^{***}$ & 1.06$^{***}$ \\ 
+ \hline \\[-1.8ex] 
+Observations & 15 & 14 & 13 & 15 & 14 & 13 \\ 
+R$^{2}$ & 0.02 & 0.04 & 0.01 & 0.52 & 0.04 & 0.0004 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
+
 \begin{figure}
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
-\caption{\label{fig:unnamed-chunk-6}\label{fig:lobsters}Time series of
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-8-1.pdf}
+\caption{\label{fig:unnamed-chunk-8}\label{fig:lobsters}Time series of
 lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
@@
 
-\subsection{Governance}\label{governance-1}
+\subsection{Governance}\label{governance}
+
+Although we have little information on the social dimension of these
+fisheries, using the SES framework indicators (Table
+\ref{table:gov_ind}), we can analyze the performance of each governance
+system with respect to MR enforcement (Table \ref{table:gov_res}). In
+general, the presence and success of conservation initiatives depends on
+the incentives of local communities to maintain a healthy status of the
+resources they depend upon \citep{jupiter_2017}. The enabling conditions
+for conservation seem to be strongly present in Isla Natividad. Due to
+the clarity of access rights and isolation, the benefits of conservation
+directly benefit the members of the fishing cooperative. These
+conditions have favored the development of an efficient community based
+enforcement system. In contrast, the communities of Maria Elena and
+Punta Herrero are located near other fishing communities and cities. In
+Maria Elena, the fishing pressure caused by outsiders can be reduced by
+implementing a strong enforcement system (in water and land) supported
+by CSOs and the local government (CONANP). Lastly, the community of
+Punta Herrero shows the highest levels of illegal activities which can
+be attributed to its connectedness to other communities and the lack of
+appropriate technologies for enforcement.
+
+\begin{table}
+
+\caption{\label{tab:unnamed-chunk-9}\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \citep{mcginnis_2014}.}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}|l|l|l}
+\hline
+ & Indicator & Isla Natividad & Maria Elena & Punta Herrero\\
+\hline
+\textbf{Resource systems (RS)} & \textbf{} & \textbf{} & \textbf{}\\
+\hline
+RS2 – Clarity of system boundaries & TURF presence & High & High & High\\
+\hline
+RS3 – Size of resource system &  &  &  & \\
+\hline
+RS5 – Productivity of system & Type of ecosystem & High & High & High\\
+\hline
+RS7 – Predictability of system dynamics & Intensity of environmental disturbance & Low (ENSO) & High & High\\
+\hline
+RS9 – Location & Proximity to other communities/cities & Isolated & Not Isolated & Not Isolated\\
+\hline
+\textbf{Governance systems (GS)} & \textbf{} & \textbf{} & \textbf{}\\
+\hline
+GS1 – Government organizations & Presence of fishing cooperatives & Yes & Yes & Yes\\
+\hline
+GS2 – Nongovernment organizations & Involved actors & Yes & Yes & Yes\\
+\hline
+GS3 – Network structure & Presence of an inter-cooperative structure & Yes & No & No\\
+\hline
+GS4 – Property-rights systems & TURF presence & Yes & Yes & Yes\\
+\hline
+GS5 – Operational-choice rules & Fishing Regulations / MPA enforcement / Enforcement technolofy & Yes & Yes & Yes\\
+\hline
+GS6 – Collective-choice rules & Cooperative regulations & Yes & Yes & Yes\\
+\hline
+GS7 – Constitutional-choice rules &  &  &  & \\
+\hline
+\textbf{Resource units (RU)} & \textbf{} & \textbf{} & \textbf{}\\
+\hline
+RU1 – Resource unit mobility & Targeted species home range & Low & Medium & Medium\\
+\hline
+RU2 – Growth or replacement rate & Max age of targeted species & Low & Medium & Medium\\
+\hline
+RU4 – Economic value & Price of targeted species & high & High & high\\
+\hline
+\textbf{Actors (A)} & \textbf{} & \textbf{} & \textbf{}\\
+\hline
+A1 – Number of relevant actors &  & 98 &  & \\
+\hline
+A2 – Socioeconomic attributes &  &  &  & \\
+\hline
+A5 – Leadership/entrepreneurship & Leadership & High & High & High\\
+\hline
+A6 – Norms (trust-reciprocity)/social capital--- (Based on illegal fishing) & Level of illegal fishing & High & High & Low\\
+\hline
+A8 – Importance of resource (dependence) & Presence of alternative livelihoods & High & High & High\\
+\hline
+\end{tabular}
+\end{table}
+
+\clearpage
 
@@
 
-The biological effects of the reserve contrast the existing literature.
-Why no effect on bio? IN hypoxia. ME and PH, age? Perhaps is the use of
-causal inference methods?
+Our results indicated idiosyncratic biological effects of the reserves
+across communities and indicators, with a combination of positive and
+negative effects. However, many of these effects were not statistically
+significant, indicating no effect of the reserve \ref{fig:res_com}. The
+socioeconomic indicators pertaining to landings and revenues associated
+to those landings showed little or no temporal change before and after
+reserve implementation. These contrasting effects, however, might be
+clarified when understanding the social-ecological context in which
+these communities and their reserves sit. In this section, we discuss
+potential shortcomings in our analysis, and provide plausible
+explanations to the observed biological and socioeconomic basing on
+previous literature and our analysis of the social-ecological system.
 
-Isla natividad no funciona por hipoxya
+The contrasting biological effectiveness observed is perhaps explained
+by our approach to evaluate the temporal and spatial changes of each
+indicator. Some works have solely focused on an inside-outside
+comparison of indicators
+\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD}, which do
+not address temporal variability \citep{depalma_2018}. Other works have
+compared the trend observed within a reserve through time
+\citep{betti_2017-lq}, which cannot distinguish between the temporal
+trends in a reserve and the entire system \citep{depalma_2018}. By
+accounting for trends between sites and through times, we can control
+for time and space dynamics, and provide a better identification of the
+effect. However, it is worth looking deeper into each case, and
+identifying other plausible explanations.
 
-Maria Elena funciona ``bien''
+Age, isolation, and enforcement are important factors influencing
+effectiveness of a marine reserve \citep{edgar_2014-UO}. Isla Natividad
+has the oldest reserve, is fairly isolated, and has a well-established
+community-based enforcement system. While other communities are
+certainly within reach, these are known to be well organized fishing
+communities with successful resource management
+\citep{mccay_2017-1m,mccay_2014-CN}. The reserve at Isla Natividad
+presented the largest percentage of significantly positive changes in
+biological indicators (19\%), but an important portion of was also
+negative (15\%). With the age, relative isolation, and enforcement level
+of this reserve, it would be expected for it to be considerably
+effective. The potential gap in performance can be attributed to
+perturbations that do not distinguish reserve boundaries, such as
+environmental variability (\textbf{no recuerdo esta cita}). The region
+is known to be under the influence of recurrent hypoxia and
+high-temperature events known to cause massive adult mortalities
+\citep{micheli_2012-EU}.
 
-Punta Herrero no funciona por poaching.
+Maria Elena and Punta Herrero are relatively young reserves (See Table
+\ref{table:com_sum}). From these, the Maria Elena exhibited the highest
+performance in terms of biological indicators (15\% significantly
+positive). In contrast, Punta Herrero had a similar proportion of
+positive and negative effects.
 
-Aun con las mejores caracteristicas sociales, noe s posible vencer al
-ambiente (IN). Para el Caribe, parece ser que buena vigilancia y poca
-pesca ilegal son determinantes para obtener buenos resultados.
-Preguntarle a Stuart si ME solamente se pesca durante lanosta, y todo el
-ano es reserva.
+The way in which we measure changes in catches and revenues can not
+identify whether the observed differences are simply caused by
+pre-existing temporal trends or by the implementation of the reserve.
+Yet, there were no detectable changes in these indicators, except for
+landings in Isla Natividad. Other research has shown that reserves in
+Isla Natividad yield fishery benefits for the abalone fishery
+\citep{rossetto_2015-V0}. Since the trend was not detected in catches
+--directly related to abundance and fishing effort-- it is plausible
+that these differences are purely explained by an increase in
+market-level prices.
 
-IN captures more dynamics. Reflects hypoxia events. Citar paper Giron
+The fact that there was no detectable change in catches for Maria Elena
+and Punta Herrero can be explained by a combination of factors related
+to the design, management, age, or ecological factors. Reserves in these
+communities are relatively small and young, and may need more time for
+lobster abundances to increase enough to export larvae or adult
+organisms. Other community-based marine reserves in tropical ecosystems
+have taken up to six years to show a spillover effect
+\citep{dasilva_2015-zX}. A complimentary explanation lies in the results
+observed for the governance system. The lack of enforcement in Punta
+Herrero, for example, could explain the lack of effectiveness observed
+in their reserves.
 
-Differences between fish and inverts sheds a lights on direct and
-indirect effects (Paper que recomendo Fio).
+Our results show that community-based marine reserves can be effective
+if the environmental and social settings allow it. By studying the
+social-ecological system as a whole, we can provide a wider range of
+explanations to the patterns observed. It is interesting that even under
+the best enabling social conditions, climate variability can hinder the
+effect of a reserve --Although it is interesting to imagine what the
+state of that fishery had been if the reserve and organized cooperative
+were not present--. On the contrary, we show how under low climate
+variability, absence of proper governance structures can limit the
+effectiveness and benefits of a reserve. Whether the combination of a
+stable environment and governance structures are additive or
+multiplicative represents an interesting area for future research,
+especially under a changing climate.
 
@@
 
-JC and EA analyzed and interpreted data, discussed the results and wrote
-the manuscrip. SF and JT edited the manuscript and discussed the
-results.
+JC and EA analyzed and interpreted data, discussed the results, and
+wrote the manuscript. AS, SF and JT edited the manuscript and discussed
+the results.
 
@@
 
-Details of all funding sources should be provided, including grant
-numbers if applicable. Please ensure to add all necessary funding
-information, as after publication this is no longer possible.
+JCVD CONACyT + LAFF ASC SF JT
 
@@
 
-This is a short text to acknowledge the contributions of specific
-colleagues, institutions, or agencies that aided the efforts of the
-authors.
+The authors wish to acknowledge Arturo Hernández and Imelda Amador for
+contributions on the governance data, as well as pre-processing
+biological data. This study would have not been possible without the
+effort by members of the communities here mentioned, who collected the
+biological data.
 

### 2018-02-21 Abstract_WSSFConf.Rmd
Added abstract to the paper

---
title: "Enabling conditions for effective community based marine reserves in small scale fisheries"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Alvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
date: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA. $^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico"
output: pdf_document
---

# Abstract

Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. However, overfishing and unsustainable fishing practices threaten the marine environment and jeopardize the wellbeing of coastal communities. A common approach to protect the environment and recover overexploited stocks is to implement no-take marine reserves (areas where all extractive activities are off-limits). In small-scale fisheries, these are sometimes implemented as community-based reserves, where a group of fishers collectively agree to close an area to fishing. While we know that reductions in fishing effort are followed by a series of ecological benefits (increased biomass, abundance, and species diversity), we do not fully understand how environmental and governance dynamics influence the conservation and fisheries benefits of community-based marine reserves. In this work, we evaluate the ecological outcomes of four reserves established by three coastal communities in temperate and tropical ecosystems of Mexico. By combining causal inference techniques with an operationalization of the social-ecological systems framework, we identify the environmental and social conditions that enable reserve effectiveness. Our results show a strong interaction between environmental variation and community organization, which influences reserve effectiveness. For example, the most effective reserve had strong governance structures accompanied with low environmental variability. Thus, even when a community is well organized (and reserves are well enforced), environmental variation can hinder the benefits of a reserve, and vice versa. Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making.

# How does this paper address the transdisciplinarity and/or transformation of small-scale fisheries?

Our methodological approach uses state-of-the-art econometric techniques to evaluate the causal effect of marine reserves on a series of ecological indicators, which enable us to control for spatiotemporal dynamics. By combining changes these with catch and revenues from associated fisheries, along with the social-ecological systems framework and information generated by the communities themselves through a large-scale citizen science program, we are able further identify the conditions that enable reserve effectiveness.

#What are the key messages in your paper about the future and/or sustainability of small-scale fisheries?

-	Reserve effectiveness can be better explained by analyzing the effect of both environmental variation and governance.
-	The benefits of community-based conservation can be diminished by environmental variation even in the presence of strong governance systems.

### 2018-02-21 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Added abstract to the paper

@@
 email: "jvillasenor@bren.ucsb.edu"
-abstract: "Abstract after comments from co-authors"
+abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
+However, overfishing and unsustainable fishing practices threaten the marine environment and jeopardize the wellbeing of coastal communities.
+A common approach to protect the environment and recover overexploited stocks is to implement no-take marine reserves (areas where all extractive activities are off-limits).
+In small-scale fisheries, these are sometimes implemented as community-based reserves, where a group of fishers collectively agree to close an area to fishing.
+While we know that reductions in fishing effort are followed by a series of ecological benefits (increased biomass, abundance, and species diversity), we do not fully understand how environmental and governance dynamics influence the conservation and fisheries benefits of community-based marine reserves.
+In this work, we evaluate the ecological outcomes of four reserves established by three coastal communities in temperate and tropical ecosystems of Mexico.
+By combining causal inference techniques with an operationalization of the social-ecological systems framework, we identify the environmental and social conditions that enable reserve effectiveness.
+Our results show a strong interaction between environmental variation and community organization, which influences reserve effectiveness.
+For example, the most effective reserve had strong governance structures accompanied with low environmental variability.
+Thus, even when a community is well organized (and reserves are well enforced), environmental variation can hinder the benefits of a reserve, and vice versa.
+Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making."
 keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico, Social--Ecological Systems"
@@
 ```
+
+\clearpage
 

### 2018-02-21 Frontiers_Manuscript/Villasenor-Derbez.tex
Added abstract to the paper

@@
 
-Abstract after comments from co-authors
+Coastal marine ecosystems provide livelihoods for small-scale fishers
+and coastal communities around the world. However, overfishing and
+unsustainable fishing practices threaten the marine environment and
+jeopardize the wellbeing of coastal communities. A common approach to
+protect the environment and recover overexploited stocks is to implement
+no-take marine reserves (areas where all extractive activities are
+off-limits). In small-scale fisheries, these are sometimes implemented
+as community-based reserves, where a group of fishers collectively agree
+to close an area to fishing. While we know that reductions in fishing
+effort are followed by a series of ecological benefits (increased
+biomass, abundance, and species diversity), we do not fully understand
+how environmental and governance dynamics influence the conservation and
+fisheries benefits of community-based marine reserves. In this work, we
+evaluate the ecological outcomes of four reserves established by three
+coastal communities in temperate and tropical ecosystems of Mexico. By
+combining causal inference techniques with an operationalization of the
+social-ecological systems framework, we identify the environmental and
+social conditions that enable reserve effectiveness. Our results show a
+strong interaction between environmental variation and community
+organization, which influences reserve effectiveness. For example, the
+most effective reserve had strong governance structures accompanied with
+low environmental variability. Thus, even when a community is well
+organized (and reserves are well enforced), environmental variation can
+hinder the benefits of a reserve, and vice versa. Our results are
+particularly relevant under present changing climate conditions, as they
+can better inform management and decision making.
 
@@
 \end{abstract}
+
+\clearpage
 

### 2018-02-27 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Added xaringan presentation

@@
 ```{r load_spatial}
-
 # Determine a projection for the maps
@@
   scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
   geom_vline(xintercept = c(2006, 2012, 2013), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +

### 2018-02-27 GainesLabPresentation.Rmd
Added xaringan presentation

---
title: "Community-based marine reserves in Mexico: their effectiveness"
subtitle: "(Or their lack of?)"
author: "Juan Carlos Villaseñor-Derbez"
date: "27/02/18"
output: xaringan::moon_reader
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Background

In Mexico:

- Increased advocacy for marine reserves

--

- These are implemented as:

--

  - No-take zone within MPA (bottom-up)

--

  - Community-based (usually within a TURF)

--

  - Fishing refugia (a community-based, with legal recognition)

---

# Background

In Mexico:

- Increased advocacy for marine reserves

- These are implemented as:

    - No-take zone within MPA (bottom-up)

    - **Community-based (usually within a TURF)**

    - **Fishing refugia (a community-based, with legal recognition)**

---

```{r}
suppressPackageStartupMessages({
  library(sp)
  library(sf)
  library(ggsn)
  library(tmap)
  library(magrittr)
  library(kableExtra)
  library(broom)
  library(tidyverse)
})
```

# Where

```{r load_spatial}
# Determine a projection for the maps
proj <- CRS("+proj=longlat +datum=NAD27")

# Load coastline for Mexico and World
data(World)
W <- spTransform(World, proj) %>% 
  st_as_sf()

# Load high-res mexican coastline
coastline_mx <- readRDS("./Data/Spatial/coastline_mx.rds")

# Declare points for reference map
location_labels <- data.frame(X = c(-115.24, -87.4611756),
                              Y = c(27.85, 19.3183386),
                              Location = c("Isla Natividad", "Punta Herrero"),
                              Region = c("Pacific", "Caribbean"),
                              N = c("N = 1", "N = 3"),
                              stringsAsFactors = F)

zrp <- st_read(dsn = "./Data/Spatial", layer = "ZRP", quiet = T) %>% 
  sf::st_centroid() %>% 
  select(Name, geometry) %>% 
  mutate(Type = "FR")

rescom <- st_read(dsn = "./Data/Spatial", layer = "Res_Com", quiet = T) %>% 
  sf::st_centroid() %>% 
  select(Name = Nombre, geometry) %>% 
  mutate(Type = "CB")

reserves <- rbind(zrp, rescom)

ggplot() +
  geom_sf(data = W) +
  geom_sf(data = reserves, aes(fill = Type), shape = 21) +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = c(0.9, 0.5)) +
  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863))
```

---

# Where

```{r}
ggplot() +
  geom_sf(data = W) +
  theme_classic() +
  geom_point(data = location_labels, aes(x = X, y = Y), fill = "red", size = 3, shape = 21) +
  geom_text(data = location_labels, aes(x = X, y = Y, label = N), nudge_x = -1.5) +
  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863))
```

---

# Brief context

```{r}
data.frame(Community = c("Isla Natividad", "Maria Elena", "Punta Herrero"),
           N_reserves = c(2, 3, 3),
           N_reserves2 = c("1", "1", "2*"),
           Size = c(10.58, 483.98, 153.13),
           Implemented = c(2006, 2012, 2013)) %>% 
  knitr::kable(format = "html", col.names = c("Community", "Reserves", "Evaluated", "Size (ha)", "Implemented"))
```

--

- All have cooperatives and TURF's

--

- The objectives are *"mixed"* on paper (consrvation + fisheries)
- Ask a fisher, they'll say:

> "We want to increase lobsters populations and catches"

---

![](./Img/Imagen5.jpg)

Lobster is their most important resource

- (*Panulirus interruptus* and *P. argus*)

---
class: inverse

# Questions

--

## Are these community-based marine reserves effective?

--

## Are they effective for lobster?

--

## What are the enabling conditions?

---

# Methods

Data:

--

- Fish and invert surveys BACI

--

- Timeseries of landings and income BA

--

- Data on how the communities operate

---

# Methods

## Biological

- Biomass

- Abundances

  - Lobster abundances

- Richness

$$I_{i,t,s} = \alpha_{i} + \sum_{t = 1}^T\gamma_{t} Y_t + \beta Z_i +  \sum_{t = 1}^T\lambda_{t} Y_t\times Z_i + \sum_{s = 1}^S\sigma_s S_s + \epsilon$$

- I care about $\lambda_{t}$

---

## Fisheries side

- Change in lobster income and landings

## Social / governance data

- Thanks Ere!

  - Rules and structures under which decisions are made

  - Use this to help identify "enabling conditions"

---

# Results

Are these community-based marine reserves effective?

![](./Img/plot.png)

---

# Results

Are these community-based marine reserves effective?

![](./Img/plot_summary.png)

---

## Are they effective for lobster?

Lobster abundances

![](./Img/Lobsters.png)

---

## Are they effective for lobster?

<table style="text-align:center">
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"></td><td colspan="6"><em>Dependent variable:</em></td></tr>
<tr><td></td><td colspan="6" style="border-bottom: 1px solid black"></td></tr>
<tr><td style="text-align:left"></td><td colspan="3">Catches (tones)</td><td colspan="3">Revenue(M MXP)</td></tr>
<tr><td style="text-align:left"></td><td>(IN)</td><td>(ME)</td><td>(PH)</td><td>(IN)</td><td>(ME)</td><td>(PH)</td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">PostbPost</td><td>7.37</td><td>5.83</td><td>-1.26</td><td><strong>14.37<sup>***</sup></strong></td><td>1.24</td><td>-0.06</td></tr>
<tr><td style="text-align:left">Constant</td><td>122.68<sup>***</sup></td><td>11.41<sup>***</sup></td><td>6.06<sup>***</sup></td><td>10.89<sup>***</sup></td><td>2.04<sup>***</sup></td><td>1.06<sup>***</sup></td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Observations</td><td>15</td><td>14</td><td>13</td><td>15</td><td>14</td><td>13</td></tr>
<tr><td style="text-align:left">R<sup>2</sup></td><td>0.02</td><td>0.04</td><td>0.01</td><td>0.52</td><td>0.04</td><td>0.0004</td></tr>
<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"><em>Note:</em></td><td colspan="6" style="text-align:right"><sup>*</sup>p<0.1; <sup>**</sup>p<0.05; <sup>***</sup>p<0.01</td></tr>
</table>

---

## Are they effective for lobster?

![](./Img/econ2.png)

---

## Are they effective for lobster?

![](./Img/econ1.png)

---

# Results

Enabling conditions

```{r}
data.frame(Feature = c("No-take", "Well-enforced", "Old (> 10yrs)", "Large", "Isolated", "Presence of mgmt plan", "Fisher engange in mgmt", "Fishers make decisions"),
           IN = c(1, 1, 1, 0, 1, 0.5, 1, 1), 
           ME = c(1, 1, 0, 0, 0.5, 0.5, 1, 1),
           PH = c(1, 0.5, 0, 0, 0.5, 0, 1, 1)) %>% 
  knitr::kable(format = "html")
```

From: Edgar *et al.*, (2014); Di Franco *et al.* (2016)

---

#

### 2018-02-27 GainesLabPresentation.Rmd
Added presentation

@@
 ---
-title: "Community-based marine reserves in Mexico: their effectiveness"
-subtitle: "(Or their lack of?)"
-author: "Juan Carlos Villaseñor-Derbez"
+title: "Community-based marine reserves in Mexico: do they work?"
+author: "JC"
 date: "27/02/18"
-output: xaringan::moon_reader
+output:
+  xaringan::moon_reader:
+    nature:
+      slideNumberFormat: ""
 ---
@@
 ```
+
+# Before we start
+
+- This is onging work
+
+- Other collaborators
+
+- Help me better tell the story (if there is one)
+
+> Even when a community is very well organized, environmental variation can hinder reserve effectiveness
+
+> Even if environment is stable, lack of community organization can hinder reserve effectiveness
+
+---
 
@@
            N_reserves2 = c("1", "1", "2*"),
-           Size = c(10.58, 483.98, 153.13),
+           Size = c(10.58, 9.73, 527.79),
            Implemented = c(2006, 2012, 2013)) %>% 
@@
 
-- The objectives are *"mixed"* on paper (consrvation + fisheries)
-- Ask a fisher, they'll say:
-
-> "We want to increase lobsters populations and catches"
+- The objectives are *"mixed"* on paper (conservation + fisheries)
 
@@
 
+- Ask a fisher, they'll say:
+
+> "We want to increase lobster populations and catches"
+
 ![](./Img/Imagen5.jpg)
-
-Lobster is their most important resource
-
-- (*Panulirus interruptus* and *P. argus*)
 
@@
 
-## What are the enabling conditions?
+## How do environmental and social dynamics influence the outcomes of community-based marine reserves?
 
@@
 
-Data:
+Combine:
+
+- Causal inference
+
+- Operationalization of the social-ecological systems framework (Gracias Ere!)
 
 --
+
+Data:
 
@@
 
-## Fisheries side
+## Fisheries-side
 
@@
 ## Social / governance data
-
-- Thanks Ere!
 
@@
 
+.pull-left[
+Some general trends:
+- Positive effects in invertebrate abundances
+- Fish richness post 2010 takes a dive (IN)
+- Negative effect in invertebrate richness (IN)
+- Not much more...
+]
+
+.pull-right[
+![](./Img/plot.png)
+]
+
+---
+
+# Results
+
 Are these community-based marine reserves effective?
@@
 
-## Are they effective for lobster?
+# Results
+
+Are they effective for lobster?
 
@@
 
-## Are they effective for lobster?
+# Results
+
+Are they effective for lobster?
+
+- Landings and income
 
@@
 <tr><td></td><td colspan="6" style="border-bottom: 1px solid black"></td></tr>
-<tr><td style="text-align:left"></td><td colspan="3">Catches (tones)</td><td colspan="3">Revenue(M MXP)</td></tr>
+<tr><td style="text-align:left"></td><td colspan="3">Landings (tones)</td><td colspan="3">Revenue(M MXP)</td></tr>
 <tr><td style="text-align:left"></td><td>(IN)</td><td>(ME)</td><td>(PH)</td><td>(IN)</td><td>(ME)</td><td>(PH)</td></tr>
-<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">PostbPost</td><td>7.37</td><td>5.83</td><td>-1.26</td><td><strong>14.37<sup>***</sup></strong></td><td>1.24</td><td>-0.06</td></tr>
+<tr><td colspan="7" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Post</td><td>7.37</td><td>5.83</td><td>-1.26</td><td><strong>14.37<sup>***</sup></strong></td><td>1.24</td><td>-0.06</td></tr>
 <tr><td style="text-align:left">Constant</td><td>122.68<sup>***</sup></td><td>11.41<sup>***</sup></td><td>6.06<sup>***</sup></td><td>10.89<sup>***</sup></td><td>2.04<sup>***</sup></td><td>1.06<sup>***</sup></td></tr>
@@
 
-![](./Img/econ2.png)
+Income
+
+```{r, warning = F, message = F}
+
+load("./Data/conapesca_2000-2015.RData")
+
+cpi <- read.csv("./Data/CPI.csv", stringsAsFactors = F, strip.white = T)
+
+cpi_2014 <- cpi$CPI[cpi$Ano == 2014]
+
+conapesca %<>%
+  filter(UnidadEconomica %in% c("Scpp cozumel scl",
+                                "Scpp jose maria azcorra scl",
+                                "Scpp buzos y pescadores de la baja california scl")) %>% 
+  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
+  group_by(UnidadEconomica, Ano, NombreComun) %>%
+  summarize(PesoDesembarcado = sum(PesoDesembarcado), 
+            PesoVivo = sum(PesoVivo),
+            Valor = sum(Valor)) %>%
+  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
+  left_join(cpi, by = "Ano") %>% 
+  mutate(Valor = Valor * (CPI / cpi_2014))
+
+L_conapesca <- tibble(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl",
+                                          "Scpp cozumel scl",
+                                          "Scpp jose maria azcorra scl"),
+                      Comunidad = c("Isla Natividad",
+                                    "Maria Elena",
+                                    "Punta Herrero")) %>% 
+  left_join(conapesca, ., by = "UnidadEconomica") %>% 
+  mutate(Post = case_when(Comunidad == "Isla Natividad" && Ano <= 2006 ~ 0,
+                          Comunidad == "Maria Elena" && Ano <= 2012 ~ 0,
+                          Comunidad == "Punta Herrero" && Ano <= 2013 ~ 0,
+                          TRUE ~ 1))
+
+L_conapesca2 <- L_conapesca %>% 
+  group_by(Comunidad, Post) %>% 
+  summarize(meanC = mean(PesoVivo/1000),
+            meanI = mean(Valor/1e6))
+
+left_join(L_conapesca, L_conapesca2, by = c("Comunidad", "Post")) %>% 
+  ggplot(aes(x = Ano, y = Valor/1e6, color = Comunidad, group = paste(Comunidad, Post), fill = Comunidad)) +
+  geom_line(aes(y = meanI), size = 1) +
+  geom_point(aes(shape = Comunidad), size = 2) +
+  scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
+  geom_vline(xintercept = c(2006.5, 2012.5, 2013.5), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +
+  labs(x = "Year", y = "Income (M MXP)") +
+  theme(legend.position = "none",
+        text = element_text(size = 12)) +
+  cowplot::theme_cowplot()
+```
 
@@
 
-![](./Img/econ1.png)
+Landings
+
+```{r}
+left_join(L_conapesca, L_conapesca2, by = c("Comunidad", "Post")) %>% 
+  ggplot(aes(x = Ano, y = PesoVivo/1e3, color = Comunidad, group = paste(Comunidad, Post), fill = Comunidad)) +
+  geom_line(aes(y = meanC), size = 1) +
+  geom_point(aes(shape = Comunidad), size = 2) +
+  scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
+  geom_vline(xintercept = c(2006.5, 2012.5, 2013.5), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +
+  labs(x = "Year", y = "Landings (Tones)") +
+  theme(legend.position = "none",
+        text = element_text(size = 12)) +
+  cowplot::theme_cowplot()
+```
 
@@
 # Results
+
+Isla Natividad
+
+- Designed for invertebrates
+
+--
+
+- Very well organized community
+
+--
+
+- Relatively isolated, but still surrounded by "good fishers"
+
+--
+
+- Reserve is well enforced
+
+--
+
+- Reserve has been in place quite some time
+
+--
+
+- Should be the perfect example... but
+
+--
+
+- Has had mass mortality events due to hypoxia (2009, 2010) and warm water (~2015)
+
+---
+
+# Results
+
+Maria Elena
+
+- Also intended to protect grouper reproductive aggregations
+
+--
+
+- Well enforced
+
+--
+
+- Community is very well organized
+
+--
+
+- No hypoxia or other environmental disturbances
+
+--
+
+- Small (~ 10 ha)
+
+--
+
+- Fairly young (5 years of data)
+
+---
+
+# Results
+
+Punta Herrero
+
+--
+
+- Also intended to protect grouper reproductive aggregations
+
+--
+
+- Good size (~ 500 ha)
+
+--
+
+- Low enforcement
+
+--
+
+- Community is not very well organized
+
+--
+
+- No hypoxia or other environmental disturbances
+
+--
+
+- Fairly young (4 years of data)
+
+---
+
+# Telling the story
+
+> Even when a community is very well organized, environmental variation can hinder reserve effectiveness
+
+> Even if environment is stable, lack of community organization can hinder reserve effectiveness
+
+## Missing pieces
+
+- Data on environmental disturbances (can get DO, S, T > 2009 for N)
+
+---
+background-image: url(https://raw.githubusercontent.com/jcvdav/ReserveEffect/master/Img/Imagen5.jpg)
+class: inverse, center, middle
+background-size: cover
+
+# Thanks!
+
+--
+
+## Questions?
+
+--
+
+## Suggestions?
+
+--
+
+## Comments?
+
+--
+
+### Anyone??
+
+--
+
+# Help!
+
+---
+
+---
+
+# Extras
 
@@
 From: Edgar *et al.*, (2014); Di Franco *et al.* (2016)
-
----
-
-#

### 2018-02-27 GainesLabPresentation.Rmd
small changes to presentation

@@
 
-- Operationalization of the social-ecological systems framework (Gracias Ere!)
+- Operationalization of the social-ecological systems framework
+  - (Gracias Ere!)
 
@@
 
+- Should be the perfect example:
+
+--
+
   - Designed for invertebrates
@@
 
-- Should be the perfect example... but
+- But:
 
---
-
-- Has had mass mortality events due to hypoxia (2009, 2010) and warm water (~2015)
+  - Mortality events due to hypoxia (2009, 2010) and warm water (~2015)
 
@@
 
-- Also intended to protect grouper reproductive aggregations
+- This is:
 
@@
 
+- But not a great design:
+
+  - Also intended to protect grouper reproductive aggregations
+
+--
+
   - Small (~ 10 ha)
@@
 
---
-
-- Also intended to protect grouper reproductive aggregations
+- Opposite to Isla Natividad:
 
@@
   - Good size (~ 500 ha)
+
+--
+
+  - No hypoxia or other environmental disturbances
+
+--
+
+- But:
+
+--
+
+  - Also intended to protect grouper reproductive aggregations
 
@@
   - Community is not very well organized
-
---
-
-- No hypoxia or other environmental disturbances
 

### 2018-03-20 docs/WSSFConf_abstract/Abstract_WSSFConf.Rmd
moved abstract to its own folde under docs

--- 2018-02-21 Abstract_WSSFConf.Rmd
@@
 ---
-title: "Enabling conditions for effective community based marine reserves in small scale fisheries"
+title: "Enabling conditions for effective community-based marine reserves in small-scale fisheries"
 author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Alvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"

### 2018-03-20 docs/paper/Villasenor-Derbez.Rmd
Added manuscript to its own folder

--- 2018-02-27 Frontiers_Manuscript/Villasenor-Derbez.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+
 options(knitr.table.format = "latex")
@@
 
-```{r loadp_kgs}
+```{r load_pkgs}
 suppressPackageStartupMessages({
@@
 # Load indicator functions
-source(file = "../Scripts/abundance.R")
-source(file = "../Scripts/richness.R")
-source(file = "../Scripts/shannon.R")
-source(file = "../Scripts/biomass.R")
-source(file = "../Scripts/trophic.R")
+source(file = here::here("scripts", "abundance.R")) #Calculate abundance per transect
+source(file = here::here("scripts", "richness.R")) # Calculate richness per transect
+source(file = here::here("scripts", "shannon.R")) #Calculate shannon's index per transect
+source(file = here::here("scripts", "biomass.R")) # Calculate fish biomass per transect
+source(file = here::here("scripts", "trophic.R")) # Calculate mean trophic level per transect
 
 # Load analysis functions
-source(file = "../Scripts/model.R") #fits the model
-source(file = "../Scripts/robust_se.R") #calculates heteroskedastic-robust Standard Errors
-source(file = "../Scripts/plot_model.R") #plots effect sizes
+source(file = here::here("scripts", "model.R")) #fits the model
+source(file = here::here("scripts", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
+source(file = here::here("scripts", "plot_model.R")) #plots effect sizes
 ```
@@
 # Load high-res mexican coastline
-coastline_mx <- readRDS("../Data/Spatial/coastline_mx.rds")
+coastline_mx <- readRDS(here::here("data", "spatial", "coastline_mx.rds"))
 
@@
 # Load ZRP shapefile
-zrp <- st_read(dsn = "../Data/Spatial", layer = "ZRP", quiet = T) %>% 
+zrp <- st_read(dsn = here::here("data", "spatial"), layer = "ZRP", quiet = T) %>% 
   filter(Name %in% c("El Cabezo", "El Faro")) %>% 
@@
 # Load Community reserve shapefile
-rescom <- st_read(dsn = "../Data/Spatial", layer = "Res_Com", quiet = T) %>% 
+rescom <- st_read(dsn = here::here("data", "spatial"), layer = "Res_Com", quiet = T) %>% 
   filter(Nombre == "La_Plana") %>% 
@@
 ```{r load_data}
-fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F) %>% 
+fish <- read.csv(here::here("data", "fish_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
   filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
@@
 
-invert <- read.csv("../Data/invert_data.csv", strip.white = T, stringsAsFactors = F) %>% 
+invert <- read.csv(here::here("data", "invert_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
   filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
@@
 
-load("../Data/conapesca_2000-2015.RData")
+load(here::here("data","conapesca_2000-2015.RData"))
 
-cpi <- read.csv("../Data/CPI.csv", stringsAsFactors = F, strip.white = T)
+cpi <- read.csv(here::here("data", "CPI.csv"), stringsAsFactors = F, strip.white = T)
 
@@
 
-gob <- read.csv("../Data/governance.csv", strip.white = T, stringsAsFactors = T)
+gob <- read.csv(here::here("data" ,"governance.csv"), strip.white = T, stringsAsFactors = T)
 ```
@@
 ```{r}
-read.csv("../Data/gov_ind.csv") %>% 
+read.csv(here::here("data", "gov_ind.csv")) %>% 
   knitr::kable(caption = "\\label{table:gov_ind}Indicators used for the operationalization of the SES framework \\citep{leslie_2015-na}", col.names = c("Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero"), format = "latex") %>%
@@
 ```{r}
-read.csv("../Data/gov_res.csv") %>% 
+read.csv(here::here("data","gov_res.csv")) %>% 
   knitr::kable(caption = "\\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \\citep{mcginnis_2014}.",
@@
 # Discussion
+
+## Summary of main findings
 
@@
 Whether the combination of a stable environment and governance structures are additive or multiplicative represents an interesting area for future research, especially under a changing climate.
+
+## Limitations
+
+## Conclusions
+
+# Thoughts from GainesLab Meeting
+
+## General thoughts
+
+A nice story to tell is that reserves need to have optimal social and environmental conditions to work.
+But they also need to be opttimally designed, accounting for size and location.
+It is worth mentioning that even when these reserves did not have nice effects on the ecological or fisheries side, they were a project that brought together the community, and also gave fishers access to a network of resources facilitated by the NGO.
+I still need to get data that support these.
+
+Creo que hay que agregar informacion de la variabilidad ambienta. Posiblemente SST o algo similar. Ver si se puede obtener datos de oxigeno disuelto... Tal vez algo de ENSO / ONI / NAO o algo asi
+
+Quitar analisis de riqueza (la metodologia es restrictiva) y diversidad (aunque shannon provee une buena mezcla del cambio en abundancias y el cambio en riquezas)
+
+revisar biomasa!
+
+Revisar poisson o binomial negativa
+
+Mapa de El Manchon
+
+Las graficas y el analisis de Profit y Catches tienen que ser de langosta entera fresca solamente, y revisar que no haya habido un cambio de commodity a traves del tiempo para justificar usar solamente entera fresca...
+Tratar de obtener valores hasta 2017
+
+Pedir datos biologicos hasta 2017
+
+Distinguis profit / revenue / income
+
+## Comentarios del Lab
+
+Owen
+
+Are there confounding effects because reserves are so close to control sites?
+
+Dan
+
+Are the data good enough to answer these questions?
+
+Karly
+
+Think about other social indicators that do not depend on catches or revenue from these
+
+Alexa
+
+Daniel
+
+Sebas
+
+Look at other indicators of success (adaptive capacity) in the social / governance data
+
+Steve
+
+Would we expect to see effects in the first five years? (ME and PH)
+
+Are the designs good enough to see effects?
+
+Becky
+
+Ignacia
+
+Perhaps interesting follow-up projects include comparing perceptions vs. reality
+
+Estimating the benefits of being friends with an NGO
 

### 2018-03-20 docs/paper/Villasenor-Derbez.tex
Added manuscript to its own folder

--- 2018-02-21 Frontiers_Manuscript/Villasenor-Derbez.tex
@@
 
+\subsection{Summary of main findings}\label{summary-of-main-findings}
+
 Our results indicated idiosyncratic biological effects of the reserves
@@
 
+\subsection{Limitations}\label{limitations}
+
+\subsection{Conclusions}\label{conclusions}
+
+\section{Thoughts from GainesLab
+Meeting}\label{thoughts-from-gaineslab-meeting}
+
+\subsection{General thoughts}\label{general-thoughts}
+
+A nice story to tell is that reserves need to have optimal social and
+environmental conditions to work. But they also need to be opttimally
+designed, accounting for size and location. It is worth mentioning that
+even when these reserves did not have nice effects on the ecological or
+fisheries side, they were a project that brought together the community,
+and also gave fishers access to a network of resources facilitated by
+the NGO. I still need to get data that support these.
+
+Creo que hay que agregar informacion de la variabilidad ambienta.
+Posiblemente SST o algo similar. Ver si se puede obtener datos de
+oxigeno disuelto\ldots{} Tal vez algo de ENSO / ONI / NAO o algo asi
+
+Quitar analisis de riqueza (la metodologia es restrictiva) y diversidad
+(aunque shannon provee une buena mezcla del cambio en abundancias y el
+cambio en riquezas)
+
+revisar biomasa!
+
+Revisar poisson o binomial negativa
+
+Mapa de El Manchon
+
+Las graficas y el analisis de Profit y Catches tienen que ser de
+langosta entera fresca solamente, y revisar que no haya habido un cambio
+de commodity a traves del tiempo para justificar usar solamente entera
+fresca\ldots{} Tratar de obtener valores hasta 2017
+
+Pedir datos biologicos hasta 2017
+
+Distinguis profit / revenue / income
+
+\subsection{Comentarios del Lab}\label{comentarios-del-lab}
+
+Owen
+
+Are there confounding effects because reserves are so close to control
+sites?
+
+Dan
+
+Are the data good enough to answer these questions?
+
+Karly
+
+Think about other social indicators that do not depend on catches or
+revenue from these
+
+Alexa
+
+Daniel
+
+Sebas
+
+Look at other indicators of success (adaptive capacity) in the social /
+governance data
+
+Steve
+
+Would we expect to see effects in the first five years? (ME and PH)
+
+Are the designs good enough to see effects?
+
+Becky
+
+Ignacia
+
+Perhaps interesting follow-up projects include comparing perceptions
+vs.~reality
+
+Estimating the benefits of being friends with an NGO
+
 \section*{Conflict of Interest Statement}

### 2018-03-20 docs/technical_report/Report.Rmd
Moved technical report files to its own folder

--- 2018-01-27 TechnicalReport/Manuscript.Rmd
@@
 
-xy1 <- data.frame(X = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087), Y = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957))
+xy1 <- data.frame(X = c(-115.24, -87.5129043, -87.4611756, -111.377856,-112.7294087),
+                  Y = c(27.85, 19.3722361, 19.3183386, 27.979079, 29.8904957))
 coordinates(xy1) <- c("X", "Y")
@@
 
-load(file = "../Data/coastline_mx.Rda")
+load(file = here::here("data", "coastline_mx.Rda"))
 
@@
 
-sam <- tm_shape(baja2, bbox = bb(xlim = c(-88, -86.726716),
+sam <- tm_shape(baja2, bbox = bb(xlim = c(-88, -86.8),
                                  ylim = c(19.25, 19.5))) +
@@
   tm_shape(res1) +
-  tm_markers(shape = 21, col = "red", border.col = "black", size = 0.5) +
+  tm_markers() +
   tm_compass() +
@@
 ```{r}
-read.csv("../Data/governance_summary.csv") %>% 
+read.csv(here::here("data", "governance_summary.csv")) %>% 
   knitr::kable(booktabs = T, col.names = c("Comunidad", "Reservas", "Año Implementación", "Año Reconocimiento", "Tipo", "Manejo de recursos", "Grado de pesca ilegal", "Inspección y Vigilancia", "Organizaciones pesqueras", "Edad"), caption = "Características de gobernanza generales de las cinco comunidades estudiadas.") %>%
@@
 ```{r}
-fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F)
+fish <- read.csv(here::here("data","fish_data.csv"), strip.white = T, stringsAsFactors = F)
 
-invert <- read.csv("../Data/invert_data.csv", strip.white = T, stringsAsFactors = F)
+invert <- read.csv(here::here("data","invert_data.csv"), strip.white = T, stringsAsFactors = F)
 
-load("../Data/conapesca_2000-2015.RData")
+load(here::here("data", "conapesca_2000-2015.RData"))
 
-cpi <- read.csv("../Data/CPI.csv", stringsAsFactors = F, strip.white = T)
+cpi <- read.csv(here::here("data", "CPI.csv"), stringsAsFactors = F, strip.white = T)
 
@@
 
-gob <- read.csv("../Data/governance.csv", strip.white = T, stringsAsFactors = T)
+gob <- read.csv(here::here("data", "governance.csv"), strip.white = T, stringsAsFactors = T)
 ```
@@
 El tamaño del círculo indica el número de transectos totales realizados por año."}
-esfuerzo_peces <- filter(fish, Comunidad %in% c("Isla Natividad", "Puerto Libertad", "Guaymas", "Maria Elena", "Punta Herrero")) %>% 
+esfuerzo_peces <- fish %>% 
+  filter(Comunidad %in% c("Isla Natividad", "Puerto Libertad", "Guaymas", "Maria Elena", "Punta Herrero")) %>% 
   group_by(Comunidad, Ano, Zona, Transecto) %>% 
@@
   summarize(Transectos = n()) %>% 
+  ungroup() %>% 
   mutate(Class = "Peces")
@@
   summarize(Transectos = n()) %>% 
+  ungroup() %>% 
   mutate(Class = "Invertebrados") %>% 
@@
 ```{r}
-source(file = "analisis.R")
-source(file = "modelo.R")
-source(file = "robust_se.R")
-source(file = "extract_effect.R")
+source(file = here::here("scripts", "analisis.R"))
+source(file = here::here("scripts", "modelo.R"))
+source(file = here::here("scripts", "robust_se.R"))
+source(file = here::here("scripts", "extract_effect.R"))
 

### 2018-03-20 TechnicalReport/robust_se.R
Added all functions and scripts to the scripts folder

@@ -1,8 +1,0 @@
-robust_se <- function(model){
-  
-  RSE <- model %>%
-    lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC1")) %>%
-    broom::tidy()
-  
-  return(RSE)
-}

### 2018-03-20 slides/GainesLabPresentation.Rmd
Added new paths to the presentation so that it renders

--- 2018-02-27 GainesLabPresentation.Rmd
@@
 # Load high-res mexican coastline
-coastline_mx <- readRDS("./Data/Spatial/coastline_mx.rds")
+coastline_mx <- readRDS(here::here("data", "spatial","coastline_mx.rds"))
 
@@
 
-zrp <- st_read(dsn = "./Data/Spatial", layer = "ZRP", quiet = T) %>% 
+zrp <- st_read(dsn = here::here("data", "spatial"), layer = "ZRP", quiet = T) %>% 
   sf::st_centroid() %>% 
@@
 
-rescom <- st_read(dsn = "./Data/Spatial", layer = "Res_Com", quiet = T) %>% 
+rescom <- st_read(dsn = here::here("data", "spatial"), layer = "Res_Com", quiet = T) %>% 
   sf::st_centroid() %>% 
@@
 
-![](./Img/Imagen5.jpg)
+![](./img/Imagen5.jpg)
 
@@
 
-![](./Img/plot.png)
+![](./img/plot.png)
 
@@
 .pull-right[
-![](./Img/plot.png)
+![](./img/plot.png)
 ]
@@
 
-![](./Img/plot_summary.png)
+![](./img/plot_summary.png)
 
@@
 
-![](./Img/Lobsters.png)
+![](./img/Lobsters.png)
 
@@
 
-load("./Data/conapesca_2000-2015.RData")
+load(here::here("data","conapesca_2000-2015.RData"))
 
-cpi <- read.csv("./Data/CPI.csv", stringsAsFactors = F, strip.white = T)
+cpi <- read.csv(here::here("data","CPI.csv"), stringsAsFactors = F, strip.white = T)
 
@@
 ---
-background-image: url(https://raw.githubusercontent.com/jcvdav/ReserveEffect/master/Img/Imagen5.jpg)
+background-image: url(https://raw.githubusercontent.com/jcvdav/ReserveEffect/master/img/Imagen5.jpg)
 class: inverse, center, middle

### 2018-03-21 docs/paper/Villasenor-Derbez.Rmd
Updated title and a couple of graphs

@@
 ---
-title: "Enabling conditions for effective community based marine reserves in small scale fisheries"
+title: "Enabling conditions for effective community-based marine reserves in small-scale fisheries"
 rtitle: "Marine reserve effectiveness"
 author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Álvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
-institution: "
-$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
-$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico
-"
+institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
+$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico"
 corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
@@
 Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making."
-keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico, Social--Ecological Systems"
+keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, Mexico, Social-Ecological Systems"
 output:
@@
 # Determine a projection for the maps
-proj <- CRS("+proj=longlat +datum=NAD27")
+proj <- "+proj=longlat +datum=NAD27"
 
@@
 ```{r model_fit}
-# Fish
-
-## Abundance
-N_IN <- fish_complete %>% 
-  abundance(com = "Isla Natividad") %>% 
-  model(sfe = F, family = "quasipoisson")
-
-N_ME <- fish_complete %>% 
-  abundance(com = "Maria Elena") %>% 
-  model(sfe = F, family = "quasipoisson")
-
-N_PH <- fish_complete %>% 
-  abundance(com = "Punta Herrero") %>% 
-  model(sfe = F, family = "quasipoisson")
-
-## Richness
-S_IN <- fish_complete %>% 
-  richness(com = "Isla Natividad") %>% 
-  model(sfe = F, spfe = F, family = "quasipoisson")
-
-S_ME <- fish_complete %>% 
-  richness(com = "Maria Elena") %>% 
-  model(sfe = F, spfe = F, family = "quasipoisson")
-
-S_PH <- fish_complete %>% 
-  richness(com = "Punta Herrero") %>% 
-  model(sfe = F, spfe = F, family = "quasipoisson")
-
-## Shannon
-H_IN <- fish_complete %>% 
-  shannon(com = "Isla Natividad") %>% 
-  model(sfe = F, spfe = F)
-
-H_ME <- fish_complete %>% 
-  shannon(com = "Maria Elena") %>% 
-  model(sfe = F, spfe = F)
-
-H_PH <- fish_complete %>% 
-  shannon(com = "Punta Herrero") %>% 
-  model(sfe = F, spfe = F)
-
-## Biomass
-B_IN <- fish %>% 
-  biomass(com = "Isla Natividad") %>% 
-  model(sfe = F)
-
-B_ME <- fish%>% 
-  biomass(com = "Maria Elena") %>% 
-  model(sfe = F)
-
-B_PH <- fish %>% 
-  biomass(com = "Punta Herrero") %>% 
-  model(sfe = F)
 
 # Invertebrates
-
-## Abundance
-Ni_IN <- invert_complete %>% 
-  abundance(com = "Isla Natividad") %>% 
-  model(sfe = F, family = "quasipoisson")
-
-Ni_ME <- invert_complete %>% 
-  abundance(com = "Maria Elena") %>% 
-  model(sfe = F, family = "quasipoisson")
-
-Ni_PH <- invert_complete %>% 
-  abundance(com = "Punta Herrero") %>% 
-  model(sfe = F, family = "quasipoisson")
-
-## Richness
-Si_IN <- invert_complete %>% 
-  richness(com = "Isla Natividad") %>% 
-  model(sfe = F, spfe = F, family = "quasipoisson")
-
-Si_ME <- invert_complete %>% 
-  richness(com = "Maria Elena") %>% 
-  model(sfe = F, spfe = F, family = "quasipoisson")
-
-Si_PH <- invert_complete %>% 
-  richness(com = "Punta Herrero") %>% 
-  model(sfe = F, spfe = F, family = "quasipoisson")
-
-## Shannon
-Hi_IN <- invert_complete %>% 
-  shannon(com = "Isla Natividad") %>% 
-  model(sfe = F, spfe = F)
-
-Hi_ME <- invert_complete %>% 
-  shannon(com = "Maria Elena") %>% 
-  model(sfe = F, spfe = F)
-
-Hi_PH <- invert_complete %>% 
-  shannon(com = "Punta Herrero") %>% 
-  model(sfe = F, spfe = F)
 
@@
   model(sfe = F, spfe = F, family = "quasipoisson")
+
+## Abundance
+Ni_IN <- invert_complete %>% 
+  abundance(com = "Isla Natividad") %>% 
+  model(sfe = F, family = "quasipoisson")
+
+Ni_ME <- invert_complete %>% 
+  abundance(com = "Maria Elena") %>% 
+  model(sfe = F, family = "quasipoisson")
+
+Ni_PH <- invert_complete %>% 
+  abundance(com = "Punta Herrero") %>% 
+  model(sfe = F, family = "quasipoisson")
+
+# Fish
+
+## Biomass
+B_IN <- fish %>% 
+  biomass(com = "Isla Natividad") %>% 
+  model(sfe = F)
+
+B_ME <- fish%>% 
+  biomass(com = "Maria Elena") %>% 
+  model(sfe = F)
+
+B_PH <- fish %>% 
+  biomass(com = "Punta Herrero") %>% 
+  model(sfe = F)
+
+## Abundance
+N_IN <- fish_complete %>% 
+  abundance(com = "Isla Natividad") %>% 
+  model(sfe = F, family = "quasipoisson")
+
+N_ME <- fish_complete %>% 
+  abundance(com = "Maria Elena") %>% 
+  model(sfe = F, family = "quasipoisson")
+
+N_PH <- fish_complete %>% 
+  abundance(com = "Punta Herrero") %>% 
+  model(sfe = F, family = "quasipoisson")
+
 ```
@@
 
-```{r, fig.width = 6.5, fig.height = 8, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for community-levell indicators.
+```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for community-levell indicators.
 Plots are ordered by survey type (left: fish; right: invertebrates) and indicators: Abundance (A, B), Richness (C, D), Shannon's diversity index (E, F),fish biomass (G), and lobster (*Panulirus spp*) abundances (H).
@@
 
-# Fish plots
-F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
+# Invert plots
+## Lobsters
+L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
                      various = T,
-                  model_names = c("IN", "ME", "PH"), legend = T)  +
+                     model_names = c("IS", "ME", "PH"))  +
   theme(legend.justification = c(0, 0),
-        legend.position = c(0, 0),
+        legend.position = c(0, 0.4),
         legend.text = element_text(size = 8),
@@
         legend.direction = "horizontal") +
-  scale_colour_brewer(guide = guide_legend(title.position = "top", nrow = 1), palette = "Set1")
+  scale_colour_brewer(guide = guide_legend(title.position = "top", nrow = 3), palette = "Set1")
 
-F_S <- plot_model(model = list(S_IN, S_ME, S_PH),
+## Abundance
+I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
                   various = T,
@@
 
-F_H <- plot_model(model = list(H_IN, H_ME, H_PH),
-                  various = T,
-                  model_names = c("IN", "ME", "PH"))
-
+# Fish plots
+## Biomass
 F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
@@
 
-# Invert plots
-I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
+## Abundance
+F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
                   various = T,
-                  model_names = c("IN", "ME", "PH"))
+                  model_names = c("IN", "ME", "PH"), legend = F)
 
-I_S <- plot_model(model = list(Si_IN, Si_ME, Si_PH),
-                  various = T,
-                  model_names = c("IN", "ME", "PH"))
-
-I_H <- plot_model(model = list(Hi_IN, Hi_ME, Hi_PH),
-                  various = T,
-                  model_names = c("IN", "ME", "PH"))
-
-## Lobsters
-L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
-                     various = T,
-                     model_names = c("IS", "ME", "PH"))
-
-cowplot::plot_grid(plotlist = list(F_N, I_N, F_S, I_S, F_H, I_H, F_B, L_plot),
+cowplot::plot_grid(plotlist = list(L_plot, F_B, I_N, F_N),
                    labels = "AUTO",
                    ncol = 2)
+
 ```
@@
 
-IN <- list(N_IN, Ni_IN, S_IN, Si_IN, H_IN, Hi_IN, B_IN, L_IN) %>% map_df(.f = observe) %>%
+IN <- list(N_IN, Ni_IN, B_IN, L_IN) %>% map_df(.f = observe) %>%
   as.tibble() %>% 
@@
 
-ME <- list(N_ME, Ni_ME, S_ME, Si_ME, H_ME, Hi_ME, B_ME, L_ME) %>% map_df(.f = observe) %>%
+ME <- list(N_ME, Ni_ME, B_ME, L_ME) %>% map_df(.f = observe) %>%
   as.tibble() %>% 
@@
 
-PH <- list(N_PH, Ni_PH, S_PH, Si_PH, H_PH, Hi_PH, B_PH, L_PH) %>% map_df(.f = observe) %>%
+PH <- list(N_PH, Ni_PH, B_PH, L_PH) %>% map_df(.f = observe) %>%
   as.tibble() %>% 
@@
   labs(x = "Community", y = "Percentage of total coefficients") +
-  scale_y_continuous(breaks = c(-0.5, -0.25, 0, 0.25, 0.5, 0.75), limits = c(-0.5, 0.75),
+  scale_y_continuous(breaks = seq(-0.65, 0.9, by = 0.25), limits = c(-0.65, 0.9),
                      labels = scales::percent) +

### 2018-03-21 docs/paper/Villasenor-Derbez.tex
Updated title and a couple of graphs

@@
 
-\title[Marine reserve effectiveness]{Enabling conditions for effective community based marine reserves in
-small scale fisheries}
+\title[Marine reserve effectiveness]{Enabling conditions for effective community-based marine reserves in
+small-scale fisheries}
 
@@
 \tiny
- \keyFont{ \section{Keywords:} Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen
-Science, Mexico, Social--Ecological Systems}
+ \keyFont{ \section{Keywords:} Marine Protected Areas, Marine Conservation, Small-Scale Fisheries,
+Citizen Science, Mexico, Social-Ecological Systems}
 

### 2018-03-22 Scripts/abundance.R
Modifed content in scripts

@@ -1,14 +1,16 @@
 abundance <- function(x, com, spp = NULL){
   
   if(!is.null(spp)){
-    x %<>% filter(GeneroEspecie == spp)
+    x %<>% filter(generoespecie == spp)
   }
   
   results <- x %>% 
-    filter(Comunidad == com) %>%
-    mutate(Ano = as.factor(Ano)) %>% 
-    group_by(Comunidad, RC, Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
-    summarize(Indicador = sum(Abundancia))
+    filter(comunidad == com) %>%
+    mutate(ano = as.factor(ano),
+           year = as.factor(year)) %>% 
+    group_by(rc, ano, year, sitio, zona, transecto, generoespecie) %>% 
+    summarize(indicador = sum(abundancia)) %>% 
+    ungroup()
     
     return(results)
 }

### 2018-03-22 Scripts/biomass.R

@@ -1,12 +1,12 @@
 biomass <- function(x, com){
   results <- x %>% 
-    filter(Comunidad == com) %>% 
-    filter(Abundancia > 0) %>% 
-    left_join(species_bio, by = c("GeneroEspecie")) %>% 
-    mutate(Ano = as.factor(Ano),
-           biomass = Abundancia * a * (Talla^b)/1000) %>% 
-    group_by(RC, Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
-    summarize(Indicador = sum(biomass, na.rm = T))
+    filter(comunidad == com) %>% 
+    mutate(ano = as.factor(ano),
+           year = as.factor(year),
+           biomass = abundancia * a * (talla^b)/1000) %>% 
+    group_by(rc, ano, year, sitio, zona, transecto, generoespecie) %>% 
+    summarize(indicador = sum(biomass, na.rm = T)) %>% 
+    ungroup()
   
   return(results)
 }

### 2018-03-22 Scripts/model.R

@@ -3,13 +3,13 @@ model <- function(x, family = "gaussian", sfe = T, spfe = T){
   # x %<>% filter(Comunidad == com)
   
   if(sfe){
-    if(spfe){model <- glm(Indicador ~ Ano * Zona + Sitio + GeneroEspecie, data = x, family = family)} #Site FE and Spp FE
-    if(!spfe){model <- glm(Indicador ~ Ano * Zona + Sitio, data = x, family = family)} #Site FE, no Spp FE
+    if(spfe){model <- glm(indicador ~ year * zona + sitio + generoespecie -1, data = x, family = family)} #Site FE and Spp FE
+    if(!spfe){model <- glm(indicador ~ year * zona + sitio -1, data = x, family = family)} #Site FE, no Spp FE
   }
   
   if(!sfe){
-    if(spfe){model <- glm(Indicador ~ Ano * Zona + GeneroEspecie, data = x, family = family)} #No site FE and Spp FE
-    if(!spfe){model <- glm(Indicador ~ Ano * Zona, data = x, family = family)} #No site and no Spp FE
+    if(spfe){model <- glm(indicador ~ year * zona + generoespecie -1, data = x, family = family)} #No site FE and Spp FE
+    if(!spfe){model <- glm(indicador ~ year * zona, data = x, family = family)} #No site and no Spp FE
   }
   
   return(model)

### 2018-03-22 Scripts/plot_model.R

@@ -15,16 +15,16 @@ plot_model <- function(model, various = F, model_names = NULL, legend = F){
     for(i in 1:length(model)){
       tidy_terms <- model[[i]] %>% 
         robust_se() %>%
-        filter(grepl(":ZonaReserva", term)) %$%
+        filter(grepl(":zonaReserva", term)) %$%
         str_split(term, pattern = ":", simplify = T) %>% 
         as.tibble() %>% 
-        set_colnames(c("Ano", "Zona")) %>% 
-        select(-Zona) %>% 
-        mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = "")),
+        set_colnames(c("ano", "zona")) %>% 
+        select(-zona) %>% 
+        mutate(Ano = as.numeric(gsub(ano, pattern = "year", replacement = "")),
                Community = model_names[i])
       
       p_i <- robust_se(model[[i]]) %>%
-        filter(grepl(":ZonaReserva", term)) %>%
+        filter(grepl(":zonaReserva", term)) %>%
         cbind(tidy_terms) %>%
         mutate(p = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05"))
       
@@ -40,8 +40,9 @@ plot_model <- function(model, various = F, model_names = NULL, legend = F){
       geom_hline(yintercept = 0, linetype = "dashed") +
       scale_fill_brewer(palette = "Set1") +
       theme(text = element_text(size = 12)) +
-      labs(x = "Year", y = "Effect size") +
-      scale_color_brewer(palette = "Set1")
+      labs(x = "Years since implementation", y = quo(lambda[t])) +
+      scale_color_brewer(palette = "Set1") +
+      scale_x_continuous(breaks = seq(0, 10, by = 1))
     
   }
   

### 2018-03-22 Scripts/richness.R

@@ -1,10 +1,0 @@
-richness <- function(x, com){
-  results <- x %>% 
-    filter(Comunidad == com) %>%
-    filter(Abundancia > 0) %>%
-    mutate(Ano = as.factor(Ano)) %>% 
-    group_by(RC, Ano, Sitio, Zona, Transecto) %>% 
-    summarize(Indicador = n())
-  
-  return(results)
-}

### 2018-03-22 Scripts/robust_se.R

@@ -1,7 +1,7 @@
 robust_se <- function(model){
   
   RSE <- model %>%
-    lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC1")) %>%
+    lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC2")) %>%
     broom::tidy()
   
   return(RSE)

### 2018-03-22 Scripts/shannon.R

@@ -1,38 +1,0 @@
-shannon <- function(x, com){
-  N <- x %>%
-    filter(Comunidad == com) %>%
-    filter(Abundancia > 0) %>%
-    mutate(Ano = as.factor(Ano)) %>%
-    group_by(Ano,
-             Zona,
-             Sitio,
-             Transecto) %>%
-    summarize(N = sum(Abundancia)) %>%
-    ungroup() %>%
-    mutate(ID = paste(Ano, Zona, Sitio, Transecto)) %>%
-    select(ID, N)
-
-  H <- x %>%
-    filter(Comunidad == com) %>%
-    filter(Abundancia > 0) %>%
-    mutate(Ano = as.factor(Ano)) %>%
-    group_by(Ano,
-             Zona,
-             Sitio,
-             Transecto,
-             GeneroEspecie) %>%
-    summarize(ni = sum(Abundancia)) %>%
-    ungroup() %>%
-    mutate(ID = paste(Ano, Zona, Sitio, Transecto)) %>%
-    left_join(N, by = "ID") %>%
-    mutate(pi = ni/N) %>%
-    group_by(Ano,
-             Zona,
-             Sitio,
-             Transecto) %>%
-    summarize(H = -1*sum(pi*log2(pi))) %>%
-    ungroup() %>%
-    select(Ano, Zona, Sitio, Transecto, Indicador = H)
-
-  return(as.data.frame(H))
-}

### 2018-03-22 docs/data_cleaning_logbook.Rmd
deleted old test html reg table

---
title: "Data Cleaning logbook"
author: "Juan Carlos Villaseñor-Derbez"
date: "last compilation: `r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(rfishbase)
  library(tidyverse)
})
```

# Ecological data

## Invertebrates

```{r}
invert <- read.csv(here::here("raw_data", "invert_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(abundancia > 0) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013)) %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, abundancia) %>% 
  arrange()
```

## Fish

### Species biology

```{r}
species_bio <- MPAtools::species_bio

needed_spp <- c("Kyphosus incisor",
                "Kyphosus sectatrix",
                "Pterois volitans",
                "Scarus guacamaia",
                "Chaetodon capistratus",
                "Chaetodon capistratus",
                "Chaetodon striatus",
                "Calamus pennatula")

FB_update <- rfishbase::length_weight(needed_spp, fields = c("a", "b")) %>% 
  select(GeneroEspecie = sciname, a, b) %>% 
  mutate(GeneroEspecie = case_when(GeneroEspecie %in% c("Kyphosus sectatrix", "Kyphosus incisor") ~ "Kyphosus",
                                   GeneroEspecie %in% c("Chaetodon capistratus", "Chaetodon striatus") ~ "Chaetodon",
                                   TRUE ~ GeneroEspecie)) %>% 
  group_by(GeneroEspecie) %>% 
  summarize_all(mean, na.rm = T) %>% 
  ungroup()

needed_genus <- c("Mycteroperca",
                  "Chromis",
                  "Stegastes",
                  "Haemulon",
                  "Acanthurus",
                  "Sebastes",
                  "Lutjanus",
                  "Scarus")

GenusLevelBio <- species_bio %>% 
  filter(Genus %in% needed_genus) %>%
  select(GeneroEspecie = Genus, a, b) %>% 
  group_by(GeneroEspecie) %>% 
  summarize_all(mean, na.rm = T) %>% 
  ungroup()

species_bio %<>%
  select(GeneroEspecie, a, b) %>% 
  rbind(FB_update) %>% 
  rbind(GenusLevelBio) %>% 
  janitor::clean_names()
```

### Clean data

```{r}
fish <- read.csv(here::here("raw_data", "fish_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
  janitor::clean_names() %>% 
  filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(rc),
         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
                   "Cabezo - Cabezo (Control)",
                   "Manchon - Manchon (Control)",
                   "El Faro - El Faro (Control)")) %>% 
  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
  filter(abundancia > 0) %>% 
  mutate(generoespecie = case_when(generoespecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
                                   generoespecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
                                   generoespecie == "Xcochin spp" ~ "Balistes capriscus",
                                   generoespecie == "Pterois volitans/miles" ~ "Pterois volitans",
                                   generoespecie == "Mycteroperca spp" ~ "Mycteroperca",
                                   generoespecie == "Chromis spp" ~ "Chromis",
                                   generoespecie == "Damisela spp" ~ "Stegastes",
                                   generoespecie == "Haemulon spp" ~ "Haemulon",
                                   generoespecie == "Lancero spp" ~ "Acanthurus",
                                   generoespecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
                                   generoespecie == "Sebastes spp" ~ "Sebastes",
                                   generoespecie == "Loro spp" ~ "Scarus",
                                   generoespecie == "Lutjanus spp" ~ "Lutjanus",
                                   generoespecie == "Mariposa spp" ~ "Chaetodon",
                                   generoespecie == "Pluma spp" ~ "Calamus pennatula",
                                   generoespecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
                                   generoespecie == "Vieja spp" ~ "Bodianus rufus",
                                   TRUE ~ generoespecie)) %>% 
  left_join(species_bio, by = c("generoespecie")) %>% 
  group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie, talla, a,b) %>% 
  summarize(abundancia = sum(abundancia)) %>% 
  ungroup() %>% 
  mutate(year = case_when(comunidad == "Isla Natividad" ~ ano - 2006,
                         comunidad == "Maria Elena" ~ ano - 2012,
                         TRUE ~ ano - 2013)) %>% 
  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 
  arrange()
```

# Socioeconomic data

## CPI

```{r}
cpi <- read.csv(here::here("raw_data", "CPI.csv"), stringsAsFactors = F, strip.white = T)

cpi_2014 <- cpi$CPI[cpi$Ano == 2014]
```

## Landings and revenue

```{r}
conapesca <- readRDS(here::here("raw_data","conapesca.rds")) %>% 
  filter(UnidadEconomica %in% c("Scpp cozumel scl",
                                "Scpp jose maria azcorra scl",
                                "Scpp buzos y pescadores de la baja california scl")) %>% 
  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
  group_by(UnidadEconomica, Ano, NombreComun) %>%
  summarize(PesoDesembarcado = sum(PesoDesembarcado), 
            PesoVivo = sum(PesoVivo),
            Valor = sum(Valor)) %>%
  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
  left_join(cpi, by = "Ano") %>% 
  mutate(Valor = Valor * (CPI / cpi_2014))
```

# Export all data

```{r}
write.csv(x = invert, file = here::here("data", "invertebrates.csv"), row.names = F)

write.csv(x = fish, file = here::here("data", "fish.csv"), row.names = F)

write.csv(x = conapesca, file = here::here("data", "conapesca.csv"))

```

### 2018-03-22 docs/paper/Villasenor-Derbez.Rmd
deleted old test html reg table

@@
 ```{r load_data}
-fish <- read.csv(here::here("data", "fish_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
-  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
-  filter(!is.na(RC),
-         RC %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
-                   "Cabezo - Cabezo (Control)",
-                   "Manchon - Manchon (Control)",
-                   "El Faro - El Faro (Control)")) %>% 
-  filter(!(RC %in% c("El Faro - El Faro (Control)") & Ano == 2012)) %>% 
-  mutate(GeneroEspecie = case_when(
-    GeneroEspecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
-    GeneroEspecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
-    GeneroEspecie == "Xcochin spp" ~ "Balistes capriscus",
-    GeneroEspecie == "Pterois volitans/miles" ~ "Pterois volitans",
-    GeneroEspecie == "Mycteroperca spp" ~ "Mycteroperca",
-    GeneroEspecie == "Chromis spp" ~ "Chromis",
-    GeneroEspecie == "Damisela spp" ~ "Stegastes",
-    GeneroEspecie == "Haemulon spp" ~ "Haemulon",
-    GeneroEspecie == "Lancero spp" ~ "Acanthurus",
-    GeneroEspecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
-    GeneroEspecie == "Sebastes spp" ~ "Sebastes",
-    GeneroEspecie == "Loro spp" ~ "Scarus",
-    GeneroEspecie == "Lutjanus spp" ~ "Lutjanus",
-    GeneroEspecie == "Mariposa spp" ~ "Chaetodon",
-    GeneroEspecie == "Pluma spp" ~ "Calamus pennatula",
-    GeneroEspecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
-    GeneroEspecie == "Vieja spp" ~ "Bodianus rufus",
-    TRUE ~ GeneroEspecie
-  ))
+invert <- read.csv(file = here::here("data", "invertebrates.csv"))
 
-invert <- read.csv(here::here("data", "invert_data.csv"), strip.white = T, stringsAsFactors = F) %>% 
-  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
-  filter(!is.na(RC),
-         RC %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
-                   "Cabezo - Cabezo (Control)",
-                   "Manchon - Manchon (Control)",
-                   "El Faro - El Faro (Control)")) %>% 
-  filter(!(RC %in% c("El Faro - El Faro (Control)") & Ano == 2012))
+fish <- read.csv(file = here::here("data", "fish.csv"))
 
-load(here::here("data","conapesca_2000-2015.RData"))
-
-cpi <- read.csv(here::here("data", "CPI.csv"), stringsAsFactors = F, strip.white = T)
-
-cpi_2014 <- cpi$CPI[cpi$Ano == 2014]
-
-conapesca %<>%
-  filter(UnidadEconomica %in% c("Scpp cozumel scl",
-                                "Scpp jose maria azcorra scl",
-                                "Scpp buzos y pescadores de la baja california scl")) %>% 
-  filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
-  group_by(UnidadEconomica, Ano, NombreComun) %>%
-  summarize(PesoDesembarcado = sum(PesoDesembarcado), 
-            PesoVivo = sum(PesoVivo),
-            Valor = sum(Valor)) %>%
-  filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
-  left_join(cpi, by = "Ano") %>% 
-  mutate(Valor = Valor * (CPI / cpi_2014))
+conapesca <- read.csv(x = conapesca, file = here::here("data", "conapesca.csv"))
 
@@
 ```{r}
-inv_summary <- group_by(invert, Comunidad, RC, Zona, Sitio, Transecto, Ano) %>%
+inv_summary <- group_by(invert, comunidad, rc, zona, sitio, transecto, ano) %>%
   count() %>% 
   ungroup() %>% 
-  group_by(Comunidad, RC, Zona) %>% 
+  group_by(comunidad, rc, zona) %>% 
   summarize(n = n()) %>% 
   ungroup() %>% 
-  spread(Zona, n) %>% 
-  select(Comunidad, RC, Control, Reserve = Reserva)
+  spread(zona, n) %>% 
+  select(comunidad, rc, control = Control, reserve = Reserva)
 
-group_by(fish, Comunidad, RC, Zona, Sitio, Transecto, Ano) %>%
+group_by(fish, comunidad, rc, zona, sitio, transecto, ano) %>%
   count() %>% 
   ungroup() %>% 
-  group_by(Comunidad, RC, Zona) %>% 
-  summarize(Start = min(Ano), n = n()) %>% 
+  group_by(comunidad, rc, zona) %>% 
+  summarize(ano = min(ano), n = n()) %>% 
   ungroup() %>% 
-  spread(Zona, n) %>% 
-  mutate(Implementation = c(2006, 2012, 2013, 2013)) %>% 
-  select(Comunidad, RC, Implementation, Start, Control, Reserve = Reserva) %>% 
-  left_join(inv_summary, by = c("Comunidad", "RC")) %>% 
+  spread(zona, n) %>% 
+  mutate(implementation = c(2006, 2012, 2013, 2013)) %>% 
+  select(comunidad, implementation, ano, rc, control = Control, reserve = Reserva) %>% 
+  left_join(inv_summary, by = c("comunidad", "rc")) %>% 
   knitr::kable(format = "latex",
@@
 
-\clearpage
-
-```{r, prepare_covars}
-fish_covars <- fish %>% 
-  group_by(RC, Ano, Sitio, Zona, Transecto, Comunidad) %>% 
-  count() %>% 
-  select(-n)
-
-invert_covars <- invert %>% 
-  group_by(RC, Ano, Sitio, Zona, Transecto, Comunidad) %>% 
-  count() %>% 
-  select(-n)
-```
-
-```{r complete_fish}
-fish_complete <- fish %>% 
-  select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
-  group_by(Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>% 
-  ungroup() %>% 
-  complete(Ano, Transecto, nesting(Sitio, Zona, GeneroEspecie), fill = list(Abundancia = 0L)) %>% 
-  left_join(fish_covars, by = c("Ano", "Sitio", "Zona", "Transecto")) %>% 
-  select(Comunidad, Ano, RC, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
-  arrange() %>% 
-  filter(!is.na(RC))
-
-species_bio <- MPAtools::species_bio
-
-spp <- group_by(fish_complete, GeneroEspecie) %>% 
-  count() %>% 
-  left_join(species_bio, by = "GeneroEspecie") %>% 
-  filter(is.na(a) | is.na(b))
-
-FB_update <- rfishbase::length_weight(c("Kyphosus incisor", "Kyphosus sectatrix", "Pterois volitans", "Scarus guacamaia", "Chaetodon capistratus", "Chaetodon capistratus", "Chaetodon striatus", "Calamus pennatula"), fields = c("a", "b")) %>%
-  select(GeneroEspecie = sciname, a, b) %>% 
-  mutate(GeneroEspecie = case_when(
-    GeneroEspecie %in% c("Kyphosus sectatrix", "Kyphosus incisor") ~ "Kyphosus",
-    GeneroEspecie %in% c("Chaetodon capistratus", "Chaetodon striatus") ~ "Chaetodon",
-    TRUE ~ GeneroEspecie
-  )) %>% 
-  group_by(GeneroEspecie) %>% 
-  summarize_all(mean, na.rm = T)
-
-GenusLevelBio <- species_bio %>% 
-  filter(Genus %in% c("Mycteroperca", "Chromis", "Stegastes", "Haemulon", "Acanthurus", "Sebastes", "Lutjanus", "Scarus")) %>%
-  select(GeneroEspecie = Genus, a, b) %>% 
-  group_by(GeneroEspecie) %>% 
-  summarize_all(mean, na.rm = T)
-
-species_bio %<>%
-  select(GeneroEspecie, a, b) %>% 
-  rbind(FB_update) %>% 
-  rbind(GenusLevelBio)
-```
-
-```{r complete_inverts}
-invert_complete <- invert %>%
-  select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
-  group_by(Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
-  summarize(Abundancia = sum(Abundancia)) %>% 
-  ungroup() %>% 
-  complete(Ano, Transecto, nesting(Sitio, Zona, GeneroEspecie), fill = list(Abundancia = 0L)) %>% 
-  left_join(fish_covars, by = c("Ano", "Sitio", "Zona", "Transecto")) %>% 
-  select(Comunidad, Ano, RC, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
-  arrange()
-```
-
 ```{r model_fit}
-
 # Invertebrates
@@
 ## Lobsters
-L_IN <- invert_complete %>%
-  filter(GeneroEspecie == "Panulirus interruptus") %>% 
+L_IN <- invert %>%
+  filter(generoespecie == "Panulirus interruptus") %>% 
   abundance(com = "Isla Natividad") %>% 
-  model(sfe = F, spfe = F, family = "quasipoisson")
+  model(sfe = F, spfe = F)
 
-L_ME <- invert_complete %>% 
-  filter(GeneroEspecie == "Panulirus argus") %>% 
+L_ME <- invert %>% 
+  filter(generoespecie == "Panulirus argus") %>% 
   abundance(com = "Maria Elena") %>% 
-  model(sfe = F, spfe = F, family = "quasipoisson")
+  model(sfe = F, spfe = F)
 
-L_PH <- invert_complete %>% 
-  filter(GeneroEspecie == "Panulirus argus") %>% 
+L_PH <- invert %>% 
+  filter(generoespecie == "Panulirus argus") %>% 
   abundance(com = "Punta Herrero") %>% 
-  model(sfe = F, spfe = F, family = "quasipoisson")
+  model(sfe = F, spfe = F)
 
 ## Abundance
-Ni_IN <- invert_complete %>% 
+Ni_IN <- invert %>% 
   abundance(com = "Isla Natividad") %>% 
-  model(sfe = F, family = "quasipoisson")
+  model(sfe = F)
 
-Ni_ME <- invert_complete %>% 
+Ni_ME <- invert %>% 
   abundance(com = "Maria Elena") %>% 
-  model(sfe = F, family = "quasipoisson")
+  model(sfe = F)
 
-Ni_PH <- invert_complete %>% 
+Ni_PH <- invert %>% 
   abundance(com = "Punta Herrero") %>% 
-  model(sfe = F, family = "quasipoisson")
+  model(sfe = F)
 
@@
 ## Abundance
-N_IN <- fish_complete %>% 
+N_IN <- fish %>% 
   abundance(com = "Isla Natividad") %>% 
-  model(sfe = F, family = "quasipoisson")
+  model(sfe = F)
 
-N_ME <- fish_complete %>% 
+N_ME <- fish %>% 
   abundance(com = "Maria Elena") %>% 
-  model(sfe = F, family = "quasipoisson")
+  model(sfe = F)
 
-N_PH <- fish_complete %>% 
+N_PH <- fish %>% 
   abundance(com = "Punta Herrero") %>% 
-  model(sfe = F, family = "quasipoisson")
+  model(sfe = F)
 

### 2018-03-22 docs/paper/Villasenor-Derbez.Rmd
fixed mistakes due to reorganization of repo

@@
 source(file = here::here("scripts", "abundance.R")) #Calculate abundance per transect
-source(file = here::here("scripts", "richness.R")) # Calculate richness per transect
-source(file = here::here("scripts", "shannon.R")) #Calculate shannon's index per transect
 source(file = here::here("scripts", "biomass.R")) # Calculate fish biomass per transect
@@
 # Load high-res mexican coastline
-coastline_mx <- readRDS(here::here("data", "spatial", "coastline_mx.rds"))
+coastline_mx <- readRDS(here::here("raw_data", "spatial", "coastline_mx.rds"))
 
@@
 # Load ZRP shapefile
-zrp <- st_read(dsn = here::here("data", "spatial"), layer = "ZRP", quiet = T) %>% 
+zrp <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
   filter(Name %in% c("El Cabezo", "El Faro")) %>% 
@@
 # Load Community reserve shapefile
-rescom <- st_read(dsn = here::here("data", "spatial"), layer = "Res_Com", quiet = T) %>% 
+rescom <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
   filter(Nombre == "La_Plana") %>% 
@@
 
-conapesca <- read.csv(x = conapesca, file = here::here("data", "conapesca.csv"))
+conapesca <- read.csv(file = here::here("data", "conapesca.csv"))
 
@@
   mutate(implementation = c(2006, 2012, 2013, 2013)) %>% 
-  select(comunidad, implementation, ano, rc, control = Control, reserve = Reserva) %>% 
+  select(comunidad, rc, implementation, ano, control = Control, reserve = Reserva) %>% 
   left_join(inv_summary, by = c("comunidad", "rc")) %>% 
@@
                    ncol = 2)
-
 ```
@@
 ```{r, fig.height = 3, fig.width = 6, fig.cap = "\\label{fig:res_com}Summarized effects of the marne reserves by direction (positive - negative) and significance."}
-observe <- function(model){      tidy_terms <- model %>% 
+observe <- function(model){
+  tidy_terms <- model %>% 
         robust_se() %>%
-        filter(grepl(":ZonaReserva", term)) %$%
+        filter(grepl(":zonaReserva", term)) %$%
         str_split(term, pattern = ":", simplify = T) %>% 
@@
         select(-Zona) %>% 
-        mutate(Ano = as.numeric(gsub(Ano, pattern = "Ano", replacement = "")))
+        mutate(Ano = as.numeric(gsub(Ano, pattern = "year", replacement = "")))
 
       p_i <- robust_se(model) %>%
-        filter(grepl(":ZonaReserva", term)) %>%
+        filter(grepl(":zonaReserva", term)) %>%
         cbind(tidy_terms) %>%
         mutate(ptest = ((p.value < 0.05)*-1)+0.05)
-       return(p_i)}
 
-IN <- list(N_IN, Ni_IN, B_IN, L_IN) %>% map_df(.f = observe) %>%
+       return(p_i)
+      }
+
+IN <- list(N_IN, Ni_IN, B_IN, L_IN) %>%
+  map_df(.f = observe) %>%
   as.tibble() %>% 
@@
 
-ME <- list(N_ME, Ni_ME, B_ME, L_ME) %>% map_df(.f = observe) %>%
+ME <- list(N_ME, Ni_ME, B_ME, L_ME) %>%
+  map_df(.f = observe) %>%
   as.tibble() %>% 
@@
 
-PH <- list(N_PH, Ni_PH, B_PH, L_PH) %>% map_df(.f = observe) %>%
+PH <- list(N_PH, Ni_PH, B_PH, L_PH) %>%
+  map_df(.f = observe) %>%
   as.tibble() %>% 
@@
   labs(x = "Community", y = "Percentage of total coefficients") +
-  scale_y_continuous(breaks = seq(-0.65, 0.9, by = 0.25), limits = c(-0.65, 0.9),
+  scale_y_continuous(breaks = seq(-0.5, 1, by = 0.25), limits = c(-0.5, 1),
                      labels = scales::percent) +

### 2018-03-22 Scripts/model.R

@@ -9,7 +9,7 @@ model <- function(x, family = "gaussian", sfe = T, spfe = T){
   
   if(!sfe){
     if(spfe){model <- glm(indicador ~ year * zona + generoespecie -1, data = x, family = family)} #No site FE and Spp FE
-    if(!spfe){model <- glm(indicador ~ year * zona, data = x, family = family)} #No site and no Spp FE
+    if(!spfe){model <- glm(indicador ~ year * zona -1, data = x, family = family)} #No site and no Spp FE
   }
   
   return(model)

### 2018-03-22 Scripts/robust_se.R

@@ -1,7 +1,7 @@
 robust_se <- function(model){
   
   RSE <- model %>%
-    lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC2")) %>%
+    lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC1")) %>%
     broom::tidy()
   
   return(RSE)

### 2018-04-03 docs/data_cleaning_logbook.Rmd
updated logbook

@@
   filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
-  filter(abundancia > 0) %>% 
+  filter(between(abundancia, 0, 100)) %>% 
   group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie) %>% 
@@
 ```{r}
+cooperatives_t <- c("Scpp cozumel scl",
+                    "Scpp jose maria azcorra scl",
+                    "Scpp buzos y pescadores de la baja california scl")
+
+cooperatives_c <- c("Scpp langosteros del caribe scl",
+                    "Scpp vigia chico scl",
+                    "Scpp pesc nacionales de abulon sc de rl",
+                    "Scpp la purisima scl",
+                    "Scpp bahia tortugas scl")
+
+cooperatives <- c(cooperatives_t, cooperatives_c)
+```
+
+```{r}
+L_conapesca <- tibble(UnidadEconomica = c("buzos y pescadores de la baja california",
+                                          "bahia tortugas",
+                                          "pesc nacionales de abulon",
+                                          "la purisima",
+                                          "cozumel",
+                                          "jose maria azcorra",
+                                          "vigia chico",
+                                          "langosteros del caribe"),
+                      Comunidad = c("Isla Natividad",
+                                    NA, NA, NA, 
+                                    "Maria Elena",
+                                    "Punta Herrero",
+                                    NA, NA),
+                      Zona = c("PN", "PN", "PN", "PN", "MC", "MC", "MC", "MC"),
+                      Start = c(2006, 2006, 2006, 2006, 2012, 2013, 2012, 2012))
+```
+
+```{r}
 conapesca <- readRDS(here::here("raw_data","conapesca.rds")) %>% 
-  filter(UnidadEconomica %in% c("Scpp cozumel scl",
-                                "Scpp jose maria azcorra scl",
-                                "Scpp buzos y pescadores de la baja california scl")) %>% 
+  mutate(UnidadEconomica = as.character(UnidadEconomica)) %>% 
+  mutate(UnidadEconomica = ifelse(UnidadEconomica == "Scpp bahia tortugas s.c de r.l", 
+                                  "Scpp bahia tortugas scl",
+                                  UnidadEconomica)) %>% 
+  filter(UnidadEconomica %in% cooperatives) %>%
   filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
@@
             Valor = sum(Valor)) %>%
+  ungroup() %>% 
   filter(!(UnidadEconomica == "Scpp buzos y pescadores de la baja california scl" & NombreComun == "Langosta ent. fca.")) %>% 
   left_join(cpi, by = "Ano") %>% 
-  mutate(Valor = Valor * (CPI / cpi_2014))
+  mutate(Valor = Valor * (CPI / cpi_2014),
+         Group = case_when(UnidadEconomica %in% cooperatives_t ~ "Treated",
+                           TRUE ~ "Control"),
+         UnidadEconomica = gsub(pattern = "Scpp ", replacement = "", x = UnidadEconomica),
+         UnidadEconomica = gsub(pattern = " scl", replacement = "", x = UnidadEconomica),
+         UnidadEconomica = gsub(pattern = " s.c de r.l", replacement = "", x = UnidadEconomica),
+         UnidadEconomica = gsub(pattern = " sc de rl", replacement = "", x = UnidadEconomica)) %>% 
+  left_join(L_conapesca, by = "UnidadEconomica") %>% 
+  mutate(Post = ifelse(Ano <= Start, 0, 1),
+         years_centered = Ano - Start) %>% 
+  select(Zona, Grupo = Group, Comunidad, UnidadEconomica, Ano, Post, years_centered, NombreComun, PesoVivo, PesoDesembarcado, Valor, CPI) %>% 
+  janitor::clean_names()
 ```

### 2018-04-03 docs/paper/Villasenor-Derbez.Rmd
Added render date to PDF

@@
 ```
+
+**Last_update**: `r Sys.Date()`
 
@@
 ```{r, results = "asis"}
-L_conapesca <- tibble(UnidadEconomica = c("Scpp buzos y pescadores de la baja california scl",
-                                          "Scpp cozumel scl",
-                                          "Scpp jose maria azcorra scl"),
-                      Comunidad = c("Isla Natividad",
-                                    "Maria Elena",
-                                    "Punta Herrero")) %>% 
-  left_join(conapesca, ., by = "UnidadEconomica") %>% 
-  mutate(Post = case_when(Comunidad == "Isla Natividad" && Ano <= 2006 ~ 0,
-                          Comunidad == "Maria Elena" && Ano <= 2012 ~ 0,
-                          Comunidad == "Punta Herrero" && Ano <= 2013 ~ 0,
-                          TRUE ~ 1))
-
 C_IN <- filter(L_conapesca, Comunidad == "Isla Natividad") %>% 
@@
 
+```{r}
+invert %>%
+  filter(generoespecie == "Panulirus interruptus",
+         abundancia < 100) %>% 
+  abundance(com = "Isla Natividad") %>%
+  mutate(ano = as.numeric(ano)) %>%
+  ggplot(aes(x = ano, y = indicador, color = zona)) +
+  geom_jitter(height = 0, alpha = 0.5) +
+  stat_summary(geom = "line", fun.y = mean) +
+  scale_color_brewer(palette = "Set1")
+```
+
 \paragraph*{S2 Figure}

### 2018-04-10 docs/data_cleaning_logbook.Rmd
Previous change slinked to cleaning notebook

@@
   filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
-  filter(between(abundancia, 0, 100)) %>% 
+  filter(between(abundancia, 0.1, 100)) %>% 
   group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie) %>% 
@@
 
-write.csv(x = conapesca, file = here::here("data", "conapesca.csv"))
+write.csv(x = conapesca, file = here::here("data", "conapesca.csv"), row.names = F)
 

### 2018-04-10 scripts/model.R
switched to a gaussian error - identity link GLM instead

@@ -1,15 +1,15 @@
-model <- function(x, family = "gaussian", sfe = T, spfe = T){
+model <- function(x, sfe = T, spfe = T){
   
   # x %<>% filter(Comunidad == com)
   
   if(sfe){
-    if(spfe){model <- glm(indicador ~ year * zona + sitio + generoespecie -1, data = x, family = family)} #Site FE and Spp FE
-    if(!spfe){model <- glm(indicador ~ year * zona + sitio -1, data = x, family = family)} #Site FE, no Spp FE
+    if(spfe){model <- lm(indicador ~ zona * year + sitio + generoespecie -1, data = x)} #Site FE and Spp FE
+    if(!spfe){model <- glm(indicador ~ zona * year + sitio -1, data = x)} #Site FE, no Spp FE
   }
   
   if(!sfe){
-    if(spfe){model <- glm(indicador ~ year * zona + generoespecie -1, data = x, family = family)} #No site FE and Spp FE
-    if(!spfe){model <- glm(indicador ~ year * zona -1, data = x, family = family)} #No site and no Spp FE
+    if(spfe){model <- lm(indicador ~ zona * year + generoespecie -1, data = x)} #No site FE and Spp FE
+    if(!spfe){model <- lm(indicador ~ zona * year -1, data = x)} #No site and no Spp FE
   }
   
   return(model)

### 2018-04-10 scripts/plot_model.R
purrr-ified the way in which models are plotted (term plots)

@@ -1,74 +1,29 @@
-plot_model <- function(model, various = F, model_names = NULL, legend = F){
+plot_model <- function(model, model_names = NULL, legend = F){
   
-  if(various){
-    
-    if(is.null(model_names)){model_names <- letters[1:length(model)]}
-    
-    p_all <- data.frame(term = NULL,
-                        estimate = NULL,
-                        std.error = NULL,
-                        statistic = NULL,
-                        p.value = NULL,
-                        model = NULL,
-                        stringsAsFactors = F)
-    
-    for(i in 1:length(model)){
-      tidy_terms <- model[[i]] %>% 
-        robust_se() %>%
-        filter(grepl(":zonaReserva", term)) %$%
-        str_split(term, pattern = ":", simplify = T) %>% 
-        as.tibble() %>% 
-        set_colnames(c("ano", "zona")) %>% 
-        select(-zona) %>% 
-        mutate(Ano = as.numeric(gsub(ano, pattern = "year", replacement = "")),
-               Community = model_names[i])
-      
-      p_i <- robust_se(model[[i]]) %>%
-        filter(grepl(":zonaReserva", term)) %>%
-        cbind(tidy_terms) %>%
-        mutate(p = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05"))
-      
-      p_all <- rbind(p_all,  p_i)
-    }
-    
-    pd <- position_dodge(width = 0.4)
-    
-    p <- p_all %>% 
-      ggplot(aes(x = Ano, y = estimate, color = Community)) +
-      geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, size = 1, position = pd) +
-      geom_point(aes(shape = Community), size = 4, position = pd, alpha = 0.5) +
-      geom_hline(yintercept = 0, linetype = "dashed") +
-      scale_fill_brewer(palette = "Set1") +
-      theme(text = element_text(size = 12)) +
-      labs(x = "Years since implementation", y = quo(lambda[t])) +
-      scale_color_brewer(palette = "Set1") +
-      scale_x_continuous(breaks = seq(0, 10, by = 1))
-    
+  extract_and_filter <- function(model, name){
+    model %>% 
+      robust_se() %>% 
+      filter(grepl(pattern = "zonaReserva:year", x = term)) %>% 
[…]

### 2018-04-10 docs/paper/Villasenor-Derbez.Rmd
Updated manuscript with new model and included CI for socioeco graphs

@@
 \clearpage
+
+```{r load_pkgs}
+suppressPackageStartupMessages({
+  library(sp)
+  library(sf)
+  library(ggsn)
+  library(tmap)
+  library(magrittr)
+  library(kableExtra)
+  library(broom)
+  library(cowplot)
+  library(tidyverse)
+})
+
+# Load indicator functions
+source(file = here::here("scripts", "abundance.R")) #Calculate abundance per transect
+source(file = here::here("scripts", "biomass.R")) # Calculate fish biomass per transect
+source(file = here::here("scripts", "trophic.R")) # Calculate mean trophic level per transect
+
+# Load analysis functions
+source(file = here::here("scripts", "model.R")) #fits the model
+source(file = here::here("scripts", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
+source(file = here::here("scripts", "plot_model.R")) #plots effect sizes
+```
 
@@
 
-```{r load_pkgs}
-suppressPackageStartupMessages({
-  library(sp)
-  library(sf)
-  library(ggsn)
-  library(tmap)
-  library(magrittr)
-  library(kableExtra)
-  library(broom)
-  library(cowplot)
-  library(tidyverse)
-})
+## Data collection
 
-# Load indicator functions
-source(file = here::here("scripts", "abundance.R")) #Calculate abundance per transect
-source(file = here::here("scripts", "biomass.R")) # Calculate fish biomass per transect
-source(file = here::here("scripts", "trophic.R")) # Calculate mean trophic level per transect
+```{r load_data}
+invert <- read.csv(file = here::here("data", "invertebrates.csv"),
+                stringsAsFactors = T)
 
-# Load analysis functions
-source(file = here::here("scripts", "model.R")) #fits the model
-source(file = here::here("scripts", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
-source(file = here::here("scripts", "plot_model.R")) #plots effect sizes
+fish <- read.csv(file = here::here("data", "fish.csv"),
+                stringsAsFactors = T)
+
+conapesca <- read.csv(file = here::here("data", "conapesca.csv"),
+                stringsAsFactors = T)
+
+gob <- read.csv(here::here("data" ,"governance.csv"),
+                strip.white = T,
+                stringsAsFactors = T)
 ```
+
+To perform the evaluation of these reserves we use thee sources of information.
+Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO "Comunidad y Biodiversidad" ([COBI](www.cobi.org.mx)).
+These monitorings record richness and abundances of fish and invertebrate species in the reserves and control sites.
+For fish census, size structures are also collected to derive biomass.
+We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had the same probability of being selected as reserves during the design phase.
+From all the reserves in these three communities, we use the ones that have data for reserve and control sites before and after the implementation of the reserve.
+This provides us with a Before-After-Control-Impact (*i.e.* BACI) design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
+BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
+All reserves were surveyed annually from at least one year before implementation until 2016.
+Table \ref{table:com_sum} shows a summary of the number of reserves, year of implementation, and number of transects for each reserve.
 
@@
 ```
-
-## Data collection
-
-```{r load_data}
-invert <- read.csv(file = here::here("data", "invertebrates.csv"))
-
-fish <- read.csv(file = here::here("data", "fish.csv"))
-
-conapesca <- read.csv(file = here::here("data", "conapesca.csv"))
-
-gob <- read.csv(here::here("data" ,"governance.csv"), strip.white = T, stringsAsFactors = T)
-```
-
-To perform the evaluation of these reserves we use thee sources of information.
-Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO "Comunidad y Biodiversidad" ([COBI](www.cobi.org.mx)).
-These monitorings record richness and abundances of fish and invertebrate species in the reserves and control sites.
-For fish census, size structures are also collected to derive biomass.
-We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had the same probability of being selected as reserves during the design phase.
-From all the reserves in these three communities, we use the ones that have data for reserve and control sites before and after the implementation of the reserve.
-This provides us with a Before-After-Control-Impact (*i.e.* BACI) design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
-BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
-All reserves were surveyed annually from at least one year before implementation until 2016.
-Table \ref{table:com_sum} shows a summary of the number of reserves, year of implementation, and number of transects for each reserve.
 
@@
 
+## Abundance commercially important
+
 # Fish
-
 ## Biomass
@@
 
+## Socioeconomic
+
+Lobster landings and revenue showed a increase after the implementation of the reserves for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}).
+However, the differences in catches and and revenue were not different in the periods before and after the implementation ($p > 0.05$) except for revenues in Isla Natividad, which presented a significant increase of 14.37 (M MXP; $p<0.05$).
+All regression coefficients are presented in **S4 Table**.
+
+## Governance
+
+Although we have little information on the social dimension of these fisheries, using the SES framework indicators (Table \ref{table:gov_ind}), we can analyze the performance of each governance system with respect to MR enforcement (Table \ref{table:gov_res}).
+In general, the presence and success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources they depend upon \citep{jupiter_2017}.
+The enabling conditions for conservation seem to be strongly present in Isla Natividad.
+Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperative.
+These conditions have favored the development of an efficient community based enforcement system.
+In contrast, the communities of Maria Elena and Punta Herrero are located near other fishing communities and cities.
+In Maria Elena, the fishing pressure caused by outsiders can be reduced by implementing a strong enforcement system (in water and land) supported by CSOs and the local government (CONANP).
+Lastly, the community of Punta Herrero shows the highest levels of illegal activities which can be attributed to its connectedness to other communities and the lack of appropriate technologies for enforcement.
+
+\clearpage
+
 ```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for community-levell indicators.
@@
 L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
-                     various = T,
-                     model_names = c("IS", "ME", "PH"))  +
-  theme(legend.justification = c(0, 0),
-        legend.position = c(0, 0.4),
-        legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8),
-        legend.direction = "horizontal") +
-  scale_colour_brewer(guide = guide_legend(title.position = "top", nrow = 3), palette = "Set1")
+                     model_names = c("IS", "ME", "PH"))
 
@@
 I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
-                  various = T,
                   model_names = c("IN", "ME", "PH"))
@@
 F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
-                 various = T,
                  model_names = c("IS", "ME", "PH"))
@@
 F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
-                  various = T,
-                  model_names = c("IN", "ME", "PH"), legend = F)
+                  model_names = c("IN", "ME", "PH")) +
+  theme(legend.justification = c(1, 0),
+        legend.position = c(1, 0.55),
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        legend.direction = "horizontal")
 
@@
 ```{r, fig.height = 3, fig.width = 6, fig.cap = "\\label{fig:res_com}Summarized effects of the marne reserves by direction (positive - negative) and significance."}
+
 observe <- function(model){
-  tidy_terms <- model %>% 
+  model %>% 
     robust_se() %>%
-        filter(grepl(":zonaReserva", term)) %$%
-        str_split(term, pattern = ":", simplify = T) %>% 
-        as.tibble() %>% 
-        set_colnames(c("Ano", "Zona")) %>% 
-        select(-Zona) %>% 
-        mutate(Ano = as.numeric(gsub(Ano, pattern = "year", replacement = "")))
-
-      p_i <- robust_se(model) %>%
-        filter(grepl(":zonaReserva", term)) %>%
-        cbind(tidy_terms) %>%
-        mutate(ptest = ((p.value < 0.05)*-1)+0.05)
-
-       return(p_i)
+    filter(grepl("zonaReserva:", term)) %>% 
+    mutate(Ano = as.numeric(gsub(term, pattern = "zonaReserva:year", replacement = "")),
+           ptest = ((p.value < 0.05)*-1)+0.05)
 }
@@
 IN <- list(N_IN, Ni_IN, B_IN, L_IN) %>%
-  map_df(.f = observe) %>%
+  map_df(observe) %>%
   as.tibble() %>% 
@@
   theme(legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8))
+        legend.title = element_text(size = 8),
+        text = element_text(size = 12))
 ```
 
-## Socioeconomic
+```{r, results = "asis"}
+C_IN <- conapesca %>% 
+  filter(years_centered > -10) %>% 
+  mutate(years_centered = as.factor(years_centered)) %>% 
+  filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
+  filter(zona == "PN") %>%
+  lm(pesovivo/1e3 ~ grupo * years_centered + unidadeconomica, data = .)
 
-Lobster catches and income showed a increase after the implementation of the reserves for Isla Natividad and Maria Elena, but not for Punta Herrero (Fig \ref{fig:lobsters}).
-However, the differences in catches and and revenue were not different in the periods before and after the implementation ($p > 0.05$) except for revenues in Isla Natividad, which presented a significant increase of 14.37 (M MXP; $p<0.05$).
-Table \ref{table:soc_res} presents all the regression coefficients.
+C_ME <- conapesca %>% 
+  filter(years_centered > -10) %>% 
+  mutate(years_centered = as.factor(years_centered)) %>% 
+  filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
+  filter(zona == "MC") %>%
+  lm(pesovivo/1e3 ~ grupo * years_centered + unidadeconomica, data = .)
 
-```{r, results = "asis"}
-C_IN <- filter(L_conapesca, Comunidad == "Isla Natividad") %>% 
-  lm(PesoVivo/1000~Post, data = .)
+V_IN <- conapesca %>% 
+  filter(years_centered > -10) %>% 
+  mutate(years_centered = as.factor(years_centered)) %>% 
+  filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
+  filter(zona == "PN") %>%
+  lm(valor/1e6 ~ grupo * years_centered + unidadeconomica, data = .)
 
-C_ME <- filter(L_conapesca, Comunidad == "Maria Elena") %>% 
-  lm(PesoVivo/1000~Post, data = .)
+V_ME <- conapesca %>% 
+  filter(years_centered > -10) %>% 
+  mutate(years_centered = as.factor(years_centered)) %>% 
+  filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
+  filter(zona == "MC") %>%
+  lm(valor/1e6 ~ grupo * years_centered + unidadeconomica, data = .)
 
-C_PH <- filter(L_conapesca, Comunidad == "Punta Herrero") %>% 
-  lm(PesoVivo/1000~Post, data = .)
+#soc_res se fue
 
-V_IN <- filter(L_conapesca, Comunidad == "Isla Natividad") %>% 
-  lm(Valor/1000000~Post, data = .)
+plot_socioeco <- function(model_list, model_names, ylab){
 
-V_ME <- filter(L_conapesca, Comunidad == "Maria Elena") %>% 
-  lm(Valor/1000000~Post, data = .)
+  extract_filter <- function(model, name){
+    model %>% 
+      robust_se() %>% 
+      filter(grepl(pattern = "grupoTreated:years_centered", x = term)) %>% 
+      mutate(years_centered = as.numeric(gsub(pattern = "grupoTreated:years_centered",
+                                              replacement = "",
+                                              x = term)),
+             Community = name)
+  }
 
-V_PH <- filter(L_conapesca, Comunidad == "Punta Herrero") %>% 
-  lm(Valor/1000000~Post, data = .)
+  purrr::map2_df(model_list, model_names, extract_filter) %>% 
+    ggplot(aes(x = years_centered, y = estimate, group = Community)) +
+    geom_hline(yintercept = 0,
+               linetype = "dashed",
+               size = 1) +
+    geom_vline(xintercept = 0,
+               linetype = "dashed",
+               size = 1) +
+    geom_errorbar(aes(color = Community,
+                      ymin = estimate-std.error,
+                      ymax = estimate + std.error),
+                  width = 0.1,
+                  size = 1,
+                  position = position_dodge(width = 0.4)) +
+    geom_point(aes(color = Community, shape = Community),
+               size = 4,
+               position = position_dodge(width = 0.4),
+               alpha = 0.5) +
+    scale_color_brewer(palette = "Set1") +
+    labs(x = "Years since implementation", y = ylab) +
+    theme(text = element_text(size = 12))
+}
+```
 
-stargazer::stargazer(C_IN, C_ME, C_PH, V_IN, V_ME, V_PH, digits = 2, type = "latex", header = F, omit.stat = c("adj.rsq", "ser", "f"), title = "\\label{table:soc_res} Regression coefficients for lobster catches and revenues for Isla Natividad (1, 4), Maria Elena (2, 5), and Punta Herrero (3, 6).", single.row = T, report = c("vc*"), dep.var.labels = c("Catches (tones)", "Revenue(M MXP)"))
+```{r, fig.width = 5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles)"}
+
+c_plot <- plot_socioeco(model_list = list(C_IN, C_ME),
+                        model_names = list("IN", "ME"),
+                        ylab = "Landings (K MT)")
+
+v_plot <- plot_socioeco(model_list = list(V_IN, V_ME),
+                        model_names = list("IN", "ME"),
+                        ylab = "Revenue (M MXP)")
+
+plot_grid(c_plot, v_plot, ncol = 1)
 
 ```
-
-```{r, fig.width = 3.5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Time series of lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares)."}
-
-L_catches <- ggplot(L_conapesca, aes(x = Ano, y = PesoVivo/1000, color = Comunidad, group = Comunidad, fill = Comunidad)) +
-  geom_smooth(method = "lm") +
-  geom_point(aes(shape = Comunidad), size = 2, alpha = 0.5) +
-  scale_color_brewer(palette = "Set1") +
-  scale_fill_brewer(palette = "Set1") +
-  geom_vline(xintercept = c(2006, 2012, 2013), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +
-  labs(x = "Year", y = "Catches (Tones)") +
-  theme(legend.position = "none",
-        text = element_text(size = 12))
-
-L_value <- ggplot(L_conapesca, aes(x = Ano, y = Valor/1e6, color = Comunidad, group = Comunidad, fill = Comunidad)) +
-  geom_smooth(method = "lm") +
-  geom_point(aes(shape = Comunidad), size = 2, alpha = 0.5) +
-  scale_color_brewer(palette = "Set1") +
-  scale_fill_brewer(palette = "Set1") +
-  geom_vline(xintercept = c(2006, 2012, 2013), color = c("red", "blue", "darkgreen"), linetype = "dashed", size = 0.5) +
-  labs(x = "Year", y = "Value (M MXP)") +
-  theme(legend.position = "none",
-        text = element_text(size = 12))
-
-cowplot::plot_grid(plotlist = list(L_catches, L_value),
-                   labels = "AUTO",
-                   ncol = 1)
-```
-
-## Governance
-
-Although we have little information on the social dimension of these fisheries, using the SES framework indicators (Table \ref{table:gov_ind}), we can analyze the performance of each governance system with respect to MR enforcement (Table \ref{table:gov_res}).
-In general, the presence and success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources they depend upon \citep{jupiter_2017}.
-The enabling conditions for conservation seem to be strongly present in Isla Natividad.
-Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperative.
-These conditions have favored the development of an efficient community based enforcement system.
-In contrast, the communities of Maria Elena and Punta Herrero are located near other fishing communities and cities.
-In Maria Elena, the fishing pressure caused by outsiders can be reduced by implementing a strong enforcement system (in water and land) supported by CSOs and the local government (CONANP).
-Lastly, the community of Punta Herrero shows the highest levels of illegal activities which can be attributed to its connectedness to other communities and the lack of appropriate technologies for enforcement.
 
@@
 
-## Summary of main findings
-
-Our results indicated idiosyncratic biological effects of the reserves across communities and indicators, with a combination of positive and negative effects.
+Our results show idiosyncratic biological effects of the reserves across communities and indicators.
 However, many of these effects were not statistically significant, indicating no effect of the reserve \ref{fig:res_com}.
@@
 
+**Limitations...**
+
 Our results show that community-based marine reserves can be effective if the environmental and social settings allow it.
@@
 
-## Limitations
+<!-- # Thoughts from GainesLab Meeting -->
 
-## Conclusions
+<!-- ## General thoughts -->
 
-# Thoughts from GainesLab Meeting
-
-## General thoughts
-
-A nice story to tell is that reserves need to have optimal social and environmental conditions to work.
+<!-- A nice story to tell is that reserves need to have optimal social and environmental conditions to work.
 But they also need to be opttimally designed, accounting for size and location.
@@
 I still need to get data that support these.
+-->
 
-Creo que hay que agregar informacion de la variabilidad ambienta. Posiblemente SST o algo similar. Ver si se puede obtener datos de oxigeno disuelto... Tal vez algo de ENSO / ONI / NAO o algo asi
+<!-- Creo que hay que agregar informacion de la variabilidad ambienta.
+Posiblemente SST o algo similar.
+Ver si se puede obtener datos de oxigeno disuelto...
+Tal vez algo de ENSO / ONI / NAO o algo asi -->
 
-Quitar analisis de riqueza (la metodologia es restrictiva) y diversidad (aunque shannon provee une buena mezcla del cambio en abundancias y el cambio en riquezas)
+<!-- Quitar analisis de riqueza (la metodologia es restrictiva) y diversidad (aunque shannon provee une buena mezcla del cambio en abundancias y el cambio en riquezas) -->
 
-revisar biomasa!
+<!-- revisar biomasa! -->
 
-Revisar poisson o binomial negativa
+<!-- Revisar poisson o binomial negativa -->
 
-Mapa de El Manchon
+<!-- Mapa de El Manchon -->
 
-Las graficas y el analisis de Profit y Catches tienen que ser de langosta entera fresca solamente, y revisar que no haya habido un cambio de commodity a traves del tiempo para justificar usar solamente entera fresca...
-Tratar de obtener valores hasta 2017
+<!-- Las graficas y el analisis de Profit y Catches tienen que ser de langosta entera fresca solamente, y revisar que no haya habido un cambio de commodity a traves del tiempo para justificar usar solamente entera fresca...
+Tratar de obtener valores hasta 2017 -->
 
-Pedir datos biologicos hasta 2017
+<!-- Pedir datos biologicos hasta 2017 -->
 
-Distinguis profit / revenue / income
+<!-- ## Comentarios del Lab -->
 
-## Comentarios del Lab
+<!-- Owen -->
 
-Owen
+<!-- Are there confounding effects because reserves are so close to control sites? -->
 
-Are there confounding effects because reserves are so close to control sites?
+<!-- Dan -->
 
-Dan
+<!-- Are the data good enough to answer these questions? -->
 
-Are the data good enough to answer these questions?
+<!-- Karly -->
 
-Karly
+<!-- Think about other social indicators that do not depend on catches or revenue from these -->
 
-Think about other social indicators that do not depend on catches or revenue from these
+<!-- Alexa -->
 
-Alexa
+<!-- Daniel -->
 
-Daniel
+<!-- Sebas -->
 
-Sebas
+<!-- Look at other indicators of success (adaptive capacity) in the social / governance data -->
 
-Look at other indicators of success (adaptive capacity) in the social / governance data
+<!-- Steve -->
 
-Steve
+<!-- Would we expect to see effects in the first five years? (ME and PH) -->
 
-Would we expect to see effects in the first five years? (ME and PH)
+<!-- Are the designs good enough to see effects? -->
 
-Are the designs good enough to see effects?
+<!-- Becky -->
 
-Becky
+<!-- Ignacia -->
 
-Ignacia
+<!-- Perhaps interesting follow-up projects include comparing perceptions vs. reality -->
 
-Perhaps interesting follow-up projects include comparing perceptions vs. reality
-
-Estimating the benefits of being friends with an NGO
+<!-- Estimating the benefits of being friends with an NGO -->
 
@@
 
-```{r}
+```{r, eval = F}
 invert %>%
@@
 
+```{r, eval = F}
+
+# Series de tiempo de langosta
+L_catches <- ggplot(L_conapesca,
+                    aes(x = Ano,
+                        y = PesoVivo/1000,
+                        color = Comunidad,
+                        group = Comunidad,
+                        fill = Comunidad)) +
+  geom_smooth(method = "lm") +
+  geom_point(aes(shape = Comunidad), size = 2, alpha = 0.5) +
+  scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
+  geom_vline(xintercept = c(2006, 2012, 2013),
+             color = c("red", "blue", "darkgreen"),
+             linetype = "dashed", size = 0.5) +
+  labs(x = "Year", y = "Catches (Tones)") +
+  theme(legend.position = "none",
+        text = element_text(size = 12))
+
+L_value <- ggplot(L_conapesca, aes(x = Ano,
+                                   y = Valor/1e6,
+                                   color = Comunidad,
+                                   group = Comunidad,
+                                   fill = Comunidad)) +
+  geom_smooth(method = "lm") +
+  geom_point(aes(shape = Comunidad), size = 2, alpha = 0.5) +
+  scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
+  geom_vline(xintercept = c(2006, 2012, 2013),
+             color = c("red", "blue", "darkgreen"),
+             linetype = "dashed",
+             size = 0.5) +
+  labs(x = "Year", y = "Value (M MXP)") +
+  theme(legend.position = "none",
+        text = element_text(size = 12))
+
+cowplot::plot_grid(plotlist = list(L_catches, L_value),
+                   labels = "AUTO",
+                   ncol = 1)
+```
+
 \bibliographystyle{frontiersinSCNS_ENG_HUMS}

### 2018-04-10 docs/paper/Villasenor-Derbez.tex
fixed mistakes due to reorganization of repo
Added render date to PDF
Updated manuscript with new model and included CI for socioeco graphs

--- 2018-02-21 Frontiers_Manuscript/Villasenor-Derbez.tex
@@
 \tiny
- \keyFont{ \section{Keywords:} Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen
-Science, Mexico, Social--Ecological Systems}
+ \keyFont{ \section{Keywords:} Marine Protected Areas, Marine Conservation, Small-Scale Fisheries,
+Citizen Science, Mexico, Social-Ecological Systems}
 
 \end{abstract}
+
+\textbf{Last\_update}: 2018-04-10
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Location of the three
-coastal communities studied (A). Isla Natividad (B) is located off the
-Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are
-located in the yucatan Peninsula.}
-\end{figure}
-
 \subsection{Data collection}\label{data-collection}
@@
 
+\begin{figure}
+\centering
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-1-1.pdf}
+\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Location of the three
+coastal communities studied (A). Isla Natividad (B) is located off the
+Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are
+located in the yucatan Peninsula.}
+\end{figure}
+
 \begin{table}
@@
 \hline
-Isla Natividad & La Plana / Las Cuevas - La Dulce / Babencho & 2006 & 2006 & 405 & 242 & 415 & 245\\
+Isla Natividad & La Plana / Las Cuevas - La Dulce / Babencho & 2006 & 2006 & 400 & 241 & 415 & 244\\
 \hline
@@
 \hline
-Punta Herrero & Manchon - Manchon (Control) & 2013 & 2012 & 43 & 45 & 27 & 42\\
+Punta Herrero & Manchon - Manchon (Control) & 2013 & 2012 & 43 & 45 & 27 & 41\\
 \hline
@@
 
-\clearpage
-
 \section{Results}\label{results}
@@
 
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{\label{fig:unnamed-chunk-5}\label{fig:indicators}Effect sizes for
-marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
-blue triangles), and Punta Herrero (PH; green squares) for
-community-levell indicators. Plots are ordered by survey type (left:
-fish; right: invertebrates) and indicators: Abundance (A, B), Richness
-(C, D), Shannon's diversity index (E, F),fish biomass (G), and lobster
-(\emph{Panulirus spp}) abundances (H). Points are jittered hotizontally
-to avoid overplotting. Points indicate the effect size, and errorbars
-are heteroskedastic-robust standard errors.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
-\caption{\label{fig:unnamed-chunk-6}\label{fig:res_com}Summarized effects of
-the marne reserves by direction (positive - negative) and significance.}
-\end{figure}
-
 \subsection{Socioeconomic}\label{socioeconomic}
 
-Lobster catches and income showed a increase after the implementation of
-the reserves for Isla Natividad and Maria Elena, but not for Punta
-Herrero (Fig \ref{fig:lobsters}). However, the differences in catches
-and and revenue were not different in the periods before and after the
-implementation (\(p > 0.05\)) except for revenues in Isla Natividad,
-which presented a significant increase of 14.37 (M MXP; \(p<0.05\)).
-Table \ref{table:soc_res} presents all the regression coefficients.
-
-\begin{table}[!htbp] \centering 
-  \caption{\label{table:soc_res} Regression coefficients for lobster catches and revenues for Isla Natividad (1, 4), Maria Elena (2, 5), and Punta Herrero (3, 6).} 
-  \label{} 
-\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-7} 
-\\[-1.8ex] & \multicolumn{3}{c}{Catches (tones)} & \multicolumn{3}{c}{Revenue(M MXP)} \\ 
-\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
-\hline \\[-1.8ex] 
- Post & 7.37 & 5.83 & $-$1.26 & 14.37$^{***}$ & 1.24 & $-$0.06 \\ 
-  Constant & 122.68$^{***}$ & 11.41$^{***}$ & 6.06$^{***}$ & 10.89$^{***}$ & 2.04$^{***}$ & 1.06$^{***}$ \\ 
- \hline \\[-1.8ex] 
-Observations & 15 & 14 & 13 & 15 & 14 & 13 \\ 
-R$^{2}$ & 0.02 & 0.04 & 0.01 & 0.52 & 0.04 & 0.0004 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{\label{fig:unnamed-chunk-8}\label{fig:lobsters}Time series of
-lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
-circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green
-squares).}
-\end{figure}
+Lobster landings and revenue showed a increase after the implementation
+of the reserves for Isla Natividad and Maria Elena (Fig
+\ref{fig:lobsters}). However, the differences in catches and and revenue
+were not different in the periods before and after the implementation
+(\(p > 0.05\)) except for revenues in Isla Natividad, which presented a
+significant increase of 14.37 (M MXP; \(p<0.05\)). All regression
+coefficients are presented in \textbf{S4 Table}.
 
@@
 appropriate technologies for enforcement.
+
+\clearpage
+
+\begin{figure}
+\centering
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{\label{fig:unnamed-chunk-5}\label{fig:indicators}Effect sizes for
+marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
+blue triangles), and Punta Herrero (PH; green squares) for
+community-levell indicators. Plots are ordered by survey type (left:
+fish; right: invertebrates) and indicators: Abundance (A, B), Richness
+(C, D), Shannon's diversity index (E, F),fish biomass (G), and lobster
+(\emph{Panulirus spp}) abundances (H). Points are jittered hotizontally
+to avoid overplotting. Points indicate the effect size, and errorbars
+are heteroskedastic-robust standard errors.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{\label{fig:unnamed-chunk-6}\label{fig:res_com}Summarized effects of
+the marne reserves by direction (positive - negative) and significance.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-8-1.pdf}
+\caption{\label{fig:unnamed-chunk-8}\label{fig:lobsters}Effect sizes for
+lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
+circles) and Maria Elena (ME; blue triangles)}
+\end{figure}
 
@@
 
-Our results indicated idiosyncratic biological effects of the reserves
-across communities and indicators, with a combination of positive and
-negative effects. However, many of these effects were not statistically
-significant, indicating no effect of the reserve \ref{fig:res_com}. The
-socioeconomic indicators pertaining to landings and revenues associated
-to those landings showed little or no temporal change before and after
-reserve implementation. These contrasting effects, however, might be
-clarified when understanding the social-ecological context in which
-these communities and their reserves sit. In this section, we discuss
-potential shortcomings in our analysis, and provide plausible
-explanations to the observed biological and socioeconomic basing on
-previous literature and our analysis of the social-ecological system.
+Our results show idiosyncratic biological effects of the reserves across
+communities and indicators. However, many of these effects were not
+statistically significant, indicating no effect of the reserve
+\ref{fig:res_com}. The socioeconomic indicators pertaining to landings
+and revenues associated to those landings showed little or no temporal
+change before and after reserve implementation. These contrasting
+effects, however, might be clarified when understanding the
+social-ecological context in which these communities and their reserves
+sit. In this section, we discuss potential shortcomings in our analysis,
+and provide plausible explanations to the observed biological and
+socioeconomic basing on previous literature and our analysis of the
+social-ecological system.
 
@@
 in their reserves.
+
+\textbf{Limitations\ldots{}}
 

### 2018-04-24 scripts/create_turf_shapefile.R
Added script to join TURF shapefiles and modified the model by allowing for an intercept

@@ -1,0 +1,74 @@
+# Mon Apr 23 19:41:00 2018 ------------------------------
+
+# Create an sf_object with TURF polygons
+
+## For Isla Natividad
+# Polygon vertices from http://www.dof.gob.mx/nota_detalle.php?codigo=4692631&fecha=14/10/1992
+
+library(tidyverse)
+library(sf)
+
+# Create vectors of latitudes and longitudes
+lat <- c(
+  27.96666667,
+  27.96666667,
+  27.89638889,
+  27.72083333,
+  27.795,
+  27.96666667
+)
+
+lon <- c(
+  -115.1019444,
+  -115.4,
+  -115.6311111,
+  -115.2777778,
+  -115.1255556,
+  -115.1019444
+)
+
+# Transform vectors into a simple feature
+pol <- cbind(lon, lat) %>% 
+  list() %>% 
+  st_polygon() %>% 
+  st_sfc(crs = "+proj=longlat +datum=NAD27")
+
+# Create a dataframe with the above as simple feature
+dat <- data.frame(Coop = "SCPP Buzos y Pescadores de la Baja California", stringsAsFactors = F)
+st_geometry(dat) <- pol
+
+## For the Caribbean ones
+
+# Caribbean TURFs were supplied by Stuart Fulton
+# Read them in and append the Isla Natividad Polygon
+# We also reproject and calculate the area
+
+turf <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ConcessionesQRoo") %>% 
+  st_cast(to = "POLYGON") %>% 
+  st_transform(crs = "+proj=longlat +datum=NAD27") %>% 
+  select(Coop = Cooperativ) %>% 
+  rbind(dat) %>% 
+  mutate(Area = as.numeric(st_area(.))/1000)
+
+# Visualize to inspect
+ggplot() +
+  geom_sf(data = turf, fill = "red")
+
+# Export shapefiles
+st_write(turf, dsn = here::here("raw_data", "spatial", "turfs.shp"), delete_dsn = T)
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2018-04-24 scripts/model.R

@@ -3,13 +3,13 @@ model <- function(x, sfe = T, spfe = T){
   # x %<>% filter(Comunidad == com)
   
   if(sfe){
-    if(spfe){model <- lm(indicador ~ zona * year + sitio + generoespecie -1, data = x)} #Site FE and Spp FE
-    if(!spfe){model <- glm(indicador ~ zona * year + sitio -1, data = x)} #Site FE, no Spp FE
+    if(spfe){model <- lm(indicador ~ zona * year + sitio + generoespecie, data = x)} #Site FE and Spp FE
+    if(!spfe){model <- glm(indicador ~ zona * year + sitio, data = x)} #Site FE, no Spp FE
   }
   
   if(!sfe){
-    if(spfe){model <- lm(indicador ~ zona * year + generoespecie -1, data = x)} #No site FE and Spp FE
-    if(!spfe){model <- lm(indicador ~ zona * year -1, data = x)} #No site and no Spp FE
+    if(spfe){model <- lm(indicador ~ zona * year + generoespecie, data = x)} #No site FE and Spp FE
+    if(!spfe){model <- lm(indicador ~ zona * year, data = x)} #No site and no Spp FE
   }
   
   return(model)

### 2018-04-24 docs/paper/Villasenor-Derbez.Rmd
Updated analysis in manuscript. Must still update text!

@@
 zrp <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
-  filter(Name %in% c("El Cabezo", "El Faro")) %>% 
-  st_transform(crs = "+proj=longlat +datum=NAD27")
+  filter(Name %in% c("El Cabezo",
+                     "Gallineros",
+                     "Punta Loria",
+                     "San Roman Norte",
+                     "San Roman Sur",
+                     "Anegado de Chal",
+                     "Laguna Canche Balam",
+                     "Niluc",
+                     "Mimis",
+                     "El Faro (Pesca de Langosta)",
+                     "El Faro")) %>% 
+  st_transform(crs = "+proj=longlat +datum=NAD27") %>% 
+  mutate(evaluated = ifelse(Name %in% c("El Cabezo", "El Faro (Pesca de Langosta)", "El Faro"), T, F))
 
@@
 rescom <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
-  filter(Nombre == "La_Plana") %>% 
+  filter(Nombre %in% c("La_Plana", "PuntaPrieta")) %>% 
   st_transform(crs = "+proj=longlat +datum=NAD27")
+
+turf <- st_read(dsn = here::here("raw_data", "spatial"), layer = "turfs", quiet = T) %>% 
+  filter(Coop %in% c("SCPP Buzos y Pescadores de la Baja California", "SCPP Cozumel", "SCPP Jose Maria Azcorra"))
 
@@
   scale_fill_brewer(palette = "Set1") +
-  theme(legend.position = "none") +
-  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863))
+  theme(legend.position = "none",
+        panel.grid.major = element_line(colour = "transparent"),
+        axis.text = element_text(size = 7)) +
+  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863)) +
+  annotate(geom = "text", x = -102, y = 25, label = "Mexico") +
+  annotate(geom = "text", x = -90, y = 25, label = "Gulf of\nMexico") +
+  annotate(geom = "text", x = -112, y = 20, label = "Pacific\nOcean") +
+  labs(x = "", y = "")
 
-# Map for Punta Herrero
+# Map for SAM
 ph <- ggplot() +
+  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
   geom_sf(data = coastline_mx) +
-  lims(x = c(-87.5, -87.4), y = c(19.29, 19.36)) +
-  geom_sf(data = zrp, fill = "red") +
-  theme_classic()
+  lims(x = c(-87.7, -87.35), y = c(19.15, 19.6)) +
+  geom_sf(data = zrp, fill = "red2") +
+  theme_classic() +
+  theme(panel.grid.major = element_line(colour = "transparent"),
+        axis.text = element_text(size = 7)) +
+  annotate(geom = "text", x = -87.6, y = 19.45, label = "Maria Elena") +
+  annotate(geom = "text", x = -87.64, y = 19.15, label = "Punta Herrero") +
+  labs(x = "", y = "")
 
-# Map for Maria Elena
-ma <- ggplot() +
+# Map for PBC
+pbc <- ggplot() +
+  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
+  lims(x = c(-115.65, -115.109), y = c(27.7, 28)) +
   geom_sf(data = coastline_mx) +
-  lims(x = c(-87.54, -87.44), y = c(19.38, 19.44)) +
-  geom_sf(data = zrp, fill = "red") +
-  theme_classic()
+  geom_sf(data = rescom, fill = "red2") +
+  theme_classic() +
+  theme(panel.grid.major = element_line(colour = "transparent"),
+        axis.text = element_text(size = 7)) +
+  labs(x = "", y = "")
 
-# Map for Natividad
-pbc <- ggplot() +
-  geom_sf(data = coastline_mx) +
-  lims(x = c(-115.24, -115.14), y = c(27.83, 27.90)) +
-  geom_sf(data = rescom, fill = "red") +
-  theme_classic()
+left <- plot_grid(mex, pbc, ncol = 1, labels = "AUTO")
 
-plot_grid(mex, ma, pbc, ph, labels = c("A", "C", "B", "D"), ncol = 2)
+plot_grid(left, ph, labels = c("", "C"), ncol = 2)
 ```

### 2018-05-14 scripts/abundance.R
Added wrapper function to create list of standard errors for stargazer

@@ -7,9 +7,10 @@ abundance <- function(x, com, spp = NULL){
   results <- x %>% 
     filter(comunidad == com) %>%
     mutate(ano = as.factor(ano),
-           year = as.factor(year)) %>% 
+           year = as.factor(year),
+           year = fct_relevel(year, "0")) %>% 
     group_by(rc, ano, year, sitio, zona, transecto, generoespecie) %>% 
-    summarize(indicador = sum(abundancia)) %>% 
+    summarize(indicador = sum(abundancia)/60) %>% 
     ungroup()
     
     return(results)

### 2018-05-14 scripts/biomass.R

@@ -3,9 +3,10 @@ biomass <- function(x, com){
     filter(comunidad == com) %>% 
     mutate(ano = as.factor(ano),
            year = as.factor(year),
+           year = fct_relevel(year, "0"),
            biomass = abundancia * a * (talla^b)/1000) %>% 
     group_by(rc, ano, year, sitio, zona, transecto, generoespecie) %>% 
-    summarize(indicador = sum(biomass, na.rm = T)) %>% 
+    summarize(indicador = sum(biomass, na.rm = T)/60) %>% 
     ungroup()
   
   return(results)

### 2018-05-14 scripts/model.R

@@ -4,12 +4,12 @@ model <- function(x, sfe = T, spfe = T){
   
   if(sfe){
     if(spfe){model <- lm(indicador ~ zona * year + sitio + generoespecie, data = x)} #Site FE and Spp FE
-    if(!spfe){model <- glm(indicador ~ zona * year + sitio, data = x)} #Site FE, no Spp FE
+    if(!spfe){model <- lm(indicador ~ zona * year + sitio, data = x)} #Site FE, no Spp FE
   }
   
   if(!sfe){
     if(spfe){model <- lm(indicador ~ zona * year + generoespecie, data = x)} #No site FE and Spp FE
-    if(!spfe){model <- lm(indicador ~ zona * year, data = x)} #No site and no Spp FE
+    if(!spfe){model <- lm(indicador ~ zona * year , data = x)} #No site and no Spp FE
   }
   
   return(model)

### 2018-05-14 scripts/plot_model.R

@@ -23,7 +23,7 @@ plot_model <- function(model, model_names = NULL, legend = F){
     theme(text = element_text(size = 12)) +
     labs(x = "Years since implementation", y = quo(lambda[t])) +
     scale_color_brewer(guide = guide_legend(title.position = "top", nrow = 2), palette = "Set1") +
-    scale_x_continuous(breaks = seq(0, 10, by = 1))
+    scale_x_continuous(breaks = seq(-1, 10, by = 1))
   
   if(!legend){
     p <- p +

### 2018-05-14 scripts/robust_se_list.R

@@ -1,0 +1,8 @@
+robust_se_list <- function(model){
+  robust_data_frame <- robust_se(model)
+  
+  robust_se_errors <- robust_data_frame$std.error
+  names(robust_se_errors) <- robust_data_frame$term
+  
+  return(robust_se_errors)
+}

### 2018-05-14 docs/paper/Villasenor-Derbez.Rmd
Added new version of manuscript

--- 2018-04-10 docs/paper/Villasenor-Derbez.Rmd
@@
   library(broom)
+  library(ggrepel)
   library(cowplot)
@@
 source(file = here::here("scripts", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
+source(file = here::here("scripts", "robust_se_list.R")) #heteroskedastic-robust Standard Errors for stargazer
 source(file = here::here("scripts", "plot_model.R")) #plots effect sizes
@@
 However, there is considerably less literature focusing on the relationship between socioeconomic and governance structures and their relationship to ecological effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or benefits to fisheries \citep{krueck_2017-J1}; evaluations of marine reserves rarely provide a holistic view of the social-ecological system \citep{lpezangarita_2014}.
-Here, we combine causal inference techniques \citep{depalma_2018} and the social-ecological systems framework \citep{ostrom_2009-hg} to provide a comprehensive ecological and socioeconomic evaluation of four community-based marine reserves in three coastal communities in Mexico.
+
+While ecological factors (*e.g.* habitat representation, initial state of protection, connectivity to other protected areas) can determine the success of a MR, their effectiveness also depends on the socioeconomic and governance settings under which they are implemented and managed.
+Literature shows that many non-ecological characteristics can play an equally important role in the effectiveness of MRs.
+For example, age of a reserve (*i.e.* time since its implementation) and size were key to the effectiveness of MRs in Palau \citep{friedlander_2017-oI}.
+In the Mediterranean, \citet{difranco_2016-Xw} identify that surveillance and enforcement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key factors that increased stock health and income to fishers.
+At a global level, \citet{edgar_2014-UO} indicate that enforcement, age, size, and isolation were important factors determining effectiveness of the reserves.
 
@@
 While CONANP has made efforts to have a participatory process, the implementation of these zones is still characterized by top--down approaches.
-This motivated Civil Society Organizations (CSOs) to work with coastal communities to implement community--based marine reserves \citep{uribe_2010-u2}, which are usually established within a Territorial Use Rights for Fisheries (TURFs); thus making them TURF--reserves \citep{afflerbach_2014-HP}.
+This motivated Civil Society Organizations (CSOs) to work with coastal communities to implement community--based marine reserves \citep{uribe_2010-u2}, which are usually established within a TURFs; thus making them TURF--reserves \citep{afflerbach_2014-HP}.
 This bottom--up approach allows fishers to design their own reserves, which  increases compliance and self--enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
@@
 In 2014, a new norm \citep{nom} allowed fishers to request the legal recognition of a community--based reserve under the name of "Fishing Refugia" (*Zona de Refugio Pesquero*, FR).
-This new norm thus combines bottom--up approaches to design marine reserves, along with a legal recognition of the management intervention.
-Since then, *39* FR have been implemented along the Pacific, Gulf of California, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.
+Since then, **39** FR have been implemented along the Pacific, Gulf of California, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.
 
-While there are ecological factors defining the success of a MR (*i.e.* habitat representation, initial state of protection, connectivity to other protected areas), their effectiveness also depends on the socioeconomic and governance settings under which they are implemented.
-Literature shows that many non-ecological characteristics can play an equally important role in the effectiveness of MRs.
-For example, age of a reserve (*i.e.* time since its implementation), size, and habitat contained were key to the effectiveness of MRs in Palau \citep{friedlander_2017-oI}.
-In the Mediterranean, \citet{difranco_2016-Xw} identify that surveillance and enforcement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key factors that increased stock health and income to fishers.
-At a global level, \citet{edgar_2014-UO} indicate that enforcement, age, size, and isolation were important factors determining effectiveness of the reserves.
-
+We combine causal inference techniques \citep{depalma_2018} and the social-ecological systems framework \citep{ostrom_2009-hg} to provide a comprehensive ecological and socioeconomic evaluation of community-based marine reserves in three coastal communities in Mexico.
 The objective of this work is twofold: i) Provide the first evaluation of community--based marine reserves in Mexico, and ii) provide a comprehensive evaluation of the social--ecological system to identify how socioeconomic and governance characteristics relate to ecological effectiveness.
-With the purpose of providing a holistic evaluation, we combine ecological, socioeconomic, and governance indicators.
-We use causal inference techniques to provide a measurement of the effect of the management intervention, and combine it with the social--ecological systems framework \citep{ostrom_2009-hg}.
 
@@
 
-We focus our evaluation in three coastal communities from the Pacific coast of Baja California (n = 1) and the Mesoamerican Reef System (n = 2; Fig \ref{fig:map}).
+We focus our evaluation in three coastal communities from the Pacific coast of Baja California and the Mesoamerican Reef System (Fig \ref{fig:map}).
+Fishing activities in all communities are assambled around fishing cooperatives that hold Territorial Use Rights for Fisheries (TURFs).
 Isla Natividad (IN) lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests (*Macrocystis pyrifera*) and rocky reefs are the predominant habitats.
-The island is home to a fishing cooperative (*Sociedad Cooperative de Producción Pesquera Buzos y Pescadores de la Baja California SCL*), that holds a TURF for spiny lobster (*Panulirus interruptus*).
+The island is home to a the *Sociedad Cooperative de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California*, spiny lobster (*Panulirus interruptus*) represents an important fishery.
 However, other resources like finfish (yellow-tail jack, *Seriola lalandi*), sea cucumber (*Parastichopus parvimensis*), red sea urchin (*Mesocentrotus franciscanus*), snail (*Megastraea turbanica* y *M.
@@
 In 2006, the community decided to implement two community--based marine reserves within their fishing grounds, seeking to recover depleted stocks of invertebrate species (mainly lobster and abalone).
-Until today, these reserves are yet to be legally recognized as Fishing Refugia.
+Until today, these reserves are yet to be legally recognized as Fishing Refugia but count with full support from the community.
 
-The other two communities are Maria Elena (ME; Fig \ref{fig:map}C) and Punta Herrero (PH; Fig \ref{fig:map}D) in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems.
-ME is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative.
-PH is home to the "José María Azcorra" fishing cooperative.
-The main source of income to both communities is the Caribbean spiny lobster fishery (*Panulirus argus*), which is carried out within their respective TURFs.
+The other two communities are Maria Elena (ME) and Punta Herrero (PH; Fig \ref{fig:map}C) in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems.
+ME is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*).
+PH is also home to a fishing cooperative (*SCPP José María Azcorra*).
+The main source of income to both communities is the Caribbean spiny lobster (*Panulirus argus*).
 These communities also target finfish in the off season, mainly snappers (Lutjanidae) and groupers (Serranidae).
-ME established eight marine reserves in 2012, and PH established four marine reserves in 2013.
+ME established eight marine reserves in 2012, and PH established four marine reserves in 2013 and an additional community-based (*i.e.* not legally recognized).
 All these reserves are legally recognized as Fishing Refugia.
@@
 
-To perform the evaluation of these reserves we use thee sources of information.
+We use three main sources of information to evaluate these reserves.
 Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO "Comunidad y Biodiversidad" ([COBI](www.cobi.org.mx)).
 These monitorings record richness and abundances of fish and invertebrate species in the reserves and control sites.
-For fish census, size structures are also collected to derive biomass.
+Size structures are also collected in fish census.
 We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had the same probability of being selected as reserves during the design phase.
 From all the reserves in these three communities, we use the ones that have data for reserve and control sites before and after the implementation of the reserve.
-This provides us with a Before-After-Control-Impact (*i.e.* BACI) design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
+This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
 BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
 All reserves were surveyed annually from at least one year before implementation until 2016.
-Table \ref{table:com_sum} shows a summary of the number of reserves, year of implementation, and number of transects for each reserve.
+Table \ref{table:com_sum} shows a summary of sampling effort and year of implementation of the analyzed reserves.
 
@@
 
-# Declare points for reference map
+# Load ZRP shapefile
+zrp <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
+  filter(Name %in% c("El Cabezo",
+                     "Gallineros",
+                     "Punta Loria",
+                     "San Roman Norte",
+                     "San Roman Sur",
+                     "Anegado de Chal",
+                     "Laguna Canche Balam",
+                     "Niluc",
+                     "Mimis",
+                     "El Faro (Pesca de Langosta)",
+                     "El Faro")) %>% 
+  st_transform(crs = proj) %>% 
+  mutate(evaluated = ifelse(Name %in% c("El Cabezo", "El Faro (Pesca de Langosta)", "El Faro"), T, F))
+
+# Load Community reserve shapefile
+rescom <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
+  filter(Nombre %in% c("La_Plana", "PuntaPrieta")) %>% 
+  st_transform(crs = proj)
+
+turf <- st_read(dsn = here::here("raw_data", "spatial"), layer = "turfs", quiet = T) %>% 
+  filter(Coop %in% c("SCPP Buzos y Pescadores de la Baja California", "SCPP Cozumel", "SCPP Jose Maria Azcorra"))
+
+```
+
+```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A).
+Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are located in the Yucatan Peninsula."}
+# Define reference map
+
+## Declare points for reference map
 location_labels <- data.frame(X = c(-115.24, -87.4611756),
@@
                               stringsAsFactors = F)
-coordinates(location_labels) <- c("X", "Y")
-proj4string(location_labels) <- proj
-location_labels <- spTransform(location_labels, proj) %>% 
-  st_as_sf()
 
-# Load ZRP shapefile
-zrp <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
-  filter(Name %in% c("El Cabezo", "El Faro")) %>% 
-  st_transform(crs = "+proj=longlat +datum=NAD27")
-
-# Load Community reserve shapefile
-rescom <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
-  filter(Nombre == "La_Plana") %>% 
-  st_transform(crs = "+proj=longlat +datum=NAD27")
-
-```
-
-```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A).
-Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are located in the yucatan Peninsula."}
-# Define reference map
+## Create map
 mex <- ggplot() +
   geom_sf(data = W) +
-  geom_sf(data = location_labels, color = "steelblue", size = 3) +
+  geom_point(data = location_labels, aes(x = X, y = Y), color = "black", fill = "steelblue", size = 3, shape = 21) +
   theme_classic() +
@@
   scale_fill_brewer(palette = "Set1") +
-  theme(legend.position = "none") +
-  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863))
+  theme(legend.position = "none",
+        panel.grid.major = element_line(colour = "transparent"),
+        axis.text = element_text(size = 7)) +
+  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863)) +
+  annotate(geom = "text", x = -102, y = 25, label = "Mexico") +
+  annotate(geom = "text", x = -90, y = 25, label = "Gulf of\nMexico") +
+  annotate(geom = "text", x = -112, y = 20, label = "Pacific\nOcean") +
+  labs(x = "", y = "")
 
-# Map for Punta Herrero
+# Map for SAM
+
+## Declare points for SAM
+sam_points <- data.frame(x = c(-87.4646745, -87.539104),
+                         y = c(19.3229767, 19.396729),
+                         location = c("Punta Herrero", "Maria Elena"))
+
+## Create map
 ph <- ggplot() +
+  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
   geom_sf(data = coastline_mx) +
-  lims(x = c(-87.5, -87.4), y = c(19.29, 19.36)) +
-  geom_sf(data = zrp, fill = "red") +
-  theme_classic()
+  lims(x = c(-87.7, -87.35), y = c(19.15, 19.6)) +
+  geom_sf(data = zrp, fill = "red2") +
+  theme_classic() +
+  theme(panel.grid.major = element_line(colour = "transparent"),
+        axis.text = element_text(size = 7)) +
+  geom_label_repel(data = sam_points, aes(x = x, y = y, label = location), nudge_y = 0.05, nudge_x = 0.05) +
+  labs(x = "", y = "") +
+  scalebar(data = turf[3,], dist = 5, dd2km = T, model = 'WGS84', location = "bottomright", st.size = 2, st.dist = 0.05, height = 0.1) +
+  scale_x_continuous(breaks = seq(-87.7, -87.3, 0.1), limits = c(-87.7, -87.3))
 
-# Map for Maria Elena
-ma <- ggplot() +
+# Map for PBC
+pbc <- ggplot() +
+  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
+  lims(x = c(-115.65, -115.109), y = c(27.7, 28)) +
   geom_sf(data = coastline_mx) +
-  lims(x = c(-87.54, -87.44), y = c(19.38, 19.44)) +
-  geom_sf(data = zrp, fill = "red") +
-  theme_classic()
+  geom_sf(data = rescom, fill = "red2") +
+  theme_classic() +
+  theme(panel.grid.major = element_line(colour = "transparent"),
+        axis.text = element_text(size = 7)) +
+  labs(x = "", y = "") +
+  scalebar(data = turf[4,], dist = 10, dd2km = T, model = 'WGS84', location = "bottomleft", st.size = 2, st.dist = 0.05, height = 0.1)
 
-# Map for Natividad
-pbc <- ggplot() +
-  geom_sf(data = coastline_mx) +
-  lims(x = c(-115.24, -115.14), y = c(27.83, 27.90)) +
-  geom_sf(data = rescom, fill = "red") +
-  theme_classic()
+left <- plot_grid(mex, pbc, ncol = 1, labels = "AUTO")
 
-plot_grid(mex, ma, pbc, ph, labels = c("A", "C", "B", "D"), ncol = 2)
+plot_grid(left, ph, labels = c("", "C"), ncol = 2)
 ```
@@
 ```{r}
-inv_summary <- group_by(invert, comunidad, rc, zona, sitio, transecto, ano) %>%
-  count() %>% 
-  ungroup() %>% 
-  group_by(comunidad, rc, zona) %>% 
-  summarize(n = n()) %>% 
-  ungroup() %>% 
-  spread(zona, n) %>% 
-  select(comunidad, rc, control = Control, reserve = Reserva)
-
-group_by(fish, comunidad, rc, zona, sitio, transecto, ano) %>%
-  count() %>% 
-  ungroup() %>% 
-  group_by(comunidad, rc, zona) %>% 
-  summarize(ano = min(ano), n = n()) %>% 
-  ungroup() %>% 
-  spread(zona, n) %>% 
-  mutate(implementation = c(2006, 2012, 2013, 2013)) %>% 
-  select(comunidad, rc, implementation, ano, control = Control, reserve = Reserva) %>% 
-  left_join(inv_summary, by = c("comunidad", "rc")) %>% 
+tibble(Community = c("Isla Natividad", "Maria Elena", "Punta Herrero"),
+       TURF = c(889.5, 353.1, 299.7),
+       reserve = c(1.53, 0.10, 0.43),
+       percent = reserve / TURF * 100,
+       implementation = c(2006, 2012, 2013)) %>%
   knitr::kable(format = "latex",
-               caption = "\\label{table:com_sum} Summary of commuity--based marine reserves by community.
-Imp = Year of implementation, Start = Year of first sampling, number of fish transects in control (Cf) and reserve (Rf) sites, and number of invertebrate transects in Control (Ci) and Reserve (Ri) sites.",
-               col.names = c("Community", "Reserve - Control", "Imp", "Start", "Cf", "Rf", "Ci", "Ri"))
+               caption = "\\label{table:com_sum} Summary of commuity--based marine reserves by community.",
+               col.names = c("Community", "TURF area ($km^2$)", "Reserve area ($km^2$)", "Percent as reserves","Year of implementation"),
+               escape = F)
 ```
@@
 
-Governance data were collected at the community-level.
-The information was compiled by combining key informants and the authors; experience and knowledge of the communities to collect the necessary information.
-These data contain information on the ecological system where the fishing activities develop, as well as the governance structures present in the cooperative.
-We also gathered information on the resource unit (*i.e.* lobsters) and the relevant actors present in each community \citep{leslie_2015-na}.
+We use a set of governance indicators to analyze the governance structures of each cooperative \citep{leslie_2015-na}.
+The indicators (Table \ref{table:gov_ind}) focus on the resource system (four indicators), governance system (seven indicators), resource units (three indicators) and actors (three indicators).
+These were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2012} and based on the authors' experience and knowledge of the communities.
 
@@
 
-Following a framework that relates reserve objectives to performance indicators \citep{Villasenor-Derbez_2018}, we use five biological and two socioeconomic indicators to evaluate these marine reserves Table \ref{table:indicators}.
-We also use a set of governance indicators to analyze the governance structures of each cooperative \citep{leslie_2015-na}.
-The indicators (Table \ref{table:gov_ind}) focus on the resource system (four indicators), governance system (seven indicators), resource units (three indicators) and actors (three indicators).
+We evaluate the effect thar marine reserves have had on four biological and two socioeconomic indicators (Table \ref{table:indicators}).
+Recall that reserves were implemented to protect lobster and other benthic invertebrates.
+However, we also contemplate fish species that are also targeted in the off-season and might receive collateral protection from the reserves.
 
 ```{r}
-tibble(Category = c("Biological", "","","","", "Socioeconomic", ""),
-           Indicador = c("Abundance", "Richness", "Shannon's diversity index", "Biomass", "Abundance of target species (lobsters)", "Income from target species", "Landings from target species")) %>% 
-  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex") %>% 
+tibble(Category = c("Biological", "Biological","Biological","Biological", "Socioeconomic", "Socioeconomic"),
+           Indicador = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
+       Units = c("org / $m^2$", "org / $m^2$", "org / $m^2$", "org / $m^2$", "M MXP", "Metric Tones")) %>% 
+  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
   kable_styling(latex_options = "HOLD_position")
@@
 Biological indicators are analyzed with a difference--in--differences analysis (Eq \ref{eqn:reg_bio}), which allows us to estimate the effect that the reserve has on the biological indicators by comparing trends across time and treatments \citep{moland_2013-VP,Villasenor-Derbez_2018}.
-The analysis is performed with generalized linear models of the form:
+The analysis is performed with a multiple linear regression of the form:
 
 \begin{equation}
-I_i = \alpha_{i} + \gamma_{it} Year_t + \beta Zone_i + \lambda_{it} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon
+I_{itj} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon
 \label{eqn:reg_bio}
@@
 When evaluating biomass and abundances, we include species--fixed effects ($\sigma_j$).
-For abundances and richness (*i.e.* count data) the model is estimated with a quasipoisson error distribution.
 
-Socioeconomic indicators are evaluated with a similar approach (Eq \ref{eqn:soc_reg}), where landings and income before and after the implementation of the reserve are compared:
+Socioeconomic indicators are evaluated with a similar approach (Eq \ref{eqn:soc_reg}).
+Due to data constrains, only Isla Natividad and Maria Elena are evaluated in this case.
+We constructed panel-data information with yearly (2001 - 2014) lobster landings and income of the studied and neighbouring communities that have similar management strategies, and belong to larger Cooperative Federtations \citep{mccay_2017-1m,ayer_2018}.
+Neighbouring communities are used as counterfacutals that allow us to control for unobserved time-invariants.
+Each "treated" community (Isla Natividad and Maria Elena) has three counterfactual communities.
 
 \begin{equation}
-I_i = \beta_0 + \beta_1Post
+I_i = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jComm_j +\epsilon
 \label{eqn:soc_reg}
@@
 
-This approach does not allow for a causal attribution of the observed changes to the reserve, but still allows us to draw important information that can inform our conclusions.
-For both approaches, model coefficients are estimated via ordinary least--squares and heteroskedastic--robust standard errors \citep{zeileis_2004-7n}.
+The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$) and $\sigma_jComm$ captures community-level fixed-effects.
+This approach allows for a causal attribution of the observed changes to the reserve.
+All model coefficients were estimated via ordinary least--squares and heteroskedastic--robust standard errors \citep{zeileis_2004-7n}.
+
+These regressions allows us to make a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}.
+The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the reserve site at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
 
@@
 
-Our methodological approach with biological indicators allows us to make a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}.
-The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the reserve site at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
-Here we present the effect that marine reserves had on each of the biological indicators for each coastal community, along with the trends in socioeconomic indicators of lobster catches and revenues.
-We also provide an overview of the state of the socioeconomic and governance settings of each community, and discuss how these dimensions might be intertwined with each other.
+The following sections present the effect that marine reserves had on each of the biological and socioeconomic indicators for each coastal community.
+Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* effect size).
+We also provide an overview of the governance settings of each community, and discuss how these dimensions might be related to the effectiveness of the reserves.
 
@@
 
-Effect sizes for biological indicators are shown in Figure \ref{fig:indicators}, and Figure \ref{fig:res_com} shows the summarized biological effects by community.
-Isla Natividad shows inconsistent effects across data sources (*i.e.* fish vs.
-invertebrates).
-For example, the reserve had a small effect on fish abundances (Fig \ref{fig:indicators}A), where only year 2010 showed significant effect sizes in fish abundances ($p<0.05$) and all other years oscillated above and under zero ($p > 0.05$).
-However, invertebrate abundances (Fig \ref{fig:indicators}B) presented a positive trend relative to the control site before implementation ($p < 0.05$) for all but 1 year (2008).
-Maria Elena and Punta Herrero showed no significant increase in fish and invertebrate abundances ($p< 0.05$), except for invertebrates in Punta Herrero for 2014 --right after the implementation of the reserves-- which showed a significant increase (*i.e.* $\lambda_{2014} = 2.5$, $p < 0.05$).
+All indicators show idiosyncratic responses through time for each community.
+Figure \ref{fig:indicators}A shows how lobster densities increase inside the reserve for Isla Natividad an Punta Herrero within the first years, but the effect is aroded through time.
+However, these effects are in the order of 0.2 extra organisms $m^{-2}$ and are not significantly different from zero ($p > 0.05$).
+The only case when lobster densities showed a significant positive effect ($p < 0.1$) was for Isla Natividad on the sixth year (*i.e.* 2012), a year after hypoxia events described in \citet{micheli_2012-EU} caused mass mortality of organisms.
+Biomass was only evaluated for fish data (Fig \ref{fig:indicators}B), and no changes were observed ($p > 0.1$), Punta Herrero even showed negative insignificant trends in the order of 0.01.
+Invertebrate and fish densities (\ref{fig:indicators}C-D) show similar patterns of lack of effectiveness with effect sizes oscillating around zero.
 Full tables with model coefficients are presented in the supplementary materials (**S1 Table**, **S2 Table**, **S3 Table**).
-
-While the number of fish species oscillated above and bellow zero through time for all reserves, none of these changes were statistically significant ($p > 0.05$) indicating that the reserves had no effect on fish species richness (Fig \ref{fig:indicators}C).
-For invertebrate species in Isla Natividad, all effect sizes were negative, but only significant for 2008, 2009, 2011, and 2014 ($p < 0.05$; Fig \ref{fig:indicators}D).
-For Maria Elena and Punta Herrero, the data do not show significant changes in invertebrate species richness ($p > 0.05$).
-
-Effect sizes for Shannon's diversity index for fish (Fig \ref{fig:indicators}E) in Isla Natividad oscillated between $\lambda_{2011} = -0.45$ and $\lambda_{2010} = -0.005$, but were not significantly different from null hypotheses of no change (*i.e.* $\lambda_t = 0$; $p > 0.05$).
-For invertebrates in that same community (Fig \ref{fig:indicators}F), Shannon's diversity index showed a significant decrease between 2008 and 2014, with largest decrease observed for 2011 ($\lambda_{2011} = -0.91$; $p < 0.05$).
-In the case of Maria Elena and Punta Herrero, Shannon's diversity index for fish showed increases in the order of $\lambda_t = 1$.
-For Maria Elena and Punta Herrero, these effects were only statistically significant for 2014, and 2014 and 2015 ($p < 0.05$).
-
-Biomass was only evaluated for fish data (Fig \ref{fig:indicators}G).
-In Isla Natividad, fish biomass presented a steady but small increase ($p>0.05$), and exhibited an increased variability in biomass between 2013 and 2016.
-Maria Elena and Punta Herrero also showed small, non-statistically significant increases in fish biomass ($p>0.05$).
-The last biological indicator is abundance of target species, *Panulirus interruptus* and *P.
-argus*, for the Pacific and Caribbean, respectively (Fig \ref{fig:indicators}H).
-Isla Natividad presented small constantly-positive effects but were not significantly different from the reference point of control site before the implementation of the reserve ($p > 0.05$.
-Maria Elena showed significant increases in lobster densities in the order of $\lambda_t = 10$ ($p < 0.05$).
-Finally, Punta Herrero presented alternating negative and positive effects, but these were not different from the baseline case ($p > 0.05$).
 
@@
 
-Lobster landings and revenue showed a increase after the implementation of the reserves for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}).
-However, the differences in catches and and revenue were not different in the periods before and after the implementation ($p > 0.05$) except for revenues in Isla Natividad, which presented a significant increase of 14.37 (M MXP; $p<0.05$).
-All regression coefficients are presented in **S4 Table**.
+Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}).
+For all years befor implementation, the effect sizes are close to zero, indicating that the control and treatment sites track each other well.
+However, effect sizes do not change after the implementation of the reserve.
+Again, the negative coeficient observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events.
+The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$).
+The trhee years of post-implementation data for Maria Elena do not show a significant effect of the reserve.
+Isla Natividad shows higher revenues after the iumplementation of the reserve, as compared to the control communities.
+However, these changes are also not ignificantly different and present an increased variation.
+All regression coefficients for each community and indicator are presented in **S4 Table**.
 
@@
 
-Although we have little information on the social dimension of these fisheries, using the SES framework indicators (Table \ref{table:gov_ind}), we can analyze the performance of each governance system with respect to MR enforcement (Table \ref{table:gov_res}).
+Although we have little information on the social dimension of these fisheries, using the SES framework indicators (Table \ref{table:gov_ind}), we can analyze the performance of each governance system (Table \ref{table:gov_res}).
+
+Our analysis shows that all of the systems analyzed share similarities in their Governance system which is based on cooperatives (GS5.2.3.2), with strong rules in use that include Operational rules (GS6.2), Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even Territorial use communal rights (GS6.1.4.3).
+However, we identified important differences in terms of the actors, resource systems and resource units.
+The value of lobster in isla natividad is higher than the lobster sold from Punta Herrero and Maria Elena (RU4), which can reduce the pressure on harvest.
+Lastly, in terms of actors, although all communities show a high level of leadership (A5), the level of trust (A6.1) is lower in Punta Herrero.
 In general, the presence and success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources they depend upon \citep{jupiter_2017}.
-The enabling conditions for conservation seem to be strongly present in Isla Natividad.
+The enabling conditions for conservation seem to be strongly present in all communities.
 Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperative.
-These conditions have favored the development of an efficient community based enforcement system.
-In contrast, the communities of Maria Elena and Punta Herrero are located near other fishing communities and cities.
-In Maria Elena, the fishing pressure caused by outsiders can be reduced by implementing a strong enforcement system (in water and land) supported by CSOs and the local government (CONANP).
-Lastly, the community of Punta Herrero shows the highest levels of illegal activities which can be attributed to its connectedness to other communities and the lack of appropriate technologies for enforcement.
+These conditions have favored the development of an efficient community-based enforcement systems.
 
@@
 
-```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for community-levell indicators.
-Plots are ordered by survey type (left: fish; right: invertebrates) and indicators: Abundance (A, B), Richness (C, D), Shannon's diversity index (E, F),fish biomass (G), and lobster (*Panulirus spp*) abundances (H).
+```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for lobster densities (*Panulirus spp*; A), fish biomass (B), invertebrate densities (C), and fish densities (D).
+Plots are ordered by survey type (left column: invertebrates; right column: fish).
 Points are jittered hotizontally to avoid overplotting.
-Points indicate the effect size, and errorbars are heteroskedastic-robust standard errors."}
+Points indicate the effect size, and errorbars standard errors.
+Years have been centered to year of implementation."}
 
@@
                   model_names = c("IN", "ME", "PH")) +
-  theme(legend.justification = c(1, 0),
-        legend.position = c(1, 0.55),
+  theme(legend.justification = c(1, 0.1),
+        legend.position = c(1, 0),
         legend.text = element_text(size = 8),
@@
 
-```{r, fig.height = 3, fig.width = 6, fig.cap = "\\label{fig:res_com}Summarized effects of the marne reserves by direction (positive - negative) and significance."}
-
-observe <- function(model){
-  model %>% 
-    robust_se() %>%
-    filter(grepl("zonaReserva:", term)) %>% 
-    mutate(Ano = as.numeric(gsub(term, pattern = "zonaReserva:year", replacement = "")),
-           ptest = ((p.value < 0.05)*-1)+0.05)
-}
-
-IN <- list(N_IN, Ni_IN, B_IN, L_IN) %>%
-  map_df(observe) %>%
-  as.tibble() %>% 
-  select(estimate, p.value) %>% 
-  mutate(Com = "Isla Natividad",
-         n_ind = nrow(.))
-
-ME <- list(N_ME, Ni_ME, B_ME, L_ME) %>%
-  map_df(.f = observe) %>%
-  as.tibble() %>% 
-  select(estimate, p.value) %>% 
-  mutate(Com = "Maria Elena",
-         n_ind = nrow(.))
-
-PH <- list(N_PH, Ni_PH, B_PH, L_PH) %>%
-  map_df(.f = observe) %>%
-  as.tibble() %>% 
-  select(estimate, p.value) %>% 
-  mutate(Com = "Punta Herrero",
-         n_ind = nrow(.))
-
-rbind(IN, ME, PH) %>% 
-  mutate(direction = case_when(estimate < 0 ~ "Negative",
-                               estimate > 0 ~ "Positive"),
-         Significance = case_when(p.value < 0.05 ~ "p < 0.05",
-                            p.value >= 0.05 ~ "p > 0.05")) %>%
-  group_by(Com, direction, Significance, n_ind) %>% 
-  summarize(A = n()) %>% 
-  ungroup() %>% 
-  mutate(A = A/n_ind,
-         A = ifelse(direction == "Negative", -A, A),
-         Significance = fct_relevel(Significance, "p > 0.05")) %>% 
-  ggplot(aes(x = Com, y = A, fill = Significance)) +
-  geom_col(color = "black", alpha = 0.8, width = 0.8) +
-  geom_hline(yintercept = 0, linetype = "dashed", size = 1) +
-  scale_fill_brewer(palette = "Set1", direction = -1) +
-  coord_flip() +
-  labs(x = "Community", y = "Percentage of total coefficients") +
-  scale_y_continuous(breaks = seq(-0.5, 1, by = 0.25), limits = c(-0.5, 1),
-                     labels = scales::percent) +
-  theme(legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8),
-        text = element_text(size = 12))
-```
-
 ```{r, results = "asis"}
@@
   filter(years_centered > -10) %>% 
-  mutate(years_centered = as.factor(years_centered)) %>% 
+  mutate(years_centered = as.factor(years_centered),
+         years_centered = fct_relevel(years_centered, "0")) %>% 
   filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
@@
   filter(years_centered > -10) %>% 
-  mutate(years_centered = as.factor(years_centered)) %>% 
+  mutate(years_centered = as.factor(years_centered),
+         years_centered = fct_relevel(years_centered, "0")) %>% 
   filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
@@
                         model_names = list("IN", "ME"),
-                        ylab = "Landings (K MT)")
+                        ylab = quo(lambda[t]))
 
@@
                         model_names = list("IN", "ME"),
-                        ylab = "Revenue (M MXP)")
+                        ylab = quo(lambda[t]))
 
-plot_grid(c_plot, v_plot, ncol = 1)
+plot_grid(c_plot, v_plot, ncol = 1, labels = "AUTO")
 
@@
 
-Our results show idiosyncratic biological effects of the reserves across communities and indicators.
-However, many of these effects were not statistically significant, indicating no effect of the reserve \ref{fig:res_com}.
-The socioeconomic indicators pertaining to landings and revenues associated to those landings showed little or no temporal change before and after reserve implementation.
-These contrasting effects, however, might be clarified when understanding the social-ecological context in which these communities and their reserves sit.
-In this section, we discuss potential shortcomings in our analysis, and provide plausible explanations to the observed biological and socioeconomic basing on previous literature and our analysis of the social-ecological system.
+theory tells us more than 5 years are needed, so it;s not surprise for PH and ME. But IN didnt make it either.
 
-The contrasting biological effectiveness observed is perhaps explained by our approach to evaluate the temporal and spatial changes of each indicator.
+Our results indicated idiosyncratic biological effects of the reserves across communities and indicators, with a combination of positive and negative effects.
+However, many of these effects were not statistically significant, indicating no effect of the reserve.
+The socioeconomic indicators pertaining to landings and revenues showed little or no change after reserve implementation.
+While lobster densities represent an indicator directly tied to reserve objectives, we also evaluate other biological indicators to test for additional effects and as a double-check.
+ll indicators presented a similar pattern.
+The lack of expected effectiveness poses the question, why do these communities continue to support the reserves?
+While not in the main scope of this paper, understanding the social-ecological context in which these communities and their reserves might provide insight to answer this question.
+Here, we discuss plausible explanations to lack of effectiveness observed based on literature and our social-ecological system analysis.
+We also discuss potential shortcomings in our analysis, and provide management recommendations to improve reserve effectiveness.
+
+Our approach to evaluate the temporal and spatial changes of each indicator provides a more robust measure of reserve effectiveness.
 Some works have solely focused on an inside-outside comparison of indicators \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD}, which do not address temporal variability \citep{depalma_2018}.
 Other works have compared the trend observed within a reserve through time \citep{betti_2017-lq}, which cannot distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}.
-By accounting for trends between sites and through times, we can control for time and space dynamics, and provide a better identification of the effect.
-However, it is worth looking deeper into each case, and identifying other plausible explanations.
+By accounting for trends between sites and through time, we can control for time and space dynamics, and provide a better identification of the effect
 
@@
 Isla Natividad has the oldest reserve, is fairly isolated, and has a well-established community-based enforcement system.
-While other communities are certainly within reach, these are known to be well organized fishing communities with successful resource management \citep{mccay_2017-1m,mccay_2014-CN}.
-The reserve at Isla Natividad presented the largest percentage of significantly positive changes in biological indicators (19%), but an important portion of was also negative (15%).
-With the age, relative isolation, and enforcement level of this reserve, it would be expected for it to be considerably effective.
-The potential gap in performance can be attributed to perturbations that do not distinguish reserve boundaries, such as environmental variability (**no recuerdo esta cita**).
-The region is known to be under the influence of recurrent hypoxia and high-temperature events known to cause massive adult mortalities \citep{micheli_2012-EU}.
+Neighbouring fishing communities are known to be well organized with successful resource management of their resources \citep{mccay_2017-1m,mccay_2014-CN}.
+Maria Elena and Punta Herrero are relatively young reserves (Table \ref{table:com_sum}).
+With the age, relative isolation, and enforcement level of these reserves, one would expect to observe effectiveness.
+However, another key feature of effective MRs is size \citep{edgar_2014-UO}; the lack of effectiveness is perhaps attributed to the reserves being too small \ref{table:com_sum}.
+Furthermore, perturbations that do not distinguish reserve boundaries, such as the environmental variability observed in Isla Natividad can also hinder effectiveness.
+The possibility of increasing reserve size or merging existing networks into a larger reserve should be evaluated.
 
-Maria Elena and Punta Herrero are relatively young reserves (See Table \ref{table:com_sum}).
-From these, the Maria Elena exhibited the highest performance in terms of biological indicators (15% significantly positive).
-In contrast, Punta Herrero had a similar proportion of positive and negative effects.
+Our analysis of landings and revenues does not identify detectable changes in these indicators.
+However, previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
+Abalone are sesile invertebrates with less mobility (compared to lobsters), and thus current reserve size might not be enough for lobster's range of mobility even when accounting for reserve age in Isla Natividad.
+Other community-based marine reserves in tropical ecosystems have taken up to six years to show a spillover effect \citep{dasilva_2015-zX}.
+Reserves in Maria Elena and Punta Herrero are relatively small and young, and may need more time for abundances to increase enough to export larvae or adult organisms.
 
-The way in which we measure changes in catches and revenues can not identify whether the observed differences are simply caused by pre-existing temporal trends or by the implementation of the reserve.
-Yet, there were no detectable changes in these indicators, except for landings in Isla Natividad.
-Other research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
-Since the trend was not detected in catches --directly related to abundance and fishing effort-- it is plausible that these differences are purely explained by an increase in market-level prices.
+A second explanation lies on the the mismatch of objectives during the design process.
+The Project description (*Estudios Técnicos Justificativos*) of each MR provides little information about the followed design guidelines.
+However, all reserves had to be approved by the community members.
+Depending on the communities’ capacity to look for longterm benefits, many communities may favor implementation of reserves on sites that represent a low fishing cost.
+As a consequence, the marine reserve have low impacts in reducng the fishing effort.
+Having small reserves in areas that do not compromise fishing profits might explain why this communities still support their reserves.
 
-The fact that there was no detectable change in catches for Maria Elena and Punta Herrero can be explained by a combination of factors related to the design, management, age, or ecological factors.
-Reserves in these communities are relatively small and young, and may need more time for lobster abundances to increase enough to export larvae or adult organisms.
-Other community-based marine reserves in tropical ecosystems have taken up to six years to show a spillover effect \citep{dasilva_2015-zX}.
-A complimentary explanation lies in the results observed for the governance system.
-The lack of enforcement in Punta Herrero, for example, could explain the lack of effectiveness observed in their reserves.
-
-**Limitations...**
-
-Our results show that community-based marine reserves can be effective if the environmental and social settings allow it.
-By studying the social-ecological system as a whole, we can provide a wider range of explanations to the patterns observed.
-It is interesting that even under the best enabling social conditions, climate variability can hinder the effect of a reserve --Although it is interesting to imagine what the state of that fishery had been if the reserve and organized cooperative were not present--.
-On the contrary, we show how under low climate variability, absence of proper governance structures can limit the effectiveness and benefits of a reserve.
-Whether the combination of a stable environment and governance structures are additive or multiplicative represents an interesting area for future research, especially under a changing climate.
-
-<!-- # Thoughts from GainesLab Meeting -->
-
-<!-- ## General thoughts -->
-
-<!-- A nice story to tell is that reserves need to have optimal social and environmental conditions to work.
-But they also need to be opttimally designed, accounting for size and location.
-It is worth mentioning that even when these reserves did not have nice effects on the ecological or fisheries side, they were a project that brought together the community, and also gave fishers access to a network of resources facilitated by the NGO.
-I still need to get data that support these.
--->
-
-<!-- Creo que hay que agregar informacion de la variabilidad ambienta.
-Posiblemente SST o algo similar.
-Ver si se puede obtener datos de oxigeno disuelto...
-Tal vez algo de ENSO / ONI / NAO o algo asi -->
-
-<!-- Quitar analisis de riqueza (la metodologia es restrictiva) y diversidad (aunque shannon provee une buena mezcla del cambio en abundancias y el cambio en riquezas) -->
-
-<!-- revisar biomasa! -->
-
-<!-- Revisar poisson o binomial negativa -->
-
-<!-- Mapa de El Manchon -->
-
-<!-- Las graficas y el analisis de Profit y Catches tienen que ser de langosta entera fresca solamente, y revisar que no haya habido un cambio de commodity a traves del tiempo para justificar usar solamente entera fresca...
-Tratar de obtener valores hasta 2017 -->
-
-<!-- Pedir datos biologicos hasta 2017 -->
-
-<!-- ## Comentarios del Lab -->
-
-<!-- Owen -->
-
-<!-- Are there confounding effects because reserves are so close to control sites? -->
-
-<!-- Dan -->
-
-<!-- Are the data good enough to answer these questions? -->
-
-<!-- Karly -->
-
-<!-- Think about other social indicators that do not depend on catches or revenue from these -->
-
-<!-- Alexa -->
-
-<!-- Daniel -->
-
-<!-- Sebas -->
-
-<!-- Look at other indicators of success (adaptive capacity) in the social / governance data -->
-
-<!-- Steve -->
-
-<!-- Would we expect to see effects in the first five years? (ME and PH) -->
-
-<!-- Are the designs good enough to see effects? -->
-
-<!-- Becky -->
-
-<!-- Ignacia -->
-
-<!-- Perhaps interesting follow-up projects include comparing perceptions vs. reality -->
-
-<!-- Estimating the benefits of being friends with an NGO -->
-
-<!-- This section may be divided by subheadings.
-Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.
--->
-
-<!-- Summary of main findings -->
-<!-- Limitations -->
-<!-- Conclusions -->
+Although our case studies fulfilled the social requirements for effective marine reserves ("high enforcement, presence of a management plan, fisher engagement in management, and promotion of sustainable fishing”; \citet{difranco_2016-Xw}), our results show that proper reserve design is crucial for effective marine reserves.
+The social-ecological systems framework allowed a systematic diganosis and compare all the different case studies, allowing us to identify and tease appart possible explanations \citep{basurto_2013-oq}.
 
@@
 
-```{r, eval = F}
-invert %>%
-  filter(generoespecie == "Panulirus interruptus",
-         abundancia < 100) %>% 
-  abundance(com = "Isla Natividad") %>%
-  mutate(ano = as.numeric(ano)) %>%
-  ggplot(aes(x = ano, y = indicador, color = zona)) +
-  geom_jitter(height = 0, alpha = 0.5) +
-  stat_summary(geom = "line", fun.y = mean) +
-  scale_color_brewer(palette = "Set1")
-```
-
 \paragraph*{S2 Figure}
@@
 ```{r, results = "asis", eval = F}
-stargazer::stargazer(list(N_IN, Ni_IN, S_IN, Si_IN, H_IN, Hi_IN, B_IN, L_IN),
-                     type = "latex",
+
+se <- list(robust_se_list(L_IN),
+           robust_se_list(L_ME),
+           robust_se_list(L_PH))
+
+stargazer::stargazer(list(L_IN, L_ME, L_PH),
+                     se = se,
+                     t.auto = T,
+                     p.auto = F,
+                     type = "html",
                      single.row = T,
-                     omit = "GeneroEspecie",
                      model.names = F,
-                     column.labels = c("Abundance (f)", "Abundance(i)", "Richness (f)", "Richness(i)", "Shannon (f)", "Shannon (i)", "Biomass", "Lobsters"),
+                     column.labels = c("IN", "ME", "PH"),
                      dep.var.labels.include = FALSE,
@@
                      digits = 2,
-                     intercept.bottom = F, title = "\\paragraph*{S1 Table}\\label{S1_Table}")
+                     intercept.bottom = F,
+                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
+                     out = "lobster_models.html")
 ```
@@
 ```{r, eval = F}
+se <- list(robust_se_list(B_IN),
+           robust_se_list(B_ME),
+           robust_se_list(B_PH))
 
-# Series de tiempo de langosta
-L_catches <- ggplot(L_conapesca,
-                    aes(x = Ano,
-                        y = PesoVivo/1000,
-                        color = Comunidad,
-                        group = Comunidad,
-                        fill = Comunidad)) +
-  geom_smooth(method = "lm") +
-  geom_point(aes(shape = Comunidad), size = 2, alpha = 0.5) +
-  scale_color_brewer(palette = "Set1") +
-  scale_fill_brewer(palette = "Set1") +
-  geom_vline(xintercept = c(2006, 2012, 2013),
-             color = c("red", "blue", "darkgreen"),
-             linetype = "dashed", size = 0.5) +
-  labs(x = "Year", y = "Catches (Tones)") +
-  theme(legend.position = "none",
-        text = element_text(size = 12))
+stargazer::stargazer(list(B_IN, B_ME, B_PH),
+                     se = se,
+                     t.auto = T,
+                     p.auto = F,
+                     type = "html",
+                     single.row = T,
+                     model.names = F,
+                     column.labels = c("IN", "ME", "PH"),
+                     dep.var.labels.include = FALSE,
+                     model.numbers = FALSE,
+                     digits = 2,
+                     intercept.bottom = F,
+                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
+                     out = "lobster_models.html")
+```
 
-L_value <- ggplot(L_conapesca, aes(x = Ano,
-                                   y = Valor/1e6,
-                                   color = Comunidad,
-                                   group = Comunidad,
-                                   fill = Comunidad)) +
-  geom_smooth(method = "lm") +
-  geom_point(aes(shape = Comunidad), size = 2, alpha = 0.5) +
-  scale_color_brewer(palette = "Set1") +
-  scale_fill_brewer(palette = "Set1") +
-  geom_vline(xintercept = c(2006, 2012, 2013),
-             color = c("red", "blue", "darkgreen"),
-             linetype = "dashed",
-             size = 0.5) +
-  labs(x = "Year", y = "Value (M MXP)") +
-  theme(legend.position = "none",
-        text = element_text(size = 12))
+```{r, eval = F}
+se <- list(robust_se_list(Ni_IN),
+           robust_se_list(Ni_ME),
+           robust_se_list(Ni_PH))
 
-cowplot::plot_grid(plotlist = list(L_catches, L_value),
-                   labels = "AUTO",
-                   ncol = 1)
+stargazer::stargazer(list(Ni_IN, Ni_ME, Ni_PH),
+                     se = se,
+                     t.auto = T,
+                     p.auto = F,
+                     type = "html",
+                     single.row = T,
+                     model.names = F,
+                     column.labels = c("IN", "ME", "PH"),
+                     dep.var.labels.include = FALSE,
+                     model.numbers = FALSE,
+                     digits = 2,
+                     intercept.bottom = F,
+                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
+                     out = "lobster_models.html")
+```
+
+```{r, eval = F}
+se <- list(robust_se_list(N_IN),
+           robust_se_list(N_ME),
+           robust_se_list(N_PH))
+
+stargazer::stargazer(list(N_IN, N_ME, N_PH),
+                     se = se,
+                     t.auto = T,
+                     p.auto = F,
+                     type = "html",
+                     single.row = T,
+                     model.names = F,
+                     column.labels = c("IN", "ME", "PH"),
+                     dep.var.labels.include = FALSE,
+                     model.numbers = FALSE,
+                     digits = 2,
+                     intercept.bottom = F,
+                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
+                     out = "lobster_models.html")
+```
+
+```{r, eval = F}
+se <- list(robust_se_list(C_IN),
+           robust_se_list(C_ME),
+           robust_se_list(V_IN),
+           robust_se_list(V_ME))
+
+stargazer::stargazer(list(C_IN, C_ME, V_IN, V_ME),
+                     se = se,
+                     t.auto = T,
+                     p.auto = F,
+                     type = "html",
+                     single.row = T,
+                     model.names = F,
+                     column.labels = c("Landings IN", "Landings ME", "Revenue IN", "Revenue ME"),
+                     dep.var.labels.include = FALSE,
+                     model.numbers = FALSE,
+                     digits = 2,
+                     intercept.bottom = F,
+                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
+                     out = "lobster_models.html")
 ```

### 2018-05-14 docs/paper/Villasenor-Derbez.tex
Updated analysis in manuscript. Must still update text!
Added new version of manuscript

@@
 
-\textbf{Last\_update}: 2018-04-10
+\textbf{Last\_update}: 2018-05-13
 
@@
 provide a holistic view of the social-ecological system
-\citep{lpezangarita_2014}. Here, we combine causal inference techniques
-\citep{depalma_2018} and the social-ecological systems framework
-\citep{ostrom_2009-hg} to provide a comprehensive ecological and
-socioeconomic evaluation of four community-based marine reserves in
-three coastal communities in Mexico.
+\citep{lpezangarita_2014}.
+
+While ecological factors (\emph{e.g.} habitat representation, initial
+state of protection, connectivity to other protected areas) can
+determine the success of a MR, their effectiveness also depends on the
+socioeconomic and governance settings under which they are implemented
+and managed. Literature shows that many non-ecological characteristics
+can play an equally important role in the effectiveness of MRs. For
+example, age of a reserve (\emph{i.e.} time since its implementation)
+and size were key to the effectiveness of MRs in Palau
+\citep{friedlander_2017-oI}. In the Mediterranean,
+\citet{difranco_2016-Xw} identify that surveillance and enforcement,
+presence of a management plan, and involvement of fishers in management
+and decision--making along with promotion of sustainable fishing
+practices were the key factors that increased stock health and income to
+fishers. At a global level, \citet{edgar_2014-UO} indicate that
+enforcement, age, size, and isolation were important factors determining
+effectiveness of the reserves.
 
@@
 community--based marine reserves \citep{uribe_2010-u2}, which are
-usually established within a Territorial Use Rights for Fisheries
-(TURFs); thus making them TURF--reserves \citep{afflerbach_2014-HP}.
-This bottom--up approach allows fishers to design their own reserves,
-which increases compliance and self--enforcement
+usually established within a TURFs; thus making them TURF--reserves
+\citep{afflerbach_2014-HP}. This bottom--up approach allows fishers to
+design their own reserves, which increases compliance and
+self--enforcement
 \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
@@
 legal recognition of a community--based reserve under the name of
-``Fishing Refugia'' (\emph{Zona de Refugio Pesquero}, FR). This new norm
-thus combines bottom--up approaches to design marine reserves, along
-with a legal recognition of the management intervention. Since then,
-\emph{39} FR have been implemented along the Pacific, Gulf of
+``Fishing Refugia'' (\emph{Zona de Refugio Pesquero}, FR). Since then,
+\textbf{39} FR have been implemented along the Pacific, Gulf of
 California, and Mexican Caribbean coastlines, but no formal evaluation
@@
 
-While there are ecological factors defining the success of a MR
-(\emph{i.e.} habitat representation, initial state of protection,
-connectivity to other protected areas), their effectiveness also depends
-on the socioeconomic and governance settings under which they are
-implemented. Literature shows that many non-ecological characteristics
-can play an equally important role in the effectiveness of MRs. For
-example, age of a reserve (\emph{i.e.} time since its implementation),
-size, and habitat contained were key to the effectiveness of MRs in
-Palau \citep{friedlander_2017-oI}. In the Mediterranean,
-\citet{difranco_2016-Xw} identify that surveillance and enforcement,
-presence of a management plan, and involvement of fishers in management
-and decision--making along with promotion of sustainable fishing
-practices were the key factors that increased stock health and income to
-fishers. At a global level, \citet{edgar_2014-UO} indicate that
-enforcement, age, size, and isolation were important factors determining
-effectiveness of the reserves.
-
-The objective of this work is twofold: i) Provide the first evaluation
-of community--based marine reserves in Mexico, and ii) provide a
+We combine causal inference techniques \citep{depalma_2018} and the
+social-ecological systems framework \citep{ostrom_2009-hg} to provide a
+comprehensive ecological and socioeconomic evaluation of community-based
+marine reserves in three coastal communities in Mexico. The objective of
+this work is twofold: i) Provide the first evaluation of
+community--based marine reserves in Mexico, and ii) provide a
 comprehensive evaluation of the social--ecological system to identify
 how socioeconomic and governance characteristics relate to ecological
-effectiveness. With the purpose of providing a holistic evaluation, we
-combine ecological, socioeconomic, and governance indicators. We use
-causal inference techniques to provide a measurement of the effect of
-the management intervention, and combine it with the social--ecological
-systems framework \citep{ostrom_2009-hg}.
+effectiveness.
 
@@
 We focus our evaluation in three coastal communities from the Pacific
-coast of Baja California (n = 1) and the Mesoamerican Reef System (n =
-2; Fig \ref{fig:map}). Isla Natividad (IN) lies west of the Baja
-California Peninsula (Fig \ref{fig:map}B), where kelp forests
-(\emph{Macrocystis pyrifera}) and rocky reefs are the predominant
-habitats. The island is home to a fishing cooperative (\emph{Sociedad
-Cooperative de Producción Pesquera Buzos y Pescadores de la Baja
-California SCL}), that holds a TURF for spiny lobster (\emph{Panulirus
-interruptus}). However, other resources like finfish (yellow-tail jack,
-\emph{Seriola lalandi}), sea cucumber (\emph{Parastichopus
-parvimensis}), red sea urchin (\emph{Mesocentrotus franciscanus}), snail
-(\emph{Megastraea turbanica} y \emph{M. undosa}), and abalone
-(\emph{Haliotis spp}, until 2010) are also important sources of income.
-In 2006, the community decided to implement two community--based marine
-reserves within their fishing grounds, seeking to recover depleted
-stocks of invertebrate species (mainly lobster and abalone). Until
-today, these reserves are yet to be legally recognized as Fishing
-Refugia.
+coast of Baja California and the Mesoamerican Reef System (Fig
+\ref{fig:map}). Fishing activities in all communities are assambled
+around fishing cooperatives that hold Territorial Use Rights for
+Fisheries (TURFs). Isla Natividad (IN) lies west of the Baja California
+Peninsula (Fig \ref{fig:map}B), where kelp forests (\emph{Macrocystis
+pyrifera}) and rocky reefs are the predominant habitats. The island is
+home to a the \emph{Sociedad Cooperative de Producción Pesquera (SCPP)
+Buzos y Pescadores de la Baja California}, spiny lobster
+(\emph{Panulirus interruptus}) represents an important fishery. However,
+other resources like finfish (yellow-tail jack, \emph{Seriola lalandi}),
+sea cucumber (\emph{Parastichopus parvimensis}), red sea urchin
+(\emph{Mesocentrotus franciscanus}), snail (\emph{Megastraea turbanica}
+y \emph{M. undosa}), and abalone (\emph{Haliotis spp}, until 2010) are
+also important sources of income. In 2006, the community decided to
+implement two community--based marine reserves within their fishing
+grounds, seeking to recover depleted stocks of invertebrate species
+(mainly lobster and abalone). Until today, these reserves are yet to be
+legally recognized as Fishing Refugia but count with full support from
+the community.
 
-The other two communities are Maria Elena (ME; Fig \ref{fig:map}C) and
-Punta Herrero (PH; Fig \ref{fig:map}D) in the Yucatan Peninsula, where
-coral reefs and mangroves are the representative coastal ecosystems. ME
-is a fishing camp --visited intermittently during the fishing season--
-belonging to the Cozumel fishing cooperative. PH is home to the ``José
-María Azcorra'' fishing cooperative. The main source of income to both
-communities is the Caribbean spiny lobster fishery (\emph{Panulirus
-argus}), which is carried out within their respective TURFs. These
-communities also target finfish in the off season, mainly snappers
-(Lutjanidae) and groupers (Serranidae). ME established eight marine
-reserves in 2012, and PH established four marine reserves in 2013. All
-these reserves are legally recognized as Fishing Refugia.
+The other two communities are Maria Elena (ME) and Punta Herrero (PH;
+Fig \ref{fig:map}C) in the Yucatan Peninsula, where coral reefs and
+mangroves are the representative coastal ecosystems. ME is a fishing
+camp --visited intermittently during the fishing season-- belonging to
+the Cozumel fishing cooperative (\emph{SCPP Cozumel}). PH is also home
+to a fishing cooperative (\emph{SCPP José María Azcorra}). The main
+source of income to both communities is the Caribbean spiny lobster
+(\emph{Panulirus argus}). These communities also target finfish in the
+off season, mainly snappers (Lutjanidae) and groupers (Serranidae). ME
+established eight marine reserves in 2012, and PH established four
+marine reserves in 2013 and an additional community-based (\emph{i.e.}
+not legally recognized). All these reserves are legally recognized as
+Fishing Refugia.
 
@@
 
-To perform the evaluation of these reserves we use thee sources of
-information. Ecological data come from the annual ecological monitoring
-of reserve and control areas, carried out by members from each community
-and personnel from the Mexican CSO ``Comunidad y Biodiversidad''
+We use three main sources of information to evaluate these reserves.
+Ecological data come from the annual ecological monitoring of reserve
+and control areas, carried out by members from each community and
+personnel from the Mexican CSO ``Comunidad y Biodiversidad''
 (\href{www.cobi.org.mx}{COBI}). These monitorings record richness and
 abundances of fish and invertebrate species in the reserves and control
-sites. For fish census, size structures are also collected to derive
-biomass. We define control sites as regions with habitat characteristics
-similar to the corresponding reserves, and that presumably had the same
-probability of being selected as reserves during the design phase. From
-all the reserves in these three communities, we use the ones that have
-data for reserve and control sites before and after the implementation
-of the reserve. This provides us with a Before-After-Control-Impact
-(\emph{i.e.} BACI) design that allows us to capture and control for
-temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI
-designs and causal inference techniques have proven effective to
+sites. Size structures are also collected in fish census. We define
+control sites as regions with habitat characteristics similar to the
+corresponding reserves, and that presumably had the same probability of
+being selected as reserves during the design phase. From all the
+reserves in these three communities, we use the ones that have data for
+reserve and control sites before and after the implementation of the
+reserve. This provides us with a Before-After-Control-Impact
+(\emph{i.e.} BACI) sampling design that allows us to capture and control
+for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
+BACI designs and causal inference techniques have proven effective to
 evaluate marine reserves, as they allow us to causally attribute
@@
 surveyed annually from at least one year before implementation until
-2016. Table \ref{table:com_sum} shows a summary of the number of
-reserves, year of implementation, and number of transects for each
-reserve.
+2016. Table \ref{table:com_sum} shows a summary of sampling effort and
+year of implementation of the analyzed reserves.
 
@@
 Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are
-located in the yucatan Peninsula.}
+located in the Yucatan Peninsula.}
 \end{figure}
@@
 
-\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of commuity--based marine reserves by community.
-Imp = Year of implementation, Start = Year of first sampling, number of fish transects in control (Cf) and reserve (Rf) sites, and number of invertebrate transects in Control (Ci) and Reserve (Ri) sites.}
+\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of commuity--based marine reserves by community.}
 \centering
-\begin{tabular}[t]{l|l|r|r|r|r|r|r}
+\begin{tabular}[t]{l|r|r|r|r}
 \hline
-Community & Reserve - Control & Imp & Start & Cf & Rf & Ci & Ri\\
+Community & TURF area ($km^2$) & Reserve area ($km^2$) & Percent as reserves & Year of implementation\\
 \hline
-Isla Natividad & La Plana / Las Cuevas - La Dulce / Babencho & 2006 & 2006 & 400 & 241 & 415 & 244\\
+Isla Natividad & 889.5 & 1.53 & 0.1720067 & 2006\\
 \hline
-Maria Elena & Cabezo - Cabezo (Control) & 2012 & 2012 & 44 & 45 & 27 & 21\\
+Maria Elena & 353.1 & 0.10 & 0.0283206 & 2012\\
 \hline
-Punta Herrero & El Faro - El Faro (Control) & 2013 & 2013 & 39 & 40 & 24 & 32\\
-\hline
-Punta Herrero & Manchon - Manchon (Control) & 2013 & 2012 & 43 & 45 & 27 & 41\\
+Punta Herrero & 299.7 & 0.43 & 0.1434768 & 2013\\
 \hline
@@
 
-Governance data were collected at the community-level. The information
-was compiled by combining key informants and the authors; experience and
-knowledge of the communities to collect the necessary information. These
-data contain information on the ecological system where the fishing
-activities develop, as well as the governance structures present in the
-cooperative. We also gathered information on the resource unit
-(\emph{i.e.} lobsters) and the relevant actors present in each community
-\citep{leslie_2015-na}.
+We use a set of governance indicators to analyze the governance
+structures of each cooperative \citep{leslie_2015-na}. The indicators
+(Table \ref{table:gov_ind}) focus on the resource system (four
+indicators), governance system (seven indicators), resource units (three
+indicators) and actors (three indicators). These were collected at the
+community--level from official documents used in the creation and
+designation of the marine reserves
+\citep{dof_website_2012,dof_website_2012} and based on the authors'
+experience and knowledge of the communities.
 
@@
 
-Following a framework that relates reserve objectives to performance
-indicators \citep{Villasenor-Derbez_2018}, we use five biological and
-two socioeconomic indicators to evaluate these marine reserves Table
-\ref{table:indicators}. We also use a set of governance indicators to
-analyze the governance structures of each cooperative
-\citep{leslie_2015-na}. The indicators (Table \ref{table:gov_ind}) focus
-on the resource system (four indicators), governance system (seven
-indicators), resource units (three indicators) and actors (three
-indicators).
+We evaluate the effect thar marine reserves have had on four biological
+and two socioeconomic indicators (Table \ref{table:indicators}). Recall
+that reserves were implemented to protect lobster and other benthic
+invertebrates. However, we also contemplate fish species that are also
+targeted in the off-season and might receive collateral protection from
+the reserves.
 
@@
 \centering
-\begin{tabular}[t]{l|l}
+\begin{tabular}[t]{l|l|l}
 \hline
-Category & Indicador\\
+Category & Indicador & Units\\
 \hline
-Biological & Abundance\\
+Biological & Lobster density & org / $m^2$\\
 \hline
- & Richness\\
+Biological & Invertebrate density & org / $m^2$\\
 \hline
- & Shannon's diversity index\\
+Biological & Fish biomass & org / $m^2$\\
 \hline
- & Biomass\\
+Biological & Fish density & org / $m^2$\\
 \hline
- & Abundance of target species (lobsters)\\
+Socioeconomic & Income from target species & M MXP\\
 \hline
-Socioeconomic & Income from target species\\
-\hline
- & Landings from target species\\
+Socioeconomic & Landings from target species & Metric Tones\\
 \hline
@@
 \citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis is performed
-with generalized linear models of the form:
+with a multiple linear regression of the form:
 
 \begin{equation}
-I_i = \alpha_{i} + \gamma_{it} Year_t + \beta Zone_i + \lambda_{it} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon
+I_{itj} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon
 \label{eqn:reg_bio}
@@
 \ref{table:com_sum}). When evaluating biomass and abundances, we include
-species--fixed effects (\(\sigma_j\)). For abundances and richness
-(\emph{i.e.} count data) the model is estimated with a quasipoisson
-error distribution.
+species--fixed effects (\(\sigma_j\)).
 
 Socioeconomic indicators are evaluated with a similar approach (Eq
-\ref{eqn:soc_reg}), where landings and income before and after the
-implementation of the reserve are compared:
+\ref{eqn:soc_reg}). Due to data constrains, only Isla Natividad and
+Maria Elena are evaluated in this case. We constructed panel-data
+information with yearly (2001 - 2014) lobster landings and income of the
+studied and neighbouring communities that have similar management
+strategies, and belong to larger Cooperative Federtations
+\citep{mccay_2017-1m,ayer_2018}. Neighbouring communities are used as
+counterfacutals that allow us to control for unobserved time-invariants.
+Each ``treated'' community (Isla Natividad and Maria Elena) has three
+counterfactual communities.
 
 \begin{equation}
-I_i = \beta_0 + \beta_1Post
+I_i = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jComm_j +\epsilon
 \label{eqn:soc_reg}
@@
 
-This approach does not allow for a causal attribution of the observed
-changes to the reserve, but still allows us to draw important
-information that can inform our conclusions. For both approaches, model
-coefficients are estimated via ordinary least--squares and
+The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in
+this case the \(Treated\) dummy variable indicates if the community has
+a reserve (\(Treated = 1\)) or not (\(Treated = 0\)) and
+\(\sigma_jComm\) captures community-level fixed-effects. This approach
+allows for a causal attribution of the observed changes to the reserve.
+All model coefficients were estimated via ordinary least--squares and
 heteroskedastic--robust standard errors \citep{zeileis_2004-7n}.
+
+These regressions allows us to make a causal link between the
+implementation of marine reserves and the observed trends by accounting
+for temporal and spatial dynamics \citep{depalma_2018}. The effect of
+the reserve is captured by the \(\lambda_t\) coefficient, and represents
+the difference observed between the control site before the
+implementation of the reserve and the reserve site at time \(t\) after
+controlling for other time and space variations (\emph{i.e.}
+\(\gamma_t\) and \(\beta\) respectively).
 
@@
 
-Our methodological approach with biological indicators allows us to make
-a causal link between the implementation of marine reserves and the
-observed trends by accounting for temporal and spatial dynamics
-\citep{depalma_2018}. The effect of the reserve is captured by the
-\(\lambda_t\) coefficient, and represents the difference observed
-between the control site before the implementation of the reserve and
-the reserve site at time \(t\) after controlling for other time and
-space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
-Here we present the effect that marine reserves had on each of the
-biological indicators for each coastal community, along with the trends
-in socioeconomic indicators of lobster catches and revenues. We also
-provide an overview of the state of the socioeconomic and governance
-settings of each community, and discuss how these dimensions might be
-intertwined with each other.
+The following sections present the effect that marine reserves had on
+each of the biological and socioeconomic indicators for each coastal
+community. Results are presented in terms of the difference through time
+and across sites, relative to the control site on the year of
+implementation (\emph{i.e.} effect size). We also provide an overview of
+the governance settings of each community, and discuss how these
+dimensions might be related to the effectiveness of the reserves.
 
@@
 
-Effect sizes for biological indicators are shown in Figure
-\ref{fig:indicators}, and Figure \ref{fig:res_com} shows the summarized
-biological effects by community. Isla Natividad shows inconsistent
-effects across data sources (\emph{i.e.} fish vs.~invertebrates). For
-example, the reserve had a small effect on fish abundances (Fig
-\ref{fig:indicators}A), where only year 2010 showed significant effect
-sizes in fish abundances (\(p<0.05\)) and all other years oscillated
-above and under zero (\(p > 0.05\)). However, invertebrate abundances
-(Fig \ref{fig:indicators}B) presented a positive trend relative to the
-control site before implementation (\(p < 0.05\)) for all but 1 year
-(2008). Maria Elena and Punta Herrero showed no significant increase in
-fish and invertebrate abundances (\(p< 0.05\)), except for invertebrates
-in Punta Herrero for 2014 --right after the implementation of the
-reserves-- which showed a significant increase (\emph{i.e.}
-\(\lambda_{2014} = 2.5\), \(p < 0.05\)). Full tables with model
-coefficients are presented in the supplementary materials (\textbf{S1
-Table}, \textbf{S2 Table}, \textbf{S3 Table}).
-
-While the number of fish species oscillated above and bellow zero
-through time for all reserves, none of these changes were statistically
-significant (\(p > 0.05\)) indicating that the reserves had no effect on
-fish species richness (Fig \ref{fig:indicators}C). For invertebrate
-species in Isla Natividad, all effect sizes were negative, but only
-significant for 2008, 2009, 2011, and 2014 (\(p < 0.05\); Fig
-\ref{fig:indicators}D). For Maria Elena and Punta Herrero, the data do
-not show significant changes in invertebrate species richness
-(\(p > 0.05\)).
-
-Effect sizes for Shannon's diversity index for fish (Fig
-\ref{fig:indicators}E) in Isla Natividad oscillated between
-\(\lambda_{2011} = -0.45\) and \(\lambda_{2010} = -0.005\), but were not
-significantly different from null hypotheses of no change (\emph{i.e.}
-\(\lambda_t = 0\); \(p > 0.05\)). For invertebrates in that same
-community (Fig \ref{fig:indicators}F), Shannon's diversity index showed
-a significant decrease between 2008 and 2014, with largest decrease
-observed for 2011 (\(\lambda_{2011} = -0.91\); \(p < 0.05\)). In the
-case of Maria Elena and Punta Herrero, Shannon's diversity index for
-fish showed increases in the order of \(\lambda_t = 1\). For Maria Elena
-and Punta Herrero, these effects were only statistically significant for
-2014, and 2014 and 2015 (\(p < 0.05\)).
-
-Biomass was only evaluated for fish data (Fig \ref{fig:indicators}G). In
-Isla Natividad, fish biomass presented a steady but small increase
-(\(p>0.05\)), and exhibited an increased variability in biomass between
-2013 and 2016. Maria Elena and Punta Herrero also showed small,
-non-statistically significant increases in fish biomass (\(p>0.05\)).
-The last biological indicator is abundance of target species,
-\emph{Panulirus interruptus} and \emph{P. argus}, for the Pacific and
-Caribbean, respectively (Fig \ref{fig:indicators}H). Isla Natividad
-presented small constantly-positive effects but were not significantly
-different from the reference point of control site before the
-implementation of the reserve (\(p > 0.05\). Maria Elena showed
-significant increases in lobster densities in the order of
-\(\lambda_t = 10\) (\(p < 0.05\)). Finally, Punta Herrero presented
-alternating negative and positive effects, but these were not different
-from the baseline case (\(p > 0.05\)).
+All indicators show idiosyncratic responses through time for each
+community. Figure \ref{fig:indicators}A shows how lobster densities
+increase inside the reserve for Isla Natividad an Punta Herrero within
+the first years, but the effect is aroded through time. However, these
+effects are in the order of 0.2 extra organisms \(m^{-2}\) and are not
+significantly different from zero (\(p > 0.05\)). The only case when
+lobster densities showed a significant positive effect (\(p < 0.1\)) was
+for Isla Natividad on the sixth year (\emph{i.e.} 2012), a year after
+hypoxia events described in \citet{micheli_2012-EU} caused mass
+mortality of organisms. Biomass was only evaluated for fish data (Fig
+\ref{fig:indicators}B), and no changes were observed (\(p > 0.1\)),
+Punta Herrero even showed negative insignificant trends in the order of
+0.01. Invertebrate and fish densities (\ref{fig:indicators}C-D) show
+similar patterns of lack of effectiveness with effect sizes oscillating
+around zero. Full tables with model coefficients are presented in the
+supplementary materials (\textbf{S1 Table}, \textbf{S2 Table},
+\textbf{S3 Table}).
 
@@
 
-Lobster landings and revenue showed a increase after the implementation
-of the reserves for Isla Natividad and Maria Elena (Fig
-\ref{fig:lobsters}). However, the differences in catches and and revenue
-were not different in the periods before and after the implementation
-(\(p > 0.05\)) except for revenues in Isla Natividad, which presented a
-significant increase of 14.37 (M MXP; \(p<0.05\)). All regression
-coefficients are presented in \textbf{S4 Table}.
+Lobster landings and revenue were only available for Isla Natividad and
+Maria Elena (Fig \ref{fig:lobsters}). For all years befor
+implementation, the effect sizes are close to zero, indicating that the
+control and treatment sites track each other well. However, effect sizes
+do not change after the implementation of the reserve. Again, the
+negative coeficient observed for Isla Natividad on year 5 correspond to
+the 2011 hypoxia events. The only positive change observed in lobster
+landings is for Isla Natividad in 2014 (\(p < 0.1\)). The trhee years of
+post-implementation data for Maria Elena do not show a significant
+effect of the reserve. Isla Natividad shows higher revenues after the
+iumplementation of the reserve, as compared to the control communities.
+However, these changes are also not ignificantly different and present
+an increased variation. All regression coefficients for each community
+and indicator are presented in \textbf{S4 Table}.
 
@@
 \ref{table:gov_ind}), we can analyze the performance of each governance
-system with respect to MR enforcement (Table \ref{table:gov_res}). In
-general, the presence and success of conservation initiatives depends on
-the incentives of local communities to maintain a healthy status of the
-resources they depend upon \citep{jupiter_2017}. The enabling conditions
-for conservation seem to be strongly present in Isla Natividad. Due to
-the clarity of access rights and isolation, the benefits of conservation
-directly benefit the members of the fishing cooperative. These
-conditions have favored the development of an efficient community based
-enforcement system. In contrast, the communities of Maria Elena and
-Punta Herrero are located near other fishing communities and cities. In
-Maria Elena, the fishing pressure caused by outsiders can be reduced by
-implementing a strong enforcement system (in water and land) supported
-by CSOs and the local government (CONANP). Lastly, the community of
-Punta Herrero shows the highest levels of illegal activities which can
-be attributed to its connectedness to other communities and the lack of
-appropriate technologies for enforcement.
+system (Table \ref{table:gov_res}).
+
+Our analysis shows that all of the systems analyzed share similarities
+in their Governance system which is based on cooperatives (GS5.2.3.2),
+with strong rules in use that include Operational rules (GS6.2),
+Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even
+Territorial use communal rights (GS6.1.4.3). However, we identified
+important differences in terms of the actors, resource systems and
+resource units. The value of lobster in isla natividad is higher than
+the lobster sold from Punta Herrero and Maria Elena (RU4), which can
+reduce the pressure on harvest. Lastly, in terms of actors, although all
+communities show a high level of leadership (A5), the level of trust
+(A6.1) is lower in Punta Herrero. In general, the presence and success
+of conservation initiatives depends on the incentives of local
+communities to maintain a healthy status of the resources they depend
+upon \citep{jupiter_2017}. The enabling conditions for conservation seem
+to be strongly present in all communities. Due to the clarity of access
+rights and isolation, the benefits of conservation directly benefit the
+members of the fishing cooperative. These conditions have favored the
+development of an efficient community-based enforcement systems.
 
@@
 marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
-blue triangles), and Punta Herrero (PH; green squares) for
-community-levell indicators. Plots are ordered by survey type (left:
-fish; right: invertebrates) and indicators: Abundance (A, B), Richness
-(C, D), Shannon's diversity index (E, F),fish biomass (G), and lobster
-(\emph{Panulirus spp}) abundances (H). Points are jittered hotizontally
-to avoid overplotting. Points indicate the effect size, and errorbars
-are heteroskedastic-robust standard errors.}
+blue triangles), and Punta Herrero (PH; green squares) for lobster
+densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
+densities (C), and fish densities (D). Plots are ordered by survey type
+(left column: invertebrates; right column: fish). Points are jittered
+hotizontally to avoid overplotting. Points indicate the effect size, and
+errorbars standard errors. Years have been centered to year of
+implementation.}
 \end{figure}
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
-\caption{\label{fig:unnamed-chunk-6}\label{fig:res_com}Summarized effects of
-the marne reserves by direction (positive - negative) and significance.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{\label{fig:unnamed-chunk-8}\label{fig:lobsters}Effect sizes for
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-7-1.pdf}
+\caption{\label{fig:unnamed-chunk-7}\label{fig:lobsters}Effect sizes for
 lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
@@
 
-\caption{\label{tab:unnamed-chunk-9}\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \citep{mcginnis_2014}.}
+\caption{\label{tab:unnamed-chunk-8}\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \citep{mcginnis_2014}.}
 \centering
@@
 
-Our results show idiosyncratic biological effects of the reserves across
-communities and indicators. However, many of these effects were not
-statistically significant, indicating no effect of the reserve
-\ref{fig:res_com}. The socioeconomic indicators pertaining to landings
-and revenues associated to those landings showed little or no temporal
-change before and after reserve implementation. These contrasting
-effects, however, might be clarified when understanding the
-social-ecological context in which these communities and their reserves
-sit. In this section, we discuss potential shortcomings in our analysis,
-and provide plausible explanations to the observed biological and
-socioeconomic basing on previous literature and our analysis of the
-social-ecological system.
+theory tells us more than 5 years are needed, so it;s not surprise for
+PH and ME. But IN didnt make it either.
 
-The contrasting biological effectiveness observed is perhaps explained
-by our approach to evaluate the temporal and spatial changes of each
-indicator. Some works have solely focused on an inside-outside
-comparison of indicators
+Our results indicated idiosyncratic biological effects of the reserves
+across communities and indicators, with a combination of positive and
+negative effects. However, many of these effects were not statistically
+significant, indicating no effect of the reserve. The socioeconomic
+indicators pertaining to landings and revenues showed little or no
+change after reserve implementation. While lobster densities represent
+an indicator directly tied to reserve objectives, we also evaluate other
+biological indicators to test for additional effects and as a
+double-check. ll indicators presented a similar pattern. The lack of
+expected effectiveness poses the question, why do these communities
+continue to support the reserves? While not in the main scope of this
+paper, understanding the social-ecological context in which these
+communities and their reserves might provide insight to answer this
+question. Here, we discuss plausible explanations to lack of
+effectiveness observed based on literature and our social-ecological
+system analysis. We also discuss potential shortcomings in our analysis,
+and provide management recommendations to improve reserve effectiveness.
+
+Our approach to evaluate the temporal and spatial changes of each
+indicator provides a more robust measure of reserve effectiveness. Some
+works have solely focused on an inside-outside comparison of indicators
 \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD}, which do
@@
 trends in a reserve and the entire system \citep{depalma_2018}. By
-accounting for trends between sites and through times, we can control
-for time and space dynamics, and provide a better identification of the
-effect. However, it is worth looking deeper into each case, and
-identifying other plausible explanations.
+accounting for trends between sites and through time, we can control for
+time and space dynamics, and provide a better identification of the
+effect
 
@@
 has the oldest reserve, is fairly isolated, and has a well-established
-community-based enforcement system. While other communities are
-certainly within reach, these are known to be well organized fishing
-communities with successful resource management
-\citep{mccay_2017-1m,mccay_2014-CN}. The reserve at Isla Natividad
-presented the largest percentage of significantly positive changes in
-biological indicators (19\%), but an important portion of was also
-negative (15\%). With the age, relative isolation, and enforcement level
-of this reserve, it would be expected for it to be considerably
-effective. The potential gap in performance can be attributed to
-perturbations that do not distinguish reserve boundaries, such as
-environmental variability (\textbf{no recuerdo esta cita}). The region
-is known to be under the influence of recurrent hypoxia and
-high-temperature events known to cause massive adult mortalities
-\citep{micheli_2012-EU}.
+community-based enforcement system. Neighbouring fishing communities are
+known to be well organized with successful resource management of their
+resources \citep{mccay_2017-1m,mccay_2014-CN}. Maria Elena and Punta
+Herrero are relatively young reserves (Table \ref{table:com_sum}). With
+the age, relative isolation, and enforcement level of these reserves,
+one would expect to observe effectiveness. However, another key feature
+of effective MRs is size \citep{edgar_2014-UO}; the lack of
+effectiveness is perhaps attributed to the reserves being too small
+\ref{table:com_sum}. Furthermore, perturbations that do not distinguish
+reserve boundaries, such as the environmental variability observed in
+Isla Natividad can also hinder effectiveness. The possibility of
+increasing reserve size or merging existing networks into a larger
+reserve should be evaluated.
 
-Maria Elena and Punta Herrero are relatively young reserves (See Table
-\ref{table:com_sum}). From these, the Maria Elena exhibited the highest
-performance in terms of biological indicators (15\% significantly
-positive). In contrast, Punta Herrero had a similar proportion of
-positive and negative effects.
+Our analysis of landings and revenues does not identify detectable
+changes in these indicators. However, previous research has shown that
+reserves in Isla Natividad yield fishery benefits for the abalone
+fishery \citep{rossetto_2015-V0}. Abalone are sesile invertebrates with
+less mobility (compared to lobsters), and thus current reserve size
+might not be enough for lobster's range of mobility even when accounting
+for reserve age in Isla Natividad. Other community-based marine reserves
+in tropical ecosystems have taken up to six years to show a spillover
+effect \citep{dasilva_2015-zX}. Reserves in Maria Elena and Punta
+Herrero are relatively small and young, and may need more time for
+abundances to increase enough to export larvae or adult organisms.
 
-The way in which we measure changes in catches and revenues can not
-identify whether the observed differences are simply caused by
-pre-existing temporal trends or by the implementation of the reserve.
-Yet, there were no detectable changes in these indicators, except for
-landings in Isla Natividad. Other research has shown that reserves in
-Isla Natividad yield fishery benefits for the abalone fishery
-\citep{rossetto_2015-V0}. Since the trend was not detected in catches
---directly related to abundance and fishing effort-- it is plausible
-that these differences are purely explained by an increase in
-market-level prices.
+A second explanation lies on the the mismatch of objectives during the
+design process. The Project description (\emph{Estudios Técnicos
+Justificativos}) of each MR provides little information about the
+followed design guidelines. However, all reserves had to be approved by
+the community members. Depending on the communities' capacity to look
+for longterm benefits, many communities may favor implementation of
+reserves on sites that represent a low fishing cost. As a consequence,
+the marine reserve have low impacts in reducng the fishing effort.
+Having small reserves in areas that do not compromise fishing profits
+might explain why this communities still support their reserves.
 
-The fact that there was no detectable change in catches for Maria Elena
-and Punta Herrero can be explained by a combination of factors related
-to the design, management, age, or ecological factors. Reserves in these
-communities are relatively small and young, and may need more time for
-lobster abundances to increase enough to export larvae or adult
-organisms. Other community-based marine reserves in tropical ecosystems
-have taken up to six years to show a spillover effect
-\citep{dasilva_2015-zX}. A complimentary explanation lies in the results
-observed for the governance system. The lack of enforcement in Punta
-Herrero, for example, could explain the lack of effectiveness observed
-in their reserves.
-
-\textbf{Limitations\ldots{}}
-
-Our results show that community-based marine reserves can be effective
-if the environmental and social settings allow it. By studying the
-social-ecological system as a whole, we can provide a wider range of
-explanations to the patterns observed. It is interesting that even under
-the best enabling social conditions, climate variability can hinder the
-effect of a reserve --Although it is interesting to imagine what the
-state of that fishery had been if the reserve and organized cooperative
-were not present--. On the contrary, we show how under low climate
-variability, absence of proper governance structures can limit the
-effectiveness and benefits of a reserve. Whether the combination of a
-stable environment and governance structures are additive or
-multiplicative represents an interesting area for future research,
-especially under a changing climate.
+Although our case studies fulfilled the social requirements for
+effective marine reserves (``high enforcement, presence of a management
+plan, fisher engagement in management, and promotion of sustainable
+fishing''; \citet{difranco_2016-Xw}), our results show that proper
+reserve design is crucial for effective marine reserves. The
+social-ecological systems framework allowed a systematic diganosis and
+compare all the different case studies, allowing us to identify and
+tease appart possible explanations \citep{basurto_2013-oq}.
 

### 2018-06-08 docs/paper/S1Table.Rmd
Made changes to manuscript body, just missing details on discussion

---
output: pdf_document
---

Update this table and fill with information from ERE's table

```{r setup, echo = FALSE}
library(magrittr)

read.csv(here::here("data", "gov_ind.csv")) %>% 
  knitr::kable(caption = "\\label{table:gov_ind} Indicators used to describe the social-ecological system", col.names = c("Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero"), format = "latex") %>% 
  kableExtra::column_spec(1, width = "10em") %>% 
  kableExtra::column_spec(2, width = "8em") %>% 
  kableExtra::column_spec(3, width = "8em") %>% 
  kableExtra::column_spec(4, width = "8em") %>% 
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(6 , bold = T) %>% 
  kableExtra::row_spec(14 , bold = T) %>% 
  kableExtra::row_spec(18 , bold = T)
```

### 2018-06-08 docs/paper/S5Table.Rmd
Made changes to manuscript body, just missing details on discussion

---
output: pdf_document
---

Update this table and fill with information from ERE's table

```{r, echo = F}
library(magrittr)
read.csv(here::here("data","gov_res.csv")) %>% 
  knitr::kable(caption = "\\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \\citep{mcginnis_2014}.",
               col.names = c("", "Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  kableExtra::column_spec(1, width = "10em") %>%
  kableExtra::column_spec(2, width = "10em") %>%
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(7 , bold = T) %>% 
  kableExtra::row_spec(15 , bold = T) %>% 
  kableExtra::row_spec(19 , bold = T)
```

### 2018-06-08 docs/paper/S5Table.tex
Made changes to manuscript body, just missing details on discussion

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\begin{document}

Update this table and fill with information from ERE's table

\begin{verbatim}
## Warning in kableExtra::column_spec(., 1, width = "10em"): Please specify
## format in kable. kableExtra can customize either HTML or LaTeX outputs. See
## https://haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::column_spec(., 2, width = "10em"): Please specify
## format in kable. kableExtra can customize either HTML or LaTeX outputs. See
## https://haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::row_spec(., 1, bold = T): Please specify format in
## kable. kableExtra can customize either HTML or LaTeX outputs. See https://
## haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::row_spec(., 7, bold = T): Please specify format in
## kable. kableExtra can customize either HTML or LaTeX outputs. See https://
## haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::row_spec(., 15, bold = T): Please specify format in
## kable. kableExtra can customize either HTML or LaTeX outputs. See https://
## haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{verbatim}
## Warning in kableExtra::row_spec(., 19, bold = T): Please specify format in
## kable. kableExtra can customize either HTML or LaTeX outputs. See https://
## haozhu233.github.io/kableExtra/ for details.
\end{verbatim}

\begin{longtable}[]{@{}lllll@{}}
\caption{\label{table:gov_res}Analysis of the fishing cooperatives based
on the Social-Ecological systems framework
\citep{mcginnis_2014}.}\tabularnewline
\toprule
& Indicator & Isla Natividad & Maria Elena & Punta
Herrero\tabularnewline
\midrule
\endfirsthead
\toprule
& Indicator & Isla Natividad & Maria Elena & Punta
Herrero\tabularnewline
\midrule
\endhead
Resource systems (RS) & & & &\tabularnewline
RS2 -- Clarity of system boundaries & TURF presence & High & High &
High\tabularnewline
RS3 -- Size of resource system & & & &\tabularnewline
RS5 -- Productivity of system & Type of ecosystem & High & High &
High\tabularnewline
RS7 -- Predictability of system dynamics & Intensity of environmental
disturbance & Low (ENSO) & High & High\tabularnewline
RS9 -- Location & Proximity to other communities/cities & Isolated & Not
Isolated & Not Isolated\tabularnewline
Governance systems (GS) & & & &\tabularnewline
GS1 -- Government organizations & Presence of fishing cooperatives & Yes
& Yes & Yes\tabularnewline
GS2 -- Nongovernment organizations & Involved actors & Yes & Yes &
Yes\tabularnewline
GS3 -- Network structure & Presence of an inter-cooperative structure &
Yes & No & No\tabularnewline
GS4 -- Property-rights systems & TURF presence & Yes & Yes &
Yes\tabularnewline
GS5 -- Operational-choice rules & Fishing Regulations / MPA enforcement
/ Enforcement technolofy & Yes & Yes & Yes\tabularnewline
GS6 -- Collective-choice rules & Cooperative regulations & Yes & Yes &
Yes\tabularnewline
GS7 -- Constitutional-choice rules & & & &\tabularnewline
Resource units (RU) & & & &\tabularnewline
RU1 -- Resource unit mobility & Targeted species home range & Low &
Medium & Medium\tabularnewline
RU2 -- Growth or replacement rate & Max age of targeted species & Low &
Medium & Medium\tabularnewline
RU4 -- Economic value & Price of targeted species & high & High &
high\tabularnewline
Actors (A) & & & &\tabularnewline
A1 -- Number of relevant actors & & 98 & &\tabularnewline
A2 -- Socioeconomic attributes & & & &\tabularnewline
A5 -- Leadership/entrepreneurship & Leadership & High & High &
High\tabularnewline
A6 -- Norms (trust-reciprocity)/social capital--- (Based on illegal
fishing) & Level of illegal fishing & High & High & Low\tabularnewline
A8 -- Importance of resource (dependence) & Presence of alternative
livelihoods & High & High & High\tabularnewline
\bottomrule
\end{longtable}

\end{document}

### 2018-06-08 docs/paper/Villasenor-Derbez.Rmd
Made changes to manuscript body, just missing details on discussion

@@
 ---
-title: "Enabling conditions for effective community-based marine reserves in small-scale fisheries"
-rtitle: "Marine reserve effectiveness"
+title: "Management and effectiveness of community-based marine reserves in small-scale fisheries"
+rtitle: "Community-based marine reserves"
 author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Álvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
@@
 email: "jvillasenor@bren.ucsb.edu"
-abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
+abstract: "Al finalizar revisiones"
+keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, Mexico, Social-Ecological Systems"
+output:
+  bookdown::pdf_document2:
+    keep_tex: yes
+    template: template.tex
+---
+
+<!-- Old Abstract -->
+<!-- Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
 However, overfishing and unsustainable fishing practices threaten the marine environment and jeopardize the wellbeing of coastal communities.
@@
 Thus, even when a community is well organized (and reserves are well enforced), environmental variation can hinder the benefits of a reserve, and vice versa.
-Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making."
-keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, Mexico, Social-Ecological Systems"
-output:
-  bookdown::pdf_document2:
-    keep_tex: yes
-    template: template.tex
----
+Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making.
+-->
 
@@
 
-**Last_update**: `r Sys.Date()`
+**Titulos alternativos**
+
+"Community-based marine reserves in small-scale fisheries"
+
+"Effectiveness of community-based marine reserves: lessons for their management and implementation" (no me encanta)
+
+Algun otro?
+
+**Last update**: `r Sys.Date()`
 
@@
 Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}.
-A common approach to manage the spatial distribution of fishing effort and recover stocks is through the implementation of marine reserves (*i.e.* areas where all fishing activities are off--limits; MRs) \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}.
+A common approach to manage the spatial distribution of fishing effort to recover stocks and preserve biodiversity is through the implementation of marine reserves (MRs).
+These areas allow bounded populations to recover by limiting all extractive activities \citep{halpern_2002}.
+The science of MRs has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
+Modelling studies show that the fishery benefits of marine reserves depend on initial stock status and the type of management under which the fishery operates \citep{hilborn_2006}, as well as reserve size and the amount of larvae exported from these \citep{krueck_2017-J1}.
+Other research has focused on the relationship between socioeconomic and governance structures and their relationship to ecological effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
+However, few studies simultaneously evaluate MRs from all these perspectives (*e.g.* \citet{lpezangarita_2014}).
 
-Marine reserve science has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, climate change mitigation \citep{roberts_2017-J9}, and protection from environmental variability \citep{micheli_2012-EU}.
-However, there is considerably less literature focusing on the relationship between socioeconomic and governance structures and their relationship to ecological effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or benefits to fisheries \citep{krueck_2017-J1}; evaluations of marine reserves rarely provide a holistic view of the social-ecological system \citep{lpezangarita_2014}.
-
-While ecological factors (*e.g.* habitat representation, initial state of protection, connectivity to other protected areas) can determine the success of a MR, their effectiveness also depends on the socioeconomic and governance settings under which they are implemented and managed.
+While ecological factors like habitat representation within the MR or connectivity to other areas can determine the success of a MR, its effectiveness also depends on the socioeconomic and governance settings under which they are implemented and managed.
 Literature shows that many non-ecological characteristics can play an equally important role in the effectiveness of MRs.
@@
 In the Mediterranean, \citet{difranco_2016-Xw} identify that surveillance and enforcement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key factors that increased stock health and income to fishers.
-At a global level, \citet{edgar_2014-UO} indicate that enforcement, age, size, and isolation were important factors determining effectiveness of the reserves.
+At a global level, enforcement, age, size, and isolation are important factors that determine the effectiveness of the reserves \citep{edgar_2014-UO}.
+However, MRs represent part of a more complex social-ecological systems governed by the interaction between humans and the environment.
+In this work we evaluate community-based marine reserves in Mexico using a triple bottom line approach by using ecological, socioeconomic, and governance indicators that provide a holistic view of the reserves \citep{halpern_2013}.
 
-Marine Reserves in Mexico have been commonly implemented as "core zones" within Biosphere Reserves that are administered by the National Commission of Protected Areas (*Comisión Nacional de Áreas Marinas Protegidas*, CONANP).
-While CONANP has made efforts to have a participatory process, the implementation of these zones is still characterized by top--down approaches.
-This motivated Civil Society Organizations (CSOs) to work with coastal communities to implement community--based marine reserves \citep{uribe_2010-u2}, which are usually established within a TURFs; thus making them TURF--reserves \citep{afflerbach_2014-HP}.
-This bottom--up approach allows fishers to design their own reserves, which  increases compliance and self--enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
-However, these reserves still lack legal recognition, making them vulnerable to poaching.
-In 2014, a new norm \citep{nom} allowed fishers to request the legal recognition of a community--based reserve under the name of "Fishing Refugia" (*Zona de Refugio Pesquero*, FR).
-Since then, **39** FR have been implemented along the Pacific, Gulf of California, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.
+There are three main approaches to implement MRs, each of which have its pros and cons.
+We describe them in the context of Mexican MRs, but argue that at least the first two apply elsewhere.
+MRs were historically implemented and managed by a government agency, in this case the National Commission of Protected Areas (*Comisión Nacional de Áreas Marinas Protegidas*, CONANP).
+While CONANP has made efforts to have a participatory process, the implementation of these zones is still mainly a top-down process.
+A second approach is the implementation of community-based marine reserves within areas of exclusive access (*i.e.* TURFs), thus making them TURF-reserves \citep{afflerbach_2014-HP}.
+Community-based spatial closures ocurr in other places, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
+This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
+However, these lack legal recognition and rely on the exclusive access granted by the TURF, making it impossible to enforce without one.
+In an effort to provide a legal framework for these reserves, Civil Society Organizations (CSOs) served as the bridge between fishers and government.
+In 2014, a new norm \citep{nom} allowed fishers to request the legal recognition of a community-based reserve under the name of "Fishing Refugia" (*Zona de Refugio Pesquero*).
+These can be implemented as temporal or partial eservs, which can protect one, some, or all resources within them.
+Since then, **39** of these have been implemented along the Pacific, Gulf of California, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.
 
-We combine causal inference techniques \citep{depalma_2018} and the social-ecological systems framework \citep{ostrom_2009-hg} to provide a comprehensive ecological and socioeconomic evaluation of community-based marine reserves in three coastal communities in Mexico.
-The objective of this work is twofold: i) Provide the first evaluation of community--based marine reserves in Mexico, and ii) provide a comprehensive evaluation of the social--ecological system to identify how socioeconomic and governance characteristics relate to ecological effectiveness.
+This work combines causal inference techniques and the social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico.
+The objective of this work is twofold.
+First, provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries.
+And second, perform the first formal evaluation of Fishing Refugia in Mexico and identify areas where improvement or adjustment might result in increased effectiveness.
+On both cases, we draw from the lessons learned and provide management recommendations to ensure or improve the effectiveness of community-based marine reserves.
 
@@
 
-We focus our evaluation in three coastal communities from the Pacific coast of Baja California and the Mesoamerican Reef System (Fig \ref{fig:map}).
-Fishing activities in all communities are assambled around fishing cooperatives that hold Territorial Use Rights for Fisheries (TURFs).
-Isla Natividad (IN) lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests (*Macrocystis pyrifera*) and rocky reefs are the predominant habitats.
-The island is home to a the *Sociedad Cooperative de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California*, spiny lobster (*Panulirus interruptus*) represents an important fishery.
+We evaluate community-based marine reserves from three coastal communities located in the Pacific coast of Baja California and the Mexican Caribbean (Fig \ref{fig:map}).
+All communities are organized as fishing cooperatives that hold Territorial Use Rights for Fisheries (TURFs).
+Isla Natividad (IN) lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests and rocky reefs are the predominant habitats.
+The island is home to the *Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California*, whose main resource by value is the spiny lobster (*Panulirus interruptus*).
 However, other resources like finfish (yellow-tail jack, *Seriola lalandi*), sea cucumber (*Parastichopus parvimensis*), red sea urchin (*Mesocentrotus franciscanus*), snail (*Megastraea turbanica* y *M.
-undosa*), and abalone (*Haliotis spp*, until 2010) are also important sources of income.
-In 2006, the community decided to implement two community--based marine reserves within their fishing grounds, seeking to recover depleted stocks of invertebrate species (mainly lobster and abalone).
+undosa*), and abalone (*Haliotis spp*) are also important sources of income.
+In 2006, the community decided to implement two community-based marine reserves within their fishing grounds to protect commercially important invertebrate species (mainly lobster and abalone).
 Until today, these reserves are yet to be legally recognized as Fishing Refugia but count with full support from the community.
@@
 The other two communities are Maria Elena (ME) and Punta Herrero (PH; Fig \ref{fig:map}C) in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems.
-ME is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*).
-PH is also home to a fishing cooperative (*SCPP José María Azcorra*).
-The main source of income to both communities is the Caribbean spiny lobster (*Panulirus argus*).
-These communities also target finfish in the off season, mainly snappers (Lutjanidae) and groupers (Serranidae).
+ME is a fishing camp --visited intermitently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); PH is home to the *SCPP José María Azcorra* cooperative.
+Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off season, mainly snappers (Lutjanidae) and groupers (Serranidae).
 ME established eight marine reserves in 2012, and PH established four marine reserves in 2013 and an additional community-based (*i.e.* not legally recognized).
@@
 We use three main sources of information to evaluate these reserves.
-Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO "Comunidad y Biodiversidad" ([COBI](www.cobi.org.mx)).
-These monitorings record richness and abundances of fish and invertebrate species in the reserves and control sites.
-Size structures are also collected in fish census.
-We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had the same probability of being selected as reserves during the design phase.
-From all the reserves in these three communities, we use the ones that have data for reserve and control sites before and after the implementation of the reserve.
+Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
+Trained divers record richness and abundances of fish and invertebrate species in the reserves and control sites.
+Size structures are also collected for fish evaluations.
+We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had a similar probability of being selected as reserves during the design phase.
+We focus our evaluation for sites where data are avaialable for reserve and control sites, before and after the implementation of the reserve.
 This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
 BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
-All reserves were surveyed annually from at least one year before implementation until 2016.
-Table \ref{table:com_sum} shows a summary of sampling effort and year of implementation of the analyzed reserves.
+All sites were surveyed annually, and at least once before implementation of the reserves.
+Table \ref{table:com_sum} shows a summary of the reserves included in this study.
 
@@
 ```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A).
-Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are located in the Yucatan Peninsula."}
+Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are located in the Yucatan Peninsula.
+Blue polygons represent the TURFs, and red polygons the marine reserves."}
 # Define reference map
@@
 Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA).
-Data contain monthly lobster landings (Kg) and value (MXP) from 2000 to 2014.
-This information was aggregated by year, and economic values were adjusted by the Consumer Price Index \citep{oecd_2017-VV} via Eq \ref{eqn:cpi}.
+Data contain monthly lobster landings (Kg) and value (MXP) from 2000 to 2014 for cooperatives with and without marine reserves (**Fig S1**).
+This information was aggregated by year, and economic values were adjusted by the Consumer Price Index \citep{oecd_2017-VV} as:
 
@@
 
-We use a set of governance indicators to analyze the governance structures of each cooperative \citep{leslie_2015-na}.
-The indicators (Table \ref{table:gov_ind}) focus on the resource system (four indicators), governance system (seven indicators), resource units (three indicators) and actors (three indicators).
-These were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2012} and based on the authors' experience and knowledge of the communities.
+Data for the qualitative analysis of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013} and based on the authors' experience and knowledge of the communities.
+These include information on the resource system, the resource units, actors, and the governance system itself (**S1 Table**).
 
@@
 
-We evaluate the effect thar marine reserves have had on four biological and two socioeconomic indicators (Table \ref{table:indicators}).
+We evaluate the effect that marine reserves have had on four biological and two socioeconomic indicators (Table \ref{table:indicators}).
 Recall that reserves were implemented to protect lobster and other benthic invertebrates.
-However, we also contemplate fish species that are also targeted in the off-season and might receive collateral protection from the reserves.
+However, we also use the available fish data, since finfish are targeted in the off-season and because there might be associated cobenefits of the reserve.
 
@@
            Indicador = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
-       Units = c("org / $m^2$", "org / $m^2$", "org / $m^2$", "org / $m^2$", "M MXP", "Metric Tones")) %>% 
+       Units = c("org / $m^2$", "org / $m^2$", "Kg / $m^2$", "org / $m^2$", "M MXP", "Metric Tonnes")) %>% 
   knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
@@
 
-```{r}
-read.csv(here::here("data", "gov_ind.csv")) %>% 
-  knitr::kable(caption = "\\label{table:gov_ind}Indicators used for the operationalization of the SES framework \\citep{leslie_2015-na}", col.names = c("Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero"), format = "latex") %>%
-  kableExtra::column_spec(1, width = "10em") %>% 
-  kableExtra::column_spec(2, width = "8em") %>% 
-  kableExtra::column_spec(3, width = "8em") %>% 
-  kableExtra::column_spec(4, width = "8em") %>% 
-  kableExtra::row_spec(1 , bold = T) %>% 
-  kableExtra::row_spec(6 , bold = T) %>% 
-  kableExtra::row_spec(14 , bold = T) %>% 
-  kableExtra::row_spec(18 , bold = T)
-```
-
-Biological indicators are analyzed with a difference--in--differences analysis (Eq \ref{eqn:reg_bio}), which allows us to estimate the effect that the reserve has on the biological indicators by comparing trends across time and treatments \citep{moland_2013-VP,Villasenor-Derbez_2018}.
+We use a difference-in-differences analysis to evaluate the biological indicators.
+This approach allows us to estimate the effect that the reserve has on the biological indicators by comparing trends across time and treatments (*i.e.* reserve / control sites \citet{moland_2013-VP,Villasenor-Derbez_2018}).
 The analysis is performed with a multiple linear regression of the form:
@@
 
-Where year--fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
-The interaction term $\lambda_{it} Year_t\times Zone_i$ represents represent the mean change in the indicator inside the reserve, for year $t$, with respect to the first year of evaluation in the control site (See Table \ref{table:com_sum}).
-When evaluating biomass and abundances, we include species--fixed effects ($\sigma_j$).
+Where year-fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
+The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site (See Table \ref{table:com_sum}).
+When evaluating biomass and densities of the entire benthic or fsh communities, we include $\sigma_j$ to control for species-fixed effects.
 
@@
 \begin{equation}
-I_i = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jComm_j +\epsilon
+I = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jCom_j +\epsilon
 \label{eqn:soc_reg}
@@
 
-The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$) and $\sigma_jComm$ captures community-level fixed-effects.
-This approach allows for a causal attribution of the observed changes to the reserve.
-All model coefficients were estimated via ordinary least--squares and heteroskedastic--robust standard errors \citep{zeileis_2004-7n}.
-
+The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$) and $\sigma_jCom$ captures community-level fixed-effects.
 These regressions allows us to make a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}.
-The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the reserve site at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
+The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
+All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}.
+All analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep(R_2018).
 
@@
 The following sections present the effect that marine reserves had on each of the biological and socioeconomic indicators for each coastal community.
-Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* effect size).
+Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* effect size $\lambda_t$).
 We also provide an overview of the governance settings of each community, and discuss how these dimensions might be related to the effectiveness of the reserves.
@@
 
-All indicators show idiosyncratic responses through time for each community.
-Figure \ref{fig:indicators}A shows how lobster densities increase inside the reserve for Isla Natividad an Punta Herrero within the first years, but the effect is aroded through time.
-However, these effects are in the order of 0.2 extra organisms $m^{-2}$ and are not significantly different from zero ($p > 0.05$).
-The only case when lobster densities showed a significant positive effect ($p < 0.1$) was for Isla Natividad on the sixth year (*i.e.* 2012), a year after hypoxia events described in \citet{micheli_2012-EU} caused mass mortality of organisms.
-Biomass was only evaluated for fish data (Fig \ref{fig:indicators}B), and no changes were observed ($p > 0.1$), Punta Herrero even showed negative insignificant trends in the order of 0.01.
-Invertebrate and fish densities (\ref{fig:indicators}C-D) show similar patterns of lack of effectiveness with effect sizes oscillating around zero.
-Full tables with model coefficients are presented in the supplementary materials (**S1 Table**, **S2 Table**, **S3 Table**).
+Indicators showed idiosyncratic responses through time for each community.
+Figure \ref{fig:indicators}A shows positive effect sizes for lobster densities for Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time.
+These effects are in the order of 0.2 extra organisms $m^{-2}$ but are not significantly different from zero ($p > 0.05$).
+Lobster densities were only significantly positive for Isla Natividad on the sixth year (*i.e.* 2012; $p < 0.05$), a year after the hypoxia events described by \citet{micheli_2012-EU} caused mass mortality of organisms.
+Likewise, no changes were detected in fish biomass or invertebrate and fish densities (\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
+Full tables with model coefficients are presented in the supplementary materials (**S2 Table**, **S3 Table**, **S4 Table**).
 
@@
 Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}).
-For all years befor implementation, the effect sizes are close to zero, indicating that the control and treatment sites track each other well.
+For all years befor implementation, the effect sizes are close to zero, indicating that the control and treatment sites track each other well and that these are plausible controls.
 However, effect sizes do not change after the implementation of the reserve.
@@
 However, these changes are also not ignificantly different and present an increased variation.
-All regression coefficients for each community and indicator are presented in **S4 Table**.
+All regression coefficients for each community and indicator are presented in **S5 Table**.
 
@@
 
-Although we have little information on the social dimension of these fisheries, using the SES framework indicators (Table \ref{table:gov_ind}), we can analyze the performance of each governance system (Table \ref{table:gov_res}).
+Although we have little information on the social dimension of these fisheries, we can use the social-ecologycal systems framework (**S1 Table**) to analyze the performance of each governance system (**S6 Table**).
 
@@
 
-```{r}
-read.csv(here::here("data","gov_res.csv")) %>% 
-  knitr::kable(caption = "\\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \\citep{mcginnis_2014}.",
-               col.names = c("", "Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
-  kableExtra::column_spec(1, width = "10em") %>%
-  kableExtra::column_spec(2, width = "10em") %>%
-  kableExtra::row_spec(1 , bold = T) %>% 
-  kableExtra::row_spec(7 , bold = T) %>% 
-  kableExtra::row_spec(15 , bold = T) %>% 
-  kableExtra::row_spec(19 , bold = T)
-```
-
 \clearpage
@@
 
-theory tells us more than 5 years are needed, so it;s not surprise for PH and ME. But IN didnt make it either.
-
-Our results indicated idiosyncratic biological effects of the reserves across communities and indicators, with a combination of positive and negative effects.
-However, many of these effects were not statistically significant, indicating no effect of the reserve.
-The socioeconomic indicators pertaining to landings and revenues showed little or no change after reserve implementation.
-While lobster densities represent an indicator directly tied to reserve objectives, we also evaluate other biological indicators to test for additional effects and as a double-check.
-ll indicators presented a similar pattern.
+Our results indicate that marine reserves are ineffective in increaseing population sizes of lobsters in the three communities.
+The socioeconomic indicators pertaining to landings and revenues showed little to no change after reserve implementation.
+While lobster densities represent an indicator directly tied to reserve objectives, we also evaluate other biological indicators to test for additional effects and fail to detect any effect other then the already reported buffering effect that Natividad reserves can have to environmental variability \citep{micheli_2012-EU}.
 The lack of expected effectiveness poses the question, why do these communities continue to support the reserves?
-While not in the main scope of this paper, understanding the social-ecological context in which these communities and their reserves might provide insight to answer this question.
-Here, we discuss plausible explanations to lack of effectiveness observed based on literature and our social-ecological system analysis.
+Understanding the social-ecological context in which these communities and their reserves operate might provide insights to this question.
+Here, we discuss plausible explanations to lack of effectiveness observed onthe biological and socioeconomic indicators based on our social-ecological system analysis.
 We also discuss potential shortcomings in our analysis, and provide management recommendations to improve reserve effectiveness.
@@
 Other works have compared the trend observed within a reserve through time \citep{betti_2017-lq}, which cannot distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}.
-By accounting for trends between sites and through time, we can control for time and space dynamics, and provide a better identification of the effect
+By accounting for trends between sites and through time, we can control for time and space dynamics, and provide a better identification of the effect.
 
@@
 
+There are two plausible explanations for the observed lack of effectiveness.
+First, marine reserves are only likely to provide fisheries benefits if initial population sizes are low \citep{hilborn_2006} and are poorly managed.
+However, both stocks were at some point certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
+The fishery is managed via species-specific minimum catch sizes, seasonal closures, and escapement windows on traps \cite{dof_website_1993}.
+A second plausible explanation lies in the design of these areas.
+Small reserves are unlikely to have a significant effect on population size if they fail to protect a significant portion of it.
+Intuitvely, small reserves have little or no displacement of fishing effort, and fishers do not perceive the short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
+On either case, these might explain why reserves were not effective, but not why they still receive community support.
+
+While reserves fail to provide conservation or economic benefits, they provide the community with access to funding from the filantrhopic sector.
+Furthermore, reserves can provide a joint enterprise to bring a community together, which promotes social cohesion and social capital.
+
 A second explanation lies on the the mismatch of objectives during the design process.
@@
 The social-ecological systems framework allowed a systematic diganosis and compare all the different case studies, allowing us to identify and tease appart possible explanations \citep{basurto_2013-oq}.
+
+Citar Fulton Marine Policy
 
@@
 \label{S1_Figure}
-Timeseries of indicators for IN
+Map of control and treated sites in A and control and treated landings in B
 
@@
 \label{S2_Figure}
+Timeseries of indicators for IN
+
+\paragraph*{S3 Figure}
+\label{S3_Figure}
 Timeseries of indicators for ME
 
-\paragraph*{S3 Figure}
+\paragraph*{S4 Figure}
 \label{S3_Figure}

### 2018-06-08 docs/paper/Villasenor-Derbez.tex
Made changes to manuscript body, just missing details on discussion

@@
 
-\title[Marine reserve effectiveness]{Enabling conditions for effective community-based marine reserves in
+\title[Community-based marine reserves]{Management and effectiveness of community-based marine reserves in
 small-scale fisheries}
@@
 
-Coastal marine ecosystems provide livelihoods for small-scale fishers
-and coastal communities around the world. However, overfishing and
-unsustainable fishing practices threaten the marine environment and
-jeopardize the wellbeing of coastal communities. A common approach to
-protect the environment and recover overexploited stocks is to implement
-no-take marine reserves (areas where all extractive activities are
-off-limits). In small-scale fisheries, these are sometimes implemented
-as community-based reserves, where a group of fishers collectively agree
-to close an area to fishing. While we know that reductions in fishing
-effort are followed by a series of ecological benefits (increased
-biomass, abundance, and species diversity), we do not fully understand
-how environmental and governance dynamics influence the conservation and
-fisheries benefits of community-based marine reserves. In this work, we
-evaluate the ecological outcomes of four reserves established by three
-coastal communities in temperate and tropical ecosystems of Mexico. By
-combining causal inference techniques with an operationalization of the
-social-ecological systems framework, we identify the environmental and
-social conditions that enable reserve effectiveness. Our results show a
-strong interaction between environmental variation and community
-organization, which influences reserve effectiveness. For example, the
-most effective reserve had strong governance structures accompanied with
-low environmental variability. Thus, even when a community is well
-organized (and reserves are well enforced), environmental variation can
-hinder the benefits of a reserve, and vice versa. Our results are
-particularly relevant under present changing climate conditions, as they
-can better inform management and decision making.
+Al finalizar revisiones
 
@@
 
-\textbf{Last\_update}: 2018-05-13
+\textbf{Last update}: 2018-06-08
 
@@
 \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. A common approach to
-manage the spatial distribution of fishing effort and recover stocks is
-through the implementation of marine reserves (\emph{i.e.} areas where
-all fishing activities are off--limits; MRs)
-\citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}.
+manage the spatial distribution of fishing effort to recover stocks and
+preserve biodiversity is through the implementation of marine reserves
+(MRs). These areas allow bounded populations to recover by limiting all
+extractive activities \citep{halpern_2002}. The science of MRs has
+largely focused on understanding the ecological effects of these areas,
+which include increased biomass, richness, and densities of organisms
+within the protected regions, climate change mitigation, and protection
+from environmental variability
+\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
+Modelling studies show that the fishery benefits of marine reserves
+depend on initial stock status and the type of management under which
+the fishery operates \citep{hilborn_2006}, as well as reserve size and
+the amount of larvae exported from these \citep{krueck_2017-J1}. Other
+research has focused on the relationship between socioeconomic and
+governance structures and their relationship to ecological effectiveness
+\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, few
+studies simultaneously evaluate MRs from all these perspectives
+(\emph{e.g.} \citet{lpezangarita_2014}).
 
-Marine reserve science has largely focused on understanding the
-ecological effects of these areas, which include increased biomass,
-richness, and densities of organisms within the protected regions
-\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, climate change
-mitigation \citep{roberts_2017-J9}, and protection from environmental
-variability \citep{micheli_2012-EU}. However, there is considerably less
-literature focusing on the relationship between socioeconomic and
-governance structures and their relationship to ecological effectiveness
-\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or benefits to
-fisheries \citep{krueck_2017-J1}; evaluations of marine reserves rarely
-provide a holistic view of the social-ecological system
-\citep{lpezangarita_2014}.
-
-While ecological factors (\emph{e.g.} habitat representation, initial
-state of protection, connectivity to other protected areas) can
-determine the success of a MR, their effectiveness also depends on the
-socioeconomic and governance settings under which they are implemented
-and managed. Literature shows that many non-ecological characteristics
-can play an equally important role in the effectiveness of MRs. For
-example, age of a reserve (\emph{i.e.} time since its implementation)
-and size were key to the effectiveness of MRs in Palau
-\citep{friedlander_2017-oI}. In the Mediterranean,
+While ecological factors like habitat representation within the MR or
+connectivity to other areas can determine the success of a MR, its
+effectiveness also depends on the socioeconomic and governance settings
+under which they are implemented and managed. Literature shows that many
+non-ecological characteristics can play an equally important role in the
+effectiveness of MRs. For example, age of a reserve (\emph{i.e.} time
+since its implementation) and size were key to the effectiveness of MRs
+in Palau \citep{friedlander_2017-oI}. In the Mediterranean,
 \citet{difranco_2016-Xw} identify that surveillance and enforcement,
@@
 practices were the key factors that increased stock health and income to
-fishers. At a global level, \citet{edgar_2014-UO} indicate that
-enforcement, age, size, and isolation were important factors determining
-effectiveness of the reserves.
+fishers. At a global level, enforcement, age, size, and isolation are
+important factors that determine the effectiveness of the reserves
+\citep{edgar_2014-UO}. However, MRs represent part of a more complex
+social-ecological systems governed by the interaction between humans and
+the environment. In this work we evaluate community-based marine
+reserves in Mexico using a triple bottom line approach by using
+ecological, socioeconomic, and governance indicators that provide a
+holistic view of the reserves \citep{halpern_2013}.
 
-Marine Reserves in Mexico have been commonly implemented as ``core
-zones'' within Biosphere Reserves that are administered by the National
-Commission of Protected Areas (\emph{Comisión Nacional de Áreas Marinas
-Protegidas}, CONANP). While CONANP has made efforts to have a
-participatory process, the implementation of these zones is still
-characterized by top--down approaches. This motivated Civil Society
-Organizations (CSOs) to work with coastal communities to implement
-community--based marine reserves \citep{uribe_2010-u2}, which are
-usually established within a TURFs; thus making them TURF--reserves
-\citep{afflerbach_2014-HP}. This bottom--up approach allows fishers to
-design their own reserves, which increases compliance and
-self--enforcement
+There are three main approaches to implement MRs, each of which have its
+pros and cons. We describe them in the context of Mexican MRs, but argue
+that at least the first two apply elsewhere. MRs were historically
+implemented and managed by a government agency, in this case the
+National Commission of Protected Areas (\emph{Comisión Nacional de Áreas
+Marinas Protegidas}, CONANP). While CONANP has made efforts to have a
+participatory process, the implementation of these zones is still mainly
+a top-down process. A second approach is the implementation of
+community-based marine reserves within areas of exclusive access
+(\emph{i.e.} TURFs), thus making them TURF-reserves
+\citep{afflerbach_2014-HP}. Community-based spatial closures ocurr in
+other places, like the \emph{kapu} or \emph{ra'ui} areas in the Pacific
+Islands \citep{bohnsack_2004,johannes_2002}. This bottom-up approach
+increases compliance and self-enforcement
 \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
-these reserves still lack legal recognition, making them vulnerable to
-poaching. In 2014, a new norm \citep{nom} allowed fishers to request the
-legal recognition of a community--based reserve under the name of
-``Fishing Refugia'' (\emph{Zona de Refugio Pesquero}, FR). Since then,
-\textbf{39} FR have been implemented along the Pacific, Gulf of
-California, and Mexican Caribbean coastlines, but no formal evaluation
-of their effectiveness has taken place.
+these lack legal recognition and rely on the exclusive access granted by
+the TURF, making it impossible to enforce without one. In an effort to
+provide a legal framework for these reserves, Civil Society
+Organizations (CSOs) served as the bridge between fishers and
+government. In 2014, a new norm \citep{nom} allowed fishers to request
+the legal recognition of a community-based reserve under the name of
+``Fishing Refugia'' (\emph{Zona de Refugio Pesquero}). These can be
+implemented as temporal or partial eservs, which can protect one, some,
+or all resources within them. Since then, \textbf{39} of these have been
+implemented along the Pacific, Gulf of California, and Mexican Caribbean
+coastlines, but no formal evaluation of their effectiveness has taken
+place.
 
-We combine causal inference techniques \citep{depalma_2018} and the
-social-ecological systems framework \citep{ostrom_2009-hg} to provide a
-comprehensive ecological and socioeconomic evaluation of community-based
+This work combines causal inference techniques and the social-ecological
+systems framework to provide a holistic evaluation of community-based
 marine reserves in three coastal communities in Mexico. The objective of
-this work is twofold: i) Provide the first evaluation of
-community--based marine reserves in Mexico, and ii) provide a
-comprehensive evaluation of the social--ecological system to identify
-how socioeconomic and governance characteristics relate to ecological
-effectiveness.
+this work is twofold. First, provide a triple bottom line evaluation of
+the effectiveness of community-based marine reserves that can inform
+similar processes in other countries. And second, perform the first
+formal evaluation of Fishing Refugia in Mexico and identify areas where
+improvement or adjustment might result in increased effectiveness. On
+both cases, we draw from the lessons learned and provide management
+recommendations to ensure or improve the effectiveness of
+community-based marine reserves.
 
@@
 
-We focus our evaluation in three coastal communities from the Pacific
-coast of Baja California and the Mesoamerican Reef System (Fig
-\ref{fig:map}). Fishing activities in all communities are assambled
-around fishing cooperatives that hold Territorial Use Rights for
-Fisheries (TURFs). Isla Natividad (IN) lies west of the Baja California
-Peninsula (Fig \ref{fig:map}B), where kelp forests (\emph{Macrocystis
-pyrifera}) and rocky reefs are the predominant habitats. The island is
-home to a the \emph{Sociedad Cooperative de Producción Pesquera (SCPP)
-Buzos y Pescadores de la Baja California}, spiny lobster
-(\emph{Panulirus interruptus}) represents an important fishery. However,
-other resources like finfish (yellow-tail jack, \emph{Seriola lalandi}),
-sea cucumber (\emph{Parastichopus parvimensis}), red sea urchin
-(\emph{Mesocentrotus franciscanus}), snail (\emph{Megastraea turbanica}
-y \emph{M. undosa}), and abalone (\emph{Haliotis spp}, until 2010) are
-also important sources of income. In 2006, the community decided to
-implement two community--based marine reserves within their fishing
-grounds, seeking to recover depleted stocks of invertebrate species
-(mainly lobster and abalone). Until today, these reserves are yet to be
-legally recognized as Fishing Refugia but count with full support from
-the community.
+We evaluate community-based marine reserves from three coastal
+communities located in the Pacific coast of Baja California and the
+Mexican Caribbean (Fig \ref{fig:map}). All communities are organized as
+fishing cooperatives that hold Territorial Use Rights for Fisheries
+(TURFs). Isla Natividad (IN) lies west of the Baja California Peninsula
+(Fig \ref{fig:map}B), where kelp forests and rocky reefs are the
+predominant habitats. The island is home to the \emph{Sociedad
+Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja
+California}, whose main resource by value is the spiny lobster
+(\emph{Panulirus interruptus}). However, other resources like finfish
+(yellow-tail jack, \emph{Seriola lalandi}), sea cucumber
+(\emph{Parastichopus parvimensis}), red sea urchin (\emph{Mesocentrotus
+franciscanus}), snail (\emph{Megastraea turbanica} y \emph{M. undosa}),
+and abalone (\emph{Haliotis spp}) are also important sources of income.
+In 2006, the community decided to implement two community-based marine
+reserves within their fishing grounds to protect commercially important
+invertebrate species (mainly lobster and abalone). Until today, these
+reserves are yet to be legally recognized as Fishing Refugia but count
+with full support from the community.
 
@@
 mangroves are the representative coastal ecosystems. ME is a fishing
-camp --visited intermittently during the fishing season-- belonging to
-the Cozumel fishing cooperative (\emph{SCPP Cozumel}). PH is also home
-to a fishing cooperative (\emph{SCPP José María Azcorra}). The main
-source of income to both communities is the Caribbean spiny lobster
-(\emph{Panulirus argus}). These communities also target finfish in the
-off season, mainly snappers (Lutjanidae) and groupers (Serranidae). ME
-established eight marine reserves in 2012, and PH established four
-marine reserves in 2013 and an additional community-based (\emph{i.e.}
-not legally recognized). All these reserves are legally recognized as
-Fishing Refugia.
+camp --visited intermitently during the fishing season-- belonging to
+the Cozumel fishing cooperative (\emph{SCPP Cozumel}); PH is home to the
+\emph{SCPP José María Azcorra} cooperative. Their main fishery is the
+Caribbean spiny lobster (\emph{Panulirus argus}), but they also target
+finfish in the off season, mainly snappers (Lutjanidae) and groupers
+(Serranidae). ME established eight marine reserves in 2012, and PH
+established four marine reserves in 2013 and an additional
+community-based (\emph{i.e.} not legally recognized). All these reserves
+are legally recognized as Fishing Refugia.
 
@@
 and control areas, carried out by members from each community and
-personnel from the Mexican CSO ``Comunidad y Biodiversidad''
-(\href{www.cobi.org.mx}{COBI}). These monitorings record richness and
+personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
+(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
 abundances of fish and invertebrate species in the reserves and control
-sites. Size structures are also collected in fish census. We define
-control sites as regions with habitat characteristics similar to the
-corresponding reserves, and that presumably had the same probability of
-being selected as reserves during the design phase. From all the
-reserves in these three communities, we use the ones that have data for
-reserve and control sites before and after the implementation of the
-reserve. This provides us with a Before-After-Control-Impact
-(\emph{i.e.} BACI) sampling design that allows us to capture and control
-for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
-BACI designs and causal inference techniques have proven effective to
-evaluate marine reserves, as they allow us to causally attribute
-observed changes to the intervention
-\citep{moland_2013-VP,Villasenor-Derbez_2018}. All reserves were
-surveyed annually from at least one year before implementation until
-2016. Table \ref{table:com_sum} shows a summary of sampling effort and
-year of implementation of the analyzed reserves.
+sites. Size structures are also collected for fish evaluations. We
+define control sites as regions with habitat characteristics similar to
+the corresponding reserves, and that presumably had a similar
+probability of being selected as reserves during the design phase. We
+focus our evaluation for sites where data are avaialable for reserve and
+control sites, before and after the implementation of the reserve. This
+provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
+sampling design that allows us to capture and control for temporal and
+spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and
+causal inference techniques have proven effective to evaluate marine
+reserves, as they allow us to causally attribute observed changes to the
+intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites
+were surveyed annually, and at least once before implementation of the
+reserves. Table \ref{table:com_sum} shows a summary of the reserves
+included in this study.
 
@@
 Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are
-located in the Yucatan Peninsula.}
+located in the Yucatan Peninsula. Blue polygons represent the TURFs, and
+red polygons the marine reserves.}
 \end{figure}
@@
 Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
-(Kg) and value (MXP) from 2000 to 2014. This information was aggregated
-by year, and economic values were adjusted by the Consumer Price Index
-\citep{oecd_2017-VV} via Eq \ref{eqn:cpi}.
+(Kg) and value (MXP) from 2000 to 2014 for cooperatives with and without
+marine reserves (\textbf{Fig S1}). This information was aggregated by
+year, and economic values were adjusted by the Consumer Price Index
+\citep{oecd_2017-VV} as:
 
@@
 
-We use a set of governance indicators to analyze the governance
-structures of each cooperative \citep{leslie_2015-na}. The indicators
-(Table \ref{table:gov_ind}) focus on the resource system (four
-indicators), governance system (seven indicators), resource units (three
-indicators) and actors (three indicators). These were collected at the
-community--level from official documents used in the creation and
-designation of the marine reserves
-\citep{dof_website_2012,dof_website_2012} and based on the authors'
-experience and knowledge of the communities.
+Data for the qualitative analysis of the social-ecological system were
+collected at the community--level from official documents used in the
+creation and designation of the marine reserves
+\citep{dof_website_2012,dof_website_2013} and based on the authors'
+experience and knowledge of the communities. These include information
+on the resource system, the resource units, actors, and the governance
+system itself (\textbf{S1 Table}).
 
@@
 
-We evaluate the effect thar marine reserves have had on four biological
+We evaluate the effect that marine reserves have had on four biological
 and two socioeconomic indicators (Table \ref{table:indicators}). Recall
 that reserves were implemented to protect lobster and other benthic
-invertebrates. However, we also contemplate fish species that are also
-targeted in the off-season and might receive collateral protection from
-the reserves.
+invertebrates. However, we also use the available fish data, since
+finfish are targeted in the off-season and because there might be
+associated cobenefits of the reserve.
 
@@
 \hline
-Biological & Fish biomass & org / $m^2$\\
+Biological & Fish biomass & Kg / $m^2$\\
 \hline
@@
 \hline
-Socioeconomic & Landings from target species & Metric Tones\\
+Socioeconomic & Landings from target species & Metric Tonnes\\
 \hline
@@
 
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-4}\label{table:gov_ind}Indicators used for the operationalization of the SES framework \citep{leslie_2015-na}}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{8em}|>{\raggedright\arraybackslash}p{8em}|>{\raggedright\arraybackslash}p{8em}}
-\hline
-Indicator & Isla Natividad & Maria Elena & Punta Herrero\\
-\hline
-\textbf{Resource systems (RS)} & \textbf{} & \textbf{}\\
-\hline
-TURF presence & Yes & Yes & Yes\\
-\hline
-Type of ecosystem & Kelp Forest / Rocky Reefs & Coral Reef & Coral Reef\\
-\hline
-Intensity of environmental Disturbance & El nino event & Hurricanes & Hurricanes\\
-\hline
-Location & Island & Coastal & Coastal\\
-\hline
-\textbf{Governance systems (GS)} & \textbf{} & \textbf{}\\
-\hline
-Fishing cooperative & Yes & Yes & Yes\\
-\hline
-Involved actors & COBI, Stanford, REBIVI & Alianza Kanan Kay, COBI, CONANP, Coop, CONAPESCA, Oceanus, FCyRH, FHMM, & Alianza Kanan Kay, COBI, CONANP, Coop, CONAPESCA, Oceanus, FCyRH, FHMM,\\
-\hline
-Presence of an inter-cooperative structure & Fedecoop & Non & Non\\
-\hline
-Fishing Regulations & Size limits, seasonal closures, quotas & Size limits, seasonal closures & Size limits, seasonal closures\\
-\hline
-Enforcement technology & Boats & Boats & Land enforcement\\
-\hline
-MR enforcement &  &  & \\
-\hline
-Cooperative regulations &  &  & \\
-\hline
-\textbf{Resource Units (RU)} & \textbf{} & \textbf{}\\
-\hline
-Adult targeted species mobility & 1km & 30km & 30km\\
-\hline
-Targeted species longevity (years) &  &  & \\
-\hline
-Price of targeted species &  &  & \\
-\hline
-\textbf{Actors (A)} & \textbf{} & \textbf{}\\
-\hline
-Leadership &  &  & \\
-\hline
-Level of illegal fishing & 1 & 1 & 3\\
-\hline
-Presence of alternative livelihoods &  &  & \\
-\hline
-\end{tabular}
-\end{table}
-
-Biological indicators are analyzed with a difference--in--differences
-analysis (Eq \ref{eqn:reg_bio}), which allows us to estimate the effect
-that the reserve has on the biological indicators by comparing trends
-across time and treatments
-\citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis is performed
-with a multiple linear regression of the form:
+We use a difference-in-differences analysis to evaluate the biological
+indicators. This approach allows us to estimate the effect that the
+reserve has on the biological indicators by comparing trends across time
+and treatments (\emph{i.e.} reserve / control sites
+\citet{moland_2013-VP,Villasenor-Derbez_2018}). The analysis is
+performed with a multiple linear regression of the form:
 
@@
 
-Where year--fixed effects are represented by \(\gamma_{it} Year_t\), and
+Where year-fixed effects are represented by \(\gamma_{it} Year_t\), and
 \(\beta Zone_i\) captures the difference between reserve (\(Zone = 1\))
 and control (\(Zone = 0\)) sites. The interaction term
-\(\lambda_{it} Year_t\times Zone_i\) represents represent the mean
-change in the indicator inside the reserve, for year \(t\), with respect
-to the first year of evaluation in the control site (See Table
-\ref{table:com_sum}). When evaluating biomass and abundances, we include
-species--fixed effects (\(\sigma_j\)).
+\(\lambda_{it} Year_t\times Zone_i\) represents the mean change in the
+indicator inside the reserve, for year \(t\), with respect to the year
+of implementation in the control site (See Table \ref{table:com_sum}).
+When evaluating biomass and densities of the entire benthic or fsh
+communities, we include \(\sigma_j\) to control for species-fixed
+effects.
 
@@
 this case the \(Treated\) dummy variable indicates if the community has
-a reserve (\(Treated = 1\)) or not (\(Treated = 0\)) and
-\(\sigma_jComm\) captures community-level fixed-effects. This approach
-allows for a causal attribution of the observed changes to the reserve.
-All model coefficients were estimated via ordinary least--squares and
-heteroskedastic--robust standard errors \citep{zeileis_2004-7n}.
-
-These regressions allows us to make a causal link between the
-implementation of marine reserves and the observed trends by accounting
-for temporal and spatial dynamics \citep{depalma_2018}. The effect of
-the reserve is captured by the \(\lambda_t\) coefficient, and represents
-the difference observed between the control site before the
-implementation of the reserve and the reserve site at time \(t\) after
-controlling for other time and space variations (\emph{i.e.}
-\(\gamma_t\) and \(\beta\) respectively).
+a reserve (\(Treated = 1\)) or not (\(Treated = 0\)) and \(\sigma_jCom\)
+captures community-level fixed-effects. These regressions allows us to
+make a causal link between the implementation of marine reserves and the
+observed trends by accounting for temporal and spatial dynamics
+\citep{depalma_2018}. The effect of the reserve is captured by the
+\(\lambda_t\) coefficient, and represents the difference observed
+between the control site before the implementation of the reserve and
+the treated sites at time \(t\) after controlling for other time and
+space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
+All model coefficients were estimated via ordinary least-squares and
+heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
+analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep(R\_2018).
 
@@
 and across sites, relative to the control site on the year of
-implementation (\emph{i.e.} effect size). We also provide an overview of
-the governance settings of each community, and discuss how these
-dimensions might be related to the effectiveness of the reserves.
+implementation (\emph{i.e.} effect size \(\lambda_t\)). We also provide
+an overview of the governance settings of each community, and discuss
+how these dimensions might be related to the effectiveness of the
+reserves.
 
@@
 
-All indicators show idiosyncratic responses through time for each
-community. Figure \ref{fig:indicators}A shows how lobster densities
-increase inside the reserve for Isla Natividad an Punta Herrero within
-the first years, but the effect is aroded through time. However, these
-effects are in the order of 0.2 extra organisms \(m^{-2}\) and are not
-significantly different from zero (\(p > 0.05\)). The only case when
-lobster densities showed a significant positive effect (\(p < 0.1\)) was
-for Isla Natividad on the sixth year (\emph{i.e.} 2012), a year after
-hypoxia events described in \citet{micheli_2012-EU} caused mass
-mortality of organisms. Biomass was only evaluated for fish data (Fig
-\ref{fig:indicators}B), and no changes were observed (\(p > 0.1\)),
-Punta Herrero even showed negative insignificant trends in the order of
-0.01. Invertebrate and fish densities (\ref{fig:indicators}C-D) show
-similar patterns of lack of effectiveness with effect sizes oscillating
-around zero. Full tables with model coefficients are presented in the
-supplementary materials (\textbf{S1 Table}, \textbf{S2 Table},
-\textbf{S3 Table}).
+Indicators showed idiosyncratic responses through time for each
+community. Figure \ref{fig:indicators}A shows positive effect sizes for
+lobster densities for Isla Natividad and Punta Herrero during the first
+years, but the effect is eroded through time. These effects are in the
+order of 0.2 extra organisms \(m^{-2}\) but are not significantly
+different from zero (\(p > 0.05\)). Lobster densities were only
+significantly positive for Isla Natividad on the sixth year (\emph{i.e.}
+2012; \(p < 0.05\)), a year after the hypoxia events described by
+\citet{micheli_2012-EU} caused mass mortality of organisms. Likewise, no
+changes were detected in fish biomass or invertebrate and fish densities
+(\ref{fig:indicators}B-D), where effect sizes oscillated around zero
+without clear trends. Full tables with model coefficients are presented
+in the supplementary materials (\textbf{S2 Table}, \textbf{S3 Table},
+\textbf{S4 Table}).
 
@@
 implementation, the effect sizes are close to zero, indicating that the
-control and treatment sites track each other well. However, effect sizes
-do not change after the implementation of the reserve. Again, the
-negative coeficient observed for Isla Natividad on year 5 correspond to
-the 2011 hypoxia events. The only positive change observed in lobster
-landings is for Isla Natividad in 2014 (\(p < 0.1\)). The trhee years of
-post-implementation data for Maria Elena do not show a significant
-effect of the reserve. Isla Natividad shows higher revenues after the
-iumplementation of the reserve, as compared to the control communities.
-However, these changes are also not ignificantly different and present
-an increased variation. All regression coefficients for each community
-and indicator are presented in \textbf{S4 Table}.
+control and treatment sites track each other well and that these are
+plausible controls. However, effect sizes do not change after the
+implementation of the reserve. Again, the negative coeficient observed
+for Isla Natividad on year 5 correspond to the 2011 hypoxia events. The
+only positive change observed in lobster landings is for Isla Natividad
+in 2014 (\(p < 0.1\)). The trhee years of post-implementation data for
+Maria Elena do not show a significant effect of the reserve. Isla
+Natividad shows higher revenues after the iumplementation of the
+reserve, as compared to the control communities. However, these changes
+are also not ignificantly different and present an increased variation.
+All regression coefficients for each community and indicator are
+presented in \textbf{S5 Table}.
 
@@
 Although we have little information on the social dimension of these
-fisheries, using the SES framework indicators (Table
-\ref{table:gov_ind}), we can analyze the performance of each governance
-system (Table \ref{table:gov_res}).
+fisheries, we can use the social-ecologycal systems framework
+(\textbf{S1 Table}) to analyze the performance of each governance system
+(\textbf{S6 Table}).
 
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{\label{fig:unnamed-chunk-5}\label{fig:indicators}Effect sizes for
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{\label{fig:unnamed-chunk-4}\label{fig:indicators}Effect sizes for
 marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-7-1.pdf}
-\caption{\label{fig:unnamed-chunk-7}\label{fig:lobsters}Effect sizes for
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{\label{fig:unnamed-chunk-6}\label{fig:lobsters}Effect sizes for
 lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
@@
 \end{figure}
-
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-8}\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \citep{mcginnis_2014}.}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}|l|l|l}
-\hline
- & Indicator & Isla Natividad & Maria Elena & Punta Herrero\\
-\hline
-\textbf{Resource systems (RS)} & \textbf{} & \textbf{} & \textbf{}\\
-\hline
-RS2 – Clarity of system boundaries & TURF presence & High & High & High\\
-\hline
-RS3 – Size of resource system &  &  &  & \\
-\hline
-RS5 – Productivity of system & Type of ecosystem & High & High & High\\
-\hline
-RS7 – Predictability of system dynamics & Intensity of environmental disturbance & Low (ENSO) & High & High\\
-\hline
-RS9 – Location & Proximity to other communities/cities & Isolated & Not Isolated & Not Isolated\\
-\hline
-\textbf{Governance systems (GS)} & \textbf{} & \textbf{} & \textbf{}\\
-\hline
-GS1 – Government organizations & Presence of fishing cooperatives & Yes & Yes & Yes\\
-\hline
-GS2 – Nongovernment organizations & Involved actors & Yes & Yes & Yes\\
-\hline
-GS3 – Network structure & Presence of an inter-cooperative structure & Yes & No & No\\
-\hline
-GS4 – Property-rights systems & TURF presence & Yes & Yes & Yes\\
-\hline
-GS5 – Operational-choice rules & Fishing Regulations / MPA enforcement / Enforcement technolofy & Yes & Yes & Yes\\
-\hline
-GS6 – Collective-choice rules & Cooperative regulations & Yes & Yes & Yes\\
-\hline
-GS7 – Constitutional-choice rules &  &  &  & \\
-\hline
-\textbf{Resource units (RU)} & \textbf{} & \textbf{} & \textbf{}\\
-\hline
-RU1 – Resource unit mobility & Targeted species home range & Low & Medium & Medium\\
-\hline
-RU2 – Growth or replacement rate & Max age of targeted species & Low & Medium & Medium\\
-\hline
-RU4 – Economic value & Price of targeted species & high & High & high\\
-\hline
-\textbf{Actors (A)} & \textbf{} & \textbf{} & \textbf{}\\
-\hline
-A1 – Number of relevant actors &  & 98 &  & \\
-\hline
-A2 – Socioeconomic attributes &  &  &  & \\
-\hline
-A5 – Leadership/entrepreneurship & Leadership & High & High & High\\
-\hline
-A6 – Norms (trust-reciprocity)/social capital--- (Based on illegal fishing) & Level of illegal fishing & High & High & Low\\
-\hline
-A8 – Importance of resource (dependence) & Presence of alternative livelihoods & High & High & High\\
-\hline
-\end{tabular}
-\end{table}
 
@@
 
-theory tells us more than 5 years are needed, so it;s not surprise for
-PH and ME. But IN didnt make it either.
-
-Our results indicated idiosyncratic biological effects of the reserves
-across communities and indicators, with a combination of positive and
-negative effects. However, many of these effects were not statistically
-significant, indicating no effect of the reserve. The socioeconomic
-indicators pertaining to landings and revenues showed little or no
+Our results indicate that marine reserves are ineffective in increaseing
+population sizes of lobsters in the three communities. The socioeconomic
+indicators pertaining to landings and revenues showed little to no
 change after reserve implementation. While lobster densities represent
 an indicator directly tied to reserve objectives, we also evaluate other
-biological indicators to test for additional effects and as a
-double-check. ll indicators presented a similar pattern. The lack of
-expected effectiveness poses the question, why do these communities
-continue to support the reserves? While not in the main scope of this
-paper, understanding the social-ecological context in which these
-communities and their reserves might provide insight to answer this
-question. Here, we discuss plausible explanations to lack of
-effectiveness observed based on literature and our social-ecological
-system analysis. We also discuss potential shortcomings in our analysis,
-and provide management recommendations to improve reserve effectiveness.
+biological indicators to test for additional effects and fail to detect
+any effect other then the already reported buffering effect that
+Natividad reserves can have to environmental variability
+\citep{micheli_2012-EU}. The lack of expected effectiveness poses the
+question, why do these communities continue to support the reserves?
+Understanding the social-ecological context in which these communities
+and their reserves operate might provide insights to this question.
+Here, we discuss plausible explanations to lack of effectiveness
+observed onthe biological and socioeconomic indicators based on our
+social-ecological system analysis. We also discuss potential
+shortcomings in our analysis, and provide management recommendations to
+improve reserve effectiveness.
 
@@
 time and space dynamics, and provide a better identification of the
-effect
+effect.
 
@@
 abundances to increase enough to export larvae or adult organisms.
+
+There are two plausible explanations for the observed lack of
+effectiveness. First, marine reserves are only likely to provide
+fisheries benefits if initial population sizes are low
+\citep{hilborn_2006} and are poorly managed. However, both stocks were
+at some point certified by the Marine Stewardship Council
+\citep{prezramrez_2016-J1}. The fishery is managed via species-specific
+minimum catch sizes, seasonal closures, and escapement windows on traps
+\cite{dof_website_1993}. A second plausible explanation lies in the
+design of these areas. Small reserves are unlikely to have a significant
+effect on population size if they fail to protect a significant portion
+of it. Intuitvely, small reserves have little or no displacement of
+fishing effort, and fishers do not perceive the short-term costs
+associated to the first years of reserve implementation
+\citep{ovando_2016-Wg}. On either case, these might explain why reserves
+were not effective, but not why they still receive community support.
+
+While reserves fail to provide conservation or economic benefits, they
+provide the community with access to funding from the filantrhopic
+sector. Furthermore, reserves can provide a joint enterprise to bring a
+community together, which promotes social cohesion and social capital.
 
@@
 
-Timeseries of indicators for IN
+Map of control and treated sites in A and control and treated landings
+in B
 
@@
 
+Timeseries of indicators for IN
+
+\paragraph*{S3 Figure}
+\label{S3_Figure}
+
 Timeseries of indicators for ME
 
-\paragraph*{S3 Figure}
+\paragraph*{S4 Figure}
 \label{S3_Figure}

### 2018-06-11 docs/paper/Villasenor-Derbez.Rmd
Updated manuscript and sent it to co-authors

@@
 ---
-title: "Management and effectiveness of community-based marine reserves in small-scale fisheries"
+title: "Effectiveness of community-based marine reserves in small-scale fisheries"
 rtitle: "Community-based marine reserves"
@@
 **Titulos alternativos**
+
+"Management and effectiveness of community-based marine reserves in small-scale fisheries"
 
@@
 Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}.
-A common approach to manage the spatial distribution of fishing effort to recover stocks and preserve biodiversity is through the implementation of marine reserves (MRs).
+A common approach to manage the spatial distribution of fishing effort to recover stocks and preserve biodiversity is through the implementation of marine reserves.
 These areas allow bounded populations to recover by limiting all extractive activities \citep{halpern_2002}.
-The science of MRs has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
-Modelling studies show that the fishery benefits of marine reserves depend on initial stock status and the type of management under which the fishery operates \citep{hilborn_2006}, as well as reserve size and the amount of larvae exported from these \citep{krueck_2017-J1}.
-Other research has focused on the relationship between socioeconomic and governance structures and their relationship to ecological effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
-However, few studies simultaneously evaluate MRs from all these perspectives (*e.g.* \citet{lpezangarita_2014}).
 
-While ecological factors like habitat representation within the MR or connectivity to other areas can determine the success of a MR, its effectiveness also depends on the socioeconomic and governance settings under which they are implemented and managed.
+The science of marine reserves has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
+Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1}.
+Other research has focused on the relationship between socioeconomic and governance structures and their relationship to reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
+However, few studies simultaneously evaluate reserves from all these perspectives (*e.g.* \citet{lpezangarita_2014}).
+
+<!-- While ecological factors like habitat representation within the MR or connectivity to other areas can determine the success of a MR, its effectiveness also depends on the socioeconomic and governance settings under which they are implemented and managed.
 Literature shows that many non-ecological characteristics can play an equally important role in the effectiveness of MRs.
@@
 In this work we evaluate community-based marine reserves in Mexico using a triple bottom line approach by using ecological, socioeconomic, and governance indicators that provide a holistic view of the reserves \citep{halpern_2013}.
+-->
 
-There are three main approaches to implement MRs, each of which have its pros and cons.
-We describe them in the context of Mexican MRs, but argue that at least the first two apply elsewhere.
-MRs were historically implemented and managed by a government agency, in this case the National Commission of Protected Areas (*Comisión Nacional de Áreas Marinas Protegidas*, CONANP).
-While CONANP has made efforts to have a participatory process, the implementation of these zones is still mainly a top-down process.
-A second approach is the implementation of community-based marine reserves within areas of exclusive access (*i.e.* TURFs), thus making them TURF-reserves \citep{afflerbach_2014-HP}.
-Community-based spatial closures ocurr in other places, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
+Implementation of marine reserves can be done by following one of three main approaches in terms of implementation.
+We describe these in the context of Mexican MRs, but argue that at least the first two apply elsewhere.
+Marine reserves are usually implemented and managed by a government agency, in this case the National Commission of Natural Protected Areas (*Comisión Nacional de Áreas Naturales Protegidas*, CONANP).
+While CONANP has made efforts to incorporate stakeholders in the design and implementation phases, these are still characterized by a heavy top-down process.
+A second approach is the implementation of community-based marine reserves, usually placed within areas of exclusive access (*i.e.* TURFs).
+Community-based spatial closures occur in other places, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
 This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
-However, these lack legal recognition and rely on the exclusive access granted by the TURF, making it impossible to enforce without one.
-In an effort to provide a legal framework for these reserves, Civil Society Organizations (CSOs) served as the bridge between fishers and government.
-In 2014, a new norm \citep{nom} allowed fishers to request the legal recognition of a community-based reserve under the name of "Fishing Refugia" (*Zona de Refugio Pesquero*).
-These can be implemented as temporal or partial eservs, which can protect one, some, or all resources within them.
-Since then, **39** of these have been implemented along the Pacific, Gulf of California, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.
+However, without legal recognition these are difficult to enforce and fishers rely on the exclusive access granted by the TURF.
+In an effort to bridge this normative gap, Civil Society Organizations (CSOs) served as the link between fishers and government, and set out to create a legal framework that solve this governance issue.
+In 2014, a new norm was created, allowing fishers to request the legal recognition of a community-based reserve under the name of "Fishing Refugia" \citep{nom}.
+These can be implemented as temporal or partial reserves, which can protect one, some, or all resources within them.
+Since then, **41** of community-based marine reserves along the Pacific, Gulf of California, and Mexican Caribbean coastlines have gained legal recognition, but no formal evaluation of their effectiveness has taken place.
 
@@
 And second, perform the first formal evaluation of Fishing Refugia in Mexico and identify areas where improvement or adjustment might result in increased effectiveness.
-On both cases, we draw from the lessons learned and provide management recommendations to ensure or improve the effectiveness of community-based marine reserves.
+On both cases, we draw from the lessons learned and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries.
 
@@
 All communities are organized as fishing cooperatives that hold Territorial Use Rights for Fisheries (TURFs).
-Isla Natividad (IN) lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests and rocky reefs are the predominant habitats.
-The island is home to the *Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California*, whose main resource by value is the spiny lobster (*Panulirus interruptus*).
-However, other resources like finfish (yellow-tail jack, *Seriola lalandi*), sea cucumber (*Parastichopus parvimensis*), red sea urchin (*Mesocentrotus franciscanus*), snail (*Megastraea turbanica* y *M.
-undosa*), and abalone (*Haliotis spp*) are also important sources of income.
+Isla Natividad lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests and rocky reefs are the predominant habitats.
+The island is home to the *Buzos y Pescadores de la Baja California* fishing cooperative, whose main resource by value is the spiny lobster (*Panulirus interruptus*).
+Other resources like finfish, sea cucumber, red sea urchin, snail, and abalone are also important sources of income.
 In 2006, the community decided to implement two community-based marine reserves within their fishing grounds to protect commercially important invertebrate species (mainly lobster and abalone).
-Until today, these reserves are yet to be legally recognized as Fishing Refugia but count with full support from the community.
+These reserves obtained legal recognition in 2018, but were well enforced since their implementation in 2006.
 
-The other two communities are Maria Elena (ME) and Punta Herrero (PH; Fig \ref{fig:map}C) in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems.
-ME is a fishing camp --visited intermitently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); PH is home to the *SCPP José María Azcorra* cooperative.
-Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off season, mainly snappers (Lutjanidae) and groupers (Serranidae).
-ME established eight marine reserves in 2012, and PH established four marine reserves in 2013 and an additional community-based (*i.e.* not legally recognized).
-All these reserves are legally recognized as Fishing Refugia.
+The other two communities are Maria Elena and Punta Herrero, which are located in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems( Fig \ref{fig:map}C).
+Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); Punta Herrero is home to the *SCPP José María Azcorra* cooperative.
+Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season.
+Maria Elena and Punta Herrero established eight marine reserves in 2012, and four marine reserves in 2013, respectively.
+All these reserves are legally recognized as Fishing Refugia since their creation.
 
@@
 
-We use three main sources of information to evaluate these reserves.
+We use three main sources of information to evaluate these reserves across the ecological, socioeconomic, and governance dimensions.
 Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
 Trained divers record richness and abundances of fish and invertebrate species in the reserves and control sites.
-Size structures are also collected for fish evaluations.
+Size structures are also collected during fish surveys.
 We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had a similar probability of being selected as reserves during the design phase.
-We focus our evaluation for sites where data are avaialable for reserve and control sites, before and after the implementation of the reserve.
+We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
 This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
@@
 Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA).
-Data contain monthly lobster landings (Kg) and value (MXP) from 2000 to 2014 for cooperatives with and without marine reserves (**Fig S1**).
-This information was aggregated by year, and economic values were adjusted by the Consumer Price Index \citep{oecd_2017-VV} as:
+Data contain monthly lobster landings (Kg) and revenues (MXP) from 2000 to 2014 for cooperatives with and without marine reserves(**Fig S1**).
+All cooperatives of each region (*i.e.* Pacific and Caribbean) incorporated in this analysis, belong to larger Cooperative Federations, and are exposed to the same markets and institutional frameworks \citep{mccay_2017-1m,ayer_2018}, making them plausible controls.
+Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted by the Consumer Price Index for Mexico \citep{oecd_2017-VV} as:
 
@@
 
-Data for the qualitative analysis of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013} and based on the authors' experience and knowledge of the communities.
+Data for the qualitative analysis of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities.
 These include information on the resource system, the resource units, actors, and the governance system itself (**S1 Table**).
@@
 
-We evaluate the effect that marine reserves have had on four biological and two socioeconomic indicators (Table \ref{table:indicators}).
+We evaluate the effect that marine reserves have had on four ecological and two socioeconomic indicators (Table \ref{table:indicators}).
 Recall that reserves were implemented to protect lobster and other benthic invertebrates.
-However, we also use the available fish data, since finfish are targeted in the off-season and because there might be associated cobenefits of the reserve.
+However, we also use the available fish datato test for associated co-benefits.
 
@@
            Indicador = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
-       Units = c("org / $m^2$", "org / $m^2$", "Kg / $m^2$", "org / $m^2$", "M MXP", "Metric Tonnes")) %>% 
+       Units = c("org $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "Kg $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "M MXP", "Metric Tonnes")) %>% 
   knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
@@
 
-We use a difference-in-differences analysis to evaluate the biological indicators.
-This approach allows us to estimate the effect that the reserve has on the biological indicators by comparing trends across time and treatments (*i.e.* reserve / control sites \citet{moland_2013-VP,Villasenor-Derbez_2018}).
-The analysis is performed with a multiple linear regression of the form:
+We use a difference-in-differences analysis to evaluate these indicators.
+This approach allows us to estimate the effect that the reserve had by comparing trends across time and treatments (*i.e.* reserve / control sites \citet{moland_2013-VP,Villasenor-Derbez_2018}).
+The analysis of ecological indicators is performed with a multiple linear regression of the form:
 
@@
 The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site (See Table \ref{table:com_sum}).
-When evaluating biomass and densities of the entire benthic or fsh communities, we include $\sigma_j$ to control for species-fixed effects.
+When evaluating biomass and densities of the entire benthic or fish communities, we include $\sigma_j$ to control for species-fixed effects.
 
-Socioeconomic indicators are evaluated with a similar approach (Eq \ref{eqn:soc_reg}).
-Due to data constrains, only Isla Natividad and Maria Elena are evaluated in this case.
-We constructed panel-data information with yearly (2001 - 2014) lobster landings and income of the studied and neighbouring communities that have similar management strategies, and belong to larger Cooperative Federtations \citep{mccay_2017-1m,ayer_2018}.
-Neighbouring communities are used as counterfacutals that allow us to control for unobserved time-invariants.
+Socioeconomic indicators are evaluated with a similar approach.
+Due to data constrains, we only evaluate socioeconomic data for Isla Natividad and Maria Elena.
+Neighboring communities are used as counterfacutals that allow us to control for unobserved time-invariants.
 Each "treated" community (Isla Natividad and Maria Elena) has three counterfactual communities.
@@
 All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}.
-All analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep(R_2018).
+All analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.
 
@@
 Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* effect size $\lambda_t$).
-We also provide an overview of the governance settings of each community, and discuss how these dimensions might be related to the effectiveness of the reserves.
+We also provide an overview of the governance settings of each community, and discuss how these might be related to the effectiveness and performance of the reserves.
 
@@
 
-Indicators showed idiosyncratic responses through time for each community.
-Figure \ref{fig:indicators}A shows positive effect sizes for lobster densities for Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time.
-These effects are in the order of 0.2 extra organisms $m^{-2}$ but are not significantly different from zero ($p > 0.05$).
+Indicators showed ambiguous responses through time for each reserve.
+Figure \ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time.
+These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ but are not significantly different from zero ($p > 0.05$).
 Lobster densities were only significantly positive for Isla Natividad on the sixth year (*i.e.* 2012; $p < 0.05$), a year after the hypoxia events described by \citet{micheli_2012-EU} caused mass mortality of organisms.
@@
 Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}).
-For all years befor implementation, the effect sizes are close to zero, indicating that the control and treatment sites track each other well and that these are plausible controls.
+For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, asuggesting that these are plausible controls.
 However, effect sizes do not change after the implementation of the reserve.
-Again, the negative coeficient observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events.
+Again, the negative coefficient observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events.
 The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$).
-The trhee years of post-implementation data for Maria Elena do not show a significant effect of the reserve.
-Isla Natividad shows higher revenues after the iumplementation of the reserve, as compared to the control communities.
-However, these changes are also not ignificantly different and present an increased variation.
+The three years of post-implementation data for Maria Elena do not show a significant effect of the reserve.
+Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities.
+However, these changes are not significant and are associated to increased variation.
 All regression coefficients for each community and indicator are presented in **S5 Table**.
@@
 
-Although we have little information on the social dimension of these fisheries, we can use the social-ecologycal systems framework (**S1 Table**) to analyze the performance of each governance system (**S6 Table**).
-
+Although we have little information on the social dimension of these fisheries, we can use the social-ecological systems framework (**S1 Table**) to analyze the performance of each governance system (**S6 Table**).
 Our analysis shows that all of the systems analyzed share similarities in their Governance system which is based on cooperatives (GS5.2.3.2), with strong rules in use that include Operational rules (GS6.2), Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even Territorial use communal rights (GS6.1.4.3).
-However, we identified important differences in terms of the actors, resource systems and resource units.
-The value of lobster in isla natividad is higher than the lobster sold from Punta Herrero and Maria Elena (RU4), which can reduce the pressure on harvest.
-Lastly, in terms of actors, although all communities show a high level of leadership (A5), the level of trust (A6.1) is lower in Punta Herrero.
+However, we identified important differences in terms of the actors, resource systems, and resource units.
+Although all communities show a high level of leadership (A5), the level of trust (A6.1) is lower in Punta Herrero.
 In general, the presence and success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources they depend upon \citep{jupiter_2017}.
@@
 
-Our results indicate that marine reserves are ineffective in increaseing population sizes of lobsters in the three communities.
-The socioeconomic indicators pertaining to landings and revenues showed little to no change after reserve implementation.
-While lobster densities represent an indicator directly tied to reserve objectives, we also evaluate other biological indicators to test for additional effects and fail to detect any effect other then the already reported buffering effect that Natividad reserves can have to environmental variability \citep{micheli_2012-EU}.
-The lack of expected effectiveness poses the question, why do these communities continue to support the reserves?
+Our results indicate that marine reserves do not achieve the main objective of increasing lobster densities.
+No co-benefits were identified when using other ecological indicators other than the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \citep{micheli_2012-EU}.
+The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
+The lack of expected effectiveness poses the question: why do these communities continue to support the reserves?
 Understanding the social-ecological context in which these communities and their reserves operate might provide insights to this question.
-Here, we discuss plausible explanations to lack of effectiveness observed onthe biological and socioeconomic indicators based on our social-ecological system analysis.
-We also discuss potential shortcomings in our analysis, and provide management recommendations to improve reserve effectiveness.
+Here we touch on potential shortcomings in our analysis, and discuss our findings in the context of existing literature and our social-ecological system analysis.
+Finally, we provide recommendations to guide the implementation of new (and improve effectiveness of) community-based marine reserves in small-scale fisheries.
 
-Our approach to evaluate the temporal and spatial changes of each indicator provides a more robust measure of reserve effectiveness.
-Some works have solely focused on an inside-outside comparison of indicators \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD}, which do not address temporal variability \citep{depalma_2018}.
-Other works have compared the trend observed within a reserve through time \citep{betti_2017-lq}, which cannot distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}.
-By accounting for trends between sites and through time, we can control for time and space dynamics, and provide a better identification of the effect.
+Some works evaluate marine reserves by performing inside-outside \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or before-after comparisons \citep{betti_2017-lq}.
+The first approach does not address temporal variability, and the second can not distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}.
+Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness.
+However, this method assumes control sites are a plausible counterfactual for treated sites.
+This supposed that treated sites would have followed the same trend as control sites, had the reserves not been implemented.
+Nonetheless, overall trends for each site don't show any significant increases, supporting our findings of lack of change in the indicators used (**S2 Figure**, **S3 Figure**, **S4 Figure**, **S5 Figure**, **S6 Figure**).
 
-Age, isolation, and enforcement are important factors influencing effectiveness of a marine reserve \citep{edgar_2014-UO}.
-Isla Natividad has the oldest reserve, is fairly isolated, and has a well-established community-based enforcement system.
-Neighbouring fishing communities are known to be well organized with successful resource management of their resources \citep{mccay_2017-1m,mccay_2014-CN}.
-Maria Elena and Punta Herrero are relatively young reserves (Table \ref{table:com_sum}).
-With the age, relative isolation, and enforcement level of these reserves, one would expect to observe effectiveness.
-However, another key feature of effective MRs is size \citep{edgar_2014-UO}; the lack of effectiveness is perhaps attributed to the reserves being too small \ref{table:com_sum}.
-Furthermore, perturbations that do not distinguish reserve boundaries, such as the environmental variability observed in Isla Natividad can also hinder effectiveness.
-The possibility of increasing reserve size or merging existing networks into a larger reserve should be evaluated.
+Literature shows that age, isolation, and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO}.
+Isla Natividad has the oldest reserve, and all communities are fairly isolated, and have a well-established community-based enforcement system.
+With these characteristics, one would expect the reserves to be effective.
+However, another key condition for effectiveness is reserve is size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to reserves being too small.
+Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
+Abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sesile invertebrates, but not lobsters.
+Maria Elena and Punta Herrero are relatively young reserves, and it is known that community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}.
+Work by \citet{ayer_2018} shows these communities support the implementation of marine reserves.
+Nevertheless, fishers may favor implementation of reserves that pose low fishing costs either because of their location or size.
+The economic data support our claim that reserves are small, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
 
-Our analysis of landings and revenues does not identify detectable changes in these indicators.
-However, previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
-Abalone are sesile invertebrates with less mobility (compared to lobsters), and thus current reserve size might not be enough for lobster's range of mobility even when accounting for reserve age in Isla Natividad.
-Other community-based marine reserves in tropical ecosystems have taken up to six years to show a spillover effect \citep{dasilva_2015-zX}.
-Reserves in Maria Elena and Punta Herrero are relatively small and young, and may need more time for abundances to increase enough to export larvae or adult organisms.
+Even if reserves had appropriate sizes, there are other plausible explanations for the observed lack of effectiveness.
+Marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2006}.
+However, both lobster fisheries were, at some point, certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
+Additionally, lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \cite{dof_website_1993}.
+It is uncertain whether that such a well-managed fishery will experience additional benefits from marine reserves.
 
-There are two plausible explanations for the observed lack of effectiveness.
-First, marine reserves are only likely to provide fisheries benefits if initial population sizes are low \citep{hilborn_2006} and are poorly managed.
-However, both stocks were at some point certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
-The fishery is managed via species-specific minimum catch sizes, seasonal closures, and escapement windows on traps \cite{dof_website_1993}.
-A second plausible explanation lies in the design of these areas.
-Small reserves are unlikely to have a significant effect on population size if they fail to protect a significant portion of it.
-Intuitvely, small reserves have little or no displacement of fishing effort, and fishers do not perceive the short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
-On either case, these might explain why reserves were not effective, but not why they still receive community support.
+While reserves fail to provide fishery benefits, there are a number of additional ecological, fisheries, and social benefits.
+Marine reserves provide protection to a wider range of species and vulnerable habitat, like coral reefs.
+These sites can serve as an insurance against environmental shocks or mistakes in fisheries management \citep{hilborn_2004,hilborn_2006,micheli_2012-EU}.
+Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital.
+Furthermore, showing commitment to marine conservation allows fishers to have greater bargaining power and leverage over fisheries management.
 
-While reserves fail to provide conservation or economic benefits, they provide the community with access to funding from the filantrhopic sector.
-Furthermore, reserves can provide a joint enterprise to bring a community together, which promotes social cohesion and social capital.
-
-A second explanation lies on the the mismatch of objectives during the design process.
-The Project description (*Estudios Técnicos Justificativos*) of each MR provides little information about the followed design guidelines.
-However, all reserves had to be approved by the community members.
-Depending on the communities’ capacity to look for longterm benefits, many communities may favor implementation of reserves on sites that represent a low fishing cost.
-As a consequence, the marine reserve have low impacts in reducng the fishing effort.
-Having small reserves in areas that do not compromise fishing profits might explain why this communities still support their reserves.
-
-Although our case studies fulfilled the social requirements for effective marine reserves ("high enforcement, presence of a management plan, fisher engagement in management, and promotion of sustainable fishing”; \citet{difranco_2016-Xw}), our results show that proper reserve design is crucial for effective marine reserves.
-The social-ecological systems framework allowed a systematic diganosis and compare all the different case studies, allowing us to identify and tease appart possible explanations \citep{basurto_2013-oq}.
-
-Citar Fulton Marine Policy
+Community-based marine reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented.
+Lessons learned from these cases can guide implementation of community-based marine reserves elsewhere.
+For the particular case of the marine reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
+At the broader scale, having full community support surely represents an advantage, but it is important for marine reserves to meet essential design principles such as size and placement.
+Community-based marine reserves might have more benefits that result from indirect effects of the reserves, which should be taken into account when evaluating the outcomes of similar projects.
 
@@
 
-JC and EA analyzed and interpreted data, discussed the results, and wrote the manuscript. AS, SF and JT edited the manuscript and discussed the results.
+JC and EA analyzed and interpreted data, discussed the results, and wrote the first draft. AS, SF and JT discussed the results and edited the manuscript.
 
 \section*{Funding}
-JCVD CONACyT + LAFF
-ASC
-SF
-JT
+
+JCVD CONACyT (Beca de Posgrados en el extranjero, CVU 669403) + LAFF + El proyecto de COBI que haya pagado esto
+
+**ASC**
+
+**SF**
+
+**JT**
+
+**Que fuentes de financiamiento incluyo par ustedes?**
 
@@
 \label{S2_Figure}
-Timeseries of indicators for IN
+Time series of biological indicators for IN
 
@@
 \label{S3_Figure}
-Timeseries of indicators for ME
+Time series of biological indicators for  ME
 
 \paragraph*{S4 Figure}
-\label{S3_Figure}
-Timeseries of indicators for PH
+\label{S4_Figure}
+Time series of biological indicators for  PH
+
+\paragraph*{S5 Figure}
+\label{S5_Figure}
+Time series of economic indicators for ME
+
+\paragraph*{S6 Figure}
+\label{S6_Figure}
+Time series of economic indicators for PH
 
@@
 \label{S1_Table}
-Coefficient estimates for Isla Natividad
+Coefficient estimates for biological indicators in Isla Natividad
 
@@
 \label{S2_Table}
-Coefficient estimates for Maria Elena
+Coefficient estimates for biological indicators in Maria Elena
 
@@
 \label{S3_Table}
-Coefficient estimates for Punta Herrero
+Coefficient estimates for biological indicators in Punta Herrero
+
+\paragraph*{S4 Table}
+\label{S4_Table}
+Coefficient estimates for economic indicators
 

### 2018-06-11 docs/paper/Villasenor-Derbez.tex
Updated manuscript and sent it to co-authors

@@
 
-\usepackage{amsthm}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{proposition}{Proposition}[section]
-\theoremstyle{definition}
-\newtheorem{example}{Example}[section]
-\theoremstyle{definition}
-\newtheorem{exercise}{Exercise}[section]
-\theoremstyle{remark}
-\newtheorem*{remark}{Remark}
-\newtheorem*{solution}{Solution}
 \begin{document}
@@
 
-\title[Community-based marine reserves]{Management and effectiveness of community-based marine reserves in
-small-scale fisheries}
+\title[Community-based marine reserves]{Effectiveness of community-based marine reserves in small-scale
+fisheries}
 
@@
 
-\textbf{Last update}: 2018-06-08
+\textbf{Titulos alternativos}
+
+``Management and effectiveness of community-based marine reserves in
+small-scale fisheries''
+
+``Community-based marine reserves in small-scale fisheries''
+
+``Effectiveness of community-based marine reserves: lessons for their
+management and implementation'' (no me encanta)
+
+Algun otro?
+
+\textbf{Last update}: 2018-06-11
 
@@
 manage the spatial distribution of fishing effort to recover stocks and
-preserve biodiversity is through the implementation of marine reserves
-(MRs). These areas allow bounded populations to recover by limiting all
-extractive activities \citep{halpern_2002}. The science of MRs has
-largely focused on understanding the ecological effects of these areas,
-which include increased biomass, richness, and densities of organisms
-within the protected regions, climate change mitigation, and protection
-from environmental variability
+preserve biodiversity is through the implementation of marine reserves.
+These areas allow bounded populations to recover by limiting all
+extractive activities \citep{halpern_2002}.
+
+The science of marine reserves has largely focused on understanding the
+ecological effects of these areas, which include increased biomass,
+richness, and densities of organisms within the protected regions,
+climate change mitigation, and protection from environmental variability
 \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
-Modelling studies show that the fishery benefits of marine reserves
-depend on initial stock status and the type of management under which
-the fishery operates \citep{hilborn_2006}, as well as reserve size and
-the amount of larvae exported from these \citep{krueck_2017-J1}. Other
-research has focused on the relationship between socioeconomic and
-governance structures and their relationship to ecological effectiveness
+Modelling studies show that fishery benefits of marine reserves depend
+on initial stock status and the management under which the fishery
+operates, as well as reserve size and the amount of larvae exported from
+these \citep{hilborn_2006,krueck_2017-J1}. Other research has focused on
+the relationship between socioeconomic and governance structures and
+their relationship to reserve effectiveness
 \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, few
-studies simultaneously evaluate MRs from all these perspectives
+studies simultaneously evaluate reserves from all these perspectives
 (\emph{e.g.} \citet{lpezangarita_2014}).
 
-While ecological factors like habitat representation within the MR or
-connectivity to other areas can determine the success of a MR, its
-effectiveness also depends on the socioeconomic and governance settings
-under which they are implemented and managed. Literature shows that many
-non-ecological characteristics can play an equally important role in the
-effectiveness of MRs. For example, age of a reserve (\emph{i.e.} time
-since its implementation) and size were key to the effectiveness of MRs
-in Palau \citep{friedlander_2017-oI}. In the Mediterranean,
-\citet{difranco_2016-Xw} identify that surveillance and enforcement,
-presence of a management plan, and involvement of fishers in management
-and decision--making along with promotion of sustainable fishing
-practices were the key factors that increased stock health and income to
-fishers. At a global level, enforcement, age, size, and isolation are
-important factors that determine the effectiveness of the reserves
-\citep{edgar_2014-UO}. However, MRs represent part of a more complex
-social-ecological systems governed by the interaction between humans and
-the environment. In this work we evaluate community-based marine
-reserves in Mexico using a triple bottom line approach by using
-ecological, socioeconomic, and governance indicators that provide a
-holistic view of the reserves \citep{halpern_2013}.
-
-There are three main approaches to implement MRs, each of which have its
-pros and cons. We describe them in the context of Mexican MRs, but argue
-that at least the first two apply elsewhere. MRs were historically
-implemented and managed by a government agency, in this case the
-National Commission of Protected Areas (\emph{Comisión Nacional de Áreas
-Marinas Protegidas}, CONANP). While CONANP has made efforts to have a
-participatory process, the implementation of these zones is still mainly
-a top-down process. A second approach is the implementation of
-community-based marine reserves within areas of exclusive access
-(\emph{i.e.} TURFs), thus making them TURF-reserves
-\citep{afflerbach_2014-HP}. Community-based spatial closures ocurr in
-other places, like the \emph{kapu} or \emph{ra'ui} areas in the Pacific
-Islands \citep{bohnsack_2004,johannes_2002}. This bottom-up approach
-increases compliance and self-enforcement
+Implementation of marine reserves can be done by following one of three
+main approaches in terms of implementation. We describe these in the
+context of Mexican MRs, but argue that at least the first two apply
+elsewhere. Marine reserves are usually implemented and managed by a
+government agency, in this case the National Commission of Natural
+Protected Areas (\emph{Comisión Nacional de Áreas Naturales Protegidas},
+CONANP). While CONANP has made efforts to incorporate stakeholders in
+the design and implementation phases, these are still characterized by a
+heavy top-down process. A second approach is the implementation of
+community-based marine reserves, usually placed within areas of
+exclusive access (\emph{i.e.} TURFs). Community-based spatial closures
+occur in other places, like the \emph{kapu} or \emph{ra'ui} areas in the
+Pacific Islands \citep{bohnsack_2004,johannes_2002}. This bottom-up
+approach increases compliance and self-enforcement
 \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
-these lack legal recognition and rely on the exclusive access granted by
-the TURF, making it impossible to enforce without one. In an effort to
-provide a legal framework for these reserves, Civil Society
-Organizations (CSOs) served as the bridge between fishers and
-government. In 2014, a new norm \citep{nom} allowed fishers to request
-the legal recognition of a community-based reserve under the name of
-``Fishing Refugia'' (\emph{Zona de Refugio Pesquero}). These can be
-implemented as temporal or partial eservs, which can protect one, some,
-or all resources within them. Since then, \textbf{39} of these have been
-implemented along the Pacific, Gulf of California, and Mexican Caribbean
-coastlines, but no formal evaluation of their effectiveness has taken
+without legal recognition these are difficult to enforce and fishers
+rely on the exclusive access granted by the TURF. In an effort to bridge
+this normative gap, Civil Society Organizations (CSOs) served as the
+link between fishers and government, and set out to create a legal
+framework that solve this governance issue. In 2014, a new norm was
+created, allowing fishers to request the legal recognition of a
+community-based reserve under the name of ``Fishing Refugia''
+\citep{nom}. These can be implemented as temporal or partial reserves,
+which can protect one, some, or all resources within them. Since then,
+\textbf{41} of community-based marine reserves along the Pacific, Gulf
+of California, and Mexican Caribbean coastlines have gained legal
+recognition, but no formal evaluation of their effectiveness has taken
 place.
@@
 both cases, we draw from the lessons learned and provide management
-recommendations to ensure or improve the effectiveness of
-community-based marine reserves.
+recommendations to maximize the effectiveness of community-based marine
+reserves in small-scale fisheries.
 
@@
 fishing cooperatives that hold Territorial Use Rights for Fisheries
-(TURFs). Isla Natividad (IN) lies west of the Baja California Peninsula
-(Fig \ref{fig:map}B), where kelp forests and rocky reefs are the
-predominant habitats. The island is home to the \emph{Sociedad
-Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja
-California}, whose main resource by value is the spiny lobster
-(\emph{Panulirus interruptus}). However, other resources like finfish
-(yellow-tail jack, \emph{Seriola lalandi}), sea cucumber
-(\emph{Parastichopus parvimensis}), red sea urchin (\emph{Mesocentrotus
-franciscanus}), snail (\emph{Megastraea turbanica} y \emph{M. undosa}),
-and abalone (\emph{Haliotis spp}) are also important sources of income.
-In 2006, the community decided to implement two community-based marine
-reserves within their fishing grounds to protect commercially important
-invertebrate species (mainly lobster and abalone). Until today, these
-reserves are yet to be legally recognized as Fishing Refugia but count
-with full support from the community.
+(TURFs). Isla Natividad lies west of the Baja California Peninsula (Fig
+\ref{fig:map}B), where kelp forests and rocky reefs are the predominant
+habitats. The island is home to the \emph{Buzos y Pescadores de la Baja
+California} fishing cooperative, whose main resource by value is the
+spiny lobster (\emph{Panulirus interruptus}). Other resources like
+finfish, sea cucumber, red sea urchin, snail, and abalone are also
+important sources of income. In 2006, the community decided to implement
+two community-based marine reserves within their fishing grounds to
+protect commercially important invertebrate species (mainly lobster and
+abalone). These reserves obtained legal recognition in 2018, but were
+well enforced since their implementation in 2006.
 
-The other two communities are Maria Elena (ME) and Punta Herrero (PH;
-Fig \ref{fig:map}C) in the Yucatan Peninsula, where coral reefs and
-mangroves are the representative coastal ecosystems. ME is a fishing
-camp --visited intermitently during the fishing season-- belonging to
-the Cozumel fishing cooperative (\emph{SCPP Cozumel}); PH is home to the
-\emph{SCPP José María Azcorra} cooperative. Their main fishery is the
-Caribbean spiny lobster (\emph{Panulirus argus}), but they also target
-finfish in the off season, mainly snappers (Lutjanidae) and groupers
-(Serranidae). ME established eight marine reserves in 2012, and PH
-established four marine reserves in 2013 and an additional
-community-based (\emph{i.e.} not legally recognized). All these reserves
-are legally recognized as Fishing Refugia.
+The other two communities are Maria Elena and Punta Herrero, which are
+located in the Yucatan Peninsula, where coral reefs and mangroves are
+the representative coastal ecosystems( Fig \ref{fig:map}C). Maria Elena
+is a fishing camp --visited intermittently during the fishing season--
+belonging to the Cozumel fishing cooperative (\emph{SCPP Cozumel});
+Punta Herrero is home to the \emph{SCPP José María Azcorra} cooperative.
+Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
+argus}), but they also target finfish in the off-season. Maria Elena and
+Punta Herrero established eight marine reserves in 2012, and four marine
+reserves in 2013, respectively. All these reserves are legally
+recognized as Fishing Refugia since their creation.
 
@@
 
-We use three main sources of information to evaluate these reserves.
+We use three main sources of information to evaluate these reserves
+across the ecological, socioeconomic, and governance dimensions.
 Ecological data come from the annual ecological monitoring of reserve
@@
 abundances of fish and invertebrate species in the reserves and control
-sites. Size structures are also collected for fish evaluations. We
-define control sites as regions with habitat characteristics similar to
-the corresponding reserves, and that presumably had a similar
-probability of being selected as reserves during the design phase. We
-focus our evaluation for sites where data are avaialable for reserve and
-control sites, before and after the implementation of the reserve. This
-provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
-sampling design that allows us to capture and control for temporal and
-spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and
-causal inference techniques have proven effective to evaluate marine
-reserves, as they allow us to causally attribute observed changes to the
-intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites
-were surveyed annually, and at least once before implementation of the
-reserves. Table \ref{table:com_sum} shows a summary of the reserves
-included in this study.
+sites. Size structures are also collected during fish surveys. We define
+control sites as regions with habitat characteristics similar to the
+corresponding reserves, and that presumably had a similar probability of
+being selected as reserves during the design phase. We focus our
+evaluation on sites where data are available for reserve and control
+sites, before and after the implementation of the reserve. This provides
+us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design
+that allows us to capture and control for temporal and spatial dynamics
+\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
+techniques have proven effective to evaluate marine reserves, as they
+allow us to causally attribute observed changes to the intervention
+\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
+annually, and at least once before implementation of the reserves. Table
+\ref{table:com_sum} shows a summary of the reserves included in this
+study.
 
@@
 Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
-(Kg) and value (MXP) from 2000 to 2014 for cooperatives with and without
-marine reserves (\textbf{Fig S1}). This information was aggregated by
-year, and economic values were adjusted by the Consumer Price Index
+(Kg) and revenues (MXP) from 2000 to 2014 for cooperatives with and
+without marine reserves(\textbf{Fig S1}). All cooperatives of each
+region (\emph{i.e.} Pacific and Caribbean) incorporated in this
+analysis, belong to larger Cooperative Federations, and are exposed to
+the same markets and institutional frameworks
+\citep{mccay_2017-1m,ayer_2018}, making them plausible controls.
+Landings and revenues were aggregated at the cooperative-year level, and
+revenues were adjusted by the Consumer Price Index for Mexico
 \citep{oecd_2017-VV} as:
@@
 creation and designation of the marine reserves
-\citep{dof_website_2012,dof_website_2013} and based on the authors'
-experience and knowledge of the communities. These include information
-on the resource system, the resource units, actors, and the governance
-system itself (\textbf{S1 Table}).
+\citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on
+the authors' experience and knowledge of the communities. These include
+information on the resource system, the resource units, actors, and the
+governance system itself (\textbf{S1 Table}).
 
@@
 
-We evaluate the effect that marine reserves have had on four biological
+We evaluate the effect that marine reserves have had on four ecological
 and two socioeconomic indicators (Table \ref{table:indicators}). Recall
 that reserves were implemented to protect lobster and other benthic
-invertebrates. However, we also use the available fish data, since
-finfish are targeted in the off-season and because there might be
-associated cobenefits of the reserve.
+invertebrates. However, we also use the available fish datato test for
+associated co-benefits.
 
@@
 \hline
-Biological & Lobster density & org / $m^2$\\
+Biological & Lobster density & org $\mathrm{m}^{-2}$\\
 \hline
-Biological & Invertebrate density & org / $m^2$\\
+Biological & Invertebrate density & org $\mathrm{m}^{-2}$\\
 \hline
-Biological & Fish biomass & Kg / $m^2$\\
+Biological & Fish biomass & Kg $\mathrm{m}^{-2}$\\
 \hline
-Biological & Fish density & org / $m^2$\\
+Biological & Fish density & org $\mathrm{m}^{-2}$\\
 \hline
@@
 
-We use a difference-in-differences analysis to evaluate the biological
+We use a difference-in-differences analysis to evaluate these
 indicators. This approach allows us to estimate the effect that the
-reserve has on the biological indicators by comparing trends across time
-and treatments (\emph{i.e.} reserve / control sites
-\citet{moland_2013-VP,Villasenor-Derbez_2018}). The analysis is
-performed with a multiple linear regression of the form:
+reserve had by comparing trends across time and treatments (\emph{i.e.}
+reserve / control sites \citet{moland_2013-VP,Villasenor-Derbez_2018}).
+The analysis of ecological indicators is performed with a multiple
+linear regression of the form:
 
@@
 of implementation in the control site (See Table \ref{table:com_sum}).
-When evaluating biomass and densities of the entire benthic or fsh
+When evaluating biomass and densities of the entire benthic or fish
 communities, we include \(\sigma_j\) to control for species-fixed
@@
 
-Socioeconomic indicators are evaluated with a similar approach (Eq
-\ref{eqn:soc_reg}). Due to data constrains, only Isla Natividad and
-Maria Elena are evaluated in this case. We constructed panel-data
-information with yearly (2001 - 2014) lobster landings and income of the
-studied and neighbouring communities that have similar management
-strategies, and belong to larger Cooperative Federtations
-\citep{mccay_2017-1m,ayer_2018}. Neighbouring communities are used as
-counterfacutals that allow us to control for unobserved time-invariants.
-Each ``treated'' community (Isla Natividad and Maria Elena) has three
+Socioeconomic indicators are evaluated with a similar approach. Due to
+data constrains, we only evaluate socioeconomic data for Isla Natividad
+and Maria Elena. Neighboring communities are used as counterfacutals
+that allow us to control for unobserved time-invariants. Each
+``treated'' community (Isla Natividad and Maria Elena) has three
 counterfactual communities.
@@
 heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
-analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep(R\_2018).
+analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.
 
@@
 an overview of the governance settings of each community, and discuss
-how these dimensions might be related to the effectiveness of the
+how these might be related to the effectiveness and performance of the
 reserves.
@@
 
-Indicators showed idiosyncratic responses through time for each
-community. Figure \ref{fig:indicators}A shows positive effect sizes for
-lobster densities for Isla Natividad and Punta Herrero during the first
-years, but the effect is eroded through time. These effects are in the
-order of 0.2 extra organisms \(m^{-2}\) but are not significantly
+Indicators showed ambiguous responses through time for each reserve.
+Figure \ref{fig:indicators}A shows positive effect sizes for lobster
+densities in Isla Natividad and Punta Herrero during the first years,
+but the effect is eroded through time. These effects are in the order of
+0.2 extra organisms \(\mathrm{m}^{-2}\) but are not significantly
 different from zero (\(p > 0.05\)). Lobster densities were only
@@
 Lobster landings and revenue were only available for Isla Natividad and
-Maria Elena (Fig \ref{fig:lobsters}). For all years befor
+Maria Elena (Fig \ref{fig:lobsters}). For all years before
 implementation, the effect sizes are close to zero, indicating that the
-control and treatment sites track each other well and that these are
-plausible controls. However, effect sizes do not change after the
-implementation of the reserve. Again, the negative coeficient observed
-for Isla Natividad on year 5 correspond to the 2011 hypoxia events. The
-only positive change observed in lobster landings is for Isla Natividad
-in 2014 (\(p < 0.1\)). The trhee years of post-implementation data for
-Maria Elena do not show a significant effect of the reserve. Isla
-Natividad shows higher revenues after the iumplementation of the
-reserve, as compared to the control communities. However, these changes
-are also not ignificantly different and present an increased variation.
-All regression coefficients for each community and indicator are
-presented in \textbf{S5 Table}.
+control and treatment sites have similar pre-treatment trends,
+asuggesting that these are plausible controls. However, effect sizes do
+not change after the implementation of the reserve. Again, the negative
+coefficient observed for Isla Natividad on year 5 correspond to the 2011
+hypoxia events. The only positive change observed in lobster landings is
+for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
+post-implementation data for Maria Elena do not show a significant
+effect of the reserve. Isla Natividad shows higher revenues after the
+implementation of the reserve, as compared to the control communities.
+However, these changes are not significant and are associated to
+increased variation. All regression coefficients for each community and
+indicator are presented in \textbf{S5 Table}.
 
@@
 Although we have little information on the social dimension of these
-fisheries, we can use the social-ecologycal systems framework
+fisheries, we can use the social-ecological systems framework
 (\textbf{S1 Table}) to analyze the performance of each governance system
-(\textbf{S6 Table}).
-
-Our analysis shows that all of the systems analyzed share similarities
-in their Governance system which is based on cooperatives (GS5.2.3.2),
-with strong rules in use that include Operational rules (GS6.2),
-Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even
-Territorial use communal rights (GS6.1.4.3). However, we identified
-important differences in terms of the actors, resource systems and
-resource units. The value of lobster in isla natividad is higher than
-the lobster sold from Punta Herrero and Maria Elena (RU4), which can
-reduce the pressure on harvest. Lastly, in terms of actors, although all
+(\textbf{S6 Table}). Our analysis shows that all of the systems analyzed
+share similarities in their Governance system which is based on
+cooperatives (GS5.2.3.2), with strong rules in use that include
+Operational rules (GS6.2), Collective-choice rules (GS6.3),
+Constitutional rules (GS6.3), and even Territorial use communal rights
+(GS6.1.4.3). However, we identified important differences in terms of
+the actors, resource systems, and resource units. Although all
 communities show a high level of leadership (A5), the level of trust
@@
 
-Our results indicate that marine reserves are ineffective in increaseing
-population sizes of lobsters in the three communities. The socioeconomic
-indicators pertaining to landings and revenues showed little to no
-change after reserve implementation. While lobster densities represent
-an indicator directly tied to reserve objectives, we also evaluate other
-biological indicators to test for additional effects and fail to detect
-any effect other then the already reported buffering effect that
-Natividad reserves can have to environmental variability
-\citep{micheli_2012-EU}. The lack of expected effectiveness poses the
-question, why do these communities continue to support the reserves?
-Understanding the social-ecological context in which these communities
-and their reserves operate might provide insights to this question.
-Here, we discuss plausible explanations to lack of effectiveness
-observed onthe biological and socioeconomic indicators based on our
-social-ecological system analysis. We also discuss potential
-shortcomings in our analysis, and provide management recommendations to
-improve reserve effectiveness.
+Our results indicate that marine reserves do not achieve the main
+objective of increasing lobster densities. No co-benefits were
+identified when using other ecological indicators other than the
+previously reported buffering effect that reserves can have to
+environmental variability in Isla Natividad \citep{micheli_2012-EU}. The
+socioeconomic indicators pertaining landings and revenues showed little
+to no change after reserve implementation. The lack of expected
+effectiveness poses the question: why do these communities continue to
+support the reserves? Understanding the social-ecological context in
+which these communities and their reserves operate might provide
+insights to this question. Here we touch on potential shortcomings in
+our analysis, and discuss our findings in the context of existing
+literature and our social-ecological system analysis. Finally, we
+provide recommendations to guide the implementation of new (and improve
+effectiveness of) community-based marine reserves in small-scale
+fisheries.
 
-Our approach to evaluate the temporal and spatial changes of each
-indicator provides a more robust measure of reserve effectiveness. Some
-works have solely focused on an inside-outside comparison of indicators
-\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD}, which do
-not address temporal variability \citep{depalma_2018}. Other works have
-compared the trend observed within a reserve through time
-\citep{betti_2017-lq}, which cannot distinguish between the temporal
-trends in a reserve and the entire system \citep{depalma_2018}. By
-accounting for trends between sites and through time, we can control for
-time and space dynamics, and provide a better identification of the
-effect.
+Some works evaluate marine reserves by performing inside-outside
+\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or
+before-after comparisons \citep{betti_2017-lq}. The first approach does
+not address temporal variability, and the second can not distinguish
+between the temporal trends in a reserve and the entire system
+\citep{depalma_2018}. Our approach to evaluate the temporal and spatial
+changes provides a more robust measure of reserve effectiveness.
+However, this method assumes control sites are a plausible
+counterfactual for treated sites. This supposed that treated sites would
+have followed the same trend as control sites, had the reserves not been
+implemented. Nonetheless, overall trends for each site don't show any
+significant increases, supporting our findings of lack of change in the
+indicators used (\textbf{S2 Figure}, \textbf{S3 Figure}, \textbf{S4
+Figure}, \textbf{S5 Figure}, \textbf{S6 Figure}).
 
-Age, isolation, and enforcement are important factors influencing
-effectiveness of a marine reserve \citep{edgar_2014-UO}. Isla Natividad
-has the oldest reserve, is fairly isolated, and has a well-established
-community-based enforcement system. Neighbouring fishing communities are
-known to be well organized with successful resource management of their
-resources \citep{mccay_2017-1m,mccay_2014-CN}. Maria Elena and Punta
-Herrero are relatively young reserves (Table \ref{table:com_sum}). With
-the age, relative isolation, and enforcement level of these reserves,
-one would expect to observe effectiveness. However, another key feature
-of effective MRs is size \citep{edgar_2014-UO}; the lack of
-effectiveness is perhaps attributed to the reserves being too small
-\ref{table:com_sum}. Furthermore, perturbations that do not distinguish
-reserve boundaries, such as the environmental variability observed in
-Isla Natividad can also hinder effectiveness. The possibility of
-increasing reserve size or merging existing networks into a larger
-reserve should be evaluated.
+Literature shows that age, isolation, and enforcement are important
+factors that influence reserve effectiveness \citep{edgar_2014-UO}. Isla
+Natividad has the oldest reserve, and all communities are fairly
+isolated, and have a well-established community-based enforcement
+system. With these characteristics, one would expect the reserves to be
+effective. However, another key condition for effectiveness is reserve
+is size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps
+be attributed to reserves being too small. Previous research has shown
+that reserves in Isla Natividad yield fishery benefits for the abalone
+fishery \citep{rossetto_2015-V0}. Abalone are less mobile than lobsters,
+and perhaps the reserves provide enough protection to these sesile
+invertebrates, but not lobsters. Maria Elena and Punta Herrero are
+relatively young reserves, and it is known that community-based marine
+reserves in tropical ecosystems may take up to six years to show a
+spillover effect \citep{dasilva_2015-zX}. Work by \citet{ayer_2018}
+shows these communities support the implementation of marine reserves.
+Nevertheless, fishers may favor implementation of reserves that pose low
+fishing costs either because of their location or size. The economic
+data support our claim that reserves are small, as neither landings nor
+revenues showed the expected short-term costs associated to the first
+years of reserve implementation \citep{ovando_2016-Wg}.
 
-Our analysis of landings and revenues does not identify detectable
-changes in these indicators. However, previous research has shown that
-reserves in Isla Natividad yield fishery benefits for the abalone
-fishery \citep{rossetto_2015-V0}. Abalone are sesile invertebrates with
-less mobility (compared to lobsters), and thus current reserve size
-might not be enough for lobster's range of mobility even when accounting
-for reserve age in Isla Natividad. Other community-based marine reserves
-in tropical ecosystems have taken up to six years to show a spillover
-effect \citep{dasilva_2015-zX}. Reserves in Maria Elena and Punta
-Herrero are relatively small and young, and may need more time for
-abundances to increase enough to export larvae or adult organisms.
+Even if reserves had appropriate sizes, there are other plausible
+explanations for the observed lack of effectiveness. Marine reserves are
+only likely to provide fisheries benefits if initial population sizes
+are low and the fishery is poorly managed \citep{hilborn_2006}. However,
+both lobster fisheries were, at some point, certified by the Marine
+Stewardship Council \citep{prezramrez_2016-J1}. Additionally, lobster
+fisheries are managed via species-specific minimum catch sizes, seasonal
+closures, protection of ``berried'' females, and escapement windows
+where traps are allowed \cite{dof_website_1993}. It is uncertain whether
+that such a well-managed fishery will experience additional benefits
+from marine reserves.
 
-There are two plausible explanations for the observed lack of
-effectiveness. First, marine reserves are only likely to provide
-fisheries benefits if initial population sizes are low
-\citep{hilborn_2006} and are poorly managed. However, both stocks were
-at some point certified by the Marine Stewardship Council
-\citep{prezramrez_2016-J1}. The fishery is managed via species-specific
-minimum catch sizes, seasonal closures, and escapement windows on traps
-\cite{dof_website_1993}. A second plausible explanation lies in the
-design of these areas. Small reserves are unlikely to have a significant
-effect on population size if they fail to protect a significant portion
-of it. Intuitvely, small reserves have little or no displacement of
-fishing effort, and fishers do not perceive the short-term costs
-associated to the first years of reserve implementation
-\citep{ovando_2016-Wg}. On either case, these might explain why reserves
-were not effective, but not why they still receive community support.
+While reserves fail to provide fishery benefits, there are a number of
+additional ecological, fisheries, and social benefits. Marine reserves
+provide protection to a wider range of species and vulnerable habitat,
+like coral reefs. These sites can serve as an insurance against
+environmental shocks or mistakes in fisheries management
+\citep{hilborn_2004,hilborn_2006,micheli_2012-EU}. Embarking in a marine
+conservation project can bring the community together, which promotes
+social cohesion and builds social capital. Furthermore, showing
+commitment to marine conservation allows fishers to have greater
+bargaining power and leverage over fisheries management.
 
-While reserves fail to provide conservation or economic benefits, they
-provide the community with access to funding from the filantrhopic
-sector. Furthermore, reserves can provide a joint enterprise to bring a
-community together, which promotes social cohesion and social capital.
-
-A second explanation lies on the the mismatch of objectives during the
-design process. The Project description (\emph{Estudios Técnicos
-Justificativos}) of each MR provides little information about the
-followed design guidelines. However, all reserves had to be approved by
-the community members. Depending on the communities' capacity to look
-for longterm benefits, many communities may favor implementation of
-reserves on sites that represent a low fishing cost. As a consequence,
-the marine reserve have low impacts in reducng the fishing effort.
-Having small reserves in areas that do not compromise fishing profits
-might explain why this communities still support their reserves.
-
-Although our case studies fulfilled the social requirements for
-effective marine reserves (``high enforcement, presence of a management
-plan, fisher engagement in management, and promotion of sustainable
-fishing''; \citet{difranco_2016-Xw}), our results show that proper
-reserve design is crucial for effective marine reserves. The
-social-ecological systems framework allowed a systematic diganosis and
-compare all the different case studies, allowing us to identify and
-tease appart possible explanations \citep{basurto_2013-oq}.
+Community-based marine reserves in small-scale fisheries can be helpful
+conservation and fishery management tools when appropriately
+implemented. Lessons learned from these cases can guide implementation
+of community-based marine reserves elsewhere. For the particular case of
+the marine reserves that we evaluate, the possibility of expanding
+reserves or merging existing polygons into larger areas should be
+evaluated and proposed to the communities. At the broader scale, having
+full community support surely represents an advantage, but it is
+important for marine reserves to meet essential design principles such
+as size and placement. Community-based marine reserves might have more
+benefits that result from indirect effects of the reserves, which should
+be taken into account when evaluating the outcomes of similar projects.
 
@@
 JC and EA analyzed and interpreted data, discussed the results, and
-wrote the manuscript. AS, SF and JT edited the manuscript and discussed
-the results.
+wrote the first draft. AS, SF and JT discussed the results and edited
+the manuscript.
 
@@
 
-JCVD CONACyT + LAFF ASC SF JT
+JCVD CONACyT (Beca de Posgrados en el extranjero, CVU 669403) + LAFF +
+El proyecto de COBI que haya pagado esto
+
+\textbf{ASC}
+
+\textbf{SF}
+
+\textbf{JT}
+
+\textbf{Que fuentes de financiamiento incluyo par ustedes?}
 
@@
 
-Timeseries of indicators for IN
+Time series of biological indicators for IN
 
@@
 
-Timeseries of indicators for ME
+Time series of biological indicators for ME
 
 \paragraph*{S4 Figure}
-\label{S3_Figure}
+\label{S4_Figure}
 
-Timeseries of indicators for PH
+Time series of biological indicators for PH
+
+\paragraph*{S5 Figure}
+\label{S5_Figure}
+
+Time series of economic indicators for ME
+
+\paragraph*{S6 Figure}
+\label{S6_Figure}
+
+Time series of economic indicators for PH
 
@@
 
-Coefficient estimates for Isla Natividad
+Coefficient estimates for biological indicators in Isla Natividad
 
@@
 
-Coefficient estimates for Maria Elena
+Coefficient estimates for biological indicators in Maria Elena
 
@@
 
-Coefficient estimates for Punta Herrero
+Coefficient estimates for biological indicators in Punta Herrero
+
+\paragraph*{S4 Table}
+\label{S4_Table}
+
+Coefficient estimates for economic indicators
 

### 2018-06-30 docs/paper/Villasenor-Derbez.Rmd
Added comments by coauthors

@@
 rtitle: "Community-based marine reserves"
-author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Álvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
+author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
 institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
@@
 
-<!-- Old Abstract -->
-<!-- Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
-However, overfishing and unsustainable fishing practices threaten the marine environment and jeopardize the wellbeing of coastal communities.
-A common approach to protect the environment and recover overexploited stocks is to implement no-take marine reserves (areas where all extractive activities are off-limits).
-In small-scale fisheries, these are sometimes implemented as community-based reserves, where a group of fishers collectively agree to close an area to fishing.
-While we know that reductions in fishing effort are followed by a series of ecological benefits (increased biomass, abundance, and species diversity), we do not fully understand how environmental and governance dynamics influence the conservation and fisheries benefits of community-based marine reserves.
-In this work, we evaluate the ecological outcomes of four reserves established by three coastal communities in temperate and tropical ecosystems of Mexico.
-By combining causal inference techniques with an operationalization of the social-ecological systems framework, we identify the environmental and social conditions that enable reserve effectiveness.
-Our results show a strong interaction between environmental variation and community organization, which influences reserve effectiveness.
-For example, the most effective reserve had strong governance structures accompanied with low environmental variability.
-Thus, even when a community is well organized (and reserves are well enforced), environmental variation can hinder the benefits of a reserve, and vice versa.
-Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making.
--->
-
 ```{r setup, include=FALSE}
@@
 ```
-
-**Titulos alternativos**
-
-"Management and effectiveness of community-based marine reserves in small-scale fisheries"
-
-"Community-based marine reserves in small-scale fisheries"
-
-"Effectiveness of community-based marine reserves: lessons for their management and implementation" (no me encanta)
-
-Algun otro?
 
@@
 Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}.
-A common approach to manage the spatial distribution of fishing effort to recover stocks and preserve biodiversity is through the implementation of marine reserves.
-These areas allow bounded populations to recover by limiting all extractive activities \citep{halpern_2002}.
+In particular, artisanal fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
+Recent research shows that combining Territorial Use Rights for Fisheries (TURFs) with no-take marine reserves (MR) can greatly improve the performance of coastal fisheries and the health of the local resouces \citep{costello_2010-Ix,lester_2017}.
+Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintainance of healthy resources \citep{afflerbach_2014-HP,lester_2017}.
 
+Marine reserves allow bounded populations to recover by limiting all extractive activities \citep{halpern_2002}, while the TURF controls the harvest levels outside the MPA.
+Although in theory these systems are successful \citep{costello_2010-Ix}, little empirical evidence evidence exists of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017}.
+The performance of these systems depends on how environmental and social factors work combined.
 The science of marine reserves has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
@@
 Other research has focused on the relationship between socioeconomic and governance structures and their relationship to reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
-However, few studies simultaneously evaluate reserves from all these perspectives (*e.g.* \citet{lpezangarita_2014}).
+However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.
 
-<!-- While ecological factors like habitat representation within the MR or connectivity to other areas can determine the success of a MR, its effectiveness also depends on the socioeconomic and governance settings under which they are implemented and managed.
-Literature shows that many non-ecological characteristics can play an equally important role in the effectiveness of MRs.
-For example, age of a reserve (*i.e.* time since its implementation) and size were key to the effectiveness of MRs in Palau \citep{friedlander_2017-oI}.
-In the Mediterranean, \citet{difranco_2016-Xw} identify that surveillance and enforcement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key factors that increased stock health and income to fishers.
-At a global level, enforcement, age, size, and isolation are important factors that determine the effectiveness of the reserves \citep{edgar_2014-UO}.
-However, MRs represent part of a more complex social-ecological systems governed by the interaction between humans and the environment.
-In this work we evaluate community-based marine reserves in Mexico using a triple bottom line approach by using ecological, socioeconomic, and governance indicators that provide a holistic view of the reserves \citep{halpern_2013}.
--->
+In Mexico many of these marine reserves are created as community-based marine reserves.
+Community-based spatial closures occur in other places, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
+This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
+However, without legal recognition these are difficult to enforce and fishers rely on the exclusive access granted by the TURF.
+In an effort to bridge this normative gap, Civil Society Organizations (CSOs) served as the link between fishers and government, and set out to create a legal framework that solve this governance issue.
+In 2014, a new norm was created, allowing fishers to request the legal recognition of a community-based reserve under the name of "Fish Refuge" \citep{nom}.
+These can be implemented as temporal or partial reserves, which can protect one, some, or all resources within them.
+Since then, 45 of community-based marine reserves along the Pacific, Gulf of California, and Mexican Caribbean coastlines have gained legal recognition, but their effectiveness has not been reported in the scientific literature.
 
-Implementation of marine reserves can be done by following one of three main approaches in terms of implementation.
+This work combines causal inference techniques and the social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico.
+The objective of this work is twofold.
+First, provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries.
+And second, evaluate the effectiveness of TURF-reserves established as Fishing Refugia in Mexico to identify areas where improvement or adjustment might result in increased effectiveness.
+On both cases, we draw from the lessons learned and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries.
+
+<!-- Implementation of marine reserves can be done by following one of three main approaches in terms of implementation.
 We describe these in the context of Mexican MRs, but argue that at least the first two apply elsewhere.
@@
 A second approach is the implementation of community-based marine reserves, usually placed within areas of exclusive access (*i.e.* TURFs).
-Community-based spatial closures occur in other places, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
-This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
-However, without legal recognition these are difficult to enforce and fishers rely on the exclusive access granted by the TURF.
-In an effort to bridge this normative gap, Civil Society Organizations (CSOs) served as the link between fishers and government, and set out to create a legal framework that solve this governance issue.
-In 2014, a new norm was created, allowing fishers to request the legal recognition of a community-based reserve under the name of "Fishing Refugia" \citep{nom}.
-These can be implemented as temporal or partial reserves, which can protect one, some, or all resources within them.
-Since then, **41** of community-based marine reserves along the Pacific, Gulf of California, and Mexican Caribbean coastlines have gained legal recognition, but no formal evaluation of their effectiveness has taken place.
-
-This work combines causal inference techniques and the social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico.
-The objective of this work is twofold.
-First, provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries.
-And second, perform the first formal evaluation of Fishing Refugia in Mexico and identify areas where improvement or adjustment might result in increased effectiveness.
-On both cases, we draw from the lessons learned and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries.
+-->
 
@@
 
-We evaluate community-based marine reserves from three coastal communities located in the Pacific coast of Baja California and the Mexican Caribbean (Fig \ref{fig:map}).
-All communities are organized as fishing cooperatives that hold Territorial Use Rights for Fisheries (TURFs).
-Isla Natividad lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests and rocky reefs are the predominant habitats.
-The island is home to the *Buzos y Pescadores de la Baja California* fishing cooperative, whose main resource by value is the spiny lobster (*Panulirus interruptus*).
-Other resources like finfish, sea cucumber, red sea urchin, snail, and abalone are also important sources of income.
-In 2006, the community decided to implement two community-based marine reserves within their fishing grounds to protect commercially important invertebrate species (mainly lobster and abalone).
-These reserves obtained legal recognition in 2018, but were well enforced since their implementation in 2006.
+We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A).
+The first one was created by the *Buzos y Pescadores de la Baja California* fishing cooperative, located in Isla Natividad in the Baja Peninsula (Fig \ref{fig:map}B).
+The main fishery in the island is the spiny lobster (*Panulirus interruptus*), but other resources like finfish, sea cucumber, read sea urchin, snail, and abalone are also an important source of income.
+In 2006, the community decided to implement two marine reserves within their fishing grounds to protect commercially important invertebrate species; mainly lobster and abalone.
+The reserves obtained legal recognition in 2018, but have been well enforced since their implementation.
 
-The other two communities are Maria Elena and Punta Herrero, which are located in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems( Fig \ref{fig:map}C).
+The other two TURF-reserves are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C).
 Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); Punta Herrero is home to the *SCPP José María Azcorra* cooperative.
@@
 invert <- read.csv(file = here::here("data", "invertebrates.csv"),
-                stringsAsFactors = T)
+                stringsAsFactors = F)
 
 fish <- read.csv(file = here::here("data", "fish.csv"),
-                stringsAsFactors = T)
+                stringsAsFactors = F)
 
 conapesca <- read.csv(file = here::here("data", "conapesca.csv"),
-                stringsAsFactors = T)
+                stringsAsFactors = F)
 
@@
                 strip.white = T,
-                stringsAsFactors = T)
+                stringsAsFactors = F)
 ```
@@
 Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
-Trained divers record richness and abundances of fish and invertebrate species in the reserves and control sites.
+Trained divers record richness and abundances of fish and invertebrate species in the reserves and control sites \citep{fulton_2018}.
 Size structures are also collected during fish surveys.
@@
 ```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A).
-Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are located in the Yucatan Peninsula.
+Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena and Punta Herrero (C) are located in the Yucatan Peninsula.
 Blue polygons represent the TURFs, and red polygons the marine reserves."}
@@
 Recall that reserves were implemented to protect lobster and other benthic invertebrates.
-However, we also use the available fish datato test for associated co-benefits.
+However, we also use the available fish data to test for associated co-benefits.
 
@@
 L_IN <- invert %>% 
+  spread(generoespecie, abundancia, fill = 0) %>% 
+  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>%
   filter(generoespecie == "Panulirus interruptus") %>% 
@@
 L_ME <- invert %>% 
+  spread(generoespecie, abundancia, fill = 0) %>% 
+  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>% 
   filter(generoespecie == "Panulirus argus") %>% 
@@
 L_PH <- invert %>% 
+  spread(generoespecie, abundancia, fill = 0) %>% 
+  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>% 
   filter(generoespecie == "Panulirus argus") %>% 
@@
 Figure \ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time.
-These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ but are not significantly different from zero ($p > 0.05$).
-Lobster densities were only significantly positive for Isla Natividad on the sixth year (*i.e.* 2012; $p < 0.05$), a year after the hypoxia events described by \citet{micheli_2012-EU} caused mass mortality of organisms.
+In the case of Maria Elena, positive changes were observed in the third and forth year.
+These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ for Isla Natividad and Punta Herrero, and 0.01 organisms $\mathrm{m}^{-2}$ for Maria Elena, but are not significantly different from zero ($p > 0.05$).
+The rapid increase observed for changes in lobster densities for Isla Natividad on the sixth year (*i.e.* 2012) occur a year after the hypoxia events described by \citet{micheli_2012-EU} caused mass mortality of organisms.
 Likewise, no changes were detected in fish biomass or invertebrate and fish densities (\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
@@
   filter(zona == "PN") %>%
-  lm(pesovivo/1e3 ~ grupo * years_centered + unidadeconomica, data = .)
+  lm(peso_vivo/1e3 ~ grupo * years_centered + unidad_economica, data = .)
 
@@
   filter(zona == "MC") %>%
-  lm(pesovivo/1e3 ~ grupo * years_centered + unidadeconomica, data = .)
+  lm(peso_vivo/1e3 ~ grupo * years_centered + unidad_economica, data = .)
 
@@
   filter(zona == "PN") %>%
-  lm(valor/1e6 ~ grupo * years_centered + unidadeconomica, data = .)
+  lm(valor/1e6 ~ grupo * years_centered + unidad_economica, data = .)
 
@@
   filter(zona == "MC") %>%
-  lm(valor/1e6 ~ grupo * years_centered + unidadeconomica, data = .)
+  lm(valor/1e6 ~ grupo * years_centered + unidad_economica, data = .)
 
@@
 
-Our results indicate that marine reserves do not achieve the main objective of increasing lobster densities.
-No co-benefits were identified when using other ecological indicators other than the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \citep{micheli_2012-EU}.
+Our results indicate that these TURF-reserves have not increased lobster densities.
+Additionally, no co-benefits were identified when using other ecological indicators other than the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \citep{micheli_2012-EU}.
 The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
-The lack of expected effectiveness poses the question: why do these communities continue to support the reserves?
-Understanding the social-ecological context in which these communities and their reserves operate might provide insights to this question.
-Here we touch on potential shortcomings in our analysis, and discuss our findings in the context of existing literature and our social-ecological system analysis.
-Finally, we provide recommendations to guide the implementation of new (and improve effectiveness of) community-based marine reserves in small-scale fisheries.
+The lack of evidence of the effectiveness of these reserves is surprising since most of the communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}.
+Analyzing the shortcomings of our study and understanding the social-ecological context in which these communities and their reserves operate might provide insights to this question.
 
@@
 
-Literature shows that age, isolation, and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO}.
-Isla Natividad has the oldest reserve, and all communities are fairly isolated, and have a well-established community-based enforcement system.
+Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO}.
+Isla Natividad has the oldest reserve, and our SES analysis suggests that all communities have a well-established community-based enforcement system.
 With these characteristics, one would expect the reserves to be effective.
-However, another key condition for effectiveness is reserve is size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to reserves being too small.
+Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 5 years old); other community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}.
+
+Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to reserves being too small.
 Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
 Abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sesile invertebrates, but not lobsters.
-Maria Elena and Punta Herrero are relatively young reserves, and it is known that community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}.
-Work by \citet{ayer_2018} shows these communities support the implementation of marine reserves.
-Nevertheless, fishers may favor implementation of reserves that pose low fishing costs either because of their location or size.
-The economic data support our claim that reserves are small, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
+Design principles developed by \citet{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
+Furthermore, may favor implementation of reserves that pose low fishing costs due to their small size or location.
+Our analysis of economic data supports this, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
+Small reserves can serve as a way to self-regulate the spatial distribution of fishing effort
 
-Even if reserves had appropriate sizes, there are other plausible explanations for the observed lack of effectiveness.
-Marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2006}.
-However, both lobster fisheries were, at some point, certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
+Even if reserves had appropriate sizes and were placed in optimal locations, there are other plausible explanations for the observed patterns.
+For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2006}.
+Both lobster fisheries were, at some point, certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
 Additionally, lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \cite{dof_website_1993}.
-It is uncertain whether that such a well-managed fishery will experience additional benefits from marine reserves.
+It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves.
 
@@
 These sites can serve as an insurance against environmental shocks or mistakes in fisheries management \citep{hilborn_2004,hilborn_2006,micheli_2012-EU}.
+Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
 Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital.
@@
 
-JCVD CONACyT (Beca de Posgrados en el extranjero, CVU 669403) + LAFF + El proyecto de COBI que haya pagado esto
+JCVD: CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the Latin American Fisheries Fellowship.
 
-**ASC**
-
-**SF**
-
-**JT**
-
-**Que fuentes de financiamiento incluyo par ustedes?**
+AS, SF and JT Walton Family Foundation, Summit Foundation, and Oak Foundation.
 

### 2018-06-30 docs/paper/Villasenor-Derbez.tex
Added comments by coauthors

@@
 \def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
-Aceves-Bueno\(^{1,*}\), Álvin Suarez-Castillo\(^{2}\), Stuart
-Fulton\(^{2}\), Jorge Torre\(^{2}\)}
+Aceves-Bueno\(^{1,*}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
+Jorge Torre\(^{2}\)}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 
-\textbf{Titulos alternativos}
-
-``Management and effectiveness of community-based marine reserves in
-small-scale fisheries''
-
-``Community-based marine reserves in small-scale fisheries''
-
-``Effectiveness of community-based marine reserves: lessons for their
-management and implementation'' (no me encanta)
-
-Algun otro?
-
-\textbf{Last update}: 2018-06-11
+\textbf{Last update}: 2018-06-29
 
@@
 overfishing and unsustainable fishing practices
-\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. A common approach to
-manage the spatial distribution of fishing effort to recover stocks and
-preserve biodiversity is through the implementation of marine reserves.
-These areas allow bounded populations to recover by limiting all
-extractive activities \citep{halpern_2002}.
+\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
+artisanal fisheries face great challenges since they tend to be hard to
+monitor and enforce \citep{costello_2012}. Recent research shows that
+combining Territorial Use Rights for Fisheries (TURFs) with no-take
+marine reserves (MR) can greatly improve the performance of coastal
+fisheries and the health of the local resouces
+\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
+these systems increase the benefits of spatial access rights allowing
+the maintainance of healthy resources
+\citep{afflerbach_2014-HP,lester_2017}.
 
-The science of marine reserves has largely focused on understanding the
+Marine reserves allow bounded populations to recover by limiting all
+extractive activities \citep{halpern_2002}, while the TURF controls the
+harvest levels outside the MPA. Although in theory these systems are
+successful \citep{costello_2010-Ix}, little empirical evidence evidence
+exists of their effectiveness and the drivers of their success
+\citep{afflerbach_2014-HP,lester_2017}. The performance of these systems
+depends on how environmental and social factors work combined. The
+science of marine reserves has largely focused on understanding the
 ecological effects of these areas, which include increased biomass,
@@
 their relationship to reserve effectiveness
-\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, few
-studies simultaneously evaluate reserves from all these perspectives
-(\emph{e.g.} \citet{lpezangarita_2014}).
+\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
+knowledge, no studies exist that evaluate TURF-reserves from both a
+social and ecological perspective.
 
-Implementation of marine reserves can be done by following one of three
-main approaches in terms of implementation. We describe these in the
-context of Mexican MRs, but argue that at least the first two apply
-elsewhere. Marine reserves are usually implemented and managed by a
-government agency, in this case the National Commission of Natural
-Protected Areas (\emph{Comisión Nacional de Áreas Naturales Protegidas},
-CONANP). While CONANP has made efforts to incorporate stakeholders in
-the design and implementation phases, these are still characterized by a
-heavy top-down process. A second approach is the implementation of
-community-based marine reserves, usually placed within areas of
-exclusive access (\emph{i.e.} TURFs). Community-based spatial closures
-occur in other places, like the \emph{kapu} or \emph{ra'ui} areas in the
-Pacific Islands \citep{bohnsack_2004,johannes_2002}. This bottom-up
-approach increases compliance and self-enforcement
+In Mexico many of these marine reserves are created as community-based
+marine reserves. Community-based spatial closures occur in other places,
+like the \emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
+\citep{bohnsack_2004,johannes_2002}. This bottom-up approach increases
+compliance and self-enforcement
 \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
@@
 created, allowing fishers to request the legal recognition of a
-community-based reserve under the name of ``Fishing Refugia''
-\citep{nom}. These can be implemented as temporal or partial reserves,
-which can protect one, some, or all resources within them. Since then,
-\textbf{41} of community-based marine reserves along the Pacific, Gulf
-of California, and Mexican Caribbean coastlines have gained legal
-recognition, but no formal evaluation of their effectiveness has taken
-place.
+community-based reserve under the name of ``Fish Refuge'' \citep{nom}.
+These can be implemented as temporal or partial reserves, which can
+protect one, some, or all resources within them. Since then, 45 of
+community-based marine reserves along the Pacific, Gulf of California,
+and Mexican Caribbean coastlines have gained legal recognition, but
+their effectiveness has not been reported in the scientific literature.
 
@@
 the effectiveness of community-based marine reserves that can inform
-similar processes in other countries. And second, perform the first
-formal evaluation of Fishing Refugia in Mexico and identify areas where
-improvement or adjustment might result in increased effectiveness. On
-both cases, we draw from the lessons learned and provide management
-recommendations to maximize the effectiveness of community-based marine
-reserves in small-scale fisheries.
+similar processes in other countries. And second, evaluate the
+effectiveness of TURF-reserves established as Fishing Refugia in Mexico
+to identify areas where improvement or adjustment might result in
+increased effectiveness. On both cases, we draw from the lessons learned
+and provide management recommendations to maximize the effectiveness of
+community-based marine reserves in small-scale fisheries.
 
@@
 
-We evaluate community-based marine reserves from three coastal
-communities located in the Pacific coast of Baja California and the
-Mexican Caribbean (Fig \ref{fig:map}). All communities are organized as
-fishing cooperatives that hold Territorial Use Rights for Fisheries
-(TURFs). Isla Natividad lies west of the Baja California Peninsula (Fig
-\ref{fig:map}B), where kelp forests and rocky reefs are the predominant
-habitats. The island is home to the \emph{Buzos y Pescadores de la Baja
-California} fishing cooperative, whose main resource by value is the
-spiny lobster (\emph{Panulirus interruptus}). Other resources like
-finfish, sea cucumber, red sea urchin, snail, and abalone are also
-important sources of income. In 2006, the community decided to implement
-two community-based marine reserves within their fishing grounds to
-protect commercially important invertebrate species (mainly lobster and
-abalone). These reserves obtained legal recognition in 2018, but were
-well enforced since their implementation in 2006.
+We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A). The
+first one was created by the \emph{Buzos y Pescadores de la Baja
+California} fishing cooperative, located in Isla Natividad in the Baja
+Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the
+spiny lobster (\emph{Panulirus interruptus}), but other resources like
+finfish, sea cucumber, read sea urchin, snail, and abalone are also an
+important source of income. In 2006, the community decided to implement
+two marine reserves within their fishing grounds to protect commercially
+important invertebrate species; mainly lobster and abalone. The reserves
+obtained legal recognition in 2018, but have been well enforced since
+their implementation.
 
-The other two communities are Maria Elena and Punta Herrero, which are
-located in the Yucatan Peninsula, where coral reefs and mangroves are
-the representative coastal ecosystems( Fig \ref{fig:map}C). Maria Elena
-is a fishing camp --visited intermittently during the fishing season--
+The other two TURF-reserves are located in Maria Elena and Punta
+Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). Maria Elena is a
+fishing camp --visited intermittently during the fishing season--
 belonging to the Cozumel fishing cooperative (\emph{SCPP Cozumel});
@@
 abundances of fish and invertebrate species in the reserves and control
-sites. Size structures are also collected during fish surveys. We define
-control sites as regions with habitat characteristics similar to the
-corresponding reserves, and that presumably had a similar probability of
-being selected as reserves during the design phase. We focus our
-evaluation on sites where data are available for reserve and control
-sites, before and after the implementation of the reserve. This provides
-us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design
-that allows us to capture and control for temporal and spatial dynamics
+sites \citep{fulton_2018}. Size structures are also collected during
+fish surveys. We define control sites as regions with habitat
+characteristics similar to the corresponding reserves, and that
+presumably had a similar probability of being selected as reserves
+during the design phase. We focus our evaluation on sites where data are
+available for reserve and control sites, before and after the
+implementation of the reserve. This provides us with a
+Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
+allows us to capture and control for temporal and spatial dynamics
 \citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
@@
 coastal communities studied (A). Isla Natividad (B) is located off the
-Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are
-located in the Yucatan Peninsula. Blue polygons represent the TURFs, and
-red polygons the marine reserves.}
+Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
+in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
+polygons the marine reserves.}
 \end{figure}
@@
 that reserves were implemented to protect lobster and other benthic
-invertebrates. However, we also use the available fish datato test for
+invertebrates. However, we also use the available fish data to test for
 associated co-benefits.
@@
 densities in Isla Natividad and Punta Herrero during the first years,
-but the effect is eroded through time. These effects are in the order of
-0.2 extra organisms \(\mathrm{m}^{-2}\) but are not significantly
-different from zero (\(p > 0.05\)). Lobster densities were only
-significantly positive for Isla Natividad on the sixth year (\emph{i.e.}
-2012; \(p < 0.05\)), a year after the hypoxia events described by
-\citet{micheli_2012-EU} caused mass mortality of organisms. Likewise, no
-changes were detected in fish biomass or invertebrate and fish densities
+but the effect is eroded through time. In the case of Maria Elena,
+positive changes were observed in the third and forth year. These
+effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
+Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
+for Maria Elena, but are not significantly different from zero
+(\(p > 0.05\)). The rapid increase observed for changes in lobster
+densities for Isla Natividad on the sixth year (\emph{i.e.} 2012) occur
+a year after the hypoxia events described by \citet{micheli_2012-EU}
+caused mass mortality of organisms. Likewise, no changes were detected
+in fish biomass or invertebrate and fish densities
 (\ref{fig:indicators}B-D), where effect sizes oscillated around zero
@@
 
-Our results indicate that marine reserves do not achieve the main
-objective of increasing lobster densities. No co-benefits were
-identified when using other ecological indicators other than the
-previously reported buffering effect that reserves can have to
-environmental variability in Isla Natividad \citep{micheli_2012-EU}. The
-socioeconomic indicators pertaining landings and revenues showed little
-to no change after reserve implementation. The lack of expected
-effectiveness poses the question: why do these communities continue to
-support the reserves? Understanding the social-ecological context in
-which these communities and their reserves operate might provide
-insights to this question. Here we touch on potential shortcomings in
-our analysis, and discuss our findings in the context of existing
-literature and our social-ecological system analysis. Finally, we
-provide recommendations to guide the implementation of new (and improve
-effectiveness of) community-based marine reserves in small-scale
-fisheries.
+Our results indicate that these TURF-reserves have not increased lobster
+densities. Additionally, no co-benefits were identified when using other
+ecological indicators other than the previously reported buffering
+effect that reserves can have to environmental variability in Isla
+Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
+pertaining landings and revenues showed little to no change after
+reserve implementation. The lack of evidence of the effectiveness of
+these reserves is surprising since most of the communities show a
+positive perception about their performance and continue to support
+their presence \citep{ayer_2018}. Analyzing the shortcomings of our
+study and understanding the social-ecological context in which these
+communities and their reserves operate might provide insights to this
+question.
 
@@
 
-Literature shows that age, isolation, and enforcement are important
-factors that influence reserve effectiveness \citep{edgar_2014-UO}. Isla
-Natividad has the oldest reserve, and all communities are fairly
-isolated, and have a well-established community-based enforcement
-system. With these characteristics, one would expect the reserves to be
-effective. However, another key condition for effectiveness is reserve
-is size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps
-be attributed to reserves being too small. Previous research has shown
-that reserves in Isla Natividad yield fishery benefits for the abalone
+Literature shows that age and enforcement are important factors that
+influence reserve effectiveness \citep{edgar_2014-UO}. Isla Natividad
+has the oldest reserve, and our SES analysis suggests that all
+communities have a well-established community-based enforcement system.
+With these characteristics, one would expect the reserves to be
+effective. Maria Elena and Punta Herrero are relatively young reserves
+(\emph{i.e.} \textless{} 5 years old); other community-based marine
+reserves in tropical ecosystems may take up to six years to show a
+spillover effect \citep{dasilva_2015-zX}.
+
+Another key condition for effectiveness is reserve size
+\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
+attributed to reserves being too small. Previous research has shown that
+reserves in Isla Natividad yield fishery benefits for the abalone
 fishery \citep{rossetto_2015-V0}. Abalone are less mobile than lobsters,
 and perhaps the reserves provide enough protection to these sesile
-invertebrates, but not lobsters. Maria Elena and Punta Herrero are
-relatively young reserves, and it is known that community-based marine
-reserves in tropical ecosystems may take up to six years to show a
-spillover effect \citep{dasilva_2015-zX}. Work by \citet{ayer_2018}
-shows these communities support the implementation of marine reserves.
-Nevertheless, fishers may favor implementation of reserves that pose low
-fishing costs either because of their location or size. The economic
-data support our claim that reserves are small, as neither landings nor
-revenues showed the expected short-term costs associated to the first
-years of reserve implementation \citep{ovando_2016-Wg}.
+invertebrates, but not lobsters. Design principles developed by
+\citet{green_2017} for marine reserves in the Caribbean state that
+reserves ``should be more than twice the size of the home range of
+adults and juveniles'', and suggest that reserves seeking to protect
+spiny lobsters should have at least 14 km across. Furthermore, may favor
+implementation of reserves that pose low fishing costs due to their
+small size or location. Our analysis of economic data supports this, as
+neither landings nor revenues showed the expected short-term costs
+associated to the first years of reserve implementation
+\citep{ovando_2016-Wg}. Small reserves can serve as a way to
+self-regulate the spatial distribution of fishing effort
 
-Even if reserves had appropriate sizes, there are other plausible
-explanations for the observed lack of effectiveness. Marine reserves are
-only likely to provide fisheries benefits if initial population sizes
-are low and the fishery is poorly managed \citep{hilborn_2006}. However,
-both lobster fisheries were, at some point, certified by the Marine
-Stewardship Council \citep{prezramrez_2016-J1}. Additionally, lobster
-fisheries are managed via species-specific minimum catch sizes, seasonal
-closures, protection of ``berried'' females, and escapement windows
-where traps are allowed \cite{dof_website_1993}. It is uncertain whether
-that such a well-managed fishery will experience additional benefits
-from marine reserves.
+Even if reserves had appropriate sizes and were placed in optimal
+locations, there are other plausible explanations for the observed
+patterns. For instance, marine reserves are only likely to provide
+fisheries benefits if initial population sizes are low and the fishery
+is poorly managed \citep{hilborn_2006}. Both lobster fisheries were, at
+some point, certified by the Marine Stewardship Council
+\citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed
+via species-specific minimum catch sizes, seasonal closures, protection
+of ``berried'' females, and escapement windows where traps are allowed
+\cite{dof_website_1993}. It is uncertain whether such a well-managed
+fishery will experience additional benefits from marine reserves.
 
@@
 environmental shocks or mistakes in fisheries management
-\citep{hilborn_2004,hilborn_2006,micheli_2012-EU}. Embarking in a marine
-conservation project can bring the community together, which promotes
-social cohesion and builds social capital. Furthermore, showing
-commitment to marine conservation allows fishers to have greater
-bargaining power and leverage over fisheries management.
+\citep{hilborn_2004,hilborn_2006,micheli_2012-EU}. Self-regulation of
+fishing effort (\emph{i.e.} reduction in harvest) can serve as a way to
+compensate for future declines associated to environmental variation
+\citep{finkbeiner_2018}. Embarking in a marine conservation project can
+bring the community together, which promotes social cohesion and builds
+social capital. Furthermore, showing commitment to marine conservation
+allows fishers to have greater bargaining power and leverage over
+fisheries management.
 
@@
 
-JCVD CONACyT (Beca de Posgrados en el extranjero, CVU 669403) + LAFF +
-El proyecto de COBI que haya pagado esto
+JCVD: CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the
+Latin American Fisheries Fellowship.
 
-\textbf{ASC}
-
-\textbf{SF}
-
-\textbf{JT}
-
-\textbf{Que fuentes de financiamiento incluyo par ustedes?}
+AS, SF and JT Walton Family Foundation, Summit Foundation, and Oak
+Foundation.
 

### 2018-06-30 scripts/data_cleaning_logbook.Rmd
Updated data cleaning
Added comments by coauthors

--- 2018-04-10 docs/data_cleaning_logbook.Rmd
@@
   janitor::clean_names() %>% 
+  rename(generoespecie = genero_especie) %>% 
   filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
@@
 ```{r}
-species_bio <- MPAtools::species_bio
+species_bio <- MPAtools::species_bio %>% 
+  mutate(Genus = as.character(str_split(GeneroEspecie, pattern = " ", simplify = T)[,1]),
+         Genus = str_replace(Genus, " ", ""))
 
@@
   rbind(GenusLevelBio) %>% 
-  janitor::clean_names()
+  janitor::clean_names() %>% 
+    rename(generoespecie = genero_especie)
 ```
@@
   select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 
-  arrange()
+  arrange() %>% 
+  drop_na(a, b)
 ```

### 2018-07-06 docs/paper/Villasenor-Derbez.Rmd
Added small changes before sending out to Fio

@@
 title: "Effectiveness of community-based marine reserves in small-scale fisheries"
-rtitle: "Community-based marine reserves"
+rtitle: "Community-based marine reserves: `r Sys.Date()`"
 author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
@@
 email: "jvillasenor@bren.ucsb.edu"
-abstract: "Al finalizar revisiones"
-keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, Mexico, Social-Ecological Systems"
+abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
+Artisanal fisheries face great challenges since they are difficult to monitor, enforce, and manage.
+Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves is believed to improve the performance of small-scale fisheries.
+Mexico has seen a proliferation in their implementation of, with 45 TURF-reserves implemented in the last 4 years after a new regulation was published in 2014.
+However, their effectiveness has not been formally evaluated accounting for the intricate social-ecological dimensions and their governance context.
+This work combines causal inference techniques and the social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico.
+We identify that while reserves don’t achieve their stated goals of increasing lobster densities, they continue to receive significant support from the fishing communities.
+Our triple-bottom evaluation suggests that the lack of effect is a combination of factors.
+First, the lobster fisheries are already well managed, and it is unlikely that reserves might have a detectable effect.
+Second, reserves are not big enough to protect lobster’s home range.
+Third, some of these reserves might be too young for the effects to show.
+However, we argue that these are not failed enterprises, since reserves can still serve as an insurance mechanism against errors and uncertainty in management, as well as environmental variation.
+Furthermore, these reserves have shaped small-scale fisher’s way of thinking about marine reserves, which can provide solid grounds to implement more or larger marine reserves."
+keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, TURF-reserves, Social-Ecological Systems"
 output:
@@
 ```
-
-**Last update**: `r Sys.Date()`
-
-\clearpage
 
@@
 
+\clearpage
+
 # Introduction
@@
 Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintainance of healthy resources \citep{afflerbach_2014-HP,lester_2017}.
+Although in theory these systems are successful \citep{costello_2010-Ix}, little empirical evidence evidence exists of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017}.
 
-Marine reserves allow bounded populations to recover by limiting all extractive activities \citep{halpern_2002}, while the TURF controls the harvest levels outside the MPA.
-Although in theory these systems are successful \citep{costello_2010-Ix}, little empirical evidence evidence exists of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017}.
 The performance of these systems depends on how environmental and social factors work combined.
@@
 
-In Mexico many of these marine reserves are created as community-based marine reserves.
+TURF-reserves can be created as community-based marine reserves.
+This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
 Community-based spatial closures occur in other places, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
-This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
 However, without legal recognition these are difficult to enforce and fishers rely on the exclusive access granted by the TURF.
@@
 First, provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries.
-And second, evaluate the effectiveness of TURF-reserves established as Fishing Refugia in Mexico to identify areas where improvement or adjustment might result in increased effectiveness.
+And second, evaluate the effectiveness of TURF-reserves established as Fishing Refugia in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
 On both cases, we draw from the lessons learned and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries.
@@
 In 2006, the community decided to implement two marine reserves within their fishing grounds to protect commercially important invertebrate species; mainly lobster and abalone.
-The reserves obtained legal recognition in 2018, but have been well enforced since their implementation.
+While these reserves obtained legal recognition until 2018, they have been well enforced since their implementation.
 
@@
 All sites were surveyed annually, and at least once before implementation of the reserves.
-Table \ref{table:com_sum} shows a summary of the reserves included in this study.
+Table \ref{table:com_sum} shows a summary of the TURF-reserves included in this study.
+
+\clearpage
 
@@
                caption = "\\label{table:com_sum} Summary of commuity--based marine reserves by community.",
-               col.names = c("Community", "TURF area ($km^2$)", "Reserve area ($km^2$)", "Percent as reserves","Year of implementation"),
+               col.names = c("Community", "TURF area ($\\mathrm{km}^{-2}$)", "Reserve area ($\\mathrm{km}^{-2}$)", "Percent as reserves","Year of implementation"),
                escape = F)
@@
 Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA).
-Data contain monthly lobster landings (Kg) and revenues (MXP) from 2000 to 2014 for cooperatives with and without marine reserves(**Fig S1**).
-All cooperatives of each region (*i.e.* Pacific and Caribbean) incorporated in this analysis, belong to larger Cooperative Federations, and are exposed to the same markets and institutional frameworks \citep{mccay_2017-1m,ayer_2018}, making them plausible controls.
-Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted by the Consumer Price Index for Mexico \citep{oecd_2017-VV} as:
+Data contain monthly lobster landings (Kg) and revenues (MXP) for cooperatives with and without marine reserves(**Fig S1**).
+Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017-1m,ayer_2018}.
+Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV} as:
 
@@
 We use a difference-in-differences analysis to evaluate these indicators.
-This approach allows us to estimate the effect that the reserve had by comparing trends across time and treatments (*i.e.* reserve / control sites \citet{moland_2013-VP,Villasenor-Derbez_2018}).
+This approach allows us to estimate the effect that the reserve had by comparing trends across time and treatments \citep{moland_2013-VP,Villasenor-Derbez_2018}.
 The analysis of ecological indicators is performed with a multiple linear regression of the form:
@@
 The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site (See Table \ref{table:com_sum}).
-When evaluating biomass and densities of the entire benthic or fish communities, we include $\sigma_j$ to control for species-fixed effects.
+When evaluating biomass and densities of the invertebrate or fish communities, we include $\sigma_j$ to control for species-fixed effects.
 
 Socioeconomic indicators are evaluated with a similar approach.
-Due to data constrains, we only evaluate socioeconomic data for Isla Natividad and Maria Elena.
+Due to data constrains, we only evaluate socioeconomic data for Isla Natividad (2000 - 2014) and Maria Elena (2006 - 2013).
 Neighboring communities are used as counterfacutals that allow us to control for unobserved time-invariants.
@@
 All analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.
+Data and code are available on [github.com](https://github.com/jcvdav/ReserveEffect).
 
@@
 Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}).
-For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, asuggesting that these are plausible controls.
+For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, suggesting that these are plausible controls.
 However, effect sizes do not change after the implementation of the reserve.
-Again, the negative coefficient observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events.
+Interestingly, the negative effect observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events.
 The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$).
@@
 Although we have little information on the social dimension of these fisheries, we can use the social-ecological systems framework (**S1 Table**) to analyze the performance of each governance system (**S6 Table**).
-Our analysis shows that all of the systems analyzed share similarities in their Governance system which is based on cooperatives (GS5.2.3.2), with strong rules in use that include Operational rules (GS6.2), Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even Territorial use communal rights (GS6.1.4.3).
+Our analysis shows that all communities analyzed share similarities in their Governance system which is based on cooperatives (GS5.2.3.2), with strong rules in use that include Operational rules (GS6.2), Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even Territorial use communal rights (GS6.1.4.3).
 However, we identified important differences in terms of the actors, resource systems, and resource units.
@@
 Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperative.
-These conditions have favored the development of an efficient community-based enforcement systems.
+These conditions have favored the development of efficient community-based enforcement systems.
 
@@
 
-```{r, fig.width = 5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles)"}
+```{r, fig.width = 5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles).
+Points indicate the effect size, and errorbars standard errors.
+Years have been centered to year of implementation."}
 
@@
 However, this method assumes control sites are a plausible counterfactual for treated sites.
-This supposed that treated sites would have followed the same trend as control sites, had the reserves not been implemented.
+This implies that treated sites would have followed the same trend as control sites, had the reserves not been implemented.
 Nonetheless, overall trends for each site don't show any significant increases, supporting our findings of lack of change in the indicators used (**S2 Figure**, **S3 Figure**, **S4 Figure**, **S5 Figure**, **S6 Figure**).
 
+A first possible explanation for the lack of effectivenes may be the young age of the reserves.
 Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO}.
@@
 Our analysis of economic data supports this, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
-Small reserves can serve as a way to self-regulate the spatial distribution of fishing effort
 
@@
 Both lobster fisheries were, at some point, certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
-Additionally, lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \cite{dof_website_1993}.
+Additionally, lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \citep{dof_website_1993}.
 It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves.
+Additionally, \citet{gelcich_2008} has shown that TURFs alone can have greater biomass and richness than areas operating under open acces.
+These increased attributes perhaps minimize the difference between TURF and reserve.
+Further research should focus on evaluating sites in the reserve, TURF, and open access areas.
 
@@
 Marine reserves provide protection to a wider range of species and vulnerable habitat, like coral reefs.
-These sites can serve as an insurance against environmental shocks or mistakes in fisheries management \citep{hilborn_2004,hilborn_2006,micheli_2012-EU}.
+These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as environmental shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU}.
 Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
 Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital.
-Furthermore, showing commitment to marine conservation allows fishers to have greater bargaining power and leverage over fisheries management.
+Furthermore, showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
 
+Previous studies have evaluated the potential of implementing marine reserves in Baja California and connect them to the existing network in California \citep{arafehdalmau_2017}.
 Community-based marine reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented.
@@
 For the particular case of the marine reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
-At the broader scale, having full community support surely represents an advantage, but it is important for marine reserves to meet essential design principles such as size and placement.
+At the broader scale, having full community support surely represents an advantage, but it is important that marine reserves meet essential design principles such as size and placement.
 Community-based marine reserves might have more benefits that result from indirect effects of the reserves, which should be taken into account when evaluating the outcomes of similar projects.

### 2018-07-06 docs/paper/Villasenor-Derbez.tex
Added small changes before sending out to Fio

@@
 \def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
-\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
-Aceves-Bueno\(^{1,*}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
-Jorge Torre\(^{2}\)}
+\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira Aceves-Bueno\(^{1}\),
+Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\), Jorge Torre\(^{2}\)}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 
-\title[Community-based marine reserves]{Effectiveness of community-based marine reserves in small-scale
+\title[Community-based marine reserves: 2018-07-05]{Effectiveness of community-based marine reserves in small-scale
 fisheries}
@@
 
-Al finalizar revisiones
+Coastal marine ecosystems provide livelihoods for small-scale fishers
+and coastal communities around the world. Artisanal fisheries face great
+challenges since they are difficult to monitor, enforce, and manage.
+Combining territorial user rights for fisheries (TURF) with no-take
+marine reserves to create TURF-reserves is believed to improve the
+performance of small-scale fisheries. Mexico has seen a proliferation in
+their implementation of, with 45 TURF-reserves implemented in the last 4
+years after a new regulation was published in 2014. However, their
+effectiveness has not been formally evaluated accounting for the
+intricate social-ecological dimensions and their governance context.
+This work combines causal inference techniques and the social-ecological
+systems framework to provide a holistic evaluation of community-based
+marine reserves in three coastal communities in Mexico. We identify that
+while reserves don't achieve their stated goals of increasing lobster
+densities, they continue to receive significant support from the fishing
+communities. Our triple-bottom evaluation suggests that the lack of
+effect is a combination of factors. First, the lobster fisheries are
+already well managed, and it is unlikely that reserves might have a
+detectable effect. Second, reserves are not big enough to protect
+lobster's home range. Third, some of these reserves might be too young
+for the effects to show. However, we argue that these are not failed
+enterprises, since reserves can still serve as an insurance mechanism
+against errors and uncertainty in management, as well as environmental
+variation. Furthermore, these reserves have shaped small-scale fisher's
+way of thinking about marine reserves, which can provide solid grounds
+to implement more or larger marine reserves.
 
@@
  \keyFont{ \section{Keywords:} Marine Protected Areas, Marine Conservation, Small-Scale Fisheries,
-Citizen Science, Mexico, Social-Ecological Systems}
+Citizen Science, TURF-reserves, Social-Ecological Systems}
 
 \end{abstract}
-
-\textbf{Last update}: 2018-06-29
 
@@
 the maintainance of healthy resources
+\citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
+are successful \citep{costello_2010-Ix}, little empirical evidence
+evidence exists of their effectiveness and the drivers of their success
 \citep{afflerbach_2014-HP,lester_2017}.
 
-Marine reserves allow bounded populations to recover by limiting all
-extractive activities \citep{halpern_2002}, while the TURF controls the
-harvest levels outside the MPA. Although in theory these systems are
-successful \citep{costello_2010-Ix}, little empirical evidence evidence
-exists of their effectiveness and the drivers of their success
-\citep{afflerbach_2014-HP,lester_2017}. The performance of these systems
-depends on how environmental and social factors work combined. The
-science of marine reserves has largely focused on understanding the
-ecological effects of these areas, which include increased biomass,
-richness, and densities of organisms within the protected regions,
-climate change mitigation, and protection from environmental variability
+The performance of these systems depends on how environmental and social
+factors work combined. The science of marine reserves has largely
+focused on understanding the ecological effects of these areas, which
+include increased biomass, richness, and densities of organisms within
+the protected regions, climate change mitigation, and protection from
+environmental variability
 \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
@@
 
-In Mexico many of these marine reserves are created as community-based
-marine reserves. Community-based spatial closures occur in other places,
-like the \emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
-\citep{bohnsack_2004,johannes_2002}. This bottom-up approach increases
-compliance and self-enforcement
-\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
-without legal recognition these are difficult to enforce and fishers
-rely on the exclusive access granted by the TURF. In an effort to bridge
-this normative gap, Civil Society Organizations (CSOs) served as the
-link between fishers and government, and set out to create a legal
-framework that solve this governance issue. In 2014, a new norm was
-created, allowing fishers to request the legal recognition of a
-community-based reserve under the name of ``Fish Refuge'' \citep{nom}.
-These can be implemented as temporal or partial reserves, which can
-protect one, some, or all resources within them. Since then, 45 of
-community-based marine reserves along the Pacific, Gulf of California,
-and Mexican Caribbean coastlines have gained legal recognition, but
-their effectiveness has not been reported in the scientific literature.
+TURF-reserves can be created as community-based marine reserves. This
+bottom-up approach increases compliance and self-enforcement
+\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
+Community-based spatial closures occur in other places, like the
+\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
+\citep{bohnsack_2004,johannes_2002}. However, without legal recognition
+these are difficult to enforce and fishers rely on the exclusive access
+granted by the TURF. In an effort to bridge this normative gap, Civil
+Society Organizations (CSOs) served as the link between fishers and
+government, and set out to create a legal framework that solve this
+governance issue. In 2014, a new norm was created, allowing fishers to
+request the legal recognition of a community-based reserve under the
+name of ``Fish Refuge'' \citep{nom}. These can be implemented as
+temporal or partial reserves, which can protect one, some, or all
+resources within them. Since then, 45 of community-based marine reserves
+along the Pacific, Gulf of California, and Mexican Caribbean coastlines
+have gained legal recognition, but their effectiveness has not been
+reported in the scientific literature.
 
@@
 effectiveness of TURF-reserves established as Fishing Refugia in Mexico
-to identify areas where improvement or adjustment might result in
+to identify opportunities where improvement or adjustment might lead to
 increased effectiveness. On both cases, we draw from the lessons learned
@@
 two marine reserves within their fishing grounds to protect commercially
-important invertebrate species; mainly lobster and abalone. The reserves
-obtained legal recognition in 2018, but have been well enforced since
-their implementation.
+important invertebrate species; mainly lobster and abalone. While these
+reserves obtained legal recognition until 2018, they have been well
+enforced since their implementation.
 
@@
 annually, and at least once before implementation of the reserves. Table
-\ref{table:com_sum} shows a summary of the reserves included in this
-study.
+\ref{table:com_sum} shows a summary of the TURF-reserves included in
+this study.
+
+\clearpage
 
@@
 \hline
-Community & TURF area ($km^2$) & Reserve area ($km^2$) & Percent as reserves & Year of implementation\\
+Community & TURF area ($\mathrm{km}^{-2}$) & Reserve area ($\mathrm{km}^{-2}$) & Percent as reserves & Year of implementation\\
 \hline
@@
 Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
-(Kg) and revenues (MXP) from 2000 to 2014 for cooperatives with and
-without marine reserves(\textbf{Fig S1}). All cooperatives of each
-region (\emph{i.e.} Pacific and Caribbean) incorporated in this
-analysis, belong to larger Cooperative Federations, and are exposed to
-the same markets and institutional frameworks
-\citep{mccay_2017-1m,ayer_2018}, making them plausible controls.
-Landings and revenues were aggregated at the cooperative-year level, and
-revenues were adjusted by the Consumer Price Index for Mexico
+(Kg) and revenues (MXP) for cooperatives with and without marine
+reserves(\textbf{Fig S1}). Cooperatives incorporated in this analysis
+belong to larger regional-level Cooperative Federations, and are exposed
+to the same markets and institutional frameworks, making them plausible
+controls \citep{mccay_2017-1m,ayer_2018}. Landings and revenues were
+aggregated at the cooperative-year level, and revenues were adjusted to
+represent 2014 values by the Consumer Price Index for Mexico
 \citep{oecd_2017-VV} as:
@@
 indicators. This approach allows us to estimate the effect that the
-reserve had by comparing trends across time and treatments (\emph{i.e.}
-reserve / control sites \citet{moland_2013-VP,Villasenor-Derbez_2018}).
-The analysis of ecological indicators is performed with a multiple
-linear regression of the form:
+reserve had by comparing trends across time and treatments
+\citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis of
+ecological indicators is performed with a multiple linear regression of
+the form:
 
@@
 of implementation in the control site (See Table \ref{table:com_sum}).
-When evaluating biomass and densities of the entire benthic or fish
+When evaluating biomass and densities of the invertebrate or fish
 communities, we include \(\sigma_j\) to control for species-fixed
@@
 data constrains, we only evaluate socioeconomic data for Isla Natividad
-and Maria Elena. Neighboring communities are used as counterfacutals
-that allow us to control for unobserved time-invariants. Each
-``treated'' community (Isla Natividad and Maria Elena) has three
-counterfactual communities.
+(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
+used as counterfacutals that allow us to control for unobserved
+time-invariants. Each ``treated'' community (Isla Natividad and Maria
+Elena) has three counterfactual communities.
 
@@
 analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.
+Data and code are available on
+\href{https://github.com/jcvdav/ReserveEffect}{github.com}.
 
@@
 control and treatment sites have similar pre-treatment trends,
-asuggesting that these are plausible controls. However, effect sizes do
-not change after the implementation of the reserve. Again, the negative
-coefficient observed for Isla Natividad on year 5 correspond to the 2011
-hypoxia events. The only positive change observed in lobster landings is
-for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
+suggesting that these are plausible controls. However, effect sizes do
+not change after the implementation of the reserve. Interestingly, the
+negative effect observed for Isla Natividad on year 5 correspond to the
+2011 hypoxia events. The only positive change observed in lobster
+landings is for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
 post-implementation data for Maria Elena do not show a significant
@@
 (\textbf{S1 Table}) to analyze the performance of each governance system
-(\textbf{S6 Table}). Our analysis shows that all of the systems analyzed
+(\textbf{S6 Table}). Our analysis shows that all communities analyzed
 share similarities in their Governance system which is based on
@@
 members of the fishing cooperative. These conditions have favored the
-development of an efficient community-based enforcement systems.
+development of efficient community-based enforcement systems.
 
@@
 lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
-circles) and Maria Elena (ME; blue triangles)}
+circles) and Maria Elena (ME; blue triangles). Points indicate the
+effect size, and errorbars standard errors. Years have been centered to
+year of implementation.}
 \end{figure}
@@
 However, this method assumes control sites are a plausible
-counterfactual for treated sites. This supposed that treated sites would
+counterfactual for treated sites. This implies that treated sites would
 have followed the same trend as control sites, had the reserves not been
@@
 
-Literature shows that age and enforcement are important factors that
-influence reserve effectiveness \citep{edgar_2014-UO}. Isla Natividad
-has the oldest reserve, and our SES analysis suggests that all
-communities have a well-established community-based enforcement system.
-With these characteristics, one would expect the reserves to be
-effective. Maria Elena and Punta Herrero are relatively young reserves
-(\emph{i.e.} \textless{} 5 years old); other community-based marine
-reserves in tropical ecosystems may take up to six years to show a
-spillover effect \citep{dasilva_2015-zX}.
+A first possible explanation for the lack of effectivenes may be the
+young age of the reserves. Literature shows that age and enforcement are
+important factors that influence reserve effectiveness
+\citep{edgar_2014-UO}. Isla Natividad has the oldest reserve, and our
+SES analysis suggests that all communities have a well-established
+community-based enforcement system. With these characteristics, one
+would expect the reserves to be effective. Maria Elena and Punta Herrero
+are relatively young reserves (\emph{i.e.} \textless{} 5 years old);
+other community-based marine reserves in tropical ecosystems may take up
+to six years to show a spillover effect \citep{dasilva_2015-zX}.
 
@@
 associated to the first years of reserve implementation
-\citep{ovando_2016-Wg}. Small reserves can serve as a way to
-self-regulate the spatial distribution of fishing effort
+\citep{ovando_2016-Wg}.
 
@@
 of ``berried'' females, and escapement windows where traps are allowed
-\cite{dof_website_1993}. It is uncertain whether such a well-managed
+\citep{dof_website_1993}. It is uncertain whether such a well-managed
 fishery will experience additional benefits from marine reserves.
+Additionally, \citet{gelcich_2008} has shown that TURFs alone can have
+greater biomass and richness than areas operating under open acces.
+These increased attributes perhaps minimize the difference between TURF
+and reserve. Further research should focus on evaluating sites in the
+reserve, TURF, and open access areas.
 
@@
 like coral reefs. These sites can serve as an insurance against
-environmental shocks or mistakes in fisheries management
-\citep{hilborn_2004,hilborn_2006,micheli_2012-EU}. Self-regulation of
-fishing effort (\emph{i.e.} reduction in harvest) can serve as a way to
-compensate for future declines associated to environmental variation
-\citep{finkbeiner_2018}. Embarking in a marine conservation project can
-bring the community together, which promotes social cohesion and builds
-social capital. Furthermore, showing commitment to marine conservation
+uncertainty and errors in fisheries management, as well as environmental
+shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU}.
+Self-regulation of fishing effort (\emph{i.e.} reduction in harvest) can
+serve as a way to compensate for future declines associated to
+environmental variation \citep{finkbeiner_2018}. Embarking in a marine
+conservation project can bring the community together, which promotes
+social cohesion and builds social capital. Furthermore, showing
+commitment to marine conservation and sustainable fishing practices
 allows fishers to have greater bargaining power and leverage over
-fisheries management.
+fisheries management \citep{prezramrez_2012}.
 
-Community-based marine reserves in small-scale fisheries can be helpful
-conservation and fishery management tools when appropriately
-implemented. Lessons learned from these cases can guide implementation
-of community-based marine reserves elsewhere. For the particular case of
-the marine reserves that we evaluate, the possibility of expanding
-reserves or merging existing polygons into larger areas should be
-evaluated and proposed to the communities. At the broader scale, having
-full community support surely represents an advantage, but it is
-important for marine reserves to meet essential design principles such
-as size and placement. Community-based marine reserves might have more
-benefits that result from indirect effects of the reserves, which should
-be taken into account when evaluating the outcomes of similar projects.
+Previous studies have evaluated the potential of implementing marine
+reserves in Baja California and connect them to the existing network in
+California \citep{arafehdalmau_2017}. Community-based marine reserves in
+small-scale fisheries can be helpful conservation and fishery management
+tools when appropriately implemented. Lessons learned from these cases
+can guide implementation of community-based marine reserves elsewhere.
+For the particular case of the marine reserves that we evaluate, the
+possibility of expanding reserves or merging existing polygons into
+larger areas should be evaluated and proposed to the communities. At the
+broader scale, having full community support surely represents an
+advantage, but it is important that marine reserves meet essential
+design principles such as size and placement. Community-based marine
+reserves might have more benefits that result from indirect effects of
+the reserves, which should be taken into account when evaluating the
+outcomes of similar projects.
 

### 2018-07-06 scripts/data_cleaning_logbook.Rmd
Added small changes before sending out to Fio

@@
 conapesca <- readRDS(here::here("raw_data","conapesca.rds")) %>% 
-  mutate(UnidadEconomica = as.character(UnidadEconomica)) %>% 
+  mutate(UnidadEconomica = as.character(UnidadEconomica),
+         LugarDeCaptura = as.character(LugarDeCaptura)) %>% 
   mutate(UnidadEconomica = ifelse(UnidadEconomica == "Scpp bahia tortugas s.c de r.l", 
@@
   filter(NombreComun %in% c("Langosta roja ent. fca.", "Langosta ent. fca.")) %>%
+  filter(!(UnidadEconomica == "Scpp cozumel scl" & !LugarDeCaptura %in% c("Bahia espiritu santo",
+                                                                          "Bahia del del espiritu santo",
+                                                                          "Bahia espiritu santo"))) %>% 
   group_by(UnidadEconomica, Ano, NombreComun) %>%

### 2018-08-16 docs/paper/Villasenor-Derbez.Rmd
Incorporated Fio's comments

@@
 Artisanal fisheries face great challenges since they are difficult to monitor, enforce, and manage.
-Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves is believed to improve the performance of small-scale fisheries.
-Mexico has seen a proliferation in their implementation of, with 45 TURF-reserves implemented in the last 4 years after a new regulation was published in 2014.
-However, their effectiveness has not been formally evaluated accounting for the intricate social-ecological dimensions and their governance context.
-This work combines causal inference techniques and the social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico.
-We identify that while reserves don’t achieve their stated goals of increasing lobster densities, they continue to receive significant support from the fishing communities.
-Our triple-bottom evaluation suggests that the lack of effect is a combination of factors.
+Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves is believed to improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
+In the last 4 years, 45 TURF reserves have been implemented in Mexico, but their effectiveness has not been formally evaluated.
+We combine causal inference techniques and a social-ecological systems framework to provide a holistic evaluation of community-based TURF reserves in three coastal communities in Mexico.
+We find that while reserves have not yet achieved their stated goal of increasing lobster densities, they continue to receive significant support from the fishing communities.
+A lack of ecological and socioeconomic effects or reserves likely results from a combination of factors.
 First, the lobster fisheries are already well managed, and it is unlikely that reserves might have a detectable effect.
-Second, reserves are not big enough to protect lobster’s home range.
+Second, reserves are not large enough to protect lobster’s home ranges.
 Third, some of these reserves might be too young for the effects to show.
-However, we argue that these are not failed enterprises, since reserves can still serve as an insurance mechanism against errors and uncertainty in management, as well as environmental variation.
-Furthermore, these reserves have shaped small-scale fisher’s way of thinking about marine reserves, which can provide solid grounds to implement more or larger marine reserves."
+However, these reserves have shaped small-scale fishers’ way of thinking about marine management, which can provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
 keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, TURF-reserves, Social-Ecological Systems"
@@
     template: template.tex
+header-includes:
+  - \usepackage{float}
+  - \floatplacement{figure}{H}
 ---
@@
 Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintainance of healthy resources \citep{afflerbach_2014-HP,lester_2017}.
-Although in theory these systems are successful \citep{costello_2010-Ix}, little empirical evidence evidence exists of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017}.
+Although in theory these systems are successful \citep{costello_2010-Ix}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017}.
 
-The performance of these systems depends on how environmental and social factors work combined.
-The science of marine reserves has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
-Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1}.
-Other research has focused on the relationship between socioeconomic and governance structures and their relationship to reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
+The performance of these systems depends on how environmental and social factors combine and interact.
+The science of marine reserves has largely focused on understanding the ecological effects of these areas, which include increased biomass, species richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
+Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1} [**add De Leo and Micheli 2015**](http://dx.doi.org/10.1098/rstb.2014.0276).
+Other research has focused on the relationship between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
 However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.
+This is especially important in social-ecological coastal systems dominated by close interaction and feedbacks between people and natural resources \citep{ostrom_2009-hg}.
 
-TURF-reserves can be created as community-based marine reserves.
+TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities.
 This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
-Community-based spatial closures occur in other places, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
-However, without legal recognition these are difficult to enforce and fishers rely on the exclusive access granted by the TURF.
-In an effort to bridge this normative gap, Civil Society Organizations (CSOs) served as the link between fishers and government, and set out to create a legal framework that solve this governance issue.
-In 2014, a new norm was created, allowing fishers to request the legal recognition of a community-based reserve under the name of "Fish Refuge" \citep{nom}.
-These can be implemented as temporal or partial reserves, which can protect one, some, or all resources within them.
-Since then, 45 of community-based marine reserves along the Pacific, Gulf of California, and Mexican Caribbean coastlines have gained legal recognition, but their effectiveness has not been reported in the scientific literature.
+Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
+However, without legal recognition no-take regulations are difficult to enforce and fishers rely on the exclusive access granted by the TURF.
+In an effort to bridge this normative gap, Civil Society Organizations (CSOs) have served as a link between fishers and government, and set out to create a legal framework that solve this governance issue.
+In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuge” (*Zona de Refugio Pesquero* ;\citet{nom}).
+Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
+Since 2014, 45 of community-based marine reserves along the Pacific, Gulf of California, and Mexican Caribbean coastlines have gained legal recognition.
+However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
 
-This work combines causal inference techniques and the social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico.
+Here, we combine causal inference techniques and a social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico.
+These three case studies span a range of ecological and social conditions representative of different regions of Mexico.
 The objective of this work is twofold.
-First, provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries.
-And second, evaluate the effectiveness of TURF-reserves established as Fishing Refugia in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
-On both cases, we draw from the lessons learned and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries.
-
-<!-- Implementation of marine reserves can be done by following one of three main approaches in terms of implementation.
-We describe these in the context of Mexican MRs, but argue that at least the first two apply elsewhere.
-Marine reserves are usually implemented and managed by a government agency, in this case the National Commission of Natural Protected Areas (*Comisión Nacional de Áreas Naturales Protegidas*, CONANP).
-While CONANP has made efforts to incorporate stakeholders in the design and implementation phases, these are still characterized by a heavy top-down process.
-A second approach is the implementation of community-based marine reserves, usually placed within areas of exclusive access (*i.e.* TURFs).
--->
+First, to provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries.
+Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
+We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world that are using this tool to manage and rebuild their coastal fisheries.
 
@@
 In 2006, the community decided to implement two marine reserves within their fishing grounds to protect commercially important invertebrate species; mainly lobster and abalone.
-While these reserves obtained legal recognition until 2018, they have been well enforced since their implementation.
+While these reserves obtained legal recognition only in 2018, they have been well enforced since their implementation.
 
 The other two TURF-reserves are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C).
-Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); Punta Herrero is home to the *SCPP José María Azcorra* cooperative.
+In contrast with Isla Nativdad, which hosts a well established fishing community, Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); Punta Herrero is home to the *SCPP José María Azcorra* cooperative, and similar to Isla Natividad hosts a local community.
 Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season.
@@
 
-## Data collection
+<!-- Revise the paragraph bellow. Add citations for CC system (Ramires-Valdez?), QRoo (Fulton, 2018, Caamal, half-way?). Cite noms as "opperate under similar fishing regulations". -->
 
-```{r load_data}
-invert <- read.csv(file = here::here("data", "invertebrates.csv"),
-                stringsAsFactors = F)
-
-fish <- read.csv(file = here::here("data", "fish.csv"),
-                stringsAsFactors = F)
-
-conapesca <- read.csv(file = here::here("data", "conapesca.csv"),
-                stringsAsFactors = F)
-
-gob <- read.csv(here::here("data" ,"governance.csv"),
-                strip.white = T,
-                stringsAsFactors = F)
-```
-
-We use three main sources of information to evaluate these reserves across the ecological, socioeconomic, and governance dimensions.
-Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
-Trained divers record richness and abundances of fish and invertebrate species in the reserves and control sites \citep{fulton_2018}.
-Size structures are also collected during fish surveys.
-We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had a similar probability of being selected as reserves during the design phase.
-We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
-This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
-BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
-All sites were surveyed annually, and at least once before implementation of the reserves.
-Table \ref{table:com_sum} shows a summary of the TURF-reserves included in this study.
+These communities are representative of their region in terms of ecology, socioeconomic, and governance aspects.
+Isla Natividad, for example, is part of a greater group of fishing cooperatives that assamble into a Federation of Cooperatives (**FEDECOOP**).
+This group of X communities has been identified as a cohesive group that often cooperate to better manage their resources \citep{mccay_2017-1m,mccay_2014-CN} (**Aceves, 2018?**).
+All these communities occur in the sourthenmost raange distribution of Giant Kelp, and are still under the influence of the California Current System.
+Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean.
+Together, these three communities provide an accurate representation of all other fishing communities in Mexico.
+While each region has more communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth analysis that we undertake.
+Yet, given the similarities among communities and the socioeconomic and governance setting under which they opperate, it is sage to generalize our results to other communities in Mexico and other regions around the world.
 
@@
 
+## Data collection
+
+```{r load_data}
+invert <- read.csv(file = here::here("data", "invertebrates.csv"),
+                stringsAsFactors = F)
+
+fish <- read.csv(file = here::here("data", "fish.csv"),
+                stringsAsFactors = F)
+
+conapesca <- read.csv(file = here::here("data", "conapesca.csv"),
+                stringsAsFactors = F)
+
+gob <- read.csv(here::here("data" ,"governance.csv"),
+                strip.white = T,
+                stringsAsFactors = F)
+```
+
+We use three main sources of information to evaluate these reserves across the ecological, socioeconomic, and governance dimensions.
+Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
+Trained divers record richness and abundances of fish and invertebrate species along replicate transects (30x 2 m each) at depths 5-20 m in the reserves and control sites \citep{fulton_2018}.
+Size structures are also collected during fish surveys.
+We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had a similar probability of being selected as reserves during the design phase.
+We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
+This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
+BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
+All sites were surveyed annually, and at least once before implementation of the reserves.
+Table \ref{table:com_sum} shows a summary of the TURF-reserves included in this study.
+
+\clearpage
+
 ```{r}
@@
                col.names = c("Community", "TURF area ($\\mathrm{km}^{-2}$)", "Reserve area ($\\mathrm{km}^{-2}$)", "Percent as reserves","Year of implementation"),
-               escape = F)
+               escape = F) %>% 
+  kable_styling(latex_options = "HOLD_position")
 ```
@@
 tibble(Category = c("Biological", "Biological","Biological","Biological", "Socioeconomic", "Socioeconomic"),
-           Indicador = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
+           Indicator = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
        Units = c("org $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "Kg $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "M MXP", "Metric Tonnes")) %>% 
@@
 Due to data constrains, we only evaluate socioeconomic data for Isla Natividad (2000 - 2014) and Maria Elena (2006 - 2013).
-Neighboring communities are used as counterfacutals that allow us to control for unobserved time-invariants.
-Each "treated" community (Isla Natividad and Maria Elena) has three counterfactual communities.
+Neighboring communities are used as counterfactuals that allow us to control for unobserved time-invariants.
+Each focal community (Isla Natividad and Maria Elena) has three counterfactual communities.
 
@@
 The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$) and $\sigma_jCom$ captures community-level fixed-effects.
-These regressions allows us to make a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}.
+These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}.
 The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
@@
 
-## Biological
+## Biological effects
 
@@
 These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ for Isla Natividad and Punta Herrero, and 0.01 organisms $\mathrm{m}^{-2}$ for Maria Elena, but are not significantly different from zero ($p > 0.05$).
-The rapid increase observed for changes in lobster densities for Isla Natividad on the sixth year (*i.e.* 2012) occur a year after the hypoxia events described by \citet{micheli_2012-EU} caused mass mortality of organisms.
 Likewise, no changes were detected in fish biomass or invertebrate and fish densities (\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
@@
 
-## Socioeconomic
+## Socioeconomic effects
 
@@
 All regression coefficients for each community and indicator are presented in **S5 Table**.
+
+<!-- Revenues show a positive trend post implementation for I Natividad. Because landings have not changed, this might be due to an increase in price? -->
 
@@
 Points are jittered hotizontally to avoid overplotting.
-Points indicate the effect size, and errorbars standard errors.
+Points indicate the effect size and standard errors.
 Years have been centered to year of implementation."}
@@
 
-cowplot::plot_grid(plotlist = list(L_plot, F_B, I_N, F_N),
+plot2 <- cowplot::plot_grid(plotlist = list(L_plot, F_B, I_N, F_N),
                             labels = "AUTO",
                             ncol = 2)
+
+plot2
 ```
@@
 ```{r, fig.width = 5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles).
-Points indicate the effect size, and errorbars standard errors.
+Points indicate the effect size and standard errors.
 Years have been centered to year of implementation."}
@@
 
+plot3 <- plot_grid(c_plot + theme(legend.position = "none"), v_plot, ncol = 2, labels = "AUTO")
+
 ```
 
-\clearpage
+# Discussion
 
-# Discussion
+<!-- From Fio: One aspect that is important to discuss is the concurrent effect of environmental variability.
+In isla Natividad, and the whole Pacifico Norte region, the decade following the establishment of the reserves has seen extreme ENSO events and hypoxia, which have impacted populations above and beyond the effects of fishing and conservation.
+-->
 
@@
 The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
-The lack of evidence of the effectiveness of these reserves is surprising since most of the communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}.
-Analyzing the shortcomings of our study and understanding the social-ecological context in which these communities and their reserves operate might provide insights to this question.
+The coastal ecosystems where these reserves are located have been profoundly affected by climatic and oceanographic extremes, including warming events, extreme storms and prolonged hypoxia \citep{micheli_2012-EU}, (**Beas et al.
+in prep., Woodson et al in review**).
+Despite the lack of evidence of the effectiveness of these reserves, most of the communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}.
+Understanding the social-ecological context in which these communities and their reserves operate might provide insights to this.
 
@@
 Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness.
-However, this method assumes control sites are a plausible counterfactual for treated sites.
+For example, we capture previously described patterns like the rapid increase observed for changes in lobster densities for Isla Natividad on the sixth year (*i.e.* 2012; Fig.
+\ref{fig:indicators}A) occur a year after the hypoxia events described by \citet{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
+Yet, our empirical approach assumes control sites are a plausible counterfactual for treated sites.
 This implies that treated sites would have followed the same trend as control sites, had the reserves not been implemented.
@@
 A first possible explanation for the lack of effectivenes may be the young age of the reserves.
-Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO}.
-Isla Natividad has the oldest reserve, and our SES analysis suggests that all communities have a well-established community-based enforcement system.
+Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO} **Babcock PNAS 2010**.
+Isla Natividad has the oldest reserves, and our SES analysis suggests that all communities have a well-established community-based enforcement system.
 With these characteristics, one would expect the reserves to be effective.
-Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 5 years old); other community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}.
+Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 5 years old) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; other community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}.
 
-Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to reserves being too small.
+Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve dsign (*sensu* \citet{rees_2018}.
 Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
-Abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sesile invertebrates, but not lobsters.
+Abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary invertebrates, but not lobsters.
 Design principles developed by \citet{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
@@
 For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2006}.
-Both lobster fisheries were, at some point, certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
+Both lobster fisheries were certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
 Additionally, lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \citep{dof_website_1993}.
@@
 
-While reserves fail to provide fishery benefits, there are a number of additional ecological, fisheries, and social benefits.
+While reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits.
 Marine reserves provide protection to a wider range of species and vulnerable habitat, like coral reefs.
-These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as environmental shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU}.
+These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as environmental shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (**, Aalto et al, in press Am Nat**).
 Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.

### 2018-08-16 docs/paper/Villasenor-Derbez.tex
Incorporated Fio's comments

@@
 
-\title[Community-based marine reserves: 2018-07-05]{Effectiveness of community-based marine reserves in small-scale
+\title[Community-based marine reserves: 2018-08-16]{Effectiveness of community-based marine reserves in small-scale
 fisheries}
@@
 marine reserves to create TURF-reserves is believed to improve the
-performance of small-scale fisheries. Mexico has seen a proliferation in
-their implementation of, with 45 TURF-reserves implemented in the last 4
-years after a new regulation was published in 2014. However, their
-effectiveness has not been formally evaluated accounting for the
-intricate social-ecological dimensions and their governance context.
-This work combines causal inference techniques and the social-ecological
-systems framework to provide a holistic evaluation of community-based
-marine reserves in three coastal communities in Mexico. We identify that
-while reserves don't achieve their stated goals of increasing lobster
-densities, they continue to receive significant support from the fishing
-communities. Our triple-bottom evaluation suggests that the lack of
-effect is a combination of factors. First, the lobster fisheries are
-already well managed, and it is unlikely that reserves might have a
-detectable effect. Second, reserves are not big enough to protect
-lobster's home range. Third, some of these reserves might be too young
-for the effects to show. However, we argue that these are not failed
-enterprises, since reserves can still serve as an insurance mechanism
-against errors and uncertainty in management, as well as environmental
-variation. Furthermore, these reserves have shaped small-scale fisher's
-way of thinking about marine reserves, which can provide solid grounds
-to implement more or larger marine reserves.
+performance of small-scale fisheries by buffering fisheries from
+environmental variability and management errors, while ensuring that
+fishers reap the benefits of conservation investments. In the last 4
+years, 45 TURF reserves have been implemented in Mexico, but their
+effectiveness has not been formally evaluated. We combine causal
+inference techniques and a social-ecological systems framework to
+provide a holistic evaluation of community-based TURF reserves in three
+coastal communities in Mexico. We find that while reserves have not yet
+achieved their stated goal of increasing lobster densities, they
+continue to receive significant support from the fishing communities. A
+lack of ecological and socioeconomic effects or reserves likely results
+from a combination of factors. First, the lobster fisheries are already
+well managed, and it is unlikely that reserves might have a detectable
+effect. Second, reserves are not large enough to protect lobster's home
+ranges. Third, some of these reserves might be too young for the effects
+to show. However, these reserves have shaped small-scale fishers' way of
+thinking about marine management, which can provide a foundation for
+establishing additional, larger marine reserves needed to effectively
+conserve mobile species.
 
@@
 \citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
-are successful \citep{costello_2010-Ix}, little empirical evidence
-evidence exists of their effectiveness and the drivers of their success
+are successful \citep{costello_2010-Ix}, there is little empirical
+evidence of their effectiveness and the drivers of their success
 \citep{afflerbach_2014-HP,lester_2017}.
@@
 The performance of these systems depends on how environmental and social
-factors work combined. The science of marine reserves has largely
+factors combine and interact. The science of marine reserves has largely
 focused on understanding the ecological effects of these areas, which
-include increased biomass, richness, and densities of organisms within
-the protected regions, climate change mitigation, and protection from
-environmental variability
+include increased biomass, species richness, and densities of organisms
+within the protected regions, climate change mitigation, and protection
+from environmental variability
 \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
@@
 operates, as well as reserve size and the amount of larvae exported from
-these \citep{hilborn_2006,krueck_2017-J1}. Other research has focused on
-the relationship between socioeconomic and governance structures and
-their relationship to reserve effectiveness
+these \citep{hilborn_2006,krueck_2017-J1}
+\href{http://dx.doi.org/10.1098/rstb.2014.0276}{\textbf{add De Leo and
+Micheli 2015}}. Other research has focused on the relationship between
+socioeconomic and governance structures and reserve effectiveness
 \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
 knowledge, no studies exist that evaluate TURF-reserves from both a
-social and ecological perspective.
+social and ecological perspective. This is especially important in
+social-ecological coastal systems dominated by close interaction and
+feedbacks between people and natural resources \citep{ostrom_2009-hg}.
 
-TURF-reserves can be created as community-based marine reserves. This
+TURF-reserves can be created as community-based marine reserves,
+voluntarily established and enforced by local communities. This
 bottom-up approach increases compliance and self-enforcement
 \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
-Community-based spatial closures occur in other places, like the
+Community-based spatial closures occur in different contexts, like the
 \emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
 \citep{bohnsack_2004,johannes_2002}. However, without legal recognition
-these are difficult to enforce and fishers rely on the exclusive access
-granted by the TURF. In an effort to bridge this normative gap, Civil
-Society Organizations (CSOs) served as the link between fishers and
-government, and set out to create a legal framework that solve this
-governance issue. In 2014, a new norm was created, allowing fishers to
-request the legal recognition of a community-based reserve under the
-name of ``Fish Refuge'' \citep{nom}. These can be implemented as
-temporal or partial reserves, which can protect one, some, or all
-resources within them. Since then, 45 of community-based marine reserves
+no-take regulations are difficult to enforce and fishers rely on the
+exclusive access granted by the TURF. In an effort to bridge this
+normative gap, Civil Society Organizations (CSOs) have served as a link
+between fishers and government, and set out to create a legal framework
+that solve this governance issue. In Mexico, a new norm was created in
+2014 allowing fishers to request the legal recognition of
+community-based reserves as ``Fish Refuge'' (\emph{Zona de Refugio
+Pesquero} ;\citet{nom}). Fish refuges can be implemented as temporal or
+partial reserves, which can protect one, some, or all resources within
+their boundaries. Since 2014, 45 of community-based marine reserves
 along the Pacific, Gulf of California, and Mexican Caribbean coastlines
-have gained legal recognition, but their effectiveness has not been
-reported in the scientific literature.
+have gained legal recognition. However, their effectiveness has not yet
+been formally evaluated and reported in the scientific literature.
 
-This work combines causal inference techniques and the social-ecological
+Here, we combine causal inference techniques and a social-ecological
 systems framework to provide a holistic evaluation of community-based
-marine reserves in three coastal communities in Mexico. The objective of
-this work is twofold. First, provide a triple bottom line evaluation of
-the effectiveness of community-based marine reserves that can inform
-similar processes in other countries. And second, evaluate the
-effectiveness of TURF-reserves established as Fishing Refugia in Mexico
-to identify opportunities where improvement or adjustment might lead to
-increased effectiveness. On both cases, we draw from the lessons learned
-and provide management recommendations to maximize the effectiveness of
-community-based marine reserves in small-scale fisheries.
+marine reserves in three coastal communities in Mexico. These three case
+studies span a range of ecological and social conditions representative
+of different regions of Mexico. The objective of this work is twofold.
+First, to provide a triple bottom line evaluation of the effectiveness
+of community-based marine reserves that can inform similar processes in
+other countries. Second, to evaluate the effectiveness of TURF-reserves
+established as Fish Refuges in Mexico to identify opportunities where
+improvement or adjustment might lead to increased effectiveness. We draw
+from lessons learned in these three case studies and provide management
+recommendations to maximize the effectiveness of community-based marine
+reserves in small-scale fisheries in Mexico and in other regions around
+the world that are using this tool to manage and rebuild their coastal
+fisheries.
 
@@
 important invertebrate species; mainly lobster and abalone. While these
-reserves obtained legal recognition until 2018, they have been well
+reserves obtained legal recognition only in 2018, they have been well
 enforced since their implementation.
@@
 The other two TURF-reserves are located in Maria Elena and Punta
-Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). Maria Elena is a
-fishing camp --visited intermittently during the fishing season--
-belonging to the Cozumel fishing cooperative (\emph{SCPP Cozumel});
-Punta Herrero is home to the \emph{SCPP José María Azcorra} cooperative.
+Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
+Isla Nativdad, which hosts a well established fishing community, Maria
+Elena is a fishing camp --visited intermittently during the fishing
+season-- belonging to the Cozumel fishing cooperative (\emph{SCPP
+Cozumel}); Punta Herrero is home to the \emph{SCPP José María Azcorra}
+cooperative, and similar to Isla Natividad hosts a local community.
 Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
@@
 
-\subsection{Data collection}\label{data-collection}
-
-We use three main sources of information to evaluate these reserves
-across the ecological, socioeconomic, and governance dimensions.
-Ecological data come from the annual ecological monitoring of reserve
-and control areas, carried out by members from each community and
-personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
-(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
-abundances of fish and invertebrate species in the reserves and control
-sites \citep{fulton_2018}. Size structures are also collected during
-fish surveys. We define control sites as regions with habitat
-characteristics similar to the corresponding reserves, and that
-presumably had a similar probability of being selected as reserves
-during the design phase. We focus our evaluation on sites where data are
-available for reserve and control sites, before and after the
-implementation of the reserve. This provides us with a
-Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
-allows us to capture and control for temporal and spatial dynamics
-\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
-techniques have proven effective to evaluate marine reserves, as they
-allow us to causally attribute observed changes to the intervention
-\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
-annually, and at least once before implementation of the reserves. Table
-\ref{table:com_sum} shows a summary of the TURF-reserves included in
-this study.
+These communities are representative of their region in terms of
+ecology, socioeconomic, and governance aspects. Isla Natividad, for
+example, is part of a greater group of fishing cooperatives that
+assamble into a Federation of Cooperatives (\textbf{FEDECOOP}). This
+group of X communities has been identified as a cohesive group that
+often cooperate to better manage their resources
+\citep{mccay_2017-1m,mccay_2014-CN} (\textbf{Aceves, 2018?}). All these
+communities occur in the sourthenmost raange distribution of Giant Kelp,
+and are still under the influence of the California Current System.
+Likewise, Maria Elena and Punta Herrero are representative of fishing
+cooperatives in the Mexican Caribbean. Together, these three communities
+provide an accurate representation of all other fishing communities in
+Mexico. While each region has more communities that have established
+community-based TURF-reserves, available data would not allow us to
+perform the in-depth analysis that we undertake. Yet, given the
+similarities among communities and the socioeconomic and governance
+setting under which they opperate, it is sage to generalize our results
+to other communities in Mexico and other regions around the world.
 
@@
 
-\begin{table}
+\subsection{Data collection}\label{data-collection}
+
+We use three main sources of information to evaluate these reserves
+across the ecological, socioeconomic, and governance dimensions.
+Ecological data come from the annual ecological monitoring of reserve
+and control areas, carried out by members from each community and
+personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
+(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
+abundances of fish and invertebrate species along replicate transects
+(30x 2 m each) at depths 5-20 m in the reserves and control sites
+\citep{fulton_2018}. Size structures are also collected during fish
+surveys. We define control sites as regions with habitat characteristics
+similar to the corresponding reserves, and that presumably had a similar
+probability of being selected as reserves during the design phase. We
+focus our evaluation on sites where data are available for reserve and
+control sites, before and after the implementation of the reserve. This
+provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
+sampling design that allows us to capture and control for temporal and
+spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and
+causal inference techniques have proven effective to evaluate marine
+reserves, as they allow us to causally attribute observed changes to the
+intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites
+were surveyed annually, and at least once before implementation of the
+reserves. Table \ref{table:com_sum} shows a summary of the TURF-reserves
+included in this study.
+
+\clearpage
+
+\begin{table}[H]
 
@@
 \hline
-Category & Indicador & Units\\
+Category & Indicator & Units\\
 \hline
@@
 \end{table}
+
+\clearpage
 
@@
 (2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
-used as counterfacutals that allow us to control for unobserved
-time-invariants. Each ``treated'' community (Isla Natividad and Maria
-Elena) has three counterfactual communities.
+used as counterfactuals that allow us to control for unobserved
+time-invariants. Each focal community (Isla Natividad and Maria Elena)
+has three counterfactual communities.
 
@@
 a reserve (\(Treated = 1\)) or not (\(Treated = 0\)) and \(\sigma_jCom\)
-captures community-level fixed-effects. These regressions allows us to
-make a causal link between the implementation of marine reserves and the
-observed trends by accounting for temporal and spatial dynamics
+captures community-level fixed-effects. These regression models allow us
+to establish a causal link between the implementation of marine reserves
+and the observed trends by accounting for temporal and spatial dynamics
 \citep{depalma_2018}. The effect of the reserve is captured by the
@@
 
-\subsection{Biological}\label{biological}
+\subsection{Biological effects}\label{biological-effects}
 
@@
 for Maria Elena, but are not significantly different from zero
-(\(p > 0.05\)). The rapid increase observed for changes in lobster
-densities for Isla Natividad on the sixth year (\emph{i.e.} 2012) occur
-a year after the hypoxia events described by \citet{micheli_2012-EU}
-caused mass mortality of organisms. Likewise, no changes were detected
-in fish biomass or invertebrate and fish densities
-(\ref{fig:indicators}B-D), where effect sizes oscillated around zero
-without clear trends. Full tables with model coefficients are presented
-in the supplementary materials (\textbf{S2 Table}, \textbf{S3 Table},
-\textbf{S4 Table}).
+(\(p > 0.05\)). Likewise, no changes were detected in fish biomass or
+invertebrate and fish densities (\ref{fig:indicators}B-D), where effect
+sizes oscillated around zero without clear trends. Full tables with
+model coefficients are presented in the supplementary materials
+(\textbf{S2 Table}, \textbf{S3 Table}, \textbf{S4 Table}).
 
-\subsection{Socioeconomic}\label{socioeconomic}
+\subsection{Socioeconomic effects}\label{socioeconomic-effects}
 
@@
 (left column: invertebrates; right column: fish). Points are jittered
-hotizontally to avoid overplotting. Points indicate the effect size, and
-errorbars standard errors. Years have been centered to year of
-implementation.}
+hotizontally to avoid overplotting. Points indicate the effect size and
+standard errors. Years have been centered to year of implementation.}
 \end{figure}
@@
 circles) and Maria Elena (ME; blue triangles). Points indicate the
-effect size, and errorbars standard errors. Years have been centered to
-year of implementation.}
+effect size and standard errors. Years have been centered to year of
+implementation.}
 \end{figure}
-
-\clearpage
 
@@
 pertaining landings and revenues showed little to no change after
-reserve implementation. The lack of evidence of the effectiveness of
-these reserves is surprising since most of the communities show a
-positive perception about their performance and continue to support
-their presence \citep{ayer_2018}. Analyzing the shortcomings of our
-study and understanding the social-ecological context in which these
-communities and their reserves operate might provide insights to this
-question.
+reserve implementation. The coastal ecosystems where these reserves are
+located have been profoundly affected by climatic and oceanographic
+extremes, including warming events, extreme storms and prolonged hypoxia
+\citep{micheli_2012-EU}, (\textbf{Beas et al. in prep., Woodson et al in
+review}). Despite the lack of evidence of the effectiveness of these
+reserves, most of the communities show a positive perception about their
+performance and continue to support their presence \citep{ayer_2018}.
+Understanding the social-ecological context in which these communities
+and their reserves operate might provide insights to this.
 
@@
 \citep{depalma_2018}. Our approach to evaluate the temporal and spatial
-changes provides a more robust measure of reserve effectiveness.
-However, this method assumes control sites are a plausible
-counterfactual for treated sites. This implies that treated sites would
-have followed the same trend as control sites, had the reserves not been
-implemented. Nonetheless, overall trends for each site don't show any
-significant increases, supporting our findings of lack of change in the
-indicators used (\textbf{S2 Figure}, \textbf{S3 Figure}, \textbf{S4
-Figure}, \textbf{S5 Figure}, \textbf{S6 Figure}).
+changes provides a more robust measure of reserve effectiveness. For
+example, we capture previously described patterns like the rapid
+increase observed for changes in lobster densities for Isla Natividad on
+the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A) occur a
+year after the hypoxia events described by \citet{micheli_2012-EU},
+which caused mass mortality of sedentary organisms such as abalone and
+sea urchins, but not lobster and finfish. Yet, our empirical approach
+assumes control sites are a plausible counterfactual for treated sites.
+This implies that treated sites would have followed the same trend as
+control sites, had the reserves not been implemented. Nonetheless,
+overall trends for each site don't show any significant increases,
+supporting our findings of lack of change in the indicators used
+(\textbf{S2 Figure}, \textbf{S3 Figure}, \textbf{S4 Figure}, \textbf{S5
+Figure}, \textbf{S6 Figure}).
 
@@
 important factors that influence reserve effectiveness
-\citep{edgar_2014-UO}. Isla Natividad has the oldest reserve, and our
-SES analysis suggests that all communities have a well-established
-community-based enforcement system. With these characteristics, one
-would expect the reserves to be effective. Maria Elena and Punta Herrero
-are relatively young reserves (\emph{i.e.} \textless{} 5 years old);
-other community-based marine reserves in tropical ecosystems may take up
-to six years to show a spillover effect \citep{dasilva_2015-zX}.
+\citep{edgar_2014-UO} \textbf{Babcock PNAS 2010}. Isla Natividad has the
+oldest reserves, and our SES analysis suggests that all communities have
+a well-established community-based enforcement system. With these
+characteristics, one would expect the reserves to be effective. Maria
+Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
+\textless{} 5 years old) and effects may not yet be evident due to the
+short duration of protection, relative to the life histories of the
+protected species; other community-based marine reserves in tropical
+ecosystems may take up to six years to show a spillover effect
+\citep{dasilva_2015-zX}.
 
@@
 \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
-attributed to reserves being too small. Previous research has shown that
-reserves in Isla Natividad yield fishery benefits for the abalone
-fishery \citep{rossetto_2015-V0}. Abalone are less mobile than lobsters,
-and perhaps the reserves provide enough protection to these sesile
+attributed to poor ecological coherence in reserve dsign (\emph{sensu}
+\citet{rees_2018}. Previous research has shown that reserves in Isla
+Natividad yield fishery benefits for the abalone fishery
+\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
+perhaps the reserves provide enough protection to these sedentary
 invertebrates, but not lobsters. Design principles developed by
@@
 fisheries benefits if initial population sizes are low and the fishery
-is poorly managed \citep{hilborn_2006}. Both lobster fisheries were, at
-some point, certified by the Marine Stewardship Council
-\citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed
-via species-specific minimum catch sizes, seasonal closures, protection
-of ``berried'' females, and escapement windows where traps are allowed
-\citep{dof_website_1993}. It is uncertain whether such a well-managed
-fishery will experience additional benefits from marine reserves.
-Additionally, \citet{gelcich_2008} has shown that TURFs alone can have
-greater biomass and richness than areas operating under open acces.
-These increased attributes perhaps minimize the difference between TURF
-and reserve. Further research should focus on evaluating sites in the
-reserve, TURF, and open access areas.
+is poorly managed \citep{hilborn_2006}. Both lobster fisheries were
+certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
+Additionally, lobster fisheries are managed via species-specific minimum
+catch sizes, seasonal closures, protection of ``berried'' females, and
+escapement windows where traps are allowed \citep{dof_website_1993}. It
+is uncertain whether such a well-managed fishery will experience
+additional benefits from marine reserves. Additionally,
+\citet{gelcich_2008} has shown that TURFs alone can have greater biomass
+and richness than areas operating under open acces. These increased
+attributes perhaps minimize the difference between TURF and reserve.
+Further research should focus on evaluating sites in the reserve, TURF,
+and open access areas.
 
-While reserves fail to provide fishery benefits, there are a number of
-additional ecological, fisheries, and social benefits. Marine reserves
-provide protection to a wider range of species and vulnerable habitat,
-like coral reefs. These sites can serve as an insurance against
-uncertainty and errors in fisheries management, as well as environmental
-shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU}.
-Self-regulation of fishing effort (\emph{i.e.} reduction in harvest) can
-serve as a way to compensate for future declines associated to
-environmental variation \citep{finkbeiner_2018}. Embarking in a marine
-conservation project can bring the community together, which promotes
-social cohesion and builds social capital. Furthermore, showing
-commitment to marine conservation and sustainable fishing practices
-allows fishers to have greater bargaining power and leverage over
-fisheries management \citep{prezramrez_2012}.
+While reserves have failed to provide fishery benefits up to now, there
+are a number of additional ecological, fisheries, and social benefits.
+Marine reserves provide protection to a wider range of species and
+vulnerable habitat, like coral reefs. These sites can serve as an
+insurance against uncertainty and errors in fisheries management, as
+well as environmental shocks
+\citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (\textbf{, Aalto et
+al, in press Am Nat}). Self-regulation of fishing effort (\emph{i.e.}
+reduction in harvest) can serve as a way to compensate for future
+declines associated to environmental variation \citep{finkbeiner_2018}.
+Embarking in a marine conservation project can bring the community
+together, which promotes social cohesion and builds social capital.
+Furthermore, showing commitment to marine conservation and sustainable
+fishing practices allows fishers to have greater bargaining power and
+leverage over fisheries management \citep{prezramrez_2012}.
 

### 2018-08-16 scripts/map_for_poster.R
Added script that created a map for the poster

@@ -1,0 +1,46 @@
+library(sf)
+library(ggplot2)
+
+coast <- rnaturalearth::ne_countries(continent = "north america", returnclass = "sf", scale = "medium")
+
+zrp <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
+  st_centroid()
+
+rescom <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
+  st_centroid()
+
+zrp_this <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
+  filter(Name %in% c("El Cabezo",
+                     "Gallineros",
+                     "Punta Loria",
+                     "San Roman Norte",
+                     "San Roman Sur",
+                     "Anegado de Chal",
+                     "Laguna Canche Balam",
+                     "Niluc",
+                     "Mimis",
+                     "El Faro (Pesca de Langosta)",
+                     "El Faro")) %>% 
+  head(1) %>% 
+  st_centroid()
+
+rescom_this <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
+  filter(Nombre %in% c("La_Plana", "PuntaPrieta")) %>% 
+  head(1) %>% 
+  st_centroid()
+
+
+map <- ggplot() +
+  geom_sf(data = coast, fill = "grey50", color = "black", size = 0.1) +
+  geom_sf(data = zrp, fill = "red", color = "black", shape = 21, size = 2) +
+  geom_sf(data = rescom, fill = "red", color = "black", shape = 21, size = 2) +
+  geom_sf(data = zrp_this, fill = "steelblue", color = "black", shape = 21, size = 2) +
+  geom_sf(data = rescom_this, fill = "steelblue", color = "black", shape = 21, size = 2) +
+  annotate(geom = "text", x = -102, y = 25, label = "Mexico") +
+  annotate(geom = "text", x = -90, y = 25, label = "Gulf of\nMexico") +
+  annotate(geom = "text", x = -112, y = 15, label = "Pacific\nOcean") +
+  lims(x = c(-120, -80), y = c(10, 35)) +
+  theme(panel.grid.major = element_line(color = "transparent")) +
+  labs(x = "", y = "")
+
+ggsave(plot = map, filename = "figures/map.png", width = 6, height = 4, dpi = 600)  

### 2018-09-02 docs/paper/Villasenor-Derbez.Rmd
Added Arturo to ms

@@
 rtitle: "Community-based marine reserves: `r Sys.Date()`"
-author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
+author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Arturo Hernández-Velasco$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
 institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline

### 2018-09-02 docs/paper/Villasenor-Derbez.tex
Added Arturo to ms

@@
 \def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira Aceves-Bueno\(^{1}\),
-Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\), Jorge Torre\(^{2}\)}
+Álvin Suarez\(^{2}\), Arturo Hernández-Velasco\(^{2}\), Stuart
+Fulton\(^{2}\), Jorge Torre\(^{2}\)}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 
-\title[Community-based marine reserves: 2018-08-16]{Effectiveness of community-based marine reserves in small-scale
+\title[Community-based marine reserves: 2018-08-30]{Effectiveness of community-based marine reserves in small-scale
 fisheries}

### 2018-09-13 docs/paper/Villasenor-Derbez.Rmd
Last changes before sending out to authors for last round of reviews

--- 2018-08-16 docs/paper/Villasenor-Derbez.Rmd
@@
 rtitle: "Community-based marine reserves: `r Sys.Date()`"
-author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
+author: "Juan Carlos Villaseñor-Derbez$^{1,2,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Arturo Hernández-Velasco$^{2}$, Jorge Torre$^{2}$, Fiorenza Micheli&^{3}^&"
 institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
-$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico"
+$^{2}$Comunidad y Biodiversidad A.C., Guaymas, Mexico\\newline
+$^{3}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, California"
 corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
@@
 Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintainance of healthy resources \citep{afflerbach_2014-HP,lester_2017}.
-Although in theory these systems are successful \citep{costello_2010-Ix}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017}.
+Although in theory these systems are successful \citep{costello_2010-Ix}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017,smallhornwest_2018}.
 
@@
 The science of marine reserves has largely focused on understanding the ecological effects of these areas, which include increased biomass, species richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
-Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1} [**add De Leo and Micheli 2015**](http://dx.doi.org/10.1098/rstb.2014.0276).
+Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}.
 Other research has focused on the relationship between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
@@
 These communities are representative of their region in terms of ecology, socioeconomic, and governance aspects.
-Isla Natividad, for example, is part of a greater group of fishing cooperatives that assamble into a Federation of Cooperatives (**FEDECOOP**).
-This group of X communities has been identified as a cohesive group that often cooperate to better manage their resources \citep{mccay_2017-1m,mccay_2014-CN} (**Aceves, 2018?**).
-All these communities occur in the sourthenmost raange distribution of Giant Kelp, and are still under the influence of the California Current System.
-Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean.
-Together, these three communities provide an accurate representation of all other fishing communities in Mexico.
+Isla Natividad, for example, is part of a greater group of fishing cooperatives belonging to a Federation of Fishing Cooperatives.
+This group has been identified as a cohesive group that often cooperates to better manage their resources \citep{mccay_2017-1m,mccay_2014-CN,acevesbueno_2017}.
+Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation.
+Together, these three communities provide an accurate representation of other fishing communities in each of their regions.
 While each region has more communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth analysis that we undertake.
-Yet, given the similarities among communities and the socioeconomic and governance setting under which they opperate, it is sage to generalize our results to other communities in Mexico and other regions around the world.
+Yet, given the similarities among communities and the socioeconomic and governance setting under which they opperate, it is safe to cautiously generalize our results to other communities in Mexico and other regions around the world.
 
@@
 
-Data for the qualitative analysis of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities.
+Data for the operationalization of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities.
 These include information on the resource system, the resource units, actors, and the governance system itself (**S1 Table**).
@@
 Data and code are available on [github.com](https://github.com/jcvdav/ReserveEffect).
+
+We use the social-ecologycal system (SES) framework to evaluate each community as a means of providing an explanation to the biological and socioeconomic results.
+The use the SES framework standardizes our analysis and allows us to communicate our results in a common language across fields.
+We based our variable selection primarily on \citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed Mexican fishing cooperatives using this framework.
+**S1 Table** shows the selected variables, the definition and selected indicators and scoring system.
+
+We followed \citet{leslie_2015-na} to operationalize the implementation of the SES framework.
+We analyzed 18 variables, that fitted within the first, second and third tires of the framework.
+Categorical variables were scored as binary data (0 or 1).
+Numerical variables were given a continuous score, and then standardized between 0 and 1.
+For example, the score on number of actors was relative to the smallest values (21 fishers) which was given a score of one since theoretically smaller groups facilitate coordination \citep{viana_2018}.
+Each component of the SES framework had a cumulative value of 1 and was equally weighted to calculate the final score.
 
@@
 A first possible explanation for the lack of effectivenes may be the young age of the reserves.
-Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO} **Babcock PNAS 2010**.
+Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO,babcock_2010}.
 Isla Natividad has the oldest reserves, and our SES analysis suggests that all communities have a well-established community-based enforcement system.
@@
 
-JC and EA analyzed and interpreted data, discussed the results, and wrote the first draft. AS, SF and JT discussed the results and edited the manuscript.
+JC and EA analyzed and interpreted data, discussed the results, and wrote the first draft.
+AHV coordinated fieldwork and collected the data.
+AS, AHV, SF, JT, and FM discussed the results and edited the manuscript.
 
@@
 
+**Tell me about your funding**
+
 JCVD: CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the Latin American Fisheries Fellowship.
 
-AS, SF and JT Walton Family Foundation, Summit Foundation, and Oak Foundation.
+AS, AHV, SF and JT Walton Family Foundation, Summit Foundation, and Oak Foundation.
 
 \section*{Acknowledgments}
-The authors wish to acknowledge Arturo Hernández and Imelda Amador for contributions on the governance data, as well as pre-processing biological data.
-This study would have not been possible without the effort by members of the communities here mentioned, who collected the biological data.
+The authors wish to acknowledge Imelda Amador for contributions on the governance data, as well as pre-processing biological data.
+This study would have not been possible without the effort by members of the fishing communities here mentioned, who participated in the data-collection process.
 

### 2018-09-13 docs/paper/Villasenor-Derbez.tex
Last changes before sending out to authors for last round of reviews

--- 2018-08-16 docs/paper/Villasenor-Derbez.tex
@@
 \def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
-\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira Aceves-Bueno\(^{1}\),
-Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\), Jorge Torre\(^{2}\)}
+\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,2,*}\), Eréndira
+Aceves-Bueno\(^{1}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
+Arturo Hernández-Velasco\(^{2}\), Jorge Torre\(^{2}\), Fiorenza
+Micheli\&\textsuperscript{\{3\}}\&}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 of California, Santa Barbara, Santa Barbara, CA,
-USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Mexico}
+USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas,
+Mexico\newline \(^{3}\)Hopkins Marine Station and Center for Ocean
+Solutions, Stanford University, Pacific Grove, California}
 % The Corresponding Author should be marked with an asterisk
@@
 
-\title[Community-based marine reserves: 2018-08-16]{Effectiveness of community-based marine reserves in small-scale
+\title[Community-based marine reserves: 2018-09-13]{Effectiveness of community-based marine reserves in small-scale
 fisheries}
@@
 evidence of their effectiveness and the drivers of their success
-\citep{afflerbach_2014-HP,lester_2017}.
+\citep{afflerbach_2014-HP,lester_2017,smallhornwest_2018}.
 
@@
 operates, as well as reserve size and the amount of larvae exported from
-these \citep{hilborn_2006,krueck_2017-J1}
-\href{http://dx.doi.org/10.1098/rstb.2014.0276}{\textbf{add De Leo and
-Micheli 2015}}. Other research has focused on the relationship between
-socioeconomic and governance structures and reserve effectiveness
+these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}. Other research has
+focused on the relationship between socioeconomic and governance
+structures and reserve effectiveness
 \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
@@
 ecology, socioeconomic, and governance aspects. Isla Natividad, for
-example, is part of a greater group of fishing cooperatives that
-assamble into a Federation of Cooperatives (\textbf{FEDECOOP}). This
-group of X communities has been identified as a cohesive group that
-often cooperate to better manage their resources
-\citep{mccay_2017-1m,mccay_2014-CN} (\textbf{Aceves, 2018?}). All these
-communities occur in the sourthenmost raange distribution of Giant Kelp,
-and are still under the influence of the California Current System.
-Likewise, Maria Elena and Punta Herrero are representative of fishing
-cooperatives in the Mexican Caribbean. Together, these three communities
-provide an accurate representation of all other fishing communities in
-Mexico. While each region has more communities that have established
-community-based TURF-reserves, available data would not allow us to
-perform the in-depth analysis that we undertake. Yet, given the
-similarities among communities and the socioeconomic and governance
-setting under which they opperate, it is sage to generalize our results
-to other communities in Mexico and other regions around the world.
+example, is part of a greater group of fishing cooperatives belonging to
+a Federation of Fishing Cooperatives. This group has been identified as
+a cohesive group that often cooperates to better manage their resources
+\citep{mccay_2017-1m,mccay_2014-CN,acevesbueno_2017}. Likewise, Maria
+Elena and Punta Herrero are representative of fishing cooperatives in
+the Mexican Caribbean, which are also part of a regional Federation.
+Together, these three communities provide an accurate representation of
+other fishing communities in each of their regions. While each region
+has more communities that have established community-based
+TURF-reserves, available data would not allow us to perform the in-depth
+analysis that we undertake. Yet, given the similarities among
+communities and the socioeconomic and governance setting under which
+they opperate, it is safe to cautiously generalize our results to other
+communities in Mexico and other regions around the world.
 
@@
 
-Data for the qualitative analysis of the social-ecological system were
+Data for the operationalization of the social-ecological system were
 collected at the community--level from official documents used in the
@@
 \href{https://github.com/jcvdav/ReserveEffect}{github.com}.
+
+We use the social-ecologycal system (SES) framework to evaluate each
+community as a means of providing an explanation to the biological and
+socioeconomic results. The use the SES framework standardizes our
+analysis and allows us to communicate our results in a common language
+across fields. We based our variable selection primarily on
+\citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed
+Mexican fishing cooperatives using this framework. \textbf{S1 Table}
+shows the selected variables, the definition and selected indicators and
+scoring system.
+
+We followed \citet{leslie_2015-na} to operationalize the implementation
+of the SES framework. We analyzed 18 variables, that fitted within the
+first, second and third tires of the framework. Categorical variables
+were scored as binary data (0 or 1). Numerical variables were given a
+continuous score, and then standardized between 0 and 1. For example,
+the score on number of actors was relative to the smallest values (21
+fishers) which was given a score of one since theoretically smaller
+groups facilitate coordination \citep{viana_2018}. Each component of the
+SES framework had a cumulative value of 1 and was equally weighted to
+calculate the final score.
 
@@
 important factors that influence reserve effectiveness
-\citep{edgar_2014-UO} \textbf{Babcock PNAS 2010}. Isla Natividad has the
-oldest reserves, and our SES analysis suggests that all communities have
-a well-established community-based enforcement system. With these
+\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
+reserves, and our SES analysis suggests that all communities have a
+well-established community-based enforcement system. With these
 characteristics, one would expect the reserves to be effective. Maria
@@
 JC and EA analyzed and interpreted data, discussed the results, and
-wrote the first draft. AS, SF and JT discussed the results and edited
-the manuscript.
+wrote the first draft. AHV coordinated fieldwork and collected the data.
+AS, AHV, SF, JT, and FM discussed the results and edited the manuscript.
 
 \section*{Funding}
+
+\textbf{Tell me about your funding}
 
@@
 
-AS, SF and JT Walton Family Foundation, Summit Foundation, and Oak
+AS, AHV, SF and JT Walton Family Foundation, Summit Foundation, and Oak
 Foundation.
@@
 
-The authors wish to acknowledge Arturo Hernández and Imelda Amador for
-contributions on the governance data, as well as pre-processing
-biological data. This study would have not been possible without the
-effort by members of the communities here mentioned, who collected the
-biological data.
+The authors wish to acknowledge Imelda Amador for contributions on the
+governance data, as well as pre-processing biological data. This study
+would have not been possible without the effort by members of the
+fishing communities here mentioned, who participated in the
+data-collection process.
 

### 2018-09-18 docs/paper/Villasenor-Derbez.Rmd
Updated manuscript, almost ready for submission

@@
 title: "Effectiveness of community-based marine reserves in small-scale fisheries"
-rtitle: "Community-based marine reserves: `r Sys.Date()`"
-author: "Juan Carlos Villaseñor-Derbez$^{1,2,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Arturo Hernández-Velasco$^{2}$, Jorge Torre$^{2}$, Fiorenza Micheli&^{3}^&"
+rtitle: "Community-based marine reserves"
+author: "Juan Carlos Villaseñor-Derbez$^{1,2,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Arturo Hernández-Velasco$^{2}$, Jorge Torre$^{2}$, Fiorenza Micheli$^{3}$"
 institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
-$^{2}$Comunidad y Biodiversidad A.C., Guaymas, Mexico\\newline
-$^{3}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, California"
+$^{2}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico\\newline
+$^{3}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA"
 corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
@@
 Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves is believed to improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
-In the last 4 years, 45 TURF reserves have been implemented in Mexico, but their effectiveness has not been formally evaluated.
+In the last six years, and following a 2012 regulation, 18 TURF-reserves have been implemented in Mexico; their effectiveness has not been formally evaluated.
 We combine causal inference techniques and a social-ecological systems framework to provide a holistic evaluation of community-based TURF reserves in three coastal communities in Mexico.
 We find that while reserves have not yet achieved their stated goal of increasing lobster densities, they continue to receive significant support from the fishing communities.
-A lack of ecological and socioeconomic effects or reserves likely results from a combination of factors.
+A lack of ecological and socioeconomic effects likely results from a combination of factors.
 First, the lobster fisheries are already well managed, and it is unlikely that reserves might have a detectable effect.
-Second, reserves are not large enough to protect lobster’s home ranges.
+Second, some of the reserves are not large enough to protect lobsters' home ranges.
 Third, some of these reserves might be too young for the effects to show.
-However, these reserves have shaped small-scale fishers’ way of thinking about marine management, which can provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
-keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, TURF-reserves, Social-Ecological Systems"
+However, these reserves have shaped small-scale fishers’ way of thinking about marine conservation, which can provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
+keywords: "TURF-reserves, Causal Inference, Social-Ecological Systems, Marine Protected Areas, Marine Conservation, Small-Scale Fisheries"
 output:
@@
 In an effort to bridge this normative gap, Civil Society Organizations (CSOs) have served as a link between fishers and government, and set out to create a legal framework that solve this governance issue.
-In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuge” (*Zona de Refugio Pesquero* ;\citet{nom}).
+In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuge” (*Zona de Refugio Pesquero*; \citet{nom}).
 Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
-Since 2014, 45 of community-based marine reserves along the Pacific, Gulf of California, and Mexican Caribbean coastlines have gained legal recognition.
+Since 2012, 45 of Fish Refuges have been created along the Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18 of them implemented as TURF-reserves.
 However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
@@
 
-# Materials and Methods
+# Methods
 
@@
 Maria Elena and Punta Herrero established eight marine reserves in 2012, and four marine reserves in 2013, respectively.
-All these reserves are legally recognized as Fishing Refugia since their creation.
-
-<!-- Revise the paragraph bellow. Add citations for CC system (Ramires-Valdez?), QRoo (Fulton, 2018, Caamal, half-way?). Cite noms as "opperate under similar fishing regulations". -->
+All these reserves have been legally recognized as Fishing Refuges since their creation \citep{dof_website_2012,dof_website_2013}.
 
@@
 Together, these three communities provide an accurate representation of other fishing communities in each of their regions.
-While each region has more communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth analysis that we undertake.
+While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth analysis that we undertake.
 Yet, given the similarities among communities and the socioeconomic and governance setting under which they opperate, it is safe to cautiously generalize our results to other communities in Mexico and other regions around the world.
@@
 location_labels <- data.frame(X = c(-115.24, -87.4611756),
-                              Y = c(27.85, 19.3183386),
+                              Y = c(27.85, 27.85),
                               Location = c("Isla Natividad", "Punta Herrero"),
@@
   geom_sf(data = W) +
-  geom_point(data = location_labels, aes(x = X, y = Y), color = "black", fill = "steelblue", size = 3, shape = 21) +
+  geom_point(data = location_labels,
+             aes(x = X, y = Y),
+             color = "black",
+             fill = "steelblue",
+             size = 3,
+             shape = 21) +
   theme_classic() +
@@
   annotate(geom = "text", x = -112, y = 20, label = "Pacific\nOcean") +
+  annotate(geom = "text", x = -116.4, y = 27.5, label = "B") +
+  annotate(geom = "text", x = -86.8, y = 18.2, label = "C") +
   labs(x = "", y = "")
@@
         axis.text = element_text(size = 7)) +
-  geom_label_repel(data = sam_points, aes(x = x, y = y, label = location), nudge_y = 0.05, nudge_x = 0.05) +
+  geom_label_repel(data = sam_points,
+                   aes(x = x, y = y, label = location),
+                   nudge_x = c(0.1, -0.05),
+                   nudge_y = c(-0.05, 0.06)) +
   labs(x = "", y = "") +
@@
 Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
-Trained divers record richness and abundances of fish and invertebrate species along replicate transects (30x 2 m each) at depths 5-20 m in the reserves and control sites \citep{fulton_2018}.
+Trained divers record richness and abundances of fish and invertebrate species along replicate transects (30x 2 m each) at depths 5-20 m in the reserves and control sites \citep{fulton_2018,fulton_2019}.
 Size structures are also collected during fish surveys.
@@
        implementation = c(2006, 2012, 2013)) %>%
+  mutate(percent = paste0(formatC(percent, digits = 2, format = "f"), "\\%")) %>% 
   knitr::kable(format = "latex",
-               caption = "\\label{table:com_sum} Summary of commuity--based marine reserves by community.",
+               caption = "\\label{table:com_sum} Summary of analyzed TURF-reserves by community.",
                col.names = c("Community", "TURF area ($\\mathrm{km}^{-2}$)", "Reserve area ($\\mathrm{km}^{-2}$)", "Percent as reserves","Year of implementation"),
@@
 Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA).
-Data contain monthly lobster landings (Kg) and revenues (MXP) for cooperatives with and without marine reserves(**Fig S1**).
+Data contain monthly lobster landings (Kg) and revenues (MXP) for cooperatives with and without marine reserves.
 Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017-1m,ayer_2018}.
@@
 Data for the operationalization of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities.
-These include information on the resource system, the resource units, actors, and the governance system itself (**S1 Table**).
+These include information on the resource system, the resource units, actors, and the governance system itself (Table \ref{tab:ses}).
 
@@
 We based our variable selection primarily on \citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed Mexican fishing cooperatives using this framework.
-**S1 Table** shows the selected variables, the definition and selected indicators and scoring system.
+Table \ref{tab:ses} shows the selected variables, the definition and selected indicators, and scoring system.
 
@@
 Likewise, no changes were detected in fish biomass or invertebrate and fish densities (\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
-Full tables with model coefficients are presented in the supplementary materials (**S2 Table**, **S3 Table**, **S4 Table**).
+Full tables with model coefficients are presented in the supplementary materials (S1 Table, S2 Table, S3 Table).
 
@@
 However, these changes are not significant and are associated to increased variation.
-All regression coefficients for each community and indicator are presented in **S5 Table**.
+Full tables with model coefficients are presented in the supplementary materials (S4 Table, S5 Table.
 
@@
 
-Although we have little information on the social dimension of these fisheries, we can use the social-ecological systems framework (**S1 Table**) to analyze the performance of each governance system (**S6 Table**).
+Although we have little information on the social dimension of these fisheries, we can use the social-ecological systems framework to analyze the performance of each governance system (Table \ref{tab:ses}).
 Our analysis shows that all communities analyzed share similarities in their Governance system which is based on cooperatives (GS5.2.3.2), with strong rules in use that include Operational rules (GS6.2), Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even Territorial use communal rights (GS6.1.4.3).
@@
 
-```{r, results = "asis"}
-C_IN <- conapesca %>% 
+```{r}
+conapesca %<>%
    filter(years_centered > -10) %>% 
   mutate(years_centered = as.factor(years_centered),
-         years_centered = fct_relevel(years_centered, "0")) %>% 
+         year = fct_relevel(years_centered, "0"),
+         peso_vivo = peso_vivo/1e3,
+         valor = valor/1e6,
+         loc = zona,
+         zona = ifelse(grupo == "Treated", "Reserva", "Control"),
+         grupo = loc)
+```
+
+```{r}
+C_IN <- conapesca %>% 
   filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
-  filter(zona == "PN") %>%
-  lm(peso_vivo/1e3 ~ grupo * years_centered + unidad_economica, data = .)
+  filter(grupo == "PN") %>%
+  lm(peso_vivo ~ zona * year + unidad_economica, data = .)
 
 C_ME <- conapesca %>% 
-  filter(years_centered > -10) %>% 
-  mutate(years_centered = as.factor(years_centered),
-         years_centered = fct_relevel(years_centered, "0")) %>% 
   filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
-  filter(zona == "MC") %>%
-  lm(peso_vivo/1e3 ~ grupo * years_centered + unidad_economica, data = .)
+  filter(grupo == "MC") %>%
+  lm(peso_vivo ~ zona * year + unidad_economica, data = .)
 
 V_IN <- conapesca %>% 
-  filter(years_centered > -10) %>% 
-  mutate(years_centered = as.factor(years_centered),
-         years_centered = fct_relevel(years_centered, "0")) %>% 
   filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
-  filter(zona == "PN") %>%
-  lm(valor/1e6 ~ grupo * years_centered + unidad_economica, data = .)
+  filter(grupo == "PN") %>%
+  lm(valor ~ zona * year + unidad_economica, data = .)
 
 V_ME <- conapesca %>% 
-  filter(years_centered > -10) %>% 
-  mutate(years_centered = as.factor(years_centered),
-         years_centered = fct_relevel(years_centered, "0")) %>% 
   filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
-  filter(zona == "MC") %>%
-  lm(valor/1e6 ~ grupo * years_centered + unidad_economica, data = .)
+  filter(grupo == "MC") %>%
+  lm(valor ~ zona * year + unidad_economica, data = .)
 
@@
       robust_se() %>% 
-      filter(grepl(pattern = "grupoTreated:years_centered", x = term)) %>% 
-      mutate(years_centered = as.numeric(gsub(pattern = "grupoTreated:years_centered",
+      filter(grepl(pattern = "zonaReserva:year", x = term)) %>% 
+      mutate(years_centered = as.numeric(gsub(pattern = "zonaReserva:year",
                                               replacement = "",
@@
 
+```{r}
+save(L_IN, L_ME, L_PH, B_IN, B_ME, B_PH, Ni_IN, Ni_ME, Ni_PH, N_IN, N_ME, N_PH, C_IN, C_ME, V_IN, V_ME, file = here::here("data", "results", "results.RData"))
+```
+
 # Discussion
@@
 This implies that treated sites would have followed the same trend as control sites, had the reserves not been implemented.
-Nonetheless, overall trends for each site don't show any significant increases, supporting our findings of lack of change in the indicators used (**S2 Figure**, **S3 Figure**, **S4 Figure**, **S5 Figure**, **S6 Figure**).
+Nonetheless, temporal trends for each site don't show any significant increases, supporting our findings of lack of change in the indicators used.
 
@@
 
-Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve dsign (*sensu* \citet{rees_2018}.
+Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (*sensu* \citet{rees_2018}.
 Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
@@
 
-While reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits.
+While the evaluated reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits.
 Marine reserves provide protection to a wider range of species and vulnerable habitat, like coral reefs.
-These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as environmental shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (**, Aalto et al, in press Am Nat**).
+These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as environmental shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (**,Aalto et al, in press**).
 Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
-Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital.
+Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}.
 Furthermore, showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
@@
 
-**Tell me about your funding**
-
-JCVD: CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the Latin American Fisheries Fellowship.
-
-AS, AHV, SF and JT Walton Family Foundation, Summit Foundation, and Oak Foundation.
+JCVD recevied funding from CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the Latin American Fisheries Fellowship Program.
+AS, AHV, SF and JT received funding from Marisla Foundation, Packard Foundation, Walton Family Foundation, Summit Foundation, and Oak Foundation.
 
@@
 
-\section*{Supplemental Data}
- \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
-LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder
-
-\paragraph*{S1 Figure}
-\label{S1_Figure}
-Map of control and treated sites in A and control and treated landings in B
-
-\paragraph*{S2 Figure}
-\label{S2_Figure}
-Time series of biological indicators for IN
-
-\paragraph*{S3 Figure}
-\label{S3_Figure}
-Time series of biological indicators for  ME
-
-\paragraph*{S4 Figure}
-\label{S4_Figure}
-Time series of biological indicators for  PH
-
-\paragraph*{S5 Figure}
-\label{S5_Figure}
-Time series of economic indicators for ME
-
-\paragraph*{S6 Figure}
-\label{S6_Figure}
-Time series of economic indicators for PH
-
-\paragraph*{S1 Table}
-\label{S1_Table}
-Coefficient estimates for biological indicators in Isla Natividad
-
-\paragraph*{S2 Table}
-\label{S2_Table}
-Coefficient estimates for biological indicators in Maria Elena
-
-\paragraph*{S3 Table}
-\label{S3_Table}
-Coefficient estimates for biological indicators in Punta Herrero
-
-\paragraph*{S4 Table}
-\label{S4_Table}
-Coefficient estimates for economic indicators
-
-```{r, results = "asis", eval = F}
-
-se <- list(robust_se_list(L_IN),
-           robust_se_list(L_ME),
-           robust_se_list(L_PH))
-
-stargazer::stargazer(list(L_IN, L_ME, L_PH),
-                     se = se,
-                     t.auto = T,
-                     p.auto = F,
-                     type = "html",
-                     single.row = T,
-                     model.names = F,
-                     column.labels = c("IN", "ME", "PH"),
-                     dep.var.labels.include = FALSE,
-                     model.numbers = FALSE,
-                     digits = 2,
-                     intercept.bottom = F,
-                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
-                     out = "lobster_models.html")
-```
-
-```{r, eval = F}
-se <- list(robust_se_list(B_IN),
-           robust_se_list(B_ME),
-           robust_se_list(B_PH))
-
-stargazer::stargazer(list(B_IN, B_ME, B_PH),
-                     se = se,
-                     t.auto = T,
-                     p.auto = F,
-                     type = "html",
-                     single.row = T,
-                     model.names = F,
-                     column.labels = c("IN", "ME", "PH"),
-                     dep.var.labels.include = FALSE,
-                     model.numbers = FALSE,
-                     digits = 2,
-                     intercept.bottom = F,
-                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
-                     out = "lobster_models.html")
-```
-
-```{r, eval = F}
-se <- list(robust_se_list(Ni_IN),
-           robust_se_list(Ni_ME),
-           robust_se_list(Ni_PH))
-
-stargazer::stargazer(list(Ni_IN, Ni_ME, Ni_PH),
-                     se = se,
-                     t.auto = T,
-                     p.auto = F,
-                     type = "html",
-                     single.row = T,
-                     model.names = F,
-                     column.labels = c("IN", "ME", "PH"),
-                     dep.var.labels.include = FALSE,
-                     model.numbers = FALSE,
-                     digits = 2,
-                     intercept.bottom = F,
-                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
-                     out = "lobster_models.html")
-```
-
-```{r, eval = F}
-se <- list(robust_se_list(N_IN),
-           robust_se_list(N_ME),
-           robust_se_list(N_PH))
-
-stargazer::stargazer(list(N_IN, N_ME, N_PH),
-                     se = se,
-                     t.auto = T,
-                     p.auto = F,
-                     type = "html",
-                     single.row = T,
-                     model.names = F,
-                     column.labels = c("IN", "ME", "PH"),
-                     dep.var.labels.include = FALSE,
-                     model.numbers = FALSE,
-                     digits = 2,
-                     intercept.bottom = F,
-                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
-                     out = "lobster_models.html")
-```
-
-```{r, eval = F}
-se <- list(robust_se_list(C_IN),
-           robust_se_list(C_ME),
-           robust_se_list(V_IN),
-           robust_se_list(V_ME))
-
-stargazer::stargazer(list(C_IN, C_ME, V_IN, V_ME),
-                     se = se,
-                     t.auto = T,
-                     p.auto = F,
-                     type = "html",
-                     single.row = T,
-                     model.names = F,
-                     column.labels = c("Landings IN", "Landings ME", "Revenue IN", "Revenue ME"),
-                     dep.var.labels.include = FALSE,
-                     model.numbers = FALSE,
-                     digits = 2,
-                     intercept.bottom = F,
-                     title = "\\paragraph*{S1 Table}\\label{S1_Table}",
-                     out = "lobster_models.html")
-```
-
 \bibliographystyle{frontiersinSCNS_ENG_HUMS}

### 2018-09-18 docs/paper/Villasenor-Derbez.tex
Updated manuscript, almost ready for submission

@@
 Arturo Hernández-Velasco\(^{2}\), Jorge Torre\(^{2}\), Fiorenza
-Micheli\&\textsuperscript{\{3\}}\&}
+Micheli\(^{3}\)}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 of California, Santa Barbara, Santa Barbara, CA,
-USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas,
+USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Sonora,
 Mexico\newline \(^{3}\)Hopkins Marine Station and Center for Ocean
-Solutions, Stanford University, Pacific Grove, California}
+Solutions, Stanford University, Pacific Grove, CA, USA}
 % The Corresponding Author should be marked with an asterisk
@@
 
-\title[Community-based marine reserves: 2018-09-13]{Effectiveness of community-based marine reserves in small-scale
+\title[Community-based marine reserves]{Effectiveness of community-based marine reserves in small-scale
 fisheries}
@@
 environmental variability and management errors, while ensuring that
-fishers reap the benefits of conservation investments. In the last 4
-years, 45 TURF reserves have been implemented in Mexico, but their
-effectiveness has not been formally evaluated. We combine causal
-inference techniques and a social-ecological systems framework to
-provide a holistic evaluation of community-based TURF reserves in three
-coastal communities in Mexico. We find that while reserves have not yet
-achieved their stated goal of increasing lobster densities, they
-continue to receive significant support from the fishing communities. A
-lack of ecological and socioeconomic effects or reserves likely results
-from a combination of factors. First, the lobster fisheries are already
-well managed, and it is unlikely that reserves might have a detectable
-effect. Second, reserves are not large enough to protect lobster's home
-ranges. Third, some of these reserves might be too young for the effects
-to show. However, these reserves have shaped small-scale fishers' way of
-thinking about marine management, which can provide a foundation for
+fishers reap the benefits of conservation investments. In the last six
+years, and following a 2012 regulation, 18 TURF-reserves have been
+implemented in Mexico; their effectiveness has not been formally
+evaluated. We combine causal inference techniques and a
+social-ecological systems framework to provide a holistic evaluation of
+community-based TURF reserves in three coastal communities in Mexico. We
+find that while reserves have not yet achieved their stated goal of
+increasing lobster densities, they continue to receive significant
+support from the fishing communities. A lack of ecological and
+socioeconomic effects likely results from a combination of factors.
+First, the lobster fisheries are already well managed, and it is
+unlikely that reserves might have a detectable effect. Second, some of
+the reserves are not large enough to protect lobsters' home ranges.
+Third, some of these reserves might be too young for the effects to
+show. However, these reserves have shaped small-scale fishers' way of
+thinking about marine conservation, which can provide a foundation for
 establishing additional, larger marine reserves needed to effectively
@@
 \tiny
- \keyFont{ \section{Keywords:} Marine Protected Areas, Marine Conservation, Small-Scale Fisheries,
-Citizen Science, TURF-reserves, Social-Ecological Systems}
+ \keyFont{ \section{Keywords:} TURF-reserves, Causal Inference, Social-Ecological Systems, Marine
+Protected Areas, Marine Conservation, Small-Scale Fisheries}
 
@@
 community-based reserves as ``Fish Refuge'' (\emph{Zona de Refugio
-Pesquero} ;\citet{nom}). Fish refuges can be implemented as temporal or
+Pesquero}; \citet{nom}). Fish refuges can be implemented as temporal or
 partial reserves, which can protect one, some, or all resources within
-their boundaries. Since 2014, 45 of community-based marine reserves
-along the Pacific, Gulf of California, and Mexican Caribbean coastlines
-have gained legal recognition. However, their effectiveness has not yet
-been formally evaluated and reported in the scientific literature.
+their boundaries. Since 2012, 45 of Fish Refuges have been created along
+the Pacific, Gulf of California, and Mexican Caribbean coastlines, with
+18 of them implemented as TURF-reserves. However, their effectiveness
+has not yet been formally evaluated and reported in the scientific
+literature.
 
@@
 
-\section{Materials and Methods}\label{materials-and-methods}
+\section{Methods}\label{methods}
 
@@
 Punta Herrero established eight marine reserves in 2012, and four marine
-reserves in 2013, respectively. All these reserves are legally
-recognized as Fishing Refugia since their creation.
+reserves in 2013, respectively. All these reserves have been legally
+recognized as Fishing Refuges since their creation
+\citep{dof_website_2012,dof_website_2013}.
 
@@
 other fishing communities in each of their regions. While each region
-has more communities that have established community-based
+has additional communities that have established community-based
 TURF-reserves, available data would not allow us to perform the in-depth
@@
 (30x 2 m each) at depths 5-20 m in the reserves and control sites
-\citep{fulton_2018}. Size structures are also collected during fish
-surveys. We define control sites as regions with habitat characteristics
-similar to the corresponding reserves, and that presumably had a similar
-probability of being selected as reserves during the design phase. We
-focus our evaluation on sites where data are available for reserve and
-control sites, before and after the implementation of the reserve. This
-provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
-sampling design that allows us to capture and control for temporal and
-spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and
-causal inference techniques have proven effective to evaluate marine
-reserves, as they allow us to causally attribute observed changes to the
-intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites
-were surveyed annually, and at least once before implementation of the
-reserves. Table \ref{table:com_sum} shows a summary of the TURF-reserves
-included in this study.
+\citep{fulton_2018,fulton_2019}. Size structures are also collected
+during fish surveys. We define control sites as regions with habitat
+characteristics similar to the corresponding reserves, and that
+presumably had a similar probability of being selected as reserves
+during the design phase. We focus our evaluation on sites where data are
+available for reserve and control sites, before and after the
+implementation of the reserve. This provides us with a
+Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
+allows us to capture and control for temporal and spatial dynamics
+\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
+techniques have proven effective to evaluate marine reserves, as they
+allow us to causally attribute observed changes to the intervention
+\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
+annually, and at least once before implementation of the reserves. Table
+\ref{table:com_sum} shows a summary of the TURF-reserves included in
+this study.
 
@@
 
-\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of commuity--based marine reserves by community.}
+\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of analyzed TURF-reserves by community.}
 \centering
-\begin{tabular}[t]{l|r|r|r|r}
+\begin{tabular}[t]{l|r|r|l|r}
 \hline
@@
 \hline
-Isla Natividad & 889.5 & 1.53 & 0.1720067 & 2006\\
+Isla Natividad & 889.5 & 1.53 & 0.17\% & 2006\\
 \hline
-Maria Elena & 353.1 & 0.10 & 0.0283206 & 2012\\
+Maria Elena & 353.1 & 0.10 & 0.03\% & 2012\\
 \hline
-Punta Herrero & 299.7 & 0.43 & 0.1434768 & 2013\\
+Punta Herrero & 299.7 & 0.43 & 0.14\% & 2013\\
 \hline
@@
 (Kg) and revenues (MXP) for cooperatives with and without marine
-reserves(\textbf{Fig S1}). Cooperatives incorporated in this analysis
-belong to larger regional-level Cooperative Federations, and are exposed
-to the same markets and institutional frameworks, making them plausible
-controls \citep{mccay_2017-1m,ayer_2018}. Landings and revenues were
-aggregated at the cooperative-year level, and revenues were adjusted to
-represent 2014 values by the Consumer Price Index for Mexico
-\citep{oecd_2017-VV} as:
+reserves. Cooperatives incorporated in this analysis belong to larger
+regional-level Cooperative Federations, and are exposed to the same
+markets and institutional frameworks, making them plausible controls
+\citep{mccay_2017-1m,ayer_2018}. Landings and revenues were aggregated
+at the cooperative-year level, and revenues were adjusted to represent
+2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}
+as:
 
@@
 information on the resource system, the resource units, actors, and the
-governance system itself (\textbf{S1 Table}).
+governance system itself (Table \ref{tab:ses}).
 
@@
 \citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed
-Mexican fishing cooperatives using this framework. \textbf{S1 Table}
-shows the selected variables, the definition and selected indicators and
-scoring system.
+Mexican fishing cooperatives using this framework. Table \ref{tab:ses}
+shows the selected variables, the definition and selected indicators,
+and scoring system.
 
@@
 sizes oscillated around zero without clear trends. Full tables with
-model coefficients are presented in the supplementary materials
-(\textbf{S2 Table}, \textbf{S3 Table}, \textbf{S4 Table}).
+model coefficients are presented in the supplementary materials (S1
+Table, S2 Table, S3 Table).
 
@@
 However, these changes are not significant and are associated to
-increased variation. All regression coefficients for each community and
-indicator are presented in \textbf{S5 Table}.
+increased variation. Full tables with model coefficients are presented
+in the supplementary materials (S4 Table, S5 Table.
 
@@
 Although we have little information on the social dimension of these
-fisheries, we can use the social-ecological systems framework
-(\textbf{S1 Table}) to analyze the performance of each governance system
-(\textbf{S6 Table}). Our analysis shows that all communities analyzed
-share similarities in their Governance system which is based on
-cooperatives (GS5.2.3.2), with strong rules in use that include
-Operational rules (GS6.2), Collective-choice rules (GS6.3),
-Constitutional rules (GS6.3), and even Territorial use communal rights
-(GS6.1.4.3). However, we identified important differences in terms of
-the actors, resource systems, and resource units. Although all
-communities show a high level of leadership (A5), the level of trust
-(A6.1) is lower in Punta Herrero. In general, the presence and success
-of conservation initiatives depends on the incentives of local
-communities to maintain a healthy status of the resources they depend
-upon \citep{jupiter_2017}. The enabling conditions for conservation seem
-to be strongly present in all communities. Due to the clarity of access
-rights and isolation, the benefits of conservation directly benefit the
-members of the fishing cooperative. These conditions have favored the
-development of efficient community-based enforcement systems.
+fisheries, we can use the social-ecological systems framework to analyze
+the performance of each governance system (Table \ref{tab:ses}). Our
+analysis shows that all communities analyzed share similarities in their
+Governance system which is based on cooperatives (GS5.2.3.2), with
+strong rules in use that include Operational rules (GS6.2),
+Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even
+Territorial use communal rights (GS6.1.4.3). However, we identified
+important differences in terms of the actors, resource systems, and
+resource units. Although all communities show a high level of leadership
+(A5), the level of trust (A6.1) is lower in Punta Herrero. In general,
+the presence and success of conservation initiatives depends on the
+incentives of local communities to maintain a healthy status of the
+resources they depend upon \citep{jupiter_2017}. The enabling conditions
+for conservation seem to be strongly present in all communities. Due to
+the clarity of access rights and isolation, the benefits of conservation
+directly benefit the members of the fishing cooperative. These
+conditions have favored the development of efficient community-based
+enforcement systems.
 
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
-\caption{\label{fig:unnamed-chunk-6}\label{fig:lobsters}Effect sizes for
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-7-1.pdf}
+\caption{\label{fig:unnamed-chunk-7}\label{fig:lobsters}Effect sizes for
 lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
@@
 control sites, had the reserves not been implemented. Nonetheless,
-overall trends for each site don't show any significant increases,
-supporting our findings of lack of change in the indicators used
-(\textbf{S2 Figure}, \textbf{S3 Figure}, \textbf{S4 Figure}, \textbf{S5
-Figure}, \textbf{S6 Figure}).
+temporal trends for each site don't show any significant increases,
+supporting our findings of lack of change in the indicators used.
 
@@
 \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
-attributed to poor ecological coherence in reserve dsign (\emph{sensu}
+attributed to poor ecological coherence in reserve design (\emph{sensu}
 \citet{rees_2018}. Previous research has shown that reserves in Isla
@@
 
-While reserves have failed to provide fishery benefits up to now, there
-are a number of additional ecological, fisheries, and social benefits.
-Marine reserves provide protection to a wider range of species and
-vulnerable habitat, like coral reefs. These sites can serve as an
-insurance against uncertainty and errors in fisheries management, as
-well as environmental shocks
-\citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (\textbf{, Aalto et
-al, in press Am Nat}). Self-regulation of fishing effort (\emph{i.e.}
-reduction in harvest) can serve as a way to compensate for future
-declines associated to environmental variation \citep{finkbeiner_2018}.
-Embarking in a marine conservation project can bring the community
-together, which promotes social cohesion and builds social capital.
+While the evaluated reserves have failed to provide fishery benefits up
+to now, there are a number of additional ecological, fisheries, and
+social benefits. Marine reserves provide protection to a wider range of
+species and vulnerable habitat, like coral reefs. These sites can serve
+as an insurance against uncertainty and errors in fisheries management,
+as well as environmental shocks
+\citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (\textbf{,Aalto et al,
+in press}). Self-regulation of fishing effort (\emph{i.e.} reduction in
+harvest) can serve as a way to compensate for future declines associated
+to environmental variation \citep{finkbeiner_2018}. Embarking in a
+marine conservation project can bring the community together, which
+promotes social cohesion and builds social capital \citep{fulton_2019}.
 Furthermore, showing commitment to marine conservation and sustainable
@@
 
-\textbf{Tell me about your funding}
-
-JCVD: CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the
-Latin American Fisheries Fellowship.
-
-AS, AHV, SF and JT Walton Family Foundation, Summit Foundation, and Oak
+JCVD recevied funding from CONACyT (Beca de Posgrados en el extranjero,
+CVU 669403) and the Latin American Fisheries Fellowship Program. AS,
+AHV, SF and JT received funding from Marisla Foundation, Packard
+Foundation, Walton Family Foundation, Summit Foundation, and Oak
 Foundation.
@@
 
-\section*{Supplemental Data}
-
-\href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material}
-should be uploaded separately on submission, if there are Supplementary
-Figures, please include the caption in the same file as the figure.
-LaTeX Supplementary Material templates can be found in the Frontiers
-LaTeX folder
-
-\paragraph*{S1 Figure}
-\label{S1_Figure}
-
-Map of control and treated sites in A and control and treated landings
-in B
-
-\paragraph*{S2 Figure}
-\label{S2_Figure}
-
-Time series of biological indicators for IN
-
-\paragraph*{S3 Figure}
-\label{S3_Figure}
-
-Time series of biological indicators for ME
-
-\paragraph*{S4 Figure}
-\label{S4_Figure}
-
-Time series of biological indicators for PH
-
-\paragraph*{S5 Figure}
-\label{S5_Figure}
-
-Time series of economic indicators for ME
-
-\paragraph*{S6 Figure}
-\label{S6_Figure}
-
-Time series of economic indicators for PH
-
-\paragraph*{S1 Table}
-\label{S1_Table}
-
-Coefficient estimates for biological indicators in Isla Natividad
-
-\paragraph*{S2 Table}
-\label{S2_Table}
-
-Coefficient estimates for biological indicators in Maria Elena
-
-\paragraph*{S3 Table}
-\label{S3_Table}
-
-Coefficient estimates for biological indicators in Punta Herrero
-
-\paragraph*{S4 Table}
-\label{S4_Table}
-
-Coefficient estimates for economic indicators
-
 \bibliographystyle{frontiersinSCNS_ENG_HUMS}\bibliography{references}

### 2018-09-20 docs/paper/Villasenor-Derbez.Rmd
Updated SES table and manuscript

@@
 We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A).
-The first one was created by the *Buzos y Pescadores de la Baja California* fishing cooperative, located in Isla Natividad in the Baja Peninsula (Fig \ref{fig:map}B).
+The first one was created by the *Buzos y Pescadores de la Baja California* fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig \ref{fig:map}B).
 The main fishery in the island is the spiny lobster (*Panulirus interruptus*), but other resources like finfish, sea cucumber, read sea urchin, snail, and abalone are also an important source of income.
@@
   st_transform(crs = proj) %>% 
-  mutate(evaluated = ifelse(Name %in% c("El Cabezo", "El Faro (Pesca de Langosta)", "El Faro"), T, F))
+  mutate(evaluated = ifelse(Name %in% c("El Cabezo", "El Faro (Pesca de Langosta)", "El Faro"), T, F),
+         Comunidad = ifelse(Name %in% c("El Faro", "Laguna Canche Balam", "El Faro (Pesca de Langosta)", "Anegado de Chal"), "Puna Herrero", "Maria Elena"))
 
@@
 All sites were surveyed annually, and at least once before implementation of the reserves.
-Table \ref{table:com_sum} shows a summary of the TURF-reserves included in this study.
-
-\clearpage
-
-```{r}
-tibble(Community = c("Isla Natividad", "Maria Elena", "Punta Herrero"),
-       TURF = c(889.5, 353.1, 299.7),
-       reserve = c(1.53, 0.10, 0.43),
-       percent = reserve / TURF * 100,
-       implementation = c(2006, 2012, 2013)) %>%
-  mutate(percent = paste0(formatC(percent, digits = 2, format = "f"), "\\%")) %>% 
-  knitr::kable(format = "latex",
-               caption = "\\label{table:com_sum} Summary of analyzed TURF-reserves by community.",
-               col.names = c("Community", "TURF area ($\\mathrm{km}^{-2}$)", "Reserve area ($\\mathrm{km}^{-2}$)", "Percent as reserves","Year of implementation"),
-               escape = F) %>% 
-  kable_styling(latex_options = "HOLD_position")
-```
 
@@
 Data for the operationalization of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities.
-These include information on the resource system, the resource units, actors, and the governance system itself (Table \ref{tab:ses}).
+These include information on the resource system, the resource units, actors, and the governance system itself (Table \ref{table:ses}).
 
@@
 We based our variable selection primarily on \citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed Mexican fishing cooperatives using this framework.
-Table \ref{tab:ses} shows the selected variables, the definition and selected indicators, and scoring system.
+We also incorporate other relevant information following \citet{difranco_2016-Xw,edgar_2014-UO} Table \ref{table:ses} shows the selected variables, their definition and selected indicators.
 
-We followed \citet{leslie_2015-na} to operationalize the implementation of the SES framework.
-We analyzed 18 variables, that fitted within the first, second and third tires of the framework.
-Categorical variables were scored as binary data (0 or 1).
-Numerical variables were given a continuous score, and then standardized between 0 and 1.
-For example, the score on number of actors was relative to the smallest values (21 fishers) which was given a score of one since theoretically smaller groups facilitate coordination \citep{viana_2018}.
-Each component of the SES framework had a cumulative value of 1 and was equally weighted to calculate the final score.
+```{r}
+read.csv(here::here("raw_data", "ses.csv")) %>% 
+  select(-Dim) %>% 
+  knitr::kable(caption = "\\label{table:ses}Variables used in our Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).",
+               format = "latex",
+               escape = F) %>% 
+  kable_styling(latex_options = "HOLD_position") %>% 
+  column_spec(column = 1, width = "6cm") %>% 
+  column_spec(column = 2, width = "12cm") %>% 
+  group_rows(group_label = "Resource System", start_row = 1, end_row = 4) %>% 
+  group_rows(group_label = "Resource Unit", start_row = 5, end_row = 6) %>% 
+  group_rows(group_label = "Actors", start_row = 7, end_row = 8) %>% 
+  group_rows(group_label = "Governance system", start_row = 9, end_row = 12)
+```
 
@@
 Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness.
-For example, we capture previously described patterns like the rapid increase observed for changes in lobster densities for Isla Natividad on the sixth year (*i.e.* 2012; Fig.
-\ref{fig:indicators}A) occur a year after the hypoxia events described by \citet{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
+For example, we capture previously described patterns like the rapid increase observed for lobster densities in Isla Natividad on the sixth year (*i.e.* 2012; Fig.
+\ref{fig:indicators}A) -a year after the hypoxia events described by \citet{micheli_2012-EU}- which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
 Yet, our empirical approach assumes control sites are a plausible counterfactual for treated sites.
 This implies that treated sites would have followed the same trend as control sites, had the reserves not been implemented.
-Nonetheless, temporal trends for each site don't show any significant increases, supporting our findings of lack of change in the indicators used.
+Nonetheless, temporal trends for each site don't show any significant increases (S1 Table, S2 Table, S3 Table), supporting our findings of lack of change in the indicators used.
 
@@
 With these characteristics, one would expect the reserves to be effective.
-Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 5 years old) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; other community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}.
+Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 6 years old) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; other community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}.
 
@@
 Design principles developed by \citet{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
-Furthermore, may favor implementation of reserves that pose low fishing costs due to their small size or location.
+Furthermore, fishers may favor implementation of reserves that pose low fishing costs due to their small size or location.
 Our analysis of economic data supports this, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
@@
 It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves.
-Additionally, \citet{gelcich_2008} has shown that TURFs alone can have greater biomass and richness than areas operating under open acces.
-These increased attributes perhaps minimize the difference between TURF and reserve.
-Further research should focus on evaluating sites in the reserve, TURF, and open access areas.
+Additionally, \citet{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open acces.
+These increased attributes perhaps minimize the difference between TURF and reserve, making it difficult to detect such a small difference.
+Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of ther eserves might be larger.
 
@@
 Furthermore, showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
+Furthermore, the lack of effectiveness observed in these reserves may not be generalizable to other reserves established under the same legal frameowrk (*i.e.* Fish Refges) in Mexico, and future research should aim at evaluating other areas that have also been established as bottom-up processes bt without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), or others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}).
 

### 2018-09-20 docs/paper/Villasenor-Derbez.tex
Updated SES table and manuscript

@@
 California} fishing cooperative, located in Isla Natividad in the Baja
-Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the
-spiny lobster (\emph{Panulirus interruptus}), but other resources like
-finfish, sea cucumber, read sea urchin, snail, and abalone are also an
-important source of income. In 2006, the community decided to implement
-two marine reserves within their fishing grounds to protect commercially
-important invertebrate species; mainly lobster and abalone. While these
-reserves obtained legal recognition only in 2018, they have been well
-enforced since their implementation.
+California Peninsula (Fig \ref{fig:map}B). The main fishery in the
+island is the spiny lobster (\emph{Panulirus interruptus}), but other
+resources like finfish, sea cucumber, read sea urchin, snail, and
+abalone are also an important source of income. In 2006, the community
+decided to implement two marine reserves within their fishing grounds to
+protect commercially important invertebrate species; mainly lobster and
+abalone. While these reserves obtained legal recognition only in 2018,
+they have been well enforced since their implementation.
 
@@
 \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
-annually, and at least once before implementation of the reserves. Table
-\ref{table:com_sum} shows a summary of the TURF-reserves included in
-this study.
-
-\clearpage
-
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of analyzed TURF-reserves by community.}
-\centering
-\begin{tabular}[t]{l|r|r|l|r}
-\hline
-Community & TURF area ($\mathrm{km}^{-2}$) & Reserve area ($\mathrm{km}^{-2}$) & Percent as reserves & Year of implementation\\
-\hline
-Isla Natividad & 889.5 & 1.53 & 0.17\% & 2006\\
-\hline
-Maria Elena & 353.1 & 0.10 & 0.03\% & 2012\\
-\hline
-Punta Herrero & 299.7 & 0.43 & 0.14\% & 2013\\
-\hline
-\end{tabular}
-\end{table}
+annually, and at least once before implementation of the reserves.
 
@@
 information on the resource system, the resource units, actors, and the
-governance system itself (Table \ref{tab:ses}).
+governance system itself (Table \ref{table:ses}).
 
@@
 
-\caption{\label{tab:unnamed-chunk-3}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
+\caption{\label{tab:unnamed-chunk-2}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
 \centering
@@
 \citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed
-Mexican fishing cooperatives using this framework. Table \ref{tab:ses}
-shows the selected variables, the definition and selected indicators,
-and scoring system.
+Mexican fishing cooperatives using this framework. We also incorporate
+other relevant information following
+\citet{difranco_2016-Xw,edgar_2014-UO} Table \ref{table:ses} shows the
+selected variables, their definition and selected indicators.
 
-We followed \citet{leslie_2015-na} to operationalize the implementation
-of the SES framework. We analyzed 18 variables, that fitted within the
-first, second and third tires of the framework. Categorical variables
-were scored as binary data (0 or 1). Numerical variables were given a
-continuous score, and then standardized between 0 and 1. For example,
-the score on number of actors was relative to the smallest values (21
-fishers) which was given a score of one since theoretically smaller
-groups facilitate coordination \citep{viana_2018}. Each component of the
-SES framework had a cumulative value of 1 and was equally weighted to
-calculate the final score.
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-3}\label{table:ses}Variables used in our Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6cm}|>{\raggedright\arraybackslash}p{12cm}}
+\hline
+Variable & Narrative\\
+\hline
+\multicolumn{2}{l}{\textbf{Resource System}}\\
+\hline
+\hspace{1em}Clarity of system boundaries - Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates.
+Reserve placement is decided by the community, and members trace reserve boundaries themselves.
+Fishers use GPS units and landmarks.\\
+\hline
+Size of resource system - TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
+\hline
+Size of resource system - Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
+\hline
+\hspace{1em}Age of reserves - Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
+\hline
+\multicolumn{2}{l}{\textbf{Resource Unit}}\\
+\hline
+\hspace{1em}Number of units (catch diversity) - Diversity of targeted species & Lobster is their main fishery of these three communities, but they also target finfish.
+Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates (sea cucumber, urchin, snail, abalone).\\
+\hline
+\hspace{1em}Stock status - Status of the main fishery & Lobster stocks are well managed; along with neighboring cooperatives, the lobster fishery is MSC certified.\\
+\hline
+\multicolumn{2}{l}{\textbf{Actors}}\\
+\hline
+\hspace{1em}Number of relevant actors - Number of fishers & IN = 98; ME = 80; PH = 21\\
+\hline
+\hspace{1em}Isolation - Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers.\\
+\hline
+\multicolumn{2}{l}{\textbf{Governance system}}\\
+\hline
+\hspace{1em}Property rights - Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster.
+These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
+\hline
+\hspace{1em}Operational rules - Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandate.
+These include larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.
+They've established penalties for breaking these rules or fishing inside the reserves.\\
+\hline
+\hspace{1em}Social monitoring - Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a designated group that monitors and enforces formal and internal operational rules.
+They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
+\hline
+\hspace{1em}Biophysical monitoring - Monitoring of biological resources, including targeted species & Every year since the creation of the reserves, along with Comunidad y Biodiversidad, fishers perform standardized underwater surveys in the reserves and fishing grounds.
+In more recent years, they have installed oceanographic sensors to monitor for changes in oceanographic variables (e.g.
+Temperature, Dissolved Oxygen, Salinity, pH).\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 example, we capture previously described patterns like the rapid
-increase observed for changes in lobster densities for Isla Natividad on
-the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A) occur a
-year after the hypoxia events described by \citet{micheli_2012-EU},
-which caused mass mortality of sedentary organisms such as abalone and
-sea urchins, but not lobster and finfish. Yet, our empirical approach
-assumes control sites are a plausible counterfactual for treated sites.
-This implies that treated sites would have followed the same trend as
-control sites, had the reserves not been implemented. Nonetheless,
-temporal trends for each site don't show any significant increases,
-supporting our findings of lack of change in the indicators used.
+increase observed for lobster densities in Isla Natividad on the sixth
+year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A) -a year after the
+hypoxia events described by \citet{micheli_2012-EU}- which caused mass
+mortality of sedentary organisms such as abalone and sea urchins, but
+not lobster and finfish. Yet, our empirical approach assumes control
+sites are a plausible counterfactual for treated sites. This implies
+that treated sites would have followed the same trend as control sites,
+had the reserves not been implemented. Nonetheless, temporal trends for
+each site don't show any significant increases (S1 Table, S2 Table, S3
+Table), supporting our findings of lack of change in the indicators
+used.
 
@@
 Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
-\textless{} 5 years old) and effects may not yet be evident due to the
+\textless{} 6 years old) and effects may not yet be evident due to the
 short duration of protection, relative to the life histories of the
@@
 adults and juveniles'', and suggest that reserves seeking to protect
-spiny lobsters should have at least 14 km across. Furthermore, may favor
-implementation of reserves that pose low fishing costs due to their
-small size or location. Our analysis of economic data supports this, as
-neither landings nor revenues showed the expected short-term costs
-associated to the first years of reserve implementation
+spiny lobsters should have at least 14 km across. Furthermore, fishers
+may favor implementation of reserves that pose low fishing costs due to
+their small size or location. Our analysis of economic data supports
+this, as neither landings nor revenues showed the expected short-term
+costs associated to the first years of reserve implementation
 \citep{ovando_2016-Wg}.
@@
 additional benefits from marine reserves. Additionally,
-\citet{gelcich_2008} has shown that TURFs alone can have greater biomass
-and richness than areas operating under open acces. These increased
-attributes perhaps minimize the difference between TURF and reserve.
-Further research should focus on evaluating sites in the reserve, TURF,
-and open access areas.
+\citet{gelcich_2008} have shown that TURFs alone can have greater
+biomass and richness than areas operating under open acces. These
+increased attributes perhaps minimize the difference between TURF and
+reserve, making it difficult to detect such a small difference. Further
+research should focus on evaluating sites in the reserve, TURF, and open
+access areas or similar Fish Refuges established without the presence of
+TURFs where the impact of ther eserves might be larger.
 
@@
 fishing practices allows fishers to have greater bargaining power and
-leverage over fisheries management \citep{prezramrez_2012}.
+leverage over fisheries management \citep{prezramrez_2012}. Furthermore,
+the lack of effectiveness observed in these reserves may not be
+generalizable to other reserves established under the same legal
+frameowrk (\emph{i.e.} Fish Refges) in Mexico, and future research
+should aim at evaluating other areas that have also been established as
+bottom-up processes bt without the presence of TURFs (\emph{e.g.}
+\citet{dof_websiteC_2012}), or others established through a top-down
+process (\emph{i.e.} \citet{dof_websiteU_2018}).
 

### 2018-09-24 docs/paper/Villasenor-Derbez.Rmd
Finalized manuscript. Waiting for Ere's adscription

@@
 corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
-email: "jvillasenor@bren.ucsb.edu"
+email: "juancarlos@ucsb.edu"
 abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
@@
     template: template.tex
-header-includes:
-  - \usepackage{float}
-  - \floatplacement{figure}{H}
 ---
@@
 
-\clearpage
-
 ```{r load_spatial}
@@
 
-```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A).
-Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena and Punta Herrero (C) are located in the Yucatan Peninsula.
-Blue polygons represent the TURFs, and red polygons the marine reserves."}
+```{r}
 # Define reference map
@@
 
-plot_grid(left, ph, labels = c("", "C"), ncol = 2)
+plot1 <- plot_grid(left, ph, labels = c("", "C"), ncol = 2)
 ```
@@
 
-```{r}
-tibble(Category = c("Biological", "Biological","Biological","Biological", "Socioeconomic", "Socioeconomic"),
-           Indicator = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
-       Units = c("org $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "Kg $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "M MXP", "Metric Tonnes")) %>% 
-  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
-  kable_styling(latex_options = "HOLD_position")
-```
-
-\clearpage
-
 We use a difference-in-differences analysis to evaluate these indicators.
@@
 Where year-fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
-The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site (See Table \ref{table:com_sum}).
+The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site.
 When evaluating biomass and densities of the invertebrate or fish communities, we include $\sigma_j$ to control for species-fixed effects.
@@
 We based our variable selection primarily on \citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed Mexican fishing cooperatives using this framework.
-We also incorporate other relevant information following \citet{difranco_2016-Xw,edgar_2014-UO} Table \ref{table:ses} shows the selected variables, their definition and selected indicators.
-
-```{r}
-read.csv(here::here("raw_data", "ses.csv")) %>% 
-  select(-Dim) %>% 
-  knitr::kable(caption = "\\label{table:ses}Variables used in our Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).",
-               format = "latex",
-               escape = F) %>% 
-  kable_styling(latex_options = "HOLD_position") %>% 
-  column_spec(column = 1, width = "6cm") %>% 
-  column_spec(column = 2, width = "12cm") %>% 
-  group_rows(group_label = "Resource System", start_row = 1, end_row = 4) %>% 
-  group_rows(group_label = "Resource Unit", start_row = 5, end_row = 6) %>% 
-  group_rows(group_label = "Actors", start_row = 7, end_row = 8) %>% 
-  group_rows(group_label = "Governance system", start_row = 9, end_row = 12)
-```
+We also incorporate other relevant information following \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.
+Table \ref{table:ses} shows the selected variables, their definition and selected indicators.
 
@@
 However, these changes are not significant and are associated to increased variation.
-Full tables with model coefficients are presented in the supplementary materials (S4 Table, S5 Table.
-
-<!-- Revenues show a positive trend post implementation for I Natividad. Because landings have not changed, this might be due to an increase in price? -->
+Full tables with model coefficients are presented in the supplementary materials (S4 Table, S5 Table).
 
@@
 
-Although we have little information on the social dimension of these fisheries, we can use the social-ecological systems framework to analyze the performance of each governance system (Table \ref{tab:ses}).
-Our analysis shows that all communities analyzed share similarities in their Governance system which is based on cooperatives (GS5.2.3.2), with strong rules in use that include Operational rules (GS6.2), Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even Territorial use communal rights (GS6.1.4.3).
-However, we identified important differences in terms of the actors, resource systems, and resource units.
-Although all communities show a high level of leadership (A5), the level of trust (A6.1) is lower in Punta Herrero.
-In general, the presence and success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources they depend upon \citep{jupiter_2017}.
-The enabling conditions for conservation seem to be strongly present in all communities.
-Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperative.
-These conditions have favored the development of efficient community-based enforcement systems.
+Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed communities share similarities known to foster sustainable resource management and increase reserve effectiveness.
+For example, fishers opperate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves.
+Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have solid governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management.
+In general, success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources upon which they depend \citep{jupiter_2017}.
+Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperatives, which have favored the development of efficient community-based enforcement systems.
+However, our SES analysis also highlights factors that might hinder reserve performance or mask outcomes While total reserve size ranges from 0.2% to 3.7% of the TURF area, individual reserves are often small (RS3), and relatively young (RS5).
+Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely that marine reserves will result in increased catches.
 
-\clearpage
-
-```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for lobster densities (*Panulirus spp*; A), fish biomass (B), invertebrate densities (C), and fish densities (D).
-Plots are ordered by survey type (left column: invertebrates; right column: fish).
-Points are jittered hotizontally to avoid overplotting.
-Points indicate the effect size and standard errors.
-Years have been centered to year of implementation."}
+```{r}
 
@@
                             ncol = 2)
-
-plot2
 ```
@@
 
-```{r, fig.width = 5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles).
-Points indicate the effect size and standard errors.
-Years have been centered to year of implementation."}
+```{r}
 
@@
 
-plot_grid(c_plot, v_plot, ncol = 1, labels = "AUTO")
-
-plot3 <- plot_grid(c_plot + theme(legend.position = "none"), v_plot, ncol = 2, labels = "AUTO")
-
+plot3 <- plot_grid(c_plot, v_plot, ncol = 1, labels = "AUTO")
 ```
@@
 The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
-The coastal ecosystems where these reserves are located have been profoundly affected by climatic and oceanographic extremes, including warming events, extreme storms and prolonged hypoxia \citep{micheli_2012-EU}, (**Beas et al.
-in prep., Woodson et al in review**).
+The coastal ecosystems where these reserves are located have been profoundly affected by climatic and oceanographic extremes, including warming events, extreme storms and prolonged hypoxia \citep{micheli_2012-EU,woodson}.
 Despite the lack of evidence of the effectiveness of these reserves, most of the communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}.
@@
 Marine reserves provide protection to a wider range of species and vulnerable habitat, like coral reefs.
-These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as environmental shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (**,Aalto et al, in press**).
+These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as environmental shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU,aalto}.
 Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
 Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}.
-Furthermore, showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
+Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
 Furthermore, the lack of effectiveness observed in these reserves may not be generalizable to other reserves established under the same legal frameowrk (*i.e.* Fish Refges) in Mexico, and future research should aim at evaluating other areas that have also been established as bottom-up processes bt without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), or others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}).
@@
 
+\clearpage
+
 \bibliographystyle{frontiersinSCNS_ENG_HUMS}
@@
 
+\clearpage
+
 \section*{Figure captions}
+
+```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A).
+Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena and Punta Herrero (C) are located in the Yucatan Peninsula.
+Blue polygons represent the TURFs, and red polygons the marine reserves."}
+plot1
+```
+
+```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for lobster densities (*Panulirus spp*; A), fish biomass (B), invertebrate densities (C), and fish densities (D).
+Plots are ordered by survey type (left column: invertebrates; right column: fish).
+Points are jittered hotizontally to avoid overplotting.
+Points indicate the effect size and standard errors.
+Years have been centered to year of implementation."}
+plot2
+```
+
+```{r, fig.width = 5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles).
+Points indicate the effect size and standard errors.
+Years have been centered to year of implementation."}
+plot3
+```
+
+```{r}
+tibble(Category = c("Biological", "Biological","Biological","Biological", "Socioeconomic", "Socioeconomic"),
+           Indicator = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
+       Units = c("org $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "Kg $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "M MXP", "Metric Tonnes")) %>% 
+  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
+  kable_styling(latex_options = "HOLD_position")
+```
+
+```{r}
+read.csv(here::here("raw_data", "ses.csv")) %>% 
+  select(-Dim) %>% 
+  knitr::kable(caption = "\\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).
+Alphanumeric codes follow \\citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \\citet{difranco_2016-Xw} and \\citet{edgar_2014-UO}.",
+               format = "latex",
+               escape = F) %>% 
+  kable_styling(latex_options = c("HOLD_position", "scale_down"),
+                font_size = 11) %>% 
+  column_spec(column = 1, width = "6.5cm") %>% 
+  column_spec(column = 2, width = "12cm") %>% 
+  group_rows(group_label = "Resource System (RS)", start_row = 1, end_row = 5) %>% 
+  group_rows(group_label = "Resource Unit (RU)", start_row = 6, end_row = 6) %>% 
+  group_rows(group_label = "Actors (A)", start_row = 7, end_row = 8) %>% 
+  group_rows(group_label = "Governance system (G)", start_row = 9, end_row = 12)
+```

### 2018-09-24 docs/paper/Villasenor-Derbez.tex
Finalized manuscript. Waiting for Ere's adscription

@@
 
-\def\corrEmail{\href{mailto:jvillasenor@bren.ucsb.edu}{\nolinkurl{jvillasenor@bren.ucsb.edu}}}
+\def\corrEmail{\href{mailto:juancarlos@ucsb.edu}{\nolinkurl{juancarlos@ucsb.edu}}}
 
@@
 
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Location of the three
-coastal communities studied (A). Isla Natividad (B) is located off the
-Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
-in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
-polygons the marine reserves.}
-\end{figure}
-
 \subsection{Data collection}\label{data-collection}
@@
 
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-2}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
-\centering
-\begin{tabular}[t]{l|l|l}
-\hline
-Category & Indicator & Units\\
-\hline
-Biological & Lobster density & org $\mathrm{m}^{-2}$\\
-\hline
-Biological & Invertebrate density & org $\mathrm{m}^{-2}$\\
-\hline
-Biological & Fish biomass & Kg $\mathrm{m}^{-2}$\\
-\hline
-Biological & Fish density & org $\mathrm{m}^{-2}$\\
-\hline
-Socioeconomic & Income from target species & M MXP\\
-\hline
-Socioeconomic & Landings from target species & Metric Tonnes\\
-\hline
-\end{tabular}
-\end{table}
-
-\clearpage
-
 We use a difference-in-differences analysis to evaluate these
@@
 indicator inside the reserve, for year \(t\), with respect to the year
-of implementation in the control site (See Table \ref{table:com_sum}).
-When evaluating biomass and densities of the invertebrate or fish
-communities, we include \(\sigma_j\) to control for species-fixed
-effects.
+of implementation in the control site. When evaluating biomass and
+densities of the invertebrate or fish communities, we include
+\(\sigma_j\) to control for species-fixed effects.
 
@@
 Mexican fishing cooperatives using this framework. We also incorporate
-other relevant information following
-\citet{difranco_2016-Xw,edgar_2014-UO} Table \ref{table:ses} shows the
-selected variables, their definition and selected indicators.
-
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-3}\label{table:ses}Variables used in our Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).}
-\centering
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6cm}|>{\raggedright\arraybackslash}p{12cm}}
-\hline
-Variable & Narrative\\
-\hline
-\multicolumn{2}{l}{\textbf{Resource System}}\\
-\hline
-\hspace{1em}Clarity of system boundaries - Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates.
-Reserve placement is decided by the community, and members trace reserve boundaries themselves.
-Fishers use GPS units and landmarks.\\
-\hline
-Size of resource system - TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
-\hline
-Size of resource system - Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
-\hline
-\hspace{1em}Age of reserves - Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
-\hline
-\multicolumn{2}{l}{\textbf{Resource Unit}}\\
-\hline
-\hspace{1em}Number of units (catch diversity) - Diversity of targeted species & Lobster is their main fishery of these three communities, but they also target finfish.
-Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates (sea cucumber, urchin, snail, abalone).\\
-\hline
-\hspace{1em}Stock status - Status of the main fishery & Lobster stocks are well managed; along with neighboring cooperatives, the lobster fishery is MSC certified.\\
-\hline
-\multicolumn{2}{l}{\textbf{Actors}}\\
-\hline
-\hspace{1em}Number of relevant actors - Number of fishers & IN = 98; ME = 80; PH = 21\\
-\hline
-\hspace{1em}Isolation - Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers.\\
-\hline
-\multicolumn{2}{l}{\textbf{Governance system}}\\
-\hline
-\hspace{1em}Property rights - Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster.
-These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
-\hline
-\hspace{1em}Operational rules - Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandate.
-These include larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.
-They've established penalties for breaking these rules or fishing inside the reserves.\\
-\hline
-\hspace{1em}Social monitoring - Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a designated group that monitors and enforces formal and internal operational rules.
-They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
-\hline
-\hspace{1em}Biophysical monitoring - Monitoring of biological resources, including targeted species & Every year since the creation of the reserves, along with Comunidad y Biodiversidad, fishers perform standardized underwater surveys in the reserves and fishing grounds.
-In more recent years, they have installed oceanographic sensors to monitor for changes in oceanographic variables (e.g.
-Temperature, Dissolved Oxygen, Salinity, pH).\\
-\hline
-\end{tabular}
-\end{table}
+other relevant information following \citet{difranco_2016-Xw} and
+\citet{edgar_2014-UO}. Table \ref{table:ses} shows the selected
+variables, their definition and selected indicators.
 
@@
 increased variation. Full tables with model coefficients are presented
-in the supplementary materials (S4 Table, S5 Table.
+in the supplementary materials (S4 Table, S5 Table).
 
@@
 
-Although we have little information on the social dimension of these
-fisheries, we can use the social-ecological systems framework to analyze
-the performance of each governance system (Table \ref{tab:ses}). Our
-analysis shows that all communities analyzed share similarities in their
-Governance system which is based on cooperatives (GS5.2.3.2), with
-strong rules in use that include Operational rules (GS6.2),
-Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even
-Territorial use communal rights (GS6.1.4.3). However, we identified
-important differences in terms of the actors, resource systems, and
-resource units. Although all communities show a high level of leadership
-(A5), the level of trust (A6.1) is lower in Punta Herrero. In general,
-the presence and success of conservation initiatives depends on the
-incentives of local communities to maintain a healthy status of the
-resources they depend upon \citep{jupiter_2017}. The enabling conditions
-for conservation seem to be strongly present in all communities. Due to
-the clarity of access rights and isolation, the benefits of conservation
-directly benefit the members of the fishing cooperative. These
-conditions have favored the development of efficient community-based
-enforcement systems.
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{\label{fig:unnamed-chunk-4}\label{fig:indicators}Effect sizes for
-marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
-blue triangles), and Punta Herrero (PH; green squares) for lobster
-densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
-densities (C), and fish densities (D). Plots are ordered by survey type
-(left column: invertebrates; right column: fish). Points are jittered
-hotizontally to avoid overplotting. Points indicate the effect size and
-standard errors. Years have been centered to year of implementation.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-7-1.pdf}
-\caption{\label{fig:unnamed-chunk-7}\label{fig:lobsters}Effect sizes for
-lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
-circles) and Maria Elena (ME; blue triangles). Points indicate the
-effect size and standard errors. Years have been centered to year of
-implementation.}
-\end{figure}
+Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed
+communities share similarities known to foster sustainable resource
+management and increase reserve effectiveness. For example, fishers
+opperate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide
+exclusive access to resources and reserves. Along with their relatively
+small groups (A1 - Number of relevant actors), Isolation (A3),
+Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated
+sanctions (GS10.1), these fisheries have solid governance structures
+that enable them to monitor their resources and enforce rules to ensure
+sustainable management. In general, success of conservation initiatives
+depends on the incentives of local communities to maintain a healthy
+status of the resources upon which they depend \citep{jupiter_2017}. Due
+to the clarity of access rights and isolation, the benefits of
+conservation directly benefit the members of the fishing cooperatives,
+which have favored the development of efficient community-based
+enforcement systems. However, our SES analysis also highlights factors
+that might hinder reserve performance or mask outcomes While total
+reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual
+reserves are often small (RS3), and relatively young (RS5).
+Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely
+that marine reserves will result in increased catches.
 
@@
 extremes, including warming events, extreme storms and prolonged hypoxia
-\citep{micheli_2012-EU}, (\textbf{Beas et al. in prep., Woodson et al in
-review}). Despite the lack of evidence of the effectiveness of these
-reserves, most of the communities show a positive perception about their
-performance and continue to support their presence \citep{ayer_2018}.
-Understanding the social-ecological context in which these communities
-and their reserves operate might provide insights to this.
+\citep{micheli_2012-EU,woodson}. Despite the lack of evidence of the
+effectiveness of these reserves, most of the communities show a positive
+perception about their performance and continue to support their
+presence \citep{ayer_2018}. Understanding the social-ecological context
+in which these communities and their reserves operate might provide
+insights to this.
 
@@
 as well as environmental shocks
-\citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (\textbf{,Aalto et al,
-in press}). Self-regulation of fishing effort (\emph{i.e.} reduction in
-harvest) can serve as a way to compensate for future declines associated
-to environmental variation \citep{finkbeiner_2018}. Embarking in a
-marine conservation project can bring the community together, which
-promotes social cohesion and builds social capital \citep{fulton_2019}.
-Furthermore, showing commitment to marine conservation and sustainable
-fishing practices allows fishers to have greater bargaining power and
-leverage over fisheries management \citep{prezramrez_2012}. Furthermore,
-the lack of effectiveness observed in these reserves may not be
-generalizable to other reserves established under the same legal
-frameowrk (\emph{i.e.} Fish Refges) in Mexico, and future research
-should aim at evaluating other areas that have also been established as
-bottom-up processes bt without the presence of TURFs (\emph{e.g.}
-\citet{dof_websiteC_2012}), or others established through a top-down
-process (\emph{i.e.} \citet{dof_websiteU_2018}).
+\citep{hilborn_2004,hilborn_2006,micheli_2012-EU,aalto}. Self-regulation
+of fishing effort (\emph{i.e.} reduction in harvest) can serve as a way
+to compensate for future declines associated to environmental variation
+\citep{finkbeiner_2018}. Embarking in a marine conservation project can
+bring the community together, which promotes social cohesion and builds
+social capital \citep{fulton_2019}. Showing commitment to marine
+conservation and sustainable fishing practices allows fishers to have
+greater bargaining power and leverage over fisheries management
+\citep{prezramrez_2012}. Furthermore, the lack of effectiveness observed
+in these reserves may not be generalizable to other reserves established
+under the same legal frameowrk (\emph{i.e.} Fish Refges) in Mexico, and
+future research should aim at evaluating other areas that have also been
+established as bottom-up processes bt without the presence of TURFs
+(\emph{e.g.} \citet{dof_websiteC_2012}), or others established through a
+top-down process (\emph{i.e.} \citet{dof_websiteU_2018}).
 
@@
 
+\clearpage
+
 \bibliographystyle{frontiersinSCNS_ENG_HUMS}\bibliography{references}
+
+\clearpage
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-7-1.pdf}
+\caption{\label{fig:unnamed-chunk-7}\label{fig:map}Location of the three
+coastal communities studied (A). Isla Natividad (B) is located off the
+Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
+in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
+polygons the marine reserves.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-8-1.pdf}
+\caption{\label{fig:unnamed-chunk-8}\label{fig:indicators}Effect sizes for
+marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
+blue triangles), and Punta Herrero (PH; green squares) for lobster
+densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
+densities (C), and fish densities (D). Plots are ordered by survey type
+(left column: invertebrates; right column: fish). Points are jittered
+hotizontally to avoid overplotting. Points indicate the effect size and
+standard errors. Years have been centered to year of implementation.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-9-1.pdf}
+\caption{\label{fig:unnamed-chunk-9}\label{fig:lobsters}Effect sizes for
+lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
+circles) and Maria Elena (ME; blue triangles). Points indicate the
+effect size and standard errors. Years have been centered to year of
+implementation.}
+\end{figure}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-10}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
+\centering
+\begin{tabular}[t]{l|l|l}
+\hline
+Category & Indicator & Units\\
+\hline
+Biological & Lobster density & org $\mathrm{m}^{-2}$\\
+\hline
+Biological & Invertebrate density & org $\mathrm{m}^{-2}$\\
+\hline
+Biological & Fish biomass & Kg $\mathrm{m}^{-2}$\\
+\hline
+Biological & Fish density & org $\mathrm{m}^{-2}$\\
+\hline
+Socioeconomic & Income from target species & M MXP\\
+\hline
+Socioeconomic & Landings from target species & Metric Tonnes\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-11}\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).
+Alphanumeric codes follow \citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.}
+\centering
+\resizebox{\linewidth}{!}{\fontsize{11}{13}\selectfont
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12cm}}
+\hline
+Variable & Narrative\\
+\hline
+\multicolumn{2}{l}{\textbf{Resource System (RS)}}\\
+\hline
+\hspace{1em}RS2 - Clarity of system boundaries: Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates.
+Reserve placement is decided by the community.
+Fishers use GPS units and landmarks.\\
+\hline
+RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
+\hline
+RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
+\hline
+\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) orhave been (ME, PH) MSC certified.\\
+\hline
+\hspace{1em}*RS5 - Age of reserves: Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
+\hline
+\multicolumn{2}{l}{\textbf{Resource Unit (RU)}}\\
+\hline
+\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish.
+Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates.\\
+\hline
+\multicolumn{2}{l}{\textbf{Actors (A)}}\\
+\hline
+\hspace{1em}A1 - Number of relevant actors: Number of fishers & IN = 98; ME = 80; PH = 21\\
+\hline
+\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers.\\
+\hline
+\multicolumn{2}{l}{\textbf{Governance system (G)}}\\
+\hline
+\hspace{1em}GS6.1.4.3 - Territorial use communal rights : Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster.
+These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
+\hline
+\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandate.
+These include larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
+\hline
+\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group that monitors and enforces formal and internalrules.
+They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
+\hline
+\hspace{1em}GS9.2 - Biophysical monitoring:Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds.
+Recently, they have installed oceanographic sensors to monitor oceanographic variables.\\
+\hline
+GS10.1 - Graduated sanctions & Fishers have penalties for breaking collective-choice rules or fishing inside the reserves.
+These may range from scoldings and warnings to not being allowed to harvest a particular resource or being expelled from the cooperative.\\
+\hline
+\end{tabular}}
+\end{table}
+
 \end{document}

### 2018-09-24 docs/paper/Villasenor-Derbez.Rmd
Updated address, ready for final proof read and submit

@@
 rtitle: "Community-based marine reserves"
-author: "Juan Carlos Villaseñor-Derbez$^{1,2,*}$, Eréndira Aceves-Bueno$^{1}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Arturo Hernández-Velasco$^{2}$, Jorge Torre$^{2}$, Fiorenza Micheli$^{3}$"
+author: "Juan Carlos Villaseñor-Derbez$^{1,2,*}$, Eréndira Aceves-Bueno$^{1,3}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Arturo Hernández-Velasco$^{2}$, Jorge Torre$^{2}$, Fiorenza Micheli$^{4}$"
 institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
 $^{2}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico\\newline
-$^{3}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA"
+$^{3}$Nicholas School of the Environment, Duke University, Beaufort, NC, USA\\newline
+$^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA"
 corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"

### 2018-09-24 docs/paper/Villasenor-Derbez.tex
Updated address, ready for final proof read and submit

@@
 \def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,2,*}\), Eréndira
-Aceves-Bueno\(^{1}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
+Aceves-Bueno\(^{1,3}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
 Arturo Hernández-Velasco\(^{2}\), Jorge Torre\(^{2}\), Fiorenza
-Micheli\(^{3}\)}
+Micheli\(^{4}\)}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Sonora,
-Mexico\newline \(^{3}\)Hopkins Marine Station and Center for Ocean
-Solutions, Stanford University, Pacific Grove, CA, USA}
+Mexico\newline \(^{3}\)Nicholas School of the Environment, Duke
+University, Beaufort, NC, USA\newline \(^{4}\)Hopkins Marine Station and
+Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
 % The Corresponding Author should be marked with an asterisk

### 2018-09-26 docs/paper/Villasenor-Derbez.Rmd
Fixed table 1 where rows were not groupped

@@
 Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves is believed to improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
-In the last six years, and following a 2012 regulation, 18 TURF-reserves have been implemented in Mexico; their effectiveness has not been formally evaluated.
-We combine causal inference techniques and a social-ecological systems framework to provide a holistic evaluation of community-based TURF reserves in three coastal communities in Mexico.
+In the last six years, 18 TURF-reserves have been implemented in Mexico; their effectiveness has not been formally evaluated.
+We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
 We find that while reserves have not yet achieved their stated goal of increasing lobster densities, they continue to receive significant support from the fishing communities.
@@
 In particular, artisanal fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
-Recent research shows that combining Territorial Use Rights for Fisheries (TURFs) with no-take marine reserves (MR) can greatly improve the performance of coastal fisheries and the health of the local resouces \citep{costello_2010-Ix,lester_2017}.
+Recent research shows that combining Territorial Use Rights for Fisheries (TURFs) with no-take marine reserves (MRs) can greatly improve the performance of coastal fisheries and the health of the local resouces \citep{costello_2010-Ix,lester_2017}.
 Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintainance of healthy resources \citep{afflerbach_2014-HP,lester_2017}.
-Although in theory these systems are successful \citep{costello_2010-Ix}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017,smallhornwest_2018}.
+Although in theory these systems are successful \citep{costello_2010-Ix,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017}.
 
@@
 TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities.
-This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
+This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8,smallhornwest_2018}.
 Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
-However, without legal recognition no-take regulations are difficult to enforce and fishers rely on the exclusive access granted by the TURF.
-In an effort to bridge this normative gap, Civil Society Organizations (CSOs) have served as a link between fishers and government, and set out to create a legal framework that solve this governance issue.
-In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuge” (*Zona de Refugio Pesquero*; \citet{nom}).
+However, MRs are difficult to enforce if they are not legally recognized, and fishers rely on the exclusive access granted by the TURF.
+In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue.
+In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuged” (*Zona de Refugio Pesquero*; \citet{nom}).
 Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
-Since 2012, 45 of Fish Refuges have been created along the Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18 of them implemented as TURF-reserves.
+Since 2012, 45 Fish Refuges have been created along the Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18 of them implemented as TURF-reserves.
 However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
 
-Here, we combine causal inference techniques and a social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico.
+Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
 These three case studies span a range of ecological and social conditions representative of different regions of Mexico.
 The objective of this work is twofold.
-First, to provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries.
+First, to provide a triple bottom line evaluation of the effectiveness of community-based marine reserves, which may inform similar processes in other countries.
 Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
-We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world that are using this tool to manage and rebuild their coastal fisheries.
+We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world that use this tool to manage and rebuild their coastal fisheries.
 
@@
 The other two TURF-reserves are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C).
-In contrast with Isla Nativdad, which hosts a well established fishing community, Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); Punta Herrero is home to the *SCPP José María Azcorra* cooperative, and similar to Isla Natividad hosts a local community.
+In contrast with Isla Nativdad, which hosts a well established fishing community, Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the *Cozumel* fishing cooperative; Punta Herrero is home to the *José María Azcorra* fishing cooperative, and similar to Isla Natividad hosts a local community.
 Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season.
-Maria Elena and Punta Herrero established eight marine reserves in 2012, and four marine reserves in 2013, respectively.
+Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively.
 All these reserves have been legally recognized as Fishing Refuges since their creation \citep{dof_website_2012,dof_website_2013}.
@@
 Together, these three communities provide an accurate representation of other fishing communities in each of their regions.
-While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth analysis that we undertake.
+While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth causal inference analysis that we undertake.
 Yet, given the similarities among communities and the socioeconomic and governance setting under which they opperate, it is safe to cautiously generalize our results to other communities in Mexico and other regions around the world.
@@
 location_labels <- data.frame(X = c(-115.24, -87.4611756),
-                              Y = c(27.85, 27.85),
+                              Y = c(27.85, 19),
                               Location = c("Isla Natividad", "Punta Herrero"),
@@
 ph <- ggplot() +
-  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
+  geom_sf(data = turf[3,], fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
   geom_sf(data = coastline_mx) +
@@
   labs(x = "", y = "") +
-  scalebar(data = turf[3,], dist = 5, dd2km = T, model = 'WGS84', location = "bottomright", st.size = 2, st.dist = 0.05, height = 0.1) +
+  scalebar(data = turf[3,],
+           dist = 5,
+           dd2km = T,
+           model = 'WGS84',
+           location = "bottomright",
+           st.size = 2,
+           st.dist = 0.05,
+           height = 0.1) +
   scale_x_continuous(breaks = seq(-87.7, -87.3, 0.1), limits = c(-87.7, -87.3))
@@
 Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017-1m,ayer_2018}.
-Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV} as:
+Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.
 
-\begin{equation}
-I_t = RI_t\times\frac{CPI_t}{CPI_T}
-\label{eqn:cpi}
-\end{equation}
-
-Where $I_t$ represents the adjusted income for year $t$ as the product between the reported income for that year and the ratio between the consumer price index in that year ($CPI_t$) to the most recent year’s consumer price index ($CPI_T$).
-
-Data for the operationalization of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities.
-These include information on the resource system, the resource units, actors, and the governance system itself (Table \ref{table:ses}).
+Data for the evaluation of the SES were collected at the community-level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities.
+These include information on the Resource Systems, Resource Units, Actors, and Governance System (Table \ref{table:ses}).
 
@@
 
-Where year-fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
+Where year-level fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
 The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site.
-When evaluating biomass and densities of the invertebrate or fish communities, we include $\sigma_j$ to control for species-fixed effects.
+When evaluating biomass and densities of the invertebrate or fish communities, we include $\sigma_j$ to control for species-level fixed effects.
 
@@
 All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}.
-All analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.
-Data and code are available on [github.com](https://github.com/jcvdav/ReserveEffect).
+All analyses were performed in `r version$version.string` and R Studio version 1.1.456 \citep{R_2018}.
 
-We use the social-ecologycal system (SES) framework to evaluate each community as a means of providing an explanation to the biological and socioeconomic results.
+We use the SES framework to evaluate each community as a means of providing an explanation to the biological and socioeconomic results.
 The use the SES framework standardizes our analysis and allows us to communicate our results in a common language across fields.
-We based our variable selection primarily on \citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed Mexican fishing cooperatives using this framework.
-We also incorporate other relevant information following \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.
+We based our variable selection primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework.
+We also incorporate other relevant variables known to influence reserve performance following \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.
 Table \ref{table:ses} shows the selected variables, their definition and selected indicators.
@@
 Figure \ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time.
-In the case of Maria Elena, positive changes were observed in the third and forth year.
+In the case of Maria Elena, positive changes were observed in the third and fourth year.
 These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ for Isla Natividad and Punta Herrero, and 0.01 organisms $\mathrm{m}^{-2}$ for Maria Elena, but are not significantly different from zero ($p > 0.05$).
-Likewise, no changes were detected in fish biomass or invertebrate and fish densities (\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
+Likewise, no changes were detected in fish biomass or invertebrate and fish densities (Fig.
+\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
 Full tables with model coefficients are presented in the supplementary materials (S1 Table, S2 Table, S3 Table).
@@
 Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities.
-However, these changes are not significant and are associated to increased variation.
+However, these changes are not significant and are associated with increased variation.
 Full tables with model coefficients are presented in the supplementary materials (S4 Table, S5 Table).
@@
 Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperatives, which have favored the development of efficient community-based enforcement systems.
-However, our SES analysis also highlights factors that might hinder reserve performance or mask outcomes While total reserve size ranges from 0.2% to 3.7% of the TURF area, individual reserves are often small (RS3), and relatively young (RS5).
+However, our SES analysis also highlights factors that might hinder reserve performance or mask outcomes.
+While total reserve size ranges from 0.2% to 3.7% of the TURF area, individual reserves are often small (RS3), and relatively young (RS5).
 Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely that marine reserves will result in increased catches.
@@
 Our results indicate that these TURF-reserves have not increased lobster densities.
-Additionally, no co-benefits were identified when using other ecological indicators other than the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \citep{micheli_2012-EU}.
+Additionally, no co-benefits were identified when using other ecological indicators aside from the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \citep{micheli_2012-EU}.
 The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
@@
 Despite the lack of evidence of the effectiveness of these reserves, most of the communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}.
-Understanding the social-ecological context in which these communities and their reserves operate might provide insights to this.
+Understanding the social-ecological context in which these communities and their reserves operate might provide insights as to why this happens.
 
@@
 For example, we capture previously described patterns like the rapid increase observed for lobster densities in Isla Natividad on the sixth year (*i.e.* 2012; Fig.
-\ref{fig:indicators}A) -a year after the hypoxia events described by \citet{micheli_2012-EU}- which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
+\ref{fig:indicators}A), a year after the hypoxia events described by \citet{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
 Yet, our empirical approach assumes control sites are a plausible counterfactual for treated sites.
@@
 
-Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (*sensu* \citet{rees_2018}.
+Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (*sensu* \citet{rees_2018}).
 Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
@@
 Additionally, \citet{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open acces.
-These increased attributes perhaps minimize the difference between TURF and reserve, making it difficult to detect such a small difference.
+This might reduce the difference between indicators from the TURF and reserve sites, making it difficult to detect such a small change.
 Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of ther eserves might be larger.
@@
 Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
-Furthermore, the lack of effectiveness observed in these reserves may not be generalizable to other reserves established under the same legal frameowrk (*i.e.* Fish Refges) in Mexico, and future research should aim at evaluating other areas that have also been established as bottom-up processes bt without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), or others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}).
+Furthermore, the lack of effectiveness observed in these reserves may not be generalizable to other reserves established under the same legal frameowrk (*i.e.* Fish Refges) in Mexico, and future research should aim at evaluating other areas that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), or others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}).
 
-Previous studies have evaluated the potential of implementing marine reserves in Baja California and connect them to the existing network in California \citep{arafehdalmau_2017}.
 Community-based marine reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented.
@@
 ```{r}
-tibble(Category = c("Biological", "Biological","Biological","Biological", "Socioeconomic", "Socioeconomic"),
-           Indicator = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
-       Units = c("org $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "Kg $\\mathrm{m}^{-2}$", "org $\\mathrm{m}^{-2}$", "M MXP", "Metric Tonnes")) %>% 
-  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
-  kable_styling(latex_options = "HOLD_position")
+data.frame(Indicator = c("Lobster density",
+                         "Invertebrate density",
+                         "Fish biomass",
+                         "Fish density",
+                         "Income from target species",
+                         "Landings from target species"),
+           Units = c("org $\\mathrm{m}^{-2}$",
+                     "org $\\mathrm{m}^{-2}$",
+                     "Kg $\\mathrm{m}^{-2}$",
+                     "org $\\mathrm{m}^{-2}$",
+                     "M MXP",
+                     "Metric Tonnes")) %>% 
+  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.",
+               format = "latex",
+               escape = F) %>% 
+  kable_styling(latex_options = "HOLD_position") %>% 
+  group_rows(group_label = "Biological", start_row = 1, end_row = 4) %>% 
+  group_rows(group_label = "Socioeconomic", start_row = 5, end_row = 6)
 ```
@@
                escape = F) %>% 
-  kable_styling(latex_options = c("HOLD_position", "scale_down"),
-                font_size = 11) %>% 
+  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>% 
   column_spec(column = 1, width = "6.5cm") %>% 

### 2018-09-26 docs/paper/Villasenor-Derbez.tex
Fixed table 1 where rows were not groupped

@@
 fishers reap the benefits of conservation investments. In the last six
-years, and following a 2012 regulation, 18 TURF-reserves have been
-implemented in Mexico; their effectiveness has not been formally
-evaluated. We combine causal inference techniques and a
-social-ecological systems framework to provide a holistic evaluation of
-community-based TURF reserves in three coastal communities in Mexico. We
-find that while reserves have not yet achieved their stated goal of
-increasing lobster densities, they continue to receive significant
-support from the fishing communities. A lack of ecological and
-socioeconomic effects likely results from a combination of factors.
-First, the lobster fisheries are already well managed, and it is
-unlikely that reserves might have a detectable effect. Second, some of
-the reserves are not large enough to protect lobsters' home ranges.
-Third, some of these reserves might be too young for the effects to
-show. However, these reserves have shaped small-scale fishers' way of
-thinking about marine conservation, which can provide a foundation for
-establishing additional, larger marine reserves needed to effectively
-conserve mobile species.
+years, 18 TURF-reserves have been implemented in Mexico; their
+effectiveness has not been formally evaluated. We combine causal
+inference techniques and the Social-Ecological Systems framework to
+provide a holistic evaluation of community-based TURF-reserves in three
+coastal communities in Mexico. We find that while reserves have not yet
+achieved their stated goal of increasing lobster densities, they
+continue to receive significant support from the fishing communities. A
+lack of ecological and socioeconomic effects likely results from a
+combination of factors. First, the lobster fisheries are already well
+managed, and it is unlikely that reserves might have a detectable
+effect. Second, some of the reserves are not large enough to protect
+lobsters' home ranges. Third, some of these reserves might be too young
+for the effects to show. However, these reserves have shaped small-scale
+fishers' way of thinking about marine conservation, which can provide a
+foundation for establishing additional, larger marine reserves needed to
+effectively conserve mobile species.
 
@@
 combining Territorial Use Rights for Fisheries (TURFs) with no-take
-marine reserves (MR) can greatly improve the performance of coastal
+marine reserves (MRs) can greatly improve the performance of coastal
 fisheries and the health of the local resouces
@@
 \citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
-are successful \citep{costello_2010-Ix}, there is little empirical
-evidence of their effectiveness and the drivers of their success
-\citep{afflerbach_2014-HP,lester_2017,smallhornwest_2018}.
+are successful \citep{costello_2010-Ix,smallhornwest_2018}, there is
+little empirical evidence of their effectiveness and the drivers of
+their success \citep{afflerbach_2014-HP,lester_2017}.
 
@@
 voluntarily established and enforced by local communities. This
-bottom-up approach increases compliance and self-enforcement
-\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
+bottom-up approach increases compliance and self-enforcement, and
+reserves can yield benefits similar to systematically-designed reserves
+\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8,smallhornwest_2018}.
 Community-based spatial closures occur in different contexts, like the
 \emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
-\citep{bohnsack_2004,johannes_2002}. However, without legal recognition
-no-take regulations are difficult to enforce and fishers rely on the
+\citep{bohnsack_2004,johannes_2002}. However, MRs are difficult to
+enforce if they are not legally recognized, and fishers rely on the
 exclusive access granted by the TURF. In an effort to bridge this
-normative gap, Civil Society Organizations (CSOs) have served as a link
-between fishers and government, and set out to create a legal framework
-that solve this governance issue. In Mexico, a new norm was created in
-2014 allowing fishers to request the legal recognition of
-community-based reserves as ``Fish Refuge'' (\emph{Zona de Refugio
-Pesquero}; \citet{nom}). Fish refuges can be implemented as temporal or
-partial reserves, which can protect one, some, or all resources within
-their boundaries. Since 2012, 45 of Fish Refuges have been created along
-the Pacific, Gulf of California, and Mexican Caribbean coastlines, with
-18 of them implemented as TURF-reserves. However, their effectiveness
-has not yet been formally evaluated and reported in the scientific
+normative gap, Mexican Civil Society Organizations (CSOs) served as a
+link between fishers and government, and created a legal framework that
+solves this governance issue. In Mexico, a new norm was created in 2014
+allowing fishers to request the legal recognition of community-based
+reserves as ``Fish Refuged'' (\emph{Zona de Refugio Pesquero};
+\citet{nom}). Fish refuges can be implemented as temporal or partial
+reserves, which can protect one, some, or all resources within their
+boundaries. Since 2012, 45 Fish Refuges have been created along the
+Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18
+of them implemented as TURF-reserves. However, their effectiveness has
+not yet been formally evaluated and reported in the scientific
 literature.
 
-Here, we combine causal inference techniques and a social-ecological
-systems framework to provide a holistic evaluation of community-based
-marine reserves in three coastal communities in Mexico. These three case
-studies span a range of ecological and social conditions representative
-of different regions of Mexico. The objective of this work is twofold.
-First, to provide a triple bottom line evaluation of the effectiveness
-of community-based marine reserves that can inform similar processes in
-other countries. Second, to evaluate the effectiveness of TURF-reserves
-established as Fish Refuges in Mexico to identify opportunities where
-improvement or adjustment might lead to increased effectiveness. We draw
-from lessons learned in these three case studies and provide management
-recommendations to maximize the effectiveness of community-based marine
-reserves in small-scale fisheries in Mexico and in other regions around
-the world that are using this tool to manage and rebuild their coastal
-fisheries.
+Here, we combine causal inference techniques and the Social-Ecological
+Systems (SES) framework to provide a holistic evaluation of
+community-based TURF-reserves in three coastal communities in Mexico.
+These three case studies span a range of ecological and social
+conditions representative of different regions of Mexico. The objective
+of this work is twofold. First, to provide a triple bottom line
+evaluation of the effectiveness of community-based marine reserves,
+which may inform similar processes in other countries. Second, to
+evaluate the effectiveness of TURF-reserves established as Fish Refuges
+in Mexico to identify opportunities where improvement or adjustment
+might lead to increased effectiveness. We draw from lessons learned in
+these three case studies and provide management recommendations to
+maximize the effectiveness of community-based marine reserves in
+small-scale fisheries in Mexico and in other regions around the world
+that use this tool to manage and rebuild their coastal fisheries.
 
@@
 Elena is a fishing camp --visited intermittently during the fishing
-season-- belonging to the Cozumel fishing cooperative (\emph{SCPP
-Cozumel}); Punta Herrero is home to the \emph{SCPP José María Azcorra}
-cooperative, and similar to Isla Natividad hosts a local community.
-Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
-argus}), but they also target finfish in the off-season. Maria Elena and
-Punta Herrero established eight marine reserves in 2012, and four marine
-reserves in 2013, respectively. All these reserves have been legally
-recognized as Fishing Refuges since their creation
-\citep{dof_website_2012,dof_website_2013}.
+season-- belonging to the \emph{Cozumel} fishing cooperative; Punta
+Herrero is home to the \emph{José María Azcorra} fishing cooperative,
+and similar to Isla Natividad hosts a local community. Their main
+fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but
+they also target finfish in the off-season. Maria Elena and Punta
+Herrero established eight and four marine reserves in 2012 and 2013,
+respectively. All these reserves have been legally recognized as Fishing
+Refuges since their creation \citep{dof_website_2012,dof_website_2013}.
 
@@
 TURF-reserves, available data would not allow us to perform the in-depth
-analysis that we undertake. Yet, given the similarities among
-communities and the socioeconomic and governance setting under which
-they opperate, it is safe to cautiously generalize our results to other
-communities in Mexico and other regions around the world.
+causal inference analysis that we undertake. Yet, given the similarities
+among communities and the socioeconomic and governance setting under
+which they opperate, it is safe to cautiously generalize our results to
+other communities in Mexico and other regions around the world.
 
@@
 at the cooperative-year level, and revenues were adjusted to represent
-2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}
-as:
+2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.
 
-\begin{equation}
-I_t = RI_t\times\frac{CPI_t}{CPI_T}
-\label{eqn:cpi}
-\end{equation}
-
-Where \(I_t\) represents the adjusted income for year \(t\) as the
-product between the reported income for that year and the ratio between
-the consumer price index in that year (\(CPI_t\)) to the most recent
-year's consumer price index (\(CPI_T\)).
-
-Data for the operationalization of the social-ecological system were
-collected at the community--level from official documents used in the
-creation and designation of the marine reserves
+Data for the evaluation of the SES were collected at the community-level
+from official documents used in the creation and designation of the
+marine reserves
 \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on
 the authors' experience and knowledge of the communities. These include
-information on the resource system, the resource units, actors, and the
-governance system itself (Table \ref{table:ses}).
+information on the Resource Systems, Resource Units, Actors, and
+Governance System (Table \ref{table:ses}).
 
@@
 
-Where year-fixed effects are represented by \(\gamma_{it} Year_t\), and
-\(\beta Zone_i\) captures the difference between reserve (\(Zone = 1\))
-and control (\(Zone = 0\)) sites. The interaction term
-\(\lambda_{it} Year_t\times Zone_i\) represents the mean change in the
-indicator inside the reserve, for year \(t\), with respect to the year
-of implementation in the control site. When evaluating biomass and
-densities of the invertebrate or fish communities, we include
-\(\sigma_j\) to control for species-fixed effects.
+Where year-level fixed effects are represented by
+\(\gamma_{it} Year_t\), and \(\beta Zone_i\) captures the difference
+between reserve (\(Zone = 1\)) and control (\(Zone = 0\)) sites. The
+interaction term \(\lambda_{it} Year_t\times Zone_i\) represents the
+mean change in the indicator inside the reserve, for year \(t\), with
+respect to the year of implementation in the control site. When
+evaluating biomass and densities of the invertebrate or fish
+communities, we include \(\sigma_j\) to control for species-level fixed
+effects.
 
@@
 heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
-analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep{R_2018}.
-Data and code are available on
-\href{https://github.com/jcvdav/ReserveEffect}{github.com}.
+analyses were performed in R version 3.5.1 (2018-07-02) and R Studio
+version 1.1.456 \citep{R_2018}.
 
-We use the social-ecologycal system (SES) framework to evaluate each
-community as a means of providing an explanation to the biological and
-socioeconomic results. The use the SES framework standardizes our
-analysis and allows us to communicate our results in a common language
-across fields. We based our variable selection primarily on
-\citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed
-Mexican fishing cooperatives using this framework. We also incorporate
-other relevant information following \citet{difranco_2016-Xw} and
-\citet{edgar_2014-UO}. Table \ref{table:ses} shows the selected
-variables, their definition and selected indicators.
+We use the SES framework to evaluate each community as a means of
+providing an explanation to the biological and socioeconomic results.
+The use the SES framework standardizes our analysis and allows us to
+communicate our results in a common language across fields. We based our
+variable selection primarily on \citet{leslie_2015-na} and
+\citet{basurto_2013-oq}, who operationalized and analyzed Mexican
+fishing cooperatives using this framework. We also incorporate other
+relevant variables known to influence reserve performance following
+\citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table
+\ref{table:ses} shows the selected variables, their definition and
+selected indicators.
 
@@
 but the effect is eroded through time. In the case of Maria Elena,
-positive changes were observed in the third and forth year. These
+positive changes were observed in the third and fourth year. These
 effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
@@
 (\(p > 0.05\)). Likewise, no changes were detected in fish biomass or
-invertebrate and fish densities (\ref{fig:indicators}B-D), where effect
-sizes oscillated around zero without clear trends. Full tables with
-model coefficients are presented in the supplementary materials (S1
+invertebrate and fish densities (Fig. \ref{fig:indicators}B-D), where
+effect sizes oscillated around zero without clear trends. Full tables
+with model coefficients are presented in the supplementary materials (S1
 Table, S2 Table, S3 Table).
@@
 implementation of the reserve, as compared to the control communities.
-However, these changes are not significant and are associated to
+However, these changes are not significant and are associated with
 increased variation. Full tables with model coefficients are presented
@@
 enforcement systems. However, our SES analysis also highlights factors
-that might hinder reserve performance or mask outcomes While total
+that might hinder reserve performance or mask outcomes. While total
 reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual
@@
 densities. Additionally, no co-benefits were identified when using other
-ecological indicators other than the previously reported buffering
+ecological indicators aside from the previously reported buffering
 effect that reserves can have to environmental variability in Isla
@@
 in which these communities and their reserves operate might provide
-insights to this.
+insights as to why this happens.
 
@@
 increase observed for lobster densities in Isla Natividad on the sixth
-year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A) -a year after the
-hypoxia events described by \citet{micheli_2012-EU}- which caused mass
+year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), a year after the
+hypoxia events described by \citet{micheli_2012-EU}, which caused mass
 mortality of sedentary organisms such as abalone and sea urchins, but
@@
 attributed to poor ecological coherence in reserve design (\emph{sensu}
-\citet{rees_2018}. Previous research has shown that reserves in Isla
+\citet{rees_2018}). Previous research has shown that reserves in Isla
 Natividad yield fishery benefits for the abalone fishery
@@
 \citet{gelcich_2008} have shown that TURFs alone can have greater
-biomass and richness than areas operating under open acces. These
-increased attributes perhaps minimize the difference between TURF and
-reserve, making it difficult to detect such a small difference. Further
+biomass and richness than areas operating under open acces. This might
+reduce the difference between indicators from the TURF and reserve
+sites, making it difficult to detect such a small change. Further
 research should focus on evaluating sites in the reserve, TURF, and open
@@
 future research should aim at evaluating other areas that have also been
-established as bottom-up processes bt without the presence of TURFs
+established as bottom-up processes but without the presence of TURFs
 (\emph{e.g.} \citet{dof_websiteC_2012}), or others established through a
@@
 
-Previous studies have evaluated the potential of implementing marine
-reserves in Baja California and connect them to the existing network in
-California \citep{arafehdalmau_2017}. Community-based marine reserves in
-small-scale fisheries can be helpful conservation and fishery management
-tools when appropriately implemented. Lessons learned from these cases
-can guide implementation of community-based marine reserves elsewhere.
-For the particular case of the marine reserves that we evaluate, the
-possibility of expanding reserves or merging existing polygons into
-larger areas should be evaluated and proposed to the communities. At the
-broader scale, having full community support surely represents an
-advantage, but it is important that marine reserves meet essential
-design principles such as size and placement. Community-based marine
-reserves might have more benefits that result from indirect effects of
-the reserves, which should be taken into account when evaluating the
-outcomes of similar projects.
+Community-based marine reserves in small-scale fisheries can be helpful
+conservation and fishery management tools when appropriately
+implemented. Lessons learned from these cases can guide implementation
+of community-based marine reserves elsewhere. For the particular case of
+the marine reserves that we evaluate, the possibility of expanding
+reserves or merging existing polygons into larger areas should be
+evaluated and proposed to the communities. At the broader scale, having
+full community support surely represents an advantage, but it is
+important that marine reserves meet essential design principles such as
+size and placement. Community-based marine reserves might have more
+benefits that result from indirect effects of the reserves, which should
+be taken into account when evaluating the outcomes of similar projects.
 
@@
 \centering
-\begin{tabular}[t]{l|l|l}
+\begin{tabular}[t]{l|l}
 \hline
-Category & Indicator & Units\\
+Indicator & Units\\
 \hline
-Biological & Lobster density & org $\mathrm{m}^{-2}$\\
+\multicolumn{2}{l}{\textbf{Biological}}\\
 \hline
-Biological & Invertebrate density & org $\mathrm{m}^{-2}$\\
+\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
 \hline
-Biological & Fish biomass & Kg $\mathrm{m}^{-2}$\\
+\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
 \hline
-Biological & Fish density & org $\mathrm{m}^{-2}$\\
+\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
 \hline
-Socioeconomic & Income from target species & M MXP\\
+\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
 \hline
-Socioeconomic & Landings from target species & Metric Tonnes\\
+\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
+\hline
+\hspace{1em}Income from target species & M MXP\\
+\hline
+\hspace{1em}Landings from target species & Metric Tonnes\\
 \hline
@@
 \centering
-\resizebox{\linewidth}{!}{\fontsize{11}{13}\selectfont
+\resizebox{\linewidth}{!}{
 \begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12cm}}
@@
 \hline
-RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
+\hspace{1em}RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
 \hline
-RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
+\hspace{1em}RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
 \hline
@@
 \hline
-\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandate.
+\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandates.
 These include larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
 \hline
-\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group that monitors and enforces formal and internalrules.
+\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group that monitors and enforces formal and internal rules.
 They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\

### 2018-10-01 docs/paper/Villasenor-Derbez.Rmd
Manuscript ready for submission

@@
 rtitle: "Community-based marine reserves"
-author: "Juan Carlos Villaseñor-Derbez$^{1,2,*}$, Eréndira Aceves-Bueno$^{1,3}$, Álvin Suarez$^{2}$, Stuart Fulton$^{2}$, Arturo Hernández-Velasco$^{2}$, Jorge Torre$^{2}$, Fiorenza Micheli$^{4}$"
+author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,2}$, Stuart Fulton$^{3}$, Álvin Suarez$^{3}$, Arturo Hernández-Velasco$^{3}$, Jorge Torre$^{3}$, Fiorenza Micheli$^{4}$"
 institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
-$^{2}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico\\newline
-$^{3}$Nicholas School of the Environment, Duke University, Beaufort, NC, USA\\newline
+$^{2}$Nicholas School of the Environment, Duke University, Beaufort, NC, USA\\newline
+$^{3}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico\\newline
 $^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA"
@@
 
+Words in text: 3821 | Words in headers: 35 | Words outside text (captions, etc.): 192
+
 \clearpage
@@
 Recent research shows that combining Territorial Use Rights for Fisheries (TURFs) with no-take marine reserves (MRs) can greatly improve the performance of coastal fisheries and the health of the local resouces \citep{costello_2010-Ix,lester_2017}.
-Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintainance of healthy resources \citep{afflerbach_2014-HP,lester_2017}.
+Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014-HP,lester_2017}.
 Although in theory these systems are successful \citep{costello_2010-Ix,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017}.
@@
 In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue.
-In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuged” (*Zona de Refugio Pesquero*; \citet{nom}).
+In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}).
 Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
@@
 Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
-We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world that use this tool to manage and rebuild their coastal fisheries.
+We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world where this tool is used to manage and rebuild their coastal fisheries.
 
@@
 Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively.
-All these reserves have been legally recognized as Fishing Refuges since their creation \citep{dof_website_2012,dof_website_2013}.
+These reserves have been legally recognized as Fishing Refuges since their creation \citep{dof_website_2012,dof_website_2013}.
 
@@
 While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth causal inference analysis that we undertake.
-Yet, given the similarities among communities and the socioeconomic and governance setting under which they opperate, it is safe to cautiously generalize our results to other communities in Mexico and other regions around the world.
+Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our results to other communities in Mexico and other regions around the world.
 
@@
 Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
-Trained divers record richness and abundances of fish and invertebrate species along replicate transects (30x 2 m each) at depths 5-20 m in the reserves and control sites \citep{fulton_2018,fulton_2019}.
+Trained divers record richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \citep{fulton_2018,fulton_2019,suman_2010-ez}.
 Size structures are also collected during fish surveys.
@@
 Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed communities share similarities known to foster sustainable resource management and increase reserve effectiveness.
-For example, fishers opperate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves.
+For example, fishers operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves.
 Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have solid governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management.
@@
 
-JC and EA analyzed and interpreted data, discussed the results, and wrote the first draft.
-AHV coordinated fieldwork and collected the data.
-AS, AHV, SF, JT, and FM discussed the results and edited the manuscript.
+JC and AS conceived the idea.
+JC and EA analyzed data, discussed the results, and wrote the first draft.
+FM, SF, AS, JT, and AHV discussed the results and edited the manuscript.
+All authors provided valuable contributions.
 

### 2018-10-01 docs/paper/Villasenor-Derbez.tex
Manuscript ready for submission

@@
 \def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
-\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,2,*}\), Eréndira
-Aceves-Bueno\(^{1,3}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
-Arturo Hernández-Velasco\(^{2}\), Jorge Torre\(^{2}\), Fiorenza
+\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
+Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
+Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
 Micheli\(^{4}\)}
@@
 of California, Santa Barbara, Santa Barbara, CA,
-USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Sonora,
-Mexico\newline \(^{3}\)Nicholas School of the Environment, Duke
-University, Beaufort, NC, USA\newline \(^{4}\)Hopkins Marine Station and
+USA\newline \(^{2}\)Nicholas School of the Environment, Duke University,
+Beaufort, NC, USA\newline \(^{3}\)Comunidad y Biodiversidad A.C.,
+Guaymas, Sonora, Mexico\newline \(^{4}\)Hopkins Marine Station and
 Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
@@
 
+Words in text: 3821 \textbar{} Words in headers: 35 \textbar{} Words
+outside text (captions, etc.): 192
+
 \clearpage
@@
 these systems increase the benefits of spatial access rights allowing
-the maintainance of healthy resources
+the maintenance of healthy resources
 \citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
@@
 allowing fishers to request the legal recognition of community-based
-reserves as ``Fish Refuged'' (\emph{Zona de Refugio Pesquero};
+reserves as ``Fish Refuges'' (\emph{Zona de Refugio Pesquero};
 \citet{nom}). Fish refuges can be implemented as temporal or partial
@@
 small-scale fisheries in Mexico and in other regions around the world
-that use this tool to manage and rebuild their coastal fisheries.
+where this tool is used to manage and rebuild their coastal fisheries.
 
@@
 Herrero established eight and four marine reserves in 2012 and 2013,
-respectively. All these reserves have been legally recognized as Fishing
+respectively. These reserves have been legally recognized as Fishing
 Refuges since their creation \citep{dof_website_2012,dof_website_2013}.
@@
 among communities and the socioeconomic and governance setting under
-which they opperate, it is safe to cautiously generalize our results to
+which they operate, it is safe to cautiously generalize our results to
 other communities in Mexico and other regions around the world.
@@
 abundances of fish and invertebrate species along replicate transects
-(30x 2 m each) at depths 5-20 m in the reserves and control sites
-\citep{fulton_2018,fulton_2019}. Size structures are also collected
-during fish surveys. We define control sites as regions with habitat
-characteristics similar to the corresponding reserves, and that
-presumably had a similar probability of being selected as reserves
+(30 \(\times\) 2 m each) at depths 5-20 m in the reserves and control
+sites \citep{fulton_2018,fulton_2019,suman_2010-ez}. Size structures are
+also collected during fish surveys. We define control sites as regions
+with habitat characteristics similar to the corresponding reserves, and
+that presumably had a similar probability of being selected as reserves
 during the design phase. We focus our evaluation on sites where data are
@@
 management and increase reserve effectiveness. For example, fishers
-opperate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide
+operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide
 exclusive access to resources and reserves. Along with their relatively
@@
 
-JC and EA analyzed and interpreted data, discussed the results, and
-wrote the first draft. AHV coordinated fieldwork and collected the data.
-AS, AHV, SF, JT, and FM discussed the results and edited the manuscript.
+JC and AS conceived the idea. JC and EA analyzed data, discussed the
+results, and wrote the first draft. FM, SF, AS, JT, and AHV discussed
+the results and edited the manuscript. All authors provided valuable
+contributions.
 
@@
 \hline
-\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) orhave been (ME, PH) MSC certified.\\
+\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) MSC certified.\\
 \hline
@@
 \hline
-\hspace{1em}GS9.2 - Biophysical monitoring:Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds.
+\hspace{1em}GS9.2 - Biophysical monitoring: Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds.
 Recently, they have installed oceanographic sensors to monitor oceanographic variables.\\

### 2018-10-01 docs/paper/Villasenor-Derbez.Rmd
Incorporated Fios comments, ready for submission?

@@
 abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
-Artisanal fisheries face great challenges since they are difficult to monitor, enforce, and manage.
-Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves is believed to improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
-In the last six years, 18 TURF-reserves have been implemented in Mexico; their effectiveness has not been formally evaluated.
+Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage.
+Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
+In the last six years, 18 TURF-reserves have been implemented in Mexico following a new regulation; their effectiveness has not been formally evaluated.
 We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
-We find that while reserves have not yet achieved their stated goal of increasing lobster densities, they continue to receive significant support from the fishing communities.
-A lack of ecological and socioeconomic effects likely results from a combination of factors.
-First, the lobster fisheries are already well managed, and it is unlikely that reserves might have a detectable effect.
-Second, some of the reserves are not large enough to protect lobsters' home ranges.
+We find that while reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates, they continue to receive significant support from the fishing communities.
+A lack of clear ecological and socioeconomic effects likely results from a combination of factors.
+First, lobster and other invertebrate fisheries are already well managed, and it is unlikely that reserves might have a detectable effect.
+Second, some of the reserves are not large enough to protect mobile species, like lobster.
 Third, some of these reserves might be too young for the effects to show.
+Fourth, variable and extreme oceanographic conditions have impacted harvested populations.
 However, these reserves have shaped small-scale fishers’ way of thinking about marine conservation, which can provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
@@
 
-Words in text: 3821 | Words in headers: 35 | Words outside text (captions, etc.): 192
+Words in text: 3718 | Words in headers: 35 | Words outside text (captions, etc.): 192
 
@@
 Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}.
-In particular, artisanal fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
-Recent research shows that combining Territorial Use Rights for Fisheries (TURFs) with no-take marine reserves (MRs) can greatly improve the performance of coastal fisheries and the health of the local resouces \citep{costello_2010-Ix,lester_2017}.
+In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
+Recent research shows that combining Territorial Use Rights for Fisheries (TURFs) with no-take marine reserves (MRs) can greatly improve the performance of coastal fisheries and the health of the local resources \citep{costello_2010-Ix,lester_2017}.
 Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014-HP,lester_2017}.
@@
 However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.
-This is especially important in social-ecological coastal systems dominated by close interaction and feedbacks between people and natural resources \citep{ostrom_2009-hg}.
+This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}.
 
@@
 Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
-Since 2012, 45 Fish Refuges have been created along the Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18 of them implemented as TURF-reserves.
+Since 2006, 45 Fish Refuges have been created along the Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18 of them implemented as TURF-reserves legally recognized as "Fish Refuges".
 However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
@@
 
-We use the SES framework to evaluate each community as a means of providing an explanation to the biological and socioeconomic results.
-The use the SES framework standardizes our analysis and allows us to communicate our results in a common language across fields.
+We use the SES framework to evaluate each community.
+The use the SES framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators.
 We based our variable selection primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework.
@@
 These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ for Isla Natividad and Punta Herrero, and 0.01 organisms $\mathrm{m}^{-2}$ for Maria Elena, but are not significantly different from zero ($p > 0.05$).
-Likewise, no changes were detected in fish biomass or invertebrate and fish densities (Fig.
+Likewise, no significant changes were detected in fish biomass or invertebrate and fish densities (Fig.
 \ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
@@
 The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
-The coastal ecosystems where these reserves are located have been profoundly affected by climatic and oceanographic extremes, including warming events, extreme storms and prolonged hypoxia \citep{micheli_2012-EU,woodson}.
 Despite the lack of evidence of the effectiveness of these reserves, most of the communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}.
@@
 
-A first possible explanation for the lack of effectivenes may be the young age of the reserves.
+A first possible explanation for the lack of effectiveness may be the young age of the reserves.
 Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO,babcock_2010}.
@@
 Furthermore, fishers may favor implementation of reserves that pose low fishing costs due to their small size or location.
-Our analysis of economic data supports this, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
+Our analysis of economic data supports this hypothesis, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
 
 Even if reserves had appropriate sizes and were placed in optimal locations, there are other plausible explanations for the observed patterns.
-For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2006}.
+For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2004,hilborn_2006}.
 Both lobster fisheries were certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
@@
 It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves.
-Additionally, \citet{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open acces.
+Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open access.
 This might reduce the difference between indicators from the TURF and reserve sites, making it difficult to detect such a small change.
-Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of ther eserves might be larger.
+Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of the reserves might be larger.
+
+Finally, extreme conditions, including prolonged hypoxia, heat waves, and storms have affected both the Pacific and Caribbean regions, with large negative impacts of coastal marine species and ecosystems \citep{cavole_2016,hughes_2018,breitburg_2018}.
+The coastal ecosystems where these reserves are located have been profoundly affected by these events \citep{micheli_2012-EU,woodson}.
+Effects of protection might be eliminated by the mortalities associated with these extreme conditions.
 
 While the evaluated reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits.
-Marine reserves provide protection to a wider range of species and vulnerable habitat, like coral reefs.
-These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as environmental shocks \citep{hilborn_2004,hilborn_2006,micheli_2012-EU,aalto}.
+Marine reserves provide protection to a wider range of species and vulnerable habitat.
+These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as mild environmental shocks \citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
 Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
@@
 Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
-Furthermore, the lack of effectiveness observed in these reserves may not be generalizable to other reserves established under the same legal frameowrk (*i.e.* Fish Refges) in Mexico, and future research should aim at evaluating other areas that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), or others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}).
+Furthermore, the lack of effectiveness observed in these reserves should not be generalizable to other reserves established under the same legal framework (*i.e.* Fish Refuges) in Mexico, and future research should aim at evaluating other areas that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), or others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}).
 
@@
 At the broader scale, having full community support surely represents an advantage, but it is important that marine reserves meet essential design principles such as size and placement.
-Community-based marine reserves might have more benefits that result from indirect effects of the reserves, which should be taken into account when evaluating the outcomes of similar projects.
+Community-based marine reserves might have more benefits that result from indirect effects of the reserves, particularly providing resilience to shocks and management errors, and promoting social cohesion, which should be taken into account when evaluating the outcomes of TURF-reserves.
 
@@
 
-JCVD recevied funding from CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the Latin American Fisheries Fellowship Program.
+JCVD received funding from UCMexus - CONACyT Doctoral Fellowship (CVU 669403) and the Latin American Fisheries Fellowship Program.
 AS, AHV, SF and JT received funding from Marisla Foundation, Packard Foundation, Walton Family Foundation, Summit Foundation, and Oak Foundation.
+FM was supported by NSF-CNH and NSF BioOce (grants DEB-1212124 and 1736830).
 
@@
                          "Invertebrate density",
+                         "Fish density",
                          "Fish biomass",
-                         "Fish density",
                          "Income from target species",
@@
                      "org $\\mathrm{m}^{-2}$",
+                     "org $\\mathrm{m}^{-2}$",
                      "Kg $\\mathrm{m}^{-2}$",
-                     "org $\\mathrm{m}^{-2}$",
                      "M MXP",

### 2018-10-01 docs/paper/Villasenor-Derbez.tex
Incorporated Fios comments, ready for submission?

@@
 Coastal marine ecosystems provide livelihoods for small-scale fishers
-and coastal communities around the world. Artisanal fisheries face great
-challenges since they are difficult to monitor, enforce, and manage.
-Combining territorial user rights for fisheries (TURF) with no-take
-marine reserves to create TURF-reserves is believed to improve the
+and coastal communities around the world. Small-scale fisheries face
+great challenges since they are difficult to monitor, enforce, and
+manage. Combining territorial user rights for fisheries (TURF) with
+no-take marine reserves to create TURF-reserves can improve the
 performance of small-scale fisheries by buffering fisheries from
@@
 fishers reap the benefits of conservation investments. In the last six
-years, 18 TURF-reserves have been implemented in Mexico; their
-effectiveness has not been formally evaluated. We combine causal
-inference techniques and the Social-Ecological Systems framework to
-provide a holistic evaluation of community-based TURF-reserves in three
-coastal communities in Mexico. We find that while reserves have not yet
-achieved their stated goal of increasing lobster densities, they
-continue to receive significant support from the fishing communities. A
-lack of ecological and socioeconomic effects likely results from a
-combination of factors. First, the lobster fisheries are already well
-managed, and it is unlikely that reserves might have a detectable
-effect. Second, some of the reserves are not large enough to protect
-lobsters' home ranges. Third, some of these reserves might be too young
-for the effects to show. However, these reserves have shaped small-scale
-fishers' way of thinking about marine conservation, which can provide a
-foundation for establishing additional, larger marine reserves needed to
-effectively conserve mobile species.
+years, 18 TURF-reserves have been implemented in Mexico following a new
+regulation; their effectiveness has not been formally evaluated. We
+combine causal inference techniques and the Social-Ecological Systems
+framework to provide a holistic evaluation of community-based
+TURF-reserves in three coastal communities in Mexico. We find that while
+reserves have not yet achieved their stated goal of increasing the
+density of lobster and other benthic invertebrates, they continue to
+receive significant support from the fishing communities. A lack of
+clear ecological and socioeconomic effects likely results from a
+combination of factors. First, lobster and other invertebrate fisheries
+are already well managed, and it is unlikely that reserves might have a
+detectable effect. Second, some of the reserves are not large enough to
+protect mobile species, like lobster. Third, some of these reserves
+might be too young for the effects to show. Fourth, variable and extreme
+oceanographic conditions have impacted harvested populations. However,
+these reserves have shaped small-scale fishers' way of thinking about
+marine conservation, which can provide a foundation for establishing
+additional, larger marine reserves needed to effectively conserve mobile
+species.
 
@@
 
-Words in text: 3821 \textbar{} Words in headers: 35 \textbar{} Words
+Words in text: 3718 \textbar{} Words in headers: 35 \textbar{} Words
 outside text (captions, etc.): 192
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
-artisanal fisheries face great challenges since they tend to be hard to
-monitor and enforce \citep{costello_2012}. Recent research shows that
+small-scale fisheries face great challenges since they tend to be hard
+to monitor and enforce \citep{costello_2012}. Recent research shows that
 combining Territorial Use Rights for Fisheries (TURFs) with no-take
 marine reserves (MRs) can greatly improve the performance of coastal
-fisheries and the health of the local resouces
+fisheries and the health of the local resources
 \citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
@@
 social and ecological perspective. This is especially important in
-social-ecological coastal systems dominated by close interaction and
+social-ecological coastal systems dominated by close interactions and
 feedbacks between people and natural resources \citep{ostrom_2009-hg}.
@@
 reserves, which can protect one, some, or all resources within their
-boundaries. Since 2012, 45 Fish Refuges have been created along the
+boundaries. Since 2006, 45 Fish Refuges have been created along the
 Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18
-of them implemented as TURF-reserves. However, their effectiveness has
-not yet been formally evaluated and reported in the scientific
-literature.
+of them implemented as TURF-reserves legally recognized as ``Fish
+Refuges''. However, their effectiveness has not yet been formally
+evaluated and reported in the scientific literature.
 
@@
 
-\section{Methods}\label{methods}
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
-\subsection{Study area}\label{study-area}
+\hypertarget{study-area}{%
+\subsection{Study area}\label{study-area}}
 
@@
 
-\subsection{Data collection}\label{data-collection}
+\hypertarget{data-collection}{%
+\subsection{Data collection}\label{data-collection}}
 
@@
 
-\subsection{Data analysis}\label{data-analysis}
+\hypertarget{data-analysis}{%
+\subsection{Data analysis}\label{data-analysis}}
 
@@
 
-We use the SES framework to evaluate each community as a means of
-providing an explanation to the biological and socioeconomic results.
-The use the SES framework standardizes our analysis and allows us to
-communicate our results in a common language across fields. We based our
-variable selection primarily on \citet{leslie_2015-na} and
-\citet{basurto_2013-oq}, who operationalized and analyzed Mexican
-fishing cooperatives using this framework. We also incorporate other
-relevant variables known to influence reserve performance following
-\citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table
-\ref{table:ses} shows the selected variables, their definition and
-selected indicators.
+We use the SES framework to evaluate each community. The use the SES
+framework standardizes our analysis and allows us to communicate our
+results in a common language across fields by using a set of previously
+defined variables and indicators. We based our variable selection
+primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who
+operationalized and analyzed Mexican fishing cooperatives using this
+framework. We also incorporate other relevant variables known to
+influence reserve performance following \citet{difranco_2016-Xw} and
+\citet{edgar_2014-UO}. Table \ref{table:ses} shows the selected
+variables, their definition and selected indicators.
 
-\section{Results}\label{results}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
@@
 
-\subsection{Biological effects}\label{biological-effects}
+\hypertarget{biological-effects}{%
+\subsection{Biological effects}\label{biological-effects}}
 
@@
 for Maria Elena, but are not significantly different from zero
-(\(p > 0.05\)). Likewise, no changes were detected in fish biomass or
-invertebrate and fish densities (Fig. \ref{fig:indicators}B-D), where
-effect sizes oscillated around zero without clear trends. Full tables
-with model coefficients are presented in the supplementary materials (S1
-Table, S2 Table, S3 Table).
+(\(p > 0.05\)). Likewise, no significant changes were detected in fish
+biomass or invertebrate and fish densities (Fig.
+\ref{fig:indicators}B-D), where effect sizes oscillated around zero
+without clear trends. Full tables with model coefficients are presented
+in the supplementary materials (S1 Table, S2 Table, S3 Table).
 
-\subsection{Socioeconomic effects}\label{socioeconomic-effects}
+\hypertarget{socioeconomic-effects}{%
+\subsection{Socioeconomic effects}\label{socioeconomic-effects}}
 
@@
 
-\subsection{Governance}\label{governance}
+\hypertarget{governance}{%
+\subsection{Governance}\label{governance}}
 
@@
 
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 pertaining landings and revenues showed little to no change after
-reserve implementation. The coastal ecosystems where these reserves are
-located have been profoundly affected by climatic and oceanographic
-extremes, including warming events, extreme storms and prolonged hypoxia
-\citep{micheli_2012-EU,woodson}. Despite the lack of evidence of the
+reserve implementation. Despite the lack of evidence of the
 effectiveness of these reserves, most of the communities show a positive
@@
 
-A first possible explanation for the lack of effectivenes may be the
+A first possible explanation for the lack of effectiveness may be the
 young age of the reserves. Literature shows that age and enforcement are
@@
 their small size or location. Our analysis of economic data supports
-this, as neither landings nor revenues showed the expected short-term
-costs associated to the first years of reserve implementation
+this hypothesis, as neither landings nor revenues showed the expected
+short-term costs associated to the first years of reserve implementation
 \citep{ovando_2016-Wg}.
@@
 fisheries benefits if initial population sizes are low and the fishery
-is poorly managed \citep{hilborn_2006}. Both lobster fisheries were
-certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
-Additionally, lobster fisheries are managed via species-specific minimum
-catch sizes, seasonal closures, protection of ``berried'' females, and
-escapement windows where traps are allowed \citep{dof_website_1993}. It
-is uncertain whether such a well-managed fishery will experience
-additional benefits from marine reserves. Additionally,
-\citet{gelcich_2008} have shown that TURFs alone can have greater
-biomass and richness than areas operating under open acces. This might
-reduce the difference between indicators from the TURF and reserve
-sites, making it difficult to detect such a small change. Further
-research should focus on evaluating sites in the reserve, TURF, and open
-access areas or similar Fish Refuges established without the presence of
-TURFs where the impact of ther eserves might be larger.
+is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
+fisheries were certified by the Marine Stewardship Council
+\citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed
+via species-specific minimum catch sizes, seasonal closures, protection
+of ``berried'' females, and escapement windows where traps are allowed
+\citep{dof_website_1993}. It is uncertain whether such a well-managed
+fishery will experience additional benefits from marine reserves.
+Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have
+greater biomass and richness than areas operating under open access.
+This might reduce the difference between indicators from the TURF and
+reserve sites, making it difficult to detect such a small change.
+Further research should focus on evaluating sites in the reserve, TURF,
+and open access areas or similar Fish Refuges established without the
+presence of TURFs where the impact of the reserves might be larger.
+
+Finally, extreme conditions, including prolonged hypoxia, heat waves,
+and storms have affected both the Pacific and Caribbean regions, with
+large negative impacts of coastal marine species and ecosystems
+\citep{cavole_2016,hughes_2018,breitburg_2018}. The coastal ecosystems
+where these reserves are located have been profoundly affected by these
+events \citep{micheli_2012-EU,woodson}. Effects of protection might be
+eliminated by the mortalities associated with these extreme conditions.
 
@@
 social benefits. Marine reserves provide protection to a wider range of
-species and vulnerable habitat, like coral reefs. These sites can serve
-as an insurance against uncertainty and errors in fisheries management,
-as well as environmental shocks
-\citep{hilborn_2004,hilborn_2006,micheli_2012-EU,aalto}. Self-regulation
-of fishing effort (\emph{i.e.} reduction in harvest) can serve as a way
-to compensate for future declines associated to environmental variation
-\citep{finkbeiner_2018}. Embarking in a marine conservation project can
-bring the community together, which promotes social cohesion and builds
-social capital \citep{fulton_2019}. Showing commitment to marine
-conservation and sustainable fishing practices allows fishers to have
-greater bargaining power and leverage over fisheries management
-\citep{prezramrez_2012}. Furthermore, the lack of effectiveness observed
-in these reserves may not be generalizable to other reserves established
-under the same legal frameowrk (\emph{i.e.} Fish Refges) in Mexico, and
-future research should aim at evaluating other areas that have also been
-established as bottom-up processes but without the presence of TURFs
-(\emph{e.g.} \citet{dof_websiteC_2012}), or others established through a
-top-down process (\emph{i.e.} \citet{dof_websiteU_2018}).
+species and vulnerable habitat. These sites can serve as an insurance
+against uncertainty and errors in fisheries management, as well as mild
+environmental shocks
+\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
+Self-regulation of fishing effort (\emph{i.e.} reduction in harvest) can
+serve as a way to compensate for future declines associated to
+environmental variation \citep{finkbeiner_2018}. Embarking in a marine
+conservation project can bring the community together, which promotes
+social cohesion and builds social capital \citep{fulton_2019}. Showing
+commitment to marine conservation and sustainable fishing practices
+allows fishers to have greater bargaining power and leverage over
+fisheries management \citep{prezramrez_2012}. Furthermore, the lack of
+effectiveness observed in these reserves should not be generalizable to
+other reserves established under the same legal framework (\emph{i.e.}
+Fish Refuges) in Mexico, and future research should aim at evaluating
+other areas that have also been established as bottom-up processes but
+without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
+or others established through a top-down process (\emph{i.e.}
+\citet{dof_websiteU_2018}).
 
@@
 size and placement. Community-based marine reserves might have more
-benefits that result from indirect effects of the reserves, which should
-be taken into account when evaluating the outcomes of similar projects.
+benefits that result from indirect effects of the reserves, particularly
+providing resilience to shocks and management errors, and promoting
+social cohesion, which should be taken into account when evaluating the
+outcomes of TURF-reserves.
 
@@
 
-JCVD recevied funding from CONACyT (Beca de Posgrados en el extranjero,
-CVU 669403) and the Latin American Fisheries Fellowship Program. AS,
-AHV, SF and JT received funding from Marisla Foundation, Packard
-Foundation, Walton Family Foundation, Summit Foundation, and Oak
-Foundation.
+JCVD received funding from UCMexus - CONACyT Doctoral Fellowship (CVU
+669403) and the Latin American Fisheries Fellowship Program. AS, AHV, SF
+and JT received funding from Marisla Foundation, Packard Foundation,
+Walton Family Foundation, Summit Foundation, and Oak Foundation. FM was
+supported by NSF-CNH and NSF BioOce (grants DEB-1212124 and 1736830).
 
@@
 
-\bibliographystyle{frontiersinSCNS_ENG_HUMS}\bibliography{references}
+\bibliographystyle{frontiersinSCNS_ENG_HUMS}
+\bibliography{references}
 
@@
 \hline
+\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
+\hline
 \hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
-\hline
-\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
 \hline

### 2018-10-01 docs/paper/Villasenor-Derbez.Rmd
Manuscript ready for submission!

@@
 Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
-In the last six years, 18 TURF-reserves have been implemented in Mexico following a new regulation; their effectiveness has not been formally evaluated.
+In the last 12 years, 18 community-based TURF-reserves gained legal recognition thanks to a 2014 regulation; their effectiveness has not been formally evaluated.
 We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
@@
 A lack of clear ecological and socioeconomic effects likely results from a combination of factors.
-First, lobster and other invertebrate fisheries are already well managed, and it is unlikely that reserves might have a detectable effect.
+First, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect.
 Second, some of the reserves are not large enough to protect mobile species, like lobster.
@@
 
-Words in text: 3718 | Words in headers: 35 | Words outside text (captions, etc.): 192
+Words in text: 3714 | Words in headers: 35 | Words outside text (captions, etc.): 192
 
@@
 Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
-Since 2006, 45 Fish Refuges have been created along the Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18 of them implemented as TURF-reserves legally recognized as "Fish Refuges".
+Since 2012, old and new marine reserves have gained legal recognition as Fishing Refuges.
+Oh these, 18 were originally implemented as community-based TURF-reserves.
 However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
@@
 In 2006, the community decided to implement two marine reserves within their fishing grounds to protect commercially important invertebrate species; mainly lobster and abalone.
-While these reserves obtained legal recognition only in 2018, they have been well enforced since their implementation.
+While these reserves obtained legal recognition only in 2018 \citep{dof_website_2018}, they have been well enforced since their implementation.
 
@@
 Isla Natividad, for example, is part of a greater group of fishing cooperatives belonging to a Federation of Fishing Cooperatives.
-This group has been identified as a cohesive group that often cooperates to better manage their resources \citep{mccay_2017-1m,mccay_2014-CN,acevesbueno_2017}.
+This group has been identified as a cohesive group that often cooperates to better manage their resources \citep{mccay_2014-CN,mccay_2017-1m,acevesbueno_2017}.
 Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation.
@@
 While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth causal inference analysis that we undertake.
-Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our results to other communities in Mexico and other regions around the world.
+Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our insights to other similar reserves in Mexico and elsewhere around the world.
 
@@
 \begin{equation}
-I_{itj} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon
+I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon_{i,t,j}
 \label{eqn:reg_bio}
@@
 
-Where year-level fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
+Where year-level fixed effects are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
 The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site.
 When evaluating biomass and densities of the invertebrate or fish communities, we include $\sigma_j$ to control for species-level fixed effects.
+$\epsilon_{i,t,j}$ represents the error term of the regression.
 
@@
 \begin{equation}
-I = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jCom_j +\epsilon
+I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jCom_j +\epsilon_{i,t,j}
 \label{eqn:soc_reg}
@@
 We use the SES framework to evaluate each community.
-The use the SES framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators.
+The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators.
 We based our variable selection primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework.
 We also incorporate other relevant variables known to influence reserve performance following \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.
-Table \ref{table:ses} shows the selected variables, their definition and selected indicators.
+Table \ref{table:ses} shows the selected variables, their definition and values.
 
@@
 With these characteristics, one would expect the reserves to be effective.
-Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 6 years old) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; other community-based marine reserves in tropical ecosystems may take up to six years to show a spillover effect \citep{dasilva_2015-zX}.
+Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 6 years old) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \citep{dasilva_2015-zX}.
 
@@
 For the particular case of the marine reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
-At the broader scale, having full community support surely represents an advantage, but it is important that marine reserves meet essential design principles such as size and placement.
 Community-based marine reserves might have more benefits that result from indirect effects of the reserves, particularly providing resilience to shocks and management errors, and promoting social cohesion, which should be taken into account when evaluating the outcomes of TURF-reserves.
+Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.
 

### 2018-10-01 docs/paper/Villasenor-Derbez.tex
Manuscript ready for submission!

@@
 environmental variability and management errors, while ensuring that
-fishers reap the benefits of conservation investments. In the last six
-years, 18 TURF-reserves have been implemented in Mexico following a new
-regulation; their effectiveness has not been formally evaluated. We
-combine causal inference techniques and the Social-Ecological Systems
-framework to provide a holistic evaluation of community-based
-TURF-reserves in three coastal communities in Mexico. We find that while
-reserves have not yet achieved their stated goal of increasing the
-density of lobster and other benthic invertebrates, they continue to
-receive significant support from the fishing communities. A lack of
-clear ecological and socioeconomic effects likely results from a
-combination of factors. First, lobster and other invertebrate fisheries
-are already well managed, and it is unlikely that reserves might have a
-detectable effect. Second, some of the reserves are not large enough to
-protect mobile species, like lobster. Third, some of these reserves
-might be too young for the effects to show. Fourth, variable and extreme
-oceanographic conditions have impacted harvested populations. However,
-these reserves have shaped small-scale fishers' way of thinking about
-marine conservation, which can provide a foundation for establishing
-additional, larger marine reserves needed to effectively conserve mobile
-species.
+fishers reap the benefits of conservation investments. In the last 12
+years, 18 community-based TURF-reserves gained legal recognition thanks
+to a 2014 regulation; their effectiveness has not been formally
+evaluated. We combine causal inference techniques and the
+Social-Ecological Systems framework to provide a holistic evaluation of
+community-based TURF-reserves in three coastal communities in Mexico. We
+find that while reserves have not yet achieved their stated goal of
+increasing the density of lobster and other benthic invertebrates, they
+continue to receive significant support from the fishing communities. A
+lack of clear ecological and socioeconomic effects likely results from a
+combination of factors. First, local fisheries are already well managed,
+and it is unlikely that reserves might have a detectable effect. Second,
+some of the reserves are not large enough to protect mobile species,
+like lobster. Third, some of these reserves might be too young for the
+effects to show. Fourth, variable and extreme oceanographic conditions
+have impacted harvested populations. However, these reserves have shaped
+small-scale fishers' way of thinking about marine conservation, which
+can provide a foundation for establishing additional, larger marine
+reserves needed to effectively conserve mobile species.
 
@@
 
-Words in text: 3718 \textbar{} Words in headers: 35 \textbar{} Words
+Words in text: 3714 \textbar{} Words in headers: 35 \textbar{} Words
 outside text (captions, etc.): 192
@@
 reserves, which can protect one, some, or all resources within their
-boundaries. Since 2006, 45 Fish Refuges have been created along the
-Pacific, Gulf of California, and Mexican Caribbean coastlines, with 18
-of them implemented as TURF-reserves legally recognized as ``Fish
-Refuges''. However, their effectiveness has not yet been formally
-evaluated and reported in the scientific literature.
+boundaries. Since 2012, old and new marine reserves have gained legal
+recognition as Fishing Refuges. Oh these, 18 were originally implemented
+as community-based TURF-reserves. However, their effectiveness has not
+yet been formally evaluated and reported in the scientific literature.
 
@@
 protect commercially important invertebrate species; mainly lobster and
-abalone. While these reserves obtained legal recognition only in 2018,
-they have been well enforced since their implementation.
+abalone. While these reserves obtained legal recognition only in 2018
+\citep{dof_website_2018}, they have been well enforced since their
+implementation.
 
@@
 a cohesive group that often cooperates to better manage their resources
-\citep{mccay_2017-1m,mccay_2014-CN,acevesbueno_2017}. Likewise, Maria
+\citep{mccay_2014-CN,mccay_2017-1m,acevesbueno_2017}. Likewise, Maria
 Elena and Punta Herrero are representative of fishing cooperatives in
@@
 among communities and the socioeconomic and governance setting under
-which they operate, it is safe to cautiously generalize our results to
-other communities in Mexico and other regions around the world.
+which they operate, it is safe to cautiously generalize our insights to
+other similar reserves in Mexico and elsewhere around the world.
 
@@
 
-Where year-level fixed effects are represented by
-\(\gamma_{it} Year_t\), and \(\beta Zone_i\) captures the difference
-between reserve (\(Zone = 1\)) and control (\(Zone = 0\)) sites. The
-interaction term \(\lambda_{it} Year_t\times Zone_i\) represents the
-mean change in the indicator inside the reserve, for year \(t\), with
-respect to the year of implementation in the control site. When
-evaluating biomass and densities of the invertebrate or fish
-communities, we include \(\sigma_j\) to control for species-level fixed
-effects.
+Where year-level fixed effects are represented by \(\gamma_t Year_t\),
+and \(\beta Zone_i\) captures the difference between reserve
+(\(Zone = 1\)) and control (\(Zone = 0\)) sites. The interaction term
+\(\lambda_{it} Year_t\times Zone_i\) represents the mean change in the
+indicator inside the reserve, for year \(t\), with respect to the year
+of implementation in the control site. When evaluating biomass and
+densities of the invertebrate or fish communities, we include
+\(\sigma_j\) to control for species-level fixed effects.
+\(\epsilon_{i,t,j}\) represents the error term of the regression.
 
@@
 
-We use the SES framework to evaluate each community. The use the SES
+We use the SES framework to evaluate each community. The use of this
 framework standardizes our analysis and allows us to communicate our
@@
 \citet{edgar_2014-UO}. Table \ref{table:ses} shows the selected
-variables, their definition and selected indicators.
+variables, their definition and values.
 
@@
 short duration of protection, relative to the life histories of the
-protected species; other community-based marine reserves in tropical
-ecosystems may take up to six years to show a spillover effect
+protected species; community-based marine reserves in tropical
+ecosystems may take six years or more to show a spillover effect
 \citep{dasilva_2015-zX}.
@@
 reserves or merging existing polygons into larger areas should be
-evaluated and proposed to the communities. At the broader scale, having
-full community support surely represents an advantage, but it is
-important that marine reserves meet essential design principles such as
-size and placement. Community-based marine reserves might have more
-benefits that result from indirect effects of the reserves, particularly
-providing resilience to shocks and management errors, and promoting
-social cohesion, which should be taken into account when evaluating the
-outcomes of TURF-reserves.
+evaluated and proposed to the communities. Community-based marine
+reserves might have more benefits that result from indirect effects of
+the reserves, particularly providing resilience to shocks and management
+errors, and promoting social cohesion, which should be taken into
+account when evaluating the outcomes of TURF-reserves. Having full
+community support surely represents an advantage, but it is important
+that community-based TURF-reserves meet essential design principles such
+as size and placement so as to maximize their effectiveness.
 

### 2018-10-02 docs/paper/Villasenor-Derbez.Rmd
Manuscript submited

@@
 Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
-In the last 12 years, 18 community-based TURF-reserves gained legal recognition thanks to a 2014 regulation; their effectiveness has not been formally evaluated.
+In the last 12 years, 18 old and new community-based TURF-reserves gained legal recognition thanks to a 2014 regulation; their effectiveness has not been formally evaluated.
 We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.

### 2018-10-02 docs/paper/Villasenor-Derbez.tex
Manuscript submited

@@
 fishers reap the benefits of conservation investments. In the last 12
-years, 18 community-based TURF-reserves gained legal recognition thanks
-to a 2014 regulation; their effectiveness has not been formally
-evaluated. We combine causal inference techniques and the
+years, 18 old and new community-based TURF-reserves gained legal
+recognition thanks to a 2014 regulation; their effectiveness has not
+been formally evaluated. We combine causal inference techniques and the
 Social-Ecological Systems framework to provide a holistic evaluation of

### 2018-11-29 docs/paper/reviewer_comments.Rmd
Added PDF file with reviewer comments

---
title: "Reviewer Report"
output: pdf_document
---

# Letter from the editor

There are some rather loose statements in the ms, starting with the Abstract (see also Rev 1) where it is implied that a 'well-managed' fishery should result in no effect in a reserve. This is false. Most 'well-managed' fisheries aim to bring B0 to around 50% of unfished biomass, so reserves should on average contain double the density of fished areas. There is one unsubstantiated statement (line 81) of good enforcement, but could it not be just as likely that there is fishing activity in these reserves? Poaching is an equally valid explanation as any examine din the Discussion.

As per Reviewer 1, much more clarity is needed in Methods. It is unclear what data are being tested and what the sampling design was, especially the spatial arrangement of controls. The modeling approach was also confusing to me - you have categorical responses so why 'multiple regression'? Your reporting indicates a log-linear model (especially the use of lambda) which is probably the most appropriate structure for fish data.

There is also some fairly lazy scholarship here: e.g. BACI designs originate in papers other than those cited on line 112-113, and were demonstrated in papers well before those in 114-115, and again in 133-134. The Moland et al. paper may not be a good example anyway since their sampling was spatially confounded. Discussion of spatial comparisons and before-after studies seems to imply nobody has ever used monitoring time series before. 
I am also confused by lines 250-252 where it is implied that this does study does not in fact have fished controls in the design.

**Recommendation:** This ms requires major revision.

# Reviewer 1

**Recommendation:** Substantial revisions

## Q1 - Please summarize the main findings of the study.

The manuscript provides an analysis of ecological and economic fishery indicators across TURF-reserves in three Mexican communities, finding no substantial changes in target species or associated fisheries, or in value or quantity of landings. This is not necessarily surprising and the likely reasons are discussed. This study provides an example (adding to an existing literature) of community based reserves that have local support but do not necessarily deliver substantial fishery benefits.

## Q2 - Please highlight the limitations and strengths.

The question at the heart of this manuscript is an interesting one because the combination of TURFS and reserves offers the possibility of securing rights for fishers while also implementing conservation areas, thereby potentially avoiding some of the conflicts associated with exclusion from fishing grounds that can arise with MPA designation. However, I didn’t feel that the paper fully lived up to its potential in addressing this question (see Q3). Beyond the key findings as described above (Q1), there is a lot of additional material in the manuscript that didn’t seem to add much to the key message. If the focus on TURF-reserves is the key area of novelty, I think this should be further emphasised and explained – see more detailed comments below.

## Q3 - Please provide your detailed review report to the editor and authors (including any comments on the Q4 Check List):

This review focuses on the more substantial points, and I have not commented on minor issues.

### ABSTRACT

- I found the abstract somewhat misleading after reading the full ms because it implies that the findings include data on the level of support among fishing communities (L14), but this data is not presented in the results. In line with this the last sentence of the abstract seems to be speculation, and is not supported by evidence presented in the manuscript.

### INTRODUCTION

- The introduction is brief and not easy to follow. It could be usefully expanded to develop the argument for this study and demonstrate the novel elements. My suggestion would be to expand it to more clearly explain what TURFS are, and what benefits TURF-reserves are expected to deliver in comparison to TURFS or reserves alone.

- I found the introduction confusing in parts because some statements say there is little evidence around TURFs, while others seem to refer to documented benefits. There is quite a lot of switching between references to studies of TURFS, reserves, and TURF-reserves. A clearer explanation of the existing science would help convey the knowledge gap. One general issue is that there are several places in the ms where it was difficult to distinguish between theoretical/expected outcomes or relationships, and those supported by evidence from this or other studies – it would be helpful if the authors could clarify these areas.

- In the third paragraph it wasn’t clear to me whether the authors see TURF-reserves as synonymous with community-based marine reserves, or if not what the differences are. As the authors note, who these rights are conferred and recognised by is an important consideration and needs further explanation. In line with this the authors should perhaps revisit the title, which refers to community-based marine reserves, but not to TURFs.

- I suggest moving the specific details of Mexican TURFs to the case study section to provide context for the study sites, and focus on the broader issues in the introduction.

### METHODS

- In the study area section more information would be useful on the wider context (some of this is in the introduction

– e.g. the TURF system in Mexico and governance arrangements for this system and the reserves within it).

- The claims that the communities are representative and that insights could be generalised to elsewhere in the world could be toned down. Some evidence that the communities share similarities with other fishing communities is useful, but the authors themselves suggest in the discussion that wide generalisations are not possible.

- More information on data collection would help the reader to have greater confidence in the data and make the study more replicable – e.g. numbers of surveys/data points, location of control sites and counterfactuals, processes of site selection, etc. Some of this could be added to Table 1 (or perhaps to supplementary info?).

- With regard to the evaluation of SES – it wasn’t clear to me what the purpose of this evaluation was, and the data sources (and analysis of these) seemed sparse. There is little explanation of why these factors are important in relation to TURF-reserves, and only minimal reference to the original work that led to this framework.

- My assumption is that the statistical analysis is appropriate to the data and questions. This is not really my area of expertise, though as someone who is reasonably conversant with statistical modelling I did find it difficult to follow the explanation of the analysis.

### RESULTS

- Results are concisely presented. The key findings are clear but because the figures only show effect size it’s not easy to get a feel for the data and its variability across sites and controls.

- The presentation of governance information is brief. It seems that the purpose is to assess whether the characteristics of the system are conducive to effective local management of the resource, but this is not very informative given the depth of analysis – the same broad descriptions are presented for all three sites for many of the variables. I would question the value of this somewhat ‘tick box approach’ to SES systems across a small number of case studies. What are the questions driving this analysis and what does it add to the paper? It seems that much of this could be presented as a contextual summary of the study sites rather than as analysis.

### DISCUSSION

- The key message seems to be that there are no significant changes in the species targeted for protection, nor other fishery-related species or fishery economic benefits. There are a number of plausible explanations presented that are all well established in the MPA literature (reserve size, duration, quality of existing fisheries management and environmental variability). While I appreciate the importance of publishing negative results in the MPA literature, it’s a shame that some of these factors weren’t anticipated and explored in more detail.

- I found the second paragraph of the discussion difficult to follow – the key points here could be more clearly explained as they seem to point to the novelty of the statistical analysis.

- The latter part of the discussion focuses on the additional potential benefits of reserves including the positive perceptions of fishers noted in another study. My view is that these benefits are over-emphasised here, given that there is no data in the manuscript to support them. Furthermore the potential counter-arguments (e.g. that resources invested in these reserves might be put to better use) are not considered, which leads to a fairly one-sided view.

- It would be worth discussing fully the limitations of the data e.g. presumably the lack of data on fishing effort restricts interpretation of the time series, and the indicators representing socio-economic benefits of the MPAs are limited

## Q4 - Please comment on the methods, results and data interpretation. If there are any objective errors, or if the conclusions are not supported, you should detail your concerns.

Please see Q3

## Q5 - Check List

### Is the English language of sufficient quality?

- Yes

### Is the quality of the figures and tables satisfactory?

- No

#### Does the reference list cover the relevant literature adequately and in an unbiased manner?

- Yes

#### Are the statistical methods valid and correctly applied? (e.g. sample size, choice of test)

No answer given.

#### Are the methods sufficiently documented to allow replication studies?

- No

#### Are the data underlying the study available in either the article, supplement, or deposited in a repository? (Sequence/expression data, protein/molecule characterizations, annotations, and taxonomy data are required to be deposited in public repositories prior to publication) (yes/no/not applicable)

- No

# Reviewer 2

**Recommendation:** Moderate revisions

## Q1 - Please summarize the main findings of the study.

This paper assesses the effectiveness of 3 TURF-reserves using a Before-After-Control-Impact design. The results indicate that the reserves have not increased the density of targeted species. Besides the biological analysis, the authors also evaluate the social-ecological system at each study community based on selected indicators. They conclude that in order to maximise biological effectiveness, community-based TURF reserves need to have full community support as well as meet other reserve design criteria, such as appropriate size and location.

## Q2 - Please highlight the limitations and strengths.

Strengths: The study evaluates both the biological as well as social and governance system of marine reserves, which is not addressed simultaneously in many studies assessing marine reserve effectiveness.

Limitations:
The connection between the SES and biological evaluations need to be strengthened. At the same time, the SES provides a static picture of governance and resource conditions, whereas the biological assessment reflects changes through time. To bring these two components together, there is a need to provide information about how the SES has changed through time. Also the SES analysis lacks evaluation criteria. This inhibits a full picture of the combined biological and social contribution of TURF reserves.

See further comments in Q3 below.

## Q3 - Please provide your detailed review report to the editor and authors (including any comments on the Q4 Check List):

- This paper assesses the effectiveness of 3 TURF-reserves using a Before-After-Control-Impact design. The results indicate that the reserves have not increased the density of targeted species. Besides the biological analysis, the authors also evaluate the social-ecological system at each study community based on selected indicators. While it is encouraging to see the authors consider the social and governance aspects of marine reserves, it appears that this is secondary to the biological assessment. As it is presented now, I don’t see the added insights that the SES analysis brings to the study. The effectiveness of reserves are primarily assessed based on the biological indicators only. This is due to the lack of evaluation criteria for the SES component. I’d suggest using some rating system to rate each of the SES indicators. E.g., for RS4.1 stock status – range of possible values would be overfished, underfished, etc. Where would each study site fall along the gradient of underfished to overfished stocks?

- By doing this you can assess each site in terms of biological performance and SES conditions. This then allows you to establish a stronger connection on how biological and social factors together act to contribute to better protection and effectiveness of marine reserves. This to me, is where the study can make a contribution to the literature.

- Further to my comment above, my understanding of this paper is that the authors are making an argument for TURF-reserves. However the rationale for this does not come out clearly in the paper. There needs to be stronger justification provided for why the authors expect the 18 TURF-reserves to perform better than i) non-TURF community based marine reserves, and ii) TURF fisheries (but not necessarily marine reserves). Without this clarification, there is nothing to distinguish this study from others which assess the fishery impacts of marine reserves.

- Does fishing occur inside the marine reserves, or outside the marine reserves? If the fishing occurs inside the marine reserve then please clarify that the marine reserves are NOT no-take marine reserves.

- Were the fisheries at each study site managed as TURFs prior to the establishment of the marine reserves?

- L67. What is a triple bottom line evaluation of the effectiveness of community based marine reserves? Please clarify triple bottom line and what the criteria for this are.

### Methods

- Please provide the size of each of the reserves.

- Reserve size is an important factor affecting ecological outcomes, especially when comparing comparatively mobile vs non-mobile species (as noted by the authors in the Discussion). Why was the effect of reserve size not assessed?

- Are the control sites managed with TURFs even though they are non-reserves?

The socioeconomic analysis needs to address these points:

- Provide the ex-vessel prices of lobster and invertebrates at each study site.

- Did you control for difference in prices between sites?

- Would observed differences in revenue inside/outside reserves be due to difference in landings or due to price differences?

### Results

- Please provide results for biological and socioeconomic indicators - Average density/biomass, landings and revenue for lobster, invertebrates and fish by year in and outside marine reserves (control sites). If space is an issue then include these results in the Supplementary Materials.

- What was the species and community effect on biological and socioeconomic indicators, respectively?

- Table 2: Some variables need further clarification.

- RU5 – Number of targeted species is the variable, but this isn’t specified in the narrative. In addition to noting that lobster is the main target species please enter the total number of targeted species.

- A1 – Number of fishers. I assume that the numbers listed refer only to the number of fishing co-op members. Are there other non-coop fishers in the fishing communities? If yes, what is the ratio of co-op to non co-op fishers? The point of this question is to see how the benefits of TURFs are distributed among the entire community, not just within the fishing co-op.

- A3 – Level of isolation. The narrative says that fishing grounds and reserves are away from dense urban centers – can you provide the distance from each study site to the nearest urban centre (km).

- GS6.2 Which rules (larger min catch size, lower quotas etc) pertain to each site?

- GS9.1 Social monitoring – Do fishing coops at all 3 sites have the same monitoring and enforcement regime in place?

- GS9.2 Biophysical monitoring – Are all the listed monitoring activities performed at all 3 sites?

- GS 10.1 Graduated sanctions – Again, which of the penalties are relevant to each site?

- As it is presented now, the governance variables are a bit unclear as to which type of rules/regulations are practiced at each of the three sites. I recognise that this is because these variables are more qualitative in nature. However to make it clearer, the Narrative column could be split into 3 - one column per study site, and the relevant information entered under each site column.

- Table 2. As per my comment about rating the SES indicators, add a column in between the “Variable” and “Narrative” column to describe the range of potential values/conditions for each “Variable”. E.g., for Governance system Social monitoring can range from “Strong” to “Weak”. You could attach points to a gradient of “Strong” to “Weak” to produce a quantitative assessment, or enter a qualitative assessment. Then each site will either be assessed with points or a qualitative descriptor. The main point is to provide the reader with a better sense of what a “strong” or “weak” governance system looks like, and where each site falls along this gradient. This is missing from the current table.

- Fig 2B and 2D. Why are there big fluctuations in fish biomass from Years 7-10, but density stays pretty much the same for the same time period?

### Discussion

- L238-240: “Our analysis of economic data supports this hypothesis….reserve implementation”. I don’t understand what the authors are trying to convey in this sentence. What are the expected short-term costs associated to the first years of reserve implementation? Please clarify.

- L208-209 – The authors state that “Understanding the socio-ecological context ….provide insights as to why this happens”, but then don’t provide further elaboration. The reader is not getting any insight as to why the communities continue to support the reserves. Even though this is presumably covered in the provided reference to Ayer et al. 2018 a brief explanation is required here. Please add this to the Discussion.

- Clearly state the lessons learned. L274 simply says that “Lessons learned from these cases can guide implementation…” But what are these lessons?

- L277 the authors state that “Community-based marine reserves might have more benefits…” Are the authors referring to the reserves assessed in this study, or to community based marine reserves in general? If referring to this study, the conclusion is not supported by the results, as the authors did not compare community-based to non-community based reserves. Need to clarify this point.

- L268-269 “Furthermore, the lack of effectiveness observed in these reserves should not be generalizable to other reserves established under the same legal framework”. This appear to contradict L99-100 which says that “it is safe to cautiously generalize our insights to other similar reserves…”

- L196. “…the benefits of conservation directly benefit the members of the fishing cooperatives…” But what about benefits to the community as a whole? Going back to my question about number of actors in Table 2 – I assume that the invertebrate fisheries are high value fisheries. As such, it’s important to consider how the benefits of TURF-reserves are distributed among the entire community. Please provide some discussion about the distribution of fishery benefits – do the number of TURF fishers make up only a small proportion of total active fishers in the community? Or are all fishers in each community able to participate in the TURFs?

- The Discussion is focused mainly on ecological factors which may have affected the effectiveness of the TURF-reserves. What about potential changes in social-economic and governance conditions over the assessed time period which may have affected the reserve? For instance was there a breakdown in enforcement which may have led to illegal fishing, thereby resulting in lower density of target species? What about market drivers that may have motivated increased fishing pressure?

- Elaboration on these factors will strengthen the authors’ attempt to bring in the SES framework to this primarily biological analysis.

### Minor comments:

Figure 2 caption: first line (IN; red circles) “circles” is spelled wrong.

## Q4 - Please comment on the methods, results and data interpretation. If there are any objective errors, or if the conclusions are not supported, you should detail your concerns

- These comments are covered in Q3 above.

## Q5 - Check List

### Is the English language of sufficient quality?

- Yes

### Is the quality of the figures and tables satisfactory?

- Yes

### Does the reference list cover the relevant literature adequately and in an unbiased manner?

- Yes

### Are the statistical methods valid and correctly applied? (e.g. sample size, choice of test)

- Yes

### Are the methods sufficiently documented to allow replication studies?

- Yes

### Are the data underlying the study available in either the article, supplement, or deposited in a repository? (Sequence/expression data, protein/molecule characterizations, annotations, and taxonomy data are required to be deposited in public repositories prior to publication) (yes/no/not applicable)

- Yes

### 2018-11-30 docs/paper/Villasenor-Derbez.Rmd
Added missing polygon of TURF in fig 1

@@
 ph <- ggplot() +
-  geom_sf(data = turf[3,], fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
+  geom_sf(data = turf[c(1,3),], fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
   geom_sf(data = coastline_mx) +

### 2018-12-03 docs/paper/Villasenor-Derbez.Rmd
Modified title, clearified abstract, and documented setup chunk close #11, related to #19

@@
 ---
-title: "Effectiveness of community-based marine reserves in small-scale fisheries"
+title: "Effectiveness of community-based TURF-reserves in small-scale fisheries"
 rtitle: "Community-based marine reserves"
 author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,2}$, Stuart Fulton$^{3}$, Álvin Suarez$^{3}$, Arturo Hernández-Velasco$^{3}$, Jorge Torre$^{3}$, Fiorenza Micheli$^{4}$"
-institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
-$^{2}$Nicholas School of the Environment, Duke University, Beaufort, NC, USA\\newline
-$^{3}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico\\newline
-$^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA"
+institution: |
+  $^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA \newline
+  $^{2}$Nicholas School of the Environment, Duke University, Beaufort, NC, USA \newline
+  $^{3}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico \newline
+  $^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA" \newline
 corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
@@
 A lack of clear ecological and socioeconomic effects likely results from a combination of factors.
-First, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect.
-Second, some of the reserves are not large enough to protect mobile species, like lobster.
-Third, some of these reserves might be too young for the effects to show.
-Fourth, variable and extreme oceanographic conditions have impacted harvested populations.
+First, some of these reserves might be too young for the effects to show.
+Second, the reserves are not large enough to protect mobile species, like lobster.
+Third, variable and extreme oceanographic conditions have impacted harvested populations.
+Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in landings.
 However, these reserves have shaped small-scale fishers’ way of thinking about marine conservation, which can provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+# Setu up options
+knitr::opts_chunk$set(echo = FALSE, # Hides code from manuscript
+                      warning = FALSE, # Hides warning messages when loading packages
+                      message = FALSE, # Hides messages
+                      eval = TRUE) # Set to FALSE to compile manuscript without running code
 
+# Specify that the output of knitr should be LaTex
 options(knitr.table.format = "latex")

### 2018-12-03 docs/paper/Villasenor-Derbez.Rmd
Added packrat libraries and docs helpers

@@
   $^{3}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico \newline
-  $^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA" \newline
+  $^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
 corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
@@
 
-Words in text: 3714 | Words in headers: 35 | Words outside text (captions, etc.): 192
+<!-- Words in text: 3714 | Words in headers: 35 | Words outside text (captions, etc.): 192 -->
 
-\clearpage
+<!-- \clearpage -->
 
@@
 data(World)
-W <- spTransform(World, proj) %>% 
-  st_as_sf()
+World %<>% st_transform(proj)
 
@@
 mex <- ggplot() +
-  geom_sf(data = W) +
+  geom_sf(data = World) +
   geom_point(data = location_labels,
@@
   labs(x = "", y = "") +
-  scalebar(data = turf[3,],
+  ggsn::scalebar(data = turf[c(1:3),1],
            dist = 5,
@@
 pbc <- ggplot() +
-  geom_sf(data = turf, fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
+  geom_sf(data = turf[4,], fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
   lims(x = c(-115.65, -115.109), y = c(27.7, 28)) +
@@
   labs(x = "", y = "") +
-  scalebar(data = turf[4,], dist = 10, dd2km = T, model = 'WGS84', location = "bottomleft", st.size = 2, st.dist = 0.05, height = 0.1)
+  scalebar(data = turf[c(1:4),1], dist = 10, dd2km = T, model = 'WGS84', location = "bottomleft", st.size = 2, st.dist = 0.05, height = 0.1)
 

### 2018-12-03 docs/test.Rmd
Added packrat libraries and docs helpers

--- 2018-10-02 docs/paper/Villasenor-Derbez.tex
@@
-\documentclass{frontiersSCNS}
-\usepackage{url,hyperref,lineno,microtype,subcaption}
-\usepackage[onehalfspacing]{setspace}
-\usepackage{float}
-
-\linenumbers
-
-\usepackage[utf8]{inputenc}
-\floatplacement{figure}{H}
-
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
-\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
-Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
-Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
-Micheli\(^{4}\)}
-% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
-% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
-of California, Santa Barbara, Santa Barbara, CA,
-USA\newline \(^{2}\)Nicholas School of the Environment, Duke University,
-Beaufort, NC, USA\newline \(^{3}\)Comunidad y Biodiversidad A.C.,
-Guaymas, Sonora, Mexico\newline \(^{4}\)Hopkins Marine Station and
-Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
-% The Corresponding Author should be marked with an asterisk
-% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
-Santa Barbara, Santa Barbara, CA, 93106}
-
-\def\corrEmail{\href{mailto:juancarlos@ucsb.edu}{\nolinkurl{juancarlos@ucsb.edu}}}
-
-\begin{document}
-\onecolumn
-\firstpage{1}
-
-\title[Community-based marine reserves]{Effectiveness of community-based marine reserves in small-scale
-fisheries}
-
-\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-\extraAuth{}
-
-\maketitle
-
-\begin{abstract}
-
-Coastal marine ecosystems provide livelihoods for small-scale fishers
-and coastal communities around the world. Small-scale fisheries face
-great challenges since they are difficult to monitor, enforce, and
-manage. Combining territorial user rights for fisheries (TURF) with
-no-take marine reserves to create TURF-reserves can improve the
-performance of small-scale fisheries by buffering fisheries from
-environmental variability and management errors, while ensuring that
-fishers reap the benefits of conservation investments. In the last 12
-years, 18 old and new community-based TURF-reserves gained legal
-recognition thanks to a 2014 regulation; their effectiveness has not
-been formally evaluated. We combine causal inference techniques and the
-Social-Ecological Systems framework to provide a holistic evaluation of
-community-based TURF-reserves in three coastal communities in Mexico. We
-find that while reserves have not yet achieved their stated goal of
-increasing the density of lobster and other benthic invertebrates, they
-continue to receive significant support from the fishing communities. A
-lack of clear ecological and socioeconomic effects likely results from a
-combination of factors. First, local fisheries are already well managed,
-and it is unlikely that reserves might have a detectable effect. Second,
-some of the reserves are not large enough to protect mobile species,
-like lobster. Third, some of these reserves might be too young for the
-effects to show. Fourth, variable and extreme oceanographic conditions
-have impacted harvested populations. However, these reserves have shaped
-small-scale fishers' way of thinking about marine conservation, which
-can provide a foundation for establishing additional, larger marine
-reserves needed to effectively conserve mobile species.
-
-\medskip
-\tiny
- \keyFont{ \section{Keywords:} TURF-reserves, Causal Inference, Social-Ecological Systems, Marine
-Protected Areas, Marine Conservation, Small-Scale Fisheries}
-
-\end{abstract}
-
-Words in text: 3714 \textbar{} Words in headers: 35 \textbar{} Words
-outside text (captions, etc.): 192
-
-\clearpage
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Marine ecosystems around the world sustain significant impacts due to
-overfishing and unsustainable fishing practices
-\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
-small-scale fisheries face great challenges since they tend to be hard
-to monitor and enforce \citep{costello_2012}. Recent research shows that
-combining Territorial Use Rights for Fisheries (TURFs) with no-take
-marine reserves (MRs) can greatly improve the performance of coastal
-fisheries and the health of the local resources
-\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
-these systems increase the benefits of spatial access rights allowing
-the maintenance of healthy resources
-\citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
-are successful \citep{costello_2010-Ix,smallhornwest_2018}, there is
-little empirical evidence of their effectiveness and the drivers of
-their success \citep{afflerbach_2014-HP,lester_2017}.
-
-The performance of these systems depends on how environmental and social
-factors combine and interact. The science of marine reserves has largely
-focused on understanding the ecological effects of these areas, which
-include increased biomass, species richness, and densities of organisms
-within the protected regions, climate change mitigation, and protection
-from environmental variability
-\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
-Modelling studies show that fishery benefits of marine reserves depend
-on initial stock status and the management under which the fishery
-operates, as well as reserve size and the amount of larvae exported from
-these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}. Other research has
-focused on the relationship between socioeconomic and governance
-structures and reserve effectiveness
-\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
-knowledge, no studies exist that evaluate TURF-reserves from both a
-social and ecological perspective. This is especially important in
-social-ecological coastal systems dominated by close interactions and
-feedbacks between people and natural resources \citep{ostrom_2009-hg}.
-
-TURF-reserves can be created as community-based marine reserves,
-voluntarily established and enforced by local communities. This
-bottom-up approach increases compliance and self-enforcement, and
-reserves can yield benefits similar to systematically-designed reserves
-\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8,smallhornwest_2018}.
-Community-based spatial closures occur in different contexts, like the
-\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
-\citep{bohnsack_2004,johannes_2002}. However, MRs are difficult to
-enforce if they are not legally recognized, and fishers rely on the
-exclusive access granted by the TURF. In an effort to bridge this
-normative gap, Mexican Civil Society Organizations (CSOs) served as a
-link between fishers and government, and created a legal framework that
-solves this governance issue. In Mexico, a new norm was created in 2014
-allowing fishers to request the legal recognition of community-based
-reserves as ``Fish Refuges'' (\emph{Zona de Refugio Pesquero};
-\citet{nom}). Fish refuges can be implemented as temporal or partial
-reserves, which can protect one, some, or all resources within their
-boundaries. Since 2012, old and new marine reserves have gained legal
-recognition as Fishing Refuges. Oh these, 18 were originally implemented
-as community-based TURF-reserves. However, their effectiveness has not
-yet been formally evaluated and reported in the scientific literature.
-
-Here, we combine causal inference techniques and the Social-Ecological
-Systems (SES) framework to provide a holistic evaluation of
-community-based TURF-reserves in three coastal communities in Mexico.
-These three case studies span a range of ecological and social
-conditions representative of different regions of Mexico. The objective
-of this work is twofold. First, to provide a triple bottom line
-evaluation of the effectiveness of community-based marine reserves,
-which may inform similar processes in other countries. Second, to
-evaluate the effectiveness of TURF-reserves established as Fish Refuges
-in Mexico to identify opportunities where improvement or adjustment
-might lead to increased effectiveness. We draw from lessons learned in
-these three case studies and provide management recommendations to
-maximize the effectiveness of community-based marine reserves in
-small-scale fisheries in Mexico and in other regions around the world
-where this tool is used to manage and rebuild their coastal fisheries.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{study-area}{%
-\subsection{Study area}\label{study-area}}
-
-We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A). The
-first one was created by the \emph{Buzos y Pescadores de la Baja
-California} fishing cooperative, located in Isla Natividad in the Baja
-California Peninsula (Fig \ref{fig:map}B). The main fishery in the
-island is the spiny lobster (\emph{Panulirus interruptus}), but other
-resources like finfish, sea cucumber, read sea urchin, snail, and
-abalone are also an important source of income. In 2006, the community
-decided to implement two marine reserves within their fishing grounds to
-protect commercially important invertebrate species; mainly lobster and
-abalone. While these reserves obtained legal recognition only in 2018
-\citep{dof_website_2018}, they have been well enforced since their
-implementation.
-
-The other two TURF-reserves are located in Maria Elena and Punta
-Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
-Isla Nativdad, which hosts a well established fishing community, Maria
-Elena is a fishing camp --visited intermittently during the fishing
-season-- belonging to the \emph{Cozumel} fishing cooperative; Punta
-Herrero is home to the \emph{José María Azcorra} fishing cooperative,
-and similar to Isla Natividad hosts a local community. Their main
-fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but
-they also target finfish in the off-season. Maria Elena and Punta
-Herrero established eight and four marine reserves in 2012 and 2013,
-respectively. These reserves have been legally recognized as Fishing
-Refuges since their creation \citep{dof_website_2012,dof_website_2013}.
-
-These communities are representative of their region in terms of
-ecology, socioeconomic, and governance aspects. Isla Natividad, for
-example, is part of a greater group of fishing cooperatives belonging to
-a Federation of Fishing Cooperatives. This group has been identified as
-a cohesive group that often cooperates to better manage their resources
-\citep{mccay_2014-CN,mccay_2017-1m,acevesbueno_2017}. Likewise, Maria
-Elena and Punta Herrero are representative of fishing cooperatives in
-the Mexican Caribbean, which are also part of a regional Federation.
-Together, these three communities provide an accurate representation of
-other fishing communities in each of their regions. While each region
-has additional communities that have established community-based
-TURF-reserves, available data would not allow us to perform the in-depth
-causal inference analysis that we undertake. Yet, given the similarities
-among communities and the socioeconomic and governance setting under
-which they operate, it is safe to cautiously generalize our insights to
-other similar reserves in Mexico and elsewhere around the world.
-
-\hypertarget{data-collection}{%
-\subsection{Data collection}\label{data-collection}}
-
-We use three main sources of information to evaluate these reserves
-across the ecological, socioeconomic, and governance dimensions.
-Ecological data come from the annual ecological monitoring of reserve
-and control areas, carried out by members from each community and
-personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
-(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
-abundances of fish and invertebrate species along replicate transects
-(30 \(\times\) 2 m each) at depths 5-20 m in the reserves and control
-sites \citep{fulton_2018,fulton_2019,suman_2010-ez}. Size structures are
-also collected during fish surveys. We define control sites as regions
-with habitat characteristics similar to the corresponding reserves, and
-that presumably had a similar probability of being selected as reserves
-during the design phase. We focus our evaluation on sites where data are
-available for reserve and control sites, before and after the
-implementation of the reserve. This provides us with a
-Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
-allows us to capture and control for temporal and spatial dynamics
-\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
-techniques have proven effective to evaluate marine reserves, as they
-allow us to causally attribute observed changes to the intervention
-\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
-annually, and at least once before implementation of the reserves.
-
-Socioeconomic data come from landing receipts reported to the National
-Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
-Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
-(Kg) and revenues (MXP) for cooperatives with and without marine
-reserves. Cooperatives incorporated in this analysis belong to larger
-regional-level Cooperative Federations, and are exposed to the same
-markets and institutional frameworks, making them plausible controls
-\citep{mccay_2017-1m,ayer_2018}. Landings and revenues were aggregated
-at the cooperative-year level, and revenues were adjusted to represent
-2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.
-
-Data for the evaluation of the SES were collected at the community-level
-from official documents used in the creation and designation of the
-marine reserves
-\citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on
-the authors' experience and knowledge of the communities. These include
-information on the Resource Systems, Resource Units, Actors, and
-Governance System (Table \ref{table:ses}).
-
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
-
-We evaluate the effect that marine reserves have had on four ecological
-and two socioeconomic indicators (Table \ref{table:indicators}). Recall
-that reserves were implemented to protect lobster and other benthic
-invertebrates. However, we also use the available fish data to test for
-associated co-benefits.
-
-We use a difference-in-differences analysis to evaluate these
-indicators. This approach allows us to estimate the effect that the
-reserve had by comparing trends across time and treatments
-\citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis of
-ecological indicators is performed with a multiple linear regression of
-the form:
-
-\begin{equation}
-I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon_{i,t,j}
-\label{eqn:reg_bio}
-\end{equation}
-
-Where year-level fixed effects are represented by \(\gamma_t Year_t\),
-and \(\beta Zone_i\) captures the difference between reserve
-(\(Zone = 1\)) and control (\(Zone = 0\)) sites. The interaction term
-\(\lambda_{it} Year_t\times Zone_i\) represents the mean change in the
-indicator inside the reserve, for year \(t\), with respect to the year
-of implementation in the control site. When evaluating biomass and
-densities of the invertebrate or fish communities, we include
-\(\sigma_j\) to control for species-level fixed effects.
-\(\epsilon_{i,t,j}\) represents the error term of the regression.
-
-Socioeconomic indicators are evaluated with a similar approach. Due to
-data constrains, we only evaluate socioeconomic data for Isla Natividad
-(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
-used as counterfactuals that allow us to control for unobserved
-time-invariants. Each focal community (Isla Natividad and Maria Elena)
-has three counterfactual communities.
-
-\begin{equation}
-I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jCom_j +\epsilon_{i,t,j}
-\label{eqn:soc_reg}
-\end{equation}
-
-The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in
-this case the \(Treated\) dummy variable indicates if the community has
-a reserve (\(Treated = 1\)) or not (\(Treated = 0\)) and \(\sigma_jCom\)
-captures community-level fixed-effects. These regression models allow us
-to establish a causal link between the implementation of marine reserves
-and the observed trends by accounting for temporal and spatial dynamics
-\citep{depalma_2018}. The effect of the reserve is captured by the
-\(\lambda_t\) coefficient, and represents the difference observed
-between the control site before the implementation of the reserve and
-the treated sites at time \(t\) after controlling for other time and
-space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
-All model coefficients were estimated via ordinary least-squares and
-heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
-analyses were performed in R version 3.5.1 (2018-07-02) and R Studio
-version 1.1.456 \citep{R_2018}.
-
-We use the SES framework to evaluate each community. The use of this
-framework standardizes our analysis and allows us to communicate our
-results in a common language across fields by using a set of previously
-defined variables and indicators. We based our variable selection
-primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who
-operationalized and analyzed Mexican fishing cooperatives using this
-framework. We also incorporate other relevant variables known to
-influence reserve performance following \citet{difranco_2016-Xw} and
-\citet{edgar_2014-UO}. Table \ref{table:ses} shows the selected
-variables, their definition and values.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-The following sections present the effect that marine reserves had on
-each of the biological and socioeconomic indicators for each coastal
-community. Results are presented in terms of the difference through time
-and across sites, relative to the control site on the year of
-implementation (\emph{i.e.} effect size \(\lambda_t\)). We also provide
-an overview of the governance settings of each community, and discuss
-how these might be related to the effectiveness and performance of the
-reserves.
-
-\hypertarget{biological-effects}{%
-\subsection{Biological effects}\label{biological-effects}}
-
-Indicators showed ambiguous responses through time for each reserve.
-Figure \ref{fig:indicators}A shows positive effect sizes for lobster
-densities in Isla Natividad and Punta Herrero during the first years,
-but the effect is eroded through time. In the case of Maria Elena,
-positive changes were observed in the third and fourth year. These
-effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
-Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
-for Maria Elena, but are not significantly different from zero
-(\(p > 0.05\)). Likewise, no significant changes were detected in fish
-biomass or invertebrate and fish densities (Fig.
-\ref{fig:indicators}B-D), where effect sizes oscillated around zero
-without clear trends. Full tables with model coefficients are presented
-in the supplementary materials (S1 Table, S2 Table, S3 Table).
-
-\hypertarget{socioeconomic-effects}{%
-\subsection{Socioeconomic effects}\label{socioeconomic-effects}}
-
-Lobster landings and revenue were only available for Isla Natividad and
-Maria Elena (Fig \ref{fig:lobsters}). For all years before
-implementation, the effect sizes are close to zero, indicating that the
-control and treatment sites have similar pre-treatment trends,
-suggesting that these are plausible controls. However, effect sizes do
-not change after the implementation of the reserve. Interestingly, the
-negative effect observed for Isla Natividad on year 5 correspond to the
-2011 hypoxia events. The only positive change observed in lobster
-landings is for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
-post-implementation data for Maria Elena do not show a significant
-effect of the reserve. Isla Natividad shows higher revenues after the
-implementation of the reserve, as compared to the control communities.
-However, these changes are not significant and are associated with
-increased variation. Full tables with model coefficients are presented
-in the supplementary materials (S4 Table, S5 Table).
-
-\hypertarget{governance}{%
-\subsection{Governance}\label{governance}}
-
-Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed
-communities share similarities known to foster sustainable resource
-management and increase reserve effectiveness. For example, fishers
-operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide
-exclusive access to resources and reserves. Along with their relatively
-small groups (A1 - Number of relevant actors), Isolation (A3),
-Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated
-sanctions (GS10.1), these fisheries have solid governance structures
-that enable them to monitor their resources and enforce rules to ensure
-sustainable management. In general, success of conservation initiatives
-depends on the incentives of local communities to maintain a healthy
-status of the resources upon which they depend \citep{jupiter_2017}. Due
-to the clarity of access rights and isolation, the benefits of
-conservation directly benefit the members of the fishing cooperatives,
-which have favored the development of efficient community-based
-enforcement systems. However, our SES analysis also highlights factors
-that might hinder reserve performance or mask outcomes. While total
-reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual
-reserves are often small (RS3), and relatively young (RS5).
-Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely
-that marine reserves will result in increased catches.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Our results indicate that these TURF-reserves have not increased lobster
-densities. Additionally, no co-benefits were identified when using other
-ecological indicators aside from the previously reported buffering
-effect that reserves can have to environmental variability in Isla
-Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
-pertaining landings and revenues showed little to no change after
-reserve implementation. Despite the lack of evidence of the
-effectiveness of these reserves, most of the communities show a positive
-perception about their performance and continue to support their
-presence \citep{ayer_2018}. Understanding the social-ecological context
-in which these communities and their reserves operate might provide
-insights as to why this happens.
-
-Some works evaluate marine reserves by performing inside-outside
-\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or
-before-after comparisons \citep{betti_2017-lq}. The first approach does
-not address temporal variability, and the second can not distinguish
-between the temporal trends in a reserve and the entire system
-\citep{depalma_2018}. Our approach to evaluate the temporal and spatial
-changes provides a more robust measure of reserve effectiveness. For
-example, we capture previously described patterns like the rapid
-increase observed for lobster densities in Isla Natividad on the sixth
-year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), a year after the
-hypoxia events described by \citet{micheli_2012-EU}, which caused mass
-mortality of sedentary organisms such as abalone and sea urchins, but
-not lobster and finfish. Yet, our empirical approach assumes control
-sites are a plausible counterfactual for treated sites. This implies
-that treated sites would have followed the same trend as control sites,
-had the reserves not been implemented. Nonetheless, temporal trends for
-each site don't show any significant increases (S1 Table, S2 Table, S3
-Table), supporting our findings of lack of change in the indicators
-used.
-
-A first possible explanation for the lack of effectiveness may be the
-young age of the reserves. Literature shows that age and enforcement are
-important factors that influence reserve effectiveness
-\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
-reserves, and our SES analysis suggests that all communities have a
-well-established community-based enforcement system. With these
-characteristics, one would expect the reserves to be effective. Maria
-Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
-\textless{} 6 years old) and effects may not yet be evident due to the
-short duration of protection, relative to the life histories of the
-protected species; community-based marine reserves in tropical
-ecosystems may take six years or more to show a spillover effect
-\citep{dasilva_2015-zX}.
-
-Another key condition for effectiveness is reserve size
-\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
-attributed to poor ecological coherence in reserve design (\emph{sensu}
-\citet{rees_2018}). Previous research has shown that reserves in Isla
-Natividad yield fishery benefits for the abalone fishery
-\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
-perhaps the reserves provide enough protection to these sedentary
-invertebrates, but not lobsters. Design principles developed by
-\citet{green_2017} for marine reserves in the Caribbean state that
-reserves ``should be more than twice the size of the home range of
-adults and juveniles'', and suggest that reserves seeking to protect
-spiny lobsters should have at least 14 km across. Furthermore, fishers
-may favor implementation of reserves that pose low fishing costs due to
-their small size or location. Our analysis of economic data supports
-this hypothesis, as neither landings nor revenues showed the expected
-short-term costs associated to the first years of reserve implementation
-\citep{ovando_2016-Wg}.
-
-Even if reserves had appropriate sizes and were placed in optimal
-locations, there are other plausible explanations for the observed
-patterns. For instance, marine reserves are only likely to provide
-fisheries benefits if initial population sizes are low and the fishery
-is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
-fisheries were certified by the Marine Stewardship Council
-\citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed
-via species-specific minimum catch sizes, seasonal closures, protection
-of ``berried'' females, and escapement windows where traps are allowed
-\citep{dof_website_1993}. It is uncertain whether such a well-managed
-fishery will experience additional benefits from marine reserves.
-Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have
-greater biomass and richness than areas operating under open access.
-This might reduce the difference between indicators from the TURF and
-reserve sites, making it difficult to detect such a small change.
-Further research should focus on evaluating sites in the reserve, TURF,
-and open access areas or similar Fish Refuges established without the
-presence of TURFs where the impact of the reserves might be larger.
-
-Finally, extreme conditions, including prolonged hypoxia, heat waves,
-and storms have affected both the Pacific and Caribbean regions, with
-large negative impacts of coastal marine species and ecosystems
-\citep{cavole_2016,hughes_2018,breitburg_2018}. The coastal ecosystems
-where these reserves are located have been profoundly affected by these
-events \citep{micheli_2012-EU,woodson}. Effects of protection might be
-eliminated by the mortalities associated with these extreme conditions.
-
-While the evaluated reserves have failed to provide fishery benefits up
-to now, there are a number of additional ecological, fisheries, and
-social benefits. Marine reserves provide protection to a wider range of
-species and vulnerable habitat. These sites can serve as an insurance
-against uncertainty and errors in fisheries management, as well as mild
-environmental shocks
-\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
-Self-regulation of fishing effort (\emph{i.e.} reduction in harvest) can
-serve as a way to compensate for future declines associated to
-environmental variation \citep{finkbeiner_2018}. Embarking in a marine
-conservation project can bring the community together, which promotes
-social cohesion and builds social capital \citep{fulton_2019}. Showing
-commitment to marine conservation and sustainable fishing practices
-allows fishers to have greater bargaining power and leverage over
-fisheries management \citep{prezramrez_2012}. Furthermore, the lack of
-effectiveness observed in these reserves should not be generalizable to
-other reserves established under the same legal framework (\emph{i.e.}
-Fish Refuges) in Mexico, and future research should aim at evaluating
-other areas that have also been established as bottom-up processes but
-without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
-or others established through a top-down process (\emph{i.e.}
-\citet{dof_websiteU_2018}).
-
-Community-based marine reserves in small-scale fisheries can be helpful
-conservation and fishery management tools when appropriately
-implemented. Lessons learned from these cases can guide implementation
-of community-based marine reserves elsewhere. For the particular case of
-the marine reserves that we evaluate, the possibility of expanding
-reserves or merging existing polygons into larger areas should be
-evaluated and proposed to the communities. Community-based marine
-reserves might have more benefits that result from indirect effects of
-the reserves, particularly providing resilience to shocks and management
-errors, and promoting social cohesion, which should be taken into
-account when evaluating the outcomes of TURF-reserves. Having full
-community support surely represents an advantage, but it is important
-that community-based TURF-reserves meet essential design principles such
-as size and placement so as to maximize their effectiveness.
-
-\section*{Conflict of Interest Statement}
-
-The authors declare that the research was conducted in the absence of
-any commercial or financial relationships that could be construed as a
-potential conflict of interest.
-
-\section*{Author Contributions}
-
-JC and AS conceived the idea. JC and EA analyzed data, discussed the
-results, and wrote the first draft. FM, SF, AS, JT, and AHV discussed
-the results and edited the manuscript. All authors provided valuable
-contributions.
-
-\section*{Funding}
-
-JCVD received funding from UCMexus - CONACyT Doctoral Fellowship (CVU
-669403) and the Latin American Fisheries Fellowship Program. AS, AHV, SF
-and JT received funding from Marisla Foundation, Packard Foundation,
-Walton Family Foundation, Summit Foundation, and Oak Foundation. FM was
-supported by NSF-CNH and NSF BioOce (grants DEB-1212124 and 1736830).
-
-\section*{Acknowledgments}
-
-The authors wish to acknowledge Imelda Amador for contributions on the
-governance data, as well as pre-processing biological data. This study
-would have not been possible without the effort by members of the
-fishing communities here mentioned, who participated in the
-data-collection process.
-
-\clearpage
-
-\bibliographystyle{frontiersinSCNS_ENG_HUMS}
-\bibliography{references}
-
-\clearpage
-
-\section*{Figure captions}
-
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-7-1.pdf}
-\caption{\label{fig:unnamed-chunk-7}\label{fig:map}Location of the three
-coastal communities studied (A). Isla Natividad (B) is located off the
-Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
-in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
-polygons the marine reserves.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{\label{fig:unnamed-chunk-8}\label{fig:indicators}Effect sizes for
-marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
-blue triangles), and Punta Herrero (PH; green squares) for lobster
-densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
-densities (C), and fish densities (D). Plots are ordered by survey type
-(left column: invertebrates; right column: fish). Points are jittered
-hotizontally to avoid overplotting. Points indicate the effect size and
-standard errors. Years have been centered to year of implementation.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-9-1.pdf}
-\caption{\label{fig:unnamed-chunk-9}\label{fig:lobsters}Effect sizes for
-lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
-circles) and Maria Elena (ME; blue triangles). Points indicate the
-effect size and standard errors. Years have been centered to year of
-implementation.}
-\end{figure}
-
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-10}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
-\centering
-\begin{tabular}[t]{l|l}
-\hline
-Indicator & Units\\
-\hline
-\multicolumn{2}{l}{\textbf{Biological}}\\
-\hline
-\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
-\hline
-\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
-\hline
-\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
-\hline
-\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
-\hline
-\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
-\hline
-\hspace{1em}Income from target species & M MXP\\
-\hline
-\hspace{1em}Landings from target species & Metric Tonnes\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-11}\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).
-Alphanumeric codes follow \citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12cm}}
-\hline
-Variable & Narrative\\
-\hline
-\multicolumn{2}{l}{\textbf{Resource System (RS)}}\\
-\hline
-\hspace{1em}RS2 - Clarity of system boundaries: Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates.
-Reserve placement is decided by the community.
-Fishers use GPS units and landmarks.\\
-\hline
-\hspace{1em}RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
-\hline
-\hspace{1em}RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
-\hline
-\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) MSC certified.\\
-\hline
-\hspace{1em}*RS5 - Age of reserves: Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
-\hline
-\multicolumn{2}{l}{\textbf{Resource Unit (RU)}}\\
-\hline
-\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish.
-Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates.\\
-\hline
-\multicolumn{2}{l}{\textbf{Actors (A)}}\\
-\hline
-\hspace{1em}A1 - Number of relevant actors: Number of fishers & IN = 98; ME = 80; PH = 21\\
-\hline
-\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers.\\
-\hline
-\multicolumn{2}{l}{\textbf{Governance system (G)}}\\
-\hline
-\hspace{1em}GS6.1.4.3 - Territorial use communal rights : Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster.
-These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
-\hline
-\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandates.
-These include larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
-\hline
-\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group that monitors and enforces formal and internal rules.
-They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
-\hline
-\hspace{1em}GS9.2 - Biophysical monitoring: Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds.
-Recently, they have installed oceanographic sensors to monitor oceanographic variables.\\
-\hline
-GS10.1 - Graduated sanctions & Fishers have penalties for breaking collective-choice rules or fishing inside the reserves.
-These may range from scoldings and warnings to not being allowed to harvest a particular resource or being expelled from the cooperative.\\
-\hline
-\end{tabular}}
-\end{table}
-
-\end{document}
+---
+title: "Effectiveness of community-based TURF-reserves in small-scale fisheries"
+rtitle: "Community-based marine reserves"
+author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,2}$, Stuart Fulton$^{3}$, Álvin Suarez$^{3}$, Arturo Hernández-Velasco$^{3}$, Jorge Torre$^{3}$, Fiorenza Micheli$^{4}$"
+institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
+$^{2}$Nicholas School of the Environment, Duke University, Beaufort, NC, USA\\newline
+$^{3}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico\\newline
+$^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA"
+corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
+email: "juancarlos@ucsb.edu"
+abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
+Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage.
+Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
+In the last 12 years, 18 old and new community-based TURF-reserves gained legal recognition thanks to a 2014 regulation; their effectiveness has not been formally evaluated.
+We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
+We find that while reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates, they continue to receive significant support from the fishing communities.
+A lack of clear ecological and socioeconomic effects likely results from a combination of factors.
+First, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect.
+Second, some of the reserves are not large enough to protect mobile species, like lobster.
+Third, some of these reserves might be too young for the effects to show.
+Fourth, variable and extreme oceanographic conditions have impacted harvested populations.
+However, these reserves have shaped small-scale fishers’ way of thinking about marine conservation, which can provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
+keywords: "TURF-reserves, Causal Inference, Social-Ecological Systems, Marine Protected Areas, Marine Conservation, Small-Scale Fisheries"
+output:
+  bookdown::pdf_document2:
+    keep_tex: yes
+    template: docs/template.tex
+---

### 2018-12-03 docs/paper/Villasenor-Derbez.tex

@@
 \def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
-of California, Santa Barbara, Santa Barbara, CA,
-USA\newline \(^{2}\)Nicholas School of the Environment, Duke University,
-Beaufort, NC, USA\newline \(^{3}\)Comunidad y Biodiversidad A.C.,
-Guaymas, Sonora, Mexico\newline \(^{4}\)Hopkins Marine Station and
-Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
+of California, Santa Barbara, Santa Barbara, CA, USA \newline
+\(^{2}\)Nicholas School of the Environment, Duke University, Beaufort,
+NC, USA \newline \(^{3}\)Comunidad y Biodiversidad A.C., Guaymas,
+Sonora, Mexico \newline \(^{4}\)Hopkins Marine Station and Center for
+Ocean Solutions, Stanford University, Pacific Grove, CA, USA" \newline}
 % The Corresponding Author should be marked with an asterisk
@@
 
-\title[Community-based marine reserves]{Effectiveness of community-based marine reserves in small-scale
-fisheries}
+\title[Community-based marine reserves]{Effectiveness of community-based TURF-reserves in small-scale fisheries}
 
@@
 lack of clear ecological and socioeconomic effects likely results from a
-combination of factors. First, local fisheries are already well managed,
-and it is unlikely that reserves might have a detectable effect. Second,
-some of the reserves are not large enough to protect mobile species,
-like lobster. Third, some of these reserves might be too young for the
-effects to show. Fourth, variable and extreme oceanographic conditions
-have impacted harvested populations. However, these reserves have shaped
-small-scale fishers' way of thinking about marine conservation, which
-can provide a foundation for establishing additional, larger marine
-reserves needed to effectively conserve mobile species.
+combination of factors. First, some of these reserves might be too young
+for the effects to show. Second, the reserves are not large enough to
+protect mobile species, like lobster. Third, variable and extreme
+oceanographic conditions have impacted harvested populations. Fourth,
+local fisheries are already well managed, and it is unlikely that
+reserves might have a detectable effect in landings. However, these
+reserves have shaped small-scale fishers' way of thinking about marine
+conservation, which can provide a foundation for establishing
+additional, larger marine reserves needed to effectively conserve mobile
+species.
 
@@
 \end{abstract}
-
-Words in text: 3714 \textbar{} Words in headers: 35 \textbar{} Words
-outside text (captions, etc.): 192
-
-\clearpage
 

### 2018-12-03 slides/GainesLabPresentation.Rmd

@@
 
-zrp <- st_read(dsn = here::here("data", "spatial"), layer = "ZRP", quiet = T) %>% 
+zrp <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
   sf::st_centroid() %>% 
@@
 
-rescom <- st_read(dsn = here::here("data", "spatial"), layer = "Res_Com", quiet = T) %>% 
+rescom <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
   sf::st_centroid() %>% 
@@
   geom_sf(data = W) +
-  geom_sf(data = reserves, aes(fill = Type), shape = 21) +
+  geom_sf(data = reserves, shape = 21, fill = "red", color = "black") +
   theme_classic() +

### 2018-12-03 docs/paper/Villasenor-Derbez.Rmd
Started to add references related to #14

--- 2018-11-30 docs/paper/Villasenor-Derbez.Rmd
@@
 ph <- ggplot() +
-  geom_sf(data = turf[c(1,3),], fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
+  geom_sf(data = turf[3,], fill = "steelblue", alpha = 0.5, size = 1, color = "black") +
   geom_sf(data = coastline_mx) +
@@
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
-This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}.
-BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}.
-All sites were surveyed annually, and at least once before implementation of the reserves.
+This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{stewartoaten_1986,depalma_2018}.
 
@@
 ## Data analysis
+
+BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.
+All sites were surveyed annually, and at least once before implementation of the reserves.
+
+\{guidetti_2002,ferraro_2006}
 

### 2018-12-03 docs/paper/reviewer_comments.Rmd
Started to add references related to #14

@@
 
-- In the study area section more information would be useful on the wider context (some of this is in the introduction
-
-– e.g. the TURF system in Mexico and governance arrangements for this system and the reserves within it).
+- In the study area section more information would be useful on the wider context (some of this is in the introduction e.g.
+the TURF system in Mexico and governance arrangements for this system and the reserves within it).
 
@@
 - It would be worth discussing fully the limitations of the data e.g.
-presumably the lack of data on fishing effort restricts interpretation of the time series, and the indicators representing socio-economic benefits of the MPAs are limited
+presumably the lack of data on fishing effort restricts interpretation of the time series, and the indicators representing socio-economic benefits of the MPAs are limited.
 
@@
 Where would each study site fall along the gradient of underfished to overfished stocks?
-
-- By doing this you can assess each site in terms of biological performance and SES conditions.
+ By doing this you can assess each site in terms of biological performance and SES conditions.
 This then allows you to establish a stronger connection on how biological and social factors together act to contribute to better protection and effectiveness of marine reserves.
@@
 - Please provide the size of each of the reserves.
-
-- Reserve size is an important factor affecting ecological outcomes, especially when comparing comparatively mobile vs non-mobile species (as noted by the authors in the Discussion).
+Reserve size is an important factor affecting ecological outcomes, especially when comparing comparatively mobile vs non-mobile species (as noted by the authors in the Discussion).
 Why was the effect of reserve size not assessed?
@@
 
-- Table 2: Some variables need further clarification.
+#### Table 2: Some variables need further clarification.
 
@@
 What about market drivers that may have motivated increased fishing pressure?
-
-- Elaboration on these factors will strengthen the authors’ attempt to bring in the SES framework to this primarily biological analysis.
+Elaboration on these factors will strengthen the authors’ attempt to bring in the SES framework to this primarily biological analysis.
 

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
started working in the introduction towards #14 #17 #18

@@
 In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
-Recent research shows that combining Territorial Use Rights for Fisheries (TURFs) with no-take marine reserves (MRs) can greatly improve the performance of coastal fisheries and the health of the local resources \citep{costello_2010-Ix,lester_2017}.
-Commonly known as TURF-Reserves, these systems increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014-HP,lester_2017}.
-Although in theory these systems are successful \citep{costello_2010-Ix,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014-HP,lester_2017}.
+One of the many important ways to improve the performance of coastal fisheries and health of the local resources is through the implementation of no-take marine reserves (marine reserves from hereinafter) within Territorial Use Rights for Fisheries (TURFs) to create TURF-reserves \citep{costello_2010,afflerbach_2014,lester_2017}.
 
-The performance of these systems depends on how environmental and social factors combine and interact.
-The science of marine reserves has largely focused on understanding the ecological effects of these areas, which include increased biomass, species richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
-Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}.
-Other research has focused on the relationship between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}.
+TURFs are a way of fisheries management in which a well defined group of fishes have exclusive access to an explicitly delimited portion of the ocean.
+TURFs promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{mccay_2017}.
+On the other hand, marine reserves are areas where all extractive activities are off-limits.
+These can be implemented to protect biodiversity but also as fishery management tools that limit or restrict fishing effort and gears and therefore aid in the recovery of marine stocks.
+Commonly known as TURF-Reserves, the combination of these tools increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,lester_2017}.
+
+Research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access \citep{gelcich_2008,mccay_2014,mccay_2017}.
+Likewise, marine reserve science has shown that these lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+The benefits resulting from reserves established within TURFs may be captured by the group of fishers with exclusive access.
+Although in theory these systems are successful \citep{costello_2010,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014,lester_2017}.
+
+The performance of TURF-reserve systems must depend on how environmental and social factors combine and interact.
+This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}.
+There is a growing body of literature focusing on the relations between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}.
 However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.
-This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}.
+
+<!-- Have lead to the creation of decission support tools to guide the design and implementation of TURF-reserve systems\citep{oyanedel_2017}. -->
+
+<!-- Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}.
+-->
 
@@
 However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
+
+While many ecology studies incporporate BACI\citep{stewartoaten_1986}, few MPA studies do the same.
+Importance of evidence-based conservation mediterranean and as a movement \{guidetti_2002,ferraro_2006}.
+Whole paragraph here.
 
@@
 Isla Natividad, for example, is part of a greater group of fishing cooperatives belonging to a Federation of Fishing Cooperatives.
-This group has been identified as a cohesive group that often cooperates to better manage their resources \citep{mccay_2014-CN,mccay_2017-1m,acevesbueno_2017}.
+This group has been identified as a cohesive group that often cooperates to better manage their resources \citep{mccay_2014-CN,mccay_2017,acevesbueno_2017}.
 Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation.
@@
 Size structures are also collected during fish surveys.
-We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had a similar probability of being selected as reserves during the design phase.
+We define control sites as regions where: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, and iii) are located within the TURF and therefore fishing occurs.
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
@@
 Data contain monthly lobster landings (Kg) and revenues (MXP) for cooperatives with and without marine reserves.
-Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017-1m,ayer_2018}.
+Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017,ayer_2018}.
 Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.
@@
 All sites were surveyed annually, and at least once before implementation of the reserves.
-
-\{guidetti_2002,ferraro_2006}
-
 We evaluate the effect that marine reserves have had on four ecological and two socioeconomic indicators (Table \ref{table:indicators}).

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Expanded on TURFs and TURF-reserves close #17 close #18 close #21

@@
 In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
-One of the many important ways to improve the performance of coastal fisheries and health of the local resources is through the implementation of no-take marine reserves (marine reserves from hereinafter) within Territorial Use Rights for Fisheries (TURFs) to create TURF-reserves \citep{costello_2010,afflerbach_2014,lester_2017}.
+One of the many important ways to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves within them, thus creating TURF-reserve systems \citep{costello_2010,afflerbach_2014,lester_2017}.
 
-TURFs are a way of fisheries management in which a well defined group of fishes have exclusive access to an explicitly delimited portion of the ocean.
-TURFs promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{mccay_2017}.
-On the other hand, marine reserves are areas where all extractive activities are off-limits.
-These can be implemented to protect biodiversity but also as fishery management tools that limit or restrict fishing effort and gears and therefore aid in the recovery of marine stocks.
-Commonly known as TURF-Reserves, the combination of these tools increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,lester_2017}.
+TURFs are a fisheries management tool in which a well defined group of fishers have exclusive access to an explicitly delimited portion of the ocean.
+TURFs promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,mccay_2017}.
+On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits.
+These can be implemented to protect biodiversity but also as fishery management tools that restrict fishing effort and gears and therefore aid in the recovery of marine stocks.
+Commonly known as TURF-reserves, the combination of these tools increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,lester_2017}.
 
 Research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access \citep{gelcich_2008,mccay_2014,mccay_2017}.
-Likewise, marine reserve science has shown that these lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
-The benefits resulting from reserves established within TURFs may be captured by the group of fishers with exclusive access.
+Likewise, conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+The benefits resulting from reserves established within TURFs should be captured exclusively by the group of fishers with exclusive access.
 Although in theory these systems are successful \citep{costello_2010,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014,lester_2017}.
 
-The performance of TURF-reserve systems must depend on how environmental and social factors combine and interact.
+TURF-reserve systems are inherently intrincate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact.
 This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}.
@@
 However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.
+
+Moreover, a new Mexican norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}).
+Since 2012, old and new marine reserves have gained legal recognition as Fishing Refuges.
+Of these, 18 were originally implemented as community-based TURF-reserves.
+However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
+
+Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
+These three case studies span a range of ecological and social conditions representative of different regions of Mexico.
+The objective of this work is twofold.
+First, to provide a triple bottom line evaluation of the effectiveness of community-based TURF-reserves, which may inform similar processes in other countries.
+Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
+We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world where this tool is used to manage and rebuild their coastal fisheries.
 
@@
 
-TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities.
-This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8,smallhornwest_2018}.
-Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
-However, MRs are difficult to enforce if they are not legally recognized, and fishers rely on the exclusive access granted by the TURF.
-In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue.
-In Mexico, a new norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}).
-Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
-Since 2012, old and new marine reserves have gained legal recognition as Fishing Refuges.
-Oh these, 18 were originally implemented as community-based TURF-reserves.
-However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
-
-While many ecology studies incporporate BACI\citep{stewartoaten_1986}, few MPA studies do the same.
+<!-- While many ecology studies incporporate BACI\citep{stewartoaten_1986}, few MPA studies do the same.
 Importance of evidence-based conservation mediterranean and as a movement \{guidetti_2002,ferraro_2006}.
 Whole paragraph here.
-
-Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
-These three case studies span a range of ecological and social conditions representative of different regions of Mexico.
-The objective of this work is twofold.
-First, to provide a triple bottom line evaluation of the effectiveness of community-based marine reserves, which may inform similar processes in other countries.
-Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
-We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world where this tool is used to manage and rebuild their coastal fisheries.
+-->
 
 # Methods
+
+Before discussing our data collection methods and describing our analyses, our case studies warrant some more background.
+TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities.
+This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{gelcich_2015-Gw,beger_2004-Y8,smallhornwest_2018}.
+Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
+However, marine reserves are difficult to enforce if they are not legally recognized, and fishers rely on the exclusive access granted by the TURF.
+
+In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue (*i.e.* Fish Refuges; \citet{nom}).
+Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
+One of the ways in which fishing communities have takane advantage of this new tool is by implementing marine reserves within their TURFs.
+Our work focuses on some of these community-based TURF-reserves that occur in small-scale fisheries.
+
+The common setup of community-based TURF-reserves in Mexico is the following.
+Fishers from a given community are assambled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (*i.e.* TURFs shown as blue polygons in Fig \ref{fig:map}A).
+Each TURF is exlusively fished by one cooperative, and each community usually hosts no more than one cooperative.
+Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (*i.e.* TURF-reserves).
+Fishers then ask the government to grant legal recognition to their TURF-reserves under the name of Fish refuges following a series of studies outlined in \citet{nom}.
 

### 2018-12-04 docs/paper/Villasenor-Derbez.tex
Expanded on TURFs and TURF-reserves close #17 close #18 close #21

@@
 Sonora, Mexico \newline \(^{4}\)Hopkins Marine Station and Center for
-Ocean Solutions, Stanford University, Pacific Grove, CA, USA" \newline}
+Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
 % The Corresponding Author should be marked with an asterisk
@@
 small-scale fisheries face great challenges since they tend to be hard
-to monitor and enforce \citep{costello_2012}. Recent research shows that
-combining Territorial Use Rights for Fisheries (TURFs) with no-take
-marine reserves (MRs) can greatly improve the performance of coastal
-fisheries and the health of the local resources
-\citep{costello_2010-Ix,lester_2017}. Commonly known as TURF-Reserves,
-these systems increase the benefits of spatial access rights allowing
-the maintenance of healthy resources
-\citep{afflerbach_2014-HP,lester_2017}. Although in theory these systems
-are successful \citep{costello_2010-Ix,smallhornwest_2018}, there is
-little empirical evidence of their effectiveness and the drivers of
-their success \citep{afflerbach_2014-HP,lester_2017}.
+to monitor and enforce \citep{costello_2012}. One of the many important
+ways to improve the performance of coastal fisheries and health of the
+local resources is through the implementation of Territorial Use Rights
+for Fisheries (TURFs) that contain no-take marine reserves within them,
+thus creating TURF-reserve systems
+\citep{costello_2010,afflerbach_2014,lester_2017}.
 
-The performance of these systems depends on how environmental and social
-factors combine and interact. The science of marine reserves has largely
-focused on understanding the ecological effects of these areas, which
-include increased biomass, species richness, and densities of organisms
-within the protected regions, climate change mitigation, and protection
-from environmental variability
-\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}.
-Modelling studies show that fishery benefits of marine reserves depend
-on initial stock status and the management under which the fishery
-operates, as well as reserve size and the amount of larvae exported from
-these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}. Other research has
-focused on the relationship between socioeconomic and governance
-structures and reserve effectiveness
-\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, to our
-knowledge, no studies exist that evaluate TURF-reserves from both a
-social and ecological perspective. This is especially important in
+TURFs are a fisheries management tool in which a well defined group of
+fishers have exclusive access to an explicitly delimited portion of the
+ocean. TURFs promote a sense of stewardship and incentivise resource
+users to sustainably manage their resources
+\citep{gelcich_2008,mccay_2017}. On the other hand, no-take marine
+reserves (marine reserves from hereinafter) are areas where all
+extractive activities are off-limits. These can be implemented to
+protect biodiversity but also as fishery management tools that restrict
+fishing effort and gears and therefore aid in the recovery of marine
+stocks. Commonly known as TURF-reserves, the combination of these tools
+increase the benefits of spatial access rights allowing the maintenance
+of healthy resources \citep{afflerbach_2014,lester_2017}.
+
+Research on TURFs has shown that these areas have higher abundance of
+targeted species than sites operating under open access
+\citep{gelcich_2008,mccay_2014,mccay_2017}. Likewise, conservation
+science has shown how marine reserves lead to increased biomass, species
+richness, and abundance within the protected regions
+\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may
+have a series of additional benefits like climate change mitigation,
+protection from environmental variability, and fisheries benefits
+\cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. The benefits
+resulting from reserves established within TURFs are captured by the
+group of fishers with exclusive access. Although in theory these systems
+are successful \citep{costello_2010,smallhornwest_2018}, there is little
+empirical evidence of their effectiveness and the drivers of their
+success \citep{afflerbach_2014,lester_2017}.
+
+TURF-reserve systems are inherently intrincate social-ecological
+systems, and their effectiveness must depend on how environmental and
+social factors combine and interact. This is especially important in
 social-ecological coastal systems dominated by close interactions and
 feedbacks between people and natural resources \citep{ostrom_2009-hg}.
+There is a growing body of literature focusing on the relations between
+socioeconomic and governance structures and reserve effectiveness
+\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}.
+However, to our knowledge, no studies exist that evaluate TURF-reserves
+from both a social and ecological perspective.
 
-TURF-reserves can be created as community-based marine reserves,
-voluntarily established and enforced by local communities. This
-bottom-up approach increases compliance and self-enforcement, and
-reserves can yield benefits similar to systematically-designed reserves
-\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8,smallhornwest_2018}.
-Community-based spatial closures occur in different contexts, like the
-\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
-\citep{bohnsack_2004,johannes_2002}. However, MRs are difficult to
-enforce if they are not legally recognized, and fishers rely on the
-exclusive access granted by the TURF. In an effort to bridge this
-normative gap, Mexican Civil Society Organizations (CSOs) served as a
-link between fishers and government, and created a legal framework that
-solves this governance issue. In Mexico, a new norm was created in 2014
-allowing fishers to request the legal recognition of community-based
-reserves as ``Fish Refuges'' (\emph{Zona de Refugio Pesquero};
-\citet{nom}). Fish refuges can be implemented as temporal or partial
-reserves, which can protect one, some, or all resources within their
-boundaries. Since 2012, old and new marine reserves have gained legal
-recognition as Fishing Refuges. Oh these, 18 were originally implemented
-as community-based TURF-reserves. However, their effectiveness has not
-yet been formally evaluated and reported in the scientific literature.
+Moreover, a new Mexican norm was created in 2014 allowing fishers to
+request the legal recognition of community-based reserves as ``Fish
+Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since 2012,
+old and new marine reserves have gained legal recognition as Fishing
+Refuges. Of these, 18 were originally implemented as community-based
+TURF-reserves. However, their effectiveness has not yet been formally
+evaluated and reported in the scientific literature.
 
@@
 of this work is twofold. First, to provide a triple bottom line
-evaluation of the effectiveness of community-based marine reserves,
-which may inform similar processes in other countries. Second, to
-evaluate the effectiveness of TURF-reserves established as Fish Refuges
-in Mexico to identify opportunities where improvement or adjustment
-might lead to increased effectiveness. We draw from lessons learned in
-these three case studies and provide management recommendations to
-maximize the effectiveness of community-based marine reserves in
-small-scale fisheries in Mexico and in other regions around the world
-where this tool is used to manage and rebuild their coastal fisheries.
+evaluation of the effectiveness of community-based TURF-reserves, which
+may inform similar processes in other countries. Second, to evaluate the
+effectiveness of TURF-reserves established as Fish Refuges in Mexico to
+identify opportunities where improvement or adjustment might lead to
+increased effectiveness. We draw from lessons learned in these three
+case studies and provide management recommendations to maximize the
+effectiveness of community-based marine reserves in small-scale
+fisheries in Mexico and in other regions around the world where this
+tool is used to manage and rebuild their coastal fisheries.
 
@@
 \section{Methods}\label{methods}}
+
+TURF-reserves can be created as community-based marine reserves,
+voluntarily established and enforced by local communities. This
+bottom-up approach increases compliance and self-enforcement, and
+reserves can yield benefits similar to systematically-designed reserves
+\citep{gelcich_2015-Gw,beger_2004-Y8,smallhornwest_2018}.
+Community-based spatial closures occur in different contexts, like the
+\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
+\citep{bohnsack_2004,johannes_2002}. However, marine reserves are
+difficult to enforce if they are not legally recognized, and fishers
+rely on the exclusive access granted by the TURF.
+
+In an effort to bridge this normative gap, Mexican Civil Society
+Organizations (CSOs) served as a link between fishers and government,
+and created a legal framework that solves this governance issue
+(\emph{i.e.} Fish Refuges; \citet{nom}). Fish refuges can be implemented
+as temporal or partial reserves, which can protect one, some, or all
+resources within their boundaries. One of the ways in which fishing
+communities have takane advantage of this new tool is by implementing
+marine reserves within their TURFs.
+
+The common setup of community-based TURF-reserves in mexico is the
+following. Fishers from a given community are assambled in fishing
+cooperatives which have exclusive fishing rights over a spatially
+delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
+\ref{fig:map}A). Each TURF is exlusively fished by one cooperative, and
+each community usually hosts no more than one cooperative. Fishing
+cooperatives interested in implementing marine reserves work with CSOs
+to implement marine reserves within their TURFs (\emph{i.e.}
+TURF-reserves). Fishers then ask the government to grant legal
+recognition to their TURF-reserves under the name of Fish refuges
+following a series of studies outlined in \citet{nom}.
 
@@
 a cohesive group that often cooperates to better manage their resources
-\citep{mccay_2014-CN,mccay_2017-1m,acevesbueno_2017}. Likewise, Maria
-Elena and Punta Herrero are representative of fishing cooperatives in
-the Mexican Caribbean, which are also part of a regional Federation.
+\citep{mccay_2014-CN,mccay_2017,acevesbueno_2017}. Likewise, Maria Elena
+and Punta Herrero are representative of fishing cooperatives in the
+Mexican Caribbean, which are also part of a regional Federation.
 Together, these three communities provide an accurate representation of
@@
 also collected during fish surveys. We define control sites as regions
-with habitat characteristics similar to the corresponding reserves, and
-that presumably had a similar probability of being selected as reserves
-during the design phase. We focus our evaluation on sites where data are
-available for reserve and control sites, before and after the
+where: i) habitat characteristics are similar to the corresponding
+reserves, ii) presumably had a similar probability of being selected as
+reserves during the design phase, and iii) are located within the TURF
+and therefore fishing occurs. We focus our evaluation on sites where
+data are available for reserve and control sites, before and after the
 implementation of the reserve. This provides us with a
@@
 allows us to capture and control for temporal and spatial dynamics
-\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
-techniques have proven effective to evaluate marine reserves, as they
-allow us to causally attribute observed changes to the intervention
-\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
-annually, and at least once before implementation of the reserves.
+\citep{stewartoaten_1986,depalma_2018}.
 
@@
 markets and institutional frameworks, making them plausible controls
-\citep{mccay_2017-1m,ayer_2018}. Landings and revenues were aggregated
-at the cooperative-year level, and revenues were adjusted to represent
-2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.
+\citep{mccay_2017,ayer_2018}. Landings and revenues were aggregated at
+the cooperative-year level, and revenues were adjusted to represent 2014
+values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.
 
@@
 
-We evaluate the effect that marine reserves have had on four ecological
-and two socioeconomic indicators (Table \ref{table:indicators}). Recall
-that reserves were implemented to protect lobster and other benthic
-invertebrates. However, we also use the available fish data to test for
-associated co-benefits.
+BACI designs and causal inference techniques have proven effective to
+evaluate marine reserves, as they allow us to causally attribute
+observed changes to the intervention
+\citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}. All sites
+were surveyed annually, and at least once before implementation of the
+reserves. We evaluate the effect that marine reserves have had on four
+ecological and two socioeconomic indicators (Table
+\ref{table:indicators}). Recall that reserves were implemented to
+protect lobster and other benthic invertebrates. However, we also use
+the available fish data to test for associated co-benefits.
 

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Updated running title

@@
 title: "Effectiveness of community-based TURF-reserves in small-scale fisheries"
-rtitle: "Community-based marine reserves"
+rtitle: "Community-based TURF-reserves"
 author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,2}$, Stuart Fulton$^{3}$, Álvin Suarez$^{3}$, Arturo Hernández-Velasco$^{3}$, Jorge Torre$^{3}$, Fiorenza Micheli$^{4}$"

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Added some caveats to the TURF-reserve paragraphs in the methods

@@
 TURFs are a fisheries management tool in which a well defined group of fishers have exclusive access to an explicitly delimited portion of the ocean.
-TURFs promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,mccay_2017}.
+They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,mccay_2017}.
 On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits.
 These can be implemented to protect biodiversity but also as fishery management tools that restrict fishing effort and gears and therefore aid in the recovery of marine stocks.
-Commonly known as TURF-reserves, the combination of these tools increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,lester_2017}.
+Commonly known as TURF-reserves, the combination of these tools should in theory increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,lester_2017}.
 
 Research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access \citep{gelcich_2008,mccay_2014,mccay_2017}.
-Likewise, conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+Likewise, conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009,giakoumi_2017-V2,sala_2017-69}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
 The benefits resulting from reserves established within TURFs should be captured exclusively by the group of fishers with exclusive access.

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Fixed text

@@
 
-<!-- While many ecology studies incporporate BACI\citep{stewartoaten_1986}, few MPA studies do the same.
-Importance of evidence-based conservation mediterranean and as a movement \{guidetti_2002,ferraro_2006}.
-Whole paragraph here.
--->
-
 # Methods
@@
 TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities.
-This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{gelcich_2015-Gw,beger_2004-Y8,smallhornwest_2018}.
+This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{gelcich_2015,beger_2004,smallhornwest_2018}.
 Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Added more references and expanded discussion to close #14

@@
 The first approach does not address temporal variability, and the second can not distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}.
+While many ecology studies have used BACI sampling designs and respective analyses (*e.g.* \citep{stewartoaten_1986}), few conservations studies have done so to evaluate the effect of an intervention (*i.e.* \citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted in a call for more robust analyses in conservation science \citep{guidetti_2002,ferraro_2006}.
+
 Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness.

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Comments in #19 have been adressed by explaining how TURF-reserves are similar to CBMRs, expanding on the rights. close #19

@@
 These can be implemented to protect biodiversity but also as fishery management tools that restrict fishing effort and gears and therefore aid in the recovery of marine stocks.
-Commonly known as TURF-reserves, the combination of these tools should in theory increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,lester_2017}.
+Commonly known as TURF-reserves, the combination of these tools should in theory increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
@@
 
-TURF-reserve systems are inherently intrincate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact.
+TURF-reserve systems are inherently intrincate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{gelcich_2015}.
 This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}.
@@
 Before discussing our data collection methods and describing our analyses, our case studies warrant some more background.
-TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities.
-This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{gelcich_2015,beger_2004,smallhornwest_2018}.
+Community-based marine reserves that are implemented within TURFs are a form of TURF-reserves, voluntarily established and enforced by local communities.
+This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{beger_2004,gelcich_2015,smallhornwest_2018}.
 Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.

### 2018-12-04 docs/paper/Villasenor-Derbez.tex
Comments in #19 have been adressed by explaining how TURF-reserves are similar to CBMRs, expanding on the rights. close #19

@@
 
-\title[Community-based marine reserves]{Effectiveness of community-based TURF-reserves in small-scale fisheries}
+\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in small-scale fisheries}
 
@@
 fishers have exclusive access to an explicitly delimited portion of the
-ocean. TURFs promote a sense of stewardship and incentivise resource
+ocean. They promote a sense of stewardship and incentivise resource
 users to sustainably manage their resources
@@
 stocks. Commonly known as TURF-reserves, the combination of these tools
-increase the benefits of spatial access rights allowing the maintenance
-of healthy resources \citep{afflerbach_2014,lester_2017}.
+should in theory increase the benefits of spatial access rights allowing
+the maintenance of healthy resources
+\citep{afflerbach_2014,lester_2017}.
 
@@
 \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. The benefits
-resulting from reserves established within TURFs are captured by the
-group of fishers with exclusive access. Although in theory these systems
-are successful \citep{costello_2010,smallhornwest_2018}, there is little
-empirical evidence of their effectiveness and the drivers of their
-success \citep{afflerbach_2014,lester_2017}.
+resulting from reserves established within TURFs should be captured
+exclusively by the group of fishers with exclusive access. Although in
+theory these systems are successful
+\citep{costello_2010,smallhornwest_2018}, there is little empirical
+evidence of their effectiveness and the drivers of their success
+\citep{afflerbach_2014,lester_2017}.
 
@@
 
-TURF-reserves can be created as community-based marine reserves,
-voluntarily established and enforced by local communities. This
-bottom-up approach increases compliance and self-enforcement, and
-reserves can yield benefits similar to systematically-designed reserves
-\citep{gelcich_2015-Gw,beger_2004-Y8,smallhornwest_2018}.
-Community-based spatial closures occur in different contexts, like the
-\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
+Before discussing our data collection methods and describing our
+analyses, our case studies warrant some more background. TURF-reserves
+can be created as community-based marine reserves, voluntarily
+established and enforced by local communities. This bottom-up approach
+increases compliance and self-enforcement, and reserves can yield
+benefits similar to systematically-designed reserves
+\citep{gelcich_2015,beger_2004,smallhornwest_2018}. Community-based
+spatial closures occur in different contexts, like the \emph{kapu} or
+\emph{ra'ui} areas in the Pacific Islands
 \citep{bohnsack_2004,johannes_2002}. However, marine reserves are
@@
 communities have takane advantage of this new tool is by implementing
-marine reserves within their TURFs.
+marine reserves within their TURFs. Our work focuses on some of these
+community-based TURF-reserves that occur in small-scale fisheries.
 
-The common setup of community-based TURF-reserves in mexico is the
+The common setup of community-based TURF-reserves in Mexico is the
 following. Fishers from a given community are assambled in fishing

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Deleted speculation from abstract, close #16

@@
 Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in landings.
-However, these reserves have shaped small-scale fishers’ way of thinking about marine conservation, which can provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
+However, these reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
 keywords: "TURF-reserves, Causal Inference, Social-Ecological Systems, Marine Protected Areas, Marine Conservation, Small-Scale Fisheries"
@@
 
+The fact that these reserves were re-established and just enacted shows a comitment of support (issue \#16 on GH).
+
 ```{r}

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Added a paragraph explaining why we do each analysis to close #24 and similar to #32

@@
 TURF-reserve systems are inherently intrincate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{gelcich_2015}.
-This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}.
+This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009}.
 There is a growing body of literature focusing on the relations between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}.
@@
 
+## TURF-reserves in Mexico
+
 Before discussing our data collection methods and describing our analyses, our case studies warrant some more background.
@@
 
-## Study area
+## Study areas
 
@@
 Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our insights to other similar reserves in Mexico and elsewhere around the world.
+
+The regulation governing the implementation of Fish refuges states that these are fishery management tools intended to have biological or socioeconomic benefits \citep{nom}.
+For this reason, the main portion of our analyses focuses on a series of biological and socioeconomic indicators that may respond to reserve implementation.
+However, the effectiveness of conservation and fisheries management interventions also depends on the social and governance structures in place.
+We therefore incorporate a reduced version of the Social Ecological Systems framework \citep{ostrom_2009} and evaluate variables and indicators known to aid and hinder the effectiveness of management interventions in conservation and fisheries.
+The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather provide a discussion of potential causes of individual reserve performance.
+The following two sections describe our data collection methods and analyses of biological and socioeconomic indicators as well as the SES analysis.
 

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Added a brief text that will close #31
Clarified that fishing occurs only outside the reserves to close #34

@@
 
-We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A).
+We evaluate three community-based no-take marine reserves implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig \ref{fig:map}A).
 The first one was created by the *Buzos y Pescadores de la Baja California* fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig \ref{fig:map}B).
@@
 The fact that these reserves were re-established and just enacted shows a comitment of support (issue \#16 on GH).
+the SES provides a static picture of governance and resource conditions, whereas the biological assessment reflects changes through time (Isse \#31).
 

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Specified that all TURFs were there before reserve implementation to close #35
Changed tripple bottom or hollistic and explained what it means to close #36

@@
 
-Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
+Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico.
 These three case studies span a range of ecological and social conditions representative of different regions of Mexico.
 The objective of this work is twofold.
-First, to provide a triple bottom line evaluation of the effectiveness of community-based TURF-reserves, which may inform similar processes in other countries.
+First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries.
 Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
@@
 Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation.
-Together, these three communities provide an accurate representation of other fishing communities in each of their regions.
+Together, these three communities provide an accurate representation of other fishing communities that have been historically manged with TURFs in each of their regions.
 While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth causal inference analysis that we undertake.

### 2018-12-04 docs/Villasenor-Derbez.Rmd
Moved all manuscript files one level up into docs

--- 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
@@
 All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}.
-All analyses were performed in `r version$version.string` and R Studio version 1.1.456 \citep{R_2018}.
+All analyses were performed in R `r paste0("version ", version$major, ".", version$minor)` and R Studio version 1.1.456 \citep{R_2018}.
+All data and code ara available in a [GitHub repository](https://github.com/jcvdav/ReserveEffect).
 

### 2018-12-04 docs/Villasenor-Derbez_SupplementaryMaterial.Rmd
Moved all manuscript files one level up into docs

---
title: "Effectiveness of community-based marine reserves in small-scale fisheries"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template_sup.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

options(knitr.table.format = "latex")
```

```{r}
suppressPackageStartupMessages({
  library(magrittr)
  library(tidyverse)
})
```

```{r}
load(here::here("data", "results", "results.RData"))
```

```{r}
supp_regtable <- function(models, caption = "", fs = "tiny", bio = T){

  labels <- c(
    "Lobster abundance",
    "Fish biomass",
    "Invertebrate abundance",
    "Invertebrate biomass")

  if(!bio){
    labels <- c(
      "Landings",
      "Revenues")
  }

  stargazer::stargazer(models,
                     omit = c(
                       "generoespecie",
                       "unidad"),
                     se = commarobust::makerobustseslist(models, type = "HC1"),
                     t.auto = T,
                     p.auto = T,
                     column.labels = labels,
                     dep.var.labels = "",
                     single.row = T,
                     omit.stat = c("f", "adj.rsq"),
                     title = caption,
                     header = F,
                     font.size = fs,
                     column.sep.width = "1pt")
}
```

```{r, results = "asis"}
list(L_IN, B_IN, Ni_IN, N_IN) %>% 
  supp_regtable(caption = "Coefficient estimates of biological indicators for Isla Natividad.")
```

```{r, results = "asis"}
list(L_ME, B_ME, Ni_ME, N_ME) %>% 
  supp_regtable(caption = "Coefficient estimates of biological indicators for Maria Elena.")
```

```{r, results = "asis"}
list(L_PH, B_PH, Ni_PH, N_PH) %>% 
  supp_regtable(caption = "Coefficient estimates of biological indicators for Punta Herrero.", fs = "small")
```

```{r, results = "asis"}
list(C_IN, V_IN) %>% 
  supp_regtable(caption = "Coefficient estimates of socioeconomic indicators Isla Natividad.", bio = F)
```

```{r, results = "asis"}
list(C_ME, V_ME) %>% 
  supp_regtable(caption = "Coefficient estimates of socioeconomic indicators for Maria Elena.", bio = F)
```

### 2018-12-04 docs/Villasenor-Derbez.Rmd
Deleted unsupported part of sentence to close #12

@@
 In 2006, the community decided to implement two marine reserves within their fishing grounds to protect commercially important invertebrate species; mainly lobster and abalone.
-While these reserves obtained legal recognition only in 2018 \citep{dof_website_2018}, they have been well enforced since their implementation.
+These reserves obtained legal recognition only in 2018 \citep{dof_website_2018}.
 
@@
 Isla Natividad, for example, is part of a greater group of fishing cooperatives belonging to a Federation of Fishing Cooperatives.
-This group has been identified as a cohesive group that often cooperates to better manage their resources \citep{mccay_2014-CN,mccay_2017,acevesbueno_2017}.
+This group has been identified as a cohesive group that often cooperates to better manage their resources \citep{mccay_2014,mccay_2017,acevesbueno_2017}.
 Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation.
@@
 We use a difference-in-differences analysis to evaluate these indicators.
-This approach allows us to estimate the effect that the reserve had by comparing trends across time and treatments \citep{moland_2013-VP,Villasenor-Derbez_2018}.
+This approach allows us to estimate the effect that the reserve had by comparing trends across time and treatments \citep{moland_2013,Villasenor-Derbez_2018}.
 The analysis of ecological indicators is performed with a multiple linear regression of the form:

### 2018-12-04 docs/Villasenor-Derbez.tex
Deleted unsupported part of sentence to close #12

--- 2018-12-04 docs/paper/Villasenor-Derbez.tex
@@
 reserves might have a detectable effect in landings. However, these
-reserves have shaped small-scale fishers' way of thinking about marine
-conservation, which can provide a foundation for establishing
-additional, larger marine reserves needed to effectively conserve mobile
-species.
+reserves may provide a foundation for establishing additional, larger
+marine reserves needed to effectively conserve mobile species.
 
@@
 the maintenance of healthy resources
-\citep{afflerbach_2014,lester_2017}.
+\citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
@@
 richness, and abundance within the protected regions
-\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may
+\citep{lester_2009,giakoumi_2017-V2,sala_2017-69}, and that these may
 have a series of additional benefits like climate change mitigation,
@@
 systems, and their effectiveness must depend on how environmental and
-social factors combine and interact. This is especially important in
-social-ecological coastal systems dominated by close interactions and
-feedbacks between people and natural resources \citep{ostrom_2009-hg}.
-There is a growing body of literature focusing on the relations between
-socioeconomic and governance structures and reserve effectiveness
+social factors combine and interact \citep{gelcich_2015}. This is
+especially important in social-ecological coastal systems dominated by
+close interactions and feedbacks between people and natural resources
+\citep{ostrom_2009}. There is a growing body of literature focusing on
+the relations between socioeconomic and governance structures and
+reserve effectiveness
 \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}.
@@
 Here, we combine causal inference techniques and the Social-Ecological
-Systems (SES) framework to provide a holistic evaluation of
-community-based TURF-reserves in three coastal communities in Mexico.
-These three case studies span a range of ecological and social
-conditions representative of different regions of Mexico. The objective
-of this work is twofold. First, to provide a triple bottom line
-evaluation of the effectiveness of community-based TURF-reserves, which
+Systems (SES) framework to evaluate community-based TURF-reserves in
+three coastal communities in Mexico. These three case studies span a
+range of ecological and social conditions representative of different
+regions of Mexico. The objective of this work is twofold. First, to
+provide a holistic evaluation of the effectiveness of community-based
+TURF-reserves in terms of the changes in biological and socioeconomic
+indicators and the governance settings under which these develop, which
 may inform similar processes in other countries. Second, to evaluate the
@@
 
+\hypertarget{turf-reserves-in-mexico}{%
+\subsection{TURF-reserves in Mexico}\label{turf-reserves-in-mexico}}
+
 Before discussing our data collection methods and describing our
-analyses, our case studies warrant some more background. TURF-reserves
-can be created as community-based marine reserves, voluntarily
-established and enforced by local communities. This bottom-up approach
-increases compliance and self-enforcement, and reserves can yield
-benefits similar to systematically-designed reserves
-\citep{gelcich_2015,beger_2004,smallhornwest_2018}. Community-based
+analyses, our case studies warrant some more background. Community-based
+marine reserves that are implemented within TURFs are a form of
+TURF-reserves, voluntarily established and enforced by local
+communities. This bottom-up approach increases compliance and
+self-enforcement, and reserves can yield benefits similar to
+systematically-designed reserves
+\citep{beger_2004,gelcich_2015,smallhornwest_2018}. Community-based
 spatial closures occur in different contexts, like the \emph{kapu} or
@@
 
-\hypertarget{study-area}{%
-\subsection{Study area}\label{study-area}}
+\hypertarget{study-areas}{%
+\subsection{Study areas}\label{study-areas}}
 
-We evaluate three TURF-reserves in Mexico (Fig \ref{fig:map}A). The
-first one was created by the \emph{Buzos y Pescadores de la Baja
-California} fishing cooperative, located in Isla Natividad in the Baja
-California Peninsula (Fig \ref{fig:map}B). The main fishery in the
-island is the spiny lobster (\emph{Panulirus interruptus}), but other
-resources like finfish, sea cucumber, read sea urchin, snail, and
-abalone are also an important source of income. In 2006, the community
-decided to implement two marine reserves within their fishing grounds to
-protect commercially important invertebrate species; mainly lobster and
-abalone. While these reserves obtained legal recognition only in 2018
-\citep{dof_website_2018}, they have been well enforced since their
-implementation.
+We evaluate three community-based no-take marine reserves implemented in
+Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig
+\ref{fig:map}A). The first one was created by the \emph{Buzos y
+Pescadores de la Baja California} fishing cooperative, located in Isla
+Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The
+main fishery in the island is the spiny lobster (\emph{Panulirus
+interruptus}), but other resources like finfish, sea cucumber, read sea
+urchin, snail, and abalone are also an important source of income. In
+2006, the community decided to implement two marine reserves within
+their fishing grounds to protect commercially important invertebrate
+species; mainly lobster and abalone. These reserves obtained legal
+recognition only in 2018 \citep{dof_website_2018}.
 
@@
 a cohesive group that often cooperates to better manage their resources
-\citep{mccay_2014-CN,mccay_2017,acevesbueno_2017}. Likewise, Maria Elena
+\citep{mccay_2014,mccay_2017,acevesbueno_2017}. Likewise, Maria Elena
 and Punta Herrero are representative of fishing cooperatives in the
@@
 Together, these three communities provide an accurate representation of
-other fishing communities in each of their regions. While each region
-has additional communities that have established community-based
-TURF-reserves, available data would not allow us to perform the in-depth
-causal inference analysis that we undertake. Yet, given the similarities
-among communities and the socioeconomic and governance setting under
-which they operate, it is safe to cautiously generalize our insights to
-other similar reserves in Mexico and elsewhere around the world.
+other fishing communities that have been historically manged with TURFs
+in each of their regions. While each region has additional communities
+that have established community-based TURF-reserves, available data
+would not allow us to perform the in-depth causal inference analysis
+that we undertake. Yet, given the similarities among communities and the
+socioeconomic and governance setting under which they operate, it is
+safe to cautiously generalize our insights to other similar reserves in
+Mexico and elsewhere around the world.
+
+The regulation governing the implementation of Fish refuges states that
+these are fishery management tools intended to have biological or
+socioeconomic benefits \citep{nom}. For this reason, the main portion of
+our analyses focuses on a series of biological and socioeconomic
+indicators that may respond to reserve implementation. However, the
+effectiveness of conservation and fisheries management interventions
+also depends on the social and governance structures in place. We
+therefore incorporate a reduced version of the Social Ecological Systems
+framework \citep{ostrom_2009} and evaluate variables and indicators
+known to aid and hinder the effectiveness of management interventions in
+conservation and fisheries. The incorporation of the SES is not intended
+to relate different levels of governance with reserve effectiveness, but
+rather provide a discussion of potential causes of individual reserve
+performance. The following two sections describe our data collection
+methods and analyses of biological and socioeconomic indicators as well
+as the SES analysis.
 
@@
 reserve had by comparing trends across time and treatments
-\citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis of
-ecological indicators is performed with a multiple linear regression of
-the form:
+\citep{moland_2013,Villasenor-Derbez_2018}. The analysis of ecological
+indicators is performed with a multiple linear regression of the form:
 
@@
 heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
-analyses were performed in R version 3.5.1 (2018-07-02) and R Studio
-version 1.1.456 \citep{R_2018}.
+analyses were performed in R version 3.5.1 and R Studio version 1.1.456
+\citep{R_2018}. All data and code ara available in a
+\href{https://github.com/jcvdav/ReserveEffect}{GitHub repository}.
 
@@
 
+The fact that these reserves were re-established and just enacted shows
+a comitment of support (issue \#16 on GH). the SES provides a static
+picture of governance and resource conditions, whereas the biological
+assessment reflects changes through time (Isse \#31).
+
 \hypertarget{discussion}{%
@@
 between the temporal trends in a reserve and the entire system
-\citep{depalma_2018}. Our approach to evaluate the temporal and spatial
-changes provides a more robust measure of reserve effectiveness. For
-example, we capture previously described patterns like the rapid
-increase observed for lobster densities in Isla Natividad on the sixth
-year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), a year after the
-hypoxia events described by \citet{micheli_2012-EU}, which caused mass
-mortality of sedentary organisms such as abalone and sea urchins, but
-not lobster and finfish. Yet, our empirical approach assumes control
-sites are a plausible counterfactual for treated sites. This implies
-that treated sites would have followed the same trend as control sites,
-had the reserves not been implemented. Nonetheless, temporal trends for
-each site don't show any significant increases (S1 Table, S2 Table, S3
-Table), supporting our findings of lack of change in the indicators
-used.
+\citep{depalma_2018}. While many ecology studies have used BACI sampling
+designs and respective analyses (\emph{e.g.} \citep{stewartoaten_1986}),
+few conservations studies have done so to evaluate the effect of an
+intervention (\emph{i.e.}
+\citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted
+in a call for more robust analyses in conservation science
+\citep{guidetti_2002,ferraro_2006}.
+
+Our approach to evaluate the temporal and spatial changes provides a
+more robust measure of reserve effectiveness. For example, we capture
+previously described patterns like the rapid increase observed for
+lobster densities in Isla Natividad on the sixth year (\emph{i.e.} 2012;
+Fig. \ref{fig:indicators}A), a year after the hypoxia events described
+by \citet{micheli_2012-EU}, which caused mass mortality of sedentary
+organisms such as abalone and sea urchins, but not lobster and finfish.
+Yet, our empirical approach assumes control sites are a plausible
+counterfactual for treated sites. This implies that treated sites would
+have followed the same trend as control sites, had the reserves not been
+implemented. Nonetheless, temporal trends for each site don't show any
+significant increases (S1 Table, S2 Table, S3 Table), supporting our
+findings of lack of change in the indicators used.
 

### 2018-12-04 docs/Villasenor-Derbez.Rmd
Made more clear what data was collected and tested and expanded on the model to close #13

@@
 We use three main sources of information to evaluate these reserves across the ecological, socioeconomic, and governance dimensions.
-Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
-Trained divers record richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \citep{fulton_2018,fulton_2019,suman_2010-ez}.
-Size structures are also collected during fish surveys.
-We define control sites as regions where: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, and iii) are located within the TURF and therefore fishing occurs.
+Ecological data come from the annual ecological monitoring of reserve and control sites.
+Reserve sites are those withing the reserves, and thus no fishing takes place.
+We define control sites as areas where: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, and iii) are located within the TURF, where fishing occurs.
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
@@
 
+The biological data are collectged by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
+Trained divers record richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \citep{fulton_2018,fulton_2019,suman_2010-ez}.
+Size structures are also collected during fish surveys.
+
 Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA).
 Data contain monthly lobster landings (Kg) and revenues (MXP) for cooperatives with and without marine reserves.
+In this case our treated unit are the cooperatives (*i.e.* communities) that have implemented a reserve within their TURF, and the controls are nearby communities that have a TURF but did not implement a reserve.
 Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017,ayer_2018}.
-Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.
+Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017}.
 
@@
 All sites were surveyed annually, and at least once before implementation of the reserves.
-We evaluate the effect that marine reserves have had on four ecological and two socioeconomic indicators (Table \ref{table:indicators}).
+We evaluate the effect that marine reserves have had on four ecological and two socioeconomic indicators shown in Table \ref{table:indicators}.
 Recall that reserves were implemented to protect lobster and other benthic invertebrates.
-However, we also use the available fish data to test for associated co-benefits.
+However, we also use the available fish and invertebrate data to test for associated co-benefits.
 

### 2018-12-04 docs/Villasenor-Derbez.Rmd
Made more clear what data was collected and tested and expanded on the model to close #13

@@
 All sites were surveyed annually, and at least once before implementation of the reserves.
-We evaluate the effect that marine reserves have had on four ecological and two socioeconomic indicators shown in Table \ref{table:indicators}.
+We evaluate the effect that the TURF-reserves have had on four ecological and two socioeconomic indicators shown in Table \ref{table:indicators}.
 Recall that reserves were implemented to protect lobster and other benthic invertebrates.
@@
 This approach allows us to estimate the effect that the reserve had by comparing trends across time and treatments \citep{moland_2013,Villasenor-Derbez_2018}.
-The analysis of ecological indicators is performed with a multiple linear regression of the form:
+To perform difference-in-differences, we regress the indicator of interest on a dummy variable for treatment, a dummy variable for years, and the interaction term between these with a multiple linear regression of the form:
 

### 2018-12-05 docs/Villasenor-Derbez_SupplementaryMaterial.Rmd
stated that baci have been used for MRs before.And pointed to information on sampling effort per #23

@@
 
-```{r setup, include=FALSE}
+```{r setup, include = FALSE}
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
@@
 })
+```
+
+# Data
+
+```{r load_data}
+invert <- read.csv(file = here::here("data", "invertebrates.csv"),
+                stringsAsFactors = F)
+
+fish <- read.csv(file = here::here("data", "fish.csv"),
+                stringsAsFactors = F)
+
+conapesca <- read.csv(file = here::here("data", "conapesca.csv"),
+                stringsAsFactors = F)
+```
+
+```{r}
+invert %>% 
+  group_by(ano, comunidad, zona, sitio, transecto) %>%
+  count() %>%
+  group_by(comunidad, zona) %>%
+  count() %>% 
+  spread(zona, nn) %>% 
+  knitr::kable(caption = "Number of invertebrate transects performed in each site of each community.",
+               col.names = c("Community", "Control", "Reserve"))
+```
+
+```{r}
+invert %>%
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 
+  summarize(indicador = sum(abundancia)/60) %>% 
+  ungroup() %>% 
+  mutate(year = as.numeric(year)) %>% 
+  ggplot(aes(x = year, y = indicador, color = zona, fill = zona, group)) +
+  stat_summary(geom = "ribbon", fun.data = mean_se, alpha = 0.5) +
+  stat_summary(geom = "line", fun.y = mean) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1)
+```
+
+```{r}
+fish %>% 
+  group_by(ano, comunidad, zona, sitio, transecto) %>%
+  count() %>%
+  group_by(comunidad, zona) %>%
+  count() %>% 
+  spread(zona, nn) %>% 
+  knitr::kable(caption = "Number of fish transects performed in each site of each community.",
+               col.names = c("Community", "Control", "Reserve"))
+```
+
+```{r}
+fish %>%
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 
+  summarize(indicador = sum(abundancia)/60) %>% 
+  ungroup() %>% 
+  mutate(year = as.numeric(year)) %>% 
+  ggplot(aes(x = year, y = indicador, color = zona, fill = zona, group)) +
+  stat_summary(geom = "ribbon", fun.data = mean_se, alpha = 0.5) +
+  stat_summary(geom = "line", fun.y = mean) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1)
 ```

### 2018-12-05 docs/Villasenor-Derbez.Rmd
added information about where to find effort data per #23

@@
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
-This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{stewartoaten_1986,depalma_2018}.
+This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally attribute the changes to the reserve \citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.
 
@@
 Size structures are also collected during fish surveys.
+All sites were surveyed annually, and at least once before implementation of the reserves.
+A summary of sampling effort (number of transects per site) is shown in the supplementary materials (S1 Table and S2 Table).
 
@@
 
-BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.
-All sites were surveyed annually, and at least once before implementation of the reserves.
 We evaluate the effect that the TURF-reserves have had on four ecological and two socioeconomic indicators shown in Table \ref{table:indicators}.
@@
 We use a difference-in-differences analysis to evaluate these indicators.
-This approach allows us to estimate the effect that the reserve had by comparing trends across time and treatments \citep{moland_2013,Villasenor-Derbez_2018}.
+This approach is widely used in econometric literature to estimate the average treatment effect of an intervention \citep{card_1994,kiel_1995}.
+In our case it allows us to estimate the effect that the reserve had by comparing trends across time and treatments \citep{moland_2013,Villasenor-Derbez_2018}.
 To perform difference-in-differences, we regress the indicator of interest on a dummy variable for treatment, a dummy variable for years, and the interaction term between these with a multiple linear regression of the form:
@@
 The following sections present the effect that marine reserves had on each of the biological and socioeconomic indicators for each coastal community.
-Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* effect size $\lambda_t$).
+Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* the difference-in-differences estimate or effect size $\lambda_t$ from \ref{eqn:reg_bio}).
 We also provide an overview of the governance settings of each community, and discuss how these might be related to the effectiveness and performance of the reserves.
@@
 \ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
-Full tables with model coefficients are presented in the supplementary materials (S1 Table, S2 Table, S3 Table).
+Full tables with model coefficients are presented in the supplementary materials **(S1 Table, S2 Table, S3 Table)**.
 

### 2018-12-05 docs/Villasenor-Derbez.Rmd
Fixed figure caption close #56

@@
 
-```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for lobster densities (*Panulirus spp*; A), fish biomass (B), invertebrate densities (C), and fish densities (D).
+```{r, fig.width = 6.5, fig.height = 4, fig.cap = "\\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for lobster densities (*Panulirus spp*; A), fish biomass (B), invertebrate densities (C), and fish densities (D).
 Plots are ordered by survey type (left column: invertebrates; right column: fish).

### 2018-12-05 docs/Villasenor-Derbez.Rmd
Included mention of S3 Table and expanded explanation of Dif in Dif towards #23 #40

@@
 Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017}.
+A table with summary statistics for this data is provided in the supplementary materials **(S3 Table)**.
 
@@
 We use a difference-in-differences analysis to evaluate these indicators.
-This approach is widely used in econometric literature to estimate the average treatment effect of an intervention \citep{card_1994,kiel_1995}.
-In our case it allows us to estimate the effect that the reserve had by comparing trends across time and treatments \citep{moland_2013,Villasenor-Derbez_2018}.
+This approach is widely used in econometric literature to estimate the average treatment effect of an intervention, like the impact of minimum wage increases on employment rates \citep{card_1994}.
+In our case it allows us to estimate the effect that the reserve had on each biological and socioeconomic indicator (Table \ref{table:indicators}) by comparing trends across time and treatments \citep{moland_2013,Villasenor-Derbez_2018}.
 To perform difference-in-differences, we regress the indicator of interest on a dummy variable for treatment, a dummy variable for years, and the interaction term between these with a multiple linear regression of the form:
@@
 Where year-level fixed effects are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
-The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site.
+The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site after acounting for the temporal and spatial differences.
+Therefore, we would expect this term to be positive if the indicator increases bacause of the reserve implementation.
 When evaluating biomass and densities of the invertebrate or fish communities, we include $\sigma_j$ to control for species-level fixed effects.

### 2018-12-05 docs/Villasenor-Derbez.Rmd
toned down generalization to close #22

@@
 Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
-We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based marine reserves in small-scale fisheries in Mexico and in other regions around the world where this tool is used to manage and rebuild their coastal fisheries.
+We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild their coastal fisheries.
 

### 2018-12-05 docs/Villasenor-Derbez.Rmd
Added caveats to socioeconomic data to close #30

@@
 
+Our analyses of socioeconomic indicators has two limitations.
+First, we only look at landings and revenues by landings for communities with and without TURF-reserves.
+There are a number of other possible indicators that could show a change due to the implementation of the reserve.
+Notably, one often cited in the literature is additional benefits, such as tourism \citep{viana_2017}.
+However, it is unlikely that the evaluated communities will experience tourism benefits due to their remoteness and the lack of proper infrastructure to sustain tourism.
+
+A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve.
+For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort.
+Even without changes to revenues, a lower input of effort (*e.g.* less fishing hours, less trips, less fuel) would imply that the fishery is more profitable.
+However, fishers are assambled into fishing cooperatives, meaning that profits are shared amongst all members \citep{mccay_2014,mccay_2017}.
+Therefore, the same revenue is still shared by the same amount of fishers.
+
 A first possible explanation for the lack of effectiveness may be the young age of the reserves.

### 2018-12-06 scripts/data_cleaning_logbook.Rmd
Added missing zero's for fish and inverts. This fixes #48

@@
                          TRUE ~ ano - 2013)) %>% 
+  spread(generoespecie, abundancia, fill = 0) %>% 
+  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>% 
   select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, abundancia) %>% 
@@
                          TRUE ~ ano - 2013)) %>% 
+  spread(generoespecie, abundancia, fill = 0) %>% 
+  gather(generoespecie, abundancia, -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto", "talla", "a", "b")) %>% 
   select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 

### 2018-12-06 docs/SupplementaryMaterial.Rmd
Added tables and supplementary figures to close #23 close #26

--- 2018-12-04 docs/Villasenor-Derbez_SupplementaryMaterial.Rmd
@@
     keep_tex: yes
-    template: template_sup.tex
+    # template: template_sup.tex
 ---
 
-```{r setup, include=FALSE}
+```{r setup, include = FALSE}
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
@@
 })
+```
+
+# Data
+
+```{r load_data}
+invert <- read.csv(file = here::here("data", "invertebrates.csv"),
+                stringsAsFactors = F)
+
+fish <- read.csv(file = here::here("data", "fish.csv"),
+                stringsAsFactors = F)
+
+conapesca <- read.csv(file = here::here("data", "conapesca.csv"),
+                stringsAsFactors = F)
+```
+
+```{r}
+invert %>% 
+  group_by(ano, comunidad, zona, sitio, transecto) %>%
+  count() %>%
+  group_by(comunidad, zona) %>%
+  count() %>% 
+  spread(zona, nn) %>% 
+  knitr::kable(caption = "Number of invertebrate transects performed in each site of each community.",
+               col.names = c("Community", "Control", "Reserve"))
+```
+
+```{r}
+invert %>%
+  filter(generoespecie %in% c("Panulirus interruptus", "Panulirus argus")) %>% 
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto) %>% 
+  summarize(indicador = sum(abundancia)/60) %>%
+  ungroup() %>% 
+  mutate(year = as.numeric(year)) %>% 
+  ggplot(aes(x = year, y = indicador, color = zona, fill = zona, group)) +
+  stat_summary(geom = "ribbon", fun.data = mean_se, alpha = 0.5) +
+  stat_summary(geom = "line", fun.y = mean) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab(bquote("Density org m"^{-2})) +
+  xlab("Year since implementation") +
+  theme(legend.justification = c(0,0),
+        legend.position = c(0.5, 0.15))
+```
+
+```{r}
+invert %>%
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 
+  summarize(indicador = sum(abundancia)/60) %>%
+  ungroup() %>% 
+  mutate(year = as.numeric(year)) %>% 
+  ggplot(aes(x = year, y = indicador, color = zona, fill = zona, group)) +
+  stat_summary(geom = "ribbon", fun.data = mean_se, alpha = 0.5) +
+  stat_summary(geom = "line", fun.y = mean) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab(bquote("Density org m"^{-2})) +
+  xlab("Year since implementation") +
+  theme(legend.justification = c(0,0),
+        legend.position = c(0.5, 0.15))
+```
+
+```{r}
+fish %>% 
+  group_by(ano, comunidad, zona, sitio, transecto) %>%
+  count() %>%
+  group_by(comunidad, zona) %>%
+  count() %>% 
+  spread(zona, nn) %>% 
+  knitr::kable(caption = "Number of fish transects performed in each site of each community.",
+               col.names = c("Community", "Control", "Reserve"))
+```
+
+```{r}
+fish %>%
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 
+  summarize(indicador = sum(abundancia)/60) %>% 
+  ungroup() %>% 
+  mutate(year = as.numeric(year)) %>% 
+  ggplot(aes(x = year, y = indicador, color = zona, fill = zona, group)) +
+  stat_summary(geom = "ribbon", fun.data = mean_se, alpha = 0.5) +
+  stat_summary(geom = "line", fun.y = mean) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab(bquote("Density org m"^{-2})) +
+  xlab("Year since implementation") +
+  theme(legend.justification = c(0,0),
+        legend.position = c(0.5, 0.15))
+```
+
+```{r}
+fish %>%
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie, talla, a, b) %>% 
+  summarize(abundancia = sum(abundancia)/60) %>% 
+  ungroup() %>% 
+  mutate(year = as.numeric(year),
+         biomass = abundancia * a * (talla^b)/1000) %>% 
+  ggplot(aes(x = year, y = biomass, color = zona, fill = zona, group)) +
+  stat_summary(geom = "ribbon", fun.data = mean_se, alpha = 0.5) +
+  stat_summary(geom = "line", fun.y = mean) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab(bquote("Biomass Kg m"^{-2})) +
+  xlab("Year since implementation") +
+  theme(legend.justification = c(0,0),
+        legend.position = c(0.5, 0.15))
 ```

### 2018-12-06 docs/manuscript.Rmd
Added tables and supplementary figures to close #23 close #26

--- 2018-12-05 docs/Villasenor-Derbez.Rmd
@@
 
+<!-- mccay_2014,mccay_2017 -->
+
 # Introduction
@@
 In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
-One of the many important ways to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves within them, thus creating TURF-reserve systems \citep{costello_2010,afflerbach_2014,lester_2017}.
+One of the many important ways to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves within them, thus creating TURF-reserve systems \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
 TURFs are a fisheries management tool in which a well defined group of fishers have exclusive access to an explicitly delimited portion of the ocean.
-They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,mccay_2017}.
+They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,costello_2010,mccay_2014}.
 On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits.
@@
 
-Research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access \citep{gelcich_2008,mccay_2014,mccay_2017}.
-Likewise, conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009,giakoumi_2017-V2,sala_2017-69}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
-The benefits resulting from reserves established within TURFs should be captured exclusively by the group of fishers with exclusive access.
-Although in theory these systems are successful \citep{costello_2010,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014,lester_2017}.
+Conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \citep{gelcich_2008,gelcich_2012}.
+The benefits resulting from reserves established within TURFs (*i.e.* TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \citep{gelcich_2015}.
+Although in theory these systems are successful \citep{gelcich_2015,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014,lester_2017}.
 
@@
 All sites were surveyed annually, and at least once before implementation of the reserves.
-A summary of sampling effort (number of transects per site) is shown in the supplementary materials (S1 Table and S2 Table).
+A summary of sampling effort (number of transects per site) is shown in the supplementary materials (Tables S1-S2).
 
@@
 
-B_ME <- fish%>% 
+B_ME <- fish %>% 
   biomass(com = "Maria Elena") %>% 
@@
 \ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
-Full tables with model coefficients are presented in the supplementary materials **(S1 Table, S2 Table, S3 Table)**.
+Figures and tables with timeseries of indicators and model coefficients are presented in the supplementary materials (Figures S1-S2, Tables S3-S5).
 
@@
 Points are jittered hotizontally to avoid overplotting.
-Points indicate the effect size and standard errors.
+Points indicate the effect size and error bars are heteroskedastic-robust standard errors.
 Years have been centered to year of implementation."}
@@
 ```{r, fig.width = 5, fig.height = 4, fig.cap = "\\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles).
-Points indicate the effect size and standard errors.
+Points are jittered hotizontally to avoid overplotting.
+Points indicate the effect size and error bars are heteroskedastic-robust standard errors.
 Years have been centered to year of implementation."}

### 2018-12-06 docs/manuscript.tex
Added tables and supplementary figures to close #23 close #26

--- 2018-12-04 docs/Villasenor-Derbez.tex
@@
 thus creating TURF-reserve systems
-\citep{costello_2010,afflerbach_2014,lester_2017}.
+\citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
@@
 users to sustainably manage their resources
-\citep{gelcich_2008,mccay_2017}. On the other hand, no-take marine
-reserves (marine reserves from hereinafter) are areas where all
-extractive activities are off-limits. These can be implemented to
-protect biodiversity but also as fishery management tools that restrict
-fishing effort and gears and therefore aid in the recovery of marine
-stocks. Commonly known as TURF-reserves, the combination of these tools
-should in theory increase the benefits of spatial access rights allowing
-the maintenance of healthy resources
+\citep{gelcich_2008,costello_2010,mccay_2014}. On the other hand,
+no-take marine reserves (marine reserves from hereinafter) are areas
+where all extractive activities are off-limits. These can be implemented
+to protect biodiversity but also as fishery management tools that
+restrict fishing effort and gears and therefore aid in the recovery of
+marine stocks. Commonly known as TURF-reserves, the combination of these
+tools should in theory increase the benefits of spatial access rights
+allowing the maintenance of healthy resources
 \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
-Research on TURFs has shown that these areas have higher abundance of
-targeted species than sites operating under open access
-\citep{gelcich_2008,mccay_2014,mccay_2017}. Likewise, conservation
-science has shown how marine reserves lead to increased biomass, species
-richness, and abundance within the protected regions
-\citep{lester_2009,giakoumi_2017-V2,sala_2017-69}, and that these may
-have a series of additional benefits like climate change mitigation,
-protection from environmental variability, and fisheries benefits
-\cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. The benefits
-resulting from reserves established within TURFs should be captured
-exclusively by the group of fishers with exclusive access. Although in
-theory these systems are successful
-\citep{costello_2010,smallhornwest_2018}, there is little empirical
-evidence of their effectiveness and the drivers of their success
-\citep{afflerbach_2014,lester_2017}.
+Conservation science has shown how marine reserves lead to increased
+biomass, species richness, and abundance within the protected regions
+\citep{lester_2009}, and that these may have a series of additional
+benefits like climate change mitigation, protection from environmental
+variability, and fisheries benefits
+\cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
+research on TURFs has shown that these areas have higher abundance of
+targeted species than sites operating under open access and even similar
+to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The
+benefits resulting from reserves established within TURFs (\emph{i.e.}
+TURF-reserves) should be captured exclusively by the group of fishers
+with exclusive access \citep{gelcich_2015}. Although in theory these
+systems are successful \citep{gelcich_2015,smallhornwest_2018}, there is
+little empirical evidence of their effectiveness and the drivers of
+their success \citep{afflerbach_2014,lester_2017}.
 
@@
 case studies and provide management recommendations to maximize the
-effectiveness of community-based marine reserves in small-scale
-fisheries in Mexico and in other regions around the world where this
-tool is used to manage and rebuild their coastal fisheries.
+effectiveness of community-based TURF-reserves in small-scale fisheries
+where this tool is used to manage and rebuild their coastal fisheries.
 
@@
 Ecological data come from the annual ecological monitoring of reserve
-and control areas, carried out by members from each community and
+and control sites. Reserve sites are those withing the reserves, and
+thus no fishing takes place. We define control sites as areas where: i)
+habitat characteristics are similar to the corresponding reserves, ii)
+presumably had a similar probability of being selected as reserves
+during the design phase, and iii) are located within the TURF, where
+fishing occurs. We focus our evaluation on sites where data are
+available for reserve and control sites, before and after the
+implementation of the reserve. This provides us with a
+Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
+allows us to capture and control for temporal and spatial dynamics
+\citep{stewartoaten_1986,depalma_2018} and causally attribute the
+changes to the reserve
+\citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.
+
+The biological data are collectged by members from each community and
 personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
@@
 sites \citep{fulton_2018,fulton_2019,suman_2010-ez}. Size structures are
-also collected during fish surveys. We define control sites as regions
-where: i) habitat characteristics are similar to the corresponding
-reserves, ii) presumably had a similar probability of being selected as
-reserves during the design phase, and iii) are located within the TURF
-and therefore fishing occurs. We focus our evaluation on sites where
-data are available for reserve and control sites, before and after the
-implementation of the reserve. This provides us with a
-Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
-allows us to capture and control for temporal and spatial dynamics
-\citep{stewartoaten_1986,depalma_2018}.
+also collected during fish surveys. All sites were surveyed annually,
+and at least once before implementation of the reserves. A summary of
+sampling effort (number of transects per site) is shown in the
+supplementary materials (S1 Table and S2 Table).
 
@@
 (Kg) and revenues (MXP) for cooperatives with and without marine
-reserves. Cooperatives incorporated in this analysis belong to larger
-regional-level Cooperative Federations, and are exposed to the same
-markets and institutional frameworks, making them plausible controls
-\citep{mccay_2017,ayer_2018}. Landings and revenues were aggregated at
-the cooperative-year level, and revenues were adjusted to represent 2014
-values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}.
+reserves. In this case our treated unit are the cooperatives
+(\emph{i.e.} communities) that have implemented a reserve within their
+TURF, and the controls are nearby communities that have a TURF but did
+not implement a reserve. Cooperatives incorporated in this analysis
+belong to larger regional-level Cooperative Federations, and are exposed
+to the same markets and institutional frameworks, making them plausible
+controls \citep{mccay_2017,ayer_2018}. Landings and revenues were
+aggregated at the cooperative-year level, and revenues were adjusted to
+represent 2014 values by the Consumer Price Index for Mexico
+\citep{oecd_2017}. A table with summary statistics for this data is
+provided in the supplementary materials \textbf{(S3 Table)}.
 
@@
 
-BACI designs and causal inference techniques have proven effective to
-evaluate marine reserves, as they allow us to causally attribute
-observed changes to the intervention
-\citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}. All sites
-were surveyed annually, and at least once before implementation of the
-reserves. We evaluate the effect that marine reserves have had on four
-ecological and two socioeconomic indicators (Table
-\ref{table:indicators}). Recall that reserves were implemented to
-protect lobster and other benthic invertebrates. However, we also use
-the available fish data to test for associated co-benefits.
+We evaluate the effect that the TURF-reserves have had on four
+ecological and two socioeconomic indicators shown in Table
+\ref{table:indicators}. Recall that reserves were implemented to protect
+lobster and other benthic invertebrates. However, we also use the
+available fish and invertebrate data to test for associated co-benefits.
 
 We use a difference-in-differences analysis to evaluate these
-indicators. This approach allows us to estimate the effect that the
-reserve had by comparing trends across time and treatments
-\citep{moland_2013,Villasenor-Derbez_2018}. The analysis of ecological
-indicators is performed with a multiple linear regression of the form:
+indicators. This approach is widely used in econometric literature to
+estimate the average treatment effect of an intervention, like the
+impact of minimum wage increases on employment rates \citep{card_1994}.
+In our case it allows us to estimate the effect that the reserve had on
+each biological and socioeconomic indicator (Table
+\ref{table:indicators}) by comparing trends across time and treatments
+\citep{moland_2013,Villasenor-Derbez_2018}. To perform
+difference-in-differences, we regress the indicator of interest on a
+dummy variable for treatment, a dummy variable for years, and the
+interaction term between these with a multiple linear regression of the
+form:
 
@@
 indicator inside the reserve, for year \(t\), with respect to the year
-of implementation in the control site. When evaluating biomass and
-densities of the invertebrate or fish communities, we include
-\(\sigma_j\) to control for species-level fixed effects.
-\(\epsilon_{i,t,j}\) represents the error term of the regression.
+of implementation in the control site after acounting for the temporal
+and spatial differences. Therefore, we would expect this term to be
+positive if the indicator increases bacause of the reserve
+implementation. When evaluating biomass and densities of the
+invertebrate or fish communities, we include \(\sigma_j\) to control for
+species-level fixed effects. \(\epsilon_{i,t,j}\) represents the error
+term of the regression.
 
@@
 and across sites, relative to the control site on the year of
-implementation (\emph{i.e.} effect size \(\lambda_t\)). We also provide
-an overview of the governance settings of each community, and discuss
-how these might be related to the effectiveness and performance of the
+implementation (\emph{i.e.} the difference-in-differences estimate or
+effect size \(\lambda_t\) from \ref{eqn:reg_bio}). We also provide an
+overview of the governance settings of each community, and discuss how
+these might be related to the effectiveness and performance of the
 reserves.
@@
 without clear trends. Full tables with model coefficients are presented
-in the supplementary materials (S1 Table, S2 Table, S3 Table).
+in the supplementary materials \textbf{(S1 Table, S2 Table, S3 Table)}.
 
@@
 findings of lack of change in the indicators used.
+
+Our analyses of socioeconomic indicators has two limitations. First, we
+only look at landings and revenues by landings for communities with and
+without TURF-reserves. There are a number of other possible indicators
+that could show a change due to the implementation of the reserve.
+Notably, one often cited in the literature is additional benefits, such
+as tourism \citep{viana_2017}. However, it is unlikely that the
+evaluated communities will experience tourism benefits due to their
+remoteness and the lack of proper infrastructure to sustain tourism.
+
+A second limitation of our socioeconomic analysis is that we do not
+observe effort data, which may mask the effect of the reserve. For
+example, if catches remain relatively unchanged but fishing effort
+decreased, that would imply a larger catch per unit effort. Even without
+changes to revenues, a lower input of effort (\emph{e.g.} less fishing
+hours, less trips, less fuel) would imply that the fishery is more
+profitable. However, fishers are assambled into fishing cooperatives,
+meaning that profits are shared amongst all members
+\citep{mccay_2014,mccay_2017}. Therefore, the same revenue is still
+shared by the same amount of fishers.
 
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
 \caption{\label{fig:unnamed-chunk-7}\label{fig:map}Location of the three
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
 \caption{\label{fig:unnamed-chunk-8}\label{fig:indicators}Effect sizes for
-marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
+marine reserves from Isla Natividad (IN; red circles), Maria Elena (ME;
 blue triangles), and Punta Herrero (PH; green squares) for lobster
@@
 hotizontally to avoid overplotting. Points indicate the effect size and
-standard errors. Years have been centered to year of implementation.}
+error bars are heteroskedastic-robust standard errors. Years have been
+centered to year of implementation.}
 \end{figure}
@@
 \centering
-\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-9-1.pdf}
+\includegraphics{manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
 \caption{\label{fig:unnamed-chunk-9}\label{fig:lobsters}Effect sizes for
 lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
-circles) and Maria Elena (ME; blue triangles). Points indicate the
-effect size and standard errors. Years have been centered to year of
-implementation.}
+circles) and Maria Elena (ME; blue triangles). Points are jittered
+hotizontally to avoid overplotting. Points indicate the effect size and
+error bars are heteroskedastic-robust standard errors. Years have been
+centered to year of implementation.}
 \end{figure}

### 2018-12-07 docs/SupplementaryMaterial.Rmd
Added more information on the supplementary materials to close #40, close #39

--- 2018-12-04 docs/Villasenor-Derbez_SupplementaryMaterial.Rmd
@@
 ---
-title: "Effectiveness of community-based marine reserves in small-scale fisheries"
+title: "Effectiveness of community-based TURF-reserves in small-scale fisheries"
 output:
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+Raw data and code are available on GitHub at: [https://github.com/jcvdav/ReserveEffect](https://github.com/jcvdav/ReserveEffect)
+
+```{r setup, include = FALSE}
+knitr::opts_chunk$set(echo = FALSE,
+                      warning = FALSE,
+                      message = FALSE)
 
@@
 
-```{r}
+```{r, eval = T}
 suppressPackageStartupMessages({
@@
 ```
+
+```{r load_data, eval = T}
+invert <- read.csv(file = here::here("data", "invertebrates.csv"),
+                stringsAsFactors = F) %>% 
+  mutate(zona = ifelse(zona == "Reserva", "Reserve", zona))
+
+fish <- read.csv(file = here::here("data", "fish.csv"),
+                stringsAsFactors = F) %>% 
+  mutate(zona = ifelse(zona == "Reserva", "Reserve", zona))
+
+conapesca <- read.csv(file = here::here("data", "conapesca.csv"),
+                stringsAsFactors = F) %>% 
+  mutate(zona = ifelse(zona == "Reserva", "Reserve", zona))
+```
+
+```{r eval = T}
+invert %>% 
+  group_by(ano, comunidad, zona, sitio, transecto) %>%
+  count() %>%
+  group_by(comunidad, zona) %>%
+  count() %>% 
+  spread(zona, nn) %>% 
+  knitr::kable(caption = "Number of invertebrate transects performed in each site of each community.",
+               col.names = c("Community", "Control", "Reserve"),
+               booktabs = T)
+```
+
+```{r, eval = T, fig.cap = "Time series of lobster densities. Bars indicate standard errors."}
+invert %>%
+  filter(generoespecie %in% c("Panulirus interruptus", "Panulirus argus")) %>% 
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 
+  summarize(indicador = sum(abundancia)/60) %>%
+  ungroup() %>% 
+  mutate(year = as.numeric(year)) %>% 
+  ggplot(aes(x = year, y = indicador, color = zona)) +
+  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, size = 1) +
+  stat_summary(geom = "line", fun.y = mean) +
+  stat_summary(geom = "point", fun.y = mean, size = 2) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab(bquote("Density (org m"^{-2}~")")) +
+  xlab("Year since implementation") +
+  theme(legend.justification = c(0,0),
+        legend.position = c(0.5, 0.15)) +
+  guides(color = guide_legend(title = "Zone"))
+```
+
+```{r, fig.cap = "Time series of invertebrate densities. Bars indicate standard errors."}
+invert %>%
+  filter(abundancia > 0) %>%
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 
+  summarize(indicador = sum(abundancia)/60) %>%
+  ungroup() %>% 
+  mutate(year = as.numeric(year)) %>% 
+  ggplot(aes(x = year, y = indicador, color = zona)) +
+  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, size = 1) +
+  stat_summary(geom = "line", fun.y = mean) +
+  stat_summary(geom = "point", fun.y = mean, size = 2) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab(bquote("Density (org m"^{-2}~")")) +
+  xlab("Year since implementation") +
+  theme(legend.justification = c(0,0),
+        legend.position = c(0.5, 0.15)) +
+  guides(color = guide_legend(title = "Zone"))
+```
+
+\clearpage
+
+```{r}
+fish %>% 
+  group_by(ano, comunidad, zona, sitio, transecto) %>%
+  count() %>%
+  group_by(comunidad, zona) %>%
+  count() %>% 
+  spread(zona, nn) %>% 
+  knitr::kable(caption = "Number of fish transects performed in each site of each community.",
+               col.names = c("Community", "Control", "Reserve"),
+               booktabs = T)
+```
+
+```{r, fig.cap = "Time series of fish biomass. Bars indicate standard errors."}
+fish %>%
+  filter(abundancia > 0) %>% 
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie, talla, a, b) %>% 
+  summarize(abundancia = sum(abundancia)) %>% 
+  ungroup() %>% 
+  mutate(year = as.numeric(year),
+         biomass = abundancia * ((a * (talla^b))/1000))  %>% 
+  ggplot(aes(x = year, y = biomass, color = zona)) +
+  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, size = 1) +
+  stat_summary(geom = "line", fun.y = mean) +
+  stat_summary(geom = "point", fun.y = mean, size = 2) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab(bquote("Biomass (Kg m"^{-2}~")")) +
+  xlab("Year since implementation") +
+  theme(legend.justification = c(0,0),
+        legend.position = c(0.5, 0.15)) +
+  guides(color = guide_legend(title = "Zone"))
+
+```
+
+\clearpage
+
+```{r, fig.cap = "Time series of fish densities. Bars indicate standard errors."}
+fish %>%
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 
+  summarize(indicador = sum(abundancia)/60) %>%
+  ungroup() %>% 
+  mutate(year = as.numeric(year)) %>% 
+  ggplot(aes(x = year, y = indicador, color = zona)) +
+  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, size = 1) +
+  stat_summary(geom = "line", fun.y = mean) +
+  stat_summary(geom = "point", fun.y = mean, size = 2) +
+  theme_minimal() +
+  facet_wrap(~comunidad, scales = "free", ncol = 2) +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab(bquote("Density (org m"^{-2}~")")) +
+  xlab("Year since implementation") +
+  theme(legend.justification = c(0,0),
+        legend.position = c(0.5, 0.15)) +
+  guides(color = guide_legend(title = "Zone"))
+```
+
+\clearpage
+
+```{r}
+conapesca %>% 
+  mutate(price = valor / peso_vivo) %>% 
+  group_by(zona, grupo, post, unidad_economica) %>% 
+  summarize(price = mean(price, na.rm = T)) %>% 
+  ungroup() %>% 
+  spread(post, price) %>% 
+  knitr::kable(digits = 2,
+               col.names = c("Zone", "Group", "TURF", "Before", "After"),
+               caption = "Mean ex-vessel prices for lobster (Panulirus spp.) on each TURF (community) by zone (MC = Caribbean; PN = Pacific) and group (Control = only TURF, Treated = TURF and reserve).",
+               booktabs = T)
+```
+
+```{r, fig.cap = "Time series of landings and value of landings."}
+valor <- conapesca %>% 
+  ggplot(aes(x = years_centered, y = valor / 1000, color = grupo)) +
+  facet_wrap(~ zona, ncol = 2, scales = "free_y")  +
+  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, size = 1) +
+  stat_summary(geom = "line", fun.y = mean) +
+  stat_summary(geom = "point", fun.y = mean, size = 2) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab("Value (K MXP)") +
+  xlab("") +
+  guides(color = guide_legend(title = "", ncol = 2)) +
+  theme(legend.justification = c(0,0),
+        legend.position = c(0.55, 0.7)) +
+  geom_vline(xintercept = 0, linetype = "dashed", size = 1)
+
+peso <- conapesca %>% 
+  ggplot(aes(x = years_centered, y = peso_vivo / 1000, color = grupo)) +
+  facet_wrap(~ zona, ncol = 2, scales = "free_y")  +
+  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, size = 1) +
+  stat_summary(geom = "line", fun.y = mean) +
+  stat_summary(geom = "point", fun.y = mean, size = 2) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1", direction = -1) +
+  scale_fill_brewer(palette = "Set1", direction = -1) +
+  ylab("Landings (Tones)") +
+  xlab("Year since implementation") +
+  theme(legend.position = "none") +
+  geom_vline(xintercept = 0, linetype = "dashed", size = 1)
+
+cowplot::plot_grid(valor, peso, ncol = 1)
+
+```
+
+\clearpage
 
@@
 
+\clearpage
+
 ```{r, results = "asis"}
@@
 
+\clearpage
+
 ```{r, results = "asis"}
 list(L_PH, B_PH, Ni_PH, N_PH) %>% 
-  supp_regtable(caption = "Coefficient estimates of biological indicators for Punta Herrero.", fs = "small")
+  supp_regtable(caption = "Coefficient estimates of biological indicators for Punta Herrero.",
+                fs = "small")
 ```
+
+\clearpage
 
@@
 list(C_IN, V_IN) %>% 
-  supp_regtable(caption = "Coefficient estimates of socioeconomic indicators Isla Natividad.", bio = F)
+  supp_regtable(caption = "Coefficient estimates of socioeconomic indicators Isla Natividad.",
+                bio = F)
 ```
+
+\clearpage
 
@@
 list(C_ME, V_ME) %>% 
-  supp_regtable(caption = "Coefficient estimates of socioeconomic indicators for Maria Elena.", bio = F)
+  supp_regtable(caption = "Coefficient estimates of socioeconomic indicators for Maria Elena.",
+                bio = F)
 ```

### 2018-12-07 docs/manuscript.Rmd
Combined paragraphs of socioeco discussion and fixed ambiguity about TURF-Reserves vs reserves to close #53

@@
 While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth causal inference analysis that we undertake.
-Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our insights to other similar reserves in Mexico and elsewhere around the world.
+Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our insights to other similar community-based TURF-reserves in Mexico and elsewhere around the world.
 
@@
 This implies that treated sites would have followed the same trend as control sites, had the reserves not been implemented.
-Nonetheless, temporal trends for each site don't show any significant increases (S1 Table, S2 Table, S3 Table), supporting our findings of lack of change in the indicators used.
+Nonetheless, temporal trends for each site don't show any significant increases (Figures S1-S4), supporting our findings of lack of change in the indicators used.
 
@@
 However, it is unlikely that the evaluated communities will experience tourism benefits due to their remoteness and the lack of proper infrastructure to sustain tourism.
-
 A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve.
@@
 Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
-Furthermore, the lack of effectiveness observed in these reserves should not be generalizable to other reserves established under the same legal framework (*i.e.* Fish Refuges) in Mexico, and future research should aim at evaluating other areas that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), or others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}).
+Furthermore, our work evaluates only TURF-reserves established as Fish Refuges and is not representative of other Fish Refuges not implemented within TURFs.
+Future research should aim at evaluating other areas that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.
 

### 2018-12-07 docs/manuscript.Rmd
source socioeco plotting function

@@
 source(file = here::here("scripts", "robust_se_list.R")) #heteroskedastic-robust Standard Errors for stargazer
-source(file = here::here("scripts", "plot_model.R")) #plots effect sizes
+source(file = here::here("scripts", "plot_model.R")) #plots effect sizes for Bio
+source(file = here::here("scripts", "plot_socioeco.R")) #plots effect sizes for Socioeco
+
 ```

### 2018-12-07 docs/manuscript.Rmd
This filter is now in the data cleaning logbook

@@
 conapesca %<>%
-   filter(years_centered > -10) %>% 
   mutate(years_centered = as.factor(years_centered),

### 2018-12-07 docs/manuscript.Rmd
This function is now in scripts/plotsocioeco.R

@@
   lm(valor ~ zona * year + unidad_economica, data = .)
-
-#soc_res se fue
-
-plot_socioeco <- function(model_list, model_names, ylab){
-
-  extract_filter <- function(model, name){
-    model %>% 
-      robust_se() %>% 
-      filter(grepl(pattern = "zonaReserva:year", x = term)) %>% 
-      mutate(years_centered = as.numeric(gsub(pattern = "zonaReserva:year",
-                                              replacement = "",
-                                              x = term)),
-             Community = name)
-  }
-
-  purrr::map2_df(model_list, model_names, extract_filter) %>% 
-    ggplot(aes(x = years_centered, y = estimate, group = Community)) +
-    geom_hline(yintercept = 0,
-               linetype = "dashed",
-               size = 1) +
-    geom_vline(xintercept = 0,
-               linetype = "dashed",
-               size = 1) +
-    geom_errorbar(aes(color = Community,
-                      ymin = estimate-std.error,
-                      ymax = estimate + std.error),
-                  width = 0.1,
-                  size = 1,
-                  position = position_dodge(width = 0.4)) +
-    geom_point(aes(color = Community, shape = Community),
-               size = 4,
-               position = position_dodge(width = 0.4),
-               alpha = 0.5) +
-    scale_color_brewer(palette = "Set1") +
-    labs(x = "Years since implementation", y = ylab) +
-    theme(text = element_text(size = 12))
-}
 ```
@@
                         model_names = list("IN", "ME"),
-                        ylab = quo(lambda[t]))
+                        ylab = quo(lambda[t])) +
+  theme(legend.justification = c(0, 0),
+        legend.position = c(0, 0.55),
+        legend.title = element_blank())
 
@@
                         model_names = list("IN", "ME"),
-                        ylab = quo(lambda[t]))
+                        ylab = quo(lambda[t])) +
+  theme(legend.position = "none")
 

### 2018-12-07 docs/manuscript.Rmd
Updated discussion to talk about lessons learned

@@
 Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
-Furthermore, our work evaluates only TURF-reserves established as Fish Refuges and is not representative of other Fish Refuges not implemented within TURFs.
+Furthermore, our work evaluates only TURF-reserves established as Fish Refuges and is not representative of other Fish Refuges implemented without TURFs.
 Future research should aim at evaluating other areas that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.
 
-Community-based marine reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented.
-Lessons learned from these cases can guide implementation of community-based marine reserves elsewhere.
-For the particular case of the marine reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
+Community-based TURF-reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented.
+As we move forward in marine conservation, we must recognize the value of local knowledge to identify priority areas for conservation, but without setting design principles aside.
+Lessons learned from these cases can guide implementation of community-based TURF-reserves elsewhere.
+For the particular case of the ones we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
 Community-based marine reserves might have more benefits that result from indirect effects of the reserves, particularly providing resilience to shocks and management errors, and promoting social cohesion, which should be taken into account when evaluating the outcomes of TURF-reserves.

### 2018-12-07 docs/manuscript.tex
Updated discussion to talk about lessons learned

@@
 sampling effort (number of transects per site) is shown in the
-supplementary materials (S1 Table and S2 Table).
+supplementary materials (Tables S1-S2).
 
@@
 \ref{fig:indicators}B-D), where effect sizes oscillated around zero
-without clear trends. Full tables with model coefficients are presented
-in the supplementary materials \textbf{(S1 Table, S2 Table, S3 Table)}.
+without clear trends. Figures and tables with timeseries of indicators
+and model coefficients are presented in the supplementary materials
+(Figures S1-S2, Tables S3-S5).
 

### 2018-12-07 scripts/data_cleaning_logbook.Rmd
Updated discussion to talk about lessons learned

--- 2018-07-06 scripts/data_cleaning_logbook.Rmd
@@
 ---
+
+The commit history for this file can be seen [here](https://github.com/jcvdav/ReserveEffect/commits/master/scripts/data_cleaning_logbook.Rmd)
 
@@
   mutate(Genus = as.character(str_split(GeneroEspecie, pattern = " ", simplify = T)[,1]),
-         Genus = str_replace(Genus, " ", ""))
+         Genus = str_replace(Genus, " ", "")) %>% 
+  drop_na(a, b)
 
-needed_spp <- c("Kyphosus incisor",
+needed_spp <- c("Embiotoca jacksoni",
+                "Kyphosus incisor",
                 "Kyphosus sectatrix",
@@
 
+manual_fishbase_bayesian <- tibble(
+  GeneroEspecie = c("Chromis punctipinnis",
+                    "Oxyjulis californica",
+                    "Heterodontus francisci",
+                    "Rhacochilus vacca",
+                "Squatina californica"),
+  a = c(0.01514, 0.00447, 0.00355, 0.01950, 0.00741),
+  b = c(2.99, 3.14,3.16, 3.02, 3.04)
+)
+
 needed_genus <- c("Mycteroperca",
@@
 GenusLevelBio <- species_bio %>% 
-  filter(Genus %in% needed_genus) %>%
-  select(GeneroEspecie = Genus, a, b) %>% 
-  group_by(GeneroEspecie) %>% 
+  select(Genus, ag = a, bg = b) %>% 
+  group_by(Genus) %>% 
   summarize_all(mean, na.rm = T) %>% 
@@
   rbind(FB_update) %>% 
-  rbind(GenusLevelBio) %>% 
+  rbind(manual_fishbase_bayesian) %>% 
   janitor::clean_names() %>% 
@@
                                    TRUE ~ generoespecie)) %>% 
-  left_join(species_bio, by = c("generoespecie")) %>% 
+  mutate(Genus = as.character(str_split(generoespecie, pattern = " ", simplify = T)[,1]),
+         Genus = str_replace(Genus, " ", "")) %>% 
+  left_join(species_bio, by = "generoespecie") %>% 
+  left_join(GenusLevelBio, by = "Genus") %>%
+  mutate(a = ifelse(is.na(a), ag, a),
+         b = ifelse(is.na(b), bg, b)) %>% 
   group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie, talla, a,b) %>% 
@@
                          comunidad == "Maria Elena" ~ ano - 2012,
-                         TRUE ~ ano - 2013)) %>% 
+                         TRUE ~ ano - 2013))
+
+fish <- fish %>% 
   select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 
-  arrange() %>% 
-  drop_na(a, b)
+  arrange()
 ```
@@
   select(Zona, Grupo = Group, Comunidad, UnidadEconomica, Ano, Post, years_centered, NombreComun, PesoVivo, PesoDesembarcado, Valor, CPI) %>% 
-  janitor::clean_names()
+  janitor::clean_names() %>% 
+  filter(!(zona == "MC" & years_centered %in% c(-3, 2))) %>% 
+  filter(years_centered > -7)
 ```
-
-# Export all data
 

### 2018-12-07 scripts/plot_socioeco.R

@@ -1,0 +1,34 @@
+plot_socioeco <- function(model_list, model_names, ylab){
+  
+  extract_filter <- function(model, name){
+    model %>% 
+      robust_se() %>% 
+      filter(grepl(pattern = "zonaReserva:year", x = term)) %>% 
+      mutate(years_centered = as.numeric(gsub(pattern = "zonaReserva:year",
+                                              replacement = "",
+                                              x = term)),
+             Community = name)
+  }
+  
+  purrr::map2_df(model_list, model_names, extract_filter) %>% 
+    ggplot(aes(x = years_centered, y = estimate, group = Community)) +
+    geom_hline(yintercept = 0,
+               linetype = "dashed",
+               size = 1) +
+    geom_vline(xintercept = 0,
+               linetype = "dashed",
+               size = 1) +
+    geom_errorbar(aes(color = Community,
+                      ymin = estimate-std.error,
+                      ymax = estimate + std.error),
+                  width = 0.1,
+                  size = 1,
+                  position = position_dodge(width = 0.4)) +
+    geom_point(aes(color = Community, shape = Community),
+               size = 4,
+               position = position_dodge(width = 0.4),
+               alpha = 0.5) +
+    scale_color_brewer(palette = "Set1") +
+    labs(x = "Years since implementation", y = ylab) +
+    theme(text = element_text(size = 12))
+}

### 2018-12-07 docs/manuscript.Rmd
Explained the dips that we would expect to see if reserves were costly to close #49

@@
 Furthermore, fishers may favor implementation of reserves that pose low fishing costs due to their small size or location.
-Our analysis of economic data supports this hypothesis, as neither landings nor revenues showed the expected short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
+Our analysis of economic data supports this hypothesis, as neither landings nor revenues showed the expected short-term reductions associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
 

### 2018-12-07 docs/manuscript.Rmd
Added reference to stefan to make explicit tat we dont refer to this reserves to close #52

@@
 
-Community-based TURF-reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented.
-As we move forward in marine conservation, we must recognize the value of local knowledge to identify priority areas for conservation, but without setting design principles aside.
+Community-based TURF-reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented \citep{gelcich_2015}.
+To implement them we must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside.
 Lessons learned from these cases can guide implementation of community-based TURF-reserves elsewhere.

### 2018-12-07 docs/manuscript.Rmd
Further made clear the points in #52 and modified discussion to explain and close #50

@@
 The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
-Despite the lack of evidence of the effectiveness of these reserves, most of the communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}.
-Understanding the social-ecological context in which these communities and their reserves operate might provide insights as to why this happens.
+Lastly, the communities exhibit all the social enabling conditions for effective reserve and resource management.
+Here we discuss possible shortcommings in our analyses as well as possible explanations for the observed patterns.
 
@@
 Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
-Furthermore, our work evaluates only TURF-reserves established as Fish Refuges and is not representative of other Fish Refuges implemented without TURFs.
-Future research should aim at evaluating other areas that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through a top-down process (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.
+These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}.
 
-Community-based TURF-reserves in small-scale fisheries can be helpful conservation and fishery management tools when appropriately implemented \citep{gelcich_2015}.
+In terms of the fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs.
+Future research should aim at evaluating other Fish Refuges that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through top-down processes (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.
+
+Community-based TURF-reserves in small-scale fisheries may be helpful conservation and fishery management tools when appropriately implemented \citep{gelcich_2015}.
 To implement them we must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside.

### 2018-12-07 docs/manuscript.Rmd
Added latest pdf version to review by hand

@@
 Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
-One of the ways in which fishing communities have takane advantage of this new tool is by implementing marine reserves within their TURFs.
+One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporal marine reserves within their TURFs with a defined expiration date (often 5 years).
+After these five years have passed, fishers have the opportunity to open the reserves to fishing or continue to have them.
 Our work focuses on some of these community-based TURF-reserves that occur in small-scale fisheries.
@@
 Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively.
-These reserves have been legally recognized as Fishing Refuges since their creation \citep{dof_website_2012,dof_website_2013}.
+These reserves have been legally recognized as Fishing Refuges since their original implementation \citep{dof_website_2012,dof_website_2013} and re-enactments \citep{dof_website_2017}.
 

### 2018-12-07 docs/manuscript.tex
Added latest pdf version to review by hand

@@
 socioeconomic and governance setting under which they operate, it is
-safe to cautiously generalize our insights to other similar reserves in
-Mexico and elsewhere around the world.
+safe to cautiously generalize our insights to other similar
+community-based TURF-reserves in Mexico and elsewhere around the world.
 
@@
 pertaining landings and revenues showed little to no change after
-reserve implementation. Despite the lack of evidence of the
-effectiveness of these reserves, most of the communities show a positive
-perception about their performance and continue to support their
-presence \citep{ayer_2018}. Understanding the social-ecological context
-in which these communities and their reserves operate might provide
-insights as to why this happens.
+reserve implementation. Lastly, the communities exhibit all the social
+enabling conditions for effective reserve and resource management. Here
+we discuss possible shortcommings in our analyses as well as possible
+explanations for the observed patterns.
 
@@
 implemented. Nonetheless, temporal trends for each site don't show any
-significant increases (S1 Table, S2 Table, S3 Table), supporting our
-findings of lack of change in the indicators used.
+significant increases (Figures S1-S4), supporting our findings of lack
+of change in the indicators used.
 
@@
 evaluated communities will experience tourism benefits due to their
-remoteness and the lack of proper infrastructure to sustain tourism.
-
-A second limitation of our socioeconomic analysis is that we do not
+remoteness and the lack of proper infrastructure to sustain tourism. A
+second limitation of our socioeconomic analysis is that we do not
 observe effort data, which may mask the effect of the reserve. For
@@
 this hypothesis, as neither landings nor revenues showed the expected
-short-term costs associated to the first years of reserve implementation
-\citep{ovando_2016-Wg}.
+short-term reductions associated to the first years of reserve
+implementation \citep{ovando_2016-Wg}.
 
@@
 allows fishers to have greater bargaining power and leverage over
-fisheries management \citep{prezramrez_2012}. Furthermore, the lack of
-effectiveness observed in these reserves should not be generalizable to
-other reserves established under the same legal framework (\emph{i.e.}
-Fish Refuges) in Mexico, and future research should aim at evaluating
-other areas that have also been established as bottom-up processes but
-without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
-or others established through a top-down process (\emph{i.e.}
-\citet{dof_websiteU_2018}).
+fisheries management \citep{prezramrez_2012}. These additional benefits
+might explain why communities show a positive perception about their
+performance and continue to support their presence \citep{ayer_2018}.
 
-Community-based marine reserves in small-scale fisheries can be helpful
-conservation and fishery management tools when appropriately
-implemented. Lessons learned from these cases can guide implementation
-of community-based marine reserves elsewhere. For the particular case of
-the marine reserves that we evaluate, the possibility of expanding
-reserves or merging existing polygons into larger areas should be
-evaluated and proposed to the communities. Community-based marine
-reserves might have more benefits that result from indirect effects of
-the reserves, particularly providing resilience to shocks and management
-errors, and promoting social cohesion, which should be taken into
-account when evaluating the outcomes of TURF-reserves. Having full
-community support surely represents an advantage, but it is important
-that community-based TURF-reserves meet essential design principles such
-as size and placement so as to maximize their effectiveness.
+In terms of the fisheries regulation in Mexico, our work only evaluates
+Fish Refuges established within TURFs. Future research should aim at
+evaluating other Fish Refuges that have also been established as
+bottom-up processes but without the presence of TURFs (\emph{e.g.}
+\citet{dof_websiteC_2012}), others established through top-down
+processes (\emph{i.e.} \citet{dof_websiteU_2018}), as well as the
+relationship between governance and effectiveness across this gradient
+of approaches.
+
+Community-based TURF-reserves in small-scale fisheries may be helpful
+conservation and fishery management tools when appropriately implemented
+\citep{gelcich_2015}. To implement them we must promote bottom-up design
+and implementation processes like the ones in the evaluated reserves,
+but without setting design principles aside. Lessons learned from these
+cases can guide implementation of community-based TURF-reserves
+elsewhere. For the particular case of the ones we evaluate, the
+possibility of expanding reserves or merging existing polygons into
+larger areas should be evaluated and proposed to the communities.
+Community-based marine reserves might have more benefits that result
+from indirect effects of the reserves, particularly providing resilience
+to shocks and management errors, and promoting social cohesion, which
+should be taken into account when evaluating the outcomes of
+TURF-reserves. Having full community support surely represents an
+advantage, but it is important that community-based TURF-reserves meet
+essential design principles such as size and placement so as to maximize
+their effectiveness.
 

### 2018-12-27 docs/manuscript.Rmd
Made all changes and sent to coauthors

@@
 Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
-In the last 12 years, 18 old and new community-based TURF-reserves gained legal recognition thanks to a 2014 regulation; their effectiveness has not been formally evaluated.
+In the last 12 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a 2014 regulation; their effectiveness has not been formally evaluated.
 We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
-We find that while reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates, they continue to receive significant support from the fishing communities.
+We find that while reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates, they continue to receive support from the fishing communities.
 A lack of clear ecological and socioeconomic effects likely results from a combination of factors.
@@
 Third, variable and extreme oceanographic conditions have impacted harvested populations.
-Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in landings.
+Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches.
 However, these reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
@@
 
-<!-- Words in text: 3714 | Words in headers: 35 | Words outside text (captions, etc.): 192 -->
-
-<!-- \clearpage -->
-
-<!-- mccay_2014,mccay_2017 -->
-
 # Introduction
 
-Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}.
+Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}.
 In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
-One of the many important ways to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves within them, thus creating TURF-reserve systems \citep{afflerbach_2014,gelcich_2015,lester_2017}.
+One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves within them, thus creating TURF-reserve systems \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
@@
 These can be implemented to protect biodiversity but also as fishery management tools that restrict fishing effort and gears and therefore aid in the recovery of marine stocks.
-Commonly known as TURF-reserves, the combination of these tools should in theory increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,gelcich_2015,lester_2017}.
+These instruments can be combined by establishing a marine reserve within a TURF, thus making them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
-Conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+Conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
 Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \citep{gelcich_2008,gelcich_2012}.
 The benefits resulting from reserves established within TURFs (*i.e.* TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \citep{gelcich_2015}.
-Although in theory these systems are successful \citep{gelcich_2015,smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success \citep{afflerbach_2014,lester_2017}.
+Although in theory these systems are successful \citep{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success.
 
-TURF-reserve systems are inherently intrincate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{gelcich_2015}.
+TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{gelcich_2015}.
 This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009}.
@@
 
-Moreover, a new Mexican norm was created in 2014 allowing fishers to request the legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}).
+Recent norms in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries.
+In Mexico, a norm created in 2014 allows fishers to request legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}).
 Since 2012, old and new marine reserves have gained legal recognition as Fishing Refuges.
-Of these, 18 were originally implemented as community-based TURF-reserves.
+Of these, 18 were originally implemented as within TURFs.
 However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
@@
 First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries.
-Second, to evaluate the effectiveness of TURF-reserves established as Fish Refuges in Mexico to identify opportunities where improvement or adjustment might lead to increased effectiveness.
+Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness.
 We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild their coastal fisheries.
-
-<!-- Have lead to the creation of decission support tools to guide the design and implementation of TURF-reserve systems\citep{oyanedel_2017}. -->
-
-<!-- Modelling studies show that fishery benefits of marine reserves depend on initial stock status and the management under which the fishery operates, as well as reserve size and the amount of larvae exported from these \citep{hilborn_2006,krueck_2017-J1,deleo_2015}.
--->
 
@@
 Community-based marine reserves that are implemented within TURFs are a form of TURF-reserves, voluntarily established and enforced by local communities.
-This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{beger_2004,gelcich_2015,smallhornwest_2018}.
+This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
 Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
-However, marine reserves are difficult to enforce if they are not legally recognized, and fishers rely on the exclusive access granted by the TURF.
+However, community-based reserves can be hard to enforce if they are not legally recognized.
+In such conditions, TURF fishers must rely on the exclusive access of the TURF to maintain high levels of compliance.
 
 In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue (*i.e.* Fish Refuges; \citet{nom}).
-Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
+Fish Refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
 One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporal marine reserves within their TURFs with a defined expiration date (often 5 years).
 After these five years have passed, fishers have the opportunity to open the reserves to fishing or continue to have them.
-Our work focuses on some of these community-based TURF-reserves that occur in small-scale fisheries.
+Our work focuses on Fish Refuges implemented as community-based TURF-reserves that occur in small-scale fisheries.
 
 The common setup of community-based TURF-reserves in Mexico is the following.
-Fishers from a given community are assambled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (*i.e.* TURFs shown as blue polygons in Fig \ref{fig:map}A).
-Each TURF is exlusively fished by one cooperative, and each community usually hosts no more than one cooperative.
+Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (*i.e.* TURFs shown as blue polygons in Fig \ref{fig:map}A).
+Each TURF is elusively fished by one cooperative, and each community usually hosts no more than one cooperative.
+The profits from each TURF are shared amongst all fishers from the cooperative.
 Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (*i.e.* TURF-reserves).
-Fishers then ask the government to grant legal recognition to their TURF-reserves under the name of Fish refuges following a series of studies outlined in \citet{nom}.
+Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges following a series of studies outlined in the regulation \citep{nom}.
 
@@
 
-We evaluate three community-based no-take marine reserves implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig \ref{fig:map}A).
+We evaluate three community-based no-take TURF-reserves implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig \ref{fig:map}A).
 The first one was created by the *Buzos y Pescadores de la Baja California* fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig \ref{fig:map}B).
@@
 In 2006, the community decided to implement two marine reserves within their fishing grounds to protect commercially important invertebrate species; mainly lobster and abalone.
-These reserves obtained legal recognition only in 2018 \citep{dof_website_2018}.
+These reserves obtained legal recognition in 2018 \citep{dof_website_2018}.
 
@@
 Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively.
-These reserves have been legally recognized as Fishing Refuges since their original implementation \citep{dof_website_2012,dof_website_2013} and re-enactments \citep{dof_website_2017}.
+These reserves have been legally recognized as Fishing Refuges since their original implementation \citep{dof_website_2012,dof_website_2013} and re-enactments \citep{dof_website_2017b}.
 
@@
 Isla Natividad, for example, is part of a greater group of fishing cooperatives belonging to a Federation of Fishing Cooperatives.
-This group has been identified as a cohesive group that often cooperates to better manage their resources \citep{mccay_2014,mccay_2017,acevesbueno_2017}.
+This group has been identified as a cohesive group that cooperates to better manage their resources \citep{mccay_2014,mccay_2017,acevesbueno_2017}.
 Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation.
@@
 While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth causal inference analysis that we undertake.
-Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our insights to other similar community-based TURF-reserves in Mexico and elsewhere around the world.
+Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our insights to other similar community-based TURF-reserves in Mexico and elsewhere.
 
-The regulation governing the implementation of Fish refuges states that these are fishery management tools intended to have biological or socioeconomic benefits \citep{nom}.
+The regulation governing the implementation of Fish Refuges states that these are fishery management tools intended to have biological or socioeconomic benefits \citep{nom}.
 For this reason, the main portion of our analyses focuses on a series of biological and socioeconomic indicators that may respond to reserve implementation.
@@
 We therefore incorporate a reduced version of the Social Ecological Systems framework \citep{ostrom_2009} and evaluate variables and indicators known to aid and hinder the effectiveness of management interventions in conservation and fisheries.
-The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather provide a discussion of potential causes of individual reserve performance.
+The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather help provide context on the social-ecological system in which reserves develop.
 The following two sections describe our data collection methods and analyses of biological and socioeconomic indicators as well as the SES analysis.
@@
   mutate(evaluated = ifelse(Name %in% c("El Cabezo", "El Faro (Pesca de Langosta)", "El Faro"), T, F),
-         Comunidad = ifelse(Name %in% c("El Faro", "Laguna Canche Balam", "El Faro (Pesca de Langosta)", "Anegado de Chal"), "Puna Herrero", "Maria Elena"))
+         Comunidad = ifelse(Name %in% c("El Faro", "Laguna Canche Balam", "El Faro (Pesca de Langosta)", "Anegado de Chal"), "Punta Herrero", "Maria Elena"))
 
@@
 Ecological data come from the annual ecological monitoring of reserve and control sites.
-Reserve sites are those withing the reserves, and thus no fishing takes place.
-We define control sites as areas where: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, and iii) are located within the TURF, where fishing occurs.
+Reserve sites are those within the reserves, and thus no fishing takes place.
+Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) Are not directly adjacent to the reserves.
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
@@
 
-The biological data are collectged by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
-Trained divers record richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \citep{fulton_2018,fulton_2019,suman_2010-ez}.
+The biological data are collected by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
+Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \citep{suman_2010-ez,fulton_2018,fulton_2019}.
 Size structures are also collected during fish surveys.
 All sites were surveyed annually, and at least once before implementation of the reserves.
-A summary of sampling effort (number of transects per site) is shown in the supplementary materials (Tables S1-S2).
+A summary of sampling effort is shown in the supplementary materials (Tables S1-S2).
 
 Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA).
-Data contain monthly lobster landings (Kg) and revenues (MXP) for cooperatives with and without marine reserves.
+Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves.
 In this case our treated unit are the cooperatives (*i.e.* communities) that have implemented a reserve within their TURF, and the controls are nearby communities that have a TURF but did not implement a reserve.
@@
 Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017}.
-A table with summary statistics for this data is provided in the supplementary materials **(S3 Table)**.
+A table with summary statistics for this data is provided in the supplementary materials **(Table S3, Figure S5)**.
 
-Data for the evaluation of the SES were collected at the community-level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on the authors' experience and knowledge of the communities.
-These include information on the Resource Systems, Resource Units, Actors, and Governance System (Table \ref{table:ses}).
+Data for the evaluation of the SES were collected at the community-level from official documents used in the creation and designation of the marine reserves.
+These includes the technical studies that the cooperatives submit when they request recognition of their reserves, as well as the official enactments  \citep{dof_website_2012,dof_website_2013,dof_website_2018}.
+We also complimented information based on the authors' experience and knowledge of the communities.
+We collected information on the Resource Systems, Resource Units, Actors, and Governance System (Table \ref{table:ses}).
+The next sections further describe the variables and indicators chosen for each of these.
 
@@
 \begin{equation}
-I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \sigma_jSpp_j + \epsilon_{i,t,j}
+I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t,j}
 \label{eqn:reg_bio}
@@
 
-Where year-level fixed effects are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
-The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site after acounting for the temporal and spatial differences.
-Therefore, we would expect this term to be positive if the indicator increases bacause of the reserve implementation.
-When evaluating biomass and densities of the invertebrate or fish communities, we include $\sigma_j$ to control for species-level fixed effects.
-$\epsilon_{i,t,j}$ represents the error term of the regression.
+Where year-level fixed effects capturing a temporal trend are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
+The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
+Therefore, we would expect this term to be positive if the indicator increases because of the reserve implementation.
+Finally, $\epsilon_{i,t,j}$ represents the error term of the regression.
 
@@
 \begin{equation}
-I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jCom_j +\epsilon_{i,t,j}
+I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t,j}
 \label{eqn:soc_reg}
@@
 
-The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$) and $\sigma_jCom$ captures community-level fixed-effects.
-These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}.
-The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
+The coefficient interpretations remains as for Eq.
+\ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$).
+These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \citep{depalma_2018}.
 All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}.
 All analyses were performed in R `r paste0("version ", version$major, ".", version$minor)` and R Studio version 1.1.456 \citep{R_2018}.
-All data and code ara available in a [GitHub repository](https://github.com/jcvdav/ReserveEffect).
+All data and code needed to reproduce our analyses are available in a [GitHub repository](https://github.com/jcvdav/ReserveEffect).
 
-We use the SES framework to evaluate each community.
+We use the SES framework to evaluate each community and create a narrative that provides context for each community.
 The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators.
-We based our variable selection primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework.
+We based our variable selection primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework, and identified the key variables relevant to fishing cooperatives in Mexico.
 We also incorporate other relevant variables known to influence reserve performance following \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.
-Table \ref{table:ses} shows the selected variables, their definition and values.
+Table \ref{table:ses} shows the selected variables, along with definitions and values.
 
@@
   abundance(com = "Isla Natividad") %>% 
-  model(sfe = F)
+  model(sfe = F, spfe = F)
 
@@
   abundance(com = "Maria Elena") %>% 
-  model(sfe = F)
+  model(sfe = F, spfe = F)
 
@@
   abundance(com = "Punta Herrero") %>% 
-  model(sfe = F)
+  model(sfe = F, spfe = F)
 
@@
   biomass(com = "Isla Natividad") %>% 
-  model(sfe = F)
+  model(sfe = F, spfe = F)
 
@@
   biomass(com = "Maria Elena") %>% 
-  model(sfe = F)
+  model(sfe = F, spfe = F)
 
@@
   biomass(com = "Punta Herrero") %>% 
-  model(sfe = F)
+  model(sfe = F, spfe = F)
 
@@
 The following sections present the effect that marine reserves had on each of the biological and socioeconomic indicators for each coastal community.
-Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* the difference-in-differences estimate or effect size $\lambda_t$ from \ref{eqn:reg_bio}).
+Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* the difference-in-differences estimate or effect size $\lambda_t$ from Eqs.
+\ref{eqn:reg_bio} and \ref{eqn:soc_reg}).
 We also provide an overview of the governance settings of each community, and discuss how these might be related to the effectiveness and performance of the reserves.
@@
 \ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
-Figures and tables with timeseries of indicators and model coefficients are presented in the supplementary materials (Figures S1-S2, Tables S3-S5).
+Figures and tables with time series of indicators and model coefficients are presented in the supplementary materials (Figures S1-S4, Tables S4-S6).
 
@@
 Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities.
-However, these changes are not significant and are associated with increased variation.
-Full tables with model coefficients are presented in the supplementary materials (S4 Table, S5 Table).
+However, these changes are only significant for the third year ($p < 0.05$).
+Full tables with model coefficients are presented in the supplementary materials (Tables S4-S5).
 
@@
 However, our SES analysis also highlights factors that might hinder reserve performance or mask outcomes.
-While total reserve size ranges from 0.2% to 3.7% of the TURF area, individual reserves are often small (RS3), and relatively young (RS5).
+While total reserve size ranges from 0.2% to 3.7% of the TURF area, individual reserves are often small (RS3); the largest reserve is only 4.37 km ^2^, and the smallest one is 0.09 km ^2^.
+Reserves are also relatively young (RS5).
 Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely that marine reserves will result in increased catches.
-
-The fact that these reserves were re-established and just enacted shows a comitment of support (issue \#16 on GH).
-the SES provides a static picture of governance and resource conditions, whereas the biological assessment reflects changes through time (Isse \#31).
 
@@
   filter(grupo == "PN") %>%
-  lm(peso_vivo ~ zona * year + unidad_economica, data = .)
+  lm(peso_vivo ~ zona * year, data = .)
 
@@
   filter(grupo == "MC") %>%
-  lm(peso_vivo ~ zona * year + unidad_economica, data = .)
+  lm(peso_vivo ~ zona * year, data = .)
 
@@
   filter(grupo == "PN") %>%
-  lm(valor ~ zona * year + unidad_economica, data = .)
+  lm(valor ~ zona * year, data = .)
 
@@
   filter(grupo == "MC") %>%
-  lm(valor ~ zona * year + unidad_economica, data = .)
+  lm(valor ~ zona * year, data = .)
 ```
@@
 
-<!-- From Fio: One aspect that is important to discuss is the concurrent effect of environmental variability.
-In isla Natividad, and the whole Pacifico Norte region, the decade following the establishment of the reserves has seen extreme ENSO events and hypoxia, which have impacted populations above and beyond the effects of fishing and conservation.
--->
-
 Our results indicate that these TURF-reserves have not increased lobster densities.
@@
 Lastly, the communities exhibit all the social enabling conditions for effective reserve and resource management.
-Here we discuss possible shortcommings in our analyses as well as possible explanations for the observed patterns.
+Here we discuss possible shortcomings in our analyses as well as possible explanations for the observed patterns.
 
-Some works evaluate marine reserves by performing inside-outside \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or before-after comparisons \citep{betti_2017-lq}.
-The first approach does not address temporal variability, and the second can not distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}.
-While many ecology studies have used BACI sampling designs and respective analyses (*e.g.* \citep{stewartoaten_1986}), few conservations studies have done so to evaluate the effect of an intervention (*i.e.* \citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted in a call for more robust analyses in conservation science \citep{guidetti_2002,ferraro_2006}.
-
+While many ecology studies have used BACI sampling designs and respective analyses (*e.g.* \citet{stewartoaten_1986}), few conservation studies have done so to evaluate the effect of an intervention (*e.g.* \citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted in a call for more robust analyses in conservation science \citep{guidetti_2002,ferraro_2006}.
 Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness.
@@
 A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve.
-For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort.
-Even without changes to revenues, a lower input of effort (*e.g.* less fishing hours, less trips, less fuel) would imply that the fishery is more profitable.
-However, fishers are assambled into fishing cooperatives, meaning that profits are shared amongst all members \citep{mccay_2014,mccay_2017}.
-Therefore, the same revenue is still shared by the same amount of fishers.
+For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and higher profitability.
 
@@
 For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2004,hilborn_2006}.
-Both lobster fisheries were certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}.
-Additionally, lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \citep{dof_website_1993}.
-It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves.
+Both lobster fisheries were certified by the Marine Stewardship Council and lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \citep{dof_website_1993}.
+It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves; reserves implemented in TURFs where fishing pressure is already optimally managed will still show a trade-off between fisheries and conservation objectives \citep{lester_2017}.
 Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open access.
@@
 Finally, extreme conditions, including prolonged hypoxia, heat waves, and storms have affected both the Pacific and Caribbean regions, with large negative impacts of coastal marine species and ecosystems \citep{cavole_2016,hughes_2018,breitburg_2018}.
-The coastal ecosystems where these reserves are located have been profoundly affected by these events \citep{micheli_2012-EU,woodson}.
+The coastal ecosystems where these reserves are located have been profoundly affected by these events \citep{micheli_2012-EU,woodson_2018}.
 Effects of protection might be eliminated by the mortalities associated with these extreme conditions.
@@
 Marine reserves provide protection to a wider range of species and vulnerable habitat.
-These sites can serve as an insurance against uncertainty and errors in fisheries management, as well as mild environmental shocks \citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
-Self-regulation of fishing effort (*i.e.* reduction in harvest) can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
-Embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}.
-Showing commitment to marine conservation and sustainable fishing practices allows fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
-These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence \citep{ayer_2018}.
+Previous research focusing on these specific sites has shown that they serve as an insurance mechanism against uncertainty and errors in fisheries management, as well as mild environmental shocks \citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
+Self-regulation of fishing effort can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
+Furthermore, embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital as reported by \citet{fulton_2019}.
+Showing commitment to marine conservation and sustainable fishing practices has allowed fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
+These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-implementing the reserves \citep{ayer_2018}.
 
@@
 Community-based TURF-reserves in small-scale fisheries may be helpful conservation and fishery management tools when appropriately implemented \citep{gelcich_2015}.
-To implement them we must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside.
-Lessons learned from these cases can guide implementation of community-based TURF-reserves elsewhere.
+We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside.
+Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.
 For the particular case of the ones we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
-Community-based marine reserves might have more benefits that result from indirect effects of the reserves, particularly providing resilience to shocks and management errors, and promoting social cohesion, which should be taken into account when evaluating the outcomes of TURF-reserves.
-Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.
 
@@
   knitr::kable(caption = "\\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).
-Alphanumeric codes follow \\citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \\citet{difranco_2016-Xw} and \\citet{edgar_2014-UO}.",
+Alphanumeric codes follow \\citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \\citet{difranco_2016-Xw} and \\citet{edgar_2014-UO}.
+The presented narrative applies equally for all communities unless otherwise noted.",
                format = "latex",

### 2018-12-27 docs/manuscript.tex
Made all changes and sent to coauthors

@@
 fishers reap the benefits of conservation investments. In the last 12
-years, 18 old and new community-based TURF-reserves gained legal
+years, 18 old and new community-based Mexican TURF-reserves gained legal
 recognition thanks to a 2014 regulation; their effectiveness has not
@@
 increasing the density of lobster and other benthic invertebrates, they
-continue to receive significant support from the fishing communities. A
-lack of clear ecological and socioeconomic effects likely results from a
+continue to receive support from the fishing communities. A lack of
+clear ecological and socioeconomic effects likely results from a
 combination of factors. First, some of these reserves might be too young
@@
 local fisheries are already well managed, and it is unlikely that
-reserves might have a detectable effect in landings. However, these
+reserves might have a detectable effect in catches. However, these
 reserves may provide a foundation for establishing additional, larger
@@
 overfishing and unsustainable fishing practices
-\citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. In particular,
+\citep{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular,
 small-scale fisheries face great challenges since they tend to be hard
-to monitor and enforce \citep{costello_2012}. One of the many important
-ways to improve the performance of coastal fisheries and health of the
+to monitor and enforce \citep{costello_2012}. One of the many approaches
+taken to improve the performance of coastal fisheries and health of the
 local resources is through the implementation of Territorial Use Rights
@@
 restrict fishing effort and gears and therefore aid in the recovery of
-marine stocks. Commonly known as TURF-reserves, the combination of these
-tools should in theory increase the benefits of spatial access rights
-allowing the maintenance of healthy resources
+marine stocks. These instruments can be combined by establishing a
+marine reserve within a TURF, thus making them TURF-reserves
 \citep{afflerbach_2014,gelcich_2015,lester_2017}.
@@
 variability, and fisheries benefits
-\cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
+\citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
 research on TURFs has shown that these areas have higher abundance of
@@
 with exclusive access \citep{gelcich_2015}. Although in theory these
-systems are successful \citep{gelcich_2015,smallhornwest_2018}, there is
-little empirical evidence of their effectiveness and the drivers of
-their success \citep{afflerbach_2014,lester_2017}.
+systems are successful \citep{smallhornwest_2018}, there is little
+empirical evidence of their effectiveness and the drivers of their
+success.
 
-TURF-reserve systems are inherently intrincate social-ecological
-systems, and their effectiveness must depend on how environmental and
-social factors combine and interact \citep{gelcich_2015}. This is
-especially important in social-ecological coastal systems dominated by
-close interactions and feedbacks between people and natural resources
+TURF-reserve systems are inherently intricate social-ecological systems,
+and their effectiveness must depend on how environmental and social
+factors combine and interact \citep{gelcich_2015}. This is especially
+important in social-ecological coastal systems dominated by close
+interactions and feedbacks between people and natural resources
 \citep{ostrom_2009}. There is a growing body of literature focusing on
@@
 
-Moreover, a new Mexican norm was created in 2014 allowing fishers to
-request the legal recognition of community-based reserves as ``Fish
-Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since 2012,
-old and new marine reserves have gained legal recognition as Fishing
-Refuges. Of these, 18 were originally implemented as community-based
-TURF-reserves. However, their effectiveness has not yet been formally
-evaluated and reported in the scientific literature.
+Recent norms in fisheries regulation in Mexico provide a ripe
+opportunity to study the effectiveness of community-based TURF-reserves
+in small-scale fisheries. In Mexico, a norm created in 2014 allows
+fishers to request legal recognition of community-based reserves as
+``Fish Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since
+2012, old and new marine reserves have gained legal recognition as
+Fishing Refuges. Of these, 18 were originally implemented as within
+TURFs. However, their effectiveness has not yet been formally evaluated
+and reported in the scientific literature.
 
@@
 indicators and the governance settings under which these develop, which
-may inform similar processes in other countries. Second, to evaluate the
-effectiveness of TURF-reserves established as Fish Refuges in Mexico to
-identify opportunities where improvement or adjustment might lead to
-increased effectiveness. We draw from lessons learned in these three
-case studies and provide management recommendations to maximize the
-effectiveness of community-based TURF-reserves in small-scale fisheries
-where this tool is used to manage and rebuild their coastal fisheries.
+may inform similar processes in other countries. Second, to identify
+opportunities where improvement or adjustment might lead to increased
+effectiveness. We draw from lessons learned in these three case studies
+and provide management recommendations to maximize the effectiveness of
+community-based TURF-reserves in small-scale fisheries where this tool
+is used to manage and rebuild their coastal fisheries.
 
@@
 self-enforcement, and reserves can yield benefits similar to
-systematically-designed reserves
-\citep{beger_2004,gelcich_2015,smallhornwest_2018}. Community-based
-spatial closures occur in different contexts, like the \emph{kapu} or
-\emph{ra'ui} areas in the Pacific Islands
-\citep{bohnsack_2004,johannes_2002}. However, marine reserves are
-difficult to enforce if they are not legally recognized, and fishers
-rely on the exclusive access granted by the TURF.
+systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
+Community-based spatial closures occur in different contexts, like the
+\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
+\citep{bohnsack_2004,johannes_2002}. However, community-based reserves
+can be hard to enforce if they are not legally recognized. In such
+conditions, TURF fishers must rely on the exclusive access of the TURF
+to maintain high levels of compliance.
 
@@
 and created a legal framework that solves this governance issue
-(\emph{i.e.} Fish Refuges; \citet{nom}). Fish refuges can be implemented
+(\emph{i.e.} Fish Refuges; \citet{nom}). Fish Refuges can be implemented
 as temporal or partial reserves, which can protect one, some, or all
 resources within their boundaries. One of the ways in which fishing
-communities have takane advantage of this new tool is by implementing
-marine reserves within their TURFs. Our work focuses on some of these
-community-based TURF-reserves that occur in small-scale fisheries.
+communities have taken advantage of this new tool is by implementing
+temporal marine reserves within their TURFs with a defined expiration
+date (often 5 years). After these five years have passed, fishers have
+the opportunity to open the reserves to fishing or continue to have
+them. Our work focuses on Fish Refuges implemented as community-based
+TURF-reserves that occur in small-scale fisheries.
 
 The common setup of community-based TURF-reserves in Mexico is the
-following. Fishers from a given community are assambled in fishing
+following. Fishers from a given community are assembled in fishing
 cooperatives which have exclusive fishing rights over a spatially
 delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
-\ref{fig:map}A). Each TURF is exlusively fished by one cooperative, and
-each community usually hosts no more than one cooperative. Fishing
-cooperatives interested in implementing marine reserves work with CSOs
-to implement marine reserves within their TURFs (\emph{i.e.}
+\ref{fig:map}A). Each TURF is elusively fished by one cooperative, and
+each community usually hosts no more than one cooperative. The profits
+from each TURF are shared amongst all fishers from the cooperative.
+Fishing cooperatives interested in implementing marine reserves work
+with CSOs to implement marine reserves within their TURFs (\emph{i.e.}
 TURF-reserves). Fishers then ask the government to grant legal
-recognition to their TURF-reserves under the name of Fish refuges
-following a series of studies outlined in \citet{nom}.
+recognition to their TURF-reserves as Fish Refuges following a series of
+studies outlined in the regulation \citep{nom}.
 
@@
 
-We evaluate three community-based no-take marine reserves implemented in
+We evaluate three community-based no-take TURF-reserves implemented in
 Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig
@@
 species; mainly lobster and abalone. These reserves obtained legal
-recognition only in 2018 \citep{dof_website_2018}.
+recognition in 2018 \citep{dof_website_2018}.
 
@@
 respectively. These reserves have been legally recognized as Fishing
-Refuges since their creation \citep{dof_website_2012,dof_website_2013}.
+Refuges since their original implementation
+\citep{dof_website_2012,dof_website_2013} and re-enactments
+\citep{dof_website_2017b}.
 
@@
 a Federation of Fishing Cooperatives. This group has been identified as
-a cohesive group that often cooperates to better manage their resources
+a cohesive group that cooperates to better manage their resources
 \citep{mccay_2014,mccay_2017,acevesbueno_2017}. Likewise, Maria Elena
@@
 safe to cautiously generalize our insights to other similar
-community-based TURF-reserves in Mexico and elsewhere around the world.
+community-based TURF-reserves in Mexico and elsewhere.
 
-The regulation governing the implementation of Fish refuges states that
+The regulation governing the implementation of Fish Refuges states that
 these are fishery management tools intended to have biological or
@@
 to relate different levels of governance with reserve effectiveness, but
-rather provide a discussion of potential causes of individual reserve
-performance. The following two sections describe our data collection
-methods and analyses of biological and socioeconomic indicators as well
-as the SES analysis.
+rather help provide context on the social-ecological system in which
+reserves develop. The following two sections describe our data
+collection methods and analyses of biological and socioeconomic
+indicators as well as the SES analysis.
 
@@
 Ecological data come from the annual ecological monitoring of reserve
-and control sites. Reserve sites are those withing the reserves, and
-thus no fishing takes place. We define control sites as areas where: i)
-habitat characteristics are similar to the corresponding reserves, ii)
-presumably had a similar probability of being selected as reserves
-during the design phase, and iii) are located within the TURF, where
-fishing occurs. We focus our evaluation on sites where data are
-available for reserve and control sites, before and after the
-implementation of the reserve. This provides us with a
-Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
-allows us to capture and control for temporal and spatial dynamics
-\citep{stewartoaten_1986,depalma_2018} and causally attribute the
-changes to the reserve
+and control sites. Reserve sites are those within the reserves, and thus
+no fishing takes place. Control sites are areas that meet the following
+criteria: i) habitat characteristics are similar to the corresponding
+reserves, ii) presumably had a similar probability of being selected as
+reserves during the design phase, iii) are located within the TURF,
+where fishing occurs, and iv) Are not directly adjacent to the reserves.
+We focus our evaluation on sites where data are available for reserve
+and control sites, before and after the implementation of the reserve.
+This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
+sampling design that allows us to capture and control for temporal and
+spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally
+attribute the changes to the reserve
 \citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.
 
-The biological data are collectged by members from each community and
+The biological data are collected by members from each community and
 personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
-(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
-abundances of fish and invertebrate species along replicate transects
-(30 \(\times\) 2 m each) at depths 5-20 m in the reserves and control
-sites \citep{fulton_2018,fulton_2019,suman_2010-ez}. Size structures are
-also collected during fish surveys. All sites were surveyed annually,
-and at least once before implementation of the reserves. A summary of
-sampling effort (number of transects per site) is shown in the
-supplementary materials (Tables S1-S2).
+(\href{www.cobi.org.mx}{COBI}). Trained divers record species richness
+and abundances of fish and invertebrate species along replicate
+transects (30 \(\times\) 2 m each) at depths 5-20 m in the reserves and
+control sites \citep{suman_2010-ez,fulton_2018,fulton_2019}. Size
+structures are also collected during fish surveys. All sites were
+surveyed annually, and at least once before implementation of the
+reserves. A summary of sampling effort is shown in the supplementary
+materials (Tables S1-S2).
 
@@
 Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
-(Kg) and revenues (MXP) for cooperatives with and without marine
-reserves. In this case our treated unit are the cooperatives
+(Kg) and revenues (MXP) for TURF-managed cooperatives with and without
+marine reserves. In this case our treated unit are the cooperatives
 (\emph{i.e.} communities) that have implemented a reserve within their
@@
 \citep{oecd_2017}. A table with summary statistics for this data is
-provided in the supplementary materials \textbf{(S3 Table)}.
+provided in the supplementary materials \textbf{(Table S3, Figure S5)}.
 
@@
 from official documents used in the creation and designation of the
-marine reserves
-\citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on
-the authors' experience and knowledge of the communities. These include
-information on the Resource Systems, Resource Units, Actors, and
-Governance System (Table \ref{table:ses}).
+marine reserves. These includes the technical studies that the
+cooperatives submit when they request recognition of their reserves, as
+well as the official enactments
+\citep{dof_website_2012,dof_website_2013,dof_website_2018}. We also
+complimented information based on the authors' experience and knowledge
+of the communities. We collected information on the Resource Systems,
+Resource Units, Actors, and Governance System (Table \ref{table:ses}).
+The next sections further describe the variables and indicators chosen
+for each of these.
 
@@
 
-Where year-level fixed effects are represented by \(\gamma_t Year_t\),
-and \(\beta Zone_i\) captures the difference between reserve
-(\(Zone = 1\)) and control (\(Zone = 0\)) sites. The interaction term
-\(\lambda_{it} Year_t\times Zone_i\) represents the mean change in the
-indicator inside the reserve, for year \(t\), with respect to the year
-of implementation in the control site after acounting for the temporal
-and spatial differences. Therefore, we would expect this term to be
-positive if the indicator increases bacause of the reserve
-implementation. When evaluating biomass and densities of the
-invertebrate or fish communities, we include \(\sigma_j\) to control for
-species-level fixed effects. \(\epsilon_{i,t,j}\) represents the error
-term of the regression.
+Where year-level fixed effects capturing a temporal trend are
+represented by \(\gamma_t Year_t\), and \(\beta Zone_i\) captures the
+difference between reserve (\(Zone = 1\)) and control (\(Zone = 0\))
+sites. The effect of the reserve is captured by the \(\lambda_t\)
+coefficient, and represents the difference observed between the control
+site before the implementation of the reserve and the treated sites at
+time \(t\) after controlling for other time and space variations
+(\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Therefore, we
+would expect this term to be positive if the indicator increases because
+of the reserve implementation. Finally, \(\epsilon_{i,t,j}\) represents
+the error term of the regression.
 
@@
 
-The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in
-this case the \(Treated\) dummy variable indicates if the community has
-a reserve (\(Treated = 1\)) or not (\(Treated = 0\)) and \(\sigma_jCom\)
-captures community-level fixed-effects. These regression models allow us
-to establish a causal link between the implementation of marine reserves
-and the observed trends by accounting for temporal and spatial dynamics
-\citep{depalma_2018}. The effect of the reserve is captured by the
-\(\lambda_t\) coefficient, and represents the difference observed
-between the control site before the implementation of the reserve and
-the treated sites at time \(t\) after controlling for other time and
-space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
-All model coefficients were estimated via ordinary least-squares and
+The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio},
+but in this case the \(Treated\) dummy variable indicates if the
+community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)).
+These regression models allow us to establish a causal link between the
+implementation of marine reserves and the observed trends by accounting
+for temporal and site-specific dynamics \citep{depalma_2018}. All model
+coefficients were estimated via ordinary least-squares and
 heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
 analyses were performed in R version 3.5.1 and R Studio version 1.1.456
-\citep{R_2018}. All data and code ara available in a
-\href{https://github.com/jcvdav/ReserveEffect}{GitHub repository}.
+\citep{R_2018}. All data and code needed to reproduce our analyses are
+available in a \href{https://github.com/jcvdav/ReserveEffect}{GitHub
+repository}.
 
-We use the SES framework to evaluate each community. The use of this
+We use the SES framework to evaluate each community and create a
+narrative that provides context for each community. The use of this
 framework standardizes our analysis and allows us to communicate our
@@
 operationalized and analyzed Mexican fishing cooperatives using this
-framework. We also incorporate other relevant variables known to
-influence reserve performance following \citet{difranco_2016-Xw} and
-\citet{edgar_2014-UO}. Table \ref{table:ses} shows the selected
-variables, their definition and values.
+framework, and identified the key variables relevant to fishing
+cooperatives in Mexico. We also incorporate other relevant variables
+known to influence reserve performance following
+\citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table
+\ref{table:ses} shows the selected variables, along with definitions and
+values.
 
@@
 implementation (\emph{i.e.} the difference-in-differences estimate or
-effect size \(\lambda_t\) from \ref{eqn:reg_bio}). We also provide an
-overview of the governance settings of each community, and discuss how
-these might be related to the effectiveness and performance of the
-reserves.
+effect size \(\lambda_t\) from Eqs. \ref{eqn:reg_bio} and
+\ref{eqn:soc_reg}). We also provide an overview of the governance
+settings of each community, and discuss how these might be related to
+the effectiveness and performance of the reserves.
 
@@
 \ref{fig:indicators}B-D), where effect sizes oscillated around zero
-without clear trends. Figures and tables with timeseries of indicators
+without clear trends. Figures and tables with time series of indicators
 and model coefficients are presented in the supplementary materials
-(Figures S1-S2, Tables S3-S5).
+(Figures S1-S4, Tables S4-S6).
 
@@
 implementation of the reserve, as compared to the control communities.
-However, these changes are not significant and are associated with
-increased variation. Full tables with model coefficients are presented
-in the supplementary materials (S4 Table, S5 Table).
+However, these changes are only significant for the third year
+(\(p < 0.05\)). Full tables with model coefficients are presented in the
+supplementary materials (Tables S4-S5).
 
@@
 reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual
-reserves are often small (RS3), and relatively young (RS5).
+reserves are often small (RS3); the largest reserve is only 4.37 km
+\textsuperscript{2}, and the smallest one is 0.09 km
+\textsuperscript{2}. Reserves are also relatively young (RS5).
 Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely
 that marine reserves will result in increased catches.
-
-The fact that these reserves were re-established and just enacted shows
-a comitment of support (issue \#16 on GH). the SES provides a static
-picture of governance and resource conditions, whereas the biological
-assessment reflects changes through time (Isse \#31).
 
@@
 enabling conditions for effective reserve and resource management. Here
-we discuss possible shortcommings in our analyses as well as possible
+we discuss possible shortcomings in our analyses as well as possible
 explanations for the observed patterns.
 
-Some works evaluate marine reserves by performing inside-outside
-\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or
-before-after comparisons \citep{betti_2017-lq}. The first approach does
-not address temporal variability, and the second can not distinguish
-between the temporal trends in a reserve and the entire system
-\citep{depalma_2018}. While many ecology studies have used BACI sampling
-designs and respective analyses (\emph{e.g.} \citep{stewartoaten_1986}),
-few conservations studies have done so to evaluate the effect of an
-intervention (\emph{i.e.}
+While many ecology studies have used BACI sampling designs and
+respective analyses (\emph{e.g.} \citet{stewartoaten_1986}), few
+conservation studies have done so to evaluate the effect of an
+intervention (\emph{e.g.}
 \citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted
 in a call for more robust analyses in conservation science
-\citep{guidetti_2002,ferraro_2006}.
-
-Our approach to evaluate the temporal and spatial changes provides a
-more robust measure of reserve effectiveness. For example, we capture
-previously described patterns like the rapid increase observed for
-lobster densities in Isla Natividad on the sixth year (\emph{i.e.} 2012;
-Fig. \ref{fig:indicators}A), a year after the hypoxia events described
-by \citet{micheli_2012-EU}, which caused mass mortality of sedentary
-organisms such as abalone and sea urchins, but not lobster and finfish.
-Yet, our empirical approach assumes control sites are a plausible
-counterfactual for treated sites. This implies that treated sites would
-have followed the same trend as control sites, had the reserves not been
-implemented. Nonetheless, temporal trends for each site don't show any
-significant increases (Figures S1-S4), supporting our findings of lack
-of change in the indicators used.
+\citep{guidetti_2002,ferraro_2006}. Our approach to evaluate the
+temporal and spatial changes provides a more robust measure of reserve
+effectiveness. For example, we capture previously described patterns
+like the rapid increase observed for lobster densities in Isla Natividad
+on the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), a year
+after the hypoxia events described by \citet{micheli_2012-EU}, which
+caused mass mortality of sedentary organisms such as abalone and sea
+urchins, but not lobster and finfish. Yet, our empirical approach
+assumes control sites are a plausible counterfactual for treated sites.
+This implies that treated sites would have followed the same trend as
+control sites, had the reserves not been implemented. Nonetheless,
+temporal trends for each site don't show any significant increases
+(Figures S1-S4), supporting our findings of lack of change in the
+indicators used.
 
@@
 example, if catches remain relatively unchanged but fishing effort
-decreased, that would imply a larger catch per unit effort. Even without
-changes to revenues, a lower input of effort (\emph{e.g.} less fishing
-hours, less trips, less fuel) would imply that the fishery is more
-profitable. However, fishers are assambled into fishing cooperatives,
-meaning that profits are shared amongst all members
-\citep{mccay_2014,mccay_2017}. Therefore, the same revenue is still
-shared by the same amount of fishers.
+decreased, that would imply a larger catch per unit effort and higher
+profitability.
 
@@
 is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
-fisheries were certified by the Marine Stewardship Council
-\citep{prezramrez_2016-J1}. Additionally, lobster fisheries are managed
-via species-specific minimum catch sizes, seasonal closures, protection
-of ``berried'' females, and escapement windows where traps are allowed
-\citep{dof_website_1993}. It is uncertain whether such a well-managed
-fishery will experience additional benefits from marine reserves.
+fisheries were certified by the Marine Stewardship Council and lobster
+fisheries are managed via species-specific minimum catch sizes, seasonal
+closures, protection of ``berried'' females, and escapement windows
+where traps are allowed \citep{dof_website_1993}. It is uncertain
+whether such a well-managed fishery will experience additional benefits
+from marine reserves; reserves implemented in TURFs where fishing
+pressure is already optimally managed will still show a trade-off
+between fisheries and conservation objectives \citep{lester_2017}.
 Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have
@@
 where these reserves are located have been profoundly affected by these
-events \citep{micheli_2012-EU,woodson}. Effects of protection might be
-eliminated by the mortalities associated with these extreme conditions.
+events \citep{micheli_2012-EU,woodson_2018}. Effects of protection might
+be eliminated by the mortalities associated with these extreme
+conditions.
 
@@
 social benefits. Marine reserves provide protection to a wider range of
-species and vulnerable habitat. These sites can serve as an insurance
+species and vulnerable habitat. Previous research focusing on these
+specific sites has shown that they serve as an insurance mechanism
 against uncertainty and errors in fisheries management, as well as mild
@@
 \citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
-Self-regulation of fishing effort (\emph{i.e.} reduction in harvest) can
-serve as a way to compensate for future declines associated to
-environmental variation \citep{finkbeiner_2018}. Embarking in a marine
-conservation project can bring the community together, which promotes
-social cohesion and builds social capital \citep{fulton_2019}. Showing
-commitment to marine conservation and sustainable fishing practices
-allows fishers to have greater bargaining power and leverage over
+Self-regulation of fishing effort can serve as a way to compensate for
+future declines associated to environmental variation
+\citep{finkbeiner_2018}. Furthermore, embarking in a marine conservation
+project can bring the community together, which promotes social cohesion
+and builds social capital as reported by \citet{fulton_2019}. Showing
+commitment to marine conservation and sustainable fishing practices has
+allowed fishers to have greater bargaining power and leverage over
 fisheries management \citep{prezramrez_2012}. These additional benefits
 might explain why communities show a positive perception about their
-performance and continue to support their presence \citep{ayer_2018}.
+performance and continue to support their presence by re-implementing
+the reserves \citep{ayer_2018}.
 
@@
 conservation and fishery management tools when appropriately implemented
-\citep{gelcich_2015}. To implement them we must promote bottom-up design
-and implementation processes like the ones in the evaluated reserves,
-but without setting design principles aside. Lessons learned from these
-cases can guide implementation of community-based TURF-reserves
-elsewhere. For the particular case of the ones we evaluate, the
-possibility of expanding reserves or merging existing polygons into
-larger areas should be evaluated and proposed to the communities.
-Community-based marine reserves might have more benefits that result
-from indirect effects of the reserves, particularly providing resilience
-to shocks and management errors, and promoting social cohesion, which
-should be taken into account when evaluating the outcomes of
-TURF-reserves. Having full community support surely represents an
-advantage, but it is important that community-based TURF-reserves meet
-essential design principles such as size and placement so as to maximize
-their effectiveness.
+\citep{gelcich_2015}. We must promote bottom-up design and
+implementation processes like the ones in the evaluated reserves, but
+without setting design principles aside. Having full community support
+surely represents an advantage, but it is important that community-based
+TURF-reserves meet essential design principles such as size and
+placement so as to maximize their effectiveness. For the particular case
+of the ones we evaluate, the possibility of expanding reserves or
+merging existing polygons into larger areas should be evaluated and
+proposed to the communities.
 
@@
 \caption{\label{tab:unnamed-chunk-11}\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).
-Alphanumeric codes follow \citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.}
+Alphanumeric codes follow \citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.
+The presented narrative applies equally for all communities unless otherwise noted.}
 \centering
@@
 \hline
-\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish.
-Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates.\\
+\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish (2 spp each).
+Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates (4 spp).\\
 \hline
@@
 \hline
-\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers.\\
+\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers.
+IN lies 545 Km south from Tijuana, and ME and PH 230 Km south from Cancun, where the nearest international airports are located.\\
 \hline
@@
 \hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandates.
-These include larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
+These are: larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
 \hline
-\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group that monitors and enforces formal and internal rules.
+\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group (Consejo de vigilancia) that monitors and enforces formal and internal rules.
 They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\

### 2018-12-27 docs/reviewer_comments.Rmd
Made all changes and sent to coauthors

--- 2018-12-03 docs/paper/reviewer_comments.Rmd
@@
 
-There are some rather loose statements in the ms, starting with the Abstract (see also Rev 1) where it is implied that a 'well-managed' fishery should result in no effect in a reserve.
+- There are some rather loose statements in the ms, starting with the Abstract (see also Rev 1) where it is implied that a 'well-managed' fishery should result in no effect in a reserve.
 This is false.
@@
 
-As per Reviewer 1, much more clarity is needed in Methods.
+*That is correct, a reserve would contain higher density, biomass, and richness.
+We have corrected the abstract to reflect what we intended to say, which is that a well-managed fishery would not see increased catches by an additional intervention.
+We do not include explicit references in the abstract, but the discussion touches on this topic and cites Hilborn, Micheli and De Leo's paper that evaluates when MRs result in increased catches.*
+
+*As for the unsubstantiated statement of good enforcement, it is unlikely that these reserves experience poaching.
+The reserves are implemented, managed, and monitored by the community themselves.
+We have expanded on this when we introduce the case studies.*
+
+- As per Reviewer 1, much more clarity is needed in Methods.
 It is unclear what data are being tested and what the sampling design was, especially the spatial arrangement of controls.
@@
 
-There is also some fairly lazy scholarship here: e.g.
+*We have expanded on the methods, especially by adding to the description of case studies (TURFs, reserve implementation process) and data collection methods to clarify what the sampling design is.
+We have also expanded on the regression model.
+It is a multiple linear regression because we incorporate more than one explanatory variable.
+The lambda that we report does not make reference to the lambda parameter in a Poisson distribution, but rather to the coefficient presented in Equations 1 and 2.
+We do not use a log-linear model that assumes a discrete poison distribution because our indicators are density (i.e.
+org / m^-2^) and biomass (Kg / m^-2^), which are continuous variables.*
+
+- There is also some fairly lazy scholarship here: e.g.
 BACI designs originate in papers other than those cited on line 112-113, and were demonstrated in papers well before those in 114-115, and again in 133-134.
@@
 
-**Recommendation:** This ms requires major revision.
+*The observation that Moland's paper might have confounded estimates of reserve effectiveness is correct.
+Proximity of their reserve and control sites might imply a spillover effect from the reserves to the controls may occur, which would result in more conservative estimates of reserve effectiveness.
+While there is a long history of time series analyses in ecology, we cite Moland's paper because, to our knowledge, it is the first paper that estimated reserve effectiveness using a difference-in-differences approach.
+We have included references to older implementations of BACI in ecology, as well as the seminal econometric paper that introduces the Difference-in-Differences methodology.*
+
+\clearpage
 
 # Reviewer 1
-
-**Recommendation:** Substantial revisions
-
-## Q1 - Please summarize the main findings of the study.
-
-The manuscript provides an analysis of ecological and economic fishery indicators across TURF-reserves in three Mexican communities, finding no substantial changes in target species or associated fisheries, or in value or quantity of landings.
-This is not necessarily surprising and the likely reasons are discussed.
-This study provides an example (adding to an existing literature) of community based reserves that have local support but do not necessarily deliver substantial fishery benefits.
-
-## Q2 - Please highlight the limitations and strengths.
-
-The question at the heart of this manuscript is an interesting one because the combination of TURFS and reserves offers the possibility of securing rights for fishers while also implementing conservation areas, thereby potentially avoiding some of the conflicts associated with exclusion from fishing grounds that can arise with MPA designation.
-However, I didn’t feel that the paper fully lived up to its potential in addressing this question (see Q3).
-Beyond the key findings as described above (Q1), there is a lot of additional material in the manuscript that didn’t seem to add much to the key message.
-If the focus on TURF-reserves is the key area of novelty, I think this should be further emphasised and explained – see more detailed comments below.
-
-## Q3 - Please provide your detailed review report to the editor and authors (including any comments on the Q4 Check List):
 
@@
 In line with this the last sentence of the abstract seems to be speculation, and is not supported by evidence presented in the manuscript.
+
+*A particular characteristic of Fish Refuges is that they are implemented with a pre-determined expiration date.
+Reserves are only re-implemented if the fishers decide to do so.
+We view this willingness to renew the reserves as a continued support.
+We have expanded in the description of the case studies and discussion to explain the reserve implementation process and why we think that these reserves continue to be supported.
+Furthermore, the work by Ayer et al., 2018 explicitly looked at the perceptions that these communities have about these reserves, and their findings support our statements.
+We have modified the last sentence of the manuscript to present a solid conclusion about the recommendations for the reserves we evaluate, but without speculating about fisher's perception of marine conservation.*
 
@@
 
-- In the third paragraph it wasn’t clear to me whether the authors see TURF-reserves as synonymous with community-based marine reserves, or if not what the differences are.
-As the authors note, who these rights are conferred and recognised by is an important consideration and needs further explanation.
+- In the third paragraph it wasn't clear to me whether the authors see TURF-reserves as synonymous with community-based marine reserves, or if not what the differences are.
+As the authors note, who these rights are conferred and recognized by is an important consideration and needs further explanation.
 In line with this the authors should perhaps revisit the title, which refers to community-based marine reserves, but not to TURFs.
@@
 - I suggest moving the specific details of Mexican TURFs to the case study section to provide context for the study sites, and focus on the broader issues in the introduction.
+
+*The points raised above were amongst the most useful comments, and the fact that these were highlighted by both reviewers suggest that there was room for improvement.
+We have significantly expanded the introduction by including more background on TURFs, reserves, and TURF reserves.
+We have re-written some paragraphs to highlight the fact that TURF-reserve effectiveness has been proposed from theoretical and modelling work, but that no empirical evidence exists for reserves*
+
+*We have also included a brief motivation for the study based on Mexican regulation, but moved the bulk of the case studies to the Methods section.*
 
@@
 
-- The claims that the communities are representative and that insights could be generalised to elsewhere in the world could be toned down.
-Some evidence that the communities share similarities with other fishing communities is useful, but the authors themselves suggest in the discussion that wide generalisations are not possible.
+*We have included more information on the wider context, touching on implementation processes and TURF history.*
+
+- The claims that the communities are representative and that insights could be generalized to elsewhere in the world could be toned down.
+Some evidence that the communities share similarities with other fishing communities is useful, but the authors themselves suggest in the discussion that wide generalizations are not possible.
+
+*We have toned down the generalizations and made distinctions between our claims of when to generalize.
+Specifically, we meant to say that one may generalize our findings to other TURF-managed fisheries where community-based reserves are implemented.
+Further down in the discussion, we meant to say that no generalizations should be made to other Fish Refuges not implemented in TURF-managed fisheries.
+We believe that our current description and expanded introduction help make this more clear.*
 
@@
 
-- With regard to the evaluation of SES – it wasn’t clear to me what the purpose of this evaluation was, and the data sources (and analysis of these) seemed sparse.
+*We have included more information on data collection as supplementary materials, where we show the number of transects for each site (reserve - control) and type of survey (invertebrate - fish).
+The process of site selection is presented in the methods, where we expanded the definition of our control sites (L142)*
+
+- With regard to the evaluation of SES – it wasn't clear to me what the purpose of this evaluation was, and the data sources (and analysis of these) seemed sparse.
 There is little explanation of why these factors are important in relation to TURF-reserves, and only minimal reference to the original work that led to this framework.
+
+*We have explicitly mentioned the objective of evaluating the SES in our methods section (L134), and further explained variable selection in L203.*
 
@@
 
+*We have expanded on the methods section and further explained the Difference-in-Differences method (L178).
+We have also included a reference to the main paper (Card and Krueger, 1994) where the DiD methodology was first introduced.*
+
 ### RESULTS
@@
 - Results are concisely presented. The key findings are clear but because the figures only show effect size it’s not easy to get a feel for the data and its variability across sites and controls.
+
+*Our goal is to report the effect of the reserve on the indicator, and using effect sizes allows us to reduce the ambiguity that temporal and spatial trends may induce.
+However, we recognize the importance of observing "raw" data, and have included time series of all indicators in the supplementary materials.*
 
@@
 
+*The interpretation is correct, we intend to provide enough information to show that the characteristics of the system are conductive of effective resource management.
+The reason why similar broad descriptions are presented across sites is because these don't vary across them.
+For example, all sites must align with the same official fisheries regulation (Lobster Fishery Official Mexican Norm; NOM-006-pesc-1993).
+Our end goal is to present the information as a contextual summary.
+The use of the SES framework simply allows us to be more objective by being transparent about the variables we report across communities and using a common language.
+We believe this is now made clear in sentence starting in L134.*
+
 ### DISCUSSION
@@
 There are a number of plausible explanations presented that are all well established in the MPA literature (reserve size, duration, quality of existing fisheries management and environmental variability).
-While I appreciate the importance of publishing negative results in the MPA literature, it’s a shame that some of these factors weren’t anticipated and explored in more detail.
+While I appreciate the importance of publishing negative results in the MPA literature, it’s a shame that some of these factors weren't anticipated and explored in more detail.
 
@@
 
+*The main objective of this paragraph is to justify the use of our method, and highlight the advantages of using causal inference techniques.
+We have removed references to previous studies that used different methodologies because these distracted from the main point.
+We believe the paragraph now addresses three key messages: 1) There is a need for evidence-based conservation and the use of causal inference techniques, 2) Our study uses robust inference methods, which allow us to talk about the causal effect of reserve implementations, and 3) Provide a benchmark comparison by highlighting patterns that our analyses identifies and relate them to previously identified patterns.*
+
 - The latter part of the discussion focuses on the additional potential benefits of reserves including the positive perceptions of fishers noted in another study.
-My view is that these benefits are over-emphasised here, given that there is no data in the manuscript to support them.
+My view is that these benefits are over-emphasized here, given that there is no data in the manuscript to support them.
 Furthermore the potential counter-arguments (e.g.
 that resources invested in these reserves might be put to better use) are not considered, which leads to a fairly one-sided view.
+
+*We believe that part of the disconnect had to do with the way in which the previous text was presented, where we failed to distinguish between the evaluated TURF-reserves, other non-evaluated TURF-reserves and other reserves not implemented within TURFs.
+We have also made it clear that some of the claims (which include citations) come from conclusions of other works performed on these same sites.
+For example, the buffering to environmental variation comes from Micheli et al., 2012, who evaluated the TURF-reserves in Isla Natividad.*
 
@@
 
-## Q4 - Please comment on the methods, results and data interpretation. If there are any objective errors, or if the conclusions are not supported, you should detail your concerns.
-
-Please see Q3
-
-## Q5 - Check List
-
-### Is the English language of sufficient quality?
-
-- Yes
+*We have included a paragraph on some of these limitations in the discussion (L271)*
 
@@
 
-#### Does the reference list cover the relevant literature adequately and in an unbiased manner?
-
-- Yes
-
-#### Are the statistical methods valid and correctly applied? (e.g. sample size, choice of test)
-
-No answer given.
+*Figures and table have been modified or added to the manuscript.*
 
@@
 
+*We have expanded on the methods and expect these to be now better documented.*
+
 #### Are the data underlying the study available in either the article, supplement, or deposited in a repository?
@@
 - No
+
+*Our data and code are deposited in a public GitHub Repository. We have included a sentence in the methods sections mentioning this and a link to said repository (L201).*
+
+\clearpage
 
@@
 Besides the biological analysis, the authors also evaluate the social-ecological system at each study community based on selected indicators.
-They conclude that in order to maximise biological effectiveness, community-based TURF reserves need to have full community support as well as meet other reserve design criteria, such as appropriate size and location.
+They conclude that in order to maximize biological effectiveness, community-based TURF reserves need to have full community support as well as meet other reserve design criteria, such as appropriate size and location.
 
@@
 Where would each study site fall along the gradient of underfished to overfished stocks?
+
 By doing this you can assess each site in terms of biological performance and SES conditions.
@@
 This to me, is where the study can make a contribution to the literature.
+
+*The social evaluation is, indeed, secondary to this study.
+While we believe it is equally important, we assess reserve effectiveness based on the stated objectives, which were to increase lobster densities and landings.
+We incorporate the SES approach to provide more background about the system under which the reserves are placed, and to discard the usually-mentioned causes of lack of effectiveness which "blame" fishers (namely poaching and lack of enforcement).
+Instead, our approach shows that the most likely cause of lack of effectiveness is poor design, that ignore basic design principles such as size of reserves.*
 
@@
 
+*We believe that the re-structuring of the introduction and case study descriptions have now addressed this.*
+
 - Does fishing occur inside the marine reserves, or outside the marine reserves?
@@
 
+*No fishing activities occur within the Marine Reserves, and we have now made it clear that these are no-take marine reserves.*
+
 - Were the fisheries at each study site managed as TURFs prior to the establishment of the marine reserves?
 
+*Yes, these fisheries have been historically managed with TURFs. We have included this in the case study description (sentence starting in L120).*
+
 - L67. What is a triple bottom line evaluation of the effectiveness of community based marine reserves? Please clarify triple bottom line and what the criteria for this are.
+
+*We were using the term "triple bottom line evaluation" in the same way as Halpern 2013 did.
+We realize that this is not clear, and have now explicitly stated that we "provide a holistic evaluation of the effectiveness of community-based in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop" (L65).*
 
@@
 
+*The sizes were included in Table 2 (RS3 - Size of resource system: reserve area), but are now also mentioned in the results (L248).
+We do not directly assess the effect of reserve size because all reserves are smaller than the suggested design principles (Green et al., 2017).
+If we had data from more sites and across a wider range of reserve sizes (i.e.
+smaller and greater than minimum requirements) it would be interesting to show how time to recovery or effect sizes relate to reserve size.
+One would expect to find a positive correlation.*
+
 - Are the control sites managed with TURFs even though they are non-reserves?
+
+*Yes, all control sites are placed within the TURFs, as it is now explicitly mentioned in the case study section (list starting in L141).*
 
@@
 
+*Mean ex-vessel prices for lobster are now presented in Table S3*
+
 - Did you control for difference in prices between sites?
 
+*Yes, our Difference-in-Differences analyses controls for community-specific changes.*
+
 - Would observed differences in revenue inside/outside reserves be due to difference in landings or due to price differences?
+
+*Differences in revenue may be explained by both, a simple increase in prices due to market interactions or an increase in landings (plausibly because of the reserve).
+However, we do not observe an increase in revenues nor in landings, and ex-vessel prices remain relatively constant.*
 
@@
 
+*These were also requested by Reviewer #1 and have now been included in the Supplementary Materials.*
+
 - What was the species and community effect on biological and socioeconomic indicators, respectively?
+
+*We did not directly evaluate the community- or species-level effects on the indicators.
+Since we fit one model for each combination of indicator-community, we cannot make inference about the community- or species-level effects.
+This type of analyses would be more suited for a community ecology approach, or when seeking differential response among species (e.g.
+Babcock et al, 2010).
+Furthermore, we believe that this approach should be used in a post hoc fashion for reserves that have actually had an effect on the indicators, seeking to understand which species and communities may be driving this.*
 
@@
 
-- RU5 – Number of targeted species is the variable, but this isn’t specified in the narrative.
+- RU5 – Number of targeted species is the variable, but this isn't specified in the narrative.
 In addition to noting that lobster is the main target species please enter the total number of targeted species.
+
+*We have included the total number of species targeted*
 
@@
 
-- A3 – Level of isolation. The narrative says that fishing grounds and reserves are away from dense urban centers – can you provide the distance from each study site to the nearest urban centre (km).
+*This assumption is correct, we only list the number of fishing co-op members.
+However, all fishers that fish within the TURF belong to the cooperative, and there are no "free fishers".
+This point is now mentioned in the case study section (L90).*
+
+- A3 – Level of isolation. The narrative says that fishing grounds and reserves are away from dense urban centers – can you provide the distance from each study site to the nearest urban center (km).
+
+*We have included the linear distance to the closes urban centers in Table 2.
+Isla Natividad lies 545 Km south from Tijuana, while Maria Elena and Punta Herrero are more than 200 km south from Cancun.*
 
@@
 
+*All rules are applied for all sites. We have changed the wording from "these include" to "these are" in Table 2.*
+
 - GS9.1 Social monitoring – Do fishing coops at all 3 sites have the same monitoring and enforcement regime in place?
+
+*Yes, they have the same regime known as "Consejo de vigilancia". We have now included this particular wording in Table 2.*
 
@@
 
+*Yes, each community engages in biological monitoring (the data for this study comes from this) and oceanographic monitoring.*
+
 - GS 10.1 Graduated sanctions – Again, which of the penalties are relevant to each site?
 
+*The sanctions and penalties are homogeneous across sites. However, sanctions are administered on a case-by-case basis and voted or agreed by all cooperative members before being applied.*
+
 - As it is presented now, the governance variables are a bit unclear as to which type of rules/regulations are practiced at each of the three sites.
-I recognise that this is because these variables are more qualitative in nature.
+I recognize that this is because these variables are more qualitative in nature.
 However to make it clearer, the Narrative column could be split into 3 - one column per study site, and the relevant information entered under each site column.
+
+*We believe that splitting the narrative column into three would be beneficial if each community had very different values / narratives.
+However, these three communities have strikingly similar values.
+Splitting it into three columns would result in redundant text being repeated across columns.
+However, we do understand the point that it is difficult to recognize which values correspond to each community.
+To address this, we have modified the table caption to explain that the values apply to all communities, and when needed a specific community-level narrative is included (for example, isolation of IN vs PH and ME).*
 
@@
 
+*This is a fair point, which was raised previously.
+This would be an interesting approach.
+Initially, we had "operationalized" the SES framework and included a scoring system.
+However, since all sites operate under very similar settings scores were practically the same and obtained little contrast between them.
+We believe that this happens because the three communities represent well-managed fisheries, and all of them achieve high scores.
+The approach suggested would be much more interesting if it could be applied to a series of fisheries along a wider gradient of cooperation and co-management.
+We have included a discussion paragraph recognizing this (L330).*
+
 - Fig 2B and 2D. Why are there big fluctuations in fish biomass from Years 7-10, but density stays pretty much the same for the same time period?
+
+*This big fluctuations in biomass are explained by an opportunistic increase in sizes.
+In this particular case, a few larger-than-usual organisms were observed, but the total number of fish remained similar to other years.
+As a result, the few larger fish observed these years yield a higher biomass.
+However these differences are not significantly different from the pre-observed trend.*
 
@@
 
-- L208-209 – The authors state that “Understanding the socio-ecological context ….provide insights as to why this happens”, but then don’t provide further elaboration.
+*After the first few years of reserve implementation, one would expect landings to decrease.
+This would be explained by the relative reduction of total fishable area.
+The "dip" described in Ovando et al., (2016) would then recover as densities inside the reserve recover and spillover occurs.
+The fact that we don't see a reduction in landings over the first few years suggests that reserves are not big enough to reduce landings over the first few years.*
+
+- L208-209 – The authors state that “Understanding the socio-ecological context...
+provide insights as to why this happens”, but then don’t provide further elaboration.
 The reader is not getting any insight as to why the communities continue to support the reserves.
@@
 
+*This lose statement has been removed after the re-structuring the discussion.*
+
 - Clearly state the lessons learned. L274 simply says that “Lessons learned from these cases can guide implementation…” But what are these lessons?
+
+*We have removed this lose statement and included an explicit conclusions paragraph (L335) where we state that "We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside".*
 
@@
 
+*The previous wording of this statement was confusing, and seemed like we were referencing the reserves evaluated in this work.
+Clearly this is not true and our data would not support such claim.
+As with previous statements, we believe this is now clear after re-structuring the discussion.*
+
 - L268-269 “Furthermore, the lack of effectiveness observed in these reserves should not be generalizable to other reserves established under the same legal framework”.
 This appear to contradict L99-100 which says that “it is safe to cautiously generalize our insights to other similar reserves…”
+
+*Similar points were raised by Reviewer #1.
+We believe our current wording and expansion of study sites now makes clear when we refer to the evaluated TURF-reserves, other TURF-reserves and other reserves outside TURFs.
+This same idea that other reserves may be effective is now shown in a new paragraph (L330).*
 
@@
 
+*The expansion of the study sites and description of Communities / TURFs / Fisheries in these areas may now provide more context to avoid confusion.*
+
 - The Discussion is focused mainly on ecological factors which may have affected the effectiveness of the TURF-reserves.
@@
 
+*Our SES approach is intended to provide a snapshot of the state of each community.
+We do not include time-varying values for any indicator because all these have remained constant at least since reserve implementation.
+The only addition to their community-based management have been the biological and oceanographic monitoring (GS9.2) which started at least one year before reserve implementation.
+As per market drivers that may increase fishing pressure, our supplementary materials show little change in prices before / after the implementation of the reserves, and it is unlikely that TURF-managed fisheries (plausibly managed at optimum) would deviate from this already optimal management point.
+However, we believe some of these points have now been covered in our discussion (L271).*
+
 ### Minor comments:
@@
 
-## Q4 - Please comment on the methods, results and data interpretation. If there are any objective errors, or if the conclusions are not supported, you should detail your concerns
+*The figure caption has been corrected and now says "circles"*
 
-- These comments are covered in Q3 above.
+place-holder
 
-## Q5 - Check List
-
-### Is the English language of sufficient quality?
-
-- Yes
-
-### Is the quality of the figures and tables satisfactory?
-
-- Yes
-
-### Does the reference list cover the relevant literature adequately and in an unbiased manner?
-
-- Yes
-
-### Are the statistical methods valid and correctly applied? (e.g. sample size, choice of test)
-
-- Yes
-
-### Are the methods sufficiently documented to allow replication studies?
-
-- Yes
-
-### Are the data underlying the study available in either the article, supplement, or deposited in a repository?
-(Sequence/expression data, protein/molecule characterizations, annotations, and taxonomy data are required to be deposited in public repositories prior to publication) (yes/no/not applicable)
-
-- Yes
+*This would be an interesting approach.
+Initially, we had "operationalized" the SES framework and included a scoring system.
+However, since all sites operate under very similar settings scores were practically the same and obtained little contrast between them.
+We believe that this happens because the three communities represent well-managed fisheries, and all of them achieve high scores.
+The approach suggested would be much more interesting if it could be applied to a series of fisheries along a wider gradient of cooperation and co-management.
+We have included a discussion paragraph recognizing this (L330).*

### 2018-12-31 docs/SupplementaryMaterial.Rmd
Made last changes and re-submitted

@@
 ---
-title: "Effectiveness of community-based TURF-reserves in small-scale fisheries"
+title: "Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries"
 output:

### 2018-12-31 docs/manuscript.Rmd
Made last changes and re-submitted

@@
 ---
-title: "Effectiveness of community-based TURF-reserves in small-scale fisheries"
+title: "Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries"
 rtitle: "Community-based TURF-reserves"
@@
 Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage.
-Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
-In the last 12 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a 2014 regulation; their effectiveness has not been formally evaluated.
+Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
+In the last 12 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2014; their effectiveness has not been formally evaluated.
 We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
@@
 
-TURFs are a fisheries management tool in which a well defined group of fishers have exclusive access to an explicitly delimited portion of the ocean.
+TURFs are a fisheries management tool in which a well defined group of fishers (*e.g.* fishing cooperatives) have exclusive access to an explicitly delimited portion of the ocean.
 They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,costello_2010,mccay_2014}.
 On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits.
-These can be implemented to protect biodiversity but also as fishery management tools that restrict fishing effort and gears and therefore aid in the recovery of marine stocks.
+These can be implemented to protect biodiversity but also as fishery management tools to aid in the recovery of marine stocks.
 These instruments can be combined by establishing a marine reserve within a TURF, thus making them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.
@@
 Although in theory these systems are successful \citep{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success.
-
-TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{gelcich_2015}.
-This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009}.
-There is a growing body of literature focusing on the relations between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}.
-However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.
+Moreover, TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{ostrom_2009,gelcich_2015}.
+It is therefore important to consider not only the indicators of interest, bu also the governance settings under which the reserves occur.
 
@@
 In Mexico, a norm created in 2014 allows fishers to request legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}).
-Since 2012, old and new marine reserves have gained legal recognition as Fishing Refuges.
-Of these, 18 were originally implemented as within TURFs.
+Since 2012, old and new marine reserves have gained legal recognition as Fish Refuges.
+Of these, 18 were originally implemented within TURFs.
 However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
@@
 Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico.
-These three case studies span a range of ecological and social conditions representative of different regions of Mexico.
 The objective of this work is twofold.
@@
 
-In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue (*i.e.* Fish Refuges; \citet{nom}).
+In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue: Fish Refuges \citep{nom}.
 Fish Refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
 One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporal marine reserves within their TURFs with a defined expiration date (often 5 years).
-After these five years have passed, fishers have the opportunity to open the reserves to fishing or continue to have them.
+When the expiration date is reached, fishers can chose to open the reserves to fishing or re-establish them.
 Our work focuses on Fish Refuges implemented as community-based TURF-reserves that occur in small-scale fisheries.
@@
 Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (*i.e.* TURFs shown as blue polygons in Fig \ref{fig:map}A).
-Each TURF is elusively fished by one cooperative, and each community usually hosts no more than one cooperative.
+Each TURF is exclusively fished by one cooperative, and each community usually hosts no more than one cooperative.
 The profits from each TURF are shared amongst all fishers from the cooperative.
@@
 The main fishery in the island is the spiny lobster (*Panulirus interruptus*), but other resources like finfish, sea cucumber, read sea urchin, snail, and abalone are also an important source of income.
-In 2006, the community decided to implement two marine reserves within their fishing grounds to protect commercially important invertebrate species; mainly lobster and abalone.
+In 2006, the community decided to implement two marine reserves within their fishing grounds.
+The objective of these reserves was "to protect commercially important invertebrate species"; mainly lobster and abalone.
 These reserves obtained legal recognition in 2018 \citep{dof_website_2018}.
@@
 Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively.
-These reserves have been legally recognized as Fishing Refuges since their original implementation \citep{dof_website_2012,dof_website_2013} and re-enactments \citep{dof_website_2017b}.
+These reserves have been legally recognized as Fishing Refuges since their original implementation \citep{dof_website_2012,dof_website_2013} and subsequent re-establishments \citep{dof_website_2017b}.
 
@@
 The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather help provide context on the social-ecological system in which reserves develop.
-The following two sections describe our data collection methods and analyses of biological and socioeconomic indicators as well as the SES analysis.
+The following two sections describe our data collection methods and analyses.
 
@@
 Ecological data come from the annual ecological monitoring of reserve and control sites.
-Reserve sites are those within the reserves, and thus no fishing takes place.
+Reserve sites are areas where no fishing occurs.
 Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) Are not directly adjacent to the reserves.
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
-This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally attribute the changes to the reserve \citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.
+This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally attribute the changes to the reserve \citep{francinifilho_2008,Villasenor-Derbez_2018}.
 
@@
 Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017}.
-A table with summary statistics for this data is provided in the supplementary materials **(Table S3, Figure S5)**.
+A table with summary statistics for this data is provided in the supplementary materials (Table S3, Figure S5).
 
-Data for the evaluation of the SES were collected at the community-level from official documents used in the creation and designation of the marine reserves.
-These includes the technical studies that the cooperatives submit when they request recognition of their reserves, as well as the official enactments  \citep{dof_website_2012,dof_website_2013,dof_website_2018}.
+Data for the evaluation of the SES were collected at the community-level from official documents used in the design, creation, and designation of the marine reserves.
+These include the technical studies that the cooperatives submit when they request recognition of their reserves, as well as the official enactments  \citep{dof_website_2012,dof_website_2013,dof_website_2018}.
 We also complimented information based on the authors' experience and knowledge of the communities.
 We collected information on the Resource Systems, Resource Units, Actors, and Governance System (Table \ref{table:ses}).
-The next sections further describe the variables and indicators chosen for each of these.
+The next section further describe the variables and indicators chosen for each of these.
 
@@
 \begin{equation}
-I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t,j}
+I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t}
 \label{eqn:reg_bio}
@@
 The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).
-Therefore, we would expect this term to be positive if the indicator increases because of the reserve implementation.
-Finally, $\epsilon_{i,t,j}$ represents the error term of the regression.
+Therefore, we would expect this term to be positive if the indicator increases because of the reserve.
+Finally, $\epsilon_{i,t}$ represents the error term of the regression.
 
@@
 Neighboring communities are used as counterfactuals that allow us to control for unobserved time-invariants.
-Each focal community (Isla Natividad and Maria Elena) has three counterfactual communities.
+Each focal community (*i.e.* Isla Natividad and Maria Elena) has three counterfactual communities.
 
 \begin{equation}
-I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t,j}
+I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t}
 \label{eqn:soc_reg}
@@
 These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \citep{depalma_2018}.
-All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}.
+We fit one model per community and indicators (*e.g.* there are three models for lobster density, one for each community) for a total of 12 biological model fits and four socioeconomic model fits.
+Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard errors \citep{zeileis_2004-7n}.
 All analyses were performed in R `r paste0("version ", version$major, ".", version$minor)` and R Studio version 1.1.456 \citep{R_2018}.
-All data and code needed to reproduce our analyses are available in a [GitHub repository](https://github.com/jcvdav/ReserveEffect).
+All data and code needed to reproduce our analyses are available in a GitHub repository at: [https://github.com/jcvdav/ReserveEffect](https://github.com/jcvdav/ReserveEffect).
 
@@
 While many ecology studies have used BACI sampling designs and respective analyses (*e.g.* \citet{stewartoaten_1986}), few conservation studies have done so to evaluate the effect of an intervention (*e.g.* \citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted in a call for more robust analyses in conservation science \citep{guidetti_2002,ferraro_2006}.
-Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness.
-For example, we capture previously described patterns like the rapid increase observed for lobster densities in Isla Natividad on the sixth year (*i.e.* 2012; Fig.
-\ref{fig:indicators}A), a year after the hypoxia events described by \citet{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
-Yet, our empirical approach assumes control sites are a plausible counterfactual for treated sites.
-This implies that treated sites would have followed the same trend as control sites, had the reserves not been implemented.
-Nonetheless, temporal trends for each site don't show any significant increases (Figures S1-S4), supporting our findings of lack of change in the indicators used.
+Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness, and captures previously described patterns.
+For example, the rapid increase observed for lobster densities in Isla Natividad on the sixth year (*i.e.* 2012; Fig.
+\ref{fig:indicators}A), occurs a year after the hypoxia events described by \citet{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
+The use of causal inference techniques may help us support evidence-based conservation.
 
@@
 A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve.
-For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and higher profitability.
+For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and thus higher profitability.
 
@@
 For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2004,hilborn_2006}.
-Both lobster fisheries were certified by the Marine Stewardship Council and lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \citep{dof_website_1993}.
+Both lobster fisheries were certified by the Marine Stewardship Council and are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \citep{dof_website_1993}.
 It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves; reserves implemented in TURFs where fishing pressure is already optimally managed will still show a trade-off between fisheries and conservation objectives \citep{lester_2017}.
@@
 This might reduce the difference between indicators from the TURF and reserve sites, making it difficult to detect such a small change.
-Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of the reserves might be larger.
+Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of the reserves might be greater.
 
@@
 Self-regulation of fishing effort can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
-Furthermore, embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital as reported by \citet{fulton_2019}.
+Furthermore, embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}.
 Showing commitment to marine conservation and sustainable fishing practices has allowed fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
 These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-implementing the reserves \citep{ayer_2018}.
-
-In terms of the fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs.
-Future research should aim at evaluating other Fish Refuges that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through top-down processes (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.
 
@@
 Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.
-For the particular case of the ones we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
+Furthermore, consvation and advocacy groups should consider the opportunity costs of such interventions (*sensu* \citet{smith_2010}) and evaluate the potential of other approaches that may yield similar benefits.
+
+In terms of the fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs.
+Future research should aim at evaluating other Fish Refuges established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through top-down processes (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.
+For the particular case of the reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
 
@@
 This study would have not been possible without the effort by members of the fishing communities here mentioned, who participated in the data-collection process.
+The authors wish to acknowledge comments by the reviewrs and editor, which significantly improved the quality of this work.
 

### 2018-12-31 docs/manuscript.tex
Made last changes and re-submitted

@@
 
-\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in small-scale fisheries}
+\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in Mexican small-scale
+fisheries}
 
@@
 great challenges since they are difficult to monitor, enforce, and
-manage. Combining territorial user rights for fisheries (TURF) with
+manage. Combining territorial use rights for fisheries (TURF) with
 no-take marine reserves to create TURF-reserves can improve the
@@
 years, 18 old and new community-based Mexican TURF-reserves gained legal
-recognition thanks to a 2014 regulation; their effectiveness has not
-been formally evaluated. We combine causal inference techniques and the
-Social-Ecological Systems framework to provide a holistic evaluation of
-community-based TURF-reserves in three coastal communities in Mexico. We
-find that while reserves have not yet achieved their stated goal of
-increasing the density of lobster and other benthic invertebrates, they
-continue to receive support from the fishing communities. A lack of
-clear ecological and socioeconomic effects likely results from a
-combination of factors. First, some of these reserves might be too young
-for the effects to show. Second, the reserves are not large enough to
-protect mobile species, like lobster. Third, variable and extreme
-oceanographic conditions have impacted harvested populations. Fourth,
-local fisheries are already well managed, and it is unlikely that
-reserves might have a detectable effect in catches. However, these
-reserves may provide a foundation for establishing additional, larger
-marine reserves needed to effectively conserve mobile species.
+recognition thanks to a regulation passed in 2014; their effectiveness
+has not been formally evaluated. We combine causal inference techniques
+and the Social-Ecological Systems framework to provide a holistic
+evaluation of community-based TURF-reserves in three coastal communities
+in Mexico. We find that while reserves have not yet achieved their
+stated goal of increasing the density of lobster and other benthic
+invertebrates, they continue to receive support from the fishing
+communities. A lack of clear ecological and socioeconomic effects likely
+results from a combination of factors. First, some of these reserves
+might be too young for the effects to show. Second, the reserves are not
+large enough to protect mobile species, like lobster. Third, variable
+and extreme oceanographic conditions have impacted harvested
+populations. Fourth, local fisheries are already well managed, and it is
+unlikely that reserves might have a detectable effect in catches.
+However, these reserves may provide a foundation for establishing
+additional, larger marine reserves needed to effectively conserve mobile
+species.
 
@@
 TURFs are a fisheries management tool in which a well defined group of
-fishers have exclusive access to an explicitly delimited portion of the
-ocean. They promote a sense of stewardship and incentivise resource
-users to sustainably manage their resources
-\citep{gelcich_2008,costello_2010,mccay_2014}. On the other hand,
-no-take marine reserves (marine reserves from hereinafter) are areas
-where all extractive activities are off-limits. These can be implemented
-to protect biodiversity but also as fishery management tools that
-restrict fishing effort and gears and therefore aid in the recovery of
-marine stocks. These instruments can be combined by establishing a
-marine reserve within a TURF, thus making them TURF-reserves
-\citep{afflerbach_2014,gelcich_2015,lester_2017}.
+fishers (\emph{e.g.} fishing cooperatives) have exclusive access to an
+explicitly delimited portion of the ocean. They promote a sense of
+stewardship and incentivise resource users to sustainably manage their
+resources \citep{gelcich_2008,costello_2010,mccay_2014}. On the other
+hand, no-take marine reserves (marine reserves from hereinafter) are
+areas where all extractive activities are off-limits. These can be
+implemented to protect biodiversity but also as fishery management tools
+to aid in the recovery of marine stocks. These instruments can be
+combined by establishing a marine reserve within a TURF, thus making
+them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
@@
 empirical evidence of their effectiveness and the drivers of their
-success.
-
-TURF-reserve systems are inherently intricate social-ecological systems,
-and their effectiveness must depend on how environmental and social
-factors combine and interact \citep{gelcich_2015}. This is especially
-important in social-ecological coastal systems dominated by close
-interactions and feedbacks between people and natural resources
-\citep{ostrom_2009}. There is a growing body of literature focusing on
-the relations between socioeconomic and governance structures and
-reserve effectiveness
-\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}.
-However, to our knowledge, no studies exist that evaluate TURF-reserves
-from both a social and ecological perspective.
+success. Moreover, TURF-reserve systems are inherently intricate
+social-ecological systems, and their effectiveness must depend on how
+environmental and social factors combine and interact
+\citep{ostrom_2009,gelcich_2015}. It is therefore important to consider
+not only the indicators of interest, bu also the governance settings
+under which the reserves occur.
 
@@
 ``Fish Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since
-2012, old and new marine reserves have gained legal recognition as
-Fishing Refuges. Of these, 18 were originally implemented as within
-TURFs. However, their effectiveness has not yet been formally evaluated
-and reported in the scientific literature.
+2012, old and new marine reserves have gained legal recognition as Fish
+Refuges. Of these, 18 were originally implemented within TURFs. However,
+their effectiveness has not yet been formally evaluated and reported in
+the scientific literature.
 
@@
 Systems (SES) framework to evaluate community-based TURF-reserves in
-three coastal communities in Mexico. These three case studies span a
-range of ecological and social conditions representative of different
-regions of Mexico. The objective of this work is twofold. First, to
-provide a holistic evaluation of the effectiveness of community-based
-TURF-reserves in terms of the changes in biological and socioeconomic
-indicators and the governance settings under which these develop, which
-may inform similar processes in other countries. Second, to identify
-opportunities where improvement or adjustment might lead to increased
-effectiveness. We draw from lessons learned in these three case studies
-and provide management recommendations to maximize the effectiveness of
-community-based TURF-reserves in small-scale fisheries where this tool
-is used to manage and rebuild their coastal fisheries.
+three coastal communities in Mexico. The objective of this work is
+twofold. First, to provide a holistic evaluation of the effectiveness of
+community-based TURF-reserves in terms of the changes in biological and
+socioeconomic indicators and the governance settings under which these
+develop, which may inform similar processes in other countries. Second,
+to identify opportunities where improvement or adjustment might lead to
+increased effectiveness. We draw from lessons learned in these three
+case studies and provide management recommendations to maximize the
+effectiveness of community-based TURF-reserves in small-scale fisheries
+where this tool is used to manage and rebuild their coastal fisheries.
 
@@
 Organizations (CSOs) served as a link between fishers and government,
-and created a legal framework that solves this governance issue
-(\emph{i.e.} Fish Refuges; \citet{nom}). Fish Refuges can be implemented
-as temporal or partial reserves, which can protect one, some, or all
-resources within their boundaries. One of the ways in which fishing
-communities have taken advantage of this new tool is by implementing
-temporal marine reserves within their TURFs with a defined expiration
-date (often 5 years). After these five years have passed, fishers have
-the opportunity to open the reserves to fishing or continue to have
-them. Our work focuses on Fish Refuges implemented as community-based
-TURF-reserves that occur in small-scale fisheries.
+and created a legal framework that solves this governance issue: Fish
+Refuges \citep{nom}. Fish Refuges can be implemented as temporal or
+partial reserves, which can protect one, some, or all resources within
+their boundaries. One of the ways in which fishing communities have
+taken advantage of this new tool is by implementing temporal marine
+reserves within their TURFs with a defined expiration date (often 5
+years). When the expiration date is reached, fishers can chose to open
+the reserves to fishing or re-establish them. Our work focuses on Fish
+Refuges implemented as community-based TURF-reserves that occur in
+small-scale fisheries.
 
@@
 delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
-\ref{fig:map}A). Each TURF is elusively fished by one cooperative, and
+\ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and
 each community usually hosts no more than one cooperative. The profits
@@
 2006, the community decided to implement two marine reserves within
-their fishing grounds to protect commercially important invertebrate
-species; mainly lobster and abalone. These reserves obtained legal
-recognition in 2018 \citep{dof_website_2018}.
+their fishing grounds. The objective of these reserves was ``to protect
+commercially important invertebrate species''; mainly lobster and
+abalone. These reserves obtained legal recognition in 2018
+\citep{dof_website_2018}.
 
@@
 Refuges since their original implementation
-\citep{dof_website_2012,dof_website_2013} and re-enactments
-\citep{dof_website_2017b}.
+\citep{dof_website_2012,dof_website_2013} and subsequent
+re-establishments \citep{dof_website_2017b}.
 
@@
 reserves develop. The following two sections describe our data
-collection methods and analyses of biological and socioeconomic
-indicators as well as the SES analysis.
+collection methods and analyses.
 
@@
 Ecological data come from the annual ecological monitoring of reserve
-and control sites. Reserve sites are those within the reserves, and thus
-no fishing takes place. Control sites are areas that meet the following
-criteria: i) habitat characteristics are similar to the corresponding
-reserves, ii) presumably had a similar probability of being selected as
-reserves during the design phase, iii) are located within the TURF,
-where fishing occurs, and iv) Are not directly adjacent to the reserves.
-We focus our evaluation on sites where data are available for reserve
-and control sites, before and after the implementation of the reserve.
-This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
-sampling design that allows us to capture and control for temporal and
-spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally
-attribute the changes to the reserve
-\citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.
+and control sites. Reserve sites are areas where no fishing occurs.
+Control sites are areas that meet the following criteria: i) habitat
+characteristics are similar to the corresponding reserves, ii)
+presumably had a similar probability of being selected as reserves
+during the design phase, iii) are located within the TURF, where fishing
+occurs, and iv) Are not directly adjacent to the reserves. We focus our
+evaluation on sites where data are available for reserve and control
+sites, before and after the implementation of the reserve. This provides
+us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design
+that allows us to capture and control for temporal and spatial dynamics
+\citep{stewartoaten_1986,depalma_2018} and causally attribute the
+changes to the reserve
+\citep{francinifilho_2008,Villasenor-Derbez_2018}.
 
@@
 \citep{oecd_2017}. A table with summary statistics for this data is
-provided in the supplementary materials \textbf{(Table S3, Figure S5)}.
+provided in the supplementary materials (Table S3, Figure S5).
 
 Data for the evaluation of the SES were collected at the community-level
-from official documents used in the creation and designation of the
-marine reserves. These includes the technical studies that the
+from official documents used in the design, creation, and designation of
+the marine reserves. These include the technical studies that the
 cooperatives submit when they request recognition of their reserves, as
@@
 Resource Units, Actors, and Governance System (Table \ref{table:ses}).
-The next sections further describe the variables and indicators chosen
+The next section further describe the variables and indicators chosen
 for each of these.
@@
 would expect this term to be positive if the indicator increases because
-of the reserve implementation. Finally, \(\epsilon_{i,t,j}\) represents
-the error term of the regression.
+of the reserve. Finally, \(\epsilon_{i,t}\) represents the error term of
+the regression.
 
@@
 used as counterfactuals that allow us to control for unobserved
-time-invariants. Each focal community (Isla Natividad and Maria Elena)
-has three counterfactual communities.
+time-invariants. Each focal community (\emph{i.e.} Isla Natividad and
+Maria Elena) has three counterfactual communities.
 
 \begin{equation}
-I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t,j}
+I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t}
 \label{eqn:soc_reg}
@@
 implementation of marine reserves and the observed trends by accounting
-for temporal and site-specific dynamics \citep{depalma_2018}. All model
-coefficients were estimated via ordinary least-squares and
+for temporal and site-specific dynamics \citep{depalma_2018}. We fit one
+model per community and indicators (\emph{e.g.} there are three models
+for lobster density, one for each community) for a total of 12
+biological model fits and four socioeconomic model fits. Model
+coefficients were estimated via ordinary least-squares and used
 heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
@@
 \citep{R_2018}. All data and code needed to reproduce our analyses are
-available in a \href{https://github.com/jcvdav/ReserveEffect}{GitHub
-repository}.
+available in a GitHub repository at:
+\url{https://github.com/jcvdav/ReserveEffect}.
 
@@
 temporal and spatial changes provides a more robust measure of reserve
-effectiveness. For example, we capture previously described patterns
-like the rapid increase observed for lobster densities in Isla Natividad
-on the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), a year
-after the hypoxia events described by \citet{micheli_2012-EU}, which
-caused mass mortality of sedentary organisms such as abalone and sea
-urchins, but not lobster and finfish. Yet, our empirical approach
-assumes control sites are a plausible counterfactual for treated sites.
-This implies that treated sites would have followed the same trend as
-control sites, had the reserves not been implemented. Nonetheless,
-temporal trends for each site don't show any significant increases
-(Figures S1-S4), supporting our findings of lack of change in the
-indicators used.
+effectiveness, and captures previously described patterns. For example,
+the rapid increase observed for lobster densities in Isla Natividad on
+the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), occurs a
+year after the hypoxia events described by \citet{micheli_2012-EU},
+which caused mass mortality of sedentary organisms such as abalone and
+sea urchins, but not lobster and finfish. The use of causal inference
+techniques may help us support evidence-based conservation.
 
@@
 example, if catches remain relatively unchanged but fishing effort
-decreased, that would imply a larger catch per unit effort and higher
-profitability.
+decreased, that would imply a larger catch per unit effort and thus
+higher profitability.
 
@@
 is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
-fisheries were certified by the Marine Stewardship Council and lobster
-fisheries are managed via species-specific minimum catch sizes, seasonal
-closures, protection of ``berried'' females, and escapement windows
-where traps are allowed \citep{dof_website_1993}. It is uncertain
-whether such a well-managed fishery will experience additional benefits
-from marine reserves; reserves implemented in TURFs where fishing
-pressure is already optimally managed will still show a trade-off
-between fisheries and conservation objectives \citep{lester_2017}.
-Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have
-greater biomass and richness than areas operating under open access.
-This might reduce the difference between indicators from the TURF and
-reserve sites, making it difficult to detect such a small change.
-Further research should focus on evaluating sites in the reserve, TURF,
-and open access areas or similar Fish Refuges established without the
-presence of TURFs where the impact of the reserves might be larger.
+fisheries were certified by the Marine Stewardship Council and are
+managed via species-specific minimum catch sizes, seasonal closures,
+protection of ``berried'' females, and escapement windows where traps
+are allowed \citep{dof_website_1993}. It is uncertain whether such a
+well-managed fishery will experience additional benefits from marine
+reserves; reserves implemented in TURFs where fishing pressure is
+already optimally managed will still show a trade-off between fisheries
+and conservation objectives \citep{lester_2017}. Furthermore,
+\citet{gelcich_2008} have shown that TURFs alone can have greater
+biomass and richness than areas operating under open access. This might
+reduce the difference between indicators from the TURF and reserve
+sites, making it difficult to detect such a small change. Further
+research should focus on evaluating sites in the reserve, TURF, and open
+access areas or similar Fish Refuges established without the presence of
+TURFs where the impact of the reserves might be greater.
 
@@
 project can bring the community together, which promotes social cohesion
-and builds social capital as reported by \citet{fulton_2019}. Showing
-commitment to marine conservation and sustainable fishing practices has
-allowed fishers to have greater bargaining power and leverage over
-fisheries management \citep{prezramrez_2012}. These additional benefits
-might explain why communities show a positive perception about their
+and builds social capital \citep{fulton_2019}. Showing commitment to
+marine conservation and sustainable fishing practices has allowed
+fishers to have greater bargaining power and leverage over fisheries
+management \citep{prezramrez_2012}. These additional benefits might
+explain why communities show a positive perception about their
 performance and continue to support their presence by re-implementing
 the reserves \citep{ayer_2018}.
-
-In terms of the fisheries regulation in Mexico, our work only evaluates
-Fish Refuges established within TURFs. Future research should aim at
-evaluating other Fish Refuges that have also been established as
-bottom-up processes but without the presence of TURFs (\emph{e.g.}
-\citet{dof_websiteC_2012}), others established through top-down
-processes (\emph{i.e.} \citet{dof_websiteU_2018}), as well as the
-relationship between governance and effectiveness across this gradient
-of approaches.
 
@@
 TURF-reserves meet essential design principles such as size and
-placement so as to maximize their effectiveness. For the particular case
-of the ones we evaluate, the possibility of expanding reserves or
-merging existing polygons into larger areas should be evaluated and
-proposed to the communities.
+placement so as to maximize their effectiveness. Furthermore, consvation
+and advocacy groups should consider the opportunity costs of such
+interventions (\emph{sensu} \citet{smith_2010}) and evaluate the
+potential of other approaches that may yield similar benefits.
+
+In terms of the fisheries regulation in Mexico, our work only evaluates
+Fish Refuges established within TURFs. Future research should aim at
+evaluating other Fish Refuges established as bottom-up processes but
+without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
+others established through top-down processes (\emph{i.e.}
+\citet{dof_websiteU_2018}), as well as the relationship between
+governance and effectiveness across this gradient of approaches. For the
+particular case of the reserves that we evaluate, the possibility of
+expanding reserves or merging existing polygons into larger areas should
+be evaluated and proposed to the communities.
 
@@
 fishing communities here mentioned, who participated in the
-data-collection process.
+data-collection process. The authors wish to acknowledge comments by the
+reviewrs and editor, which significantly improved the quality of this
+work.
 

### 2018-12-31 docs/reviewer_comments.Rmd
Made last changes and re-submitted

@@
 There is one unsubstantiated statement (line 81) of good enforcement, but could it not be just as likely that there is fishing activity in these reserves?
-Poaching is an equally valid explanation as any examine din the Discussion.
+Poaching is an equally valid explanation as any examined in the Discussion.
 
 *That is correct, a reserve would contain higher density, biomass, and richness.
-We have corrected the abstract to reflect what we intended to say, which is that a well-managed fishery would not see increased catches by an additional intervention.
+We have corrected the abstract to reflect what we intended to say, which is that a well-managed fishery would not see increased aggregate catches by an additional intervention.
 We do not include explicit references in the abstract, but the discussion touches on this topic and cites Hilborn, Micheli and De Leo's paper that evaluates when MRs result in increased catches.*
@@
 *As for the unsubstantiated statement of good enforcement, it is unlikely that these reserves experience poaching.
-The reserves are implemented, managed, and monitored by the community themselves.
-We have expanded on this when we introduce the case studies.*
+The reserves are implemented, managed, and monitored by the communities themselves.
+We have expanded on this when we introduce the case studies, and our SES approch suggests high levels of self-enforcement to protect the TURF and reserves.*
 
@@
 *We have expanded on the methods, especially by adding to the description of case studies (TURFs, reserve implementation process) and data collection methods to clarify what the sampling design is.
-We have also expanded on the regression model.
-It is a multiple linear regression because we incorporate more than one explanatory variable.
+We have also expanded on the regression models.
+It is a multiple linear regression because we incorporate more than one explanatory variable (time, zone, and the interaction between these).
 The lambda that we report does not make reference to the lambda parameter in a Poisson distribution, but rather to the coefficient presented in Equations 1 and 2.
@@
 org / m^-2^) and biomass (Kg / m^-2^), which are continuous variables.*
+
+*Additionally, we realized that our equations were incorrectly formulated.
+They included species- and community-level fixed effects.
+However, our code did not incorporate these terms, as our data does not allow us to include them.
+The equations and text have been modified to reflect our actual analysis.*
 
@@
 *The observation that Moland's paper might have confounded estimates of reserve effectiveness is correct.
-Proximity of their reserve and control sites might imply a spillover effect from the reserves to the controls may occur, which would result in more conservative estimates of reserve effectiveness.
-While there is a long history of time series analyses in ecology, we cite Moland's paper because, to our knowledge, it is the first paper that estimated reserve effectiveness using a difference-in-differences approach.
-We have included references to older implementations of BACI in ecology, as well as the seminal econometric paper that introduces the Difference-in-Differences methodology.*
+Proximity of their reserve and control sites might imply a spillover effect from the reserves to the controls, which would result in downard-biased estimates of reserve effectiveness.
+While there is a long history of time series analyses in ecology, we cite Moland's paper because of the methods, not the results; to our knowledge, it is the first paper that estimated reserve effectiveness using a difference-in-differences approach.
+Furthermore, we have included references to older implementations of BACI in ecology, as well as the seminal econometric paper that introduces the difference-in-differences methodology.*
 
@@
 Reserves are only re-implemented if the fishers decide to do so.
-We view this willingness to renew the reserves as a continued support.
-We have expanded in the description of the case studies and discussion to explain the reserve implementation process and why we think that these reserves continue to be supported.
+We view this willingness to re-implement the reserves as an explicit manifestation of support.
+We have expanded in the description of the case studies and discussion to explain the reserve implementation process and why we think that these reserves continue to be supported, which may help support our point.
 Furthermore, the work by Ayer et al., 2018 explicitly looked at the perceptions that these communities have about these reserves, and their findings support our statements.
-We have modified the last sentence of the manuscript to present a solid conclusion about the recommendations for the reserves we evaluate, but without speculating about fisher's perception of marine conservation.*
+We have modified the text to present a solid conclusion about the recommendations for the reserves we evaluate, but without speculating about fisher's perception of marine conservation.*
 
@@
 
-*The points raised above were amongst the most useful comments, and the fact that these were highlighted by both reviewers suggest that there was room for improvement.
-We have significantly expanded the introduction by including more background on TURFs, reserves, and TURF reserves.
-We have re-written some paragraphs to highlight the fact that TURF-reserve effectiveness has been proposed from theoretical and modelling work, but that no empirical evidence exists for reserves*
+*The points raised above were amongst the most useful comments, and the fact that these were highlighted by both reviewers suggest that there was room for improvement on our side.
+We have significantly expanded the introduction by including more background on and distinctions between TURFs, reserves, and TURF reserves.
+We have re-written parts of the introduction to highlight the fact that TURF-reserve effectiveness has been proposed from theoretical and modelling work, but that no empirical evidence exists for reserves*
 
@@
 
-*We have included more information on the wider context, touching on implementation processes and TURF history.*
+*We have included more information on the wider context, touching on implementation processes and TURF history.We have also expanded on the way in which the evaluated reserves are implemented, which may further help the reader understand TURF systems and TURF-reserves in Mexico.*
 
@@
 Further down in the discussion, we meant to say that no generalizations should be made to other Fish Refuges not implemented in TURF-managed fisheries.
-We believe that our current description and expanded introduction help make this more clear.*
+We believe that our current description and expanded introduction help make this more clear.
+In recognition of the caution needed when generalizing, we have also re-structured the title to make reference to Mexican small-scale fisheries.*
 
@@
 *We have included more information on data collection as supplementary materials, where we show the number of transects for each site (reserve - control) and type of survey (invertebrate - fish).
-The process of site selection is presented in the methods, where we expanded the definition of our control sites (L142)*
+The process of site selection is presented in the methods, where we expanded on the definition of our control sites and their location (L135).*
 
@@
 
-*We have explicitly mentioned the objective of evaluating the SES in our methods section (L134), and further explained variable selection in L203.*
+*We have explicitly mentioned the objective of evaluating the SES in our methods section (L122), and further explained variable selection in L197.*
 
@@
 
-*We have expanded on the methods section and further explained the Difference-in-Differences method (L178).
-We have also included a reference to the main paper (Card and Krueger, 1994) where the DiD methodology was first introduced.*
+*We have expanded on the methods section and further explained the Difference-in-Differences approach (L170).
+We have also included a reference to the main paper (Card and Krueger, 1994) where the DiD methodology was first introduced in econometric literature.*
 
@@
 
-*Our goal is to report the effect of the reserve on the indicator, and using effect sizes allows us to reduce the ambiguity that temporal and spatial trends may induce.
+*Our goal is to report the effect of the reserves on each indicator, and using effect sizes allows us to reduce the ambiguity that temporal and spatial trends may induce.
 However, we recognize the importance of observing "raw" data, and have included time series of all indicators in the supplementary materials.*
@@
 *The interpretation is correct, we intend to provide enough information to show that the characteristics of the system are conductive of effective resource management.
-The reason why similar broad descriptions are presented across sites is because these don't vary across them.
+The reason why similar broad descriptions are presented across sites is because there is little variation between them.
 For example, all sites must align with the same official fisheries regulation (Lobster Fishery Official Mexican Norm; NOM-006-pesc-1993).
@@
 The use of the SES framework simply allows us to be more objective by being transparent about the variables we report across communities and using a common language.
-We believe this is now made clear in sentence starting in L134.*
+We believe this is now made clear in sentence starting in L129.*
 
@@
 
-*We believe that part of the disconnect had to do with the way in which the previous text was presented, where we failed to distinguish between the evaluated TURF-reserves, other non-evaluated TURF-reserves and other reserves not implemented within TURFs.
-We have also made it clear that some of the claims (which include citations) come from conclusions of other works performed on these same sites.
+*We believe that part of the disconnect had to do with the way in which the previous text was presented, where we failed to distinguish between the evaluated TURF-reserves and other hypothetical TURF-reserves.
+We have now made it clear that some of the claims come from conclusions of other works performed on these same sites; we have included relevant references.
 For example, the buffering to environmental variation comes from Micheli et al., 2012, who evaluated the TURF-reserves in Isla Natividad.*
+
+*The point raised about the possible use of resources for other interventions is interesting.
+This issue has been partially addressed by other literature, such as Smith et al (2010) on what economists call "Opportunity costs".
+We have included a mention of this in the discussion (L327).*
 
@@
 
-*We have included a paragraph on some of these limitations in the discussion (L271)*
+*We have included a paragraph on some of these limitations in the discussion (L264)*
 
@@
 
-*Figures and table have been modified or added to the manuscript.*
+*The requested figures and tables have been added to the Supplementary Materials.*
 
@@
 
-*We have expanded on the methods and expect these to be now better documented.*
+*We have expanded on the methods.*
 
@@
 
-*Our data and code are deposited in a public GitHub Repository. We have included a sentence in the methods sections mentioning this and a link to said repository (L201).*
+*Our data and code are deposited in a public GitHub Repository.
+We have included a sentence in the methods sections mentioning this and a link to said repository (L195).
+The same information is available within the supplementary materials and in the "data and code archiving" section of the submission.*
 
@@
 While we believe it is equally important, we assess reserve effectiveness based on the stated objectives, which were to increase lobster densities and landings.
-We incorporate the SES approach to provide more background about the system under which the reserves are placed, and to discard the usually-mentioned causes of lack of effectiveness which "blame" fishers (namely poaching and lack of enforcement).
-Instead, our approach shows that the most likely cause of lack of effectiveness is poor design, that ignore basic design principles such as size of reserves.*
+We incorporate the SES approach to provide more background and context about the system under which the reserves are placed, and to systematically address the usually-mentioned causes of lack of effectiveness which "blame" fishers (namely poaching and lack of enforcement).
+Instead, our approach shows that the most likely cause of lack of effectiveness is poor design (i.e.
+reserves are too small).*
+
+*We agree that an operationalization of the SES that allows us to match scores to effectiveness would be an interesting approach.
+Initially, we had "operationalized" the SES framework and included a scoring system.
+However, since all sites operate under very similar settings, scores were practically the same and obtained little contrast between them.
+We believe that this happens because the three communities represent well-managed fisheries, and all of them achieve high scores.
+Furthermore, they all seem to show a common cause for lack of effectivness.
+The approach suggested would be much more interesting if it could be applied to a series of fisheries along a wider gradient of cooperation and co-management.
+We have included this in our discussion (L330).*
 
@@
 
-*We believe that the re-structuring of the introduction and case study descriptions have now addressed this.*
+*We believe that the re-structuring of the introduction and case study descriptions have provided additional context needed (Paragraphs in L39 and further in L77 and L85)..*
 
@@
 
-*No fishing activities occur within the Marine Reserves, and we have now made it clear that these are no-take marine reserves.*
+*No fishing activities occur within the Marine Reserves, and we have now made it clear that these are no-take marine reserves (L34 and L135).*
 
@@
 
-*Yes, these fisheries have been historically managed with TURFs. We have included this in the case study description (sentence starting in L120).*
+*Yes, these fisheries have been historically managed with TURFs. We have included this in the case study description (sentence starting in L115).*
 
@@
 *We were using the term "triple bottom line evaluation" in the same way as Halpern 2013 did.
-We realize that this is not clear, and have now explicitly stated that we "provide a holistic evaluation of the effectiveness of community-based in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop" (L65).*
+We realize that this was not a clear term.
+Instead, we have now explicitly stated that we "provide a holistic evaluation of the effectiveness of community-based in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop" (L60).*
 
@@
 
-*The sizes were included in Table 2 (RS3 - Size of resource system: reserve area), but are now also mentioned in the results (L248).
+*The sizes were included in Table 2 (RS3 - Size of resource system: reserve area), but are now also mentioned in the results (L243).
 We do not directly assess the effect of reserve size because all reserves are smaller than the suggested design principles (Green et al., 2017).
 If we had data from more sites and across a wider range of reserve sizes (i.e.
-smaller and greater than minimum requirements) it would be interesting to show how time to recovery or effect sizes relate to reserve size.
-One would expect to find a positive correlation.*
+smaller and greater than minimum requirements) it would be interesting to show how time to recovery and effect sizes relate to reserve size.
+One would expect to find a positive correlation between these, but that plateaus as one reaches optimum size.*
 
@@
 
-*Yes, all control sites are placed within the TURFs, as it is now explicitly mentioned in the case study section (list starting in L141).*
+*Yes, all control sites are placed within the TURFs, as it is now explicitly mentioned in the case study section (list starting in L135).*
 
@@
 
-*Mean ex-vessel prices for lobster are now presented in Table S3*
+*These are now presented in Table S3*
 
@@
 
-*Yes, our Difference-in-Differences analyses controls for community-specific changes.*
+*Yes, our Difference-in-Differences analyses controls for these changes.*
 
@@
 
-*Differences in revenue may be explained by both, a simple increase in prices due to market interactions or an increase in landings (plausibly because of the reserve).
-However, we do not observe an increase in revenues nor in landings, and ex-vessel prices remain relatively constant.*
+*Differences in revenue may be explained by both, an increase in prices due to market interactions or an increase in landings (plausibly because of the reserve or increased effort).
+However, we do not observe an increase in revenues or in landings, and ex-vessel prices remain relatively constant.*
 
@@
 Since we fit one model for each combination of indicator-community, we cannot make inference about the community- or species-level effects.
-This type of analyses would be more suited for a community ecology approach, or when seeking differential response among species (e.g.
+This type of analysis would be more suited for a community ecology approach, or when seeking differential response among species (e.g.
 Babcock et al, 2010).
-Furthermore, we believe that this approach should be used in a post hoc fashion for reserves that have actually had an effect on the indicators, seeking to understand which species and communities may be driving this.*
+Furthermore, we believe that this approach should be used in a post hoc fashion for reserves that have actually had an effect on the indicators, seeking to understand which species and communities may be driving this.
+We have included part of this explanation on model fitting when the analyses section was re-structured.*
 
@@
 *This assumption is correct, we only list the number of fishing co-op members.
-However, all fishers that fish within the TURF belong to the cooperative, and there are no "free fishers".
-This point is now mentioned in the case study section (L90).*
+However, all fishers that fish within the TURF belong to the cooperative, and there are no unaffiliated fishers.
+This point is now mentioned in the case study section (L85).*
 
@@
 
-*We have included the linear distance to the closes urban centers in Table 2.
-Isla Natividad lies 545 Km south from Tijuana, while Maria Elena and Punta Herrero are more than 200 km south from Cancun.*
+*We have included the linear distance to the closest urban centers in Table 2.
+Isla Natividad lies 545 Km south from Tijuana, while Maria Elena and Punta Herrero are more than 200 km south from Cancun.
+Both cities have more than 1 Million people, international airports, and receive large amounts of tourism.*
 
@@
 
-*Yes, each community engages in biological monitoring (the data for this study comes from this) and oceanographic monitoring.*
+*Yes, each community engages in biological monitoring  and oceanographic monitoring programmes. In fact, the data for this study comes from such monitoring efforts.*
 
@@
 
-*We believe that splitting the narrative column into three would be beneficial if each community had very different values / narratives.
+*We agree that splitting the narrative column into three would be beneficial if each community had very different values / narratives.
 However, these three communities have strikingly similar values.
@@
 *This is a fair point, which was raised previously.
-This would be an interesting approach.
-Initially, we had "operationalized" the SES framework and included a scoring system.
+This would be an interesting approach if we were opreationalizing the framework.
+As before, we state: Initially, we had "operationalized" the SES framework and included a scoring system.
 However, since all sites operate under very similar settings scores were practically the same and obtained little contrast between them.
@@
 
-*This big fluctuations in biomass are explained by an opportunistic increase in sizes.
+*This big fluctuations in biomass are explained by an increase in sizes.
 In this particular case, a few larger-than-usual organisms were observed, but the total number of fish remained similar to other years.
 As a result, the few larger fish observed these years yield a higher biomass.
-However these differences are not significantly different from the pre-observed trend.*
+However these differences are not significantly different from the pre-observed trend or from other years.*
 
@@
 This would be explained by the relative reduction of total fishable area.
-The "dip" described in Ovando et al., (2016) would then recover as densities inside the reserve recover and spillover occurs.
+The "dip" described in Ovando et al., (2016) would then recover as densities inside the reserve increase and spillover occurs.
 The fact that we don't see a reduction in landings over the first few years suggests that reserves are not big enough to reduce landings over the first few years.*
@@
 
-*We have removed this lose statement and included an explicit conclusions paragraph (L335) where we state that "We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside".*
+*We have removed this lose statement and included an explicit conclusions paragraph (L322) where we state that "We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside".*
 
@@
 
-*The previous wording of this statement was confusing, and seemed like we were referencing the reserves evaluated in this work.
+*Our previous wording of this statement made it confusing, and seemed like we were referencing the reserves evaluated in this work.
 Clearly this is not true and our data would not support such claim.
@@
 *The figure caption has been corrected and now says "circles"*
-
-place-holder
-
-*This would be an interesting approach.
-Initially, we had "operationalized" the SES framework and included a scoring system.
-However, since all sites operate under very similar settings scores were practically the same and obtained little contrast between them.
-We believe that this happens because the three communities represent well-managed fisheries, and all of them achieve high scores.
-The approach suggested would be much more interesting if it could be applied to a series of fisheries along a wider gradient of cooperation and co-management.
-We have included a discussion paragraph recognizing this (L330).*

### 2019-01-27 docs/SupplementaryMaterial.Rmd
Uncorporated changes from Stuart and Rev2 and re-submitted

@@
                col.names = c("Zone", "Group", "TURF", "Before", "After"),
-               caption = "Mean ex-vessel prices for lobster (Panulirus spp.) on each TURF (community) by zone (MC = Caribbean; PN = Pacific) and group (Control = only TURF, Treated = TURF and reserve).",
+               caption = "Mean ex-vessel prices (MXP / Kg) for lobster (Panulirus spp.) on each TURF (community) by zone (MC = Caribbean; PN = Pacific) and group (Control = only TURF, Treated = TURF and reserve).",
                booktabs = T)

### 2019-01-27 docs/manuscript.Rmd
Uncorporated changes from Stuart and Rev2 and re-submitted

@@
 In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
-One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves within them, thus creating TURF-reserve systems \citep{afflerbach_2014,gelcich_2015,lester_2017}.
+One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves, thus creating TURF-reserve systems \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
-TURFs are a fisheries management tool in which a well defined group of fishers (*e.g.* fishing cooperatives) have exclusive access to an explicitly delimited portion of the ocean.
+TURFs are a fisheries management tool in which a well-defined group of fishers (*e.g.* fishing cooperatives) have exclusive access to an explicitly delimited portion of the ocean.
 They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,costello_2010,mccay_2014}.
@@
 
-Conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+Conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits such as climate change mitigation, protection from environmental variability, and fisheries benefits \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
 Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \citep{gelcich_2008,gelcich_2012}.
@@
 Moreover, TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{ostrom_2009,gelcich_2015}.
-It is therefore important to consider not only the indicators of interest, bu also the governance settings under which the reserves occur.
+It is therefore important to consider not only the indicators of interest, but also the governance settings under which the reserves operate.
 
@@
 Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness.
-We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild their coastal fisheries.
+We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.
 
@@
 
-Before discussing our data collection methods and describing our analyses, our case studies warrant some more background.
-Community-based marine reserves that are implemented within TURFs are a form of TURF-reserves, voluntarily established and enforced by local communities.
+Before discussing our data collection methods and describing our analyses, our case studies warrant some background.
+Community-based marine reserves that are implemented within TURFs are a form of TURF-reserve, voluntarily established and enforced by local communities.
 This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
-Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}.
+Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{johannes_2002,bohnsack_2004}.
 However, community-based reserves can be hard to enforce if they are not legally recognized.
@@
 
-In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue: Fish Refuges \citep{nom}.
-Fish Refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries.
-One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporal marine reserves within their TURFs with a defined expiration date (often 5 years).
+In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and helped create a legal framework that solves this governance issue: Fish Refuges \citep{nom}.
+Fish Refuges can be implemented as permanent, temporary or partial reserves, which can protect one, some, or all resources within their boundaries.
+One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporary marine reserves within their TURFs with a defined expiration date (often five years).
 When the expiration date is reached, fishers can chose to open the reserves to fishing or re-establish them.
-Our work focuses on Fish Refuges implemented as community-based TURF-reserves that occur in small-scale fisheries.
+Our work focuses on Fish Refuges implemented as community-based TURF-reserves in small-scale fisheries.
 
-The common setup of community-based TURF-reserves in Mexico is the following.
+The most common setup of community-based TURF-reserves in Mexico is the following.
 Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (*i.e.* TURFs shown as blue polygons in Fig \ref{fig:map}A).
@@
 
-We evaluate three community-based no-take TURF-reserves implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig \ref{fig:map}A).
+We evaluate three community-based no-take TURF-reserve systems implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig \ref{fig:map}A).
 The first one was created by the *Buzos y Pescadores de la Baja California* fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig \ref{fig:map}B).
-The main fishery in the island is the spiny lobster (*Panulirus interruptus*), but other resources like finfish, sea cucumber, read sea urchin, snail, and abalone are also an important source of income.
+The main fishery in the island is the spiny lobster (*Panulirus interruptus*), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income.
 In 2006, the community decided to implement two marine reserves within their fishing grounds.
 The objective of these reserves was "to protect commercially important invertebrate species"; mainly lobster and abalone.
-These reserves obtained legal recognition in 2018 \citep{dof_website_2018}.
+The reserves obtained legal recognition in 2018 \citep{dof_website_2018}.
 
-The other two TURF-reserves are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C).
-In contrast with Isla Nativdad, which hosts a well established fishing community, Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the *Cozumel* fishing cooperative; Punta Herrero is home to the *José María Azcorra* fishing cooperative, and similar to Isla Natividad hosts a local community.
+The other two TURF-reserve systems are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C).
+In contrast with Isla Nativdad, which hosts a well-established fishing community, Maria Elena is a fishing camp visited intermittently during the fishing season that belongs to the *Cozumel* fishing cooperative.
+Punta Herrero is home to the *José María Azcorra* fishing cooperative, and similar to Isla Natividad hosts a small community.
 Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season.
@@
 
-We use three main sources of information to evaluate these reserves across the ecological, socioeconomic, and governance dimensions.
+We use three main sources of information to evaluate these reserves across ecological, socioeconomic, and governance dimensions.
 Ecological data come from the annual ecological monitoring of reserve and control sites.
 Reserve sites are areas where no fishing occurs.
-Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) Are not directly adjacent to the reserves.
+Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) are not directly adjacent to the reserves.
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
@@
 We collected information on the Resource Systems, Resource Units, Actors, and Governance System (Table \ref{table:ses}).
-The next section further describe the variables and indicators chosen for each of these.
 
@@
 
-The following sections present the effect that marine reserves had on each of the biological and socioeconomic indicators for each coastal community.
-Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* the difference-in-differences estimate or effect size $\lambda_t$ from Eqs.
+The following sections present the effect that marine reserves had on the biological and socioeconomic indicators for each coastal community.
+Results are presented in terms of difference through time and across sites, relative to the control site on the year of implementation (*i.e.* the difference-in-differences estimate or effect size $\lambda_t$ from Eqs.
 \ref{eqn:reg_bio} and \ref{eqn:soc_reg}).
@@
 However, effect sizes do not change after the implementation of the reserve.
-Interestingly, the negative effect observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events.
+Interestingly, the negative effect observed for Isla Natividad on year 5 corresponds to the 2011 hypoxia events \citep{micheli_2012-EU}.
 The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$).
@@
 Reserves are also relatively young (RS5).
-Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely that marine reserves will result in increased catches.
+Additionally, fishers harvest healthy stocks (RS4.1), and it is unlikely that marine reserves will result in increased catches.
 
@@
 A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve.
-For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and thus higher profitability.
+For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and thus higher profitability, provided that cost per unit effort does not increase.
 
@@
 Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (*sensu* \citet{rees_2018}).
-Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}.
-Abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary invertebrates, but not lobsters.
+Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}, however, abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary invertebrates, but not lobsters.
 Design principles developed by \citet{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
@@
 
-Finally, extreme conditions, including prolonged hypoxia, heat waves, and storms have affected both the Pacific and Caribbean regions, with large negative impacts of coastal marine species and ecosystems \citep{cavole_2016,hughes_2018,breitburg_2018}.
+Finally, extreme conditions, including prolonged hypoxia, heat waves, and storms have affected both the Pacific and Caribbean regions, with large negative impacts on coastal marine species and ecosystems \citep{cavole_2016,hughes_2018,breitburg_2018}.
 The coastal ecosystems where these reserves are located have been profoundly affected by these events \citep{micheli_2012-EU,woodson_2018}.
@@
 
-While the evaluated reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits.
+While the evaluated reserves have failed to provide fishery benefits to date, there are a number of additional ecological, fisheries, and social benefits.
 Marine reserves provide protection to a wider range of species and vulnerable habitat.
@@
 Self-regulation of fishing effort can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
-Furthermore, embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}.
+Furthermore, embarking on a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}.
 Showing commitment to marine conservation and sustainable fishing practices has allowed fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
@@
 Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.
-Furthermore, consvation and advocacy groups should consider the opportunity costs of such interventions (*sensu* \citet{smith_2010}) and evaluate the potential of other approaches that may yield similar benefits.
+Furthermore, conservation and advocacy groups should consider the opportunity costs of such interventions (*sensu* \citet{smith_2010}) and evaluate the potential of other approaches that may yield similar benefits.
 
-In terms of the fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs.
+In terms of fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs.
 Future research should aim at evaluating other Fish Refuges established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through top-down processes (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.
@@
 This study would have not been possible without the effort by members of the fishing communities here mentioned, who participated in the data-collection process.
-The authors wish to acknowledge comments by the reviewrs and editor, which significantly improved the quality of this work.
+The authors wish to acknowledge comments by the reviewers and editor, which significantly improved the quality of this work.
 

### 2019-01-27 docs/manuscript.tex
Uncorporated changes from Stuart and Rev2 and re-submitted

@@
 local resources is through the implementation of Territorial Use Rights
-for Fisheries (TURFs) that contain no-take marine reserves within them,
-thus creating TURF-reserve systems
+for Fisheries (TURFs) that contain no-take marine reserves, thus
+creating TURF-reserve systems
 \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
-TURFs are a fisheries management tool in which a well defined group of
+TURFs are a fisheries management tool in which a well-defined group of
 fishers (\emph{e.g.} fishing cooperatives) have exclusive access to an
@@
 \citep{lester_2009}, and that these may have a series of additional
-benefits like climate change mitigation, protection from environmental
-variability, and fisheries benefits
+benefits such as climate change mitigation, protection from
+environmental variability, and fisheries benefits
 \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
@@
 \citep{ostrom_2009,gelcich_2015}. It is therefore important to consider
-not only the indicators of interest, bu also the governance settings
-under which the reserves occur.
+not only the indicators of interest, but also the governance settings
+under which the reserves operate.
 
@@
 effectiveness of community-based TURF-reserves in small-scale fisheries
-where this tool is used to manage and rebuild their coastal fisheries.
+where this tool is used to manage and rebuild coastal fisheries.
 
@@
 Before discussing our data collection methods and describing our
-analyses, our case studies warrant some more background. Community-based
+analyses, our case studies warrant some background. Community-based
 marine reserves that are implemented within TURFs are a form of
-TURF-reserves, voluntarily established and enforced by local
-communities. This bottom-up approach increases compliance and
-self-enforcement, and reserves can yield benefits similar to
-systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
-Community-based spatial closures occur in different contexts, like the
-\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
-\citep{bohnsack_2004,johannes_2002}. However, community-based reserves
-can be hard to enforce if they are not legally recognized. In such
-conditions, TURF fishers must rely on the exclusive access of the TURF
-to maintain high levels of compliance.
+TURF-reserve, voluntarily established and enforced by local communities.
+This bottom-up approach increases compliance and self-enforcement, and
+reserves can yield benefits similar to systematically-designed reserves
+\citep{beger_2004,smallhornwest_2018}. Community-based spatial closures
+occur in different contexts, like the \emph{kapu} or \emph{ra'ui} areas
+in the Pacific Islands \citep{johannes_2002,bohnsack_2004}. However,
+community-based reserves can be hard to enforce if they are not legally
+recognized. In such conditions, TURF fishers must rely on the exclusive
+access of the TURF to maintain high levels of compliance.
 
@@
 Organizations (CSOs) served as a link between fishers and government,
-and created a legal framework that solves this governance issue: Fish
-Refuges \citep{nom}. Fish Refuges can be implemented as temporal or
-partial reserves, which can protect one, some, or all resources within
-their boundaries. One of the ways in which fishing communities have
-taken advantage of this new tool is by implementing temporal marine
-reserves within their TURFs with a defined expiration date (often 5
-years). When the expiration date is reached, fishers can chose to open
-the reserves to fishing or re-establish them. Our work focuses on Fish
-Refuges implemented as community-based TURF-reserves that occur in
+and helped create a legal framework that solves this governance issue:
+Fish Refuges \citep{nom}. Fish Refuges can be implemented as permanent,
+temporary or partial reserves, which can protect one, some, or all
+resources within their boundaries. One of the ways in which fishing
+communities have taken advantage of this new tool is by implementing
+temporary marine reserves within their TURFs with a defined expiration
+date (often five years). When the expiration date is reached, fishers
+can chose to open the reserves to fishing or re-establish them. Our work
+focuses on Fish Refuges implemented as community-based TURF-reserves in
 small-scale fisheries.
 
-The common setup of community-based TURF-reserves in Mexico is the
+The most common setup of community-based TURF-reserves in Mexico is the
 following. Fishers from a given community are assembled in fishing
@@
 
-We evaluate three community-based no-take TURF-reserves implemented in
-Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig
-\ref{fig:map}A). The first one was created by the \emph{Buzos y
-Pescadores de la Baja California} fishing cooperative, located in Isla
-Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The
-main fishery in the island is the spiny lobster (\emph{Panulirus
-interruptus}), but other resources like finfish, sea cucumber, read sea
-urchin, snail, and abalone are also an important source of income. In
-2006, the community decided to implement two marine reserves within
-their fishing grounds. The objective of these reserves was ``to protect
-commercially important invertebrate species''; mainly lobster and
-abalone. These reserves obtained legal recognition in 2018
+We evaluate three community-based no-take TURF-reserve systems
+implemented in Mexican TURF-managed fisheries, therefore making them
+TURF-reserves (Fig \ref{fig:map}A). The first one was created by the
+\emph{Buzos y Pescadores de la Baja California} fishing cooperative,
+located in Isla Natividad in the Baja California Peninsula (Fig
+\ref{fig:map}B). The main fishery in the island is the spiny lobster
+(\emph{Panulirus interruptus}), but other resources like finfish, sea
+cucumber, sea urchin, snail, and abalone are also an important source of
+income. In 2006, the community decided to implement two marine reserves
+within their fishing grounds. The objective of these reserves was ``to
+protect commercially important invertebrate species''; mainly lobster
+and abalone. The reserves obtained legal recognition in 2018
 \citep{dof_website_2018}.
 
-The other two TURF-reserves are located in Maria Elena and Punta
+The other two TURF-reserve systems are located in Maria Elena and Punta
 Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
-Isla Nativdad, which hosts a well established fishing community, Maria
-Elena is a fishing camp --visited intermittently during the fishing
-season-- belonging to the \emph{Cozumel} fishing cooperative; Punta
-Herrero is home to the \emph{José María Azcorra} fishing cooperative,
-and similar to Isla Natividad hosts a local community. Their main
-fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but
-they also target finfish in the off-season. Maria Elena and Punta
-Herrero established eight and four marine reserves in 2012 and 2013,
-respectively. These reserves have been legally recognized as Fishing
-Refuges since their original implementation
-\citep{dof_website_2012,dof_website_2013} and subsequent
-re-establishments \citep{dof_website_2017b}.
+Isla Nativdad, which hosts a well-established fishing community, Maria
+Elena is a fishing camp visited intermittently during the fishing season
+that belongs to the \emph{Cozumel} fishing cooperative. Punta Herrero is
+home to the \emph{José María Azcorra} fishing cooperative, and similar
+to Isla Natividad hosts a small community. Their main fishery is the
+Caribbean spiny lobster (\emph{Panulirus argus}), but they also target
+finfish in the off-season. Maria Elena and Punta Herrero established
+eight and four marine reserves in 2012 and 2013, respectively. These
+reserves have been legally recognized as Fishing Refuges since their
+original implementation \citep{dof_website_2012,dof_website_2013} and
+subsequent re-establishments \citep{dof_website_2017b}.
 
@@
 We use three main sources of information to evaluate these reserves
-across the ecological, socioeconomic, and governance dimensions.
-Ecological data come from the annual ecological monitoring of reserve
-and control sites. Reserve sites are areas where no fishing occurs.
-Control sites are areas that meet the following criteria: i) habitat
-characteristics are similar to the corresponding reserves, ii)
-presumably had a similar probability of being selected as reserves
-during the design phase, iii) are located within the TURF, where fishing
-occurs, and iv) Are not directly adjacent to the reserves. We focus our
-evaluation on sites where data are available for reserve and control
-sites, before and after the implementation of the reserve. This provides
-us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design
-that allows us to capture and control for temporal and spatial dynamics
+across ecological, socioeconomic, and governance dimensions. Ecological
+data come from the annual ecological monitoring of reserve and control
+sites. Reserve sites are areas where no fishing occurs. Control sites
+are areas that meet the following criteria: i) habitat characteristics
+are similar to the corresponding reserves, ii) presumably had a similar
+probability of being selected as reserves during the design phase, iii)
+are located within the TURF, where fishing occurs, and iv) are not
+directly adjacent to the reserves. We focus our evaluation on sites
+where data are available for reserve and control sites, before and after
+the implementation of the reserve. This provides us with a
+Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
+allows us to capture and control for temporal and spatial dynamics
 \citep{stewartoaten_1986,depalma_2018} and causally attribute the
@@
 Resource Units, Actors, and Governance System (Table \ref{table:ses}).
-The next section further describe the variables and indicators chosen
-for each of these.
 
@@
 heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
-analyses were performed in R version 3.5.1 and R Studio version 1.1.456
+analyses were performed in R version 3.5.2 and R Studio version 1.1.456
 \citep{R_2018}. All data and code needed to reproduce our analyses are
@@
 The following sections present the effect that marine reserves had on
-each of the biological and socioeconomic indicators for each coastal
-community. Results are presented in terms of the difference through time
-and across sites, relative to the control site on the year of
-implementation (\emph{i.e.} the difference-in-differences estimate or
-effect size \(\lambda_t\) from Eqs. \ref{eqn:reg_bio} and
-\ref{eqn:soc_reg}). We also provide an overview of the governance
-settings of each community, and discuss how these might be related to
-the effectiveness and performance of the reserves.
+the biological and socioeconomic indicators for each coastal community.
+Results are presented in terms of difference through time and across
+sites, relative to the control site on the year of implementation
+(\emph{i.e.} the difference-in-differences estimate or effect size
+\(\lambda_t\) from Eqs. \ref{eqn:reg_bio} and \ref{eqn:soc_reg}). We
+also provide an overview of the governance settings of each community,
+and discuss how these might be related to the effectiveness and
+performance of the reserves.
 
@@
 not change after the implementation of the reserve. Interestingly, the
-negative effect observed for Isla Natividad on year 5 correspond to the
-2011 hypoxia events. The only positive change observed in lobster
-landings is for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
-post-implementation data for Maria Elena do not show a significant
-effect of the reserve. Isla Natividad shows higher revenues after the
-implementation of the reserve, as compared to the control communities.
-However, these changes are only significant for the third year
-(\(p < 0.05\)). Full tables with model coefficients are presented in the
-supplementary materials (Tables S4-S5).
+negative effect observed for Isla Natividad on year 5 corresponds to the
+2011 hypoxia events \citep{micheli_2012-EU}. The only positive change
+observed in lobster landings is for Isla Natividad in 2014
+(\(p < 0.1\)). The three years of post-implementation data for Maria
+Elena do not show a significant effect of the reserve. Isla Natividad
+shows higher revenues after the implementation of the reserve, as
+compared to the control communities. However, these changes are only
+significant for the third year (\(p < 0.05\)). Full tables with model
+coefficients are presented in the supplementary materials (Tables
+S4-S5).
 
@@
 \textsuperscript{2}. Reserves are also relatively young (RS5).
-Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely
+Additionally, fishers harvest healthy stocks (RS4.1), and it is unlikely
 that marine reserves will result in increased catches.
@@
 decreased, that would imply a larger catch per unit effort and thus
-higher profitability.
+higher profitability, provided that cost per unit effort does not
+increase.
 
@@
 Natividad yield fishery benefits for the abalone fishery
-\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
-perhaps the reserves provide enough protection to these sedentary
-invertebrates, but not lobsters. Design principles developed by
-\citet{green_2017} for marine reserves in the Caribbean state that
+\citep{rossetto_2015-V0}, however, abalone are less mobile than
+lobsters, and perhaps the reserves provide enough protection to these
+sedentary invertebrates, but not lobsters. Design principles developed
+by \citet{green_2017} for marine reserves in the Caribbean state that
 reserves ``should be more than twice the size of the home range of
@@
 and storms have affected both the Pacific and Caribbean regions, with
-large negative impacts of coastal marine species and ecosystems
+large negative impacts on coastal marine species and ecosystems
 \citep{cavole_2016,hughes_2018,breitburg_2018}. The coastal ecosystems
@@
 
-While the evaluated reserves have failed to provide fishery benefits up
-to now, there are a number of additional ecological, fisheries, and
-social benefits. Marine reserves provide protection to a wider range of
-species and vulnerable habitat. Previous research focusing on these
-specific sites has shown that they serve as an insurance mechanism
-against uncertainty and errors in fisheries management, as well as mild
+While the evaluated reserves have failed to provide fishery benefits to
+date, there are a number of additional ecological, fisheries, and social
+benefits. Marine reserves provide protection to a wider range of species
+and vulnerable habitat. Previous research focusing on these specific
+sites has shown that they serve as an insurance mechanism against
+uncertainty and errors in fisheries management, as well as mild
 environmental shocks
@@
 future declines associated to environmental variation
-\citep{finkbeiner_2018}. Furthermore, embarking in a marine conservation
+\citep{finkbeiner_2018}. Furthermore, embarking on a marine conservation
 project can bring the community together, which promotes social cohesion
@@
 TURF-reserves meet essential design principles such as size and
-placement so as to maximize their effectiveness. Furthermore, consvation
-and advocacy groups should consider the opportunity costs of such
-interventions (\emph{sensu} \citet{smith_2010}) and evaluate the
+placement so as to maximize their effectiveness. Furthermore,
+conservation and advocacy groups should consider the opportunity costs
+of such interventions (\emph{sensu} \citet{smith_2010}) and evaluate the
 potential of other approaches that may yield similar benefits.
 
-In terms of the fisheries regulation in Mexico, our work only evaluates
-Fish Refuges established within TURFs. Future research should aim at
+In terms of fisheries regulation in Mexico, our work only evaluates Fish
+Refuges established within TURFs. Future research should aim at
 evaluating other Fish Refuges established as bottom-up processes but
@@
 data-collection process. The authors wish to acknowledge comments by the
-reviewrs and editor, which significantly improved the quality of this
+reviewers and editor, which significantly improved the quality of this
 work.

### 2019-02-11 docs/SupplementaryMaterial.Rmd
Added changes for Rev1

--- 2018-12-31 docs/SupplementaryMaterial.Rmd
@@
 conapesca %>% 
+  filter(!unidad_economica == "jose maria azcorra") %>% 
   mutate(price = valor / peso_vivo) %>% 
@@
   spread(post, price) %>% 
+  mutate(zona = ifelse(zona == "MC", "Yucatan Peninsula", "Baja California"),
+         unidad_economica = taxize::taxize_capwords(unidad_economica)) %>% 
   knitr::kable(digits = 2,
-               col.names = c("Zone", "Group", "TURF", "Before", "After"),
-               caption = "Mean ex-vessel prices for lobster (Panulirus spp.) on each TURF (community) by zone (MC = Caribbean; PN = Pacific) and group (Control = only TURF, Treated = TURF and reserve).",
+               col.names = c("Community", "Group", "TURF", "Before", "After"),
+               caption = "Mean ex-vessel prices (MXP / Kg) for lobster (Panulirus spp.) on each TURF (community) and group (Control = only TURF, Treated = TURF and reserve).",
                booktabs = T)
@@
 valor <- conapesca %>% 
+  filter(!unidad_economica == "jose maria azcorra") %>% 
+  mutate(zona = ifelse(zona == "MC", "Yucatan Peninsula", "Baja California")) %>% 
   ggplot(aes(x = years_centered, y = valor / 1000, color = grupo)) +
@@
 peso <- conapesca %>% 
+  filter(!unidad_economica == "jose maria azcorra") %>% 
   ggplot(aes(x = years_centered, y = peso_vivo / 1000, color = grupo)) +
@@
 
-cowplot::plot_grid(valor, peso, ncol = 1)
+cowplot::plot_grid(peso, valor, ncol = 1)
 

### 2019-02-11 docs/manuscript.Rmd
Added changes for Rev1

@@
 Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
-In the last 12 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2014; their effectiveness has not been formally evaluated.
+In the last 13 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2012; their effectiveness has not been formally evaluated.
 We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
-We find that while reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates, they continue to receive support from the fishing communities.
+We find that reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates.
 A lack of clear ecological and socioeconomic effects likely results from a combination of factors.
@@
 
-Conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits such as climate change mitigation, protection from environmental variability, and fisheries benefits \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+Conservation science has shown how marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
 Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \citep{gelcich_2008,gelcich_2012}.
 The benefits resulting from reserves established within TURFs (*i.e.* TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \citep{gelcich_2015}.
-Although in theory these systems are successful \citep{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success.
+Although in theory these systems are expected to be successful \citep{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success.
 Moreover, TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{ostrom_2009,gelcich_2015}.
@@
 
-Recent norms in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries.
-In Mexico, a norm created in 2014 allows fishers to request legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}).
-Since 2012, old and new marine reserves have gained legal recognition as Fish Refuges.
+Recent changes in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries.
+In Mexico, a legal framework created in 2012 allows fishers to request legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*, described in more detail below; \citet{nom}).
+Since 2012, 45 old and new marine reserves have gained legal recognition as Fish Refuges.
 Of these, 18 were originally implemented within TURFs.
@@
 
-Before discussing our data collection methods and describing our analyses, our case studies warrant some background.
 Community-based marine reserves that are implemented within TURFs are a form of TURF-reserve, voluntarily established and enforced by local communities.
-This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
+This bottom-up approach can increase compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
 Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{johannes_2002,bohnsack_2004}.
@@
 Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (*i.e.* TURF-reserves).
-Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges following a series of studies outlined in the regulation \citep{nom}.
+Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges after conducting a series of studies outlined in the regulation, namely the creation of a baseline of ecological data and historical performance of the fishery \citep{nom}.
 
@@
 In 2006, the community decided to implement two marine reserves within their fishing grounds.
-The objective of these reserves was "to protect commercially important invertebrate species"; mainly lobster and abalone.
+The objective of these reserves was "to protect and recover stocks of commercially important invertebrate species"; mainly lobster and abalone.
 The reserves obtained legal recognition in 2018 \citep{dof_website_2018}.
@@
 
-The regulation governing the implementation of Fish Refuges states that these are fishery management tools intended to have biological or socioeconomic benefits \citep{nom}.
+The regulation governing the implementation of Fish Refuges states that these are fishery management tools intended to have conservation and fisheries benefits \citep{nom}.
 For this reason, the main portion of our analyses focuses on a series of biological and socioeconomic indicators that may respond to reserve implementation.
@@
 The biological data are collected by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)).
-Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \citep{suman_2010-ez,fulton_2018,fulton_2019}.
+Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites, where a minimum of 4 transects per site are performed \citep{suman_2010-ez,fulton_2018,fulton_2019}.
 Size structures are also collected during fish surveys.
@@
 In this case our treated unit are the cooperatives (*i.e.* communities) that have implemented a reserve within their TURF, and the controls are nearby communities that have a TURF but did not implement a reserve.
-Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017,ayer_2018}.
+These communities have TURFs adjacent to the TURFs of our focal communities.
+Cooperatives incorporated in this analysis have similar number of members, belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2014,mccay_2017,ayer_2018}.
 Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017}.
@@
 This approach is widely used in econometric literature to estimate the average treatment effect of an intervention, like the impact of minimum wage increases on employment rates \citep{card_1994}.
-In our case it allows us to estimate the effect that the reserve had on each biological and socioeconomic indicator (Table \ref{table:indicators}) by comparing trends across time and treatments \citep{moland_2013,Villasenor-Derbez_2018}.
+In our case it allows us to estimate the effect that the reserve had on each biological and socioeconomic indicator (Table \ref{table:indicators}) by comparing trends across time and treatments since reserve implementation \citep{moland_2013,Villasenor-Derbez_2018}.
 To perform difference-in-differences, we regress the indicator of interest on a dummy variable for treatment, a dummy variable for years, and the interaction term between these with a multiple linear regression of the form:
@@
 These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \citep{depalma_2018}.
-We fit one model per community and indicators (*e.g.* there are three models for lobster density, one for each community) for a total of 12 biological model fits and four socioeconomic model fits.
+For each indicator in each community, we fit one model (*e.g.* there are three models for lobster density, one for each community) for a total of 12 biological model fits and four socioeconomic model fits.
 Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard errors \citep{zeileis_2004-7n}.
@@
 The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators.
+Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as in \citet{leslie_2015-na}), we followed a similar approach to \citet{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems.
 We based our variable selection primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework, and identified the key variables relevant to fishing cooperatives in Mexico.
@@
 The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$).
-The three years of post-implementation data for Maria Elena do not show a significant effect of the reserve.
+The year of post-implementation data for Maria Elena does not show a significant effect of the reserve.
 Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities.
@@
 
-Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed communities share similarities known to foster sustainable resource management and increase reserve effectiveness.
+Our qualitative implementation of the social-ecological systems framework allowed us to systematically identify important differences between the case studies’ governance systems and incorporate other characteristics of these fisheries neglected during the process of data collection (Table \ref{table:ses}).
+We find that the analyzed communities share similarities known to foster sustainable resource management and increase reserve effectiveness.
 For example, fishers operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves.
@@
 
-Our analyses of socioeconomic indicators has two limitations.
+Our analyses of socioeconomic indicators has three limitations.
 First, we only look at landings and revenues by landings for communities with and without TURF-reserves.
@@
 For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and thus higher profitability, provided that cost per unit effort does not increase.
+Likewise, it is possible that fishing effort increased around reserves to maintain the historical levels of landings.
+A final limitation applies to Maria Elena, where we only observe landings and income for one year after reserve implementation.
+While one would not expect to observe increased landings or income in such a short period, a spatial closure might cause total catches to decline, especially if effort is held constant.
 
@@
 With these characteristics, one would expect the reserves to be effective.
-Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 6 years old) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \citep{dasilva_2015-zX}.
+Maria Elena and Punta Herrero are relatively young reserves (*i.e.* < 6 years old; RS5 in Table \ref{table:ses}) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \citep{dasilva_2015-zX}.
 
@@
 Design principles developed by \citet{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
+As shown through the SES analysis, the size of the marine reserves appears small compared to the movement capacity of the main targeted species (RU1, RS3; Table \ref{table:ses}).
 Furthermore, fishers may favor implementation of reserves that pose low fishing costs due to their small size or location.
@@
 Showing commitment to marine conservation and sustainable fishing practices has allowed fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
-These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-implementing the reserves \citep{ayer_2018}.
+These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-establishing the reserves \citep{ayer_2018}.
 
@@
   column_spec(column = 1, width = "6.5cm") %>% 
-  column_spec(column = 2, width = "12cm") %>% 
+  column_spec(column = 2, width = "12.2cm") %>% 
   group_rows(group_label = "Resource System (RS)", start_row = 1, end_row = 5) %>% 
-  group_rows(group_label = "Resource Unit (RU)", start_row = 6, end_row = 6) %>% 
-  group_rows(group_label = "Actors (A)", start_row = 7, end_row = 8) %>% 
-  group_rows(group_label = "Governance system (G)", start_row = 9, end_row = 12)
+  group_rows(group_label = "Resource Unit (RU)", start_row = 6, end_row = 7) %>% 
+  group_rows(group_label = "Actors (A)", start_row = 8, end_row = 9) %>% 
+  group_rows(group_label = "Governance system (G)", start_row = 10, end_row = 13)
 ```

### 2019-02-11 docs/manuscript.tex
Added changes for Rev1

@@
 \def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
-\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
-Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
-Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
-Micheli\(^{4}\)}
+\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\), Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza Micheli\(^{4}\)}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
 % If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
-of California, Santa Barbara, Santa Barbara, CA, USA \newline
-\(^{2}\)Nicholas School of the Environment, Duke University, Beaufort,
-NC, USA \newline \(^{3}\)Comunidad y Biodiversidad A.C., Guaymas,
-Sonora, Mexico \newline \(^{4}\)Hopkins Marine Station and Center for
-Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
+\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA \newline
+\(^{2}\)Nicholas School of the Environment, Duke University, Beaufort, NC, USA \newline
+\(^{3}\)Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico \newline
+\(^{4}\)Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
 % The Corresponding Author should be marked with an asterisk
 % Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
-Santa Barbara, Santa Barbara, CA, 93106}
+\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106}
 
@@
 
-\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in Mexican small-scale
-fisheries}
+\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries}
 
@@
 
-Coastal marine ecosystems provide livelihoods for small-scale fishers
-and coastal communities around the world. Small-scale fisheries face
-great challenges since they are difficult to monitor, enforce, and
-manage. Combining territorial use rights for fisheries (TURF) with
-no-take marine reserves to create TURF-reserves can improve the
-performance of small-scale fisheries by buffering fisheries from
-environmental variability and management errors, while ensuring that
-fishers reap the benefits of conservation investments. In the last 12
-years, 18 old and new community-based Mexican TURF-reserves gained legal
-recognition thanks to a regulation passed in 2014; their effectiveness
-has not been formally evaluated. We combine causal inference techniques
-and the Social-Ecological Systems framework to provide a holistic
-evaluation of community-based TURF-reserves in three coastal communities
-in Mexico. We find that while reserves have not yet achieved their
-stated goal of increasing the density of lobster and other benthic
-invertebrates, they continue to receive support from the fishing
-communities. A lack of clear ecological and socioeconomic effects likely
-results from a combination of factors. First, some of these reserves
-might be too young for the effects to show. Second, the reserves are not
-large enough to protect mobile species, like lobster. Third, variable
-and extreme oceanographic conditions have impacted harvested
-populations. Fourth, local fisheries are already well managed, and it is
-unlikely that reserves might have a detectable effect in catches.
-However, these reserves may provide a foundation for establishing
-additional, larger marine reserves needed to effectively conserve mobile
-species.
+Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
+Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage.
+Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
+In the last 13 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2012; their effectiveness has not been formally evaluated.
+We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
+We find that reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates.
+A lack of clear ecological and socioeconomic effects likely results from a combination of factors.
+First, some of these reserves might be too young for the effects to show.
+Second, the reserves are not large enough to protect mobile species, like lobster.
+Third, variable and extreme oceanographic conditions have impacted harvested populations.
+Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches.
+However, these reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species.
 
@@
 \tiny
- \keyFont{ \section{Keywords:} TURF-reserves, Causal Inference, Social-Ecological Systems, Marine
-Protected Areas, Marine Conservation, Small-Scale Fisheries}
+ \keyFont{ \section{Keywords:} TURF-reserves, Causal Inference, Social-Ecological Systems, Marine Protected Areas, Marine Conservation, Small-Scale Fisheries}
 
@@
 
-Marine ecosystems around the world sustain significant impacts due to
-overfishing and unsustainable fishing practices
-\citep{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular,
-small-scale fisheries face great challenges since they tend to be hard
-to monitor and enforce \citep{costello_2012}. One of the many approaches
-taken to improve the performance of coastal fisheries and health of the
-local resources is through the implementation of Territorial Use Rights
-for Fisheries (TURFs) that contain no-take marine reserves, thus
-creating TURF-reserve systems
-\citep{afflerbach_2014,gelcich_2015,lester_2017}.
+Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}.
+In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \citep{costello_2012}.
+One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves, thus creating TURF-reserve systems \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
-TURFs are a fisheries management tool in which a well-defined group of
-fishers (\emph{e.g.} fishing cooperatives) have exclusive access to an
-explicitly delimited portion of the ocean. They promote a sense of
-stewardship and incentivise resource users to sustainably manage their
-resources \citep{gelcich_2008,costello_2010,mccay_2014}. On the other
-hand, no-take marine reserves (marine reserves from hereinafter) are
-areas where all extractive activities are off-limits. These can be
-implemented to protect biodiversity but also as fishery management tools
-to aid in the recovery of marine stocks. These instruments can be
-combined by establishing a marine reserve within a TURF, thus making
-them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.
+TURFs are a fisheries management tool in which a well-defined group of fishers (\emph{e.g.} fishing cooperatives) have exclusive access to an explicitly delimited portion of the ocean.
+They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,costello_2010,mccay_2014}.
+On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits.
+These can be implemented to protect biodiversity but also as fishery management tools to aid in the recovery of marine stocks.
+These instruments can be combined by establishing a marine reserve within a TURF, thus making them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.
 
-Conservation science has shown how marine reserves lead to increased
-biomass, species richness, and abundance within the protected regions
-\citep{lester_2009}, and that these may have a series of additional
-benefits such as climate change mitigation, protection from
-environmental variability, and fisheries benefits
-\citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
-research on TURFs has shown that these areas have higher abundance of
-targeted species than sites operating under open access and even similar
-to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The
-benefits resulting from reserves established within TURFs (\emph{i.e.}
-TURF-reserves) should be captured exclusively by the group of fishers
-with exclusive access \citep{gelcich_2015}. Although in theory these
-systems are successful \citep{smallhornwest_2018}, there is little
-empirical evidence of their effectiveness and the drivers of their
-success. Moreover, TURF-reserve systems are inherently intricate
-social-ecological systems, and their effectiveness must depend on how
-environmental and social factors combine and interact
-\citep{ostrom_2009,gelcich_2015}. It is therefore important to consider
-not only the indicators of interest, but also the governance settings
-under which the reserves operate.
+Conservation science has shown how marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \citep{gelcich_2008,gelcich_2012}.
+The benefits resulting from reserves established within TURFs (\emph{i.e.} TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \citep{gelcich_2015}.
+Although in theory these systems are expected to be successful \citep{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success.
+Moreover, TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{ostrom_2009,gelcich_2015}.
+It is therefore important to consider not only the indicators of interest, but also the governance settings under which the reserves operate.
 
-Recent norms in fisheries regulation in Mexico provide a ripe
-opportunity to study the effectiveness of community-based TURF-reserves
-in small-scale fisheries. In Mexico, a norm created in 2014 allows
-fishers to request legal recognition of community-based reserves as
-``Fish Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since
-2012, old and new marine reserves have gained legal recognition as Fish
-Refuges. Of these, 18 were originally implemented within TURFs. However,
-their effectiveness has not yet been formally evaluated and reported in
-the scientific literature.
+Recent changes in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries.
+In Mexico, a legal framework created in 2012 allows fishers to request legal recognition of community-based reserves as ``Fish Refuges'' (\emph{Zona de Refugio Pesquero}, described in more detail below; \citet{nom}).
+Since 2012, 45 old and new marine reserves have gained legal recognition as Fish Refuges.
+Of these, 18 were originally implemented within TURFs.
+However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
 
-Here, we combine causal inference techniques and the Social-Ecological
-Systems (SES) framework to evaluate community-based TURF-reserves in
-three coastal communities in Mexico. The objective of this work is
-twofold. First, to provide a holistic evaluation of the effectiveness of
-community-based TURF-reserves in terms of the changes in biological and
-socioeconomic indicators and the governance settings under which these
-develop, which may inform similar processes in other countries. Second,
-to identify opportunities where improvement or adjustment might lead to
-increased effectiveness. We draw from lessons learned in these three
-case studies and provide management recommendations to maximize the
-effectiveness of community-based TURF-reserves in small-scale fisheries
-where this tool is used to manage and rebuild coastal fisheries.
+Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico.
+The objective of this work is twofold.
+First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries.
+Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness.
+We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.
 
@@
 
-Before discussing our data collection methods and describing our
-analyses, our case studies warrant some background. Community-based
-marine reserves that are implemented within TURFs are a form of
-TURF-reserve, voluntarily established and enforced by local communities.
-This bottom-up approach increases compliance and self-enforcement, and
-reserves can yield benefits similar to systematically-designed reserves
-\citep{beger_2004,smallhornwest_2018}. Community-based spatial closures
-occur in different contexts, like the \emph{kapu} or \emph{ra'ui} areas
-in the Pacific Islands \citep{johannes_2002,bohnsack_2004}. However,
-community-based reserves can be hard to enforce if they are not legally
-recognized. In such conditions, TURF fishers must rely on the exclusive
-access of the TURF to maintain high levels of compliance.
+Community-based marine reserves that are implemented within TURFs are a form of TURF-reserve, voluntarily established and enforced by local communities.
+This bottom-up approach can increase compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
+Community-based spatial closures occur in different contexts, like the \emph{kapu} or \emph{ra'ui} areas in the Pacific Islands \citep{johannes_2002,bohnsack_2004}.
+However, community-based reserves can be hard to enforce if they are not legally recognized.
+In such conditions, TURF fishers must rely on the exclusive access of the TURF to maintain high levels of compliance.
 
-In an effort to bridge this normative gap, Mexican Civil Society
-Organizations (CSOs) served as a link between fishers and government,
-and helped create a legal framework that solves this governance issue:
-Fish Refuges \citep{nom}. Fish Refuges can be implemented as permanent,
-temporary or partial reserves, which can protect one, some, or all
-resources within their boundaries. One of the ways in which fishing
-communities have taken advantage of this new tool is by implementing
-temporary marine reserves within their TURFs with a defined expiration
-date (often five years). When the expiration date is reached, fishers
-can chose to open the reserves to fishing or re-establish them. Our work
-focuses on Fish Refuges implemented as community-based TURF-reserves in
-small-scale fisheries.
+In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and helped create a legal framework that solves this governance issue: Fish Refuges \citep{nom}.
+Fish Refuges can be implemented as permanent, temporary or partial reserves, which can protect one, some, or all resources within their boundaries.
+One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporary marine reserves within their TURFs with a defined expiration date (often five years).
+When the expiration date is reached, fishers can chose to open the reserves to fishing or re-establish them.
+Our work focuses on Fish Refuges implemented as community-based TURF-reserves in small-scale fisheries.
 
-The most common setup of community-based TURF-reserves in Mexico is the
-following. Fishers from a given community are assembled in fishing
-cooperatives which have exclusive fishing rights over a spatially
-delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
-\ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and
-each community usually hosts no more than one cooperative. The profits
-from each TURF are shared amongst all fishers from the cooperative.
-Fishing cooperatives interested in implementing marine reserves work
-with CSOs to implement marine reserves within their TURFs (\emph{i.e.}
-TURF-reserves). Fishers then ask the government to grant legal
-recognition to their TURF-reserves as Fish Refuges following a series of
-studies outlined in the regulation \citep{nom}.
+The most common setup of community-based TURF-reserves in Mexico is the following.
+Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig \ref{fig:map}A).
+Each TURF is exclusively fished by one cooperative, and each community usually hosts no more than one cooperative.
+The profits from each TURF are shared amongst all fishers from the cooperative.
+Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (\emph{i.e.} TURF-reserves).
+Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges after conducting a series of studies outlined in the regulation, namely the creation of a baseline of ecological data and historical performance of the fishery \citep{nom}.
 
@@
 
-We evaluate three community-based no-take TURF-reserve systems
-implemented in Mexican TURF-managed fisheries, therefore making them
-TURF-reserves (Fig \ref{fig:map}A). The first one was created by the
-\emph{Buzos y Pescadores de la Baja California} fishing cooperative,
-located in Isla Natividad in the Baja California Peninsula (Fig
-\ref{fig:map}B). The main fishery in the island is the spiny lobster
-(\emph{Panulirus interruptus}), but other resources like finfish, sea
-cucumber, sea urchin, snail, and abalone are also an important source of
-income. In 2006, the community decided to implement two marine reserves
-within their fishing grounds. The objective of these reserves was ``to
-protect commercially important invertebrate species''; mainly lobster
-and abalone. The reserves obtained legal recognition in 2018
-\citep{dof_website_2018}.
+We evaluate three community-based no-take TURF-reserve systems implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig \ref{fig:map}A).
+The first one was created by the \emph{Buzos y Pescadores de la Baja California} fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig \ref{fig:map}B).
+The main fishery in the island is the spiny lobster (\emph{Panulirus interruptus}), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income.
+In 2006, the community decided to implement two marine reserves within their fishing grounds.
+The objective of these reserves was ``to protect and recover stocks of commercially important invertebrate species''; mainly lobster and abalone.
+The reserves obtained legal recognition in 2018 \citep{dof_website_2018}.
 
-The other two TURF-reserve systems are located in Maria Elena and Punta
-Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
-Isla Nativdad, which hosts a well-established fishing community, Maria
-Elena is a fishing camp visited intermittently during the fishing season
-that belongs to the \emph{Cozumel} fishing cooperative. Punta Herrero is
-home to the \emph{José María Azcorra} fishing cooperative, and similar
-to Isla Natividad hosts a small community. Their main fishery is the
-Caribbean spiny lobster (\emph{Panulirus argus}), but they also target
-finfish in the off-season. Maria Elena and Punta Herrero established
-eight and four marine reserves in 2012 and 2013, respectively. These
-reserves have been legally recognized as Fishing Refuges since their
-original implementation \citep{dof_website_2012,dof_website_2013} and
-subsequent re-establishments \citep{dof_website_2017b}.
+The other two TURF-reserve systems are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C).
+In contrast with Isla Nativdad, which hosts a well-established fishing community, Maria Elena is a fishing camp visited intermittently during the fishing season that belongs to the \emph{Cozumel} fishing cooperative.
+Punta Herrero is home to the \emph{José María Azcorra} fishing cooperative, and similar to Isla Natividad hosts a small community.
+Their main fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but they also target finfish in the off-season.
+Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively.
+These reserves have been legally recognized as Fishing Refuges since their original implementation \citep{dof_website_2012,dof_website_2013} and subsequent re-establishments \citep{dof_website_2017b}.
 
-These communities are representative of their region in terms of
-ecology, socioeconomic, and governance aspects. Isla Natividad, for
-example, is part of a greater group of fishing cooperatives belonging to
-a Federation of Fishing Cooperatives. This group has been identified as
-a cohesive group that cooperates to better manage their resources
-\citep{mccay_2014,mccay_2017,acevesbueno_2017}. Likewise, Maria Elena
-and Punta Herrero are representative of fishing cooperatives in the
-Mexican Caribbean, which are also part of a regional Federation.
-Together, these three communities provide an accurate representation of
-other fishing communities that have been historically manged with TURFs
-in each of their regions. While each region has additional communities
-that have established community-based TURF-reserves, available data
-would not allow us to perform the in-depth causal inference analysis
-that we undertake. Yet, given the similarities among communities and the
-socioeconomic and governance setting under which they operate, it is
-safe to cautiously generalize our insights to other similar
-community-based TURF-reserves in Mexico and elsewhere.
+These communities are representative of their region in terms of ecology, socioeconomic, and governance aspects.
+Isla Natividad, for example, is part of a greater group of fishing cooperatives belonging to a Federation of Fishing Cooperatives.
+This group has been identified as a cohesive group that cooperates to better manage their resources \citep{mccay_2014,mccay_2017,acevesbueno_2017}.
+Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation.
+Together, these three communities provide an accurate representation of other fishing communities that have been historically manged with TURFs in each of their regions.
+While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth causal inference analysis that we undertake.
+Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our insights to other similar community-based TURF-reserves in Mexico and elsewhere.
 
-The regulation governing the implementation of Fish Refuges states that
-these are fishery management tools intended to have biological or
-socioeconomic benefits \citep{nom}. For this reason, the main portion of
-our analyses focuses on a series of biological and socioeconomic
-indicators that may respond to reserve implementation. However, the
-effectiveness of conservation and fisheries management interventions
-also depends on the social and governance structures in place. We
-therefore incorporate a reduced version of the Social Ecological Systems
-framework \citep{ostrom_2009} and evaluate variables and indicators
-known to aid and hinder the effectiveness of management interventions in
-conservation and fisheries. The incorporation of the SES is not intended
-to relate different levels of governance with reserve effectiveness, but
-rather help provide context on the social-ecological system in which
-reserves develop. The following two sections describe our data
-collection methods and analyses.
+The regulation governing the implementation of Fish Refuges states that these are fishery management tools intended to have conservation and fisheries benefits \citep{nom}.
+For this reason, the main portion of our analyses focuses on a series of biological and socioeconomic indicators that may respond to reserve implementation.
+However, the effectiveness of conservation and fisheries management interventions also depends on the social and governance structures in place.
+We therefore incorporate a reduced version of the Social Ecological Systems framework \citep{ostrom_2009} and evaluate variables and indicators known to aid and hinder the effectiveness of management interventions in conservation and fisheries.
+The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather help provide context on the social-ecological system in which reserves develop.
+The following two sections describe our data collection methods and analyses.
 
@@
 
-We use three main sources of information to evaluate these reserves
-across ecological, socioeconomic, and governance dimensions. Ecological
-data come from the annual ecological monitoring of reserve and control
-sites. Reserve sites are areas where no fishing occurs. Control sites
-are areas that meet the following criteria: i) habitat characteristics
-are similar to the corresponding reserves, ii) presumably had a similar
-probability of being selected as reserves during the design phase, iii)
-are located within the TURF, where fishing occurs, and iv) are not
-directly adjacent to the reserves. We focus our evaluation on sites
-where data are available for reserve and control sites, before and after
-the implementation of the reserve. This provides us with a
-Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
-allows us to capture and control for temporal and spatial dynamics
-\citep{stewartoaten_1986,depalma_2018} and causally attribute the
-changes to the reserve
-\citep{francinifilho_2008,Villasenor-Derbez_2018}.
+We use three main sources of information to evaluate these reserves across ecological, socioeconomic, and governance dimensions.
+Ecological data come from the annual ecological monitoring of reserve and control sites.
+Reserve sites are areas where no fishing occurs.
+Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) are not directly adjacent to the reserves.
+We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
+This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally attribute the changes to the reserve \citep{francinifilho_2008,Villasenor-Derbez_2018}.
 
-The biological data are collected by members from each community and
-personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
-(\href{www.cobi.org.mx}{COBI}). Trained divers record species richness
-and abundances of fish and invertebrate species along replicate
-transects (30 \(\times\) 2 m each) at depths 5-20 m in the reserves and
-control sites \citep{suman_2010-ez,fulton_2018,fulton_2019}. Size
-structures are also collected during fish surveys. All sites were
-surveyed annually, and at least once before implementation of the
-reserves. A summary of sampling effort is shown in the supplementary
-materials (Tables S1-S2).
+The biological data are collected by members from each community and personnel from the Mexican CSO \emph{Comunidad y Biodiversidad} (\href{www.cobi.org.mx}{COBI}).
+Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 \(\times\) 2 m each) at depths 5-20 m in the reserves and control sites, where a minimum of 4 transects per site are performed \citep{suman_2010-ez,fulton_2018,fulton_2019}.
+Size structures are also collected during fish surveys.
+All sites were surveyed annually, and at least once before implementation of the reserves.
+A summary of sampling effort is shown in the supplementary materials (Tables S1-S2).
 
-Socioeconomic data come from landing receipts reported to the National
-Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
-Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
-(Kg) and revenues (MXP) for TURF-managed cooperatives with and without
-marine reserves. In this case our treated unit are the cooperatives
-(\emph{i.e.} communities) that have implemented a reserve within their
-TURF, and the controls are nearby communities that have a TURF but did
-not implement a reserve. Cooperatives incorporated in this analysis
-belong to larger regional-level Cooperative Federations, and are exposed
-to the same markets and institutional frameworks, making them plausible
-controls \citep{mccay_2017,ayer_2018}. Landings and revenues were
-aggregated at the cooperative-year level, and revenues were adjusted to
-represent 2014 values by the Consumer Price Index for Mexico
-\citep{oecd_2017}. A table with summary statistics for this data is
-provided in the supplementary materials (Table S3, Figure S5).
+Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de Acuacultura y Pesca}; CONAPESCA).
+Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves.
+In this case our treated unit are the cooperatives (\emph{i.e.} communities) that have implemented a reserve within their TURF, and the controls are nearby communities that have a TURF but did not implement a reserve.
+These communities have TURFs adjacent to the TURFs of our focal communities.
+Cooperatives incorporated in this analysis have similar number of members, belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2014,mccay_2017,ayer_2018}.
+Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017}.
+A table with summary statistics for this data is provided in the supplementary materials (Table S3, Figure S5).
 
-Data for the evaluation of the SES were collected at the community-level
-from official documents used in the design, creation, and designation of
-the marine reserves. These include the technical studies that the
-cooperatives submit when they request recognition of their reserves, as
-well as the official enactments
-\citep{dof_website_2012,dof_website_2013,dof_website_2018}. We also
-complimented information based on the authors' experience and knowledge
-of the communities. We collected information on the Resource Systems,
-Resource Units, Actors, and Governance System (Table \ref{table:ses}).
+Data for the evaluation of the SES were collected at the community-level from official documents used in the design, creation, and designation of the marine reserves.
+These include the technical studies that the cooperatives submit when they request recognition of their reserves, as well as the official enactments \citep{dof_website_2012,dof_website_2013,dof_website_2018}.
+We also complimented information based on the authors' experience and knowledge of the communities.
+We collected information on the Resource Systems, Resource Units, Actors, and Governance System (Table \ref{table:ses}).
 
@@
 
-We evaluate the effect that the TURF-reserves have had on four
-ecological and two socioeconomic indicators shown in Table
-\ref{table:indicators}. Recall that reserves were implemented to protect
-lobster and other benthic invertebrates. However, we also use the
-available fish and invertebrate data to test for associated co-benefits.
+We evaluate the effect that the TURF-reserves have had on four ecological and two socioeconomic indicators shown in Table \ref{table:indicators}.
+Recall that reserves were implemented to protect lobster and other benthic invertebrates.
+However, we also use the available fish and invertebrate data to test for associated co-benefits.
 
-We use a difference-in-differences analysis to evaluate these
-indicators. This approach is widely used in econometric literature to
-estimate the average treatment effect of an intervention, like the
-impact of minimum wage increases on employment rates \citep{card_1994}.
-In our case it allows us to estimate the effect that the reserve had on
-each biological and socioeconomic indicator (Table
-\ref{table:indicators}) by comparing trends across time and treatments
-\citep{moland_2013,Villasenor-Derbez_2018}. To perform
-difference-in-differences, we regress the indicator of interest on a
-dummy variable for treatment, a dummy variable for years, and the
-interaction term between these with a multiple linear regression of the
-form:
+We use a difference-in-differences analysis to evaluate these indicators.
+This approach is widely used in econometric literature to estimate the average treatment effect of an intervention, like the impact of minimum wage increases on employment rates \citep{card_1994}.
+In our case it allows us to estimate the effect that the reserve had on each biological and socioeconomic indicator (Table \ref{table:indicators}) by comparing trends across time and treatments since reserve implementation \citep{moland_2013,Villasenor-Derbez_2018}.
+To perform difference-in-differences, we regress the indicator of interest on a dummy variable for treatment, a dummy variable for years, and the interaction term between these with a multiple linear regression of the form:
 
@@
 
-Where year-level fixed effects capturing a temporal trend are
-represented by \(\gamma_t Year_t\), and \(\beta Zone_i\) captures the
-difference between reserve (\(Zone = 1\)) and control (\(Zone = 0\))
-sites. The effect of the reserve is captured by the \(\lambda_t\)
-coefficient, and represents the difference observed between the control
-site before the implementation of the reserve and the treated sites at
-time \(t\) after controlling for other time and space variations
-(\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Therefore, we
-would expect this term to be positive if the indicator increases because
-of the reserve. Finally, \(\epsilon_{i,t}\) represents the error term of
-the regression.
+Where year-level fixed effects capturing a temporal trend are represented by \(\gamma_t Year_t\), and \(\beta Zone_i\) captures the difference between reserve (\(Zone = 1\)) and control (\(Zone = 0\)) sites.
+The effect of the reserve is captured by the \(\lambda_t\) coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time \(t\) after controlling for other time and space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
+Therefore, we would expect this term to be positive if the indicator increases because of the reserve.
+Finally, \(\epsilon_{i,t}\) represents the error term of the regression.
 
-Socioeconomic indicators are evaluated with a similar approach. Due to
-data constrains, we only evaluate socioeconomic data for Isla Natividad
-(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
-used as counterfactuals that allow us to control for unobserved
-time-invariants. Each focal community (\emph{i.e.} Isla Natividad and
-Maria Elena) has three counterfactual communities.
+Socioeconomic indicators are evaluated with a similar approach.
+Due to data constrains, we only evaluate socioeconomic data for Isla Natividad (2000 - 2014) and Maria Elena (2006 - 2013).
+Neighboring communities are used as counterfactuals that allow us to control for unobserved time-invariants.
+Each focal community (\emph{i.e.} Isla Natividad and Maria Elena) has three counterfactual communities.
 
@@
 
-The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio},
-but in this case the \(Treated\) dummy variable indicates if the
-community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)).
-These regression models allow us to establish a causal link between the
-implementation of marine reserves and the observed trends by accounting
-for temporal and site-specific dynamics \citep{depalma_2018}. We fit one
-model per community and indicators (\emph{e.g.} there are three models
-for lobster density, one for each community) for a total of 12
-biological model fits and four socioeconomic model fits. Model
-coefficients were estimated via ordinary least-squares and used
-heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
-analyses were performed in R version 3.5.2 and R Studio version 1.1.456
-\citep{R_2018}. All data and code needed to reproduce our analyses are
-available in a GitHub repository at:
-\url{https://github.com/jcvdav/ReserveEffect}.
+The coefficient interpretations remains as for Eq.
+\ref{eqn:reg_bio}, but in this case the \(Treated\) dummy variable indicates if the community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)).
+These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \citep{depalma_2018}.
+For each indicator in each community, we fit one model (\emph{e.g.} there are three models for lobster density, one for each community) for a total of 12 biological model fits and four socioeconomic model fits.
+Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard errors \citep{zeileis_2004-7n}.
+All analyses were performed in R version 3.5.2 and R Studio version 1.1.456 \citep{R_2018}.
+All data and code needed to reproduce our analyses are available in a GitHub repository at: \url{https://github.com/jcvdav/ReserveEffect}.
 
-We use the SES framework to evaluate each community and create a
-narrative that provides context for each community. The use of this
-framework standardizes our analysis and allows us to communicate our
-results in a common language across fields by using a set of previously
-defined variables and indicators. We based our variable selection
-primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who
-operationalized and analyzed Mexican fishing cooperatives using this
-framework, and identified the key variables relevant to fishing
-cooperatives in Mexico. We also incorporate other relevant variables
-known to influence reserve performance following
-\citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table
-\ref{table:ses} shows the selected variables, along with definitions and
-values.
+We use the SES framework to evaluate each community and create a narrative that provides context for each community.
+The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators.
+Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as in \citet{leslie_2015-na}), we followed a similar approach to \citet{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems.
+We based our variable selection primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework, and identified the key variables relevant to fishing cooperatives in Mexico.
+We also incorporate other relevant variables known to influence reserve performance following \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.
+Table \ref{table:ses} shows the selected variables, along with definitions and values.
 
@@
 
-The following sections present the effect that marine reserves had on
-the biological and socioeconomic indicators for each coastal community.
-Results are presented in terms of difference through time and across
-sites, relative to the control site on the year of implementation
-(\emph{i.e.} the difference-in-differences estimate or effect size
-\(\lambda_t\) from Eqs. \ref{eqn:reg_bio} and \ref{eqn:soc_reg}). We
-also provide an overview of the governance settings of each community,
-and discuss how these might be related to the effectiveness and
-performance of the reserves.
+The following sections present the effect that marine reserves had on the biological and socioeconomic indicators for each coastal community.
+Results are presented in terms of difference through time and across sites, relative to the control site on the year of implementation (\emph{i.e.} the difference-in-differences estimate or effect size \(\lambda_t\) from Eqs.
+\ref{eqn:reg_bio} and \ref{eqn:soc_reg}).
+We also provide an overview of the governance settings of each community, and discuss how these might be related to the effectiveness and performance of the reserves.
 
@@
 Indicators showed ambiguous responses through time for each reserve.
-Figure \ref{fig:indicators}A shows positive effect sizes for lobster
-densities in Isla Natividad and Punta Herrero during the first years,
-but the effect is eroded through time. In the case of Maria Elena,
-positive changes were observed in the third and fourth year. These
-effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
-Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
-for Maria Elena, but are not significantly different from zero
-(\(p > 0.05\)). Likewise, no significant changes were detected in fish
-biomass or invertebrate and fish densities (Fig.
-\ref{fig:indicators}B-D), where effect sizes oscillated around zero
-without clear trends. Figures and tables with time series of indicators
-and model coefficients are presented in the supplementary materials
-(Figures S1-S4, Tables S4-S6).
+Figure \ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time.
+In the case of Maria Elena, positive changes were observed in the third and fourth year.
+These effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\) for Maria Elena, but are not significantly different from zero (\(p > 0.05\)).
+Likewise, no significant changes were detected in fish biomass or invertebrate and fish densities (Fig.
+\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
+Figures and tables with time series of indicators and model coefficients are presented in the supplementary materials (Figures S1-S4, Tables S4-S6).
 
@@
 
-Lobster landings and revenue were only available for Isla Natividad and
-Maria Elena (Fig \ref{fig:lobsters}). For all years before
-implementation, the effect sizes are close to zero, indicating that the
-control and treatment sites have similar pre-treatment trends,
-suggesting that these are plausible controls. However, effect sizes do
-not change after the implementation of the reserve. Interestingly, the
-negative effect observed for Isla Natividad on year 5 corresponds to the
-2011 hypoxia events \citep{micheli_2012-EU}. The only positive change
-observed in lobster landings is for Isla Natividad in 2014
-(\(p < 0.1\)). The three years of post-implementation data for Maria
-Elena do not show a significant effect of the reserve. Isla Natividad
-shows higher revenues after the implementation of the reserve, as
-compared to the control communities. However, these changes are only
-significant for the third year (\(p < 0.05\)). Full tables with model
-coefficients are presented in the supplementary materials (Tables
-S4-S5).
+Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}).
+For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, suggesting that these are plausible controls.
+However, effect sizes do not change after the implementation of the reserve.
+Interestingly, the negative effect observed for Isla Natividad on year 5 corresponds to the 2011 hypoxia events \citep{micheli_2012-EU}.
+The only positive change observed in lobster landings is for Isla Natividad in 2014 (\(p < 0.1\)).
+The year of post-implementation data for Maria Elena does not show a significant effect of the reserve.
+Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities.
+However, these changes are only significant for the third year (\(p < 0.05\)).
+Full tables with model coefficients are presented in the supplementary materials (Tables S4-S5).
 
@@
 
-Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed
-communities share similarities known to foster sustainable resource
-management and increase reserve effectiveness. For example, fishers
-operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide
-exclusive access to resources and reserves. Along with their relatively
-small groups (A1 - Number of relevant actors), Isolation (A3),
-Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated
-sanctions (GS10.1), these fisheries have solid governance structures
-that enable them to monitor their resources and enforce rules to ensure
-sustainable management. In general, success of conservation initiatives
-depends on the incentives of local communities to maintain a healthy
-status of the resources upon which they depend \citep{jupiter_2017}. Due
-to the clarity of access rights and isolation, the benefits of
-conservation directly benefit the members of the fishing cooperatives,
-which have favored the development of efficient community-based
-enforcement systems. However, our SES analysis also highlights factors
-that might hinder reserve performance or mask outcomes. While total
-reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual
-reserves are often small (RS3); the largest reserve is only 4.37 km
-\textsuperscript{2}, and the smallest one is 0.09 km
-\textsuperscript{2}. Reserves are also relatively young (RS5).
-Additionally, fishers harvest healthy stocks (RS4.1), and it is unlikely
-that marine reserves will result in increased catches.
+Our qualitative implementation of the social-ecological systems framework allowed us to systematically identify important differences between the case studies' governance systems and incorporate other characteristics of these fisheries neglected during the process of data collection (Table \ref{table:ses}).
+We find that the analyzed communities share similarities known to foster sustainable resource management and increase reserve effectiveness.
+For example, fishers operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves.
+Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have solid governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management.
+In general, success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources upon which they depend \citep{jupiter_2017}.
+Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperatives, which have favored the development of efficient community-based enforcement systems.
+However, our SES analysis also highlights factors that might hinder reserve performance or mask outcomes.
+While total reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual reserves are often small (RS3); the largest reserve is only 4.37 km \textsuperscript{2}, and the smallest one is 0.09 km \textsuperscript{2}.
+Reserves are also relatively young (RS5).
+Additionally, fishers harvest healthy stocks (RS4.1), and it is unlikely that marine reserves will result in increased catches.
 
@@
 
-Our results indicate that these TURF-reserves have not increased lobster
-densities. Additionally, no co-benefits were identified when using other
-ecological indicators aside from the previously reported buffering
-effect that reserves can have to environmental variability in Isla
-Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
-pertaining landings and revenues showed little to no change after
-reserve implementation. Lastly, the communities exhibit all the social
-enabling conditions for effective reserve and resource management. Here
-we discuss possible shortcomings in our analyses as well as possible
-explanations for the observed patterns.
+Our results indicate that these TURF-reserves have not increased lobster densities.
+Additionally, no co-benefits were identified when using other ecological indicators aside from the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \citep{micheli_2012-EU}.
+The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
+Lastly, the communities exhibit all the social enabling conditions for effective reserve and resource management.
+Here we discuss possible shortcomings in our analyses as well as possible explanations for the observed patterns.
 
-While many ecology studies have used BACI sampling designs and
-respective analyses (\emph{e.g.} \citet{stewartoaten_1986}), few
-conservation studies have done so to evaluate the effect of an
-intervention (\emph{e.g.}
-\citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted
-in a call for more robust analyses in conservation science
-\citep{guidetti_2002,ferraro_2006}. Our approach to evaluate the
-temporal and spatial changes provides a more robust measure of reserve
-effectiveness, and captures previously described patterns. For example,
-the rapid increase observed for lobster densities in Isla Natividad on
-the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), occurs a
-year after the hypoxia events described by \citet{micheli_2012-EU},
-which caused mass mortality of sedentary organisms such as abalone and
-sea urchins, but not lobster and finfish. The use of causal inference
-techniques may help us support evidence-based conservation.
+While many ecology studies have used BACI sampling designs and respective analyses (\emph{e.g.} \citet{stewartoaten_1986}), few conservation studies have done so to evaluate the effect of an intervention (\emph{e.g.} \citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted in a call for more robust analyses in conservation science \citep{guidetti_2002,ferraro_2006}.
+Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness, and captures previously described patterns.
+For example, the rapid increase observed for lobster densities in Isla Natividad on the sixth year (\emph{i.e.} 2012; Fig.
+\ref{fig:indicators}A), occurs a year after the hypoxia events described by \citet{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
+The use of causal inference techniques may help us support evidence-based conservation.
 
-Our analyses of socioeconomic indicators has two limitations. First, we
-only look at landings and revenues by landings for communities with and
-without TURF-reserves. There are a number of other possible indicators
-that could show a change due to the implementation of the reserve.
-Notably, one often cited in the literature is additional benefits, such
-as tourism \citep{viana_2017}. However, it is unlikely that the
-evaluated communities will experience tourism benefits due to their
-remoteness and the lack of proper infrastructure to sustain tourism. A
-second limitation of our socioeconomic analysis is that we do not
-observe effort data, which may mask the effect of the reserve. For
-example, if catches remain relatively unchanged but fishing effort
-decreased, that would imply a larger catch per unit effort and thus
-higher profitability, provided that cost per unit effort does not
-increase.
+Our analyses of socioeconomic indicators has three limitations.
+First, we only look at landings and revenues by landings for communities with and without TURF-reserves.
+There are a number of other possible indicators that could show a change due to the implementation of the reserve.
+Notably, one often cited in the literature is additional benefits, such as tourism \citep{viana_2017}.
+However, it is unlikely that the evaluated communities will experience tourism benefits due to their remoteness and the lack of proper infrastructure to sustain tourism.
+A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve.
+For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and thus higher profitability, provided that cost per unit effort does not increase.
+Likewise, it is possible that fishing effort increased around reserves to maintain the historical levels of landings.
+A final limitation applies to Maria Elena, where we only observe landings and income for one year after reserve implementation.
+While one would not expect to observe increased landings or income in such a short period, a spatial closure might cause total catches to decline, especially if effort is held constant.
 
-A first possible explanation for the lack of effectiveness may be the
-young age of the reserves. Literature shows that age and enforcement are
-important factors that influence reserve effectiveness
-\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
-reserves, and our SES analysis suggests that all communities have a
-well-established community-based enforcement system. With these
-characteristics, one would expect the reserves to be effective. Maria
-Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
-\textless{} 6 years old) and effects may not yet be evident due to the
-short duration of protection, relative to the life histories of the
-protected species; community-based marine reserves in tropical
-ecosystems may take six years or more to show a spillover effect
-\citep{dasilva_2015-zX}.
+A first possible explanation for the lack of effectiveness may be the young age of the reserves.
+Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO,babcock_2010}.
+Isla Natividad has the oldest reserves, and our SES analysis suggests that all communities have a well-established community-based enforcement system.
+With these characteristics, one would expect the reserves to be effective.
+Maria Elena and Punta Herrero are relatively young reserves (\emph{i.e.} \textless{} 6 years old; RS5 in Table \ref{table:ses}) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \citep{dasilva_2015-zX}.
 
-Another key condition for effectiveness is reserve size
-\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
-attributed to poor ecological coherence in reserve design (\emph{sensu}
-\citet{rees_2018}). Previous research has shown that reserves in Isla
-Natividad yield fishery benefits for the abalone fishery
-\citep{rossetto_2015-V0}, however, abalone are less mobile than
-lobsters, and perhaps the reserves provide enough protection to these
-sedentary invertebrates, but not lobsters. Design principles developed
-by \citet{green_2017} for marine reserves in the Caribbean state that
-reserves ``should be more than twice the size of the home range of
-adults and juveniles'', and suggest that reserves seeking to protect
-spiny lobsters should have at least 14 km across. Furthermore, fishers
-may favor implementation of reserves that pose low fishing costs due to
-their small size or location. Our analysis of economic data supports
-this hypothesis, as neither landings nor revenues showed the expected
-short-term reductions associated to the first years of reserve
-implementation \citep{ovando_2016-Wg}.
+Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (\emph{sensu} \citet{rees_2018}).
+Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}, however, abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary invertebrates, but not lobsters.
+Design principles developed by \citet{green_2017} for marine reserves in the Caribbean state that reserves ``should be more than twice the size of the home range of adults and juveniles'', and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
+As shown through the SES analysis, the size of the marine reserves appears small compared to the movement capacity of the main targeted species (RU1, RS3; Table \ref{table:ses}).
+Furthermore, fishers may favor implementation of reserves that pose low fishing costs due to their small size or location.
+Our analysis of economic data supports this hypothesis, as neither landings nor revenues showed the expected short-term reductions associated to the first years of reserve implementation \citep{ovando_2016-Wg}.
 
-Even if reserves had appropriate sizes and were placed in optimal
-locations, there are other plausible explanations for the observed
-patterns. For instance, marine reserves are only likely to provide
-fisheries benefits if initial population sizes are low and the fishery
-is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
-fisheries were certified by the Marine Stewardship Council and are
-managed via species-specific minimum catch sizes, seasonal closures,
-protection of ``berried'' females, and escapement windows where traps
-are allowed \citep{dof_website_1993}. It is uncertain whether such a
-well-managed fishery will experience additional benefits from marine
-reserves; reserves implemented in TURFs where fishing pressure is
-already optimally managed will still show a trade-off between fisheries
-and conservation objectives \citep{lester_2017}. Furthermore,
-\citet{gelcich_2008} have shown that TURFs alone can have greater
-biomass and richness than areas operating under open access. This might
-reduce the difference between indicators from the TURF and reserve
-sites, making it difficult to detect such a small change. Further
-research should focus on evaluating sites in the reserve, TURF, and open
-access areas or similar Fish Refuges established without the presence of
-TURFs where the impact of the reserves might be greater.
+Even if reserves had appropriate sizes and were placed in optimal locations, there are other plausible explanations for the observed patterns.
+For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2004,hilborn_2006}.
+Both lobster fisheries were certified by the Marine Stewardship Council and are managed via species-specific minimum catch sizes, seasonal closures, protection of ``berried'' females, and escapement windows where traps are allowed \citep{dof_website_1993}.
+It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves; reserves implemented in TURFs where fishing pressure is already optimally managed will still show a trade-off between fisheries and conservation objectives \citep{lester_2017}.
+Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open access.
+This might reduce the difference between indicators from the TURF and reserve sites, making it difficult to detect such a small change.
+Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of the reserves might be greater.
 
-Finally, extreme conditions, including prolonged hypoxia, heat waves,
-and storms have affected both the Pacific and Caribbean regions, with
-large negative impacts on coastal marine species and ecosystems
-\citep{cavole_2016,hughes_2018,breitburg_2018}. The coastal ecosystems
-where these reserves are located have been profoundly affected by these
-events \citep{micheli_2012-EU,woodson_2018}. Effects of protection might
-be eliminated by the mortalities associated with these extreme
-conditions.
+Finally, extreme conditions, including prolonged hypoxia, heat waves, and storms have affected both the Pacific and Caribbean regions, with large negative impacts on coastal marine species and ecosystems \citep{cavole_2016,hughes_2018,breitburg_2018}.
+The coastal ecosystems where these reserves are located have been profoundly affected by these events \citep{micheli_2012-EU,woodson_2018}.
+Effects of protection might be eliminated by the mortalities associated with these extreme conditions.
 
-While the evaluated reserves have failed to provide fishery benefits to
-date, there are a number of additional ecological, fisheries, and social
-benefits. Marine reserves provide protection to a wider range of species
-and vulnerable habitat. Previous research focusing on these specific
-sites has shown that they serve as an insurance mechanism against
-uncertainty and errors in fisheries management, as well as mild
-environmental shocks
-\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
-Self-regulation of fishing effort can serve as a way to compensate for
-future declines associated to environmental variation
-\citep{finkbeiner_2018}. Furthermore, embarking on a marine conservation
-project can bring the community together, which promotes social cohesion
-and builds social capital \citep{fulton_2019}. Showing commitment to
-marine conservation and sustainable fishing practices has allowed
-fishers to have greater bargaining power and leverage over fisheries
-management \citep{prezramrez_2012}. These additional benefits might
-explain why communities show a positive perception about their
-performance and continue to support their presence by re-implementing
-the reserves \citep{ayer_2018}.
+While the evaluated reserves have failed to provide fishery benefits to date, there are a number of additional ecological, fisheries, and social benefits.
+Marine reserves provide protection to a wider range of species and vulnerable habitat.
+Previous research focusing on these specific sites has shown that they serve as an insurance mechanism against uncertainty and errors in fisheries management, as well as mild environmental shocks \citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
+Self-regulation of fishing effort can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}.
+Furthermore, embarking on a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}.
+Showing commitment to marine conservation and sustainable fishing practices has allowed fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}.
+These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-establishing the reserves \citep{ayer_2018}.
 
-Community-based TURF-reserves in small-scale fisheries may be helpful
-conservation and fishery management tools when appropriately implemented
-\citep{gelcich_2015}. We must promote bottom-up design and
-implementation processes like the ones in the evaluated reserves, but
-without setting design principles aside. Having full community support
-surely represents an advantage, but it is important that community-based
-TURF-reserves meet essential design principles such as size and
-placement so as to maximize their effectiveness. Furthermore,
-conservation and advocacy groups should consider the opportunity costs
-of such interventions (\emph{sensu} \citet{smith_2010}) and evaluate the
-potential of other approaches that may yield similar benefits.
+Community-based TURF-reserves in small-scale fisheries may be helpful conservation and fishery management tools when appropriately implemented \citep{gelcich_2015}.
+We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside.
+Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.
+Furthermore, conservation and advocacy groups should consider the opportunity costs of such interventions (\emph{sensu} \citet{smith_2010}) and evaluate the potential of other approaches that may yield similar benefits.
 
-In terms of fisheries regulation in Mexico, our work only evaluates Fish
-Refuges established within TURFs. Future research should aim at
-evaluating other Fish Refuges established as bottom-up processes but
-without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
-others established through top-down processes (\emph{i.e.}
-\citet{dof_websiteU_2018}), as well as the relationship between
-governance and effectiveness across this gradient of approaches. For the
-particular case of the reserves that we evaluate, the possibility of
-expanding reserves or merging existing polygons into larger areas should
-be evaluated and proposed to the communities.
+In terms of fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs.
+Future research should aim at evaluating other Fish Refuges established as bottom-up processes but without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}), others established through top-down processes (\emph{i.e.} \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.
+For the particular case of the reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
 
@@
 
-The authors declare that the research was conducted in the absence of
-any commercial or financial relationships that could be construed as a
-potential conflict of interest.
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
 
@@
 
-JC and AS conceived the idea. JC and EA analyzed data, discussed the
-results, and wrote the first draft. FM, SF, AS, JT, and AHV discussed
-the results and edited the manuscript. All authors provided valuable
-contributions.
+JC and AS conceived the idea.
+JC and EA analyzed data, discussed the results, and wrote the first draft.
+FM, SF, AS, JT, and AHV discussed the results and edited the manuscript.
+All authors provided valuable contributions.
 
@@
 
-JCVD received funding from UCMexus - CONACyT Doctoral Fellowship (CVU
-669403) and the Latin American Fisheries Fellowship Program. AS, AHV, SF
-and JT received funding from Marisla Foundation, Packard Foundation,
-Walton Family Foundation, Summit Foundation, and Oak Foundation. FM was
-supported by NSF-CNH and NSF BioOce (grants DEB-1212124 and 1736830).
+JCVD received funding from UCMexus - CONACyT Doctoral Fellowship (CVU 669403) and the Latin American Fisheries Fellowship Program.
+AS, AHV, SF and JT received funding from Marisla Foundation, Packard Foundation, Walton Family Foundation, Summit Foundation, and Oak Foundation.
+FM was supported by NSF-CNH and NSF BioOce (grants DEB-1212124 and 1736830).
 
@@
 
-The authors wish to acknowledge Imelda Amador for contributions on the
-governance data, as well as pre-processing biological data. This study
-would have not been possible without the effort by members of the
-fishing communities here mentioned, who participated in the
-data-collection process. The authors wish to acknowledge comments by the
-reviewers and editor, which significantly improved the quality of this
-work.
+The authors wish to acknowledge Imelda Amador for contributions on the governance data, as well as pre-processing biological data.
+This study would have not been possible without the effort by members of the fishing communities here mentioned, who participated in the data-collection process.
+The authors wish to acknowledge comments by the reviewers and editor, which significantly improved the quality of this work.
 
@@
 \includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
-\caption{\label{fig:unnamed-chunk-7}\label{fig:map}Location of the three
-coastal communities studied (A). Isla Natividad (B) is located off the
-Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
-in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
-polygons the marine reserves.}
+\caption{\label{fig:unnamed-chunk-7}\label{fig:map}Location of the three coastal communities studied (A).
+Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena and Punta Herrero (C) are located in the Yucatan Peninsula.
+Blue polygons represent the TURFs, and red polygons the marine reserves.}
 \end{figure}
@@
 \includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{\label{fig:unnamed-chunk-8}\label{fig:indicators}Effect sizes for
-marine reserves from Isla Natividad (IN; red circles), Maria Elena (ME;
-blue triangles), and Punta Herrero (PH; green squares) for lobster
-densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
-densities (C), and fish densities (D). Plots are ordered by survey type
-(left column: invertebrates; right column: fish). Points are jittered
-hotizontally to avoid overplotting. Points indicate the effect size and
-error bars are heteroskedastic-robust standard errors. Years have been
-centered to year of implementation.}
+\caption{\label{fig:unnamed-chunk-8}\label{fig:indicators}Effect sizes for marine reserves from Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares) for lobster densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate densities (C), and fish densities (D).
+Plots are ordered by survey type (left column: invertebrates; right column: fish).
+Points are jittered hotizontally to avoid overplotting.
+Points indicate the effect size and error bars are heteroskedastic-robust standard errors.
+Years have been centered to year of implementation.}
 \end{figure}
@@
 \includegraphics{manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
-\caption{\label{fig:unnamed-chunk-9}\label{fig:lobsters}Effect sizes for
-lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
-circles) and Maria Elena (ME; blue triangles). Points are jittered
-hotizontally to avoid overplotting. Points indicate the effect size and
-error bars are heteroskedastic-robust standard errors. Years have been
-centered to year of implementation.}
+\caption{\label{fig:unnamed-chunk-9}\label{fig:lobsters}Effect sizes for lobster catches (A) and revenues (B) in at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles).
+Points are jittered hotizontally to avoid overplotting.
+Points indicate the effect size and error bars are heteroskedastic-robust standard errors.
+Years have been centered to year of implementation.}
 \end{figure}
@@
 \centering
-\begin{tabular}[t]{l|l}
+\begin{tabular}{l|l}
 \hline
@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12cm}}
+\begin{tabular}{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12.2cm}}
 \hline
@@
 \multicolumn{2}{l}{\textbf{Resource Unit (RU)}}\\
+\hline
+\hspace{1em}RU1 - Resource unit mobility & Adult spiny lobsters can move between 1 and 10 Km, while larvae can have displacements in the order of hundreds of Km (Aceves-Bueno et al., 2017; Green et al., 2017).\\
 \hline

### 2019-02-21 docs/SupplementaryMaterial.Rmd
Added all files for a test because it is not knitting

@@
 invert %>%
-  filter(generoespecie %in% c("Panulirus interruptus", "Panulirus argus")) %>% 
+  mutate(generoespecie = ifelse(generoespecie %in% c("Panulirus interruptus", "Panulirus argus"),
+                             generoespecie,
+                             "any_invert")) %>% 
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 
+  summarize(abundancia = sum(abundancia)) %>% 
+  spread(generoespecie, abundancia, fill = 0) %>% 
+  gather(generoespecie, abundancia, -c(rc, ano, year, comunidad, sitio, zona, transecto)) %>% 
+  filter(generoespecie %in% c("Panulirus interruptus","Panulirus argus")) %>% 
   group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 

### 2019-02-21 scripts/data_cleaning_logbook.Rmd
Added all files for a test because it is not knitting

@@
   filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
-  filter(!is.na(rc),
-         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
-                   "Cabezo - Cabezo (Control)",
-                   "Manchon - Manchon (Control)",
-                   "El Faro - El Faro (Control)")) %>% 
-  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
+  filter(!is.na(rc)) %>% 
+  # filter(rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
+  #                  "Cabezo - Cabezo (Control)",
+  #                  "Manchon - Manchon (Control)",
+  #                  "El Faro - El Faro (Control)")) %>% 
+  # filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
   filter(between(abundancia, 0.1, 100)) %>%
@@
                          TRUE ~ ano - 2013)) %>% 
-  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, abundancia) %>% 
-  arrange()
+  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, abundancia)
 ```
@@
   filter(comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
-  filter(!is.na(rc),
-         rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
-                   "Cabezo - Cabezo (Control)",
-                   "Manchon - Manchon (Control)",
-                   "El Faro - El Faro (Control)")) %>% 
-  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
+  filter(!is.na(rc)) %>% 
+  # filter(rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
+  #                  "Cabezo - Cabezo (Control)",
+  #                  "Manchon - Manchon (Control)",
+  #                  "El Faro - El Faro (Control)")) %>% 
+  # filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
   filter(abundancia > 0) %>% 
@@
 fish <- fish %>% 
-  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b) %>% 
-  arrange()
+  select(comunidad, ano, year, rc, sitio, zona, transecto, generoespecie, talla, abundancia, a, b)
 ```

### 2019-02-21 docs/SupplementaryMaterial.Rmd
Updated MS

@@
   count() %>% 
+  ungroup() %>% 
   group_by(comunidad, zona) %>%
   count() %>% 
+  ungroup() %>% 
   spread(zona, nn) %>% 
+  mutate(age = c(10, 4, 4)) %>% 
   knitr::kable(caption = "Number of invertebrate transects performed in each site of each community.",
-               col.names = c("Community", "Control", "Reserve"),
+               col.names = c("Community", "Control", "Reserve", "Years of monitoring"),
                booktabs = T)
@@
   count() %>% 
+  ungroup() %>% 
   group_by(comunidad, zona) %>%
   count() %>% 
+  ungroup() %>% 
   spread(zona, nn) %>% 
-  knitr::kable(caption = "Number of fish transects performed in each site of each community.",
-               col.names = c("Community", "Control", "Reserve"),
+  mutate(age = c(10, 4, 4)) %>% 
+  knitr::kable(caption = "Number of invertebrate transects performed in each site of each community.",
+               col.names = c("Community", "Control", "Reserve", "Years of monitoring"),
                booktabs = T)

### 2019-02-21 scripts/data_cleaning_logbook.Rmd
Updated MS

@@
     toc: yes
-    toc_float: yes
+    # toc_float: yes
 ---

### 2019-02-28 docs/PLoS_manuscript.tex
Migrated from Rmd to script and updated PLOS submission

% Template for PLoS
% Version 3.4 January 2017
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Fig".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

\usepackage{float}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% Add links
\usepackage{hyperref}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Fig #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juan Carlos Villaseñor-Derbez\textsuperscript{1\Yinyang*},
Eréndira Aceves-Bueno\textsuperscript{1,2\Yinyang},
Stuart Fulton\textsuperscript{3\ddag},
Álvin Suarez\textsuperscript{3\ddag},
Arturo Hernández-Velasco\textsuperscript{3\ddag},
Jorge Torre\textsuperscript{3\ddag},
Fiorenza Micheli\textsuperscript{4\ddag}
\\
\bigskip
\textbf{1} Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA
\\
\textbf{2} Nicholas School of the Environment, Duke University, Beaufort, NC, USA
\\
\textbf{3} Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico
\\
\textbf{4} Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
\ddag These authors also contributed equally to this work.

% Current address notes
%\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
%\dag Deceased

% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* juancarlos@ucsb.edu

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage. Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments. In the last 13 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2012; their effectiveness has not been formally evaluated. We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico. We find that reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates. A lack of clear ecological and socioeconomic effects likely results from a combination of factors. First, some of these reserves might be too young for the effects to show. Second, the reserves are not large enough to protect mobile species, like lobster. Third, variable and extreme oceanographic conditions have impacted harvested populations. Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches. However, these reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
%\section*{Author summary}
%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \cite{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \cite{costello_2012}. One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves, thus creating TURF-reserve systems \cite{afflerbach_2014,gelcich_2015,lester_2017}.

TURFs are a fisheries management tool in which a well-defined group of fishers (\emph{e.g.} fishing cooperatives) have exclusive access to an explicitly delimited portion of the ocean. They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \cite{gelcich_2008,costello_2010,mccay_2014}. On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits. These can be implemented to protect biodiversity but also as fishery management tools to aid in the recovery of marine stocks. These instruments can be combined by establishing a marine reserve within a TURF, thus making them TURF-reserves \cite{afflerbach_2014,gelcich_2015,lester_2017}.

Conservation science has shown how marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \cite{lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \cite{gelcich_2008,gelcich_2012}. The benefits resulting from reserves established within TURFs (\emph{i.e.} TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \cite{gelcich_2015}. Although in theory these systems are expected to be successful \cite{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success. Moreover, TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \cite{ostrom_2009,gelcich_2015}. It is therefore important to consider not only the indicators of interest, but also the governance settings under which the reserves operate.

Recent changes in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries. In Mexico, a legal framework created in 2012 allows fishers to request legal recognition of community-based reserves as “Fish Refuges” (\emph{Zona de Refugio Pesquero}, described in more detail below; \cite{nom}). Since 2012, 45 old and new marine reserves have gained legal recognition as Fish Refuges. Of these, 18 were originally implemented within TURFs. However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico. The objective of this work is twofold. First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries. Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness. We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.

\section*{Methods}

\subsection*{TURF-reserves in Mexico}

Community-based marine reserves that are implemented within TURFs are a form of TURF-reserve, voluntarily established and enforced by local communities. This bottom-up approach can increase compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \cite{beger_2004,smallhornwest_2018}. Community-based spatial closures occur in different contexts, like the \emph{kapu} or \emph{ra’ui} areas in the Pacific Islands \cite{johannes_2002,bohnsack_2004}. However, community-based reserves can be hard to enforce if they are not legally recognized. In such conditions, TURF fishers must rely on the exclusive access of the TURF to maintain high levels of compliance.

In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and helped create a legal framework that solves this governance issue: Fish Refuges \cite{nom}. Fish Refuges can be implemented as permanent, temporary or partial reserves, which can protect one, some, or all resources within their boundaries. One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporary marine reserves within their TURFs with a defined expiration date (often five years). When the expiration date is reached, fishers can chose to open the reserves to fishing or re-establish them. Our work focuses on Fish Refuges implemented as community-based TURF-reserves in small-scale fisheries.

The most common setup of community-based TURF-reserves in Mexico is the following. Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig~\ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and each community usually hosts no more than one cooperative. The profits from each TURF are shared amongst all fishers from the cooperative. Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (\emph{i.e.} TURF-reserves). Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges after conducting a series of studies outlined in the regulation, namely the creation of a baseline of ecological data and historical performance of the fishery \cite{nom}.

\subsection*{Study areas}

We evaluate three community-based no-take TURF-reserve systems implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig~\ref{fig:map}A). The first one was created by the \emph{Buzos y Pescadores de la Baja California} fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig~\ref{fig:map}B). The main fishery in the island is the spiny lobster (\emph{Panulirus interruptus}), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income. In 2006, the community decided to implement two marine reserves within their fishing grounds. The objective of these reserves was "to protect and recover stocks of commercially important invertebrate species"; mainly lobster and abalone. The reserves obtained legal recognition in 2018 \cite{dof_website_2018}.

\begin{figure}[!h]
\centering
%\includegraphics{img/fig_map.pdf}
\caption{Location of the three coastal communities studied (A). Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena and Punta Herrero (C) are located in the Yucatan Peninsula. Blue polygons represent the TURFs, and red polygons the marine reserves.}
\label{fig:map}
\end{figure}

The other two TURF-reserve systems are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig~\ref{fig:map}C). In contrast with Isla Nativdad, which hosts a well-established fishing community, Maria Elena is a fishing camp visited intermittently during the fishing season that belongs to the \emph{Cozumel} fishing cooperative. Punta Herrero is home to the \emph{José María Azcorra} fishing cooperative, and similar to Isla Natividad hosts a small community. Their main fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but they also target finfish in the off-season. Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively. These reserves have been legally recognized as Fishing Refuges since their original implementation \cite{dof_website_2012,dof_website_2013} and subsequent re-establishments \cite{dof_website_2017b}.

These communities are representative of their region in terms of ecology, socioeconomic, and governance aspects. Isla Natividad, for example, is part of a greater group of fishing cooperatives belonging to a Federation of Fishing Cooperatives. This group has been identified as a cohesive group that cooperates to better manage their resources \cite{mccay_2014,mccay_2017,acevesbueno_2017}. Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation. Together, these three communities provide an accurate representation of other fishing communities that have been historically manged with TURFs in each of their regions. While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth causal inference analysis that we undertake. Yet, given the similarities among communities and the socioeconomic and governance setting under which they operate, it is safe to cautiously generalize our insights to other similar community-based TURF-reserves in Mexico and elsewhere.

The regulation governing the implementation of Fish Refuges states that these are fishery management tools intended to have conservation and fisheries benefits \cite{nom}. For this reason, the main portion of our analyses focuses on a series of biological and socioeconomic indicators that may respond to reserve implementation. However, the effectiveness of conservation and fisheries management interventions also depends on the social and governance structures in place. We therefore incorporate a reduced version of the Social Ecological Systems framework \cite{ostrom_2009} and evaluate variables and indicators known to aid and hinder the effectiveness of management interventions in conservation and fisheries. The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather help provide context on the social-ecological system in which reserves develop. The following two sections describe our data collection methods and analyses.

\subsection*{Data collection}

We use three main sources of information to evaluate these reserves across ecological, socioeconomic, and governance dimensions. Ecological data come from the annual ecological monitoring of reserve and control sites. Reserve sites are areas where no fishing occurs. Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) are not directly adjacent to the reserves. We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve. This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \cite{stewartoaten_1986,depalma_2018} and causally attribute the changes to the reserve \cite{francinifilho_2008,Villasenor-Derbez_2018}.

The biological data are collected by members from each community and personnel from the Mexican CSO \emph{Comunidad y Biodiversidad} (\href{https://www.cobi.org.mx}{COBI}). Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites, where a minimum of 4 transects per site are performed \cite{suman_2010-ez,fulton_2018,fulton_2019}. Size structures are also collected during fish surveys. All sites were surveyed annually, and at least once before implementation of the reserves. A summary of sampling effort is shown in the supplementary materials (\nameref{S1_Table}, \nameref{S2_Table}).

Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves. In this case our treated unit are the cooperatives (\emph{i.e.} communities) that have implemented a reserve within their TURF, and the controls are nearby communities that have a TURF but did not implement a reserve. These communities have TURFs adjacent to the TURFs of our focal communities. Cooperatives incorporated in this analysis have similar number of members, belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \cite{mccay_2014,mccay_2017,ayer_2018}. Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \cite{oecd_2017}. A table with summary statistics for this data is provided in the supplementary materials (Table S3, Fig S5).

Data for the evaluation of the SES were collected at the community-level from official documents used in the design, creation, and designation of the marine reserves. These include the technical studies that the cooperatives submit when they request recognition of their reserves, as well as the official enactments  \cite{dof_website_2012,dof_website_2013,dof_website_2018}. We also complimented information based on the authors' experience and knowledge of the communities. We collected information on the Resource Systems, Resource Units, Actors, and Governance System (Table~\ref{table:ses}).

\input{tab/ses.tex}

\subsection*{Data analysis}

We evaluate the effect that the TURF-reserves have had on four ecological and two socioeconomic indicators shown in Table~\ref{table:indicators}. Recall that reserves were implemented to protect lobster and other benthic invertebrates. However, we also use the available fish and invertebrate data to test for associated co-benefits.

\input{tab/indicators.tex}

We use a difference-in-differences analysis to evaluate these indicators. This approach is widely used in econometric literature to estimate the average treatment effect of an intervention, like the impact of minimum wage increases on employment rates \cite{card_1994}. In our case it allows us to estimate the effect that the reserve had on each biological and socioeconomic indicator (Table~\ref{table:indicators}) by comparing trends across time and treatments since reserve implementation \cite{moland_2013,Villasenor-Derbez_2018}. To perform difference-in-differences, we regress the indicator of interest on a dummy variable for treatment, a dummy variable for years, and the interaction term between these with a multiple linear regression of the form:

\begin{equation}
I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t}
\label{eqn:reg_bio}
\end{equation}

Where year-level fixed effects capturing a temporal trend are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites. The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (\emph{i.e.} $\gamma_t$ and $\beta$ respectively). Therefore, we would expect this term to be positive if the indicator increases because of the reserve. Finally, $\epsilon_{i,t}$ represents the error term of the regression.

Socioeconomic indicators are evaluated with a similar approach. Due to data constrains, we only evaluate socioeconomic data for Isla Natividad (2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are used as counterfactuals that allow us to control for unobserved time-invariants. Each focal community (\emph{i.e.} Isla Natividad and Maria Elena) has three counterfactual communities.

\begin{equation}
I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t}
\label{eqn:soc_reg}
\end{equation}

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$). These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \cite{depalma_2018}. For each indicator in each community, we fit one model (\emph{e.g.} there are three models for lobster density, one for each community) for a total of 12 biological model fits and four socioeconomic model fits. Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard errors \cite{zeileis_2004-7n}. All analyses were performed in R 3.5.2 and R Studio version 1.1.456 \cite{R_2018}. All data and code needed to reproduce our analyses are available in a GitHub repository at: \url{https://github.com/jcvdav/ReserveEffect}.

We use the SES framework to evaluate each community and create a narrative that provides context for each community. The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators. Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as in \cite{leslie_2015-na}), we followed a similar approach to \cite{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems. We based our variable selection primarily on \cite{leslie_2015-na} and \cite{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework, and identified the key variables relevant to fishing cooperatives in Mexico. We also incorporate other relevant variables known to influence reserve performance following \cite{difranco_2016-Xw} and \cite{edgar_2014-UO}. Table~\ref{table:ses} shows the selected variables, along with definitions and values.

\section*{Results}

The following sections present the effect that marine reserves had on the biological and socioeconomic indicators for each coastal community. Results are presented in terms of difference through time and across sites, relative to the control site on the year of implementation (\emph{i.e.} the difference-in-differences estimate or effect size $\lambda_t$ from Eqs. \ref{eqn:reg_bio} and \ref{eqn:soc_reg}). We also provide an overview of the governance settings of each community, and discuss how these might be related to the effectiveness and performance of the reserves.

\subsection*{Biological effects}

Indicators showed ambiguous responses through time for each reserve. Fig~\ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time. In the case of Maria Elena, positive changes were observed in the third and fourth year. These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ for Isla Natividad and Punta Herrero, and 0.01 organisms $\mathrm{m}^{-2}$ for Maria Elena, but are not significantly different from zero ($p > 0.05$). Likewise, no significant changes were detected in fish biomass or invertebrate and fish densities (Fig. \ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends. Figures and tables with time series of indicators and model coefficients are presented in the supplementary materials (\nameref{S1_Fig}, \nameref{S2_Fig}, \nameref{S3_Fig}, \nameref{S4_Fig}, \nameref{S4_Table}, \nameref{S5_Table}, \nameref{S6_Table}).

\begin{figure}[!h]
\centering
%\includegraphics{img/fig_bio_results.pdf}
\caption{{\bf Effect sizes for biological indicators}
Points indicate the effect size and error bars are heteroskedastic-robust standard errors. Years have been centered to year of implementation. Colors and shapes denote communities: Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares). Letters denote indicator: A) lobster densities (\emph{Panulirus spp}, B) fish biomass, C) invertebrate densities, and D) fish densities. Plots are ordered by survey type (left column: invertebrates; right column: fish). Points are jittered hotizontally to avoid overplotting.}
\label{fig:indicators}
\end{figure}

\subsection*{Socioeconomic effects}

Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig~\ref{fig:lobsters}). For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, suggesting that these are plausible controls. However, effect sizes do not change after the implementation of the reserve. Interestingly, the negative effect observed for Isla Natividad on year 5 corresponds to the 2011 hypoxia events \cite{micheli_2012-EU}. The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$). The year of post-implementation data for Maria Elena does not show a significant effect of the reserve. Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities. However, these changes are only significant for the third year ($p < 0.05$). Full tables with model coefficients are presented in the supplementary materials (\nameref{S7_Table}, \nameref{S8_Table}).

\begin{figure}[!h]
\centering
%\includegraphics{img/fig_socioeco_results.pdf}
\caption{{\bf Effect sizes for socioeconomic indicators}
Points indicate the effect size and error bars are heteroskedastic-robust standard errors for: lobster catches (A) and revenues (B) at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles). Years have been centered to year of implementation. Points are jittered hotizontally to avoid overplotting.}
\label{fig:lobsters}
\end{figure}

\subsection*{Governance}

Our qualitative implementation of the social-ecological systems framework allowed us to systematically identify important differences between the case studies’ governance systems and incorporate other characteristics of these fisheries neglected during the process of data collection (Table~\ref{table:ses}). We find that the analyzed communities share similarities known to foster sustainable resource management and increase reserve effectiveness. For example, fishers operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves. Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have solid governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management. In general, success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources upon which they depend \cite{jupiter_2017}. Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperatives, which have favored the development of efficient community-based enforcement systems. However, our SES analysis also highlights factors that might hinder reserve performance or mask outcomes. While total reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual reserves are often small (RS3); the largest reserve is only 4.37 km\textsuperscript{2}, and the smallest one is 0.09 km\textsuperscript{2}. Reserves are also relatively young (RS5). Additionally, fishers harvest healthy stocks (RS4.1), and it is unlikely that marine reserves will result in increased catches.

\section*{Discussion}

Our results indicate that these TURF-reserves have not increased lobster densities. Additionally, no co-benefits were identified when using other ecological indicators aside from the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \cite{micheli_2012-EU}. The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation. Lastly, the communities exhibit all the social enabling conditions for effective reserve and resource management. Here we discuss possible shortcomings in our analyses as well as possible explanations for the observed patterns.

While many ecology studies have used BACI sampling designs and respective analyses (\emph{e.g.} \cite{stewartoaten_1986}), few conservation studies have done so to evaluate the effect of an intervention (\emph{e.g.} \cite{francinifilho_2008,lester_2009,moland_2013}) which has resulted in a call for more robust analyses in conservation science \cite{guidetti_2002,ferraro_2006}. Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness, and captures previously described patterns. For example, the rapid increase observed for lobster densities in Isla Natividad on the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), occurs a year after the hypoxia events described by \cite{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish. The use of causal inference techniques may help us support evidence-based conservation.

Our analyses of socioeconomic indicators has three limitations. First, we only look at landings and revenues by landings for communities with and without TURF-reserves. There are a number of other possible indicators that could show a change due to the implementation of the reserve. Notably, one often cited in the literature is additional benefits, such as tourism \cite{viana_2017}. However, it is unlikely that the evaluated communities will experience tourism benefits due to their remoteness and the lack of proper infrastructure to sustain tourism. A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve. For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and thus higher profitability, provided that cost per unit effort does not increase. Likewise, it is possible that fishing effort increased around reserves to maintain the historical levels of landings. A final limitation applies to Maria Elena, where we only observe landings and income for one year after reserve implementation. While one would not expect to observe increased landings or income in such a short period, a spatial closure might cause total catches to decline, especially if effort is held constant.

A first possible explanation for the lack of effectiveness may be the young age of the reserves. Literature shows that age and enforcement are important factors that influence reserve effectiveness \cite{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest reserves, and our SES analysis suggests that all communities have a well-established community-based enforcement system. With these characteristics, one would expect the reserves to be effective. Maria Elena and Punta Herrero are relatively young reserves (\emph{i.e.} < 6 years old; RS5 in Table~\ref{table:ses}) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \cite{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size \cite{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (\emph{sensu} \cite{rees_2018}). Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \cite{rossetto_2015-V0}, however, abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary invertebrates, but not lobsters. Design principles developed by \cite{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across. As shown through the SES analysis, the size of the marine reserves appears small compared to the movement capacity of the main targeted species (RU1, RS3; Table~\ref{table:ses}). Furthermore, fishers may favor implementation of reserves that pose low fishing costs due to their small size or location. Our analysis of economic data supports this hypothesis, as neither landings nor revenues showed the expected short-term reductions associated to the first years of reserve implementation \cite{ovando_2016-Wg}.

Even if reserves had appropriate sizes and were placed in optimal locations, there are other plausible explanations for the observed patterns. For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \cite{hilborn_2004,hilborn_2006}. Both lobster fisheries were certified by the Marine Stewardship Council and are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \cite{dof_website_1993}. It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves; reserves implemented in TURFs where fishing pressure is already optimally managed will still show a trade-off between fisheries and conservation objectives \cite{lester_2017}. Furthermore, \cite{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open access. This might reduce the difference between indicators from the TURF and reserve sites, making it difficult to detect such a small change. Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of the reserves might be greater.

Finally, extreme conditions, including prolonged hypoxia, heat waves, and storms have affected both the Pacific and Caribbean regions, with large negative impacts on coastal marine species and ecosystems \cite{cavole_2016,hughes_2018,breitburg_2018}. The coastal ecosystems where these reserves are located have been profoundly affected by these events \cite{micheli_2012-EU,woodson_2018}. Effects of protection might be eliminated by the mortalities associated with these extreme conditions.

\section*{Conclusion}

While the evaluated reserves have failed to provide fishery benefits to date, there are a number of additional ecological, fisheries, and social benefits. Marine reserves provide protection to a wider range of species and vulnerable habitat. Previous research focusing on these specific sites has shown that they serve as an insurance mechanism against uncertainty and errors in fisheries management, as well as mild environmental shocks \cite{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}. Self-regulation of fishing effort can serve as a way to compensate for future declines associated to environmental variation \cite{finkbeiner_2018}. Furthermore, embarking on a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \cite{fulton_2019}. Showing commitment to marine conservation and sustainable fishing practices has allowed fishers to have greater bargaining power and leverage over fisheries management \cite{prezramrez_2012}. These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-establishing the reserves \cite{ayer_2018}.

Community-based TURF-reserves in small-scale fisheries may be helpful conservation and fishery management tools when appropriately implemented \cite{gelcich_2015}. We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside. Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness. Furthermore, conservation and advocacy groups should consider the opportunity costs of such interventions (\emph{sensu} \cite{smith_2010}) and evaluate the potential of other approaches that may yield similar benefits.

In terms of fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs. Future research should aim at evaluating other Fish Refuges established as bottom-up processes but without the presence of TURFs (\emph{e.g.} \cite{dof_websiteC_2012}), others established through top-down processes (\emph{i.e.} Ref. \cite{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches. For the particular case of the reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.

\section*{Supporting information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Table.}
\label{S1_Table}
{\bf Invertebrate sampling effort.} Number of invertebrate transects performed in each site of each community.

\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf Mean annual lobster density.} Dots indicate mean, bars indicate standard errors.

\paragraph*{S2 Fig.}
\label{S2_Fig}
{\bf Mean annual invertebrate density.} Dots indicate mean, bars indicate standard errors.

\paragraph*{S2 Table.}
\label{S2_Table}
{\bf Fish sampling effort} Number of invertebrate transects performed in each site of each community.

\paragraph*{S3 Fig.}
\label{S3_Fig}
{\bf Mean annual fish biomass} Dots indicate mean, bars indicate standard errors.

\paragraph*{S4 Fig.}
\label{S4_Fig}
{\bf Mean annual fish density} Dots indicate mean, bars indicate standard errors.

\paragraph*{S3 Table.}
\label{S3_Table}
{\bf Summary of socioeconomic data} Mean ex-vessel prices (MXP / Kg) for lobster (\emph{Panulirus spp.}) on each TURF (community) and group (Control = only TURF, Treated = TURF and reserve).

\paragraph*{S5 Fig.}
\label{S5_Fig}
{\bf Timeseries of socioeconomic indicators.} A) Mean annual value of landings and B) Mean landings.

\paragraph*{S4 Table.}
\label{S4_Table}
{\bf Coefficient estimates of biological indicators for Isla Natividad.}

\paragraph*{S5 Table.}
\label{S5_Table}
{\bf Coefficient estimates of biological indicators for Maria Elena.}

\paragraph*{S6 Table.}
\label{S6_Table}
{\bf Coefficient estimates of biological indicators for Punta Herrero.}

\paragraph*{S7 Table.}
\label{S7_Table}
{\bf Coefficient estimates of socioeconomic indicators Isla Natividad.}

\paragraph*{S8 Table.}
\label{S8_Table}
{\bf Coefficient estimates of socioeconomic indicators Maria Elena.}

\section*{Acknowledgments}

The authors wish to acknowledge Imelda Amador for contributions on the governance data, as well as pre-processing biological data. This study would have not been possible without the effort by members of the fishing communities here mentioned, who participated in the data-collection process. The authors wish to acknowledge comments by the reviewers and editor, which significantly improved the quality of this work.

%
%% For figure citations, please use "Fig" instead of "Figure".
%Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.
%
%% Place figure captions after the first paragraph in which they are cited.
%\begin{figure}[!h]
%\caption{{\bf Bold the figure title.}
%Fig caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
%\label{fig1}
%\end{figure}
%
%% Results and Discussion can be combined.
%\section*{Results}
%Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.
%
%% Place tables after the first paragraph in which they are cited.
%\begin{table}[!ht]
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
%\centering
%\caption{
%{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
%\begin{tabular}{|l+l|l|l|l|l|l|l|}
%\hline
%\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
%$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
%$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
%$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
%\end{tabular}
%\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
%\end{flushleft}
%\label{table1}
%\end{adjustwidth}
%\end{table}
%
%
%%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
%\subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
%\subsubsection*{3rd level heading} 
%Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit. Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat. 
%
%\begin{enumerate}
%	\item{react}
%	\item{diffuse free particles}
%	\item{increment time by dt and go to 1}
%\end{enumerate}
%
%
%\subsection*{Sed ac quam id nisi malesuada congue}
%
%Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.
%
%\begin{itemize}
%	\item First bulleted item.
%	\item Second bulleted item.
%	\item Third bulleted item.
%\end{itemize}
%
%\section*{Discussion}
%Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero~\cite{bib3}.
%
%\section*{Conclusion}
%
%CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit. 
%
%Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat. Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna. Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed. Nam condimentum sem eget mollis euismod. Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex. Nunc est dui, sodales sed mauris nec, auctor sagittis leo. Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Appendix}.
%

\nolinenumbers

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for 
% step-by-step instructions.
%

\bibliography{references}

\end{document}

### 2019-02-28 docs/tab/S3_table.tex
Migrated from Rmd to script and updated PLOS submission

\begin{table}[t]

\caption{\label{tab:}Mean ex-vessel prices (MXP / Kg) for lobster (Panulirus spp.) on each TURF (community) and group (Control = only TURF, Treated = TURF and reserve).}
\centering
\begin{tabular}{lllrr}
\toprule
Community & Group & TURF & Before & After\\
\midrule
Yucatan Peninsula & Control & Langosteros Del Caribe & 165.58 & 176.72\\
Yucatan Peninsula & Control & Vigia Chico & 177.66 & 175.31\\
Yucatan Peninsula & Treated & Cozumel & 178.21 & 170.49\\
Baja California & Control & Bahia Tortugas & 155.60 & 122.46\\
Baja California & Control & La Purisima & 67.54 & 157.88\\
\addlinespace
Baja California & Control & Pesc Nacionales De Abulon & 67.16 & 88.79\\
Baja California & Treated & Buzos Y Pescadores De La Baja California & 89.35 & 190.07\\
\bottomrule
\end{tabular}
\end{table}

### 2019-02-28 docs/tab/indicators.tex
Migrated from Rmd to script and updated PLOS submission

--- 2018-12-31 docs/manuscript.tex
@@
-\documentclass{frontiersSCNS}
-\usepackage{url,hyperref,lineno,microtype,subcaption}
-\usepackage[onehalfspacing]{setspace}
-\usepackage{float}
+\begin{table}[!ht]
 
-\linenumbers
-
-\usepackage[utf8]{inputenc}
-\floatplacement{figure}{H}
-
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
-\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
-Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
-Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
-Micheli\(^{4}\)}
-% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
-% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
-of California, Santa Barbara, Santa Barbara, CA, USA \newline
-\(^{2}\)Nicholas School of the Environment, Duke University, Beaufort,
-NC, USA \newline \(^{3}\)Comunidad y Biodiversidad A.C., Guaymas,
-Sonora, Mexico \newline \(^{4}\)Hopkins Marine Station and Center for
-Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
-% The Corresponding Author should be marked with an asterisk
-% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
-Santa Barbara, Santa Barbara, CA, 93106}
-
-\def\corrEmail{\href{mailto:juancarlos@ucsb.edu}{\nolinkurl{juancarlos@ucsb.edu}}}
-
-\begin{document}
-\onecolumn
-\firstpage{1}
-
-\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in Mexican small-scale
-fisheries}
-
-\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-\extraAuth{}
-
-\maketitle
-
-\begin{abstract}
-
-Coastal marine ecosystems provide livelihoods for small-scale fishers
-and coastal communities around the world. Small-scale fisheries face
-great challenges since they are difficult to monitor, enforce, and
-manage. Combining territorial use rights for fisheries (TURF) with
-no-take marine reserves to create TURF-reserves can improve the
-performance of small-scale fisheries by buffering fisheries from
-environmental variability and management errors, while ensuring that
-fishers reap the benefits of conservation investments. In the last 12
-years, 18 old and new community-based Mexican TURF-reserves gained legal
-recognition thanks to a regulation passed in 2014; their effectiveness
-has not been formally evaluated. We combine causal inference techniques
-and the Social-Ecological Systems framework to provide a holistic
-evaluation of community-based TURF-reserves in three coastal communities
-in Mexico. We find that while reserves have not yet achieved their
-stated goal of increasing the density of lobster and other benthic
-invertebrates, they continue to receive support from the fishing
-communities. A lack of clear ecological and socioeconomic effects likely
-results from a combination of factors. First, some of these reserves
-might be too young for the effects to show. Second, the reserves are not
-large enough to protect mobile species, like lobster. Third, variable
-and extreme oceanographic conditions have impacted harvested
-populations. Fourth, local fisheries are already well managed, and it is
-unlikely that reserves might have a detectable effect in catches.
-However, these reserves may provide a foundation for establishing
-additional, larger marine reserves needed to effectively conserve mobile
-species.
-
-\medskip
-\tiny
- \keyFont{ \section{Keywords:} TURF-reserves, Causal Inference, Social-Ecological Systems, Marine
-Protected Areas, Marine Conservation, Small-Scale Fisheries}
-
-\end{abstract}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Marine ecosystems around the world sustain significant impacts due to
-overfishing and unsustainable fishing practices
-\citep{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular,
-small-scale fisheries face great challenges since they tend to be hard
-to monitor and enforce \citep{costello_2012}. One of the many approaches
-taken to improve the performance of coastal fisheries and health of the
-local resources is through the implementation of Territorial Use Rights
-for Fisheries (TURFs) that contain no-take marine reserves within them,
-thus creating TURF-reserve systems
-\citep{afflerbach_2014,gelcich_2015,lester_2017}.
-
-TURFs are a fisheries management tool in which a well defined group of
-fishers (\emph{e.g.} fishing cooperatives) have exclusive access to an
-explicitly delimited portion of the ocean. They promote a sense of
-stewardship and incentivise resource users to sustainably manage their
-resources \citep{gelcich_2008,costello_2010,mccay_2014}. On the other
-hand, no-take marine reserves (marine reserves from hereinafter) are
-areas where all extractive activities are off-limits. These can be
-implemented to protect biodiversity but also as fishery management tools
-to aid in the recovery of marine stocks. These instruments can be
-combined by establishing a marine reserve within a TURF, thus making
-them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.
-
-Conservation science has shown how marine reserves lead to increased
-biomass, species richness, and abundance within the protected regions
-\citep{lester_2009}, and that these may have a series of additional
-benefits like climate change mitigation, protection from environmental
-variability, and fisheries benefits
-\citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
-research on TURFs has shown that these areas have higher abundance of
-targeted species than sites operating under open access and even similar
-to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The
-benefits resulting from reserves established within TURFs (\emph{i.e.}
-TURF-reserves) should be captured exclusively by the group of fishers
-with exclusive access \citep{gelcich_2015}. Although in theory these
-systems are successful \citep{smallhornwest_2018}, there is little
-empirical evidence of their effectiveness and the drivers of their
-success. Moreover, TURF-reserve systems are inherently intricate
-social-ecological systems, and their effectiveness must depend on how
-environmental and social factors combine and interact
-\citep{ostrom_2009,gelcich_2015}. It is therefore important to consider
-not only the indicators of interest, bu also the governance settings
-under which the reserves occur.
-
-Recent norms in fisheries regulation in Mexico provide a ripe
-opportunity to study the effectiveness of community-based TURF-reserves
-in small-scale fisheries. In Mexico, a norm created in 2014 allows
-fishers to request legal recognition of community-based reserves as
-``Fish Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since
-2012, old and new marine reserves have gained legal recognition as Fish
-Refuges. Of these, 18 were originally implemented within TURFs. However,
-their effectiveness has not yet been formally evaluated and reported in
-the scientific literature.
-
-Here, we combine causal inference techniques and the Social-Ecological
-Systems (SES) framework to evaluate community-based TURF-reserves in
-three coastal communities in Mexico. The objective of this work is
-twofold. First, to provide a holistic evaluation of the effectiveness of
-community-based TURF-reserves in terms of the changes in biological and
-socioeconomic indicators and the governance settings under which these
-develop, which may inform similar processes in other countries. Second,
-to identify opportunities where improvement or adjustment might lead to
-increased effectiveness. We draw from lessons learned in these three
-case studies and provide management recommendations to maximize the
-effectiveness of community-based TURF-reserves in small-scale fisheries
-where this tool is used to manage and rebuild their coastal fisheries.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{turf-reserves-in-mexico}{%
-\subsection{TURF-reserves in Mexico}\label{turf-reserves-in-mexico}}
-
-Before discussing our data collection methods and describing our
-analyses, our case studies warrant some more background. Community-based
-marine reserves that are implemented within TURFs are a form of
-TURF-reserves, voluntarily established and enforced by local
-communities. This bottom-up approach increases compliance and
-self-enforcement, and reserves can yield benefits similar to
-systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
-Community-based spatial closures occur in different contexts, like the
-\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
-\citep{bohnsack_2004,johannes_2002}. However, community-based reserves
-can be hard to enforce if they are not legally recognized. In such
-conditions, TURF fishers must rely on the exclusive access of the TURF
-to maintain high levels of compliance.
-
-In an effort to bridge this normative gap, Mexican Civil Society
-Organizations (CSOs) served as a link between fishers and government,
-and created a legal framework that solves this governance issue: Fish
-Refuges \citep{nom}. Fish Refuges can be implemented as temporal or
-partial reserves, which can protect one, some, or all resources within
-their boundaries. One of the ways in which fishing communities have
-taken advantage of this new tool is by implementing temporal marine
-reserves within their TURFs with a defined expiration date (often 5
-years). When the expiration date is reached, fishers can chose to open
-the reserves to fishing or re-establish them. Our work focuses on Fish
-Refuges implemented as community-based TURF-reserves that occur in
-small-scale fisheries.
-
-The common setup of community-based TURF-reserves in Mexico is the
-following. Fishers from a given community are assembled in fishing
-cooperatives which have exclusive fishing rights over a spatially
-delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
-\ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and
-each community usually hosts no more than one cooperative. The profits
-from each TURF are shared amongst all fishers from the cooperative.
-Fishing cooperatives interested in implementing marine reserves work
-with CSOs to implement marine reserves within their TURFs (\emph{i.e.}
-TURF-reserves). Fishers then ask the government to grant legal
-recognition to their TURF-reserves as Fish Refuges following a series of
-studies outlined in the regulation \citep{nom}.
-
-\hypertarget{study-areas}{%
-\subsection{Study areas}\label{study-areas}}
-
-We evaluate three community-based no-take TURF-reserves implemented in
-Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig
-\ref{fig:map}A). The first one was created by the \emph{Buzos y
-Pescadores de la Baja California} fishing cooperative, located in Isla
-Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The
-main fishery in the island is the spiny lobster (\emph{Panulirus
-interruptus}), but other resources like finfish, sea cucumber, read sea
-urchin, snail, and abalone are also an important source of income. In
-2006, the community decided to implement two marine reserves within
-their fishing grounds. The objective of these reserves was ``to protect
-commercially important invertebrate species''; mainly lobster and
-abalone. These reserves obtained legal recognition in 2018
-\citep{dof_website_2018}.
-
-The other two TURF-reserves are located in Maria Elena and Punta
-Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
-Isla Nativdad, which hosts a well established fishing community, Maria
-Elena is a fishing camp --visited intermittently during the fishing
-season-- belonging to the \emph{Cozumel} fishing cooperative; Punta
-Herrero is home to the \emph{José María Azcorra} fishing cooperative,
-and similar to Isla Natividad hosts a local community. Their main
-fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but
-they also target finfish in the off-season. Maria Elena and Punta
-Herrero established eight and four marine reserves in 2012 and 2013,
-respectively. These reserves have been legally recognized as Fishing
-Refuges since their original implementation
-\citep{dof_website_2012,dof_website_2013} and subsequent
-re-establishments \citep{dof_website_2017b}.
-
-These communities are representative of their region in terms of
-ecology, socioeconomic, and governance aspects. Isla Natividad, for
-example, is part of a greater group of fishing cooperatives belonging to
-a Federation of Fishing Cooperatives. This group has been identified as
-a cohesive group that cooperates to better manage their resources
-\citep{mccay_2014,mccay_2017,acevesbueno_2017}. Likewise, Maria Elena
-and Punta Herrero are representative of fishing cooperatives in the
-Mexican Caribbean, which are also part of a regional Federation.
-Together, these three communities provide an accurate representation of
-other fishing communities that have been historically manged with TURFs
-in each of their regions. While each region has additional communities
-that have established community-based TURF-reserves, available data
-would not allow us to perform the in-depth causal inference analysis
-that we undertake. Yet, given the similarities among communities and the
-socioeconomic and governance setting under which they operate, it is
-safe to cautiously generalize our insights to other similar
-community-based TURF-reserves in Mexico and elsewhere.
-
-The regulation governing the implementation of Fish Refuges states that
-these are fishery management tools intended to have biological or
-socioeconomic benefits \citep{nom}. For this reason, the main portion of
-our analyses focuses on a series of biological and socioeconomic
-indicators that may respond to reserve implementation. However, the
-effectiveness of conservation and fisheries management interventions
-also depends on the social and governance structures in place. We
-therefore incorporate a reduced version of the Social Ecological Systems
-framework \citep{ostrom_2009} and evaluate variables and indicators
-known to aid and hinder the effectiveness of management interventions in
-conservation and fisheries. The incorporation of the SES is not intended
-to relate different levels of governance with reserve effectiveness, but
-rather help provide context on the social-ecological system in which
-reserves develop. The following two sections describe our data
-collection methods and analyses.
-
-\hypertarget{data-collection}{%
-\subsection{Data collection}\label{data-collection}}
-
-We use three main sources of information to evaluate these reserves
-across the ecological, socioeconomic, and governance dimensions.
-Ecological data come from the annual ecological monitoring of reserve
-and control sites. Reserve sites are areas where no fishing occurs.
-Control sites are areas that meet the following criteria: i) habitat
-characteristics are similar to the corresponding reserves, ii)
-presumably had a similar probability of being selected as reserves
-during the design phase, iii) are located within the TURF, where fishing
-occurs, and iv) Are not directly adjacent to the reserves. We focus our
-evaluation on sites where data are available for reserve and control
-sites, before and after the implementation of the reserve. This provides
-us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design
-that allows us to capture and control for temporal and spatial dynamics
-\citep{stewartoaten_1986,depalma_2018} and causally attribute the
-changes to the reserve
-\citep{francinifilho_2008,Villasenor-Derbez_2018}.
-
-The biological data are collected by members from each community and
-personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
-(\href{www.cobi.org.mx}{COBI}). Trained divers record species richness
-and abundances of fish and invertebrate species along replicate
-transects (30 \(\times\) 2 m each) at depths 5-20 m in the reserves and
-control sites \citep{suman_2010-ez,fulton_2018,fulton_2019}. Size
-structures are also collected during fish surveys. All sites were
-surveyed annually, and at least once before implementation of the
-reserves. A summary of sampling effort is shown in the supplementary
-materials (Tables S1-S2).
-
-Socioeconomic data come from landing receipts reported to the National
-Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
-Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
-(Kg) and revenues (MXP) for TURF-managed cooperatives with and without
-marine reserves. In this case our treated unit are the cooperatives
-(\emph{i.e.} communities) that have implemented a reserve within their
-TURF, and the controls are nearby communities that have a TURF but did
-not implement a reserve. Cooperatives incorporated in this analysis
-belong to larger regional-level Cooperative Federations, and are exposed
-to the same markets and institutional frameworks, making them plausible
-controls \citep{mccay_2017,ayer_2018}. Landings and revenues were
-aggregated at the cooperative-year level, and revenues were adjusted to
-represent 2014 values by the Consumer Price Index for Mexico
-\citep{oecd_2017}. A table with summary statistics for this data is
-provided in the supplementary materials (Table S3, Figure S5).
-
-Data for the evaluation of the SES were collected at the community-level
-from official documents used in the design, creation, and designation of
-the marine reserves. These include the technical studies that the
-cooperatives submit when they request recognition of their reserves, as
-well as the official enactments
-\citep{dof_website_2012,dof_website_2013,dof_website_2018}. We also
-complimented information based on the authors' experience and knowledge
-of the communities. We collected information on the Resource Systems,
-Resource Units, Actors, and Governance System (Table \ref{table:ses}).
-The next section further describe the variables and indicators chosen
-for each of these.
-
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
-
-We evaluate the effect that the TURF-reserves have had on four
-ecological and two socioeconomic indicators shown in Table
-\ref{table:indicators}. Recall that reserves were implemented to protect
-lobster and other benthic invertebrates. However, we also use the
-available fish and invertebrate data to test for associated co-benefits.
-
-We use a difference-in-differences analysis to evaluate these
-indicators. This approach is widely used in econometric literature to
-estimate the average treatment effect of an intervention, like the
-impact of minimum wage increases on employment rates \citep{card_1994}.
-In our case it allows us to estimate the effect that the reserve had on
-each biological and socioeconomic indicator (Table
-\ref{table:indicators}) by comparing trends across time and treatments
-\citep{moland_2013,Villasenor-Derbez_2018}. To perform
-difference-in-differences, we regress the indicator of interest on a
-dummy variable for treatment, a dummy variable for years, and the
-interaction term between these with a multiple linear regression of the
-form:
-
-\begin{equation}
-I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t}
-\label{eqn:reg_bio}
-\end{equation}
-
-Where year-level fixed effects capturing a temporal trend are
-represented by \(\gamma_t Year_t\), and \(\beta Zone_i\) captures the
-difference between reserve (\(Zone = 1\)) and control (\(Zone = 0\))
-sites. The effect of the reserve is captured by the \(\lambda_t\)
-coefficient, and represents the difference observed between the control
-site before the implementation of the reserve and the treated sites at
-time \(t\) after controlling for other time and space variations
-(\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Therefore, we
-would expect this term to be positive if the indicator increases because
-of the reserve. Finally, \(\epsilon_{i,t}\) represents the error term of
-the regression.
-
-Socioeconomic indicators are evaluated with a similar approach. Due to
-data constrains, we only evaluate socioeconomic data for Isla Natividad
-(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
-used as counterfactuals that allow us to control for unobserved
-time-invariants. Each focal community (\emph{i.e.} Isla Natividad and
-Maria Elena) has three counterfactual communities.
-
-\begin{equation}
-I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t}
-\label{eqn:soc_reg}
-\end{equation}
-
-The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio},
-but in this case the \(Treated\) dummy variable indicates if the
-community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)).
-These regression models allow us to establish a causal link between the
-implementation of marine reserves and the observed trends by accounting
-for temporal and site-specific dynamics \citep{depalma_2018}. We fit one
-model per community and indicators (\emph{e.g.} there are three models
-for lobster density, one for each community) for a total of 12
-biological model fits and four socioeconomic model fits. Model
-coefficients were estimated via ordinary least-squares and used
-heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
-analyses were performed in R version 3.5.1 and R Studio version 1.1.456
-\citep{R_2018}. All data and code needed to reproduce our analyses are
-available in a GitHub repository at:
-\url{https://github.com/jcvdav/ReserveEffect}.
-
-We use the SES framework to evaluate each community and create a
-narrative that provides context for each community. The use of this
-framework standardizes our analysis and allows us to communicate our
-results in a common language across fields by using a set of previously
-defined variables and indicators. We based our variable selection
-primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who
-operationalized and analyzed Mexican fishing cooperatives using this
-framework, and identified the key variables relevant to fishing
-cooperatives in Mexico. We also incorporate other relevant variables
-known to influence reserve performance following
-\citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table
-\ref{table:ses} shows the selected variables, along with definitions and
-values.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-The following sections present the effect that marine reserves had on
-each of the biological and socioeconomic indicators for each coastal
-community. Results are presented in terms of the difference through time
-and across sites, relative to the control site on the year of
-implementation (\emph{i.e.} the difference-in-differences estimate or
-effect size \(\lambda_t\) from Eqs. \ref{eqn:reg_bio} and
-\ref{eqn:soc_reg}). We also provide an overview of the governance
-settings of each community, and discuss how these might be related to
-the effectiveness and performance of the reserves.
-
-\hypertarget{biological-effects}{%
-\subsection{Biological effects}\label{biological-effects}}
-
-Indicators showed ambiguous responses through time for each reserve.
-Figure \ref{fig:indicators}A shows positive effect sizes for lobster
-densities in Isla Natividad and Punta Herrero during the first years,
-but the effect is eroded through time. In the case of Maria Elena,
-positive changes were observed in the third and fourth year. These
-effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
-Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
-for Maria Elena, but are not significantly different from zero
-(\(p > 0.05\)). Likewise, no significant changes were detected in fish
-biomass or invertebrate and fish densities (Fig.
-\ref{fig:indicators}B-D), where effect sizes oscillated around zero
-without clear trends. Figures and tables with time series of indicators
-and model coefficients are presented in the supplementary materials
-(Figures S1-S4, Tables S4-S6).
-
-\hypertarget{socioeconomic-effects}{%
-\subsection{Socioeconomic effects}\label{socioeconomic-effects}}
-
-Lobster landings and revenue were only available for Isla Natividad and
-Maria Elena (Fig \ref{fig:lobsters}). For all years before
-implementation, the effect sizes are close to zero, indicating that the
-control and treatment sites have similar pre-treatment trends,
-suggesting that these are plausible controls. However, effect sizes do
-not change after the implementation of the reserve. Interestingly, the
-negative effect observed for Isla Natividad on year 5 correspond to the
-2011 hypoxia events. The only positive change observed in lobster
-landings is for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
-post-implementation data for Maria Elena do not show a significant
-effect of the reserve. Isla Natividad shows higher revenues after the
-implementation of the reserve, as compared to the control communities.
-However, these changes are only significant for the third year
-(\(p < 0.05\)). Full tables with model coefficients are presented in the
-supplementary materials (Tables S4-S5).
-
-\hypertarget{governance}{%
-\subsection{Governance}\label{governance}}
-
-Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed
-communities share similarities known to foster sustainable resource
-management and increase reserve effectiveness. For example, fishers
-operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide
-exclusive access to resources and reserves. Along with their relatively
-small groups (A1 - Number of relevant actors), Isolation (A3),
-Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated
-sanctions (GS10.1), these fisheries have solid governance structures
-that enable them to monitor their resources and enforce rules to ensure
-sustainable management. In general, success of conservation initiatives
-depends on the incentives of local communities to maintain a healthy
-status of the resources upon which they depend \citep{jupiter_2017}. Due
-to the clarity of access rights and isolation, the benefits of
-conservation directly benefit the members of the fishing cooperatives,
-which have favored the development of efficient community-based
-enforcement systems. However, our SES analysis also highlights factors
-that might hinder reserve performance or mask outcomes. While total
-reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual
-reserves are often small (RS3); the largest reserve is only 4.37 km
-\textsuperscript{2}, and the smallest one is 0.09 km
-\textsuperscript{2}. Reserves are also relatively young (RS5).
-Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely
-that marine reserves will result in increased catches.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Our results indicate that these TURF-reserves have not increased lobster
-densities. Additionally, no co-benefits were identified when using other
-ecological indicators aside from the previously reported buffering
-effect that reserves can have to environmental variability in Isla
-Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
-pertaining landings and revenues showed little to no change after
-reserve implementation. Lastly, the communities exhibit all the social
-enabling conditions for effective reserve and resource management. Here
-we discuss possible shortcomings in our analyses as well as possible
-explanations for the observed patterns.
-
-While many ecology studies have used BACI sampling designs and
-respective analyses (\emph{e.g.} \citet{stewartoaten_1986}), few
-conservation studies have done so to evaluate the effect of an
-intervention (\emph{e.g.}
-\citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted
-in a call for more robust analyses in conservation science
-\citep{guidetti_2002,ferraro_2006}. Our approach to evaluate the
-temporal and spatial changes provides a more robust measure of reserve
-effectiveness, and captures previously described patterns. For example,
-the rapid increase observed for lobster densities in Isla Natividad on
-the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), occurs a
-year after the hypoxia events described by \citet{micheli_2012-EU},
-which caused mass mortality of sedentary organisms such as abalone and
-sea urchins, but not lobster and finfish. The use of causal inference
-techniques may help us support evidence-based conservation.
-
-Our analyses of socioeconomic indicators has two limitations. First, we
-only look at landings and revenues by landings for communities with and
-without TURF-reserves. There are a number of other possible indicators
-that could show a change due to the implementation of the reserve.
-Notably, one often cited in the literature is additional benefits, such
-as tourism \citep{viana_2017}. However, it is unlikely that the
-evaluated communities will experience tourism benefits due to their
-remoteness and the lack of proper infrastructure to sustain tourism. A
-second limitation of our socioeconomic analysis is that we do not
-observe effort data, which may mask the effect of the reserve. For
-example, if catches remain relatively unchanged but fishing effort
-decreased, that would imply a larger catch per unit effort and thus
-higher profitability.
-
-A first possible explanation for the lack of effectiveness may be the
-young age of the reserves. Literature shows that age and enforcement are
-important factors that influence reserve effectiveness
-\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
-reserves, and our SES analysis suggests that all communities have a
-well-established community-based enforcement system. With these
-characteristics, one would expect the reserves to be effective. Maria
-Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
-\textless{} 6 years old) and effects may not yet be evident due to the
-short duration of protection, relative to the life histories of the
-protected species; community-based marine reserves in tropical
-ecosystems may take six years or more to show a spillover effect
-\citep{dasilva_2015-zX}.
-
-Another key condition for effectiveness is reserve size
-\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
-attributed to poor ecological coherence in reserve design (\emph{sensu}
-\citet{rees_2018}). Previous research has shown that reserves in Isla
-Natividad yield fishery benefits for the abalone fishery
-\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
-perhaps the reserves provide enough protection to these sedentary
-invertebrates, but not lobsters. Design principles developed by
-\citet{green_2017} for marine reserves in the Caribbean state that
-reserves ``should be more than twice the size of the home range of
-adults and juveniles'', and suggest that reserves seeking to protect
-spiny lobsters should have at least 14 km across. Furthermore, fishers
-may favor implementation of reserves that pose low fishing costs due to
-their small size or location. Our analysis of economic data supports
-this hypothesis, as neither landings nor revenues showed the expected
-short-term reductions associated to the first years of reserve
-implementation \citep{ovando_2016-Wg}.
-
-Even if reserves had appropriate sizes and were placed in optimal
-locations, there are other plausible explanations for the observed
-patterns. For instance, marine reserves are only likely to provide
-fisheries benefits if initial population sizes are low and the fishery
-is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
-fisheries were certified by the Marine Stewardship Council and are
-managed via species-specific minimum catch sizes, seasonal closures,
-protection of ``berried'' females, and escapement windows where traps
-are allowed \citep{dof_website_1993}. It is uncertain whether such a
-well-managed fishery will experience additional benefits from marine
-reserves; reserves implemented in TURFs where fishing pressure is
-already optimally managed will still show a trade-off between fisheries
-and conservation objectives \citep{lester_2017}. Furthermore,
-\citet{gelcich_2008} have shown that TURFs alone can have greater
-biomass and richness than areas operating under open access. This might
-reduce the difference between indicators from the TURF and reserve
-sites, making it difficult to detect such a small change. Further
-research should focus on evaluating sites in the reserve, TURF, and open
-access areas or similar Fish Refuges established without the presence of
-TURFs where the impact of the reserves might be greater.
-
-Finally, extreme conditions, including prolonged hypoxia, heat waves,
-and storms have affected both the Pacific and Caribbean regions, with
-large negative impacts of coastal marine species and ecosystems
-\citep{cavole_2016,hughes_2018,breitburg_2018}. The coastal ecosystems
-where these reserves are located have been profoundly affected by these
-events \citep{micheli_2012-EU,woodson_2018}. Effects of protection might
-be eliminated by the mortalities associated with these extreme
-conditions.
-
-While the evaluated reserves have failed to provide fishery benefits up
-to now, there are a number of additional ecological, fisheries, and
-social benefits. Marine reserves provide protection to a wider range of
-species and vulnerable habitat. Previous research focusing on these
-specific sites has shown that they serve as an insurance mechanism
-against uncertainty and errors in fisheries management, as well as mild
-environmental shocks
-\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
-Self-regulation of fishing effort can serve as a way to compensate for
-future declines associated to environmental variation
-\citep{finkbeiner_2018}. Furthermore, embarking in a marine conservation
-project can bring the community together, which promotes social cohesion
-and builds social capital \citep{fulton_2019}. Showing commitment to
-marine conservation and sustainable fishing practices has allowed
-fishers to have greater bargaining power and leverage over fisheries
-management \citep{prezramrez_2012}. These additional benefits might
-explain why communities show a positive perception about their
-performance and continue to support their presence by re-implementing
-the reserves \citep{ayer_2018}.
-
-Community-based TURF-reserves in small-scale fisheries may be helpful
-conservation and fishery management tools when appropriately implemented
-\citep{gelcich_2015}. We must promote bottom-up design and
-implementation processes like the ones in the evaluated reserves, but
-without setting design principles aside. Having full community support
-surely represents an advantage, but it is important that community-based
-TURF-reserves meet essential design principles such as size and
-placement so as to maximize their effectiveness. Furthermore, consvation
-and advocacy groups should consider the opportunity costs of such
-interventions (\emph{sensu} \citet{smith_2010}) and evaluate the
-potential of other approaches that may yield similar benefits.
-
-In terms of the fisheries regulation in Mexico, our work only evaluates
-Fish Refuges established within TURFs. Future research should aim at
-evaluating other Fish Refuges established as bottom-up processes but
-without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
-others established through top-down processes (\emph{i.e.}
-\citet{dof_websiteU_2018}), as well as the relationship between
-governance and effectiveness across this gradient of approaches. For the
-particular case of the reserves that we evaluate, the possibility of
-expanding reserves or merging existing polygons into larger areas should
-be evaluated and proposed to the communities.
-
-\section*{Conflict of Interest Statement}
-
-The authors declare that the research was conducted in the absence of
-any commercial or financial relationships that could be construed as a
-potential conflict of interest.
-
-\section*{Author Contributions}
-
-JC and AS conceived the idea. JC and EA analyzed data, discussed the
-results, and wrote the first draft. FM, SF, AS, JT, and AHV discussed
-the results and edited the manuscript. All authors provided valuable
-contributions.
-
-\section*{Funding}
-
-JCVD received funding from UCMexus - CONACyT Doctoral Fellowship (CVU
-669403) and the Latin American Fisheries Fellowship Program. AS, AHV, SF
-and JT received funding from Marisla Foundation, Packard Foundation,
-Walton Family Foundation, Summit Foundation, and Oak Foundation. FM was
-supported by NSF-CNH and NSF BioOce (grants DEB-1212124 and 1736830).
-
-\section*{Acknowledgments}
-
-The authors wish to acknowledge Imelda Amador for contributions on the
-governance data, as well as pre-processing biological data. This study
-would have not been possible without the effort by members of the
-fishing communities here mentioned, who participated in the
-data-collection process. The authors wish to acknowledge comments by the
-reviewrs and editor, which significantly improved the quality of this
-work.
-
-\clearpage
-
-\bibliographystyle{frontiersinSCNS_ENG_HUMS}
-\bibliography{references}
-
-\clearpage
-
-\section*{Figure captions}
-
-\begin{figure}
+\caption{
+{\bf List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}}
 \centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
-\caption{\label{fig:unnamed-chunk-7}\label{fig:map}Location of the three
-coastal communities studied (A). Isla Natividad (B) is located off the
-Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
-in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
-polygons the marine reserves.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{\label{fig:unnamed-chunk-8}\label{fig:indicators}Effect sizes for
-marine reserves from Isla Natividad (IN; red circles), Maria Elena (ME;
-blue triangles), and Punta Herrero (PH; green squares) for lobster
-densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
-densities (C), and fish densities (D). Plots are ordered by survey type
-(left column: invertebrates; right column: fish). Points are jittered
-hotizontally to avoid overplotting. Points indicate the effect size and
-error bars are heteroskedastic-robust standard errors. Years have been
-centered to year of implementation.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
-\caption{\label{fig:unnamed-chunk-9}\label{fig:lobsters}Effect sizes for
-lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
-circles) and Maria Elena (ME; blue triangles). Points are jittered
-hotizontally to avoid overplotting. Points indicate the effect size and
-error bars are heteroskedastic-robust standard errors. Years have been
-centered to year of implementation.}
-\end{figure}
-
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-10}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
-\centering
-\begin{tabular}[t]{l|l}
+\begin{tabular}{l|l}
 \hline
@@
 \end{tabular}
+\label{table:indicators}
 \end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-11}\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).
-Alphanumeric codes follow \citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.
-The presented narrative applies equally for all communities unless otherwise noted.}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12cm}}
-\hline
-Variable & Narrative\\
-\hline
-\multicolumn{2}{l}{\textbf{Resource System (RS)}}\\
-\hline
-\hspace{1em}RS2 - Clarity of system boundaries: Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates.
-Reserve placement is decided by the community.
-Fishers use GPS units and landmarks.\\
-\hline
-\hspace{1em}RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
-\hline
-\hspace{1em}RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
-\hline
-\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) MSC certified.\\
-\hline
-\hspace{1em}*RS5 - Age of reserves: Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
-\hline
-\multicolumn{2}{l}{\textbf{Resource Unit (RU)}}\\
-\hline
-\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish (2 spp each).
-Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates (4 spp).\\
-\hline
-\multicolumn{2}{l}{\textbf{Actors (A)}}\\
-\hline
-\hspace{1em}A1 - Number of relevant actors: Number of fishers & IN = 98; ME = 80; PH = 21\\
-\hline
-\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers.
-IN lies 545 Km south from Tijuana, and ME and PH 230 Km south from Cancun, where the nearest international airports are located.\\
-\hline
-\multicolumn{2}{l}{\textbf{Governance system (G)}}\\
-\hline
-\hspace{1em}GS6.1.4.3 - Territorial use communal rights : Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster.
-These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
-\hline
-\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandates.
-These are: larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
-\hline
-\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group (Consejo de vigilancia) that monitors and enforces formal and internal rules.
-They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
-\hline
-\hspace{1em}GS9.2 - Biophysical monitoring: Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds.
-Recently, they have installed oceanographic sensors to monitor oceanographic variables.\\
-\hline
-GS10.1 - Graduated sanctions & Fishers have penalties for breaking collective-choice rules or fishing inside the reserves.
-These may range from scoldings and warnings to not being allowed to harvest a particular resource or being expelled from the cooperative.\\
-\hline
-\end{tabular}}
-\end{table}
-
-\end{document}

### 2019-02-28 docs/tab/ses.tex
Migrated from Rmd to script and updated PLOS submission

--- 2018-12-31 docs/manuscript.tex
@@
-\documentclass{frontiersSCNS}
-\usepackage{url,hyperref,lineno,microtype,subcaption}
-\usepackage[onehalfspacing]{setspace}
-\usepackage{float}
+\begin{table}[!ht]
 
-\linenumbers
-
-\usepackage[utf8]{inputenc}
-\floatplacement{figure}{H}
-
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
-\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
-Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
-Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
-Micheli\(^{4}\)}
-% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
-% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
-of California, Santa Barbara, Santa Barbara, CA, USA \newline
-\(^{2}\)Nicholas School of the Environment, Duke University, Beaufort,
-NC, USA \newline \(^{3}\)Comunidad y Biodiversidad A.C., Guaymas,
-Sonora, Mexico \newline \(^{4}\)Hopkins Marine Station and Center for
-Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
-% The Corresponding Author should be marked with an asterisk
-% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
-Santa Barbara, Santa Barbara, CA, 93106}
-
-\def\corrEmail{\href{mailto:juancarlos@ucsb.edu}{\nolinkurl{juancarlos@ucsb.edu}}}
-
-\begin{document}
-\onecolumn
-\firstpage{1}
-
-\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in Mexican small-scale
-fisheries}
-
-\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-\extraAuth{}
-
-\maketitle
-
-\begin{abstract}
-
-Coastal marine ecosystems provide livelihoods for small-scale fishers
-and coastal communities around the world. Small-scale fisheries face
-great challenges since they are difficult to monitor, enforce, and
-manage. Combining territorial use rights for fisheries (TURF) with
-no-take marine reserves to create TURF-reserves can improve the
-performance of small-scale fisheries by buffering fisheries from
-environmental variability and management errors, while ensuring that
-fishers reap the benefits of conservation investments. In the last 12
-years, 18 old and new community-based Mexican TURF-reserves gained legal
-recognition thanks to a regulation passed in 2014; their effectiveness
-has not been formally evaluated. We combine causal inference techniques
-and the Social-Ecological Systems framework to provide a holistic
-evaluation of community-based TURF-reserves in three coastal communities
-in Mexico. We find that while reserves have not yet achieved their
-stated goal of increasing the density of lobster and other benthic
-invertebrates, they continue to receive support from the fishing
-communities. A lack of clear ecological and socioeconomic effects likely
-results from a combination of factors. First, some of these reserves
-might be too young for the effects to show. Second, the reserves are not
-large enough to protect mobile species, like lobster. Third, variable
-and extreme oceanographic conditions have impacted harvested
-populations. Fourth, local fisheries are already well managed, and it is
-unlikely that reserves might have a detectable effect in catches.
-However, these reserves may provide a foundation for establishing
-additional, larger marine reserves needed to effectively conserve mobile
-species.
-
-\medskip
-\tiny
- \keyFont{ \section{Keywords:} TURF-reserves, Causal Inference, Social-Ecological Systems, Marine
-Protected Areas, Marine Conservation, Small-Scale Fisheries}
-
-\end{abstract}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Marine ecosystems around the world sustain significant impacts due to
-overfishing and unsustainable fishing practices
-\citep{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular,
-small-scale fisheries face great challenges since they tend to be hard
-to monitor and enforce \citep{costello_2012}. One of the many approaches
-taken to improve the performance of coastal fisheries and health of the
-local resources is through the implementation of Territorial Use Rights
-for Fisheries (TURFs) that contain no-take marine reserves within them,
-thus creating TURF-reserve systems
-\citep{afflerbach_2014,gelcich_2015,lester_2017}.
-
-TURFs are a fisheries management tool in which a well defined group of
-fishers (\emph{e.g.} fishing cooperatives) have exclusive access to an
-explicitly delimited portion of the ocean. They promote a sense of
-stewardship and incentivise resource users to sustainably manage their
-resources \citep{gelcich_2008,costello_2010,mccay_2014}. On the other
-hand, no-take marine reserves (marine reserves from hereinafter) are
-areas where all extractive activities are off-limits. These can be
-implemented to protect biodiversity but also as fishery management tools
-to aid in the recovery of marine stocks. These instruments can be
-combined by establishing a marine reserve within a TURF, thus making
-them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.
-
-Conservation science has shown how marine reserves lead to increased
-biomass, species richness, and abundance within the protected regions
-\citep{lester_2009}, and that these may have a series of additional
-benefits like climate change mitigation, protection from environmental
-variability, and fisheries benefits
-\citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
-research on TURFs has shown that these areas have higher abundance of
-targeted species than sites operating under open access and even similar
-to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The
-benefits resulting from reserves established within TURFs (\emph{i.e.}
-TURF-reserves) should be captured exclusively by the group of fishers
-with exclusive access \citep{gelcich_2015}. Although in theory these
-systems are successful \citep{smallhornwest_2018}, there is little
-empirical evidence of their effectiveness and the drivers of their
-success. Moreover, TURF-reserve systems are inherently intricate
-social-ecological systems, and their effectiveness must depend on how
-environmental and social factors combine and interact
-\citep{ostrom_2009,gelcich_2015}. It is therefore important to consider
-not only the indicators of interest, bu also the governance settings
-under which the reserves occur.
-
-Recent norms in fisheries regulation in Mexico provide a ripe
-opportunity to study the effectiveness of community-based TURF-reserves
-in small-scale fisheries. In Mexico, a norm created in 2014 allows
-fishers to request legal recognition of community-based reserves as
-``Fish Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since
-2012, old and new marine reserves have gained legal recognition as Fish
-Refuges. Of these, 18 were originally implemented within TURFs. However,
-their effectiveness has not yet been formally evaluated and reported in
-the scientific literature.
-
-Here, we combine causal inference techniques and the Social-Ecological
-Systems (SES) framework to evaluate community-based TURF-reserves in
-three coastal communities in Mexico. The objective of this work is
-twofold. First, to provide a holistic evaluation of the effectiveness of
-community-based TURF-reserves in terms of the changes in biological and
-socioeconomic indicators and the governance settings under which these
-develop, which may inform similar processes in other countries. Second,
-to identify opportunities where improvement or adjustment might lead to
-increased effectiveness. We draw from lessons learned in these three
-case studies and provide management recommendations to maximize the
-effectiveness of community-based TURF-reserves in small-scale fisheries
-where this tool is used to manage and rebuild their coastal fisheries.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{turf-reserves-in-mexico}{%
-\subsection{TURF-reserves in Mexico}\label{turf-reserves-in-mexico}}
-
-Before discussing our data collection methods and describing our
-analyses, our case studies warrant some more background. Community-based
-marine reserves that are implemented within TURFs are a form of
-TURF-reserves, voluntarily established and enforced by local
-communities. This bottom-up approach increases compliance and
-self-enforcement, and reserves can yield benefits similar to
-systematically-designed reserves \citep{beger_2004,smallhornwest_2018}.
-Community-based spatial closures occur in different contexts, like the
-\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
-\citep{bohnsack_2004,johannes_2002}. However, community-based reserves
-can be hard to enforce if they are not legally recognized. In such
-conditions, TURF fishers must rely on the exclusive access of the TURF
-to maintain high levels of compliance.
-
-In an effort to bridge this normative gap, Mexican Civil Society
-Organizations (CSOs) served as a link between fishers and government,
-and created a legal framework that solves this governance issue: Fish
-Refuges \citep{nom}. Fish Refuges can be implemented as temporal or
-partial reserves, which can protect one, some, or all resources within
-their boundaries. One of the ways in which fishing communities have
-taken advantage of this new tool is by implementing temporal marine
-reserves within their TURFs with a defined expiration date (often 5
-years). When the expiration date is reached, fishers can chose to open
-the reserves to fishing or re-establish them. Our work focuses on Fish
-Refuges implemented as community-based TURF-reserves that occur in
-small-scale fisheries.
-
-The common setup of community-based TURF-reserves in Mexico is the
-following. Fishers from a given community are assembled in fishing
-cooperatives which have exclusive fishing rights over a spatially
-delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
-\ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and
-each community usually hosts no more than one cooperative. The profits
-from each TURF are shared amongst all fishers from the cooperative.
-Fishing cooperatives interested in implementing marine reserves work
-with CSOs to implement marine reserves within their TURFs (\emph{i.e.}
-TURF-reserves). Fishers then ask the government to grant legal
-recognition to their TURF-reserves as Fish Refuges following a series of
-studies outlined in the regulation \citep{nom}.
-
-\hypertarget{study-areas}{%
-\subsection{Study areas}\label{study-areas}}
-
-We evaluate three community-based no-take TURF-reserves implemented in
-Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig
-\ref{fig:map}A). The first one was created by the \emph{Buzos y
-Pescadores de la Baja California} fishing cooperative, located in Isla
-Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The
-main fishery in the island is the spiny lobster (\emph{Panulirus
-interruptus}), but other resources like finfish, sea cucumber, read sea
-urchin, snail, and abalone are also an important source of income. In
-2006, the community decided to implement two marine reserves within
-their fishing grounds. The objective of these reserves was ``to protect
-commercially important invertebrate species''; mainly lobster and
-abalone. These reserves obtained legal recognition in 2018
-\citep{dof_website_2018}.
-
-The other two TURF-reserves are located in Maria Elena and Punta
-Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
-Isla Nativdad, which hosts a well established fishing community, Maria
-Elena is a fishing camp --visited intermittently during the fishing
-season-- belonging to the \emph{Cozumel} fishing cooperative; Punta
-Herrero is home to the \emph{José María Azcorra} fishing cooperative,
-and similar to Isla Natividad hosts a local community. Their main
-fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but
-they also target finfish in the off-season. Maria Elena and Punta
-Herrero established eight and four marine reserves in 2012 and 2013,
-respectively. These reserves have been legally recognized as Fishing
-Refuges since their original implementation
-\citep{dof_website_2012,dof_website_2013} and subsequent
-re-establishments \citep{dof_website_2017b}.
-
-These communities are representative of their region in terms of
-ecology, socioeconomic, and governance aspects. Isla Natividad, for
-example, is part of a greater group of fishing cooperatives belonging to
-a Federation of Fishing Cooperatives. This group has been identified as
-a cohesive group that cooperates to better manage their resources
-\citep{mccay_2014,mccay_2017,acevesbueno_2017}. Likewise, Maria Elena
-and Punta Herrero are representative of fishing cooperatives in the
-Mexican Caribbean, which are also part of a regional Federation.
-Together, these three communities provide an accurate representation of
-other fishing communities that have been historically manged with TURFs
-in each of their regions. While each region has additional communities
-that have established community-based TURF-reserves, available data
-would not allow us to perform the in-depth causal inference analysis
-that we undertake. Yet, given the similarities among communities and the
-socioeconomic and governance setting under which they operate, it is
-safe to cautiously generalize our insights to other similar
-community-based TURF-reserves in Mexico and elsewhere.
-
-The regulation governing the implementation of Fish Refuges states that
-these are fishery management tools intended to have biological or
-socioeconomic benefits \citep{nom}. For this reason, the main portion of
-our analyses focuses on a series of biological and socioeconomic
-indicators that may respond to reserve implementation. However, the
-effectiveness of conservation and fisheries management interventions
-also depends on the social and governance structures in place. We
-therefore incorporate a reduced version of the Social Ecological Systems
-framework \citep{ostrom_2009} and evaluate variables and indicators
-known to aid and hinder the effectiveness of management interventions in
-conservation and fisheries. The incorporation of the SES is not intended
-to relate different levels of governance with reserve effectiveness, but
-rather help provide context on the social-ecological system in which
-reserves develop. The following two sections describe our data
-collection methods and analyses.
-
-\hypertarget{data-collection}{%
-\subsection{Data collection}\label{data-collection}}
-
-We use three main sources of information to evaluate these reserves
-across the ecological, socioeconomic, and governance dimensions.
-Ecological data come from the annual ecological monitoring of reserve
-and control sites. Reserve sites are areas where no fishing occurs.
-Control sites are areas that meet the following criteria: i) habitat
-characteristics are similar to the corresponding reserves, ii)
-presumably had a similar probability of being selected as reserves
-during the design phase, iii) are located within the TURF, where fishing
-occurs, and iv) Are not directly adjacent to the reserves. We focus our
-evaluation on sites where data are available for reserve and control
-sites, before and after the implementation of the reserve. This provides
-us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design
-that allows us to capture and control for temporal and spatial dynamics
-\citep{stewartoaten_1986,depalma_2018} and causally attribute the
-changes to the reserve
-\citep{francinifilho_2008,Villasenor-Derbez_2018}.
-
-The biological data are collected by members from each community and
-personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
-(\href{www.cobi.org.mx}{COBI}). Trained divers record species richness
-and abundances of fish and invertebrate species along replicate
-transects (30 \(\times\) 2 m each) at depths 5-20 m in the reserves and
-control sites \citep{suman_2010-ez,fulton_2018,fulton_2019}. Size
-structures are also collected during fish surveys. All sites were
-surveyed annually, and at least once before implementation of the
-reserves. A summary of sampling effort is shown in the supplementary
-materials (Tables S1-S2).
-
-Socioeconomic data come from landing receipts reported to the National
-Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
-Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
-(Kg) and revenues (MXP) for TURF-managed cooperatives with and without
-marine reserves. In this case our treated unit are the cooperatives
-(\emph{i.e.} communities) that have implemented a reserve within their
-TURF, and the controls are nearby communities that have a TURF but did
-not implement a reserve. Cooperatives incorporated in this analysis
-belong to larger regional-level Cooperative Federations, and are exposed
-to the same markets and institutional frameworks, making them plausible
-controls \citep{mccay_2017,ayer_2018}. Landings and revenues were
-aggregated at the cooperative-year level, and revenues were adjusted to
-represent 2014 values by the Consumer Price Index for Mexico
-\citep{oecd_2017}. A table with summary statistics for this data is
-provided in the supplementary materials (Table S3, Figure S5).
-
-Data for the evaluation of the SES were collected at the community-level
-from official documents used in the design, creation, and designation of
-the marine reserves. These include the technical studies that the
-cooperatives submit when they request recognition of their reserves, as
-well as the official enactments
-\citep{dof_website_2012,dof_website_2013,dof_website_2018}. We also
-complimented information based on the authors' experience and knowledge
-of the communities. We collected information on the Resource Systems,
-Resource Units, Actors, and Governance System (Table \ref{table:ses}).
-The next section further describe the variables and indicators chosen
-for each of these.
-
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
-
-We evaluate the effect that the TURF-reserves have had on four
-ecological and two socioeconomic indicators shown in Table
-\ref{table:indicators}. Recall that reserves were implemented to protect
-lobster and other benthic invertebrates. However, we also use the
-available fish and invertebrate data to test for associated co-benefits.
-
-We use a difference-in-differences analysis to evaluate these
-indicators. This approach is widely used in econometric literature to
-estimate the average treatment effect of an intervention, like the
-impact of minimum wage increases on employment rates \citep{card_1994}.
-In our case it allows us to estimate the effect that the reserve had on
-each biological and socioeconomic indicator (Table
-\ref{table:indicators}) by comparing trends across time and treatments
-\citep{moland_2013,Villasenor-Derbez_2018}. To perform
-difference-in-differences, we regress the indicator of interest on a
-dummy variable for treatment, a dummy variable for years, and the
-interaction term between these with a multiple linear regression of the
-form:
-
-\begin{equation}
-I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t}
-\label{eqn:reg_bio}
-\end{equation}
-
-Where year-level fixed effects capturing a temporal trend are
-represented by \(\gamma_t Year_t\), and \(\beta Zone_i\) captures the
-difference between reserve (\(Zone = 1\)) and control (\(Zone = 0\))
-sites. The effect of the reserve is captured by the \(\lambda_t\)
-coefficient, and represents the difference observed between the control
-site before the implementation of the reserve and the treated sites at
-time \(t\) after controlling for other time and space variations
-(\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Therefore, we
-would expect this term to be positive if the indicator increases because
-of the reserve. Finally, \(\epsilon_{i,t}\) represents the error term of
-the regression.
-
-Socioeconomic indicators are evaluated with a similar approach. Due to
-data constrains, we only evaluate socioeconomic data for Isla Natividad
-(2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are
-used as counterfactuals that allow us to control for unobserved
-time-invariants. Each focal community (\emph{i.e.} Isla Natividad and
-Maria Elena) has three counterfactual communities.
-
-\begin{equation}
-I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t}
-\label{eqn:soc_reg}
-\end{equation}
-
-The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio},
-but in this case the \(Treated\) dummy variable indicates if the
-community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)).
-These regression models allow us to establish a causal link between the
-implementation of marine reserves and the observed trends by accounting
-for temporal and site-specific dynamics \citep{depalma_2018}. We fit one
-model per community and indicators (\emph{e.g.} there are three models
-for lobster density, one for each community) for a total of 12
-biological model fits and four socioeconomic model fits. Model
-coefficients were estimated via ordinary least-squares and used
-heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
-analyses were performed in R version 3.5.1 and R Studio version 1.1.456
-\citep{R_2018}. All data and code needed to reproduce our analyses are
-available in a GitHub repository at:
-\url{https://github.com/jcvdav/ReserveEffect}.
-
-We use the SES framework to evaluate each community and create a
-narrative that provides context for each community. The use of this
-framework standardizes our analysis and allows us to communicate our
-results in a common language across fields by using a set of previously
-defined variables and indicators. We based our variable selection
-primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who
-operationalized and analyzed Mexican fishing cooperatives using this
-framework, and identified the key variables relevant to fishing
-cooperatives in Mexico. We also incorporate other relevant variables
-known to influence reserve performance following
-\citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table
-\ref{table:ses} shows the selected variables, along with definitions and
-values.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-The following sections present the effect that marine reserves had on
-each of the biological and socioeconomic indicators for each coastal
-community. Results are presented in terms of the difference through time
-and across sites, relative to the control site on the year of
-implementation (\emph{i.e.} the difference-in-differences estimate or
-effect size \(\lambda_t\) from Eqs. \ref{eqn:reg_bio} and
-\ref{eqn:soc_reg}). We also provide an overview of the governance
-settings of each community, and discuss how these might be related to
-the effectiveness and performance of the reserves.
-
-\hypertarget{biological-effects}{%
-\subsection{Biological effects}\label{biological-effects}}
-
-Indicators showed ambiguous responses through time for each reserve.
-Figure \ref{fig:indicators}A shows positive effect sizes for lobster
-densities in Isla Natividad and Punta Herrero during the first years,
-but the effect is eroded through time. In the case of Maria Elena,
-positive changes were observed in the third and fourth year. These
-effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
-Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
-for Maria Elena, but are not significantly different from zero
-(\(p > 0.05\)). Likewise, no significant changes were detected in fish
-biomass or invertebrate and fish densities (Fig.
-\ref{fig:indicators}B-D), where effect sizes oscillated around zero
-without clear trends. Figures and tables with time series of indicators
-and model coefficients are presented in the supplementary materials
-(Figures S1-S4, Tables S4-S6).
-
-\hypertarget{socioeconomic-effects}{%
-\subsection{Socioeconomic effects}\label{socioeconomic-effects}}
-
-Lobster landings and revenue were only available for Isla Natividad and
-Maria Elena (Fig \ref{fig:lobsters}). For all years before
-implementation, the effect sizes are close to zero, indicating that the
-control and treatment sites have similar pre-treatment trends,
-suggesting that these are plausible controls. However, effect sizes do
-not change after the implementation of the reserve. Interestingly, the
-negative effect observed for Isla Natividad on year 5 correspond to the
-2011 hypoxia events. The only positive change observed in lobster
-landings is for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
-post-implementation data for Maria Elena do not show a significant
-effect of the reserve. Isla Natividad shows higher revenues after the
-implementation of the reserve, as compared to the control communities.
-However, these changes are only significant for the third year
-(\(p < 0.05\)). Full tables with model coefficients are presented in the
-supplementary materials (Tables S4-S5).
-
-\hypertarget{governance}{%
-\subsection{Governance}\label{governance}}
-
-Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed
-communities share similarities known to foster sustainable resource
-management and increase reserve effectiveness. For example, fishers
-operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide
-exclusive access to resources and reserves. Along with their relatively
-small groups (A1 - Number of relevant actors), Isolation (A3),
-Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated
-sanctions (GS10.1), these fisheries have solid governance structures
-that enable them to monitor their resources and enforce rules to ensure
-sustainable management. In general, success of conservation initiatives
-depends on the incentives of local communities to maintain a healthy
-status of the resources upon which they depend \citep{jupiter_2017}. Due
-to the clarity of access rights and isolation, the benefits of
-conservation directly benefit the members of the fishing cooperatives,
-which have favored the development of efficient community-based
-enforcement systems. However, our SES analysis also highlights factors
-that might hinder reserve performance or mask outcomes. While total
-reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual
-reserves are often small (RS3); the largest reserve is only 4.37 km
-\textsuperscript{2}, and the smallest one is 0.09 km
-\textsuperscript{2}. Reserves are also relatively young (RS5).
-Additionally, fishers harvest healthy stocks (RS4.1), and it's unlikely
-that marine reserves will result in increased catches.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Our results indicate that these TURF-reserves have not increased lobster
-densities. Additionally, no co-benefits were identified when using other
-ecological indicators aside from the previously reported buffering
-effect that reserves can have to environmental variability in Isla
-Natividad \citep{micheli_2012-EU}. The socioeconomic indicators
-pertaining landings and revenues showed little to no change after
-reserve implementation. Lastly, the communities exhibit all the social
-enabling conditions for effective reserve and resource management. Here
-we discuss possible shortcomings in our analyses as well as possible
-explanations for the observed patterns.
-
-While many ecology studies have used BACI sampling designs and
-respective analyses (\emph{e.g.} \citet{stewartoaten_1986}), few
-conservation studies have done so to evaluate the effect of an
-intervention (\emph{e.g.}
-\citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted
-in a call for more robust analyses in conservation science
-\citep{guidetti_2002,ferraro_2006}. Our approach to evaluate the
-temporal and spatial changes provides a more robust measure of reserve
-effectiveness, and captures previously described patterns. For example,
-the rapid increase observed for lobster densities in Isla Natividad on
-the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A), occurs a
-year after the hypoxia events described by \citet{micheli_2012-EU},
-which caused mass mortality of sedentary organisms such as abalone and
-sea urchins, but not lobster and finfish. The use of causal inference
-techniques may help us support evidence-based conservation.
-
-Our analyses of socioeconomic indicators has two limitations. First, we
-only look at landings and revenues by landings for communities with and
-without TURF-reserves. There are a number of other possible indicators
-that could show a change due to the implementation of the reserve.
-Notably, one often cited in the literature is additional benefits, such
-as tourism \citep{viana_2017}. However, it is unlikely that the
-evaluated communities will experience tourism benefits due to their
-remoteness and the lack of proper infrastructure to sustain tourism. A
-second limitation of our socioeconomic analysis is that we do not
-observe effort data, which may mask the effect of the reserve. For
-example, if catches remain relatively unchanged but fishing effort
-decreased, that would imply a larger catch per unit effort and thus
-higher profitability.
-
-A first possible explanation for the lack of effectiveness may be the
-young age of the reserves. Literature shows that age and enforcement are
-important factors that influence reserve effectiveness
-\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
-reserves, and our SES analysis suggests that all communities have a
-well-established community-based enforcement system. With these
-characteristics, one would expect the reserves to be effective. Maria
-Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
-\textless{} 6 years old) and effects may not yet be evident due to the
-short duration of protection, relative to the life histories of the
-protected species; community-based marine reserves in tropical
-ecosystems may take six years or more to show a spillover effect
-\citep{dasilva_2015-zX}.
-
-Another key condition for effectiveness is reserve size
-\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
-attributed to poor ecological coherence in reserve design (\emph{sensu}
-\citet{rees_2018}). Previous research has shown that reserves in Isla
-Natividad yield fishery benefits for the abalone fishery
-\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
-perhaps the reserves provide enough protection to these sedentary
-invertebrates, but not lobsters. Design principles developed by
-\citet{green_2017} for marine reserves in the Caribbean state that
-reserves ``should be more than twice the size of the home range of
-adults and juveniles'', and suggest that reserves seeking to protect
-spiny lobsters should have at least 14 km across. Furthermore, fishers
-may favor implementation of reserves that pose low fishing costs due to
-their small size or location. Our analysis of economic data supports
-this hypothesis, as neither landings nor revenues showed the expected
-short-term reductions associated to the first years of reserve
-implementation \citep{ovando_2016-Wg}.
-
-Even if reserves had appropriate sizes and were placed in optimal
-locations, there are other plausible explanations for the observed
-patterns. For instance, marine reserves are only likely to provide
-fisheries benefits if initial population sizes are low and the fishery
-is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster
-fisheries were certified by the Marine Stewardship Council and are
-managed via species-specific minimum catch sizes, seasonal closures,
-protection of ``berried'' females, and escapement windows where traps
-are allowed \citep{dof_website_1993}. It is uncertain whether such a
-well-managed fishery will experience additional benefits from marine
-reserves; reserves implemented in TURFs where fishing pressure is
-already optimally managed will still show a trade-off between fisheries
-and conservation objectives \citep{lester_2017}. Furthermore,
-\citet{gelcich_2008} have shown that TURFs alone can have greater
-biomass and richness than areas operating under open access. This might
-reduce the difference between indicators from the TURF and reserve
-sites, making it difficult to detect such a small change. Further
-research should focus on evaluating sites in the reserve, TURF, and open
-access areas or similar Fish Refuges established without the presence of
-TURFs where the impact of the reserves might be greater.
-
-Finally, extreme conditions, including prolonged hypoxia, heat waves,
-and storms have affected both the Pacific and Caribbean regions, with
-large negative impacts of coastal marine species and ecosystems
-\citep{cavole_2016,hughes_2018,breitburg_2018}. The coastal ecosystems
-where these reserves are located have been profoundly affected by these
-events \citep{micheli_2012-EU,woodson_2018}. Effects of protection might
-be eliminated by the mortalities associated with these extreme
-conditions.
-
-While the evaluated reserves have failed to provide fishery benefits up
-to now, there are a number of additional ecological, fisheries, and
-social benefits. Marine reserves provide protection to a wider range of
-species and vulnerable habitat. Previous research focusing on these
-specific sites has shown that they serve as an insurance mechanism
-against uncertainty and errors in fisheries management, as well as mild
-environmental shocks
-\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
-Self-regulation of fishing effort can serve as a way to compensate for
-future declines associated to environmental variation
-\citep{finkbeiner_2018}. Furthermore, embarking in a marine conservation
-project can bring the community together, which promotes social cohesion
-and builds social capital \citep{fulton_2019}. Showing commitment to
-marine conservation and sustainable fishing practices has allowed
-fishers to have greater bargaining power and leverage over fisheries
-management \citep{prezramrez_2012}. These additional benefits might
-explain why communities show a positive perception about their
-performance and continue to support their presence by re-implementing
-the reserves \citep{ayer_2018}.
-
-Community-based TURF-reserves in small-scale fisheries may be helpful
-conservation and fishery management tools when appropriately implemented
-\citep{gelcich_2015}. We must promote bottom-up design and
-implementation processes like the ones in the evaluated reserves, but
-without setting design principles aside. Having full community support
-surely represents an advantage, but it is important that community-based
-TURF-reserves meet essential design principles such as size and
-placement so as to maximize their effectiveness. Furthermore, consvation
-and advocacy groups should consider the opportunity costs of such
-interventions (\emph{sensu} \citet{smith_2010}) and evaluate the
-potential of other approaches that may yield similar benefits.
-
-In terms of the fisheries regulation in Mexico, our work only evaluates
-Fish Refuges established within TURFs. Future research should aim at
-evaluating other Fish Refuges established as bottom-up processes but
-without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
-others established through top-down processes (\emph{i.e.}
-\citet{dof_websiteU_2018}), as well as the relationship between
-governance and effectiveness across this gradient of approaches. For the
-particular case of the reserves that we evaluate, the possibility of
-expanding reserves or merging existing polygons into larger areas should
-be evaluated and proposed to the communities.
-
-\section*{Conflict of Interest Statement}
-
-The authors declare that the research was conducted in the absence of
-any commercial or financial relationships that could be construed as a
-potential conflict of interest.
-
-\section*{Author Contributions}
-
-JC and AS conceived the idea. JC and EA analyzed data, discussed the
-results, and wrote the first draft. FM, SF, AS, JT, and AHV discussed
-the results and edited the manuscript. All authors provided valuable
-contributions.
-
-\section*{Funding}
-
-JCVD received funding from UCMexus - CONACyT Doctoral Fellowship (CVU
-669403) and the Latin American Fisheries Fellowship Program. AS, AHV, SF
-and JT received funding from Marisla Foundation, Packard Foundation,
-Walton Family Foundation, Summit Foundation, and Oak Foundation. FM was
-supported by NSF-CNH and NSF BioOce (grants DEB-1212124 and 1736830).
-
-\section*{Acknowledgments}
-
-The authors wish to acknowledge Imelda Amador for contributions on the
-governance data, as well as pre-processing biological data. This study
-would have not been possible without the effort by members of the
-fishing communities here mentioned, who participated in the
-data-collection process. The authors wish to acknowledge comments by the
-reviewrs and editor, which significantly improved the quality of this
-work.
-
-\clearpage
-
-\bibliographystyle{frontiersinSCNS_ENG_HUMS}
-\bibliography{references}
-
-\clearpage
-
-\section*{Figure captions}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
-\caption{\label{fig:unnamed-chunk-7}\label{fig:map}Location of the three
-coastal communities studied (A). Isla Natividad (B) is located off the
-Baja California Peninsula, Maria Elena and Punta Herrero (C) are located
-in the Yucatan Peninsula. Blue polygons represent the TURFs, and red
-polygons the marine reserves.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{\label{fig:unnamed-chunk-8}\label{fig:indicators}Effect sizes for
-marine reserves from Isla Natividad (IN; red circles), Maria Elena (ME;
-blue triangles), and Punta Herrero (PH; green squares) for lobster
-densities (\emph{Panulirus spp}; A), fish biomass (B), invertebrate
-densities (C), and fish densities (D). Plots are ordered by survey type
-(left column: invertebrates; right column: fish). Points are jittered
-hotizontally to avoid overplotting. Points indicate the effect size and
-error bars are heteroskedastic-robust standard errors. Years have been
-centered to year of implementation.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
-\caption{\label{fig:unnamed-chunk-9}\label{fig:lobsters}Effect sizes for
-lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
-circles) and Maria Elena (ME; blue triangles). Points are jittered
-hotizontally to avoid overplotting. Points indicate the effect size and
-error bars are heteroskedastic-robust standard errors. Years have been
-centered to year of implementation.}
-\end{figure}
-
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-10}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
-\centering
-\begin{tabular}[t]{l|l}
-\hline
-Indicator & Units\\
-\hline
-\multicolumn{2}{l}{\textbf{Biological}}\\
-\hline
-\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
-\hline
-\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
-\hline
-\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
-\hline
-\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
-\hline
-\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
-\hline
-\hspace{1em}Income from target species & M MXP\\
-\hline
-\hspace{1em}Landings from target species & Metric Tonnes\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:unnamed-chunk-11}\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero).
-Alphanumeric codes follow \citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}.
-The presented narrative applies equally for all communities unless otherwise noted.}
+\caption{{\bf Variables for the Social-Ecological System analysis.}}
 \centering
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12cm}}
+\begin{tabular}{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12.2cm}}
 \hline
@@
 \end{tabular}}
+\label{table:ses}
+IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero.
+Alphanumeric codes for variables follow \cite{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \cite{difranco_2016-Xw} and \cite{edgar_2014-UO}.
+The presented narrative applies equally for all communities unless otherwise noted.
 \end{table}
-
-\end{document}

### 2019-02-28 scripts/processing/data_cleaning_logbook.Rmd
Migrated from Rmd to script and updated PLOS submission

--- 2019-02-21 scripts/data_cleaning_logbook.Rmd
@@
   filter(!is.na(rc)) %>% 
-  # filter(rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
-  #                  "Cabezo - Cabezo (Control)",
-  #                  "Manchon - Manchon (Control)",
-  #                  "El Faro - El Faro (Control)")) %>% 
-  # filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
+  filter(rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
+                   "Cabezo - Cabezo (Control)",
+                   "Manchon - Manchon (Control)",
+                   "El Faro - El Faro (Control)")) %>%
+  # filter(!(rc == "El Faro - El Faro (Control)" & ano == 2012)) %>%
   filter(between(abundancia, 0.1, 100)) %>%
@@
   filter(!is.na(rc)) %>% 
-  # filter(rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
-  #                  "Cabezo - Cabezo (Control)",
-  #                  "Manchon - Manchon (Control)",
-  #                  "El Faro - El Faro (Control)")) %>% 
-  # filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>% 
+  filter(rc %in% c("La Plana / Las Cuevas - La Dulce / Babencho",
+                   "Cabezo - Cabezo (Control)",
+                   "Manchon - Manchon (Control)",
+                   "El Faro - El Faro (Control)")) %>%
+  filter(!(rc %in% c("El Faro - El Faro (Control)") & ano == 2012)) %>%
   filter(abundancia > 0) %>% 

### 2019-02-28 scripts/abundance.R

@@ -1,17 +1,0 @@
-abundance <- function(x, com, spp = NULL){
-  
-  if(!is.null(spp)){
-    x %<>% filter(generoespecie == spp)
-  }
-  
-  results <- x %>% 
-    filter(comunidad == com) %>%
-    mutate(ano = as.factor(ano),
-           year = as.factor(year),
-           year = fct_relevel(year, "0")) %>% 
-    group_by(rc, ano, year, sitio, zona, transecto, generoespecie) %>% 
-    summarize(indicador = sum(abundancia)/60) %>% 
-    ungroup()
-    
-    return(results)
-}

### 2019-02-28 scripts/biomass.R

@@ -1,13 +1,0 @@
-biomass <- function(x, com){
-  results <- x %>% 
-    filter(comunidad == com) %>% 
-    mutate(ano = as.factor(ano),
-           year = as.factor(year),
-           year = fct_relevel(year, "0"),
-           biomass = abundancia * a * (talla^b)/1000) %>% 
-    group_by(rc, ano, year, sitio, zona, transecto, generoespecie) %>% 
-    summarize(indicador = sum(biomass, na.rm = T)/60) %>% 
-    ungroup()
-  
-  return(results)
-}

### 2019-02-28 scripts/content/fig_bio_results.R

@@ -1,0 +1,161 @@
+#########################
+##  fig_bio_results    ##
+#########################
+
+####################################################
+# Fit the ecological models and create a termplot
+####################################################
+
+
+
+## Load packages
+library(cowplot)
+library(here)
+library(tidyverse)
+
+## Load custom functions
+# Load indicator functions
+source(file = here("scripts", "functions", "abundance.R")) #Calculate abundance per transect
+source(file = here("scripts", "functions", "biomass.R")) # Calculate fish biomass per transect
+
+# Load analysis functions
+source(file = here("scripts", "functions", "model.R")) #fits the model
+source(file = here("scripts", "functions", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
+# source(file = here("scripts", "robust_se_list.R")) #heteroskedastic-robust Standard Errors for stargazer
+source(file = here("scripts", "functions", "plot_model.R")) #plots effect sizes for Bio
+# source(file = here("scripts", "plot_socioeco.R")) #plots effect sizes for Socioeco
+
+# Load data
+# # Invertebrate data
+invert <- read.csv(file = here("data", "invertebrates.csv"),
+                   stringsAsFactors = F)
+# Fish data
+fish <- read.csv(file = here("data", "fish.csv"),
+                 stringsAsFactors = F)
+
+###### Analyses ######################################################
+# Invertebrates ###################################
+## Lobsters
+### Natividad
+L_IN <- invert %>% 
+  spread(generoespecie, abundancia, fill = 0) %>% 
+  gather(generoespecie,
+         abundancia,
+         -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>%
+  filter(generoespecie == "Panulirus interruptus") %>% 
+  abundance(com = "Isla Natividad") %>% 
+  model(sfe = F, spfe = F)
+### Maria Elena
+L_ME <- invert %>% 
+  spread(generoespecie, abundancia, fill = 0) %>% 
+  gather(generoespecie,
+         abundancia,
+         -c("comunidad", "ano", "year", "rc", "sitio", "zona", "transecto")) %>% 
+  filter(generoespecie == "Panulirus argus") %>% 
[…]

### 2019-02-28 scripts/content/fig_socioeco_results.R

@@ -1,0 +1,77 @@
+###############################
+##  fig_socioeco_results    ##
+##############################
+
+####################################################
+# Fit the socioeconomic models and create a termplot
+####################################################
+
+## Load packages
+library(kableExtra)
+library(broom)
+library(cowplot)
+library(here)
+library(tidyverse)
+
+## Load custom functions
+source(file = here("scripts", "functions", "plot_socioeco.R")) #plots effect sizes for Socioeco
+
+# Load data
+conapesca <- read.csv(file = here("data", "conapesca.csv"),
+                      stringsAsFactors = F) %>% 
+  mutate(years_centered = as.factor(years_centered),
+         year = fct_relevel(years_centered, "0"),
+         peso_vivo = peso_vivo/1e3,
+         valor = valor/1e6,
+         loc = zona,
+         zona = ifelse(grupo == "Treated", "Reserva", "Control"),
+         grupo = loc)
+
+# Fit models
+C_IN <- conapesca %>% 
+  filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
+  filter(grupo == "PN") %>%
+  lm(peso_vivo ~ zona * year, data = .)
+
+C_ME <- conapesca %>% 
+  filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
+  filter(grupo == "MC") %>%
+  lm(peso_vivo ~ zona * year, data = .)
+
+V_IN <- conapesca %>% 
+  filter(comunidad == "Isla Natividad" | is.na(comunidad)) %>% 
+  filter(grupo == "PN") %>%
+  lm(valor ~ zona * year, data = .)
+
+V_ME <- conapesca %>% 
+  filter(comunidad == "Maria Elena" | is.na(comunidad)) %>% 
+  filter(grupo == "MC") %>%
+  lm(valor ~ zona * year, data = .)
+
+# Plot terms
+c_plot <- plot_socioeco(model_list = list(C_IN, C_ME),
+                        model_names = list("IN", "ME"),
+                        ylab = quo(lambda[t])) +
+  theme(legend.justification = c(0, 0),
+        legend.position = c(0, 0.55),
+        legend.title = element_blank())
+
+v_plot <- plot_socioeco(model_list = list(V_IN, V_ME),
+                        model_names = list("IN", "ME"),
+                        ylab = quo(lambda[t])) +
+  theme(legend.position = "none")
+
[…]

### 2019-02-28 scripts/content/img_map.R

@@ -1,0 +1,156 @@
+##################
+##    fig_map   ##
+##################
+
+######################################################
+# Creates a map of the study area. The approximate
+# figure caption is: Location of the three coastal
+# communities studied (A). Isla Natividad (B) is
+# located off the Baja California Peninsula, Maria
+# Elena and Punta Herrero (C) are located in the
+# Yucatan Peninsula. Blue polygons represent the TURFs,
+# and red polygons the marine reserves.
+######################################################
+
+# Load packages
+library(tmap) #To load the WORLD dataset
+library(sf)
+library(ggsn)
+library(ggrepel)
+library(cowplot)
+library(here)
+library(tidyverse)
+
+
+######## DATA ##############
+
+# Determine a projection for the maps
+proj <- "+proj=longlat +datum=NAD27"
+
+# Load coastline for Mexico and World
+data(World)
+World <- st_transform(World, proj)
+
+# Load high-res mexican coastline
+coastline_mx <- readRDS(here("raw_data", "spatial", "coastline_mx.rds"))
+
+# Load ZRP shapefile
+zrp <- st_read(dsn = here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
+  filter(Name %in% c("El Cabezo",
+                     "Gallineros",
+                     "Punta Loria",
+                     "San Roman Norte",
+                     "San Roman Sur",
+                     "Anegado de Chal",
+                     "Laguna Canche Balam",
+                     "Niluc",
+                     "Mimis",
+                     "El Faro (Pesca de Langosta)",
+                     "El Faro")) %>%
+  st_transform(crs = proj) %>%
+  mutate(evaluated = ifelse(Name %in% c("El Cabezo",
+                                        "El Faro (Pesca de Langosta)",
+                                        "El Faro"), T, F),
+         Comunidad = ifelse(Name %in% c("El Faro",
+                                        "Laguna Canche Balam",
+                                        "El Faro (Pesca de Langosta)",
+                                        "Anegado de Chal"),
+                            "Punta Herrero",
+                            "Maria Elena"))
+
+# Load Community reserve shapefile
[…]

### 2019-02-28 scripts/content/supp_information.R

@@ -1,0 +1,305 @@
+############################
+##    supp_information    ##
+############################
+
+####################################################
+# Fit the ecological models and create a termplot
+####################################################
+
+
+
+## Load packages
+library(here)
+library(magrittr)
+library(tidyverse)
+
+# Load data
+invert <- read.csv(file = here("data", "invertebrates.csv"),
+                   stringsAsFactors = F) %>% 
+  mutate(zona = ifelse(zona == "Reserva", "Reserve", zona))
+
+fish <- read.csv(file = here("data", "fish.csv"),
+                 stringsAsFactors = F) %>% 
+  mutate(zona = ifelse(zona == "Reserva", "Reserve", zona))
+
+conapesca <- read.csv(file = here("data", "conapesca.csv"),
+                      stringsAsFactors = F) %>% 
+  mutate(zona = ifelse(zona == "Reserva", "Reserve", zona))
+
+# Table of sampling effort for invertebrates
+invert %>% 
+  group_by(ano, comunidad, zona, sitio, transecto) %>%
+  count() %>% 
+  ungroup() %>% 
+  group_by(comunidad, zona) %>%
+  count() %>% 
+  ungroup() %>% 
+  spread(zona, nn) %>% 
+  mutate(age = c(10, 4, 4)) %>% 
+  knitr::kable(col.names = c("Community", "Control", "Reserve", "Years of monitoring"),
+               booktabs = T,
+               format = "latex") %>% 
+  cat(file = here("docs", "tab", "S1_table.tex"))
+
+# Time series of lobster densities. Bars indicate standard errors
+lobster_ts_plot <- invert %>%
+  mutate(generoespecie = ifelse(
+    generoespecie %in% c("Panulirus interruptus",
+                         "Panulirus argus"),
+    generoespecie,
+    "any_invert")) %>% 
+  group_by(rc, ano, year, comunidad, sitio, zona, transecto, generoespecie) %>% 
+  summarize(abundancia = sum(abundancia)) %>% 
+  spread(generoespecie, abundancia, fill = 0) %>% 
+  gather(generoespecie, abundancia, -c(rc, ano, year, comunidad, sitio, zona, transecto)) %>% 
+  filter(generoespecie %in% c("Panulirus interruptus","Panulirus argus")) %>% 
[…]

### 2019-02-28 scripts/content/tab_indicators.R

@@ -1,0 +1,34 @@
+######################
+##  tab_indicators  ##
+######################
+
+###################################
+# Create a table of indicators used
+###################################
+
+## Load packages
+library(here)
+library(knitr)
+library(kableExtra)
+library(tidyverse)
+
+# Create a dataframe and export as tex table
+data.frame(Indicator = c("Lobster density",
+                         "Invertebrate density",
+                         "Fish density",
+                         "Fish biomass",
+                         "Income from target species",
+                         "Landings from target species"),
+           Units = c("org $\\mathrm{m}^{-2}$",
+                     "org $\\mathrm{m}^{-2}$",
+                     "org $\\mathrm{m}^{-2}$",
+                     "Kg $\\mathrm{m}^{-2}$",
+                     "M MXP",
+                     "Metric Tonnes")) %>% 
+  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.",
+               format = "latex",
+               escape = F) %>% 
+  kable_styling(latex_options = "HOLD_position") %>% 
+  group_rows(group_label = "Biological", start_row = 1, end_row = 4) %>% 
+  group_rows(group_label = "Socioeconomic", start_row = 5, end_row = 6) %>% 
+  cat(file = here("docs", "tab", "indicators.tex"))

### 2019-02-28 scripts/content/tab_ses.R

@@ -1,0 +1,28 @@
+###############
+##  tab_ses  ##
+###############
+
+######################################
+# Create a table with our SES analysis
+######################################
+
+## Load packages
+library(here)
+library(knitr)
+library(kableExtra)
+library(tidyverse)
+
+# Read data and create table as tex file
+read.csv(here::here("raw_data", "ses.csv")) %>% 
+  select(-Dim) %>% 
+  knitr::kable(caption = "\\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero). Alphanumeric codes follow \\cite{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \\cite{difranco_2016-Xw} and \\cite{edgar_2014-UO}. The presented narrative applies equally for all communities unless otherwise noted.",
+               format = "latex",
+               escape = F) %>% 
+  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>% 
+  column_spec(column = 1, width = "6.5cm") %>% 
+  column_spec(column = 2, width = "12.2cm") %>% 
+  group_rows(group_label = "Resource System (RS)", start_row = 1, end_row = 5) %>% 
+  group_rows(group_label = "Resource Unit (RU)", start_row = 6, end_row = 7) %>% 
+  group_rows(group_label = "Actors (A)", start_row = 8, end_row = 9) %>% 
+  group_rows(group_label = "Governance system (G)", start_row = 10, end_row = 13) %>% 
+  cat(file = here("docs", "tab","ses.tex"))

### 2019-02-28 scripts/functions/abundance.R

@@ -1,0 +1,18 @@
+abundance <- function(x, com, spp = NULL){
+  
+  if(!is.null(spp)){
+    x <- x %>% 
+      filter(generoespecie == spp)
+  }
+  
+  results <- x %>% 
+    filter(comunidad == com) %>%
+    mutate(ano = as.factor(ano),
+           year = as.factor(year),
+           year = fct_relevel(year, "0")) %>% 
+    group_by(rc, ano, year, sitio, zona, transecto) %>% 
+    summarize(indicador = sum(abundancia)/60) %>% 
+    ungroup()
+    
+    return(results)
+}

### 2019-02-28 scripts/functions/biomass.R

@@ -1,0 +1,13 @@
+biomass <- function(x, com){
+  results <- x %>% 
+    filter(comunidad == com) %>% 
+    mutate(ano = as.factor(ano),
+           year = as.factor(year),
+           year = fct_relevel(year, "0"),
+           biomass = abundancia * a * (talla^b)/1000) %>% 
+    group_by(rc, ano, year, sitio, zona, transecto) %>% 
+    summarize(indicador = sum(biomass, na.rm = T)/60) %>% 
+    ungroup()
+  
+  return(results)
+}

### 2019-02-28 scripts/functions/plot_model.R

@@ -1,0 +1,41 @@
+plot_model <- function(model, model_names = NULL, legend = F){
+  
+  extract_and_filter <- function(model, name){
+    model %>% 
+      robust_se() %>% 
+      filter(grepl(pattern = "zonaReserva:year", x = term)) %>% 
+      mutate(Ano = as.numeric(gsub(pattern = "zonaReserva:year",
+                                   replacement = "",
+                                   x = term)),
+             Community = name)
+  }
+  
+  if(is.null(model_names)){model_names <- letters[1:length(model)]}
+  
+  pd <- position_dodge(width = 0.4)
+  
+  p <- purrr::map2_df(model, model_names, extract_and_filter) %>% 
+    ggplot(aes(x = Ano, y = estimate, color = Community)) +
+    geom_errorbar(aes(ymin = estimate - std.error,
+                      ymax = estimate + std.error),
+                  width = 0,
+                  size = 1,
+                  position = pd) +
+    geom_point(aes(shape = Community), size = 4, position = pd, alpha = 0.5) +
+    geom_hline(yintercept = 0, linetype = "dashed") +
+    scale_fill_brewer(palette = "Set1") +
+    theme(text = element_text(size = 12)) +
+    labs(x = "Years since implementation", y = quo(lambda[t])) +
+    scale_color_brewer(guide = guide_legend(title.position = "top", nrow = 2), palette = "Set1") +
+    scale_x_continuous(breaks = seq(-1, 10, by = 1))
+  
+  if(!legend){
+    p <- p +
+      theme(legend.position = "None")
+  }
+  
+  return(p)
+}
+
+
+

### 2019-02-28 scripts/plot_model.R

@@ -1,37 +1,0 @@
-plot_model <- function(model, model_names = NULL, legend = F){
-  
-  extract_and_filter <- function(model, name){
-    model %>% 
-      robust_se() %>% 
-      filter(grepl(pattern = "zonaReserva:year", x = term)) %>% 
-      mutate(Ano = as.numeric(gsub(pattern = "zonaReserva:year",
-                                   replacement = "",
-                                   x = term)),
-             Community = name)
-  }
-  
-  if(is.null(model_names)){model_names <- letters[1:length(model)]}
-  
-  pd <- position_dodge(width = 0.4)
-  
-  p <- purrr::map2_df(model, model_names, extract_and_filter) %>% 
-    ggplot(aes(x = Ano, y = estimate, color = Community)) +
-    geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0, size = 1, position = pd) +
-    geom_point(aes(shape = Community), size = 4, position = pd, alpha = 0.5) +
-    geom_hline(yintercept = 0, linetype = "dashed") +
-    scale_fill_brewer(palette = "Set1") +
-    theme(text = element_text(size = 12)) +
-    labs(x = "Years since implementation", y = quo(lambda[t])) +
-    scale_color_brewer(guide = guide_legend(title.position = "top", nrow = 2), palette = "Set1") +
-    scale_x_continuous(breaks = seq(-1, 10, by = 1))
-  
-  if(!legend){
-    p <- p +
-      theme(legend.position = "None")
-  }
-  
-  return(p)
-}
-
-
-

### 2019-03-18 docs/PLoS_manuscript.tex
Updated results in MS and scripts

@@
 Although in theory these systems are expected to be successful \cite{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success.
-Moreover, TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \cite{ostrom_2009,gelcich_2015}.
-It is therefore important to consider not only the indicators of interest, but also the governance settings under which the reserves operate.
 
@@
 First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries.
-Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness.
-We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.
+Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness of these reserves.
+%We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.
 
@@
 Community-based marine reserves that are implemented within TURFs are a form of TURF-reserve, voluntarily established and enforced by local communities.
+Community-based spatial closures occur elsewhere, like the \emph{kapu} or \emph{ra’ui} areas in the Pacific Islands \cite{johannes_2002,bohnsack_2004}.
 This bottom-up approach can increase compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \cite{beger_2004,smallhornwest_2018}.
-Community-based spatial closures occur in different contexts, like the \emph{kapu} or \emph{ra’ui} areas in the Pacific Islands \cite{johannes_2002,bohnsack_2004}.
 However, community-based reserves can be hard to enforce if they are not legally recognized.
@@
 The profits from each TURF are shared amongst all fishers from the cooperative.
-Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (\emph{i.e.} TURF-reserves).
-Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges after conducting a series of studies outlined in the regulation, namely the creation of a baseline of ecological data and historical performance of the fishery \cite{nom}.
-
-\subsection*{Study areas}
-
-We evaluate three community-based no-take TURF-reserve systems implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig~\ref{fig:map}A).
-The first one was created by the \emph{Buzos y Pescadores de la Baja California} fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig~\ref{fig:map}B).
-The main fishery in the island is the spiny lobster (\emph{Panulirus interruptus}), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income.
-In 2006, the community decided to implement two marine reserves within their fishing grounds.
-The objective of these reserves was "to protect and recover stocks of commercially important invertebrate species"; mainly lobster and abalone.
-The reserves obtained legal recognition in 2018 \cite{dof_website_2018}.
+Fishing cooperatives interested in implementing marine reserves within their TURFs (\emph{i.e.} TURF-reserves) work with CSOs to design them.
+Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges under the 2012 regulation \cite{nom}.
 
@@
 \end{figure}
+
+\subsection*{Study areas}
+
+We evaluate three community-based no-take TURF-reserve systems implemented in Mexican TURF-managed fisheries (Fig~\ref{fig:map}A).
+The first one was created by the \emph{Buzos y Pescadores de la Baja California} fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig~\ref{fig:map}B).
+The main fishery in the island is the spiny lobster (\emph{Panulirus interruptus}), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income.
+In 2006, the community decided to implement two marine reserves within their fishing grounds.
+The objective of these reserves was "to protect and recover stocks of commercially important invertebrate species"; mainly lobster and abalone.
+The reserves obtained legal recognition in 2018 \cite{dof_website_2018}.
 
@@
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
-This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \cite{stewartoaten_1986,depalma_2018} and causally attribute the changes to the reserve \cite{francinifilho_2008,Villasenor-Derbez_2018}.
+This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \cite{stewartoaten_1986,depalma_2018,kerr_2019} and causally attribute the changes to the reserve \cite{francinifilho_2008,Villasenor-Derbez_2018}.
 
 The biological data are collected by members from each community and personnel from the Mexican CSO \emph{Comunidad y Biodiversidad} (\href{https://www.cobi.org.mx}{COBI}).
-Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites, where a minimum of 4 transects per site are performed \cite{suman_2010-ez,fulton_2018,fulton_2019}.
+Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \cite{suman_2010-ez,fulton_2018,fulton_2019}.
 Size structures are also collected during fish surveys.
 All sites were surveyed annually, and at least once before implementation of the reserves.
-A summary of sampling effort is shown in the supplementary materials (\nameref{S1_Table}, \nameref{S2_Table}).
+A summary of sampling effort and time series are shown in the supplementary materials (\nameref{S1_Table}, \nameref{S2_Table}, \nameref{S1_Fig}, \nameref{S2_Fig}, \nameref{S3_Fig}, \nameref{S4_Fig}).
 
@@
 Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves.
-In this case our treated unit are the cooperatives (\emph{i.e.} communities) that have implemented a reserve within their TURF, and the controls are nearby communities that have a TURF but did not implement a reserve.
-These communities have TURFs adjacent to the TURFs of our focal communities.
+In this case our treated unit are the cooperatives (\emph{i.e.} communities) that have implemented a reserve within their TURF, and the controls are adjacent communities that have a TURF but did not implement a reserve.
 Cooperatives incorporated in this analysis have similar number of members, belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \cite{mccay_2014,mccay_2017,ayer_2018}.
 Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \cite{oecd_2017}.
-A table with summary statistics for this data is provided in the supplementary materials (Table S3, Fig S5).
+A table with summary statistics for this data is provided in the supplementary materials (\nameref{S3_Table},\nameref{S5_Fig}).
 
-Data for the evaluation of the SES were collected at the community-level from official documents used in the design, creation, and designation of the marine reserves.
+Data for the evaluation of the SES were collected at the community-level from official documents used in the design, creation, and implementation of the marine reserves.
 These include the technical studies that the cooperatives submit when they request recognition of their reserves, as well as the official enactments  \cite{dof_website_2012,dof_website_2013,dof_website_2018}.
@@
 This approach is widely used in econometric literature to estimate the average treatment effect of an intervention, like the impact of minimum wage increases on employment rates \cite{card_1994}.
-In our case it allows us to estimate the effect that the reserve had on each biological and socioeconomic indicator (Table~\ref{table:indicators}) by comparing trends across time and treatments since reserve implementation \cite{moland_2013,Villasenor-Derbez_2018}.
+In our case it allows us to estimate the effect that the reserve had on each biological and socioeconomic indicator (Table~\ref{table:indicators}) by comparing trends across time and treatments since reserve implementation \cite{moland_2013,Villasenor-Derbez_2018, kerr_2019}.
 To perform difference-in-differences, we regress the indicator of interest on a dummy variable for treatment, a dummy variable for years, and the interaction term between these with a multiple linear regression of the form:
@@
 These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \cite{depalma_2018}.
-For each indicator in each community, we fit one model (\emph{e.g.} there are three models for lobster density, one for each community) for a total of 12 biological model fits and four socioeconomic model fits.
-Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard errors \cite{zeileis_2004-7n}.
+Since we are interested in the effectiveness of each reserve system, we fit one model for each indicator in each community (\emph{e.g.} there are three models for lobster density, one for each community).
+This gives us a total of 12 biological model fits and four socioeconomic model fits.
+Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard error correction \cite{zeileis_2004-7n}.
 All analyses were performed in R 3.5.2 and R Studio version 1.1.456 \cite{R_2018}.
@@
 
-We use the SES framework to evaluate each community and create a narrative that provides context for each community.
+TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \cite{ostrom_2009,gelcich_2015}.
+When evaluating the effects of TURF-reserves, it is important to consider not only the indicators of interest, but also the governance settings under which a reserve operates.
+We use the SES framework to qualitatively evaluate each community and create a narrative that provides context for each of them.
 The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators.
-Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as in \cite{leslie_2015-na}), we followed a similar approach to \cite{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems.
-We based our variable selection primarily on \cite{leslie_2015-na} and \cite{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework, and identified the key variables relevant to fishing cooperatives in Mexico.
-We also incorporate other relevant variables known to influence reserve performance following \cite{difranco_2016-Xw} and \cite{edgar_2014-UO}.
+Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as done in \cite{leslie_2015-na}), we followed a similar approach to \cite{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems.
+We based our variable selection primarily on previous analyses of Mexican fishing cooperatives \cite{leslie_2015-na,basurto_2013-oq}.
+We also incorporate other relevant variables known to influence reserve performance following, such as age and size of reserve or isolation of the system \cite{difranco_2016-Xw,edgar_2014-UO}.
 Table~\ref{table:ses} shows the selected variables, along with definitions and values.
@@
 Fig~\ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time.
-In the case of Maria Elena, positive changes were observed in the third and fourth year.
-These effects are in the order of 0.2 extra organisms $\mathrm{m}^{-2}$ for Isla Natividad and Punta Herrero, and 0.01 organisms $\mathrm{m}^{-2}$ for Maria Elena, but are not significantly different from zero ($p > 0.05$).
+In the case of Maria Elena, positive changes were observed in the second and third years.
+These effects were in the order of 0.01 extra organisms $\mathrm{m}^{-2}$, but were only significantly different from zero for Maria Elena ($p < 0.05$).
 Likewise, no significant changes were detected in fish biomass or invertebrate and fish densities (Fig.
@@
 Lastly, the communities exhibit all the social enabling conditions for effective reserve and resource management.
-Here we discuss possible shortcomings in our analyses as well as possible explanations for the observed patterns.
+Here we discuss possible shortcomings in our analyses as well alternative explanations for the observed lack of effectiveness.
 
-While many ecology studies have used BACI sampling designs and respective analyses (\emph{e.g.} \cite{stewartoaten_1986}), few conservation studies have done so to evaluate the effect of an intervention (\emph{e.g.} \cite{francinifilho_2008,lester_2009,moland_2013}) which has resulted in a call for more robust analyses in conservation science \cite{guidetti_2002,ferraro_2006}.
+While many ecology studies have used BACI sampling designs and respective analyses (\emph{e.g.} \cite{stewartoaten_1986}), few conservation studies have done so to evaluate the effect of an intervention (\emph{e.g.} \cite{francinifilho_2008,lester_2009,moland_2013,kerr_2019}) which has resulted in a call for more robust analyses in conservation science \cite{guidetti_2002,ferraro_2006}.
 Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness, and captures previously described patterns.
@@
 \ref{fig:indicators}A), occurs a year after the hypoxia events described by \cite{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
-The use of causal inference techniques may help us support evidence-based conservation.
+
+While the use of causal inference techniques may help us support evidence-based conservation, spatial connectivity between reserve and control sites, stockpiling, and backstopping may confound the results \cite{kerr_2019}.
+Given that we find no clear evidence of reserve effectiveness, one might say that our reserve and control sites are not spatially independent.
+This would imply that the recovery within the reserve quickly results in recovery ouside the protected area.
+However, indicators show little to no temporal variation (\nameref{S1_Fig}, \nameref{S2_Fig}, \nameref{S3_Fig}, \nameref{S4_Fig}), and it is unlikely that this effect would be observed under current reserve designs, as detailed below.
 

### 2019-04-07 docs/PLoS_manuscript.tex
PLoS Submission ready

@@
 
-%% Include all macros below
-
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
-
-%% END MACROS SECTION
-
 \begin{document}
@@
 Stuart Fulton\textsuperscript{3\ddag},
-Álvin Suarez\textsuperscript{3\ddag},
+Alvin Suarez\textsuperscript{3\ddag},
 Arturo Hernández-Velasco\textsuperscript{3\ddag},
@@
 \bigskip
-\textbf{1} Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA
+\textbf{1} Bren School of Environmental Science \& Management, University of California, Santa Barbara, Santa Barbara, CA, USA
 \\
@@
 \\
-\textbf{4} Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
+\textbf{4} Hopkins Marine Station and Stanford Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
 \\
@@
 Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
-In the last 13 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2012; their effectiveness has not been formally evaluated.
+Since 2012, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2012; their effectiveness has not been formally evaluated.
 We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico.
-We find that reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates.
+We find that, overall, reserves have not yet achieved their stated goals of increasing the density of lobster and other benthic invertebrates, nor increasing lobster catches.
 A lack of clear ecological and socioeconomic effects likely results from a combination of factors.
@@
 Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches.
-However, these reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species.
+However, even small reserves are expected to provide benefits for sedentary invertebrates over longer time frames, with continued protection.
+These reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species.
 
@@
 
+\clearpage
+
 \linenumbers
@@
 In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \cite{costello_2012}.
-One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves, thus creating TURF-reserve systems \cite{afflerbach_2014,gelcich_2015,lester_2017}.
+One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves\cite{afflerbach_2014,gelcich_2015,lester_2017}.
 
@@
 First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries.
-Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness of these reserves.
+The second objective is to identify opportunities where improvement or adjustment might lead to increased effectiveness of these reserves.
 %We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.
@@
 The most common setup of community-based TURF-reserves in Mexico is the following.
-Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig~\ref{fig:map}A).
+Fishers from a given community assemble into a fishing cooperative which has exclusive fishing rights over a spatially delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig~\ref{fig:map}A).
 Each TURF is exclusively fished by one cooperative, and each community usually hosts no more than one cooperative.
@@
 Fishing cooperatives interested in implementing marine reserves within their TURFs (\emph{i.e.} TURF-reserves) work with CSOs to design them.
-Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges under the 2012 regulation \cite{nom}.
+Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges, as stated in the 2014 regulation \cite{nom}.
 
@@
 The first one was created by the \emph{Buzos y Pescadores de la Baja California} fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig~\ref{fig:map}B).
-The main fishery in the island is the spiny lobster (\emph{Panulirus interruptus}), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income.
+At present, the main fishery in the island is the spiny lobster (\emph{Panulirus interruptus}), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income.
 In 2006, the community decided to implement two marine reserves within their fishing grounds.
@@
 Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively.
-These reserves have been legally recognized as Fishing Refuges since their original implementation \cite{dof_website_2012,dof_website_2013} and subsequent re-establishments \cite{dof_website_2017b}.
+These reserves have been legally recognized as Fishing Refuges since their original implementation \cite{dof_website_2012,dof_website_2013} and subsequent re-establishments in 2017 \cite{dof_website_2017b}.
 
@@
 We therefore incorporate a reduced version of the Social Ecological Systems framework \cite{ostrom_2009} and evaluate variables and indicators known to aid and hinder the effectiveness of management interventions in conservation and fisheries.
-The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather help provide context on the social-ecological system in which reserves develop.
+The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather to provide context on the social-ecological system in which reserves develop.
 The following two sections describe our data collection methods and analyses.
@@
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
-This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \cite{stewartoaten_1986,depalma_2018,kerr_2019} and causally attribute the changes to the reserve \cite{francinifilho_2008,Villasenor-Derbez_2018}.
+This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that allows us to capture and control for temporal and spatial dynamics and causally attribute the changes to the reserve \cite{stewartoaten_1986,francinifilho_2008,depalma_2018,kerr_2019,Villasenor-Derbez_2018}.
 
@@
 All sites were surveyed annually, and at least once before implementation of the reserves.
-A summary of sampling effort and time series are shown in the supplementary materials (\nameref{S1_Table}, \nameref{S2_Table}, \nameref{S1_Fig}, \nameref{S2_Fig}, \nameref{S3_Fig}, \nameref{S4_Fig}).
+A summary of sampling effort and time series are shown in the supplementary materials (Tables A-B and Figures A-D in \nameref{S1_Text}).
 
@@
 Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \cite{oecd_2017}.
-A table with summary statistics for this data is provided in the supplementary materials (\nameref{S3_Table},\nameref{S5_Fig}).
+A table with summary statistics and time series for this data are provided in the supplementary materials (Table C and Figure E in \nameref{S1_Text}).
 
@@
 
-\input{tab/ses.tex}
+\begin{table}[!ht]
+
+\caption{{\bf Variables for the Social-Ecological System analysis.}}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12.2cm}}
+\hline
+Variable & Narrative\\
+\hline
+\multicolumn{2}{l}{\textbf{Resource System (RS)}}\\
+\hline
+\hspace{1em}RS2 - Clarity of system boundaries: Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates.
+Reserve placement is decided by the community.
+Fishers use GPS units and landmarks.\\
+\hline
+\hspace{1em}RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
+\hline
+\hspace{1em}RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
+\hline
+\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) MSC certified.\\
+\hline
+\hspace{1em}*RS5 - Age of reserves: Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
+\hline
+\multicolumn{2}{l}{\textbf{Resource Unit (RU)}}\\
+\hline
+\hspace{1em}RU1 - Resource unit mobility & Adult spiny lobsters can move between 1 and 10 Km, while larvae can have displacements in the order of hundreds of Km (Aceves-Bueno et al., 2017; Green et al., 2017).\\
+\hline
+\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish (2 spp each).
+Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates (4 spp).\\
+\hline
+\multicolumn{2}{l}{\textbf{Actors (A)}}\\
+\hline
+\hspace{1em}A1 - Number of relevant actors: Number of fishers & IN = 98; ME = 80; PH = 21\\
+\hline
+\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers.
+IN lies 545 Km south from Tijuana, and ME and PH 230 Km south from Cancun, where the nearest international airports are located.\\
+\hline
+\multicolumn{2}{l}{\textbf{Governance system (G)}}\\
+\hline
+\hspace{1em}GS6.1.4.3 - Territorial use communal rights : Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster.
+These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
+\hline
+\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandates.
+These are: larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
+\hline
+\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group (Consejo de vigilancia) that monitors and enforces formal and internal rules.
+They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
+\hline
+\hspace{1em}GS9.2 - Biophysical monitoring: Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds.
+Recently, they have installed oceanographic sensors to monitor oceanographic variables.\\
+\hline
+GS10.1 - Graduated sanctions & Fishers have penalties for breaking collective-choice rules or fishing inside the reserves.
+These may range from scoldings and warnings to not being allowed to harvest a particular resource or being expelled from the cooperative.\\
+\hline
+\end{tabular}}
+\label{table:ses}
+IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero.
+Alphanumeric codes for variables follow \cite{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \cite{difranco_2016-Xw} and \cite{edgar_2014-UO}.
+The presented narrative applies equally for all communities unless otherwise noted.
+\end{table}
 
@@
 
-\input{tab/indicators.tex}
+\begin{table}[!ht]
+
+\caption{
+{\bf List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}}
+\centering
+\begin{tabular}{l|l}
+\hline
+Indicator & Units\\
+\hline
+\multicolumn{2}{l}{\textbf{Biological}}\\
+\hline
+\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
+\hline
+\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
+\hline
+\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
+\hline
+\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
+\hline
+\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
+\hline
+\hspace{1em}Income from target species & M MXP\\
+\hline
+\hspace{1em}Landings from target species & Metric Tonnes\\
+\hline
+\end{tabular}
+\label{table:indicators}
+\end{table}
 
@@
 Where year-level fixed effects capturing a temporal trend are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites.
-The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (\emph{i.e.} $\gamma_t$ and $\beta$ respectively).
+The effect of the reserve is captured by the $\lambda_t$ vector of coefficients, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (\emph{i.e.} $\gamma_t$ and $\beta$ respectively).
 Therefore, we would expect this term to be positive if the indicator increases because of the reserve.
@@
 
-The coefficient interpretations remains as for Eq.
-\ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$).
+The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$).
+
 These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \cite{depalma_2018}.
@@
 We based our variable selection primarily on previous analyses of Mexican fishing cooperatives \cite{leslie_2015-na,basurto_2013-oq}.
-We also incorporate other relevant variables known to influence reserve performance following, such as age and size of reserve or isolation of the system \cite{difranco_2016-Xw,edgar_2014-UO}.
+We also incorporate other relevant variables known to influence reserve performance, such as age and size of reserve or isolation of the system \cite{difranco_2016-Xw,edgar_2014-UO}.
 Table~\ref{table:ses} shows the selected variables, along with definitions and values.
@@
 Likewise, no significant changes were detected in fish biomass or invertebrate and fish densities (Fig.
-\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends.
-Figures and tables with time series of indicators and model coefficients are presented in the supplementary materials (\nameref{S1_Fig}, \nameref{S2_Fig}, \nameref{S3_Fig}, \nameref{S4_Fig}, \nameref{S4_Table}, \nameref{S5_Table}, \nameref{S6_Table}).
+\ref{fig:indicators}B-D).
+Invertebrate and fish densities showed positive trends in Natividad in most years.
+Reserves in Maria Elena and Punta Herrero showed positive trends for invertebrates in some years.
+However, these were not statistically significant.
+Figures with time series of indicators and tables with model coefficients and main effects are presented in the supplementary materials (Figures A-D and Tables D-F in \nameref{S1_Text}).
 
-\begin{figure}[!h]
+\begin{figure}[h]
 \centering
@@
 The year of post-implementation data for Maria Elena does not show a significant effect of the reserve.
-Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities.
-However, these changes are only significant for the third year ($p < 0.05$).
-Full tables with model coefficients are presented in the supplementary materials (\nameref{S7_Table}, \nameref{S8_Table}).
+Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities, though these changes are only significant for the third year ($p < 0.05$).
+Full tables with model coefficients are presented in the supplementary materials (Tables G-H in \nameref{S1_Text}).
 
@@
 For example, fishers operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves.
-Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have solid governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management.
+Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have sotrong governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management.
 In general, success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources upon which they depend \cite{jupiter_2017}.
@@
 Our results indicate that these TURF-reserves have not increased lobster densities.
-Additionally, no co-benefits were identified when using other ecological indicators aside from the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \cite{micheli_2012-EU}.
+Additionally, no co-benefits were identified when using other ecological indicators aside from the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad, as well as positive trends in invertebrate and fish densities \cite{micheli_2012-EU}.
 The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
@@
 Given that we find no clear evidence of reserve effectiveness, one might say that our reserve and control sites are not spatially independent.
-This would imply that the recovery within the reserve quickly results in recovery ouside the protected area.
-However, indicators show little to no temporal variation (\nameref{S1_Fig}, \nameref{S2_Fig}, \nameref{S3_Fig}, \nameref{S4_Fig}), and it is unlikely that this effect would be observed under current reserve designs, as detailed below.
+This would imply that the recovery within the reserve quickly results in recovery outside the protected area.
+However, indicators show little to no temporal variation (Figures A-D in \nameref{S1_Text}), and it is unlikely that this effect would be observed under current reserve designs, as detailed below.
 
@@
 With these characteristics, one would expect the reserves to be effective.
-Maria Elena and Punta Herrero are relatively young reserves (\emph{i.e.} < 6 years old; RS5 in Table~\ref{table:ses}) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \cite{dasilva_2015-zX}.
+In fact, the oldest reserves shows some positive trends for invertebrate and fish densities, as well as income.
+Maria Elena and Punta Herrero are relatively young reserves (\emph{i.e.} \textless 6 years old; RS5 in Table~\ref{table:ses}) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \cite{dasilva_2015-zX}.
 
 Another key condition for effectiveness is reserve size \cite{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (\emph{sensu} \cite{rees_2018}).
-Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \cite{rossetto_2015-V0}, however, abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary invertebrates, but not lobsters.
+Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \cite{rossetto_2015-V0}, however, abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary molusks, but not lobsters.
 Design principles developed by \cite{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
-As shown through the SES analysis, the size of the marine reserves appears small compared to the movement capacity of the main targeted species (RU1, RS3; Table~\ref{table:ses}).
+As shown in the SES analysis, the size of the marine reserves appears small compared to the movement capacity of the main targeted species (RU1, RS3; Table~\ref{table:ses}).
 Furthermore, fishers may favor implementation of reserves that pose low fishing costs due to their small size or location.
@@
 Both lobster fisheries were certified by the Marine Stewardship Council and are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \cite{dof_website_1993}.
-It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves; reserves implemented in TURFs where fishing pressure is already optimally managed will still show a trade-off between fisheries and conservation objectives \cite{lester_2017}.
+It is uncertain whether such a well-managed fishery will experience additional benefits from small marine reserves; reserves implemented in TURFs where fishing pressure is already optimally managed will still show a trade-off between fisheries and conservation objectives \cite{lester_2017}.
+In contrast, invertebrate fisheries have seen declining catches and finfish fisheries are not managed under TURFs, which may explain the positive trends observed.
 Furthermore, \cite{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open access.
@@
 
-While the evaluated reserves have failed to provide fishery benefits to date, there are a number of additional ecological, fisheries, and social benefits.
+While the evaluated reserves have failed to provide clear fishery benefits to date, there are a number of additional ecological, fisheries, and social benefits.
 Marine reserves provide protection to a wider range of species and vulnerable habitat.
@@
 % Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
-\paragraph*{S1 Table.}
-\label{S1_Table}
-{\bf Invertebrate sampling effort.} Number of invertebrate transects performed in each site of each community.
-
-\paragraph*{S1 Fig.}
-\label{S1_Fig}
-{\bf Mean annual lobster density.} Dots indicate mean, bars indicate standard errors.
-
-\paragraph*{S2 Fig.}
-\label{S2_Fig}
-{\bf Mean annual invertebrate density.} Dots indicate mean, bars indicate standard errors.
-
-\paragraph*{S2 Table.}
-\label{S2_Table}
-{\bf Fish sampling effort} Number of invertebrate transects performed in each site of each community.
-
-\paragraph*{S3 Fig.}
-\label{S3_Fig}
-{\bf Mean annual fish biomass} Dots indicate mean, bars indicate standard errors.
-
-\paragraph*{S4 Fig.}
-\label{S4_Fig}
-{\bf Mean annual fish density} Dots indicate mean, bars indicate standard errors.
-
-\paragraph*{S3 Table.}
-\label{S3_Table}
-{\bf Summary of socioeconomic data} Mean ex-vessel prices (MXP / Kg) for lobster (\emph{Panulirus spp.}) on each TURF (community) and group (Control = only TURF, Treated = TURF and reserve).
-
-\paragraph*{S5 Fig.}
-\label{S5_Fig}
-{\bf Timeseries of socioeconomic indicators.} A) Mean annual value of landings and B) Mean landings.
-
-\paragraph*{S4 Table.}
-\label{S4_Table}
-{\bf Coefficient estimates of biological indicators for Isla Natividad.}
-
-\paragraph*{S5 Table.}
-\label{S5_Table}
-{\bf Coefficient estimates of biological indicators for Maria Elena.}
-
-\paragraph*{S6 Table.}
-\label{S6_Table}
-{\bf Coefficient estimates of biological indicators for Punta Herrero.}
-
-\paragraph*{S7 Table.}
-\label{S7_Table}
-{\bf Coefficient estimates of socioeconomic indicators Isla Natividad.}
-
-\paragraph*{S8 Table.}
-\label{S8_Table}
-{\bf Coefficient estimates of socioeconomic indicators Maria Elena.}
+\paragraph*{S1 Text}
+\label{S1_Text}
+{\bf Supplementary Text.} Additional figures and tables with summary information.
+Table A with invertebrate sampling effort.
+Table B with fish sampling effort.
+Table C with summary of socioeconomic data.
+Table D with coefficient estimates for biological indicators in Isla Natividad.
+Table E with coefficient estimates for biological indicators in Maria Elena.
+Table F with coefficient estimates for biological indicators in Punta Herrero.
+Table G with coefficient estimates of socioeconomic indicators in Isla Natividad.
+Table H with coefficient estimates of socioeconomic indicators in Maria Elena.
+Figure A with time series of mean annual lobster density for each community.
+Figure B with mean annual invertebrate densities for each community.
+Figure C with mean annual fish biomass for each community.
+Figure D with mean annual fish density for each community.
+Figure E with time series of socioeconomic indicators.
 
@@
 This study would have not been possible without the effort by members of the fishing communities here mentioned, who participated in the data-collection process.
-The authors wish to acknowledge comments by the reviewers and editor, which significantly improved the quality of this work.
-
-%
-%% For figure citations, please use "Fig" instead of "Figure".
-%Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero.
-%
-%% Place figure captions after the first paragraph in which they are cited.
-%\begin{figure}[!h]
-%\caption{{\bf Bold the figure title.}
-%Fig caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
-%\label{fig1}
-%\end{figure}
-%
-%% Results and Discussion can be combined.
-%\section*{Results}
-%Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero.
-%
-%% Place tables after the first paragraph in which they are cited.
-%\begin{table}[!ht]
-%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
-%\centering
-%\caption{
-%{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
-%\begin{tabular}{|l+l|l|l|l|l|l|l|}
-%\hline
-%\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
-%$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
-%$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
-%$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
-%\end{tabular}
-%\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
-Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
-%\end{flushleft}
-%\label{table1}
-%\end{adjustwidth}
-%\end{table}
-%
-%
-%%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
-%\subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
-%\subsubsection*{3rd level heading} 
-%Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
-Sed non aliquet felis.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Mauris commodo justo ac dui pretium imperdiet.
-Sed suscipit iaculis mi at feugiat.
-%
-%\begin{enumerate}
-%	\item{react}
-%	\item{diffuse free particles}
-%	\item{increment time by dt and go to 1}
-%\end{enumerate}
-%
-%
-%\subsection*{Sed ac quam id nisi malesuada congue}
-%
-%Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero.
-%
-%\begin{itemize}
-%	\item First bulleted item.
-%	\item Second bulleted item.
-%	\item Third bulleted item.
-%\end{itemize}
-%
-%\section*{Discussion}
-%Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
-Proin rutrum vel massa non gravida.
-Quisque tempor sem et dignissim rutrum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Morbi at justo vitae nulla elementum commodo eu id massa.
-In vitae diam ac augue semper tincidunt eu ut eros.
-Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
-Aliquam in enim semper, aliquam massa id, cursus neque.
-Praesent faucibus semper libero~\cite{bib3}.
-%
-%\section*{Conclusion}
-%
-%CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
-%
-%Sed non aliquet felis.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Mauris commodo justo ac dui pretium imperdiet.
-Sed suscipit iaculis mi at feugiat.
-Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna.
-Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
-Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
-Nam condimentum sem eget mollis euismod.
-Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex.
-Nunc est dui, sodales sed mauris nec, auctor sagittis leo.
-Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor.
-For more information, see \nameref{S1_Appendix}.
-%
 

### 2019-04-07 scripts/content/fig_bio_results.R

@@ -141,6 +141,12 @@ ggsave(plot = plot,
        width = 6.5,
        height = 4)
 
+# Export figure
+ggsave(plot = plot,
+       filename = here("docs", "img", "fig_bio_results.tiff"),
+       width = 6.5,
+       height = 4)
+
 # Save all model output
 save(L_IN,
      L_ME,

### 2019-04-07 scripts/content/fig_socioeco_results.R

@@ -14,6 +14,7 @@ library(here)
 library(tidyverse)
 
 ## Load custom functions
+source(file = here("scripts", "functions", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
 source(file = here("scripts", "functions", "plot_socioeco.R")) #plots effect sizes for Socioeco
 
 # Load data
@@ -66,6 +67,11 @@ plot <- plot_grid(c_plot, v_plot, ncol = 1, labels = "AUTO")
 # Export figures
 ggsave(plot = plot,
        filename = here("docs", "img", "fig_socioeco_results.pdf"),
+       width = 6.5,
+       height = 4)
+
+ggsave(plot = plot,
+       filename = here("docs", "img", "fig_socioeco_results.tiff"),
        width = 6.5,
        height = 4)
 

### 2019-04-07 scripts/content/img_map.R

@@ -150,7 +150,13 @@ left <- plot_grid(mex, pbc, ncol = 1, labels = "AUTO")
 
 plot1 <- plot_grid(left, ph, labels = c("", "C"), ncol = 2)
 
+# Save map
 ggsave(plot1,
        filename = here("docs", "img", "fig_map.pdf"),
+       width = 6.5,
+       height = 4.5)
+
+ggsave(plot1,
+       filename = here("docs", "img", "fig_map.tiff"),
        width = 6.5,
        height = 4.5)

### 2019-04-07 scripts/content/supp_information.R

@@ -38,7 +38,8 @@ invert %>%
   mutate(age = c(10, 4, 4)) %>% 
   knitr::kable(col.names = c("Community", "Control", "Reserve", "Years of monitoring"),
                booktabs = T,
-               format = "latex") %>% 
+               format = "latex",
+               caption = "{\\bf Invertebrate sampling effort.} Number of invertebrate transects performed in each site of each community.") %>% 
   cat(file = here("docs", "tab", "S1_table.tex"))
 
 # Time series of lobster densities. Bars indicate standard errors
@@ -77,6 +78,11 @@ ggsave(plot = lobster_ts_plot,
        width = 6,
        height = 4.5)
 
+ggsave(plot = lobster_ts_plot,
+       file = here("docs", "img", "S1_fig.tiff"),
+       width = 6,
+       height = 4.5)
+
 
 # Time series of invertebrate densities. Bars indicate standard errors
 invert_ts_plot <- invert %>%
@@ -105,6 +111,11 @@ ggsave(plot = invert_ts_plot,
        width = 6,
        height = 4.5)
 
+ggsave(plot = invert_ts_plot,
+       file = here("docs", "img", "S2_fig.tiff"),
+       width = 6,
+       height = 4.5)
+
 
 # Table of sampling effort for fish
 fish %>% 
@@ -118,7 +129,8 @@ fish %>%
   mutate(age = c(10, 4, 4)) %>% 
   knitr::kable(col.names = c("Community", "Control", "Reserve", "Years of monitoring"),
                booktabs = T,
-               format = "latex") %>% 
+               format = "latex",
+               caption = "{\\bf Fish sampling effort.} Number of invertebrate transects performed in each site of each community.") %>% 
   cat(file = here("docs", "tab", "S2_table.tex"))
 
 #Time series of fish biomass. Bars indicate standard errors
@@ -151,6 +163,11 @@ ggsave(plot = fish_biomass_ts_plot,
        width = 6,
        height = 4.5)
 
+ggsave(plot = fish_biomass_ts_plot,
+       file = here("docs", "img", "S3_fig.tiff"),
+       width = 6,
+       height = 4.5)
+
 
 #Time series of fish densities. Bars indicate standard errors
 fish_density_ts_plot <- fish %>%
[…]

### 2019-05-10 docs/PLoS_manuscript.tex
incorporated reviewers comments

@@
 {\Large
-\textbf\newline{Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Community-based TURF-reserves in small-scale fisheries} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \cite{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}.
-In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \cite{costello_2012}.
+In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce, given the heterogeneity in target species and number of actors involved \cite{costello_2012}.
 One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves\cite{afflerbach_2014,gelcich_2015,lester_2017}.
@@
 
-Conservation science has shown how marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \cite{lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+Conservation science has shown how well implemented and enforced marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \cite{lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
 Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \cite{gelcich_2008,gelcich_2012}.
@@
 We use three main sources of information to evaluate these reserves across ecological, socioeconomic, and governance dimensions.
-Ecological data come from the annual ecological monitoring of reserve and control sites.
+Biological data come from the annual ecological monitoring of reserve and control sites.
 Reserve sites are areas where no fishing occurs.
@@
 
-Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de Acuacultura y Pesca}; CONAPESCA).
-Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves.
+Socioeconomic data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves.
+These were requested from the National Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de Acuacultura y Pesca}; CONAPESCA) via the access to information act.
 In this case our treated unit are the cooperatives (\emph{i.e.} communities) that have implemented a reserve within their TURF, and the controls are adjacent communities that have a TURF but did not implement a reserve.
@@
 
-Data for the evaluation of the SES were collected at the community-level from official documents used in the design, creation, and implementation of the marine reserves.
+Data for the evaluation of the SES were collected at the community-level, and focused on the Resource Systems, Resource Units, Actors, and Governance System (Table~\ref{table:ses}).
+Data come from official documents used in the design, creation, and implementation of the marine reserves.
 These include the technical studies that the cooperatives submit when they request recognition of their reserves, as well as the official enactments \cite{dof_website_2012,dof_website_2013,dof_website_2018}.
-We also complimented information based on the authors' experience and knowledge of the communities.
-We collected information on the Resource Systems, Resource Units, Actors, and Governance System (Table~\ref{table:ses}).
+We also complimented information based on the authors' experience and knowledge of the communities (RS5, RU5, A3, GS6.2, GS99.1, GS9.2, and GS10.1).
 
@@
 Colors and shapes denote communities: Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares).
-Letters denote indicator: A) lobster densities (\emph{Panulirus spp}, B) fish biomass, C) invertebrate densities, and D) fish densities.
-Plots are ordered by survey type (left column: invertebrates; right column: fish).
 Points are jittered hotizontally to avoid overplotting.}
@@
 \caption{{\bf Effect sizes for socioeconomic indicators}
-Points indicate the effect size and error bars are heteroskedastic-robust standard errors for: lobster catches (A) and revenues (B) at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles).
+Points indicate the effect size and error bars are heteroskedastic-robust standard errors for lobster catches and revenues at Isla Natividad (IN; red circles) and Maria Elena (ME; blue triangles).
 Years have been centered to year of implementation.
@@
 
-Our qualitative implementation of the social-ecological systems framework allowed us to systematically identify important differences between the case studies’ governance systems and incorporate other characteristics of these fisheries neglected during the process of data collection (Table~\ref{table:ses}).
 We find that the analyzed communities share similarities known to foster sustainable resource management and increase reserve effectiveness.
@@
 The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
-Lastly, the communities exhibit all the social enabling conditions for effective reserve and resource management.
+Our qualitative implementation of the social-ecological systems framework allowed us to systematically identify important differences between the case studies’ governance systems and incorporate other characteristics of these fisheries neglected during the process of data collection (Table~\ref{table:ses}).
+These communities exhibit all the social enabling conditions for effective reserve and resource management.
 Here we discuss possible shortcomings in our analyses as well alternative explanations for the observed lack of effectiveness.
@@
 
-\section*{Conclusion}
-
 While the evaluated reserves have failed to provide clear fishery benefits to date, there are a number of additional ecological, fisheries, and social benefits.
@@
 
-Community-based TURF-reserves in small-scale fisheries may be helpful conservation and fishery management tools when appropriately implemented \cite{gelcich_2015}.
-We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside.
-Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.
-Furthermore, conservation and advocacy groups should consider the opportunity costs of such interventions (\emph{sensu} \cite{smith_2010}) and evaluate the potential of other approaches that may yield similar benefits.
-
 In terms of fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs.
@@
 \cite{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.
+
+\section*{Conclusion}
+
+Our objectives were to evaluate the effectiveness of community-based marine reserves implemented under a new regulation, and learning from this process to inform management of these reserves as well as similar processes elsewhere.
+We do not find clear evidence of an increase in lobster densities or catches, and implementation of reserves did not come at a cost (\emph{i.e.} reduction in catches or revenues).
+After identifying this lack of effectiveness, we used the SES framework to look for alternative explanations.
+The communities seem to exhibit all social and governance conditions that would result in reserve effectiveness.
+However, these reserves may not be large enough to provide the necesary protection that would lead to increases in biomass.
 For the particular case of the reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
+
+With other projects in mind, bottom-up design and implementation processes like the ones in the evaluated reserves must be promoted.
+However, these should not come at the cost of setting design principles aside.
+Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness.
+Furthermore, conservation and advocacy groups should consider the opportunity costs of such interventions (\emph{sensu} \cite{smith_2010}) and evaluate the potential of other approaches and alternative investments that may yield similar benefits.
 
@@
 This study would have not been possible without the effort by members of the fishing communities here mentioned, who participated in the data-collection process.
+The authors would like to thank comments by two reviewrers that significantly improved the quality of our work.
 

### 2019-05-10 scripts/content/fig_bio_results.R

@@ -110,16 +110,19 @@ N_PH <- fish %>%
 # Invert plots ##################################################
 ## Lobsters
 L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
-                     model_names = c("IS", "ME", "PH"))
+                     model_names = c("IS", "ME", "PH")) +
+  ggtitle("Lobster density")
 
 ## Abundance
 I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
-                  model_names = c("IN", "ME", "PH"))
+                  model_names = c("IN", "ME", "PH")) +
+  ggtitle("Invertebrate density")
 
 # Fish plots ####################################################
 ## Biomass
 F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
-                  model_names = c("IS", "ME", "PH"))
+                  model_names = c("IS", "ME", "PH")) +
+  ggtitle("Fish biomass")
 
 ## Abundance
 F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
@@ -128,24 +131,24 @@ F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
         legend.position = c(1, 0),
         legend.text = element_text(size = 8),
         legend.title = element_text(size = 8),
-        legend.direction = "horizontal")
+        legend.direction = "horizontal") +
+  ggtitle("Fish density")
 
 # Combine plots
 plot <- cowplot::plot_grid(plotlist = list(L_plot, F_B, I_N, F_N),
-                            labels = "AUTO",
                             ncol = 2)
 
 # Export figure
 ggsave(plot = plot,
        filename = here("docs", "img", "fig_bio_results.pdf"),
        width = 6.5,
-       height = 4)
+       height = 4.2)
 
 # Export figure
 ggsave(plot = plot,
        filename = here("docs", "img", "fig_bio_results.tiff"),
        width = 6.5,
-       height = 4)
+       height = 4.2)
 
 # Save all model output
 save(L_IN,

### 2019-05-10 scripts/content/fig_socioeco_results.R

@@ -55,25 +55,29 @@ c_plot <- plot_socioeco(model_list = list(C_IN, C_ME),
                         ylab = quo(lambda[t])) +
   theme(legend.justification = c(0, 0),
         legend.position = c(0, 0.55),
-        legend.title = element_blank())
+        legend.title = element_blank()) +
+  ggtitle("Lobster catches")
 
 v_plot <- plot_socioeco(model_list = list(V_IN, V_ME),
                         model_names = list("IN", "ME"),
                         ylab = quo(lambda[t])) +
-  theme(legend.position = "none")
+  theme(legend.position = "none") +
+  ggtitle("Revenues from lobster catches")
 
-plot <- plot_grid(c_plot, v_plot, ncol = 1, labels = "AUTO")
+plot <- plot_grid(c_plot,
+                  v_plot,
+                  ncol = 1)
 
 # Export figures
 ggsave(plot = plot,
        filename = here("docs", "img", "fig_socioeco_results.pdf"),
-       width = 6.5,
-       height = 4)
+       width = 6,
+       height = 4.4)
 
 ggsave(plot = plot,
        filename = here("docs", "img", "fig_socioeco_results.tiff"),
-       width = 6.5,
-       height = 4)
+       width = 6,
+       height = 4.4)
 
 # Save the models
 save(C_IN,

### 2019-05-16 docs/PLoS_manuscript.tex
Updated for appeal process

@@
 {\Large
-\textbf\newline{Community-based TURF-reserves in small-scale fisheries} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{An interdisciplinary evaluation of community-based TURF-reserves} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \cite{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}.
-In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce, given the heterogeneity in target species and number of actors involved \cite{costello_2012}.
-One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves\cite{afflerbach_2014,gelcich_2015,lester_2017}.
+In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce, given the large number of participants, heterogeneity in fleet, gears, and targeted species; as well as seasonality and geographical distribution \cite{costello_2012,jacquet_2008}.
+One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves \cite{afflerbach_2014,gelcich_2015,lester_2017}.
 
@@
 
-Conservation science has shown how well implemented and enforced marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \cite{lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
+Conservation science has shown how well implemented and enforced marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \cite{halpern_2002,lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}.
 Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \cite{gelcich_2008,gelcich_2012}.
@@
 Of these, 18 were originally implemented within TURFs.
-However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.
+However, their effectiveness has not yet been reported in the scientific literature.
 
-Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico.
+Marine reserves have largely been evaluated from an ecological perspective, focusing on ecological indicators such as biomass and species richness or spillover\cite{halpern_2002,lester_2009,krueck_2017-J1}.
+However, TURF-reserves are intricate social-ecological systems, oftem implemented with a combination of ecological and social objectives.
+A customary evaluation of TURF-reserves is unlikely to provide an accurate representation of the changes, benefits, and limitations of such an intervention.
+
 The objective of this work is twofold.
@@
 The second objective is to identify opportunities where improvement or adjustment might lead to increased effectiveness of these reserves.
-%We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.
+We combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico.
+Our work highlights the benefits of taking an interdisciplinary approach to marine reserve evaluation, and provides a robust evaluation of TURF-reserve effectiveness that may inform adaptive management and future interventions.
 

### 2019-07-23 docs/PLoS_manuscript.tex
Added figure captions and legends for reviewr 2

@@
 Marine reserves have largely been evaluated from an ecological perspective, focusing on ecological indicators such as biomass and species richness or spillover\cite{halpern_2002,lester_2009,krueck_2017-J1}.
-However, TURF-reserves are intricate social-ecological systems, oftem implemented with a combination of ecological and social objectives.
+However, TURF-reserves are intricate social-ecological systems, often implemented with a combination of ecological and social objectives.
 A customary evaluation of TURF-reserves is unlikely to provide an accurate representation of the changes, benefits, and limitations of such an intervention.
@@
 Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \cite{suman_2010-ez,fulton_2018,fulton_2019}.
-Size structures are also collected during fish surveys.
+Size structures are also collected during fish surveys, where divers estimate fish length to the nearest centimeter.
 All sites were surveyed annually, and at least once before implementation of the reserves.
@@
 
-Socioeconomic data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves.
+Socioeconomic data contain monthly lobster landings (Kg) and revenues (in Mexican Pesos; MXP) for TURF-managed cooperatives with and without marine reserves.
 These were requested from the National Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de Acuacultura y Pesca}; CONAPESCA) via the access to information act.

### 2019-07-23 scripts/content/fig_bio_results.R

@@ -111,18 +111,18 @@ N_PH <- fish %>%
 ## Lobsters
 L_plot <- plot_model(model = list(L_IN, L_ME, L_PH),
                      model_names = c("IS", "ME", "PH")) +
-  ggtitle("Lobster density")
+  ggtitle("A) Lobster density")
 
 ## Abundance
 I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
                   model_names = c("IN", "ME", "PH")) +
-  ggtitle("Invertebrate density")
+  ggtitle("C) Invertebrate density")
 
 # Fish plots ####################################################
 ## Biomass
 F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
                   model_names = c("IS", "ME", "PH")) +
-  ggtitle("Fish biomass")
+  ggtitle("B) Fish biomass")
 
 ## Abundance
 F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
@@ -132,7 +132,7 @@ F_N <- plot_model(model = list(N_IN, N_ME, N_PH),
         legend.text = element_text(size = 8),
         legend.title = element_text(size = 8),
         legend.direction = "horizontal") +
-  ggtitle("Fish density")
+  ggtitle("D) Fish density")
 
 # Combine plots
 plot <- cowplot::plot_grid(plotlist = list(L_plot, F_B, I_N, F_N),

### 2019-07-23 scripts/content/fig_socioeco_results.R

@@ -56,13 +56,13 @@ c_plot <- plot_socioeco(model_list = list(C_IN, C_ME),
   theme(legend.justification = c(0, 0),
         legend.position = c(0, 0.55),
         legend.title = element_blank()) +
-  ggtitle("Lobster catches")
+  ggtitle("A) Lobster catches")
 
 v_plot <- plot_socioeco(model_list = list(V_IN, V_ME),
                         model_names = list("IN", "ME"),
                         ylab = quo(lambda[t])) +
   theme(legend.position = "none") +
-  ggtitle("Revenues from lobster catches")
+  ggtitle("B) Revenues from lobster catches")
 
 plot <- plot_grid(c_plot,
                   v_plot,

### 2019-07-24 docs/PLoS_manuscript.tex
Resubmitted to PLoS

@@
 Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world.
-Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage.
+Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage, which may lead to overexploitation.
 Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments.
@@
 A lack of clear ecological and socioeconomic effects likely results from a combination of factors.
-First, some of these reserves might be too young for the effects to show.
+First, some of these reserves might be too young for the effects to show (reserves were 6 - 10  years old).
 Second, the reserves are not large enough to protect mobile species, like lobster.
 Third, variable and extreme oceanographic conditions have impacted harvested populations.
-Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches.
+Fourth, local fisheries are already well managed, and while reserves may protect populations within its bounded regions, it is unlikely that reserves might have a detectable effect in catches.
 However, even small reserves are expected to provide benefits for sedentary invertebrates over longer time frames, with continued protection.
@@
 Each TURF is exclusively fished by one cooperative, and each community usually hosts no more than one cooperative.
-The profits from each TURF are shared amongst all fishers from the cooperative.
+The profits from each TURF are shared amongst all fishers from the cooperative \cite{mccay_2014,mccay_2017}.
 Fishing cooperatives interested in implementing marine reserves within their TURFs (\emph{i.e.} TURF-reserves) work with CSOs to design them.
@@
 The objective of these reserves was "to protect and recover stocks of commercially important invertebrate species"; mainly lobster and abalone.
-The reserves obtained legal recognition in 2018 \cite{dof_website_2018}.
+The reserves were implemented and enforced by the community since 2006, but obtained legal recognition in 2018 \cite{dof_website_2018}.
 
@@
 Likewise, Maria Elena and Punta Herrero are representative of fishing cooperatives in the Mexican Caribbean, which are also part of a regional Federation.
-Together, these three communities provide an accurate representation of other fishing communities that have been historically manged with TURFs in each of their regions.
+Together, these three communities provide an accurate representation of other fishing communities that have been historically managed with TURFs in each of their regions.
 While each region has additional communities that have established community-based TURF-reserves, available data would not allow us to perform the in-depth causal inference analysis that we undertake.
@@
 Reserve sites are areas where no fishing occurs.
-Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) are not directly adjacent to the reserves.
+Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) are not directly adjacent to the reserves to avoid confounding due to spillover effect (sites were at least 1 km apart).
 We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve.
@@
 All sites were surveyed annually, and at least once before implementation of the reserves.
-A summary of sampling effort and time series are shown in the supplementary materials (Tables A-B and Figures A-D in \nameref{S1_Text}).
+A summary of sampling effort and time series, as well as species checklist, are shown in the supplementary materials (Tables A-B, Figures A-D, and Tables I-J in \nameref{S1_Text}).
 
@@
 \hline
-\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) MSC certified.\\
+\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) certified by the Marine Stewardship Council.\\
 \hline
@@
 The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators.
-Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as done in \cite{leslie_2015-na}), we followed a similar approach to \cite{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems.
+Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as done in Leslie et al \cite{leslie_2015-na}), we followed a similar approach to \cite{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems.
 We based our variable selection primarily on previous analyses of Mexican fishing cooperatives \cite{leslie_2015-na,basurto_2013-oq}.
@@
 \ref{fig:indicators}B-D).
-Invertebrate and fish densities showed positive trends in Natividad in most years.
-Reserves in Maria Elena and Punta Herrero showed positive trends for invertebrates in some years.
+Invertebrate and fish densities showed positive trends in all reserves for some years.
 However, these were not statistically significant.
@@
 For example, fishers operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves.
-Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have sotrong governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management.
+Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have strong governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management \cite{mccay_2014,ayer_2018}.
+
 In general, success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources upon which they depend \cite{jupiter_2017}.
@@
 Our results indicate that these TURF-reserves have not increased lobster densities.
-Additionally, no co-benefits were identified when using other ecological indicators aside from the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad, as well as positive trends in invertebrate and fish densities \cite{micheli_2012-EU}.
+Additionally, no co-benefits were identified when using other ecological indicators aside from the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad, as well as positive trends in invertebrate and fish densities \cite{micheli_2012-EU,Villasenor-Derbez_2018}.
 The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation.
@@
 These communities exhibit all the social enabling conditions for effective reserve and resource management.
-Here we discuss possible shortcomings in our analyses as well alternative explanations for the observed lack of effectiveness.
+Here we discuss possible shortcomings in our analyses as well possible explanations for the lack of effectiveness.
 
@@
 For example, the rapid increase observed for lobster densities in Isla Natividad on the sixth year (\emph{i.e.} 2012; Fig.
-\ref{fig:indicators}A), occurs a year after the hypoxia events described by \cite{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
+\ref{fig:indicators}A), occurs a year after the hypoxia events described by Micheli et al \cite{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
 
@@
 
+Lack of evidence of reserve effectiveness is commonly associated to lack of proper enforcement and compliance, as well as capacity shortfalls \cite{edgar_2014-UO,difranco_2016-Xw,gill_2017}.
+However, it is unlikely that the evaluated reserves are subject to these common problems.
+The reserves are implemented within TURFs, which provide a sense of ownership of resources and promotes sustainable management \cite{mccay_2017}.
+The same incentives that ensure sustainable management in a TURF should extend to the reserves that these contain.
+This is supported by our SES analyses, which suggests that these communities exhibit strong governance and enforcement structures.
+Instead, the lack of effectiveness observed may be driven by a combination of factors, such as age of the reserves, sub-optimal reserve design, or environmental variation.
+The next paragraphs provide a detailed discussion of these possible causes.
+
 A first possible explanation for the lack of effectiveness may be the young age of the reserves.
@@
 With these characteristics, one would expect the reserves to be effective.
-In fact, the oldest reserves shows some positive trends for invertebrate and fish densities, as well as income.
-Maria Elena and Punta Herrero are relatively young reserves (\emph{i.e.} \textless 6 years old; RS5 in Table~\ref{table:ses}) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \cite{dasilva_2015-zX}.
+In fact, the oldest reserves show some positive trends for invertebrate and fish densities, as well as income.
+Maria Elena and Punta Herrero are relatively young reserves (\emph{i.e.} \textless 6 years old; RS5 in Table~\ref{table:ses}) and effects may not yet be evident; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \cite{dasilva_2015-zX}.
 
 Another key condition for effectiveness is reserve size \cite{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (\emph{sensu} \cite{rees_2018}).
-Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \cite{rossetto_2015-V0}, however, abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary molusks, but not lobsters.
-Design principles developed by \cite{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
+In Isla Natividad, the reserves can yield fishery benefits for the abalone fishery \cite{rossetto_2015-V0}, however, abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary molusks, but not lobsters.
+Small reserves in the Mediterranean Sea have shown that the effect of a reserve is only observable for species with a home range smaller than the reserve \cite{difranco_2018}.
+However, design principles developed by Green et al \cite{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
 As shown in the SES analysis, the size of the marine reserves appears small compared to the movement capacity of the main targeted species (RU1, RS3; Table~\ref{table:ses}).
@@
 Showing commitment to marine conservation and sustainable fishing practices has allowed fishers to have greater bargaining power and leverage over fisheries management \cite{prezramrez_2012}.
-These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-establishing the reserves \cite{ayer_2018}.
+These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-establishing the reserves \cite{ayer_2018,bennett_2019}.
 
@@
 The communities seem to exhibit all social and governance conditions that would result in reserve effectiveness.
-However, these reserves may not be large enough to provide the necesary protection that would lead to increases in biomass.
+However, these reserves may not be large enough to provide the necessary protection that would lead to increases in biomass.
 For the particular case of the reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
@@
 Figure E with time series of socioeconomic indicators.
+Table I with a checklist of invertebrate species.
+Table J with a checklist of fish species.
 
@@
 This study would have not been possible without the effort by members of the fishing communities here mentioned, who participated in the data-collection process.
-The authors would like to thank comments by two reviewrers that significantly improved the quality of our work.
+The authors would like to thank comments by three reviewers that significantly improved the quality of our work.
 

### 2019-07-24 scripts/processing/data_cleaning_logbook.Rmd
Resubmitted to PLoS

@@
   filter(between(abundancia, 0.1, 100)) %>%
+  filter(!generoespecie %in% c("Leptogorgia chilensis", "Muricea californica", "Muricea spp")) %>% 
+  mutate(generoespecie = case_when(generoespecie == "Anemona spp" ~ "Anthopleura spp",
+                                   generoespecie == "Otros erizos" ~ "Other sea urchins",
+                                   T ~ generoespecie)) %>%
   group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie) %>% 
@@
 
-FB_update <- rfishbase::length_weight(needed_spp, fields = c("a", "b")) %>% 
-  select(GeneroEspecie = sciname, a, b) %>% 
+FB_update <- rfishbase::length_weight(needed_spp, fields = c("Species", "a", "b")) %>% 
+  select(GeneroEspecie = Species, a, b) %>% 
   mutate(GeneroEspecie = case_when(GeneroEspecie %in% c("Kyphosus sectatrix", "Kyphosus incisor") ~ "Kyphosus",
@@
          b = ifelse(is.na(b), bg, b)) %>%
+  mutate(generoespecie = case_when(generoespecie == "Chromis" ~ "Chromis spp",
+                                   generoespecie == "Haemulon" ~ "Haemulon spp",
+                                   generoespecie == "Kyphosus" ~ "Kyphosus spp",
+                                   generoespecie == "Lutjanus" ~ "Lutjanus spp",
+                                   generoespecie == "Scarus" ~ "Scarus spp",
+                                   generoespecie == "Sebastes" ~ "Sebastes spp",
+                                   generoespecie == "Stegastes" ~ "Stegastes spp",
+                                   T ~ generoespecie)) %>% 
   group_by(comunidad, ano, rc, sitio, zona, transecto, generoespecie, talla, a,b) %>% 

### 2019-07-24 scripts/content/fig_bio_results.R

@@ -21,9 +21,7 @@ source(file = here("scripts", "functions", "biomass.R")) # Calculate fish biomas
 # Load analysis functions
 source(file = here("scripts", "functions", "model.R")) #fits the model
 source(file = here("scripts", "functions", "robust_se.R")) #calculates heteroskedastic-robust Standard Errors
-# source(file = here("scripts", "robust_se_list.R")) #heteroskedastic-robust Standard Errors for stargazer
 source(file = here("scripts", "functions", "plot_model.R")) #plots effect sizes for Bio
-# source(file = here("scripts", "plot_socioeco.R")) #plots effect sizes for Socioeco
 
 # Load data
 # # Invertebrate data

### 2019-07-24 scripts/content/supp_information.R

@@ -34,7 +34,7 @@ invert %>%
   group_by(comunidad, zona) %>%
   count() %>% 
   ungroup() %>% 
-  spread(zona, nn) %>% 
+  spread(zona, n) %>% 
   mutate(age = c(10, 4, 4)) %>% 
   knitr::kable(col.names = c("Community", "Control", "Reserve", "Years of monitoring"),
                booktabs = T,
@@ -125,7 +125,7 @@ fish %>%
   group_by(comunidad, zona) %>%
   count() %>% 
   ungroup() %>% 
-  spread(zona, nn) %>% 
+  spread(zona, n) %>% 
   mutate(age = c(10, 4, 4)) %>% 
   knitr::kable(col.names = c("Community", "Control", "Reserve", "Years of monitoring"),
                booktabs = T,
@@ -326,4 +326,29 @@ list(C_ME, V_ME) %>%
   supp_regtable(caption = "Coefficient estimates of socioeconomic indicators for Maria Elena.",
                 bio = F,
                 filename = here("docs", "tab", "S8_table.tex"))
+
+# invert species checklist
+invert %>%
+  group_by(comunidad, generoespecie) %>%
+  summarize(n = (n() > 0) * 1) %>%
+  spread(comunidad, n, fill = 0) %>% 
+  knitr::kable(col.names = c("Species", "Isla Natividad", "Maria Elena", "Punta Herrero"),
+               booktabs = T,
+               format = "latex",
+               caption = "{\\bf Invertebrate species checklist.}") %>% 
+  cat(file = here("docs", "tab", "S9_table.tex"))
+  
+
+# fish species checklist
+fish %>%
+  group_by(comunidad, generoespecie) %>%
+  summarize(n = (n() > 0) * 1) %>% 
+  spread(comunidad, n, fill = 0) %>% 
+  knitr::kable(col.names = c("Species", "Isla Natividad", "Maria Elena", "Punta Herrero"),
+               booktabs = T,
+               format = "latex",
+               caption = "{\\bf Fish species checklist.}",
+               longtable = T) %>% 
+  kableExtra::kable_styling(latex_options = c("repeat_header")) %>% 
+  cat(file = here("docs", "tab", "S10_table.tex"))
 

### 2019-08-09 docs/PLoS_manuscript.tex
Added second revisions for PLoS RnR after appeal

@@
 Third, variable and extreme oceanographic conditions have impacted harvested populations.
-Fourth, local fisheries are already well managed, and while reserves may protect populations within its bounded regions, it is unlikely that reserves might have a detectable effect in catches.
+Fourth, local fisheries are already well managed, and while reserves may protect populations within its boundaries, it is unlikely that reserves might have a detectable effect in catches.
 However, even small reserves are expected to provide benefits for sedentary invertebrates over longer time frames, with continued protection.
@@
 The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators.
-Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as done in Leslie et al \cite{leslie_2015-na}), we followed a similar approach to \cite{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems.
+Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as done in \cite{leslie_2015-na}), we followed a similar approach to \cite{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems.
 We based our variable selection primarily on previous analyses of Mexican fishing cooperatives \cite{leslie_2015-na,basurto_2013-oq}.
@@
 For example, the rapid increase observed for lobster densities in Isla Natividad on the sixth year (\emph{i.e.} 2012; Fig.
-\ref{fig:indicators}A), occurs a year after the hypoxia events described by Micheli et al \cite{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
+\ref{fig:indicators}A), occurs a year after the hypoxia events described by \cite{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish.
 
@@
 The same incentives that ensure sustainable management in a TURF should extend to the reserves that these contain.
-This is supported by our SES analyses, which suggests that these communities exhibit strong governance and enforcement structures.
+This is supported by our SES analysis, which suggests that these communities exhibit strong governance and enforcement structures.
 Instead, the lack of effectiveness observed may be driven by a combination of factors, such as age of the reserves, sub-optimal reserve design, or environmental variation.
@@
 Small reserves in the Mediterranean Sea have shown that the effect of a reserve is only observable for species with a home range smaller than the reserve \cite{difranco_2018}.
-However, design principles developed by Green et al \cite{green_2017} for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across.
+However, design principles developed for marine reserves in the Caribbean state that reserves "should be more than twice the size of the home range of adults and juveniles", and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across \cite{green_2017}.
 As shown in the SES analysis, the size of the marine reserves appears small compared to the movement capacity of the main targeted species (RU1, RS3; Table~\ref{table:ses}).
@@
 In contrast, invertebrate fisheries have seen declining catches and finfish fisheries are not managed under TURFs, which may explain the positive trends observed.
-Furthermore, \cite{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open access.
+Furthermore, TURFs alone can have greater biomass and richness than areas operating under open access \cite{gelcich_2008}.
 This might reduce the difference between indicators from the TURF and reserve sites, making it difficult to detect such a small change.

### 2019-08-14 docs/PLoS_manuscript.tex
Added production assignment changes

@@
 
+\clearpage
+
 \subsection*{Data analysis}
@@
 
-\bibliography{references}
+% \bibliography{references}
+\begin{thebibliography}{10}
+
+\bibitem{pauly_2005-qV}
+Pauly D, Watson R, Alder J.
+\newblock Global trends in world fisheries: impacts on marine ecosystems and
+  food security.
+\newblock Philosophical Transactions of the Royal Society B: Biological
+  Sciences. 2005;360(1453):5--12.
+\newblock doi:{10.1098/rstb.2004.1574}.
+
+\bibitem{worm_2006-IB}
+Worm B, Barbier EB, Beaumont N, Duffy JE, Folke C, Halpern BS, et~al.
+\newblock Impacts of biodiversity loss on ocean ecosystem services.
+\newblock Science. 2006;314(5800):787--790.
+\newblock doi:{10.1126/science.1132294}.
+
+\bibitem{halpern_2008-dK}
+Halpern BS, Walbridge S, Selkoe KA, Kappel CV, Micheli F, D'Agrosa C, et~al.
+\newblock A global map of human impact on marine ecosystems.
+\newblock Science. 2008;319(5865):948--952.
+\newblock doi:{10.1126/science.1149345}.
+
+\bibitem{costello_2012}
+Costello C, Ovando D, Hilborn R, Gaines SD, Deschenes O, Lester SE.
+\newblock Status and solutions for the world's unassessed fisheries.
+\newblock Science. 2012;338(6106):517--520.
+\newblock doi:{10.1126/science.1223389}.
+
+\bibitem{jacquet_2008}
+Jacquet J, Pauly D.
+\newblock Funding Priorities: Big Barriers to Small-Scale Fisheries.
+\newblock Conservation Biology. 2008;22(4):832--835.
+\newblock doi:{10.1111/j.1523-1739.2008.00978.x}.
+
+\bibitem{afflerbach_2014}
+Afflerbach JC, Lester SE, Dougherty DT, Poon SE.
+\newblock A global survey of TURF-reserves, Territorial Use Rights for
+  Fisheries coupled with marine reserves.
+\newblock Global Ecology and Conservation. 2014;2:97--106.
+\newblock doi:{10.1016/j.gecco.2014.08.001}.
+
+\bibitem{gelcich_2015}
+Gelcich S, Donlan CJ.
+\newblock Incentivizing biodiversity conservation in artisanal fishing
+  communities through territorial user rights and business model innovation.
+\newblock Conserv Biol. 2015;29(4):1076--1085.
+\newblock doi:{10.1111/cobi.12477}.
+
+\bibitem{lester_2017}
+Lester S, McDonald G, Clemence M, Dougherty D, Szuwalski C.
+\newblock Impacts of TURFs and marine reserves on fisheries and conservation
+  goals: theory, empirical evidence, and modeling.
+\newblock BMS. 2017;93(1):173--198.
+\newblock doi:{10.5343/bms.2015.1083}.
+
+\bibitem{gelcich_2008}
+Gelcich S, Godoy N, Prado L, Castilla JC.
+\newblock Add-on conservation benefits of marine territorial user rights
+  fishery policies in central Chile.
+\newblock Ecol Appl. 2008;18(1):273--281.
+\newblock doi:{10.1890/06-1896.1}.
+
+\bibitem{costello_2010}
+Costello C, Kaffine DT.
+\newblock Marine protected areas in spatial property-rights fisheries.
+\newblock Australian Journal of Agricultural and Resource Economics.
+  2010;54(3):321--341.
+\newblock doi:{10.1111/j.1467-8489.2010.00495.x}.
+
+\bibitem{mccay_2014}
+McCay BJ, Micheli F, Ponce-Díaz G, Murray G, Shester G, Ramirez-Sanchez S,
+  et~al.
+\newblock Cooperatives, concessions, and co-management on the Pacific coast of
+  Mexico.
+\newblock Marine Policy. 2014;44:49--59.
+\newblock doi:{10.1016/j.marpol.2013.08.001}.
+
+\bibitem{halpern_2002}
+Halpern BS, Warner RR.
+\newblock Marine reserves have rapid and lasting effects.
+\newblock Ecology Letters. 2002;5(3):361--366.
+\newblock doi:{10.1046/j.1461-0248.2002.00326.x}.
+
+\bibitem{lester_2009}
+Lester S, Halpern B, Grorud-Colvert K, Lubchenco J, Ruttenberg B, Gaines S,
+  et~al.
+\newblock Biological effects within no-take marine reserves: a global
+  synthesis.
+\newblock Mar Ecol Prog Ser. 2009;384:33--46.
+\newblock doi:{10.3354/meps08029}.
+
+\bibitem{roberts_2017-J9}
+Roberts CM, OLeary BC, McCauley DJ, Cury PM, Duarte CM, Lubchenco J, et~al.
+\newblock Marine reserves can mitigate and promote adaptation to climate
+  change.
+\newblock Proc Natl Acad Sci USA. 2017;114(24):6167--6175.
+\newblock doi:{10.1073/pnas.1701262114}.
+
+\bibitem{micheli_2012-EU}
+Micheli F, Saenz-Arroyo A, Greenley A, Vazquez L, Espinoza~Montes JA, Rossetto
+  M, et~al.
+\newblock Evidence that marine reserves enhance resilience to climatic impacts.
+\newblock PLoS ONE. 2012;7(7):e40832.
+\newblock doi:{10.1371/journal.pone.0040832}.
+
+\bibitem{krueck_2017-J1}
+Krueck NC, Ahmadia GN, Possingham HP, Riginos C, Treml EA, Mumby PJ.
+\newblock Marine reserve targets to sustain and rebuild unregulated fisheries.
+\newblock PLoS Biol. 2017;15(1):e2000537.
+\newblock doi:{10.1371/journal.pbio.2000537}.
+
+\bibitem{gelcich_2012}
+Gelcich S, Fernández M, Godoy N, Canepa A, Prado L, Castilla JC.
+\newblock Territorial user rights for fisheries as ancillary instruments for
+  marine coastal conservation in chile.
+\newblock Conserv Biol. 2012;26(6):1005--1015.
+\newblock doi:{10.1111/j.1523-1739.2012.01928.x}.
+
+\bibitem{smallhornwest_2018}
+Smallhorn-West PF, Bridge TCL, Malimali S, Pressey RL, Jones GP.
+\newblock Predicting impact to assess the efficacy of community-based marine
+  reserve design.
+\newblock Conserv Lett. 2018; p. e12602.
+\newblock doi:{10.1111/conl.12602}.
+
+\bibitem{nom}
+NOM-049-SAG/PESC.
+\newblock NORMA Oficial Mexicana NOM-049-SAG/PESC-2014, Que determina el
+  procedimiento para establecer zonas de refugio para los recursos pesqueros en
+  aguas de jurisdicción federal de los Estados Unidos Mexicanos.
+\newblock DOF. 2014;.
+
+\bibitem{johannes_2002}
+Johannes RE.
+\newblock The renaissance of community-based marine resource management in
+  Oceania.
+\newblock Annual Review of Ecology and Systematics. 2002;33(1):317--340.
+
+\bibitem{bohnsack_2004}
+Bohnsack JA, Ault JS, Causey B.
+\newblock Why have no-take marine protected areas?
+\newblock In: American Fisheries Society Symposium. vol.~42; 2004. p. 185--193.
+
+\bibitem{beger_2004}
+Beger M, Harborne AR, Dacles TP, Solandt JL, Ledesma GL.
+\newblock A framework of lessons learned from community-based marine reserves
+  and its effectiveness in guiding a new coastal management initiative in the
+  Philippines.
+\newblock Environ Manage. 2004;34(6):786--801.
+\newblock doi:{10.1007/s00267-004-0149-z}.
+
+\bibitem{mccay_2017}
+McCay B.
+\newblock Territorial use rights in fisheries of the northern Pacific coast of
+  Mexico.
+\newblock BMS. 2017;93(1):69--81.
+\newblock doi:{10.5343/bms.2015.1091}.
+
+\bibitem{dof_website_2018}
+DOF.
+\newblock Acuerdo por el que se establece una red de dos Zonas de Refugio
+  Pesquero Parciales Permanentes en aguas marinas de jurisdicción federal
+  adyacentes a Isla Natividad, ubicada en el Municipio de Mulegé, en el Estado
+  de Baja California Sur.
+\newblock Diario Oficial de la Federación. 2018;.
+
+\bibitem{dof_website_2012}
+DOF.
+\newblock Acuerdo por el que se establece una red de zonas de refugio pesquero
+  en aguas marinas de jurisdicción federal ubicadas en el área de Sian Ka an,
+  dentro de la Bahía Espíritu Santo en el Estado de Quintana Roo.
+\newblock Diario Oficial de la Federación. 2012;.
+
+\bibitem{dof_website_2013}
+DOF.
+\newblock Acuerdo por el que se establece una red de zonas de refugio pesquero
+  en aguas marinas de jurisdicción federal ubicadas en las áreas de Banco
+  Chinchorro y Punta Herrero en el Estado de Quintana Roo.
+\newblock Diario Oficial de la Federación. 2013;.
+
+\bibitem{dof_website_2017b}
+DOF.
+\newblock Acuerdo por el que se amplía la vigencia del similar que establece
+  una red de zonas de refugio pesquero en aguas marinas de jurisdicción
+  federal ubicadas en el área de Sian Ka an, dentro de la Bahía Espíritu
+  Santo en el Estado de Quintana Roo, publicado el 30 de noviembre de 2012.
+\newblock Diario Oficial de la Federación. 2017;.
+
+\bibitem{acevesbueno_2017}
+Aceves-Bueno E, Cornejo-Donoso J, Miller SJ, Gaines SD.
+\newblock Are Territorial Use Rights in Fisheries ({TURFs}) sufficiently large?
+\newblock Marine Policy. 2017;78(1):189--195.
+\newblock doi:{10.1016/j.marpol.2017.01.024}.
+
+\bibitem{ostrom_2009}
+Ostrom E.
+\newblock A General Framework for Analyzing Sustainability of Social-Ecological
+  Systems.
+\newblock Science. 2009;325(5939):419--422.
+\newblock doi:{10.1126/science.1172133}.
+
+\bibitem{stewartoaten_1986}
+Stewart-Oaten A, Murdoch WW, Parker KR.
+\newblock Environmental impact assessment: "pseudoreplication" in time?
+\newblock Ecology. 1986;67(4):929--940.
+\newblock doi:{10.2307/1939815}.
+
+\bibitem{francinifilho_2008}
+Francini-Filho RB, Moura RL.
+\newblock Evidence for spillover of reef fishes from a no-take marine reserve:
+  An evaluation using the before-after control-impact ({BACI}) approach.
+\newblock Fisheries Research. 2008;93(3):346--356.
+\newblock doi:{10.1016/j.fishres.2008.06.011}.
+
+\bibitem{depalma_2018}
+De~Palma A, Sanchez~Ortiz K, Martin PA, Chadwick A, Gilbert G, Bates AE, et~al.
+\newblock Challenges With Inferring How Land-Use Affects Terrestrial
+  Biodiversity: Study Design, Time, Space and Synthesis.
+\newblock Advances in ecological research.
+  2018;doi:{10.1016/bs.aecr.2017.12.004}.
+
+\bibitem{kerr_2019}
+Kerr LA, Kritzer JP, Cadrin SX.
+\newblock Strengths and limitations of before–after–control–impact
+  analysis for testing the effects of marine protected areas on managed
+  populations.
+\newblock {ICES} Journal of Marine Science. 2019;doi:{10.1093/icesjms/fsz014}.
+
+\bibitem{Villasenor-Derbez_2018}
+Villaseñor-Derbez JC, Faro C, Wright M, Martínez J, Fitzgerald S, Fulton S,
+  et~al.
+\newblock A user-friendly tool to evaluate the effectiveness of no-take marine
+  reserves.
+\newblock PLOS ONE. 2018;13(1):1--21.
+\newblock doi:{10.1371/journal.pone.0191821}.
+
+\bibitem{suman_2010-ez}
+Suman CS, Saenz-Arroyo A, Dawson C, Luna MC.
+\newblock Manual de Instruccion de Reef Check California: Guia de instruccion
+  para el monitoreo del bosque de sargazo en la Peninsula de Baja California.
+\newblock Pacific Palisades, CA, USA: Reef Check Foundation; 2010.
+
+\bibitem{fulton_2018}
+Fulton S, Caamal-Madrigal J, Aguilar-Perera A, Bourillón L, Heyman WD.
+\newblock Marine Conservation Outcomes are More Likely when Fishers Participate
+  as Citizen Scientists: Case Studies from the Mexican Mesoamerican Reef.
+\newblock CSTP. 2018;3(1).
+\newblock doi:{10.5334/cstp.118}.
+
+\bibitem{fulton_2019}
+Fulton S, Hernandez-Velasco A, Suarez-Castillo A, Fernandez-Rivera~Melo F, Rojo
+  M, Saenz-Arroyo A, et~al.
+\newblock From fishing fish to fishing data: the role of artisanal fishers in
+  conservation and resource management in mexico.
+\newblock In: Salas S, Barragán-Paladines MJ, Chuenpagdee R, editors.
+  Viability and Sustainability of Small-Scale Fisheries in Latin America and
+  The Caribbean. vol.~19 of {MARE} Publication Series. Cham: Springer
+  International Publishing; 2019. p. 151--175.
+\newblock Available from:
+  \url{http://link.springer.com/10.1007/978-3-319-76078-0\_7}.
+
+\bibitem{ayer_2018}
+Ayer A, Fulton S, Caamal-Madrigal JA, Espinoza-Tenorio A.
+\newblock Halfway to sustainability: Management lessons from community-based,
+  marine no-take zones in the Mexican Caribbean.
+\newblock Marine Policy. 2018;93:22--30.
+\newblock doi:{10.1016/j.marpol.2018.03.008}.
+
+\bibitem{oecd_2017}
+OECD. Inflation {CPI}; 2017.
+\newblock Available from: \url{https://data.oecd.org/price/inflation-cpi.htm}.
+
+\bibitem{basurto_2013-oq}
+Basurto X, Gelcich S, Ostrom E.
+\newblock The social–ecological system framework as a knowledge
+  classificatory system for benthic small-scale fisheries.
+\newblock Global Environmental Change. 2013;23(6):1366--1380.
+\newblock doi:{10.1016/j.gloenvcha.2013.08.001}.
+
+\bibitem{difranco_2016-Xw}
+Di~Franco A, Thiriet P, Di~Carlo G, Dimitriadis C, Francour P, Gutiérrez NL,
+  et~al.
+\newblock Five key attributes can increase marine protected areas performance
+  for small-scale fisheries management.
+\newblock Sci Rep. 2016;6(1):38135.
+\newblock doi:{10.1038/srep38135}.
+
+\bibitem{edgar_2014-UO}
+Edgar GJ, Stuart-Smith RD, Willis TJ, Kininmonth S, Baker SC, Banks S, et~al.
+\newblock Global conservation outcomes depend on marine protected areas with
+  five key features.
+\newblock Nature. 2014;506(7487):216--220.
+\newblock doi:{10.1038/nature13022}.
+
+\bibitem{card_1994}
+Card D, Krueger AB.
+\newblock Minimum Wages and Employment: A Case Study of {theFast}-Food Industry
+  in New Jersey and Pennsylvania.
+\newblock AER. 1994;84(4):772--793.
+
+\bibitem{moland_2013}
+Moland E, Olsen EM, Knutsen H, Garrigou P, Espeland SH, Kleiven AR, et~al.
+\newblock Lobster and cod benefit from small-scale northern marine protected
+  areas: inference from an empirical before-after control-impact study.
+\newblock Proceedings of the Royal Society B: Biological Sciences.
+  2013;280(1754):20122679--20122679.
+\newblock doi:{10.1098/rspb.2012.2679}.
+
+\bibitem{zeileis_2004-7n}
+Zeileis A.
+\newblock Econometric Computing with HC and HAC Covariance Matrix Estimators.
+\newblock J Stat Softw. 2004;11(10).
+\newblock doi:{10.18637/jss.v011.i10}.
+
+\bibitem{R_2018}
+{R Core Team}. R: A Language and Environment for Statistical Computing; 2018.
+\newblock Available from: \url{https://www.R-project.org/}.
+
+\bibitem{leslie_2015-na}
+Leslie HM, Basurto X, Nenadovic M, Sievanen L, Cavanaugh KC, Cota-Nieto JJ,
+  et~al.
+\newblock Operationalizing the social-ecological systems framework to assess
+  sustainability.
+\newblock Proc Natl Acad Sci U S A. 2015;112(19):5979--5984.
+\newblock doi:{10.1073/pnas.1414640112}.
+
+\bibitem{jupiter_2017}
+Jupiter SD, Epstein G, Ban NC, Mangubhai S, Fox M, Cox M.
+\newblock A social–ecological systems approach to assessing conservation and
+  fisheries outcomes in fijian locally managed marine areas.
+\newblock Soc Nat Resour. 2017;30(9):1096--1111.
+\newblock doi:{10.1080/08941920.2017.1315654}.
+
+\bibitem{guidetti_2002}
+Guidetti P.
+\newblock The importance of experimental design in detecting the effects of
+  protection measures on fish in Mediterranean {MPAs}.
+\newblock Aquatic Conserv: Mar Freshw Ecosyst. 2002;12(6):619--634.
+\newblock doi:{10.1002/aqc.514}.
+
+\bibitem{ferraro_2006}
+Ferraro PJ, Pattanayak SK.
+\newblock Money for nothing? A call for empirical evaluation of biodiversity
+  conservation investments.
+\newblock PLoS Biol. 2006;4(4):e105.
+\newblock doi:{10.1371/journal.pbio.0040105}.
+
+\bibitem{viana_2017}
+Viana DF, Halpern BS, Gaines SD.
+\newblock Accounting for tourism benefits in marine reserve design.
+\newblock {PLoS} {ONE}. 2017;12(12):e0190187.
+\newblock doi:{10.1371/journal.pone.0190187}.
+
+\bibitem{gill_2017}
+Gill DA, Mascia MB, Ahmadia GN, Glew L, Lester SE, Barnes M, et~al.
+\newblock Capacity shortfalls hinder the performance of marine protected areas
+  globally.
+\newblock Nature. 2017;543(7647):665--669.
+\newblock doi:{10.1038/nature21708}.
+
+\bibitem{babcock_2010}
+Babcock RC, Shears NT, Alcala AC, Barrett NS, Edgar GJ, Lafferty KD, et~al.
+\newblock Decadal trends in marine reserves reveal differential rates of change
+  in direct and indirect effects.
+\newblock Proc Natl Acad Sci {USA}. 2010;107(43):18256--18261.
+\newblock doi:{10.1073/pnas.0908012107}.
+
+\bibitem{dasilva_2015-zX}
+da~Silva IM, Hill N, Shimadzu H, Soares AMVM, Dornelas M.
+\newblock Spillover effects of a community-managed marine reserve.
+\newblock PLoS ONE. 2015;10(4):e0111774.
+\newblock doi:{10.1371/journal.pone.0111774}.
+
+\bibitem{rees_2018}
+Rees SE, Pittman SJ, Foster N, Langmead O, Griffiths C, Fletcher S, et~al.
+\newblock Bridging the divide: Social–ecological coherence in Marine
+  Protected Area network design.
+\newblock Aquatic Conservation: Marine and Freshwater Ecosystems. 2018;.
+
+\bibitem{rossetto_2015-V0}
+Rossetto M, Micheli F, Saenz-Arroyo A, Montes JAE, De~Leo GA.
+\newblock No-take marine reserves can enhance population persistence and
+  support the fishery of abalone.
+\newblock Can J Fish Aquat Sci. 2015;72(10):1503--1517.
+\newblock doi:{10.1139/cjfas-2013-0623}.
+
+\bibitem{difranco_2018}
+Di~Franco A, Plass-Johnson JG, Di~Lorenzo M, Meola B, Claudet J, Gaines SD,
+  et~al.
+\newblock Linking home ranges to protected area size: The case study of the
+  Mediterranean Sea.
+\newblock Biological Conservation. 2018;221:175--181.
+\newblock doi:{10.1016/j.biocon.2018.03.012}.
+
+\bibitem{green_2017}
+Green A, Chollett I, Suarez A, Dahlgren C, Cruz S, Zepeda C, et~al.
+\newblock Biophysical Principles for Designing a Network of Replenishment Zones
+  for the Mesoamerican Reef System; 2017.
+
+\bibitem{ovando_2016-Wg}
+Ovando D, Dougherty D, Wilson JR.
+\newblock Market and design solutions to the short-term economic impacts of
+  marine reserves.
+\newblock Fish Fish. 2016;17(4):939--954.
+\newblock doi:{10.1111/faf.12153}.
+
+\bibitem{hilborn_2004}
+Hilborn R, Stokes K, Maguire JJ, Smith T, Botsford LW, Mangel M, et~al.
+\newblock When can marine reserves improve fisheries management?
+\newblock Ocean and Coastal Management. 2004;47(3):197 -- 205.
+\newblock doi:{https://doi.org/10.1016/j.ocecoaman.2004.04.001}.
+
+\bibitem{hilborn_2006}
+Hilborn R, Micheli F, De~Leo GA.
+\newblock Integrating marine protected areas with catch regulation.
+\newblock Can J Fish Aquat Sci. 2006;63(3):642--649.
+\newblock doi:{10.1139/f05-243}.
+
+\bibitem{dof_website_1993}
+DOF.
+\newblock Norma Oficial Mexicana 006-PESC-1993, para regular el aprovechamiento
+  de todas las especies de langosta en las aguas de Jurisdiccion Federal del
+  Golfo de Mexico y mar Caribe, asi como del Oceano Pacifico incluyendo el
+  Golfo de California.
+\newblock Diario Oficial de la Federación. 1993;.
+
+\bibitem{cavole_2016}
+Cavole LM, Demko AM, Diner RE, Giddings A, Koester I, Pagniello CMLS, et~al.
+\newblock Biological Impacts of the 2013–2015 Warm-Water Anomaly in the
+  Northeast Pacific: Winners, Losers, and the Future.
+\newblock Oceanography. 2016;29(2):273--285.
+
+\bibitem{hughes_2018}
+Hughes TP, Anderson KD, Connolly SR, Heron SF, Kerry JT, Lough JM, et~al.
+\newblock Spatial and temporal patterns of mass bleaching of corals in the
+  Anthropocene.
+\newblock Science. 2018;.
+
+\bibitem{breitburg_2018}
+Breitburg D, Levin LA, Oschlies A, Grégoire M, Chavez FP, Conley DJ, et~al.
+\newblock Declining oxygen in the global ocean and coastal waters.
+\newblock Science. 2018;.
+
+\bibitem{woodson_2018}
+Woodson CB, Micheli F, Boch C, Al-Najjar M, Espinoza A, Hernandez A, et~al.
+\newblock Harnessing marine microclimates for climate change adaptation and
+  marine conservation.
+\newblock Conservation Letters. 2018; p. e12609.
+\newblock doi:{10.1111/conl.12609}.
+
+\bibitem{deleo_2015}
+De~Leo GA, Micheli F.
+\newblock The good, the bad and the ugly of marine reserves for fishery yields.
+\newblock Philos Trans R Soc Lond, B, Biol Sci. 2015;370(1681).
+\newblock doi:{10.1098/rstb.2014.0276}.
+
+\bibitem{aalto}
+Aalto EA, Micheli F, Boch CA, Espinoza~Montes JA, Woodson CB, De~Leo GA.
+\newblock Catastrophic Mortality, Allee Effects, and Marine Protected Areas.
+\newblock The American Naturalist. 0;0(0):000--000.
+\newblock doi:{10.1086/701781}.
+
+\bibitem{finkbeiner_2018}
+Finkbeiner EM, Micheli F, Saenz-Arroyo A, Vazquez-Vera L, Perafan CA, Cárdenas
+  JC.
+\newblock Local response to global uncertainty: Insights from experimental
+  economics in small-scale fisheries.
+\newblock Global Environmental Change. 2018;48:151--157.
+\newblock doi:{10.1016/j.gloenvcha.2017.11.010}.
+
+\bibitem{prezramrez_2012}
+Pérez-Ramírez M, Ponce-Díaz G, Lluch-Cota S.
+\newblock The role of MSC certification in the empowerment of fishing
+  cooperatives in Mexico: The case of red rock lobster co-managed fishery.
+\newblock Ocean Coast Manag. 2012;63:24--29.
+\newblock doi:{10.1016/j.ocecoaman.2012.03.009}.
+
+\bibitem{bennett_2019}
+Bennett NJ, Di~Franco A, Calò A, Nethery E, Niccolini F, Milazzo M, et~al.
+\newblock Local support for conservation is associated with perceptions of good
+  governance, social impacts, and ecological effectiveness.
+\newblock Conservation letters. 2019; p. e12640.
+\newblock doi:{10.1111/conl.12640}.
+
+\bibitem{dof_websiteC_2012}
+DOF.
+\newblock Acuerdo por el que se establece una red de zonas de refugio en aguas
+  marinas de jurisdicción federal frente a la costa oriental del Estado de
+  Baja California Sur, en el corredor marino de San Cosme a Punta Coyote.
+\newblock Diario Oficial de la Federación. 2012;.
+
+\bibitem{dof_websiteU_2018}
+DOF.
+\newblock Acuerdo por el que se establece el área de refugio para la tortuga
+  amarilla (Caretta caretta) en el Golfo de Ulloa, en Baja California Sur.
+\newblock Diario Oficial de la Federación. 2018;.
+
+\bibitem{smith_2010}
+Smith MD, Lynham J, Sanchirico JN, Wilson JA.
+\newblock Political economy of marine reserves: understanding the role of
+  opportunity costs.
+\newblock Proc Natl Acad Sci {USA}. 2010;107(43):18300--18305.
+\newblock doi:{10.1073/pnas.0907365107}.
+
+\end{thebibliography}
 

### 2019-08-16 docs/PLoS_manuscript.tex
Added final requests from the production team

@@
 We also complimented information based on the authors' experience and knowledge of the communities (RS5, RU5, A3, GS6.2, GS99.1, GS9.2, and GS10.1).
+In Table~\ref{table:ses}, the alphanumeric codes follow \cite{basurto_2013-oq}, and an asterisk denotes variables incorporated based on \cite{difranco_2016-Xw} and \cite{edgar_2014-UO}.
 
@@
 IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero.
-Alphanumeric codes for variables follow \cite{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \cite{difranco_2016-Xw} and \cite{edgar_2014-UO}.
 The presented narrative applies equally for all communities unless otherwise noted.
+An asterisk (*) denotes variables incorporated into the framework.
 \end{table}
